### 2020-08-05 prf/lem-pgv-value-done.tex
First commit of PGV paper.

\begin{proof}
  By induction on the derivation of $\tseq[\cs{o}]{\ty{\Gamma}}{V}{T}$.
  \begin{case}[\LabTirName{T-Lam}]
    Immediately.
    \begin{mathpar}
      \inferrule*{
        \tseq[\cs{o}]{\ty{\Gamma},\tmty{x}{T}}{M}{U}
      }{\tseq[\cs{\pbot}]{\ty{\Gamma}}{\lambda x.M}{\tylolli[\cs{\pr(\ty{\Gamma})},\cs{o}]{T}{U}}}
    \end{mathpar}
  \end{case}
  \begin{case}[\LabTirName{T-Unit}]
    Immediately.
    \begin{mathpar}
      \inferrule*{
      }{\tseq[\cs{\pbot}]{\emptyenv}{\unit}{\tyunit}}
    \end{mathpar}
  \end{case}
  \begin{case}[\LabTirName{T-Pair}]
    The induction hypotheses give us $\cs{p}=\cs{q}=\cs{\pbot}$, hence $\cs{p}\sqcup\cs{q}=\cs{\pbot}$, and $\pr(\ty{\Gamma})=\pr(\ty{T})$ and $\pr(\ty{\Delta})=\pr(\ty{U})$, hence $\pr(\ty{\Gamma},\ty{\Delta})=\pr(\ty{\Gamma})\sqcap\pr(\ty{\Delta})=\pr(\ty{T})\sqcap\pr(\ty{U})=\pr(\ty{\typrod{T}{U}})$.
    \begin{mathpar}
      \inferrule*{
        \tseq[\cs{p}]{\ty{\Gamma}}{V}{T}
        \\
        \tseq[\cs{q}]{\ty{\Delta}}{W}{U}
        \\
        \cs{p}<\pr(\ty{\Delta})
      }{\tseq[\cs{p}\sqcup\cs{q}]{\ty{\Gamma},\ty{\Delta}}{\pair{V}{W}}{\typrod{T}{U}}}
    \end{mathpar}
  \end{case}
  \begin{case}[\LabTirName{T-Inl}]
    The induction hypothesis gives us $\cs{o}=\cs{\pbot}$, and $\pr(\ty{\Gamma})=\pr{\ty{T}}$. We know $\pr(\ty{T})=\pr({\ty{U}})$, hence $\pr(\ty{\Gamma})=\pr(\ty{\tysum{T}{U}})$.
    \begin{mathpar}
      \inferrule*{
        \tseq[o]{\ty{\Gamma}}{V}{T}
        \\
        \pr(\ty{T})=\pr(\ty{U})
      }{\tseq[o]{\ty{\Gamma}}{\inl{V}}{\tysum{T}{U}}}
    \end{mathpar}
  \end{case}
  \begin{case}[\LabTirName{T-Inr}]
    The induction hypothesis gives us $\cs{o}=\cs{\pbot}$, and $\pr(\ty{\Gamma})=\pr{\ty{U}}$. We know $\pr(\ty{T})=\pr({\ty{U}})$, hence $\pr(\ty{\Gamma})=\pr(\ty{\tysum{T}{U}})$.
    \begin{mathpar}
      \inferrule*{
        \tseq[o]{\ty{\Gamma}}{V}{U}
        \\
        \pr(\ty{T})=\pr(\ty{U})
      }{\tseq[o]{\ty{\Gamma}}{\inr{V}}{\tysum{T}{U}}}
    \end{mathpar}
  \end{case}
\end{proof}

%%% Local Variables:
%%% TeX-master: "../priorities"
%%% End:

### 2020-08-05 priority-cp.tex
First commit of PGV paper.

\documentclass[main.tex]{subfiles}

\begin{document}

\section{Priority CP}
\usingnamespace{pcp}

\subsection{Syntax}

\subsubsection{Priorities, types, and environments}
\[
\begin{array}{lcl}
  \ty{A}, \ty{B}, \ty{C}
  & \coloneqq & \ty{\tyone[\cs{o}]}
    \sep        \ty{\tybot[\cs{o}]}
    \sep        \ty{\tynil[\cs{o}]}
    \sep        \ty{\tytop[\cs{o}]}
    \sep        \ty{\tytens[\cs{o}]{A}{B}}
    \sep        \ty{\typarr[\cs{o}]{A}{B}}
    \sep        \ty{\typlus[\cs{o}]{A}{B}}
    \sep        \ty{\tywith[\cs{o}]{A}{B}}
  \\
  \ty{\Gamma}, \ty{\Delta}
  & \coloneqq & \ty{\emptyenv}
    \sep        \ty{\Gamma}, \tmty{x}{A}
\end{array}
\]
\[
\begin{array}{lcl}
  \ty{\co{(\tyone[\cs{o}])}} & = & \ty{\tybot[\cs{o}]} \\
  \ty{\co{(\tybot[\cs{o}])}} & = & \ty{\tyone[\cs{o}]} \\
  \ty{\co{(\tynil[\cs{o}])}} & = & \ty{\tytop[\cs{o}]} \\
  \ty{\co{(\tytop[\cs{o}])}} & = & \ty{\tynil[\cs{o}]}
\end{array}
\qquad
\begin{array}{lcl}
  \ty{\co{(\tytens[\cs{o}]{A}{B})}} & = & \ty{\typarr[\cs{o}]{\co{A}}{\co{B}}} \\
  \ty{\co{(\typarr[\cs{o}]{A}{B})}} & = & \ty{\tytens[\cs{o}]{\co{A}}{\co{B}}} \\
  \ty{\co{(\typlus[\cs{o}]{A}{B})}} & = & \ty{\tywith[\cs{o}]{\co{A}}{\co{B}}} \\
  \ty{\co{(\tywith[\cs{o}]{A}{B})}} & = & \ty{\typlus[\cs{o}]{\co{A}}{\co{B}}}
\end{array}
\]
\[
\begin{array}{lclclcl}
  \pr(\ty{\tyone[\cs{o}]})        & = & \cs{o}  \\
  \pr(\ty{\tybot[\cs{o}]})        & = & \cs{o}  \\
  \pr(\ty{\tynil[\cs{o}]})        & = & \cs{o}  \\
  \pr(\ty{\tytop[\cs{o}]})        & = & \cs{o}  \\
  \\
  \minpr(\ty{\emptyenv})          & = & \varnothing
\end{array}
\qquad
\begin{array}{lclclcl}
  \pr(\ty{\tytens[\cs{o}]{A}{B}}) & = & \cs{o}  \\
  \pr(\ty{\typarr[\cs{o}]{A}{B}}) & = & \cs{o}  \\
  \pr(\ty{\typlus[\cs{o}]{A}{B}}) & = & \cs{o}  \\
  \pr(\ty{\tywith[\cs{o}]{A}{B}}) & = & \cs{o}  \\
  \\
  \minpr(\ty{\Gamma},\tmty{x}{A}) & = & \minpr(\ty{\Gamma})\sqcap\minpr(\ty{A})
\end{array}
\]

\subsubsection{Terms}
\[
\begin{array}[t]{lcl}
  \tm{P}, \tm{Q}
  & \coloneqq & \tm{\link{x}{y}}
         \sep   \tm{\res{x}{y}{P}}
         \sep   \tm{(\ppar{P}{Q})}
         \sep   \tm{\halt}
  \\   & \sep & \tm{\send{x}{y}{P}}
         \sep   \tm{\close{x}{P}}
         \sep   \tm{\recv{x}{y}{P}}
         \sep   \tm{\wait{x}{P}}
  \\   & \sep & \tm{\inl{x}{P}}
         \sep   \tm{\inr{x}{P}}
         \sep   \tm{\offer{x}{P}{Q}}
         \sep   \tm{\absurd{x}}
\end{array}
\]

\subsubsection{Syntactic Sugar}
We define \emph{unbound} send:
\[
  \tm{\usend{x}{y}{P}}\elabarrow\tm{\send{x}{z}{(\ppar{\link{y}{z}}{P})}}
\]

\subsection{Typing}

\subsubsection{Typing Rules}
\begin{mathpar}
  \inferrule*[lab=T-Link]{
  }{\seq{\link[\ty{A}]{x}{y}}{\tmty{x}{A},\tmty{y}{\co{A}}}}

  \inferrule*[lab=T-New]{
    \seq{P}{\ty{\Gamma},\tmty{x}{A},\tmty{y}{\co{A}}}
  }{\seq{\res{x}{y}{P}}{\ty{\Gamma}}}
  \\
  \inferrule*[lab=T-Par]{
    \seq{P}{\ty{\Gamma}}
    \\
    \seq{Q}{\ty{\Delta}}
  }{\seq{\ppar{P}{Q}}{\ty{\Gamma},\ty{\Delta}}}

  \inferrule*[lab=T-Halt]{
  }{\seq{\halt}{\emptyenv}}
  \\
  \inferrule*[lab=T-Send]{
    \seq{P}{\ty{\Gamma},\tmty{y}{A},\tmty{x}{B}}
    \\
    \cs{o}<\minpr(\ty{\Gamma},\ty{A},\ty{B})
  }{\seq{\send{x}{y}{P}}{\ty{\Gamma},\tmty{x}{\tytens[\cs{o}]{A}{B}}}}

  \inferrule*[lab=T-Close]{
    \seq{P}{\ty{\Gamma}}
    \\
    \cs{o}<\minpr(\ty{\Gamma})
  }{\seq{\close{x}{P}}{\ty{\Gamma},\tmty{x}{\tyone[\cs{o}]}}}
  \\
  \inferrule*[lab=T-Recv]{
    \seq{P}{\ty{\Gamma},\tmty{y}{A},\tmty{x}{B}}
    \\
    \cs{o}<\minpr(\ty{\Gamma},\ty{A},\ty{B})
  }{\seq{\recv{x}{y}{P}}{\ty{\Gamma},\tmty{x}{\typarr[\cs{o}]{A}{B}}}}

  \inferrule*[lab=T-Wait]{
    \seq{P}{\ty{\Gamma}}
    \\
    \cs{o}<\minpr(\ty{\Gamma})
  }{\seq{\wait{x}{P}}{\ty{\Gamma},\tmty{x}{\tybot[\cs{o}]}}}
  \\
  \inferrule*[lab=T-Select-Inl]{
    \seq{P}{\ty{\Gamma},\tmty{x}{A}}
    \\
    \cs{o}<\minpr(\ty{\Gamma},\ty{A},\ty{B})
    \\
    \pr(\ty{A})=\pr(\ty{B})
  }{\seq{\inl{x}{P}}{\ty{\Gamma},\tmty{x}{\typlus[\cs{o}]{A}{B}}}}

  \inferrule*[lab=T-Select-Inr]{
    \seq{P}{\ty{\Gamma},\tmty{x}{B}}
    \\
    \cs{o}<\minpr(\ty{\Gamma},\ty{A},\ty{B})
    \\
    \pr(\ty{A})=\pr(\ty{B})
  }{\seq{\inr{x}{P}}{\ty{\Gamma},\tmty{x}{\typlus[\cs{o}]{A}{B}}}}

  \inferrule*[lab=T-Offer]{
    \seq{P}{\ty{\Gamma},\tmty{x}{A}}
    \\
    \seq{Q}{\ty{\Gamma},\tmty{x}{B}}
    \\
    \cs{o}<\minpr(\ty{\Gamma},\ty{A},\ty{B})
  }{\seq{\offer{x}{P}{Q}}{\ty{\Gamma},\tmty{x}{\tywith[\cs{o}]{A}{B}}}}

  \inferrule*[lab=T-Offer-Absurd]{
    \cs{o}<\pr(\ty{\Gamma})
  }{\seq{\absurd{x}}{\ty{\Gamma},\tmty{x}{\tytop[\cs{o}]}}}
\end{mathpar}

\subsubsection{Typing Rules for Syntactic Sugar}
\begin{mathpar}
  \inferrule*[lab=T-Unbound-Send]{
    \seq{P}{\ty{\Gamma},\tmty{x}{B}}
  }{\seq{\usend{x}{y}{P}}{\ty{\Gamma},\tmty{x}{\tytens[\cs{o}]{A}{B}},\tmty{y}{\co{A}}}}
  \elabarrow
  \inferrule*{
    \inferrule*{
      \inferrule*{
      }{\seq{\link{y}{z}}{\tmty{y}{\co{A}},\tmty{z}{A}}}
      \\
      \seq{P}{\ty{\Gamma},\tmty{x}{B}}
    }{\seq{\ppar{\link{y}{z}}{P}}{\ty{\Gamma},\tmty{x}{B},\tmty{y}{\co{A}},\tmty{z}{A}}}
    \\
    \cs{o}<\minpr(\ty{\Gamma},\ty{A},\ty{B},\ty{\co{A}})
  }{\seq{\send{x}{z}{(\ppar{\link{y}{z}}{P})}}{\ty{\Gamma},\tmty{x}{\tytens[\cs{o}]{A}{B}},\tmty{y}{\co{A}}}}
\end{mathpar}

\subsection{Label Transition Semantics}

\subsubsection{Labels}
\[
  \tm{\ell}
  \coloneqq \tm{\lablink{x}{y}}
  \sep \tm{\labsend{x}{y}}
  \sep \tm{\labclose{x}}
  \sep \tm{\labrecv{x}{y}}
  \sep \tm{\labwait{x}}
  \sep \tm{\labselinl{x}}
  \sep \tm{\labselinr{x}}
  \sep \tm{\laboffinl{x}}
  \sep \tm{\laboffinr{x}}
\]
We let $\tm{\ell_{x}}$ range over actions on the name $\tm{x}$: $\tm{\lablink{x}{y}}$, $\tm{\labsend{x}{y}}$, $\tm{\labclose{x}}$, \etc.

\subsubsection{Action Rules}
\begin{mathpar}
  \begin{array}{llcl}
    \LabTirName{Act-Link\textsubscript{1}}
    & \tm{\link{x}{y}}
    & \lto{\lablink{x}{y}}
    & \tm{\halt}
    \\
    \LabTirName{Act-Send}
    & \tm{\send{x}{y}{P}}
    & \lto{\labsend{x}{y}}
    & \tm{P}
    \\
    \LabTirName{Act-Recv}
    & \tm{\recv{x}{y}{P}}
    & \lto{\labrecv{x}{y}}
    & \tm{P}
    \\
    \LabTirName{Act-Sel-Inl}
    & \tm{\inl{x}{P}}
    & \lto{\labselinl{x}}
    & \tm{P}
    \\
    \LabTirName{Act-Sel-Inr}
    & \tm{\inr{x}{P}}
    & \lto{\labselinr{x}}
    & \tm{P}
  \end{array}
  \qquad
  \begin{array}{llcl}
    \LabTirName{Act-Link\textsubscript{2}}
    & \tm{\link{x}{y}}
    & \lto{\lablink{y}{x}}
    & \tm{\halt}
    \\
    \LabTirName{Act-Close}
    & \tm{\close{x}{P}}
    & \lto{\labclose{x}}
    & \tm{P}
    \\
    \LabTirName{Act-Wait}
    & \tm{\wait{x}{P}}
    & \lto{\labwait{x}}
    & \tm{P}
    \\
    \LabTirName{Act-Off-Inl}
    & \tm{\offer{x}{P}{Q}}
    & \lto{\laboffinl{x}}
    & \tm{P}
    \\
    \LabTirName{Act-Off-Inr}
    & \tm{\offer{x}{P}{Q}}
    & \lto{\laboffinr{x}}
    & \tm{Q}
  \end{array}
\end{mathpar}

\subsubsection{Communication Rules}
\begin{mathpar}
  \inferrule*[lab=Tau-Link]{
    \tm{P}\lto{\lablink{x}{z}}\tm{P'}
  }{\tm{\res{x}{y}{P}}\tto\tm{\subst{P'}{z}{y}}}

  \inferrule*[lab=Tau-Send]{
    \tm{P}\lto{\ppar{\labsend{x}{x'}}{\labrecv{y}{y'}}}\tm{P'}
  }{\tm{\res{x}{y}{P}}\tto\tm{\res{x}{y}{\res{x'}{y'}{P'}}}}

  \inferrule*[lab=Tau-Close]{
    \tm{P}\lto{\ppar{\labclose{x}}{\labwait{y}}}\tm{P'}
  }{\tm{\res{x}{y}{P}}\tto\tm{P'}}

  \inferrule*[lab=Tau-Inl]{
    \tm{P}\lto{\ppar{\labselinl{x}}{\laboffinl{y}}}\tm{P'}
  }{\tm{\res{x}{y}{P}}\tto\tm{\res{x}{y}{P'}}}

  \inferrule*[lab=Tau-Inr]{
    \tm{P}\lto{\ppar{\labselinr{x}}{\laboffinr{y}}}\tm{P'}
  }{\tm{\res{x}{y}{P}}\tto\tm{\res{x}{y}{P'}}}
\end{mathpar}

\subsubsection{Structural Rules}
\begin{mathpar}
  \inferrule*[lab=Str-Rename]{
    \tm{P}\aleq\tm{Q}
    \\
    \tm{Q}\lto{\ell}\tm{R}
  }{\tm{P}\lto{\ell}\tm{R}}

  \inferrule*[lab=Str-Res]{
    \tm{P}\lto{\ell}\tm{P'}
    \\
    \tm{x},\tm{y}\not\in\fn(\tm{\ell})
  }{\tm{\res{x}{y}{P}}\lto{\ell}\tm{\res{x}{y}{P'}}}

  \inferrule*[lab=Str-Par\textsubscript{1}]{
    \tm{P}\lto{\ell}\tm{P'}
    \\
    \bn(\tm{\ell})\cap\fn(\tm{Q})=\varnothing
  }{\tm{\ppar{P}{Q}}\lto{\ell}\tm{\ppar{P'}{Q}}}

  \inferrule*[lab=Str-Par\textsubscript{2}]{
    \tm{Q}\lto{\ell}\tm{Q'}
    \\
    \bn(\tm{\ell})\cap\fn(\tm{P})=\varnothing
  }{\tm{\ppar{P}{Q}}\lto{\ell}\tm{\ppar{P}{Q'}}}

  \inferrule*[lab=Str-Syn]{
    \tm{P}\lto{\ell}\tm{P'}
    \\
    \tm{Q}\lto{\ell'}\tm{Q'}
    \\
    \bn(\ell)\cap\bn(\ell')=\varnothing
  }{\tm{\ppar{P}{Q}}\lto{\ell\parallel\ell'}\tm{\ppar{P'}{Q'}}}
\end{mathpar}

\subsection{Behavioural Theory}
\begin{definition}[Strong bisimilarity]
  A~symmetric relation $\rel{R}$ on processes is a strong bisimulation if $\tm{P}\mathbin{\rel{R}}\tm{Q}$ implies that if $\tm{P}\lto{\ell}\tm{P'}$, then $\tm{Q}\lto{\ell}\tm{Q'}$ for some $\tm{Q'}$ such that $\tm{P'}\mathbin{\rel{R}}\tm{Q'}$. Strong bisimilarity is the largest relation $\sbis$ that is a strong bisimulation.
\end{definition}
\begin{definition}[Saturated transition ($\slto{}$)]
  Saturated transition $\slto{}$ is the smallest relation such that: $\tm{P}\stto\tm{P}$ for all $\tm{P}$; and if $\tm{P}\stto\tm{P'}$, $\tm{P'}\lto{\ell}\tm{Q'}$, and $\tm{Q'}\stto\tm{Q}$, then $\tm{P}\slto{\ell}\tm{Q}$.
\end{definition}
\begin{definition}[Bisimilarity]
  A~symmetric relation $\rel{R}$ on processes is a bisimulation if $\tm{P}\mathbin{\rel{R}}\tm{Q}$ implies that if $\tm{P}\slto{\ell}\tm{P'}$, then $\tm{Q}\slto{\ell}\tm{Q'}$ for some $\tm{Q'}$ such that $\tm{P'}\mathbin{\rel{R}}\tm{Q'}$. Bisimilarity is the largest relation $\bis$ that is a bisimulation.
\end{definition}

\end{document}

%%% Local Variables:
%%% TeX-master: "priorities"
%%% End:

### 2020-08-05 priority-gv.tex
First commit of PGV paper.

\documentclass[main.tex]{subfiles}

\begin{document}

\section{Priority GV}
\usingnamespace{pgv}

\subsection{Static Syntax}%
\label{sec:pgv-syntax-static}

\subsubsection{Priorities, types, and environments}
\[
\begin{array}{lcl}
  \cs{o}, \cs{p}, \cs{q}, \cs{r}
  & \in & \mathbb{N}\cup\{\cs{\pbot},\cs{\ptop}\}
  \\
  \ty{S}
  & \coloneqq & \ty{\tysend[\cs{o}]{T}{S}}
    \sep        \ty{\tyrecv[\cs{o}]{T}{S}}
    \sep        \ty{\tyends[\cs{o}]}
    \sep        \ty{\tyendr[\cs{o}]}
  \\
  \ty{T}, \ty{U}
  & \coloneqq & \ty{\tyunit}
    \sep        \ty{\tyvoid}
    \sep        \ty{\typrod{T}{U}}
    \sep        \ty{\tysum{T}{U}}
    \sep        \ty{\tylolli[\cs{p},\cs{q}]{T}{U}}
    \sep        \ty{S}
  \\
  \ty{\Gamma}, \ty{\Delta}
  & \coloneqq & \ty{\emptyenv}
    \sep        \ty{\Gamma}, \tmty{x}{T}
\end{array}
\]
\[
\begin{array}{lcl}
  \ty{\co{\tysend[\cs{o}]{T}{S}}} & = & \ty{\tyrecv[\cs{o}]{T}{\co{S}}}
  \\
  \ty{\co{\tyrecv[\cs{o}]{T}{S}}} & = & \ty{\tysend[\cs{o}]{T}{\co{S}}}
\end{array}
\qquad
\begin{array}{lcl}
  \ty{\co{\tyends[\cs{o}]}} & = & \ty{\co{\tyendr[\cs{o}]}}
  \\
  \ty{\co{\tyendr[\cs{o}]}} & = & \ty{\co{\tyends[\cs{o}]}}
\end{array}
\]
\[
\begin{array}{lcl}
  \pr(\ty{\tysend[\cs{o}]{T}{S}}) & = & \cs{o} \\
  \pr(\ty{\tyrecv[\cs{o}]{T}{S}}) & = & \cs{o} \\
  \\
  \pr(\ty{\typrod{T}{U}}) & = & \pr({\ty{T}})\sqcap\pr({\ty{U}}) \\
  \pr(\ty{\tysum{T}{U}})  & = & \pr({\ty{T}})\sqcap\pr({\ty{U}}) \\
  \pr(\ty{\tylolli[\cs{p},\cs{q}]{T}{U}}) & = & \cs{p} \\
  \\
  \pr(\ty{\Gamma}, \tmty{x}{A}) & = & \pr(\ty{\Gamma})\sqcap\pr(\ty{A})
\end{array}
\qquad
\begin{array}{lcl}
  \pr(\ty{\tyends[\cs{o}]}) & = & \cs{o} \\
  \pr(\ty{\tyendr[\cs{o}]}) & = & \cs{o} \\
  \\
  \pr(\ty{\tyunit})         & = & \ptop \\
  \pr(\ty{\tyvoid})         & = & \ptop \\
  \pr(\ty{S})               & = & \pr(\ty{S}) \\
  \\
  \pr(\ty{\emptyenv})       & = & \ptop
\end{array}
\]

\subsubsection{Terms}%
\label{sec:pgv-syntax-terms}
\[
\begin{array}{lcl}
  \tm{L}, \tm{M}, \tm{N}
  & \coloneqq & \tm{x}
    \sep        \tm{K}
    \sep        \tm{\lambda x.M}
    \sep        \tm{M\;N} \\
  & \sep      & \tm{\unit}
    \sep        \tm{\letunit{M}{N}} \\
  & \sep      & \tm{\pair{M}{N}}
    \sep        \tm{\letpair{x}{y}{M}{N}} \\
  & \sep      & \tm{\inl{M}}
    \sep        \tm{\inr{M}}
    \sep        \tm{\casesum{L}{x}{M}{y}{N}} \\
  & \sep      & \tm{\absurd{M}} \\
  \tm{K}
  & \coloneqq & \tm{\link}
    \sep        \tm{\new}
    \sep        \tm{\spawn}
    \sep        \tm{\send}
    \sep        \tm{\recv}
    \sep        \tm{\wait}
    \sep        \tm{\close}
\end{array}
\]

\subsubsection{Syntactic Sugar}%
\label{sec:pgv-syntax-sugar}
We define the type of pure functions, \ie functions which perform no communication:
\[
  \ty{\tylolli{T}{U}} \elabarrow \ty{\tylolli[\cs{\ptop},\cs{\pbot}]{T}{U}}
\]
We define sequential composition, let-binding, and various pattern matching lambda abstractions:
\[
\begin{array}{lcl}
  \tm{\andthen{M}{N}}
  & \elabarrow & \tm{\letunit{M}{N}} \\
  \tm{\letbind{x}{M}{N}}
  & \elabarrow & \tm{(\lambda x.N)\;M} \\
  \tm{\lambda\unit.M}
  & \elabarrow & \tm{\lambda z.\letunit{z}{M}} \\
  \tm{\lambda\pair{x}{y}.M}
  & \elabarrow & \tm{\lambda z.\letpair{x}{y}{z}{M}}
\end{array}
\]
We define fork, which combines channel creation and thread spawning in one construct:
\[
\begin{array}{lcl}
  \tm{\fork}
  & \elabarrow & \tm{\lambda x.\letpair{y}{z}{\new}{\andthen{\spawn\;{(\lambda\unit.x\;y)}}{z}}}
\end{array}
\]
We define types for internal and external choice, and their corresponding communication constructs:
\[
\begin{array}{lcl}
  \ty{\tyselect[\cs{o}]{S}{S'}}
  & \elabarrow & \ty{\tysend[\cs{o}]{(\tysum{\co{S}}{\co{S'}})}{\tyends[\cs{o+1}]}} \\
  \ty{\tyoffer[\cs{o}]{S}{S'}}
  & \elabarrow & \ty{\tyrecv[\cs{o}]{(\tysum{S}{S'})}{\tyendr[\cs{o+1}]}}
\end{array}
\qquad
\begin{array}{lcl}
  \ty{\tyselectemp[\cs{o}]}
  & \elabarrow & \ty{\tysend[\cs{o}]{\tyvoid}{\tyends[\cs{o+1}]}} \\
  \ty{\tyofferemp[\cs{o}]}
  & \elabarrow & \ty{\tyrecv[\cs{o}]{\tyvoid}{\tyendr[\cs{o+1}]}} \\
\end{array}
\]
\[
\begin{array}{lcl}
  \tm{\select{\ell}}
  & \elabarrow & \tm{\lambda x.\fork\;{(\lambda y.\send\;{\pair{\ell\;y}{x}})}} \\
  \tm{\offer{L}{x}{M}{y}{N}}
  & \elabarrow & \tm{\letpair{z}{w}{\recv\;{L}}{\wait\;{w};}} \\
  &            & \quad\tm{\casesum{z}{x}{M}{y}{N}} \\
  \tm{\offeremp{L}}
  & \elabarrow & \tm{\letpair{z}{w}{\recv\;L}{\andthen{\wait\;{w}}{\absurd{z}}}}
\end{array}
\]

\subsection{Static Typing}%
\label{sec:pgv-typing-static}

\subsubsection{Typing Rules}
\begin{mathpar}
  \inferrule*[lab=T-Var]{
  }{\tseq[\cs{\pbot}]{\tmty{x}{T}}{x}{T}}

  \inferrule*[lab=T-Lam]{
    \tseq[\cs{o}]{\ty{\Gamma},\tmty{x}{T}}{M}{U}
  }{\tseq[\cs{\pbot}]{\ty{\Gamma}}{\lambda x.M}{\tylolli[\cs{\pr(\ty{\Gamma})},\cs{o}]{T}{U}}}

  \inferrule*[lab=T-Const]{
  }{\tseq[\cs{\pbot}]{\emptyenv}{K}{T}}

  \inferrule*[lab=T-App]{
    \tseq[\cs{p}]{\ty{\Gamma}}{M}{\tylolli[\cs{o},\cs{r}]{T}{U}}
    \\
    \tseq[\cs{q}]{\ty{\Delta}}{N}{T}
    \\
    \cs{p}<\pr(\ty{\Delta})
    \\
    \cs{q}<\cs{o}
  }{\tseq[\cs{p}\sqcup\cs{q}\sqcup\cs{o}]{\ty{\Gamma},\ty{\Delta}}{M\;N}{U}}
  \\
  \inferrule*[lab=T-Unit]{
  }{\tseq[\cs{\pbot}]{\emptyenv}{\unit}{\tyunit}}

  \inferrule*[lab=T-LetUnit]{
    \tseq[\cs{p}]{\ty{\Gamma}}{M}{\tyunit}
    \\
    \tseq[\cs{q}]{\ty{\Delta}}{N}{T}
    \\
    \cs{p}<\pr(\ty{\Delta})
  }{\tseq[\cs{p}\sqcup\cs{q}]{\ty{\Gamma},\ty{\Delta}}{\letunit{M}{N}}{T}}

  \inferrule*[lab=T-Pair]{
    \tseq[\cs{p}]{\ty{\Gamma}}{M}{T}
    \\
    \tseq[\cs{q}]{\ty{\Delta}}{N}{U}
    \\
    \cs{p}<\pr(\ty{\Delta})
  }{\tseq[\cs{p}\sqcup\cs{q}]{\ty{\Gamma},\ty{\Delta}}{\pair{M}{N}}{\typrod{T}{U}}}

  \inferrule*[lab=T-LetPair]{
    \tseq[\cs{p}]{\ty{\Gamma}}{M}{\typrod{T}{T'}}
    \\
    \tseq[\cs{q}]{\ty{\Delta},\tmty{x}{T},\tmty{y}{T'}}{N}{U}
    \\
    \cs{p}<\pr(\ty{\Delta},\ty{T},\ty{T'})
  }{\tseq[\cs{p}\sqcup\cs{q}]{\ty{\Gamma},\ty{\Delta}}{\letpair{x}{y}{M}{N}}{U}}

  \inferrule*[lab=T-Inl]{
    \tseq[o]{\ty{\Gamma}}{M}{T}
    \\
    \pr(\ty{T})=\pr(\ty{U})
  }{\tseq[o]{\ty{\Gamma}}{\inl{M}}{\tysum{T}{U}}}

  \inferrule*[lab=T-Inr]{
    \tseq[o]{\ty{\Gamma}}{M}{U}
    \\
    \pr(\ty{T})=\pr(\ty{U})
  }{\tseq[o]{\ty{\Gamma}}{\inr{M}}{\tysum{T}{U}}}

  \inferrule*[lab=T-CaseSum]{
    \tseq[\cs{p}]{\ty{\Gamma}}{L}{\tysum{T}{T'}}
    \\
    \tseq[\cs{q}]{\ty{\Delta},\tmty{x}{T}}{M}{U}
    \\
    \tseq[\cs{q}]{\ty{\Delta},\tmty{y}{T'}}{N}{U}
    \\
    \cs{p}<\pr(\ty{\Delta})
  }{\tseq[\cs{p}\sqcup\cs{q}]{\ty{\Gamma},\ty{\Delta}}{\casesum{L}{x}{M}{y}{N}}{U}}

  \inferrule*[lab=T-Absurd]{
    \tseq[o]{\ty{\Gamma}}{M}{\tyvoid}
  }{\tseq[o]{\ty{\Gamma},\ty{\Delta}}{\absurd{M}}{T}}
\end{mathpar}
We write $\tmty{K}{T}$ for $\tseq[\cs{\pbot}]{\emptyenv}{K}{T}$ in typing derivations.

\subsubsection{Type Schemas for Constants}
\begin{mathpar}
  \tmty{\link}{\tylolli{\typrod{S}{\co{S}}}{\tyunit}}

  \tmty{\new}{\tylolli{\tyunit}{\typrod{S}{\co{S}}}}

  \tmty{\spawn}{\tylolli{(\tylolli[\cs{p},\cs{q}]{\tyunit}{\tyunit})}{\tyunit}}
  \\
  \tmty{\send}{\tylolli[\cs{\ptop},\cs{o}]{\typrod{T}{\tysend[\cs{o}]{T}{S}}}{S}}

  \tmty{\recv}{\tylolli[\cs{\ptop},\cs{o}]{\tyrecv[\cs{o}]{T}{S}}{\typrod{T}{S}}}
  \\
  \tmty{\close}{\tylolli[\cs{\ptop},\cs{o}]{\tyends[\cs{o}]}{\tyunit}}

  \tmty{\wait}{\tylolli[\cs{\ptop},\cs{o}]{\tyendr[\cs{o}]}{\tyunit}}
\end{mathpar}

\subsubsection{Typing Rules for Syntactic Sugar}
\begin{mathpar}
  \inferrule*[lab=T-Seq]{
    \tseq[\cs{p}]{\ty{\Gamma}}{M}{\tyunit}
    \\
    \tseq[\cs{q}]{\ty{\Delta}}{N}{T}
    \\
    \cs{p}<\pr(\ty{\Delta})
  }{\tseq[\cs{p}\sqcup\cs{q}]{\ty{\Gamma},\ty{\Delta}}{\andthen{M}{N}}{T}}

  \inferrule*[lab=T-Let]{
    \tseq[\cs{p}]{\ty{\Gamma}}{M}{T}
    \\
    \tseq[\cs{q}]{\ty{\Delta},\tmty{x}{T}}{N}{U}
    \\
    \cs{p}<\pr(\ty{\Delta})
  }{\tseq[\cs{p}\sqcup\cs{q}]{\ty{\Gamma},\ty{\Delta}}{\letbind{x}{M}{N}}{U}}
  \\
  \inferrule*[lab=T-LamUnit]{
    {\tseq[\cs{o}]{\ty{\Gamma}}{M}{T}}
  }{\tseq[\cs{\pbot}]
    {\ty{\Gamma}}
    {\lambda\unit.M}
    {\tylolli[\cs{\pr(\ty{\Gamma})},\cs{o}]{\tyunit}{T}}}

  \inferrule*[lab=T-LamPair]
  {\tseq[\cs{o}]
    {\ty{\Gamma},\tmty{x}{T},\tmty{y}{T'}}
    {M}
    {U}}
  {\tseq[\cs{\pbot}]
    {\ty{\Gamma}}
    {\lambda\pair{x}{y}.M}
    {\tylolli[\cs{\pr(\ty{\Gamma})},\cs{o}]{\typrod{T}{T'}}{U}}}
  \\
  \inferrule*[lab=T-Fork]{
  }{\tseq[\cs{\pbot}]
    {\emptyenv}
    {\fork}
    {\tylolli[]{(\tylolli[\cs{p},\cs{q}]{S}{\tyunit})}{\co{S}}}}
  \\
  \inferrule*[lab=T-Select-Inl]{
    \pr(\ty{S})=\pr(\ty{S'})
  }{\tseq[\cs{\pbot}]{\emptyenv}{\select{\labinl}}{\tylolli[\cs{\ptop},\cs{o}]{\tyselect[\cs{o}]{S}{S'}}{S}}}

  \inferrule*[lab=T-Select-Inr]{
    \pr(\ty{S})=\pr(\ty{S'})
  }{\tseq[\cs{\pbot}]{\emptyenv}{\select{\labinr}}{\tylolli[\cs{\ptop},\cs{o}]{\tyselect[\cs{o}]{S}{S'}}{S'}}}

  \inferrule*[lab=T-Offer]{
    {\tseq[\cs{p}]
      {\ty{\Gamma}}
      {L}
      {\tyoffer[\cs{o}]{S}{S'}}}
    \\
    {\tseq[\cs{q}]
      {\ty{\Delta},\tmty{x}{S}}
      {M}
      {T}}
    \\
    {\tseq[\cs{q}]
      {\ty{\Delta},\tmty{y}{S'}}
      {N}
      {T}}
    \\
    \cs{o}\sqcup\cs{p}<\pr(\ty{\Delta},\ty{S},\ty{S'})
  }{\tseq[\cs{o}\sqcup\cs{p}\sqcup\cs{q}]
    {\ty{\Gamma},\ty{\Delta}}
    {\offer{L}{x}{M}{y}{N}}
    {T}}

  \inferrule*[lab=T-Offer-Absurd]{
    \tseq[\cs{p}]
    {\ty{\Gamma}}
    {L}
    {\tyofferemp[\cs{o}]}
    \\
    \cs{o}\sqcup\cs{p}<\pr(\ty{\Delta})
  }{\tseq[\cs{o}\sqcup\cs{p}]
    {\ty{\Gamma},\ty{\Delta}}
    {\offeremp{L}}
    {T}}
\end{mathpar}

\subsection{Runtime Syntax}%
\label{sec:pgv-syntax-runtime}

\subsubsection{Terms}
\[
\begin{array}[t]{lcl}
  \tm{\conf{C}}, \tm{\conf{D}}, \tm{\conf{E}}
  & \coloneqq & \tm{\phi\;M}
    \sep        \tm{\ppar{\conf{C}}{\conf{D}}}
    \sep        \tm{\res{x}{y}{\conf{C}}}
  \\
  \tm{\phi}
  & \coloneqq & \tm{\main}
    \sep        \tm{\child}
  \\
  \tm{V}, \tm{W}
  & \coloneqq & \tm{\lambda x.M}
    \sep        \tm{\unit}
    \sep        \tm{\pair{M}{N}}
    \sep        \tm{\inl{M}}
    \sep        \tm{\inr{M}}
  \\
  \tm{E}
  & \coloneqq & \tm{\hole} \\
  & \sep      & \tm{E\;M}
    \sep        \tm{V\;E} \\
  & \sep      & \tm{\letunit{E}{N}} \\
  & \sep      & \tm{\pair{E}{M}}
    \sep        \tm{\pair{V}{E}}
    \sep        \tm{\letpair{x}{y}{E}{M}} \\
  & \sep      & \tm{\inl{E}}
    \sep        \tm{\inr{E}}
    \sep        \tm{\casesum{E}{x}{M}{y}{N}} \\
  \\
  \tm{\conf{F}}
  & \coloneqq & \tm{\phi\;E}
  \\
  \tm{\conf{G}}
  & \coloneqq & \tm{\hole} \\
  & \sep      & \tm{\ppar{\conf{G}}{\conf{C}}}
    \sep        \tm{\res{x}{y}{\conf{G}}}
\end{array}
\]
\[
\begin{array}{ll}
  \tm{\main}  + \tm{\child} = \tm{\main}
  \\
  \tm{\child} + \tm{\main}  = \tm{\main}
\end{array}
\quad
\begin{array}{ll}
  \tm{\child} + \tm{\child} = \tm{\child}
  \\
  \tm{\main}  + \tm{\main} \; \text{undefined}
\end{array}
\]

\subsection{Runtime Typing}%
\label{sec:pgv-typing-runtime}
\begin{mathpar}
  \inferrule*[lab=T-Main]{
    \tseq[\cs{p}]{\ty{\Gamma}}{M}{T}
  }{\cseq[\main]{\ty{\Gamma}}{\main\;M}}

  \inferrule*[lab=T-Child]{
    \tseq[\cs{p}]{\ty{\Gamma}}{M}{\tyunit}
  }{\cseq[\child]{\ty{\Gamma}}{\child\;M}}

  \inferrule*[lab=T-Res]{
    \cseq[\phi]{\ty{\Gamma},\tmty{x}{S},\tmty{y}{\co{S}}}{\conf{C}}
  }{\cseq[\phi]{\ty{\Gamma}}{\res{x}{y}{\conf{C}}}}

  \inferrule*[lab=T-Par]{
    \cseq[\phi]{\ty{\Gamma}}{\conf{C}}
    \\
    \cseq[\phi']{\ty{\Delta}}{\conf{D}}
  }{\cseq[\phi+\phi']{\ty{\Gamma},\ty{\Delta}}{\ppar{\conf{C}}{\conf{D}}}}
\end{mathpar}

\subsection{Operational semantics}%
\label{sec:pgv-semantics}

\subsubsection{Term reduction}
\begin{mathpar}
  \begin{array}{llcl}
    \LabTirName{E-Lam}  & \tm{(\lambda x.M) \; V}
                        & \tred & \tm{\subst{M}{V}{x}}
    \\
    \LabTirName{E-Unit} & \tm{\letunit{\unit}{M}}
                        & \tred & \tm{M}
    \\
    \LabTirName{E-Pair} & \tm{\letpair{x}{y}{\pair{V}{W}}{M}}
                        & \tred & \tm{\subst{\subst{M}{V}{x}}{W}{y}}
    \\
    \LabTirName{E-Inl}  & \tm{\casesum{\inl{V}}{x}{M}{y}{N}}
                        & \tred & \tm{\subst{M}{V}{x}}
    \\
    \LabTirName{E-Inr}  & \tm{\casesum{\inr{V}}{x}{M}{y}{N}}
                        & \tred & \tm{\subst{N}{V}{y}}
  \end{array}
  \\
  \inferrule*[lab=E-Lift]{
    \tm{M}\tred\tm{M'}
  }{\tm{\plug{E}{M}}\tred\tm{\plug{E}{M'}}}
\end{mathpar}

\subsubsection{Structural congruence}
\[
\begin{array}{llcl}
  \LabTirName{E-LnkComm}  & \tm{\plug{\conf{F}}{\link\;{\pair{x}{y}}}}
                          & \equiv & \tm{\plug{\conf{F}}{\link\;{\pair{y}{x}}}}
  \\
  \LabTirName{E-ResComm}  & \tm{\res{x}{y}{\res{z}{w}{\conf{C}}}}
                          & \equiv & \tm{\res{z}{w}{\res{x}{y}{\conf{C}}}}
  \\
  \LabTirName{E-ParAssoc} & \tm{\ppar{\conf{C}}{(\ppar{\conf{D}}{\conf{E}})}}
                          & \equiv & \tm{\ppar{(\ppar{\conf{C}}{\conf{D}})}{\conf{E}}}
  \\
  \LabTirName{E-ParComm}  & \tm{\ppar{\conf{C}}{\conf{D}}}
                          & \equiv & \tm{\ppar{\conf{D}}{\conf{C}}}
  \\
  \LabTirName{E-Garbage}  & \tm{\ppar{\conf{C}}{\child{\unit}}}
                          & \equiv & \tm{\conf{C}}
  \\
  \LabTirName{E-ScopeExt} & \tm{\res{x}{y}{(\ppar{\conf{C}}{\conf{D}}})}
                          & \equiv & \tm{\ppar{\conf{C}}{\res{x}{y}{\conf{D}}}},
                                     \text{ if }{\tm{x},\tm{y}\notin\fv(\tm{\conf{C}})}
\end{array}
\]

\subsubsection{Configuration reduction}
\begin{mathpar}
  \begin{array}{llcl}
    \LabTirName{E-New}   & \tm{\plug{\conf{F}}{\new\;\unit}}
                         & \cred & \tm{\res{x}{y}{\plug{\conf{F}}{\pair{x}{y}}}}
    \\
    \LabTirName{E-Spawn} & \tm{\plug{\conf{F}}{\spawn\;V}}
                         & \cred & \tm{\ppar{\plug{\conf{F}}{\unit}}{\child{(V\;\unit)}}}
    \\
    \LabTirName{E-Send}  & \tm{\res{x}{y}{(\ppar
                           {\plug{\conf{F}}{\send\;{\pair{V}{x}}}}
                           {\plug{\conf{F'}}{\recv\;{y}}})}}
                         & \cred & \tm{\res{x}{y}{(\ppar
                                   {\plug{\conf{F}}{x}}
                                   {\plug{\conf{F'}}{\pair{V}{y}}})}}
    \\
    \LabTirName{E-Close} & \tm{\res{x}{y}{(\ppar
                           {\plug{\conf{F}}{\wait\;{x}}}
                           {\plug{\conf{F'}}{\close\;{y}}})}}
                         & \cred & \tm{\ppar{\plug{\conf{F}}{\unit}}{\plug{\conf{F'}}{\unit}}}
  \end{array}
  \\
  \inferrule*[lab=E-LiftC]{
    \tm{\conf{C}}\cred\tm{\conf{C'}}
  }{\tm{\plug{\conf{G}}{\conf{C}}}\cred\tm{\plug{\conf{G}}{\conf{C'}}}}

  \inferrule*[lab=E-LiftM]{
    \tm{M}\tred\tm{M'}
  }{\tm{\plug{\conf{F}}{M}}\tred\tm{\plug{\conf{F}}{M'}}}

  \inferrule*[lab=E-LiftE]{
    \tm{\conf{C}}\equiv\tm{\conf{C'}}
    \\
    \tm{\conf{C'}}\cred\tm{\conf{D'}}
    \\
    \tm{\conf{D'}}\equiv\tm{\conf{D}}
  }{\tm{\conf{C}}\cred\tm{\conf{D}}}
\end{mathpar}

\subsection{Metatheory}
\begin{restatable}{theorem}{thmpgvsubjectelaboration}[Subject Elaboration]
  If $\tseq[\cs{o}]{\ty{\Gamma}}{M}{T}$, $\tm{M}\elabarrow\tm{N}$, and $\ty{T}\elabarrow\ty{U}$,
  then $\tseq[\cs{o}]{\ty{\Gamma}}{N}{U}$.
\end{restatable}

\begin{restatable}{lemma}{lempgvvaluedone}
  \label{lem:value-done}
  If $\tseq[\cs{o}]{\ty{\Gamma}}{V}{T}$, then $\cs{o}=\cs{\pbot}$, and $\pr(\ty{\Gamma})=\pr(\ty{T})$.
\end{restatable}

\begin{restatable}{lemma}{lempgvsubstitution}[Substitution]
  \label{lem:pgv-substitution}
  If $\tseq[\cs{p}]{\ty{\Gamma},\tmty{x}{U'}}{M}{T}$ and $\tseq[\cs{q}]{\ty{\Theta}}{V}{U'}$, then $\tseq[\cs{p}]{\ty{\Gamma},\ty{\Theta}}{\subst{M}{V}{x}}{T}$.
\end{restatable}

\begin{restatable}{theorem}{thmpgvsubjectreductionterms}[Subject Reduction ($\tred$)]
  \label{thm:pgv-subject-reduction-terms}
  If $\tseq[\cs{o}]{\ty{\Gamma}}{M}{T}$ and $\tm{M}\tred\tm{M'}$,
  then $\tseq[\cs{o}]{\ty{\Gamma}}{M'}{T}$.
\end{restatable}

\begin{restatable}{theorem}{thmpgvsubjectcongruence}[Subject Congruence ($\equiv$)]
  \label{thm:pgv-subject-congruence}
  If $\cseq[\phi]{\ty{\Gamma}}{\conf{C}}$ and $\tm{\conf{C}}\equiv\tm{\conf{C'}}$,
  then $\cseq[\phi]{\ty{\Gamma}}{\conf{C'}}$.
\end{restatable}

\begin{restatable}{theorem}{thmpgvsubjectreductionconfs}[Subject Reduction ($\cred$)]
  \label{thm:pgv-subject-reduction-confs}
  If $\cseq[\phi]{\ty{\Gamma}}{\conf{C}}$ and $\tm{\conf{C}}\cred\tm{\conf{C'}}$,
  then $\cseq[\phi]{\ty{\Gamma}}{\conf{C'}}$.
\end{restatable}

\end{document}

%%% Local Variables:
%%% TeX-master: "priorities"
%%% End:

### 2020-09-11 prf/lem-pgv-value-done.tex
Fixed case environment.

@@
   By induction on the derivation of $\tseq[\cs{o}]{\ty{\Gamma}}{V}{T}$.
-  \begin{case}[\LabTirName{T-Lam}]
+  \begin{case*}[\LabTirName{T-Lam}]
     Immediately.
@@
     \end{mathpar}
-  \end{case}
-  \begin{case}[\LabTirName{T-Unit}]
+  \end{case*}
+  \begin{case*}[\LabTirName{T-Unit}]
     Immediately.
@@
     \end{mathpar}
-  \end{case}
-  \begin{case}[\LabTirName{T-Pair}]
+  \end{case*}
+  \begin{case*}[\LabTirName{T-Pair}]
     The induction hypotheses give us $\cs{p}=\cs{q}=\cs{\pbot}$, hence $\cs{p}\sqcup\cs{q}=\cs{\pbot}$, and $\pr(\ty{\Gamma})=\pr(\ty{T})$ and $\pr(\ty{\Delta})=\pr(\ty{U})$, hence $\pr(\ty{\Gamma},\ty{\Delta})=\pr(\ty{\Gamma})\sqcap\pr(\ty{\Delta})=\pr(\ty{T})\sqcap\pr(\ty{U})=\pr(\ty{\typrod{T}{U}})$.
@@
     \end{mathpar}
-  \end{case}
-  \begin{case}[\LabTirName{T-Inl}]
+  \end{case*}
+  \begin{case*}[\LabTirName{T-Inl}]
     The induction hypothesis gives us $\cs{o}=\cs{\pbot}$, and $\pr(\ty{\Gamma})=\pr{\ty{T}}$. We know $\pr(\ty{T})=\pr({\ty{U}})$, hence $\pr(\ty{\Gamma})=\pr(\ty{\tysum{T}{U}})$.
@@
     \end{mathpar}
-  \end{case}
-  \begin{case}[\LabTirName{T-Inr}]
+  \end{case*}
+  \begin{case*}[\LabTirName{T-Inr}]
     The induction hypothesis gives us $\cs{o}=\cs{\pbot}$, and $\pr(\ty{\Gamma})=\pr{\ty{U}}$. We know $\pr(\ty{T})=\pr({\ty{U}})$, hence $\pr(\ty{\Gamma})=\pr(\ty{\tysum{T}{U}})$.
@@
     \end{mathpar}
-  \end{case}
+  \end{case*}
 \end{proof}

### 2020-09-11 priority-cp.tex
Fixed case environment.

@@
 
-\subsection{Label Transition Semantics}
-
-\subsubsection{Labels}
-\[
-  \tm{\ell}
-  \coloneqq \tm{\lablink{x}{y}}
-  \sep \tm{\labsend{x}{y}}
-  \sep \tm{\labclose{x}}
-  \sep \tm{\labrecv{x}{y}}
-  \sep \tm{\labwait{x}}
-  \sep \tm{\labselinl{x}}
-  \sep \tm{\labselinr{x}}
-  \sep \tm{\laboffinl{x}}
-  \sep \tm{\laboffinr{x}}
-\]
-We let $\tm{\ell_{x}}$ range over actions on the name $\tm{x}$: $\tm{\lablink{x}{y}}$, $\tm{\labsend{x}{y}}$, $\tm{\labclose{x}}$, \etc.
-
-\subsubsection{Action Rules}
-\begin{mathpar}
-  \begin{array}{llcl}
-    \LabTirName{Act-Link\textsubscript{1}}
-    & \tm{\link{x}{y}}
-    & \lto{\lablink{x}{y}}
-    & \tm{\halt}
-    \\
-    \LabTirName{Act-Send}
-    & \tm{\send{x}{y}{P}}
-    & \lto{\labsend{x}{y}}
-    & \tm{P}
-    \\
-    \LabTirName{Act-Recv}
-    & \tm{\recv{x}{y}{P}}
-    & \lto{\labrecv{x}{y}}
-    & \tm{P}
-    \\
-    \LabTirName{Act-Sel-Inl}
-    & \tm{\inl{x}{P}}
-    & \lto{\labselinl{x}}
-    & \tm{P}
-    \\
-    \LabTirName{Act-Sel-Inr}
-    & \tm{\inr{x}{P}}
-    & \lto{\labselinr{x}}
-    & \tm{P}
-  \end{array}
-  \qquad
-  \begin{array}{llcl}
-    \LabTirName{Act-Link\textsubscript{2}}
-    & \tm{\link{x}{y}}
-    & \lto{\lablink{y}{x}}
-    & \tm{\halt}
-    \\
-    \LabTirName{Act-Close}
-    & \tm{\close{x}{P}}
-    & \lto{\labclose{x}}
-    & \tm{P}
-    \\
-    \LabTirName{Act-Wait}
-    & \tm{\wait{x}{P}}
-    & \lto{\labwait{x}}
-    & \tm{P}
-    \\
-    \LabTirName{Act-Off-Inl}
-    & \tm{\offer{x}{P}{Q}}
-    & \lto{\laboffinl{x}}
-    & \tm{P}
-    \\
-    \LabTirName{Act-Off-Inr}
-    & \tm{\offer{x}{P}{Q}}
-    & \lto{\laboffinr{x}}
-    & \tm{Q}
-  \end{array}
-\end{mathpar}
-
-\subsubsection{Communication Rules}
-\begin{mathpar}
-  \inferrule*[lab=Tau-Link]{
-    \tm{P}\lto{\lablink{x}{z}}\tm{P'}
-  }{\tm{\res{x}{y}{P}}\tto\tm{\subst{P'}{z}{y}}}
-
-  \inferrule*[lab=Tau-Send]{
-    \tm{P}\lto{\ppar{\labsend{x}{x'}}{\labrecv{y}{y'}}}\tm{P'}
-  }{\tm{\res{x}{y}{P}}\tto\tm{\res{x}{y}{\res{x'}{y'}{P'}}}}
-
-  \inferrule*[lab=Tau-Close]{
-    \tm{P}\lto{\ppar{\labclose{x}}{\labwait{y}}}\tm{P'}
-  }{\tm{\res{x}{y}{P}}\tto\tm{P'}}
-
-  \inferrule*[lab=Tau-Inl]{
-    \tm{P}\lto{\ppar{\labselinl{x}}{\laboffinl{y}}}\tm{P'}
-  }{\tm{\res{x}{y}{P}}\tto\tm{\res{x}{y}{P'}}}
-
-  \inferrule*[lab=Tau-Inr]{
-    \tm{P}\lto{\ppar{\labselinr{x}}{\laboffinr{y}}}\tm{P'}
-  }{\tm{\res{x}{y}{P}}\tto\tm{\res{x}{y}{P'}}}
-\end{mathpar}
-
-\subsubsection{Structural Rules}
-\begin{mathpar}
-  \inferrule*[lab=Str-Rename]{
-    \tm{P}\aleq\tm{Q}
-    \\
-    \tm{Q}\lto{\ell}\tm{R}
-  }{\tm{P}\lto{\ell}\tm{R}}
-
-  \inferrule*[lab=Str-Res]{
-    \tm{P}\lto{\ell}\tm{P'}
-    \\
-    \tm{x},\tm{y}\not\in\fn(\tm{\ell})
-  }{\tm{\res{x}{y}{P}}\lto{\ell}\tm{\res{x}{y}{P'}}}
-
-  \inferrule*[lab=Str-Par\textsubscript{1}]{
-    \tm{P}\lto{\ell}\tm{P'}
-    \\
-    \bn(\tm{\ell})\cap\fn(\tm{Q})=\varnothing
-  }{\tm{\ppar{P}{Q}}\lto{\ell}\tm{\ppar{P'}{Q}}}
-
-  \inferrule*[lab=Str-Par\textsubscript{2}]{
-    \tm{Q}\lto{\ell}\tm{Q'}
-    \\
-    \bn(\tm{\ell})\cap\fn(\tm{P})=\varnothing
-  }{\tm{\ppar{P}{Q}}\lto{\ell}\tm{\ppar{P}{Q'}}}
-
-  \inferrule*[lab=Str-Syn]{
-    \tm{P}\lto{\ell}\tm{P'}
-    \\
-    \tm{Q}\lto{\ell'}\tm{Q'}
-    \\
-    \bn(\ell)\cap\bn(\ell')=\varnothing
-  }{\tm{\ppar{P}{Q}}\lto{\ell\parallel\ell'}\tm{\ppar{P'}{Q'}}}
-\end{mathpar}
-
-\subsection{Behavioural Theory}
-\begin{definition}[Strong bisimilarity]
-  A~symmetric relation $\rel{R}$ on processes is a strong bisimulation if $\tm{P}\mathbin{\rel{R}}\tm{Q}$ implies that if $\tm{P}\lto{\ell}\tm{P'}$, then $\tm{Q}\lto{\ell}\tm{Q'}$ for some $\tm{Q'}$ such that $\tm{P'}\mathbin{\rel{R}}\tm{Q'}$.
-Strong bisimilarity is the largest relation $\sbis$ that is a strong bisimulation.
-\end{definition}
-\begin{definition}[Saturated transition ($\slto{}$)]
-  Saturated transition $\slto{}$ is the smallest relation such that: $\tm{P}\stto\tm{P}$ for all $\tm{P}$; and if $\tm{P}\stto\tm{P'}$, $\tm{P'}\lto{\ell}\tm{Q'}$, and $\tm{Q'}\stto\tm{Q}$, then $\tm{P}\slto{\ell}\tm{Q}$.
-\end{definition}
-\begin{definition}[Bisimilarity]
-  A~symmetric relation $\rel{R}$ on processes is a bisimulation if $\tm{P}\mathbin{\rel{R}}\tm{Q}$ implies that if $\tm{P}\slto{\ell}\tm{P'}$, then $\tm{Q}\slto{\ell}\tm{Q'}$ for some $\tm{Q'}$ such that $\tm{P'}\mathbin{\rel{R}}\tm{Q'}$.
-Bisimilarity is the largest relation $\bis$ that is a bisimulation.
-\end{definition}
-
 \end{document}

### 2020-09-14 priority-gv.tex
Fixed case environment.
Removed usage of amsthm, thmtools, and thm-restate.
Minor changes.

@@
   \\
+  \tm{R}
+  & \coloneqq & \tm{\link\;\pair{V}{W}}
+    \sep        \tm{\new}
+    \sep        \tm{\spawn\;V} \\
+  & \sep      & \tm{\send\;\pair{V}{W}}
+    \sep        \tm{\recv\;V}
+    \sep        \tm{\close\;V}
+    \sep        \tm{\wait\;V}
+  \\
   \tm{E}
@@
 \subsection{Metatheory}
-\begin{restatable}{theorem}{thmpgvsubjectelaboration}[Subject Elaboration]
-  If $\tseq[\cs{o}]{\ty{\Gamma}}{M}{T}$, $\tm{M}\elabarrow\tm{N}$, and $\ty{T}\elabarrow\ty{U}$,
-  then $\tseq[\cs{o}]{\ty{\Gamma}}{N}{U}$.
-\end{restatable}
 
-\begin{restatable}{lemma}{lempgvvaluedone}
-  \label{lem:value-done}
+\begin{restatablelemma}{lempgvvaluedone}
+  \label{lem:pgv-value-done}
   If $\tseq[\cs{o}]{\ty{\Gamma}}{V}{T}$, then $\cs{o}=\cs{\pbot}$, and $\pr(\ty{\Gamma})=\pr(\ty{T})$.
-\end{restatable}
+\end{restatablelemma}
+\begin{proof}
+  By induction on the derivation of $\tseq[\cs{o}]{\ty{\Gamma}}{V}{T}$.
+  (Details in \cref{prf:lem-pgv-value-done}.)
+\end{proof}
 
-\begin{restatable}{lemma}{lempgvsubstitution}[Substitution]
+\begin{restatablelemma}{lempgvsubstitution}[Substitution]
   \label{lem:pgv-substitution}
+  \hfill\\%newline before theorem statement
   If $\tseq[\cs{p}]{\ty{\Gamma},\tmty{x}{U'}}{M}{T}$ and $\tseq[\cs{q}]{\ty{\Theta}}{V}{U'}$, then $\tseq[\cs{p}]{\ty{\Gamma},\ty{\Theta}}{\subst{M}{V}{x}}{T}$.
-\end{restatable}
+\end{restatablelemma}
+\begin{proof}
+  By induction on the derivation of $\tseq[\cs{p}]{\ty{\Gamma},\tmty{x}{U'}}{M}{T}$.
+Whenever the contexts in the premises and conclusion differ, we use \cref{lem:pgv-value-done} to prove that the priorities are preserved.
+  (Details in \cref{prf:lem-pgv-substitution}.)
+\end{proof}
 
-\begin{restatable}{theorem}{thmpgvsubjectreductionterms}[Subject Reduction ($\tred$)]
-  \label{thm:pgv-subject-reduction-terms}
+\begin{restatablelemma}{lempgvsubjectreductionterms}[Subject Reduction, $\tred$]
+  \label{lem:pgv-subject-reduction-terms}
+  \hfill\\%newline before theorem statement
   If $\tseq[\cs{o}]{\ty{\Gamma}}{M}{T}$ and $\tm{M}\tred\tm{M'}$,
   then $\tseq[\cs{o}]{\ty{\Gamma}}{M'}{T}$.
-\end{restatable}
+\end{restatablelemma}
+\begin{proof}
+  By induction on the derivation of $\tm{M}\tred\tm{M'}$.
+  (Details in \cref{prf:lem-pgv-subject-reduction-terms}.)
+\end{proof}
 
-\begin{restatable}{theorem}{thmpgvsubjectcongruence}[Subject Congruence ($\equiv$)]
-  \label{thm:pgv-subject-congruence}
+\begin{restatablelemma}{lempgvsubjectcongruence}[Subject Congruence, $\equiv$]
+  \label{lem:pgv-subject-congruence}
+  \hfill\\%newline before theorem statement
   If $\cseq[\phi]{\ty{\Gamma}}{\conf{C}}$ and $\tm{\conf{C}}\equiv\tm{\conf{C'}}$,
   then $\cseq[\phi]{\ty{\Gamma}}{\conf{C'}}$.
-\end{restatable}
+\end{restatablelemma}
+\begin{proof}
+  By induction on the derivation of $\tm{\conf{C}}\equiv\tm{\conf{C'}}$.
+  (Details in \cref{prf:lem-pgv-subject-congruence}.)
+\end{proof}
 
-\begin{restatable}{theorem}{thmpgvsubjectreductionconfs}[Subject Reduction ($\cred$)]
+\begin{restatabletheorem}{thmpgvsubjectreductionconfs}[Subject Reduction, $\cred$]
   \label{thm:pgv-subject-reduction-confs}
+  \hfill\\%newline before theorem statement
   If $\cseq[\phi]{\ty{\Gamma}}{\conf{C}}$ and $\tm{\conf{C}}\cred\tm{\conf{C'}}$,
   then $\cseq[\phi]{\ty{\Gamma}}{\conf{C'}}$.
-\end{restatable}
+\end{restatabletheorem}
+\begin{proof}
+  By induction on the derivation of $\tm{\conf{C}}\cred\tm{\conf{C'}}$.
+  (Details in \cref{prf:thm-pgv-subject-reduction-confs}.)
+\end{proof}
+
+\begin{restatablelemma}{lempgvopenprogressterms}[Open Progress, $\tred$]
+  \label{lem:pgv-open-progress-terms}
+  If $\tseq[\cs{o}]{\ty{\Gamma}}{M}{T}$, then either:
+  \begin{description}[labelwidth=8ex]
+  \item[Done]
+    $\tm{M}$ is a value;
+  \item[Step]
+    There exists some $\tm{N}$ such that $\tm{M}\tred\tm{N}$; or
+  \item[Blocked] \label{lem:pgv-progress-terms:blocked}
+    There exists some $\tm{E}$ and $\tm{R}$ such that $\tm{M}=\tm{\plug{E}{R}}$.
+  \end{description}
+\end{restatablelemma}
+\begin{proof}
+  \todo{Write proof.}
+\end{proof}
+
+\begin{restatabletheorem}{thmpgvopenprogressconfs}[Open Progress, $\cred$]
+  \label{thm:pgv-open-progress-confs}
+  If $\cseq[\phi]{\ty{\Gamma}}{\conf{C}}$, then either:
+  \begin{description}[labelwidth=8ex]
+  \item[Done]
+    There exists some $\tm{V}$ such that $\tm{\conf{C}}=\tm{\main\;V}$;
+  \item[Step]
+    There exists some $\tm{\conf{D}}$ such that $\tm{\conf{C}} \cred \tm{\conf{D}}$; or
+  \item[Blocked]
+    There exist $\tm{x_i}$ and $\tm{y_i}$ for $0\leq{i}\leq{n}$, and $\tm{\phi_j}$ and $\tm{R_j}$ for $0\leq{j}\leq{m}$, such that $\tm{\conf{C}}=\tm{\res{x_1y_1\dots x_ny_n}{}{(\phi_1R_1 \parallel\dots\parallel \phi_mR_m)}}$ and for each $\tm{R_j}$ there exists some $\tm{a_j}\in\fn(\ty{\Gamma})$ such that $\blocked(\tm{a_j},\tm{R_j})$.
+  \end{description}
+\end{restatabletheorem}
+\begin{proof}
+  \todo{Write proof.}
+\end{proof}
 

### 2020-09-16 priority-cp.tex
Finished theory skeleton.
Renamed SC rules, added PCP-SC.
Added reduction for PCP.
Added PCP reduction.
quick general pass to the paper
Proved lemma. Renamed rules.
some text on types

@@
 
+\begin{figure}[t]
+  \centering
+  \includegraphics[width=0.8\columnwidth]{scheduler}
+  \vspace{-4mm}
+  \caption{Cyclic Scheduler from \cite{DardhaG18}}
+  \label{fig:scheduler}
+  \vspace{-4mm}
+  \end{figure}
+
+In this section we present an updated version of Priority CP \cite{DardhaG18}, which is Wadler's Classical Processes \cite{W12} with \emph{priorities}.
+Some details of the calculus are inspired by Carbone \emph{et al.} \cite{CLMSW16} and Caires and P\'{e}rez \cite{CP17}.
+
 \subsection{Syntax}
-
 \subsubsection{Priorities, types, and environments}
+Types are based on CLL and range over $\ty{A}, \ty{B}, \ty{C}$.
+Types are annotated with priorities which range over $\cs{o} \in \mathbb{N} \sqcup \{\omega\}$.
+Let $\omega$ be a special element such that $\cs{o}<\omega$ for all $\cs{o}\in \mathbb{N}$.
+Often, we will omit $\omega$ when not necessary.
+\todo{check the omega thing is also part of our presentation}
+Types $\ty{\tyone[\cs{o}]}$ and $\ty{\tybot[\cs{o}]}$
+are used to type channel endpoints which are ready to be closed.
+Types $\ty{\tynil[\cs{o}]}$ and $\ty{\tytop[\cs{o}]}$ are used to... 
+Type $\ty{\tytens[\cs{o}]{A}{B}}$ (respectively, $\ty{\typarr[\cs{o}]{A}{B}}$) is used to type a channel endpoint that first {outputs} (respectively, {inputs}) a channel of type $\ty A$ and then proceeds as $\ty B$.
+$\ty{\typlus[\cs{o}]{A}{B}}$ is used to type a channel endpoint over which we can {select} either the \emph{left} branch and proceed as $\ty A$ or the \emph{right} branch and proceed as $\ty B$.
+$\ty{\tywith[\cs{o}]{A}{B}}$ is used to type a channel endpoint that can offer a both $\ty A$ and $\ty B$.
+A typing context $\ty\Gamma$ is a partial function associating channel names to types.
 \[
@@
 \]
+
+Duality on types is a total function and is defined below. It preserves priorities of types.
 \[
@@
 \]
+
+Finally, function $\pr(\cdot)$ defined over types returns the priority of that type; function $\minpr(\cdot)$ defined over typing contexts returns the \emph{minimum} priority of all types in that context.
+Both functions are defined as expected below.
 \[
@@
 
+\subsection{Operational Semantics}
+
+\subsubsection{Structural congruence}
+\begin{mathpar}
+  \begin{array}{llcl}
+    \LabTirName{SC-LnkSwp}   & \tm{\link{x}{y}}
+                             & \equiv & \tm{\link{y}{x}}
+    \\
+    \LabTirName{SC-ResLnk}   & \tm{\res{x}{y}{\link{x}{y}}}
+                             & \equiv & \tm{\halt}
+    \\
+    \LabTirName{SC-ResExt}   & \tm{\res{x}{y}{(\ppar{P}{Q})}}
+                             & \equiv & \tm{\ppar{\res{x}{y}{P}}{Q}},
+                                        \text{ if }{\tm{x},\tm{y}\notin\fv(\tm{Q})}
+    \\
+    \LabTirName{SC-ResSwp}   & \tm{\res{x}{y}{P}}
+                             & \equiv & \tm{\res{y}{x}{P}}
+    \\
+    \LabTirName{SC-ResCom}   & \tm{\res{x}{y}{\res{z}{w}{P}}}
+                             & \equiv & \tm{\res{z}{w}{\res{x}{y}{P}}}
+    \\
+    \LabTirName{SC-ParNil}   & \tm{\ppar{P}{\halt}}
+                             & \equiv & \tm{P}
+    \\
+    \LabTirName{SC-ParCom}   & \tm{\ppar{P}{Q}}
+                             & \equiv & \tm{\ppar{Q}{P}}
+    \\
+    \LabTirName{SC-ParAsc}   & \tm{\ppar{P}{(\ppar{Q}{R})}}
+                             & \equiv & \tm{\ppar{(\ppar{P}{Q})}{R}}
+  \end{array}
+\end{mathpar}
+
+\subsubsection{Reduction}
+\begin{mathpar}
+  \begin{array}{llcl}
+    \LabTirName{E-Lnk}     & \tm{\res{x}{y}{(\ppar{\link{w}{x}}{P})}}
+                           & \red & \tm{\subst{P}{w}{x}}
+    \\
+    \LabTirName{E-Snd}     & \tm{\res{x}{y}{(\ppar{\send{x}{z}{P}}{\recv{x}{w}{Q}})}}
+                           & \red & \tm{\res{x}{y}{\res{z}{w}{(\ppar{P}{Q})}}}
+    \\
+    \LabTirName{E-End}     & \tm{\res{x}{y}{(\ppar{\close{x}{P}}{\wait{y}{Q}})}}
+                           & \red & \tm{\ppar{P}{Q}}
+    \\
+    \LabTirName{E-Sel\textsubscript{1}}
+                           & \tm{\res{x}{y}{(\ppar{\inl{x}{P}}{\offer{x}{Q}{R}})}}
+                           & \red & \tm{\res{x}{y}{(\ppar{P}{Q})}}
+    \\
+    \LabTirName{E-Sel\textsubscript{2}}
+                           & \tm{\res{x}{y}{(\ppar{\inr{x}{P}}{\offer{x}{Q}{R}})}}
+                           & \red & \tm{\res{x}{y}{(\ppar{P}{R})}}
+  \end{array}
+  \\
+  \inferrule*[lab=E-Res]{
+    \tm{P}\red\tm{P'}
+  }{\tm{\res{x}{y}{P}}\red\tm{\res{x}{y}{P'}}}
+
+  \inferrule*[lab=E-Par]{
+    \tm{P}\red\tm{P'}
+  }{\tm{\ppar{P}{Q}}\red\tm{\ppar{P'}{Q}}}
+
+  \inferrule*[lab=E-LiftSC]{
+    \tm{P}\equiv\tm{P'}
+    \\
+    \tm{P'}\red\tm{Q'}
+    \\
+    \tm{Q'}\equiv\tm{Q}
+  }{\tm{P}\red\tm{Q}}
+
+\end{mathpar}
+
 \end{document}

### 2020-09-16 priority-gv.tex
Renamed SC rules, added PCP-SC.
Added reduction for PCP.
Minor changes.

@@
 \subsubsection{Structural congruence}
-\[
+\begin{mathpar}
   \begin{array}{llcl}
-  \LabTirName{E-LnkComm}  & \tm{\plug{\conf{F}}{\link\;{\pair{x}{y}}}}
+    \LabTirName{SC-LinkSwap}   & \tm{\plug{\conf{F}}{\link\;{\pair{x}{y}}}}
                                & \equiv & \tm{\plug{\conf{F}}{\link\;{\pair{y}{x}}}}
     \\
-  \LabTirName{E-ResComm}  & \tm{\res{x}{y}{\res{z}{w}{\conf{C}}}}
+    \LabTirName{SC-ResSwap}    & \tm{\res{x}{y}{\conf{C}}}
+                               & \equiv & \tm{\res{y}{x}{\conf{C}}}
+    \\
+    \LabTirName{SC-ResComm}    & \tm{\res{x}{y}{\res{z}{w}{\conf{C}}}}
                                & \equiv & \tm{\res{z}{w}{\res{x}{y}{\conf{C}}}}
     \\
-  \LabTirName{E-ParAssoc} & \tm{\ppar{\conf{C}}{(\ppar{\conf{D}}{\conf{E}})}}
-                          & \equiv & \tm{\ppar{(\ppar{\conf{C}}{\conf{D}})}{\conf{E}}}
+    \LabTirName{SC-ResExt}     & \tm{\res{x}{y}{(\ppar{\conf{C}}{\conf{D}}})}
+                               & \equiv & \tm{\ppar{\conf{C}}{\res{x}{y}{\conf{D}}}},
+                                          \text{ if }{\tm{x},\tm{y}\notin\fv(\tm{\conf{C}})}
     \\
-  \LabTirName{E-ParComm}  & \tm{\ppar{\conf{C}}{\conf{D}}}
+    \LabTirName{SC-ParNil}     & \tm{\ppar{\conf{C}}{\child{\unit}}}
+                               & \equiv & \tm{\conf{C}}
+    \\
+    \LabTirName{SC-ParComm}    & \tm{\ppar{\conf{C}}{\conf{D}}}
                                & \equiv & \tm{\ppar{\conf{D}}{\conf{C}}}
     \\
-  \LabTirName{E-Garbage}  & \tm{\ppar{\conf{C}}{\child{\unit}}}
-                          & \equiv & \tm{\conf{C}}
-  \\
-  \LabTirName{E-ScopeExt} & \tm{\res{x}{y}{(\ppar{\conf{C}}{\conf{D}}})}
-                          & \equiv & \tm{\ppar{\conf{C}}{\res{x}{y}{\conf{D}}}},
-                                     \text{ if }{\tm{x},\tm{y}\notin\fv(\tm{\conf{C}})}
+    \LabTirName{SC-ParAssoc}   & \tm{\ppar{\conf{C}}{(\ppar{\conf{D}}{\conf{E}})}}
+                               & \equiv & \tm{\ppar{(\ppar{\conf{C}}{\conf{D}})}{\conf{E}}}
   \end{array}
-\]
+\end{mathpar}
 
@@
     \LabTirName{E-New}   & \tm{\plug{\conf{F}}{\new\;\unit}}
-                         & \cred & \tm{\res{x}{y}{\plug{\conf{F}}{\pair{x}{y}}}}
+                         & \cred & \tm{\res{x}{y}{(\plug{\conf{F}}{\pair{x}{y}})}}
     \\
     \LabTirName{E-Spawn} & \tm{\plug{\conf{F}}{\spawn\;V}}
-                         & \cred & \tm{\ppar{\plug{\conf{F}}{\unit}}{\child{(V\;\unit)}}}
+                         & \cred & \tm{\ppar{\plug{\conf{F}}{\unit}}{\child\;V\;\unit}}
     \\
@@
 
-  \inferrule*[lab=E-LiftE]{
+  \inferrule*[lab=E-LiftSC]{
     \tm{\conf{C}}\equiv\tm{\conf{C'}}
@@
     There exists some $\tm{N}$ such that $\tm{M}\tred\tm{N}$; or
-  \item[Blocked] \label{lem:pgv-progress-terms:blocked}
+  \item[Blocked]
     There exists some $\tm{E}$ and $\tm{R}$ such that $\tm{M}=\tm{\plug{E}{R}}$.
@@
   \item[Done]
-    There exists some $\tm{V}$ such that $\tm{\conf{C}}=\tm{\main\;V}$;
+    There exists some $\tm{V}$ such that $\tm{\conf{C}}=\tm{\phi\;V}$;
   \item[Step]

### 2020-09-17 priority-gv.tex
Finished OC-Completeness.
Extended congruence, finished proof of SC correspondence.
Updated translation to configurations so OC is complete.
Restore absurd, many other changes. I should be better at atomic commits.

@@
 \label{sec:pgv-syntax-sugar}
-We define the type of pure functions, \ie functions which perform no communication:
-\[
-  \ty{\tylolli{T}{U}} \elabarrow \ty{\tylolli[\cs{\ptop},\cs{\pbot}]{T}{U}}
-\]
-We define sequential composition, let-binding, and various pattern matching lambda abstractions:
+We define the type of pure functions, \ie functions which perform no communication, and types for internal and external choice.
+Furthermore, we define term constructs for sequential composition, let-binding, various pattern matching lambda abstractions, fork, which combines channel creation and thread spawning in one construct, and internal and external choice.
 \[
 \begin{array}{lcl}
+  \ty{\tylolli{T}{U}}
+  & \elabarrow & \ty{\tylolli[\cs{\ptop},\cs{\pbot}]{T}{U}} \\
+  \ty{\tyselect[\cs{o}]{S}{S'}}
+  & \elabarrow & \ty{\tysend[\cs{o}]{(\tysum{\co{S}}{\co{S'}})}{\tyends[\cs{o+1}]}} \\
+  \ty{\tyoffer[\cs{o}]{S}{S'}}
+  & \elabarrow & \ty{\tyrecv[\cs{o}]{(\tysum{S}{S'})}{\tyendr[\cs{o+1}]}} \\
+  \ty{\tyselectemp[\cs{o}]}
+  & \elabarrow & \ty{\tysend[\cs{o}]{\tyvoid}{\tyends[\cs{o+1}]}} \\
+  \ty{\tyofferemp[\cs{o}]}
+  & \elabarrow & \ty{\tyrecv[\cs{o}]{\tyvoid}{\tyendr[\cs{o+1}]}} \\
+  \\
   \tm{\andthen{M}{N}}
@@
   \tm{\lambda\pair{x}{y}.M}
-  & \elabarrow & \tm{\lambda z.\letpair{x}{y}{z}{M}}
-\end{array}
-\]
-We define fork, which combines channel creation and thread spawning in one construct:
-\[
-\begin{array}{lcl}
+  & \elabarrow & \tm{\lambda z.\letpair{x}{y}{z}{M}} \\
   \tm{\fork}
-  & \elabarrow & \tm{\lambda x.\letpair{y}{z}{\new}{\andthen{\spawn\;{(\lambda\unit.x\;y)}}{z}}}
-\end{array}
-\]
-We define types for internal and external choice, and their corresponding communication constructs:
-\[
-\begin{array}{lcl}
-  \ty{\tyselect[\cs{o}]{S}{S'}}
-  & \elabarrow & \ty{\tysend[\cs{o}]{(\tysum{\co{S}}{\co{S'}})}{\tyends[\cs{o+1}]}} \\
-  \ty{\tyoffer[\cs{o}]{S}{S'}}
-  & \elabarrow & \ty{\tyrecv[\cs{o}]{(\tysum{S}{S'})}{\tyendr[\cs{o+1}]}}
-\end{array}
-\qquad
-\begin{array}{lcl}
-  \ty{\tyselectemp[\cs{o}]}
-  & \elabarrow & \ty{\tysend[\cs{o}]{\tyvoid}{\tyends[\cs{o+1}]}} \\
-  \ty{\tyofferemp[\cs{o}]}
-  & \elabarrow & \ty{\tyrecv[\cs{o}]{\tyvoid}{\tyendr[\cs{o+1}]}} \\
-\end{array}
-\]
-\[
-\begin{array}{lcl}
+  & \elabarrow & \tm{\lambda x.\letpair{y}{z}{\new}{\andthen{\spawn\;{(\lambda\unit.x\;y)}}{z}}} \\
   \tm{\select{\ell}}
-  & \elabarrow & \tm{\lambda x.\fork\;{(\lambda y.\send\;{\pair{\ell\;y}{x}})}} \\
-  \tm{\offer{L}{x}{M}{y}{N}}
-  & \elabarrow & \tm{\letpair{z}{w}{\recv\;{L}}{\wait\;{w};}} \\
-  &            & \quad\tm{\casesum{z}{x}{M}{y}{N}} \\
+  & \elabarrow & \tm{\lambda x.\letpair{y}{z}{\new}{\andthen{\close\;(\send\;{\pair{\ell\;y}{x}})}{z}}} \\
+  \multicolumn{3}{l}{\tm{\offer{L}{x}{M}{y}{N}}\elabarrow} \\
+  \multicolumn{3}{l}{
+  \qquad\tm{\letpair{z}{w}{\recv\;{L}}{\andthen{\wait\;{w}}{\casesum{z}{x}{M}{y}{N}}}}} \\
   \tm{\offeremp{L}}
@@
     \\
+    \LabTirName{SC-ResLink}    & \tm{\res{x}{y}{(\phi\;\link\;\pair{x}{y})}}
+                               & \equiv & \tm{\phi\;\unit}
+    \\
     \LabTirName{SC-ResSwap}    & \tm{\res{x}{y}{\conf{C}}}
@@
   \begin{array}{llcl}
+    \LabTirName{E-Link}  & \tm{\res{x}{y}{(\ppar{\plug{\conf{F}}{\link\;\pair{w}{x}}}{\conf{C}})}}
+                         & \cred & \tm{\ppar{\plug{\conf{F}}{\unit}}{\subst{\conf{C}}{w}{y}}}
+    \\
     \LabTirName{E-New}   & \tm{\plug{\conf{F}}{\new\;\unit}}
@@
 \begin{proof}
-  \todo{Write proof.}
+  \admit
 \end{proof}
@@
   \item[Blocked]
-    There exist $\tm{x_i}$ and $\tm{y_i}$ for $0\leq{i}\leq{n}$, and $\tm{\phi_j}$ and $\tm{R_j}$ for $0\leq{j}\leq{m}$, such that $\tm{\conf{C}}=\tm{\res{x_1y_1\dots x_ny_n}{}{(\phi_1R_1 \parallel\dots\parallel \phi_mR_m)}}$ and for each $\tm{R_j}$ there exists some $\tm{a_j}\in\fn(\ty{\Gamma})$ such that $\blocked(\tm{a_j},\tm{R_j})$.
+    We have $\tm{\conf{C}}\equiv\tm{\res{x_1y_1\dots x_ny_n}{}{(\phi_1R_1 \parallel\dots\parallel \phi_mR_m)}}$,\\
+    and for some $\tm{R_j}$ and $\tm{a_j}\in\fn(\ty{\Gamma})$, $\blocked(\tm{a_j},\tm{R_j})$.
   \end{description}

### 2020-09-17 priority-gv.tex
Reorganised paper.

@@
 \label{sec:pgv-syntax-sugar}
-We define the type of pure functions, \ie functions which perform no communication, and types for internal and external choice.
-Furthermore, we define term constructs for sequential composition, let-binding, various pattern matching lambda abstractions, fork, which combines channel creation and thread spawning in one construct, and internal and external choice.
+
 \[

### 2020-09-17 priority-cp.tex
Finished OC-Completeness.
Restore absurd, many other changes. I should be better at atomic commits.
Restructuring.

--- 2020-09-11 priority-cp.tex
@@
 
-\subsection{Syntax}
+\begin{figure}[t]
+  \centering
+  \includegraphics[width=0.8\columnwidth]{scheduler}
+  \vspace{-4mm}
+  \caption{Cyclic Scheduler from \cite{DardhaG18}}
+  \label{fig:scheduler}
+  \vspace{-4mm}
+  \end{figure}
 
-\subsubsection{Priorities, types, and environments}
+In this section we present an updated version of Priority CP~\cite{DardhaG18}, which is Wadler's Classical Processes~\cite{W12} with \emph{priorities}.
+Some details of the calculus are inspired by Carbone \emph{et al.}~\cite{CLMSW16} and Caires and P\'{e}rez~\cite{CP17}.
+
+\paragraph*{Priorities, types, and environments}
+Types are based on CLL and range over $\ty{A}, \ty{B}, \ty{C}$.
+Types are annotated with priorities which range over $\cs{o} \in \mathbb{N} \sqcup \{\omega\}$.
+Let $\omega$ be a special element such that $\cs{o}<\omega$ for all $\cs{o}\in \mathbb{N}$.
+Often, we will omit $\omega$ when not necessary.
+\todo{check the omega thing is also part of our presentation} Types $\ty{\tyone[\cs{o}]}$ and $\ty{\tybot[\cs{o}]}$ are used to type channel endpoints which are ready to be closed.
+Types $\ty{\tynil[\cs{o}]}$ and $\ty{\tytop[\cs{o}]}$ are used to...
+Type $\ty{\tytens[\cs{o}]{A}{B}}$ (respectively, $\ty{\typarr[\cs{o}]{A}{B}}$) is used to type a channel endpoint that first {outputs} (respectively, {inputs}) a channel of type $\ty A$ and then proceeds as $\ty B$.
+$\ty{\typlus[\cs{o}]{A}{B}}$ is used to type a channel endpoint over which we can {select} either the \emph{left} branch and proceed as $\ty A$ or the \emph{right} branch and proceed as $\ty B$.
+$\ty{\tywith[\cs{o}]{A}{B}}$ is used to type a channel endpoint that can offer a both $\ty A$ and $\ty B$.
+A typing context $\ty\Gamma$ is a partial function associating channel names to types.
 \[
@@
 
-\subsubsection{Terms}
+\paragraph*{Terms}
 \[
@@
 
-\subsubsection{Syntactic Sugar}
+\paragraph*{Syntactic Sugar}
 We define \emph{unbound} send:
@@
 \subsection{Typing}
+\input{fig/pcp-typing}
 
-\subsubsection{Typing Rules}
-\begin{mathpar}
-  \inferrule*[lab=T-Link]{
-  }{\seq{\link[\ty{A}]{x}{y}}{\tmty{x}{A},\tmty{y}{\co{A}}}}
-
-  \inferrule*[lab=T-New]{
-    \seq{P}{\ty{\Gamma},\tmty{x}{A},\tmty{y}{\co{A}}}
-  }{\seq{\res{x}{y}{P}}{\ty{\Gamma}}}
-  \\
-  \inferrule*[lab=T-Par]{
-    \seq{P}{\ty{\Gamma}}
-    \\
-    \seq{Q}{\ty{\Delta}}
-  }{\seq{\ppar{P}{Q}}{\ty{\Gamma},\ty{\Delta}}}
-
-  \inferrule*[lab=T-Halt]{
-  }{\seq{\halt}{\emptyenv}}
-  \\
-  \inferrule*[lab=T-Send]{
-    \seq{P}{\ty{\Gamma},\tmty{y}{A},\tmty{x}{B}}
-    \\
-    \cs{o}<\minpr(\ty{\Gamma},\ty{A},\ty{B})
-  }{\seq{\send{x}{y}{P}}{\ty{\Gamma},\tmty{x}{\tytens[\cs{o}]{A}{B}}}}
-
-  \inferrule*[lab=T-Close]{
-    \seq{P}{\ty{\Gamma}}
-    \\
-    \cs{o}<\minpr(\ty{\Gamma})
-  }{\seq{\close{x}{P}}{\ty{\Gamma},\tmty{x}{\tyone[\cs{o}]}}}
-  \\
-  \inferrule*[lab=T-Recv]{
-    \seq{P}{\ty{\Gamma},\tmty{y}{A},\tmty{x}{B}}
-    \\
-    \cs{o}<\minpr(\ty{\Gamma},\ty{A},\ty{B})
-  }{\seq{\recv{x}{y}{P}}{\ty{\Gamma},\tmty{x}{\typarr[\cs{o}]{A}{B}}}}
-
-  \inferrule*[lab=T-Wait]{
-    \seq{P}{\ty{\Gamma}}
-    \\
-    \cs{o}<\minpr(\ty{\Gamma})
-  }{\seq{\wait{x}{P}}{\ty{\Gamma},\tmty{x}{\tybot[\cs{o}]}}}
-  \\
-  \inferrule*[lab=T-Select-Inl]{
-    \seq{P}{\ty{\Gamma},\tmty{x}{A}}
-    \\
-    \cs{o}<\minpr(\ty{\Gamma},\ty{A},\ty{B})
-    \\
-    \pr(\ty{A})=\pr(\ty{B})
-  }{\seq{\inl{x}{P}}{\ty{\Gamma},\tmty{x}{\typlus[\cs{o}]{A}{B}}}}
-
-  \inferrule*[lab=T-Select-Inr]{
-    \seq{P}{\ty{\Gamma},\tmty{x}{B}}
-    \\
-    \cs{o}<\minpr(\ty{\Gamma},\ty{A},\ty{B})
-    \\
-    \pr(\ty{A})=\pr(\ty{B})
-  }{\seq{\inr{x}{P}}{\ty{\Gamma},\tmty{x}{\typlus[\cs{o}]{A}{B}}}}
-
-  \inferrule*[lab=T-Offer]{
-    \seq{P}{\ty{\Gamma},\tmty{x}{A}}
-    \\
-    \seq{Q}{\ty{\Gamma},\tmty{x}{B}}
-    \\
-    \cs{o}<\minpr(\ty{\Gamma},\ty{A},\ty{B})
-  }{\seq{\offer{x}{P}{Q}}{\ty{\Gamma},\tmty{x}{\tywith[\cs{o}]{A}{B}}}}
-
-  \inferrule*[lab=T-Offer-Absurd]{
-    \cs{o}<\pr(\ty{\Gamma})
-  }{\seq{\absurd{x}}{\ty{\Gamma},\tmty{x}{\tytop[\cs{o}]}}}
-\end{mathpar}
-
-\subsubsection{Typing Rules for Syntactic Sugar}
-\begin{mathpar}
-  \inferrule*[lab=T-Unbound-Send]{
-    \seq{P}{\ty{\Gamma},\tmty{x}{B}}
-  }{\seq{\usend{x}{y}{P}}{\ty{\Gamma},\tmty{x}{\tytens[\cs{o}]{A}{B}},\tmty{y}{\co{A}}}}
-  \elabarrow
-  \inferrule*{
-    \inferrule*{
-      \inferrule*{
-      }{\seq{\link{y}{z}}{\tmty{y}{\co{A}},\tmty{z}{A}}}
-      \\
-      \seq{P}{\ty{\Gamma},\tmty{x}{B}}
-    }{\seq{\ppar{\link{y}{z}}{P}}{\ty{\Gamma},\tmty{x}{B},\tmty{y}{\co{A}},\tmty{z}{A}}}
-    \\
-    \cs{o}<\minpr(\ty{\Gamma},\ty{A},\ty{B},\ty{\co{A}})
-  }{\seq{\send{x}{z}{(\ppar{\link{y}{z}}{P})}}{\ty{\Gamma},\tmty{x}{\tytens[\cs{o}]{A}{B}},\tmty{y}{\co{A}}}}
-\end{mathpar}
+\subsection{Operational Semantics}
+\input{fig/pcp-operational-semantics}
 

### 2020-09-17 priority-gv.tex
Restructuring.
Added section headers back in.
More figures!

@@
 
-\subsection{Static Syntax}%
-\label{sec:pgv-syntax-static}
+\subsection{Syntax}
 
-\subsubsection{Priorities, types, and environments}
+\paragraph*{Priorities, types, and environments}
+\[
+  \cs{o}, \cs{p}, \cs{q}, \cs{r} \in \mathbb{N}\cup\{\cs{\pbot},\cs{\ptop}\}
+\]
 \[
 \begin{array}{lcl}
-  \cs{o}, \cs{p}, \cs{q}, \cs{r}
-  & \in & \mathbb{N}\cup\{\cs{\pbot},\cs{\ptop}\}
-  \\
   \ty{S}
@@
 \begin{array}{lcl}
-  \ty{\co{\tysend[\cs{o}]{T}{S}}} & = & \ty{\tyrecv[\cs{o}]{T}{\co{S}}}
-  \\
+  \ty{\co{\tysend[\cs{o}]{T}{S}}} & = & \ty{\tyrecv[\cs{o}]{T}{\co{S}}} \\
   \ty{\co{\tyrecv[\cs{o}]{T}{S}}} & = & \ty{\tysend[\cs{o}]{T}{\co{S}}}
@@
 \begin{array}{lcl}
-  \ty{\co{\tyends[\cs{o}]}} & = & \ty{\co{\tyendr[\cs{o}]}}
-  \\
+  \ty{\co{\tyends[\cs{o}]}} & = & \ty{\co{\tyendr[\cs{o}]}} \\
   \ty{\co{\tyendr[\cs{o}]}} & = & \ty{\co{\tyends[\cs{o}]}}
@@
   \pr(\ty{\tyrecv[\cs{o}]{T}{S}}) & = & \cs{o} \\
-  \\
   \pr(\ty{\typrod{T}{U}}) & = & \pr({\ty{T}})\sqcap\pr({\ty{U}}) \\
@@
   \pr(\ty{\tylolli[\cs{p},\cs{q}]{T}{U}}) & = & \cs{p} \\
-  \\
   \pr(\ty{\Gamma}, \tmty{x}{A}) & = & \pr(\ty{\Gamma})\sqcap\pr(\ty{A})
@@
   \pr(\ty{\tyendr[\cs{o}]}) & = & \cs{o} \\
-  \\
   \pr(\ty{\tyunit})         & = & \ptop \\
@@
   \pr(\ty{S})               & = & \pr(\ty{S}) \\
-  \\
   \pr(\ty{\emptyenv})       & = & \ptop
@@
 
-\subsubsection{Terms}%
-\label{sec:pgv-syntax-terms}
+\paragraph*{Syntactic Sugar for Types}
+\[
+  \ty{\tylolli{T}{U}} \elabarrow \ty{\tylolli[\cs{\ptop},\cs{\pbot}]{T}{U}}
+\]
+
+\paragraph*{Static Term Syntax}
 \[
@@
 
-\subsubsection{Syntactic Sugar}%
-\label{sec:pgv-syntax-sugar}
-
+\paragraph*{Syntactic Sugar for Terms}
 \[
 \begin{array}{lcl}
-  \ty{\tylolli{T}{U}}
-  & \elabarrow & \ty{\tylolli[\cs{\ptop},\cs{\pbot}]{T}{U}} \\
-  \ty{\tyselect[\cs{o}]{S}{S'}}
-  & \elabarrow & \ty{\tysend[\cs{o}]{(\tysum{\co{S}}{\co{S'}})}{\tyends[\cs{o+1}]}} \\
-  \ty{\tyoffer[\cs{o}]{S}{S'}}
-  & \elabarrow & \ty{\tyrecv[\cs{o}]{(\tysum{S}{S'})}{\tyendr[\cs{o+1}]}} \\
-  \ty{\tyselectemp[\cs{o}]}
-  & \elabarrow & \ty{\tysend[\cs{o}]{\tyvoid}{\tyends[\cs{o+1}]}} \\
-  \ty{\tyofferemp[\cs{o}]}
-  & \elabarrow & \ty{\tyrecv[\cs{o}]{\tyvoid}{\tyendr[\cs{o+1}]}} \\
-  \\
   \tm{\andthen{M}{N}}
@@
   \tm{\fork}
-  & \elabarrow & \tm{\lambda x.\letpair{y}{z}{\new}{\andthen{\spawn\;{(\lambda\unit.x\;y)}}{z}}} \\
+  & \elabarrow & \tm{\lambda x.\letpair{y}{z}{\new}{\andthen{\spawn\;{(\lambda\unit.x\;y)}}{z}}}
+\end{array}
+\]
+
+\paragraph*{Syntactic Sugar for Internal and External Choice}
+\[
+\begin{array}{lcl}
+  \ty{\tyselect[\cs{o}]{S}{S'}}
+  & \elabarrow & \ty{\tysend[\cs{o}]{(\tysum{\co{S}}{\co{S'}})}{\tyends[\cs{o+1}]}} \\
+  \ty{\tyoffer[\cs{o}]{S}{S'}}
+  & \elabarrow & \ty{\tyrecv[\cs{o}]{(\tysum{S}{S'})}{\tyendr[\cs{o+1}]}} \\
+\end{array}
+\qquad
+\begin{array}{lcl}
+  \ty{\tyselectemp[\cs{o}]}
+  & \elabarrow & \ty{\tysend[\cs{o}]{\tyvoid}{\tyends[\cs{o+1}]}} \\
+  \ty{\tyofferemp[\cs{o}]}
+  & \elabarrow & \ty{\tyrecv[\cs{o}]{\tyvoid}{\tyendr[\cs{o+1}]}} \\
+\end{array}
+\]
+\[
+\begin{array}{lcl}
   \tm{\select{\ell}}
@@
 
-\subsection{Static Typing}%
-\label{sec:pgv-typing-static}
-
-\subsubsection{Typing Rules}
-\begin{mathpar}
-  \inferrule*[lab=T-Var]{
-  }{\tseq[\cs{\pbot}]{\tmty{x}{T}}{x}{T}}
-
-  \inferrule*[lab=T-Lam]{
-    \tseq[\cs{o}]{\ty{\Gamma},\tmty{x}{T}}{M}{U}
-  }{\tseq[\cs{\pbot}]{\ty{\Gamma}}{\lambda x.M}{\tylolli[\cs{\pr(\ty{\Gamma})},\cs{o}]{T}{U}}}
-
-  \inferrule*[lab=T-Const]{
-  }{\tseq[\cs{\pbot}]{\emptyenv}{K}{T}}
-
-  \inferrule*[lab=T-App]{
-    \tseq[\cs{p}]{\ty{\Gamma}}{M}{\tylolli[\cs{o},\cs{r}]{T}{U}}
-    \\
-    \tseq[\cs{q}]{\ty{\Delta}}{N}{T}
-    \\
-    \cs{p}<\pr(\ty{\Delta})
-    \\
-    \cs{q}<\cs{o}
-  }{\tseq[\cs{p}\sqcup\cs{q}\sqcup\cs{o}]{\ty{\Gamma},\ty{\Delta}}{M\;N}{U}}
-  \\
-  \inferrule*[lab=T-Unit]{
-  }{\tseq[\cs{\pbot}]{\emptyenv}{\unit}{\tyunit}}
-
-  \inferrule*[lab=T-LetUnit]{
-    \tseq[\cs{p}]{\ty{\Gamma}}{M}{\tyunit}
-    \\
-    \tseq[\cs{q}]{\ty{\Delta}}{N}{T}
-    \\
-    \cs{p}<\pr(\ty{\Delta})
-  }{\tseq[\cs{p}\sqcup\cs{q}]{\ty{\Gamma},\ty{\Delta}}{\letunit{M}{N}}{T}}
-
-  \inferrule*[lab=T-Pair]{
-    \tseq[\cs{p}]{\ty{\Gamma}}{M}{T}
-    \\
-    \tseq[\cs{q}]{\ty{\Delta}}{N}{U}
-    \\
-    \cs{p}<\pr(\ty{\Delta})
-  }{\tseq[\cs{p}\sqcup\cs{q}]{\ty{\Gamma},\ty{\Delta}}{\pair{M}{N}}{\typrod{T}{U}}}
-
-  \inferrule*[lab=T-LetPair]{
-    \tseq[\cs{p}]{\ty{\Gamma}}{M}{\typrod{T}{T'}}
-    \\
-    \tseq[\cs{q}]{\ty{\Delta},\tmty{x}{T},\tmty{y}{T'}}{N}{U}
-    \\
-    \cs{p}<\pr(\ty{\Delta},\ty{T},\ty{T'})
-  }{\tseq[\cs{p}\sqcup\cs{q}]{\ty{\Gamma},\ty{\Delta}}{\letpair{x}{y}{M}{N}}{U}}
-
-  \inferrule*[lab=T-Inl]{
-    \tseq[o]{\ty{\Gamma}}{M}{T}
-    \\
-    \pr(\ty{T})=\pr(\ty{U})
-  }{\tseq[o]{\ty{\Gamma}}{\inl{M}}{\tysum{T}{U}}}
-
-  \inferrule*[lab=T-Inr]{
-    \tseq[o]{\ty{\Gamma}}{M}{U}
-    \\
-    \pr(\ty{T})=\pr(\ty{U})
-  }{\tseq[o]{\ty{\Gamma}}{\inr{M}}{\tysum{T}{U}}}
-
-  \inferrule*[lab=T-CaseSum]{
-    \tseq[\cs{p}]{\ty{\Gamma}}{L}{\tysum{T}{T'}}
-    \\
-    \tseq[\cs{q}]{\ty{\Delta},\tmty{x}{T}}{M}{U}
-    \\
-    \tseq[\cs{q}]{\ty{\Delta},\tmty{y}{T'}}{N}{U}
-    \\
-    \cs{p}<\pr(\ty{\Delta})
-  }{\tseq[\cs{p}\sqcup\cs{q}]{\ty{\Gamma},\ty{\Delta}}{\casesum{L}{x}{M}{y}{N}}{U}}
-
-  \inferrule*[lab=T-Absurd]{
-    \tseq[o]{\ty{\Gamma}}{M}{\tyvoid}
-  }{\tseq[o]{\ty{\Gamma},\ty{\Delta}}{\absurd{M}}{T}}
-\end{mathpar}
-We write $\tmty{K}{T}$ for $\tseq[\cs{\pbot}]{\emptyenv}{K}{T}$ in typing derivations.
-
-\subsubsection{Type Schemas for Constants}
-\begin{mathpar}
-  \tmty{\link}{\tylolli{\typrod{S}{\co{S}}}{\tyunit}}
-
-  \tmty{\new}{\tylolli{\tyunit}{\typrod{S}{\co{S}}}}
-
-  \tmty{\spawn}{\tylolli{(\tylolli[\cs{p},\cs{q}]{\tyunit}{\tyunit})}{\tyunit}}
-  \\
-  \tmty{\send}{\tylolli[\cs{\ptop},\cs{o}]{\typrod{T}{\tysend[\cs{o}]{T}{S}}}{S}}
-
-  \tmty{\recv}{\tylolli[\cs{\ptop},\cs{o}]{\tyrecv[\cs{o}]{T}{S}}{\typrod{T}{S}}}
-  \\
-  \tmty{\close}{\tylolli[\cs{\ptop},\cs{o}]{\tyends[\cs{o}]}{\tyunit}}
-
-  \tmty{\wait}{\tylolli[\cs{\ptop},\cs{o}]{\tyendr[\cs{o}]}{\tyunit}}
-\end{mathpar}
-
-\subsubsection{Typing Rules for Syntactic Sugar}
-\begin{mathpar}
-  \inferrule*[lab=T-Seq]{
-    \tseq[\cs{p}]{\ty{\Gamma}}{M}{\tyunit}
-    \\
-    \tseq[\cs{q}]{\ty{\Delta}}{N}{T}
-    \\
-    \cs{p}<\pr(\ty{\Delta})
-  }{\tseq[\cs{p}\sqcup\cs{q}]{\ty{\Gamma},\ty{\Delta}}{\andthen{M}{N}}{T}}
-
-  \inferrule*[lab=T-Let]{
-    \tseq[\cs{p}]{\ty{\Gamma}}{M}{T}
-    \\
-    \tseq[\cs{q}]{\ty{\Delta},\tmty{x}{T}}{N}{U}
-    \\
-    \cs{p}<\pr(\ty{\Delta})
-  }{\tseq[\cs{p}\sqcup\cs{q}]{\ty{\Gamma},\ty{\Delta}}{\letbind{x}{M}{N}}{U}}
-  \\
-  \inferrule*[lab=T-LamUnit]{
-    {\tseq[\cs{o}]{\ty{\Gamma}}{M}{T}}
-  }{\tseq[\cs{\pbot}]
-    {\ty{\Gamma}}
-    {\lambda\unit.M}
-    {\tylolli[\cs{\pr(\ty{\Gamma})},\cs{o}]{\tyunit}{T}}}
-
-  \inferrule*[lab=T-LamPair]
-  {\tseq[\cs{o}]
-    {\ty{\Gamma},\tmty{x}{T},\tmty{y}{T'}}
-    {M}
-    {U}}
-  {\tseq[\cs{\pbot}]
-    {\ty{\Gamma}}
-    {\lambda\pair{x}{y}.M}
-    {\tylolli[\cs{\pr(\ty{\Gamma})},\cs{o}]{\typrod{T}{T'}}{U}}}
-  \\
-  \inferrule*[lab=T-Fork]{
-  }{\tseq[\cs{\pbot}]
-    {\emptyenv}
-    {\fork}
-    {\tylolli[]{(\tylolli[\cs{p},\cs{q}]{S}{\tyunit})}{\co{S}}}}
-  \\
-  \inferrule*[lab=T-Select-Inl]{
-    \pr(\ty{S})=\pr(\ty{S'})
-  }{\tseq[\cs{\pbot}]{\emptyenv}{\select{\labinl}}{\tylolli[\cs{\ptop},\cs{o}]{\tyselect[\cs{o}]{S}{S'}}{S}}}
-
-  \inferrule*[lab=T-Select-Inr]{
-    \pr(\ty{S})=\pr(\ty{S'})
-  }{\tseq[\cs{\pbot}]{\emptyenv}{\select{\labinr}}{\tylolli[\cs{\ptop},\cs{o}]{\tyselect[\cs{o}]{S}{S'}}{S'}}}
-
-  \inferrule*[lab=T-Offer]{
-    {\tseq[\cs{p}]
-      {\ty{\Gamma}}
-      {L}
-      {\tyoffer[\cs{o}]{S}{S'}}}
-    \\
-    {\tseq[\cs{q}]
-      {\ty{\Delta},\tmty{x}{S}}
-      {M}
-      {T}}
-    \\
-    {\tseq[\cs{q}]
-      {\ty{\Delta},\tmty{y}{S'}}
-      {N}
-      {T}}
-    \\
-    \cs{o}\sqcup\cs{p}<\pr(\ty{\Delta},\ty{S},\ty{S'})
-  }{\tseq[\cs{o}\sqcup\cs{p}\sqcup\cs{q}]
-    {\ty{\Gamma},\ty{\Delta}}
-    {\offer{L}{x}{M}{y}{N}}
-    {T}}
-
-  \inferrule*[lab=T-Offer-Absurd]{
-    \tseq[\cs{p}]
-    {\ty{\Gamma}}
-    {L}
-    {\tyofferemp[\cs{o}]}
-    \\
-    \cs{o}\sqcup\cs{p}<\pr(\ty{\Delta})
-  }{\tseq[\cs{o}\sqcup\cs{p}]
-    {\ty{\Gamma},\ty{\Delta}}
-    {\offeremp{L}}
-    {T}}
-\end{mathpar}
-
-\subsection{Runtime Syntax}%
-\label{sec:pgv-syntax-runtime}
-
-\subsubsection{Terms}
+\paragraph*{Runtime Term Syntax}
 \[
@@
 
-\subsection{Runtime Typing}%
-\label{sec:pgv-typing-runtime}
-\begin{mathpar}
-  \inferrule*[lab=T-Main]{
-    \tseq[\cs{p}]{\ty{\Gamma}}{M}{T}
-  }{\cseq[\main]{\ty{\Gamma}}{\main\;M}}
+\subsection{Typing}
+\input{fig/pgv-typing}
+See \cref{fig:pgv-typing}.
 
-  \inferrule*[lab=T-Child]{
-    \tseq[\cs{p}]{\ty{\Gamma}}{M}{\tyunit}
-  }{\cseq[\child]{\ty{\Gamma}}{\child\;M}}
-
-  \inferrule*[lab=T-Res]{
-    \cseq[\phi]{\ty{\Gamma},\tmty{x}{S},\tmty{y}{\co{S}}}{\conf{C}}
-  }{\cseq[\phi]{\ty{\Gamma}}{\res{x}{y}{\conf{C}}}}
-
-  \inferrule*[lab=T-Par]{
-    \cseq[\phi]{\ty{\Gamma}}{\conf{C}}
-    \\
-    \cseq[\phi']{\ty{\Delta}}{\conf{D}}
-  }{\cseq[\phi+\phi']{\ty{\Gamma},\ty{\Delta}}{\ppar{\conf{C}}{\conf{D}}}}
-\end{mathpar}
-
-\subsection{Operational semantics}%
-\label{sec:pgv-semantics}
-
-\subsubsection{Term reduction}
-\begin{mathpar}
-  \begin{array}{llcl}
-    \LabTirName{E-Lam}  & \tm{(\lambda x.M) \; V}
-                        & \tred & \tm{\subst{M}{V}{x}}
-    \\
-    \LabTirName{E-Unit} & \tm{\letunit{\unit}{M}}
-                        & \tred & \tm{M}
-    \\
-    \LabTirName{E-Pair} & \tm{\letpair{x}{y}{\pair{V}{W}}{M}}
-                        & \tred & \tm{\subst{\subst{M}{V}{x}}{W}{y}}
-    \\
-    \LabTirName{E-Inl}  & \tm{\casesum{\inl{V}}{x}{M}{y}{N}}
-                        & \tred & \tm{\subst{M}{V}{x}}
-    \\
-    \LabTirName{E-Inr}  & \tm{\casesum{\inr{V}}{x}{M}{y}{N}}
-                        & \tred & \tm{\subst{N}{V}{y}}
-  \end{array}
-  \\
-  \inferrule*[lab=E-Lift]{
-    \tm{M}\tred\tm{M'}
-  }{\tm{\plug{E}{M}}\tred\tm{\plug{E}{M'}}}
-\end{mathpar}
-
-\subsubsection{Structural congruence}
-\begin{mathpar}
-  \begin{array}{llcl}
-    \LabTirName{SC-LinkSwap}   & \tm{\plug{\conf{F}}{\link\;{\pair{x}{y}}}}
-                               & \equiv & \tm{\plug{\conf{F}}{\link\;{\pair{y}{x}}}}
-    \\
-    \LabTirName{SC-ResLink}    & \tm{\res{x}{y}{(\phi\;\link\;\pair{x}{y})}}
-                               & \equiv & \tm{\phi\;\unit}
-    \\
-    \LabTirName{SC-ResSwap}    & \tm{\res{x}{y}{\conf{C}}}
-                               & \equiv & \tm{\res{y}{x}{\conf{C}}}
-    \\
-    \LabTirName{SC-ResComm}    & \tm{\res{x}{y}{\res{z}{w}{\conf{C}}}}
-                               & \equiv & \tm{\res{z}{w}{\res{x}{y}{\conf{C}}}}
-    \\
-    \LabTirName{SC-ResExt}     & \tm{\res{x}{y}{(\ppar{\conf{C}}{\conf{D}}})}
-                               & \equiv & \tm{\ppar{\conf{C}}{\res{x}{y}{\conf{D}}}},
-                                          \text{ if }{\tm{x},\tm{y}\notin\fv(\tm{\conf{C}})}
-    \\
-    \LabTirName{SC-ParNil}     & \tm{\ppar{\conf{C}}{\child{\unit}}}
-                               & \equiv & \tm{\conf{C}}
-    \\
-    \LabTirName{SC-ParComm}    & \tm{\ppar{\conf{C}}{\conf{D}}}
-                               & \equiv & \tm{\ppar{\conf{D}}{\conf{C}}}
-    \\
-    \LabTirName{SC-ParAssoc}   & \tm{\ppar{\conf{C}}{(\ppar{\conf{D}}{\conf{E}})}}
-                               & \equiv & \tm{\ppar{(\ppar{\conf{C}}{\conf{D}})}{\conf{E}}}
-  \end{array}
-\end{mathpar}
-
-\subsubsection{Configuration reduction}
-\begin{mathpar}
-  \begin{array}{llcl}
-    \LabTirName{E-Link}  & \tm{\res{x}{y}{(\ppar{\plug{\conf{F}}{\link\;\pair{w}{x}}}{\conf{C}})}}
-                         & \cred & \tm{\ppar{\plug{\conf{F}}{\unit}}{\subst{\conf{C}}{w}{y}}}
-    \\
-    \LabTirName{E-New}   & \tm{\plug{\conf{F}}{\new\;\unit}}
-                         & \cred & \tm{\res{x}{y}{(\plug{\conf{F}}{\pair{x}{y}})}}
-    \\
-    \LabTirName{E-Spawn} & \tm{\plug{\conf{F}}{\spawn\;V}}
-                         & \cred & \tm{\ppar{\plug{\conf{F}}{\unit}}{\child\;V\;\unit}}
-    \\
-    \LabTirName{E-Send}  & \tm{\res{x}{y}{(\ppar
-                           {\plug{\conf{F}}{\send\;{\pair{V}{x}}}}
-                           {\plug{\conf{F'}}{\recv\;{y}}})}}
-                         & \cred & \tm{\res{x}{y}{(\ppar
-                                   {\plug{\conf{F}}{x}}
-                                   {\plug{\conf{F'}}{\pair{V}{y}}})}}
-    \\
-    \LabTirName{E-Close} & \tm{\res{x}{y}{(\ppar
-                           {\plug{\conf{F}}{\wait\;{x}}}
-                           {\plug{\conf{F'}}{\close\;{y}}})}}
-                         & \cred & \tm{\ppar{\plug{\conf{F}}{\unit}}{\plug{\conf{F'}}{\unit}}}
-  \end{array}
-  \\
-  \inferrule*[lab=E-LiftC]{
-    \tm{\conf{C}}\cred\tm{\conf{C'}}
-  }{\tm{\plug{\conf{G}}{\conf{C}}}\cred\tm{\plug{\conf{G}}{\conf{C'}}}}
-
-  \inferrule*[lab=E-LiftM]{
-    \tm{M}\tred\tm{M'}
-  }{\tm{\plug{\conf{F}}{M}}\tred\tm{\plug{\conf{F}}{M'}}}
-
-  \inferrule*[lab=E-LiftSC]{
-    \tm{\conf{C}}\equiv\tm{\conf{C'}}
-    \\
-    \tm{\conf{C'}}\cred\tm{\conf{D'}}
-    \\
-    \tm{\conf{D'}}\equiv\tm{\conf{D}}
-  }{\tm{\conf{C}}\cred\tm{\conf{D}}}
-\end{mathpar}
+\subsection{Operational Semantics}
+\input{fig/pgv-operational-semantics}
+See \cref{fig:pgv-operational-semantics}.
 
@@
 
+\begin{corollary}[Closed Progress, $\cred$]
+  \label{thm:pgv-closed-progress-confs}
+  If $\cseq[\phi]{\ty{\Gamma}}{\conf{C}}$, then either:
+  \begin{description}[labelwidth=8ex]
+  \item[Done]
+    There exists some $\tm{V}$ such that $\tm{\conf{C}}=\tm{\phi\;V}$;
+  \item[Step]
+    There exists some $\tm{\conf{D}}$ such that $\tm{\conf{C}} \cred \tm{\conf{D}}$; or
+  \end{description}
+\end{corollary}
+
 \end{document}

### 2020-09-17 priority-gv.tex
Minor changes.

@@
 \subsection{Typing}
+See \cref{fig:pgv-typing}.
 \input{fig/pgv-typing}
-See \cref{fig:pgv-typing}.
 
 \subsection{Operational Semantics}
+See \cref{fig:pgv-operational-semantics}.
 \input{fig/pgv-operational-semantics}
-See \cref{fig:pgv-operational-semantics}.
 
@@
   \label{thm:pgv-closed-progress-confs}
-  If $\cseq[\phi]{\ty{\Gamma}}{\conf{C}}$, then either:
+  If $\cseq[\phi]{\emptyenv}{\conf{C}}$, then either:
   \begin{description}[labelwidth=8ex]
   \item[Done]
-    There exists some $\tm{V}$ such that $\tm{\conf{C}}=\tm{\phi\;V}$;
+    There exists some $\tm{V}$ such that $\tm{\conf{C}}=\tm{\phi\;V}$; or
   \item[Step]
-    There exists some $\tm{\conf{D}}$ such that $\tm{\conf{C}} \cred \tm{\conf{D}}$; or
+    There exists some $\tm{\conf{D}}$ such that $\tm{\conf{C}} \cred \tm{\conf{D}}$.
   \end{description}

### 2020-09-17 prf/thm-pcp-to-pgv-operational-correspondence-soundness.tex
Finished OC.

\begin{proof}
  \label{prf:thm-pcp-to-pgv-operational-correspondence-soundness}
  By induction on the derivation of $\pgv{\tm{\cpgvC{P}}\cred\tm{\conf{C}}}$.
  We omit the cases which cannot occur, \ie \LabTirName{E-New}, \LabTirName{E-Spawn}, and \LabTirName{LiftM}.

  \begin{case*}[\LabTirName{E-Link}]
    \[\pgv{%
        \tm{\res{x}{x'}{(\ppar{\plug{\conf{F}}{\link\;\pair{w}{x}}}{\conf{C}})}}
        \cred
        \tm{\ppar{\plug{\conf{F}}{\unit}}{\subst{\conf{C}}{w}{x'}}}
      }\]
    The source for $\pgv{\tm{\link\;\pair{w}{x}}}$ \emph{must} be $\pcp{\tm{\link{w}{x}}}$. None of the translation rules introduce an evaluation context around the recursive call, hence $\pgv{\tm{\conf{F}}}$ must be the empty context. Let $\pcp{\tm{P}}$ be the source term for $\pgv{\tm{\conf{C}}}$, \ie $\pgv{\tm{\cpgvC{P}}=\tm{\conf{C}}}$. Hence, we have:
    \begin{mathpar}
      \begin{tikzcd}[cramped, column sep=tiny]
        \pcp{\tm{\res{x}{x'}{(\ppar{\link{w}{x}}{P})}}}
        \arrow[r, "\pcp{\red}"]
        \arrow[d, "\cpgvC{\cdot}"]
        &
        \pcp{\tm{\subst{P}{w}{x'}}}
        \arrow[dd, "\cpgvC{\cdot}"]
        \\
        \pgv{\tm{\res{x}{x'}{(\ppar{\child\;\link\;\pair{w}{x}}{\cpgvC{P}})}}}
        \arrow[d, "\pgv{\cred^+}"]
        \\
        \pgv{\tm{\subst{\cpgvC{P}}{w}{x'}}}
        \arrow[r, "\pcp{=}"]
        &
        \pgv{\tm{\cpgvC{\subst{P}{w}{x'}}}}
      \end{tikzcd}
    \end{mathpar}
  \end{case*}
  \begin{case*}[\LabTirName{E-Send}]
    \[\pgv{%
        \tm{\res{x}{x'}{(\ppar{\plug{\conf{F}}{\send\;{\pair{V}{x}}}}{\plug{\conf{F'}}{\recv\;{x'}}})}}
        \cred
        \tm{\res{x}{x'}{(\ppar{\plug{\conf{F}}{x}}{\plug{\conf{F'}}{\pair{V}{x'}}})}}
      }\]
    There are three possible sources for $\pgv{\tm{\send}}$ and $\pgv{\tm{\recv}}$: $\pcp{\tm{\send{x}{y}{P}}}$ and $\pcp{\tm{\recv{x'}{y'}{Q}}}$; $\pcp{\tm{\inl{x}{P}}}$ and $\pcp{\tm{\offer{x'}{Q}{R}}}$; or $\pcp{\tm{\inr{x}{P}}}$ and $\pcp{\tm{\offer{x'}{Q}{R}}}$.
    \begin{subcase*}[$\pcp{\tm{\send{x}{y}{P}}}$ and $\pcp{\tm{\recv{x'}{y'}{Q}}}$]
      None of the translation rules introduce an evaluation context around the recursive call, hence $\pgv{\tm{\conf{F}}}$ must be $\pgv{\tm{\child\;\letbind{x}{\hole}{\cpgvM{P}}}}$. Similarly, $\pgv{\tm{\conf{F'}}}$ must be $\pgv{\tm{\child\;\letpair{y'}{x'}{\hole}{\cpgvM{Q}}}}$. The value $\pgv{\tm{V}}$ must be an endpoint $\tm{y}$, bound by the name restriction $\pgv{\tm{\res{y}{y'}{}}}$ introduced by the translation. Hence, we have:
      \begin{mathpar}
          \begin{tikzcd}[cramped, column sep=tiny]
            \pcp{\tm{\res{x}{x'}{(\ppar{\send{x}{y}{P}}{\recv{x'}{y'}{Q}})}}}
            \arrow[d, "\cpgvC{\cdot}"]
            \arrow[r, "\pcp{\red}"]
            &
            \pcp{\tm{\res{x}{x'}{\res{y}{y'}{(\ppar{P}{Q})}}}}
            \arrow[dd, "\cpgvC{\cdot}"]
            \\
            \pgv{\tm{\res{x}{x'}{\res{y}{y'}{}}\left(
                  \begin{array}{l}
                    \child\;\letbind{x}{\send\;{\pair{y}{x}}}{\cpgvM{P}}
                    \parallel
                    \\
                    \child\;\letpair{y'}{x'}{\recv\;{x'}}{\cpgvM{Q}}
                  \end{array}
                \right)}}
            \arrow[d, "\pgv{\equiv\cred^+}"]
            \\
            \pgv{\tm{\res{x}{x'}{\res{y}{y'}{(\ppar{\child\;\cpgvM{P}}{\child\;\cpgvM{Q}})}}}}
            \arrow[r, "\pgv{\cred^\star}", "\text{(by \cref{lem:pcp-to-pgv-cpgvM-to-cpgvC})}"']
            &
            \pgv{\tm{\res{x}{x'}{\res{y}{y'}{(\ppar{\cpgvC{P}}{\cpgvC{Q}})}}}}
          \end{tikzcd}
      \end{mathpar}
    \end{subcase*}
    \begin{subcase*}[$\pcp{\tm{\inl{x}{P}}}$ and $\pcp{\tm{\offer{x'}{Q}{R}}}$]
      None of the translation rules introduce an evaluation context around the recursive call, hence $\pgv{\tm{\conf{F}}}$ must be $$\pgv{\tm{\child\;\letbind{x}{\andthen{\close\;\hole}{y}}{\cpgvM{P}}}}.$$ Similarly, $\pgv{\tm{\conf{F'}}}$ must be $$\pgv{\tm{\child\;\letpair{y'}{x'}{\hole}{\andthen{\wait\;x'}{\casesum{y'}{y'}{\cpgvM{Q}}{y'}{\cpgvM{R}}}}}}.$$ Hence, we have:
      \begin{mathpar}
        \begin{tikzcd}[cramped, column sep=tiny]
          \pcp{\tm{\res{x}{x'}{(\ppar{\inl{x}{P}}{\offer{x}{Q}{R}})}}}
          \arrow[d, "\cpgvM{\cdot}"]
          \arrow[r, "\pcp{\red}"]
          &
          \pcp{\tm{\res{x}{x'}{(\ppar{P}{Q})}}}
          \arrow[dd, "\cpgvC{\cdot}"]
          \\
          \pgv{\tm{\res{x}{x'}{}\left(
                \begin{array}{l}
                  \child\;\letbind{x}{\select{\labinl}\;{x}}{\cpgvM{P}}
                  \parallel
                  \\
                  \child\;\offer{x'}{x'}{\cpgvM{Q}}{x'}{\cpgvM{R}}
                \end{array}
              \right)}}
          \arrow[d, "\pgv{\cred^+}"]
          \\ 
          \pgv{\tm{\res{x}{x'}{(\ppar{\child\;\cpgvM{P}}{\child\;\cpgvM{Q}})}}}
          \arrow[r, "\pgv{\cred^\star}", "\text{(by \cref{lem:pcp-to-pgv-cpgvM-to-cpgvC})}"']
          & 
          \pgv{\tm{\res{x}{x'}{(\ppar{\cpgvC{P}}{\cpgvC{Q}})}}}
        \end{tikzcd}
      \end{mathpar}
    \end{subcase*}
    \begin{subcase*}[$\pcp{\tm{\inr{x}{P}}}$ and $\pcp{\tm{\offer{x'}{Q}{R}}}$]
      None of the translation rules introduce an evaluation context around the recursive call, hence $\pgv{\tm{\conf{F}}}$ must be $$\pgv{\tm{\child\;\letbind{x}{\andthen{\close\;\hole}{y}}{\cpgvM{P}}}}.$$ Similarly, $\pgv{\tm{\conf{F'}}}$ must be $$\pgv{\tm{\child\;\letpair{y'}{x'}{\hole}{\andthen{\wait\;x'}{\casesum{y'}{y'}{\cpgvM{Q}}{y'}{\cpgvM{R}}}}}}.$$ Hence, we have:
      \begin{mathpar}
        \begin{tikzcd}[cramped, column sep=tiny]
          \pcp{\tm{\res{x}{x'}{(\ppar{\inr{x}{P}}{\offer{x}{Q}{R}})}}}
          \arrow[d, "\cpgvM{\cdot}"]
          \arrow[r, "\pcp{\red}"]
          &
          \pcp{\tm{\res{x}{x'}{(\ppar{P}{Q})}}}
          \arrow[dd, "\cpgvC{\cdot}"]
          \\
          \pgv{\tm{\res{x}{x'}{}\left(
                \begin{array}{l}
                  \child\;\letbind{x}{\select{\labinr}\;{x}}{\cpgvM{P}}
                  \parallel
                  \\
                  \child\;\offer{x'}{x'}{\cpgvM{Q}}{x'}{\cpgvM{R}}
                \end{array}
              \right)}}
          \arrow[d, "\pgv{\cred^+}"]
          \\ 
          \pgv{\tm{\res{x}{x'}{(\ppar{\child\;\cpgvM{P}}{\child\;\cpgvM{R}})}}}
          \arrow[r, "\pgv{\cred^\star}", "\text{(by \cref{lem:pcp-to-pgv-cpgvM-to-cpgvC})}"']
          & 
          \pgv{\tm{\res{x}{x'}{(\ppar{\cpgvC{P}}{\cpgvC{R}})}}}
        \end{tikzcd}
      \end{mathpar}
    \end{subcase*}
  \end{case*}
  \begin{case*}[\LabTirName{E-Close}]
    \[\pgv{%
        \tm{\res{x}{x'}{(\ppar{\plug{\conf{F}}{\wait\;{x}}}{\plug{\conf{F'}}{\close\;{x'}}})}}
        \cred
        \tm{\ppar{\plug{\conf{F}}{\unit}}{\plug{\conf{F'}}{\unit}}}
      }\]
    The source for $\pgv{\tm{\wait}}$ and $\pgv{\tm{\close}}$ \emph{must} be $\pcp{\tm{\wait{x}{P}}}$ and $\pcp{\tm{\close{x'}{Q}}}$.

    (The translation for $\pcp{\tm{\offer{x}{P}{Q}}}$ also introduces a $\pgv{\tm{\wait}}$, but it is blocked on another communication, and hence cannot be the first communication on a translated term. The translations for $\pcp{\tm{\inl{x}{P}}}$ and $\pcp{\tm{\inr{x}{P}}}$ also introduce a $\pgv{\tm{\close}}$, but these are similarly blocked.)

    None of the translation rules introduce an evaluation context around the recursive call, hence $\pgv{\tm{\conf{F}}}$ must be $\pgv{\tm{\andthen{\hole}{\cpgvM{P}}}}$. Similarly, $\pgv{\tm{\conf{F'}}}$ must be $\pgv{\tm{\andthen{\hole}{\cpgvM{Q}}}}$. Hence, we have:
    \begin{mathpar}
      \begin{tikzcd}
        \pcp{\tm{\res{x}{x'}{(\ppar{\close{x}{P}}{\wait{x'}{Q}})}}} 
        \arrow[d, "\cpgvM{\cdot}"]
        \arrow[r, "\pcp{\red}"]
        &
        \pcp{\tm{\ppar{P}{Q}}}
        \arrow[dd, "\cpgvC{\cdot}"]
        \\
        \pgv{\tm{\res{x}{x'}{(\ppar
              {\child\;\andthen{\close\;{x}}{\cpgvM{P}}} 
              {\child\;\andthen{\wait\;{x'}}{\cpgvM{Q}}}
              )}}}
        \arrow[d, "\pgv{\cred^+}"]
        \\ 
        \pgv{\tm{\ppar{\child\;\cpgvM{P}}{\child\;\cpgvM{Q}}}}
        \arrow[r, "\pgv{\cred^\star}", "\text{(by \cref{lem:pcp-to-pgv-cpgvM-to-cpgvC})}"']
        & 
        \pgv{\tm{\ppar{\cpgvC{P}}{\cpgvC{Q}}}}
      \end{tikzcd}
    \end{mathpar}
  \end{case*}
  \begin{case*}[\LabTirName{E-LiftC}]
    By the induction hypothesis and \LabTirName{E-LiftC}.
  \end{case*}
  \begin{case*}[\LabTirName{E-LiftSC}]
    By the induction hypothesis, \LabTirName{E-LiftSC},
    and \cref{lem:pcp-to-pgv-confs-operational-correspondence-equiv}.
  \end{case*}
\end{proof}

%%% Local Variables:
%%% TeX-master: "../priorities"
%%% End:

### 2020-09-24 priority-gv.tex
Minor changes.

@@
 \paragraph*{Static Term Syntax}
+We use $\tm{x}$, $\tm{y}$, $\tm{z}$, and $\tm{w}$ to range over variable names. Occasionally, we use $\tm{a}$, $\tm{b}$, and $\tm{c}$ to range over \emph{free} channel endpoints.
 \[
@@
   \tm{V}, \tm{W}
-  & \coloneqq & \tm{\lambda x.M}
+  & \coloneqq & \tm{x}
+    \sep        \tm{K}
+    \sep        \tm{\lambda x.M}
     \sep        \tm{\unit}
@@
   \tm{R}
-  & \coloneqq & \tm{\link\;\pair{V}{W}}
-    \sep        \tm{\new}
-    \sep        \tm{\spawn\;V} \\
-  & \sep      & \tm{\send\;\pair{V}{W}}
-    \sep        \tm{\recv\;V}
-    \sep        \tm{\close\;V}
-    \sep        \tm{\wait\;V}
+  & \coloneqq & \tm{K\;V}
   \\
@@
     \sep        \tm{\casesum{E}{x}{M}{y}{N}} \\
+  & \sep      & \tm{\absurd{E}} \\
   \\
@@
 \begin{proof}
-  \admit
+  By induction on the derivation of $\tseq[\cs{o}]{\ty{\Gamma}}{M}{T}$.
 \end{proof}
+
+\begin{definition}[Canonical Forms]
+  \label{def:pgv-canonical-forms}
+  A~configuration $\tm{\conf{C}}$ is in canonical form if it is of the form
+  \(
+    \tm{\res{x_1}{x'_1}{\dots\res{x_n}{x'_n}{(\child\;M_1\parallel\dots\parallel\child\;M_m\parallel\main\;N)}}}
+  \)
+  and no $\tm{M_i}$ is a value.
+\end{definition}
+
+\begin{restatablelemma}{lempgvcanonicalforms}[Canonical Forms]
+  \label{lem:pgv-canonical-forms}
+  If $\cseq[\main]{\ty{\Gamma}}{\conf{C}}$, there exists some $\tm{\conf{D}}$ such that $\tm{\conf{C}}\equiv\tm{\conf{D}}$ and $\tm{\conf{D}}$ is in canonical form.
+\end{restatablelemma}
+\begin{proof}
+  We move all $\nu$-binders to the top using \LabTirName{SC-ResExt}, move the main thread to the right using \LabTirName{SC-ParComm} and \LabTirName{SC-ParAssoc}, and if any thread is of the form $\tm{\child\;\unit}$, we discard it using \LabTirName{SC-ParNil}.
+\end{proof}
+
+\begin{definition}[Ready Terms]
+  A~term $\tm{M}$ is ready on a variable $\tm{x}$, written $\ready\pair{\tm{M},\tm{x}}$, if $\tm{M}$ is of the form $\tm{\plug{E}{R}}$, where $\tm{E}$ is an evaluation context and $\tm{R}$ is $\tm{\send\;\pair{V}{x}}$, $\tm{\recv\;x}$, $\tm{\close\;x}$, or $\tm{\wait\;x}$.
+\end{definition}
+
+%%% <VERY UNDER CONSTRUCTION>
+
+\begin{definition}
+  A~term $\tm{M}$ is in a configuration $\tm{\conf{C}}$, written $\tm{M}\in\tm{\conf{C}}$, if there exists a configuration context $\tm{\conf{G}}$ such that $\tm{\conf{C}}=\tm{\plug{\conf{G}}{M}}$.
+\end{definition}
+
+\begin{definition}[Blocked Terms]
+  A~term $\tm{M}\in\tm{\conf{C}}$ is blocked on a variable $\tm{x}$, written $\blocked_{\tm{\conf{C}}}\pair{\tm{M},\tm{x}}$, if $\ready\pair{\tm{M}}{\tm{x}}$, and no other term is ready on the dual of $\tm{x}$, \ie there exist no $\tm{\conf{G}}$, $\tm{\conf{D}}$, and $\tm{N}$ such that $\tm{C}=\tm{\plug{\conf{G}}{\res{x}{x'}{\conf{D}}}}$, $\tm{M},\tm{N}\in\tm{\conf{D}}$, and $\ready\pair{\tm{N}}{\tm{x'}}$.
+We omit the subscript $\tm{\conf{C}}$ if the configuration is clear from context.
+\end{definition}
 
@@
   \label{thm:pgv-open-progress-confs}
-  If $\cseq[\phi]{\ty{\Gamma}}{\conf{C}}$, then either:
+  Suppose $\cseq[\main]{\ty{\Gamma}}{\conf{C}}$, where $\tm{\conf{C}}$ is in canonical form.
+  Let $\tm{\conf{C}}=\tm{\res{x_1}{x'_1}{\dots\res{x_n}{x'_n}{(\child\;M_1\parallel\dots\parallel\child\;M_m\parallel\main\;N)}}}$.
+  Either:
   \begin{description}[labelwidth=8ex]
   \item[Done]
-    There exists some $\tm{V}$ such that $\tm{\conf{C}}=\tm{\phi\;V}$;
+    $\tm{N}$ is a value, and each $\tm{M_i}$ is either blocked, or a link on two free endpoints, \ie either $\blocked\pair{\tm{M_i}}{\tm{z}}$ for some $\tm{z}\in\{\tm{x_1},\tm{x'_1},\dots,\tm{x_n},\tm{x'_n}\}\cup\fv(\ty{\Gamma})$, or $\tm{M}=\tm{\plug{E}{\link\;\pair{a}{b}}}$ for some $\tm{a},\tm{b}\in\fv(\ty{\Gamma})$.
   \item[Step]
-    There exists some $\tm{\conf{D}}$ such that $\tm{\conf{C}} \cred \tm{\conf{D}}$; or
+    There exists some $\tm{\conf{D}}$ such that $\tm{\conf{C}} \cred \tm{\conf{D}}$.
   \item[Blocked]
-    We have $\tm{\conf{C}}\equiv\tm{\res{x_1y_1\dots x_ny_n}{}{(\phi_1R_1 \parallel\dots\parallel \phi_mR_m)}}$,\\
-    and for some $\tm{R_j}$ and $\tm{a_j}\in\fn(\ty{\Gamma})$, $\blocked(\tm{a_j},\tm{R_j})$.
+    $\tm{N}$ and each $\tm{M_i}$ is either blocked, or a link on two free endpoints, \ie either $\blocked\pair{\tm{M_i}}{\tm{z}}$ for some $\tm{z}\in\{\tm{x_1},\tm{x'_1},\dots,\tm{x_n},\tm{x'_n}\}\cup\fv(\ty{\Gamma})$, or $\tm{M}=\tm{\plug{E}{\link\;\pair{a}{b}}}$ for some $\tm{a},\tm{b}\in\fv(\ty{\Gamma})$.
   \end{description}
@@
 \begin{proof}
-  \admit
+  We apply \cref{lem:pgv-canonical-forms} to $\tm{\conf{C}}$ to get an equivalent configuration in canonical form with flag $\phi$, $\tm{\res{x_1}{x'_1}{\dots\res{x_n}{x'_n}{(\child\;M_1\parallel\dots\parallel\child\;M_m\parallel\phi\;N)}}}$.
+  We apply \cref{lem:pgv-open-progress-terms} to each term $\tm{M_i}$ and $\tm{N}$.
 \end{proof}
+
+%%% </VERY UNDER CONSTRUCTION>
 
@@
   \label{thm:pgv-closed-progress-confs}
-  If $\cseq[\phi]{\emptyenv}{\conf{C}}$, then either:
+  Suppose $\cseq[\main]{\emptyenv}{\conf{C}}$, where $\tm{\conf{C}}$ is in canonical form.
+  Let $\tm{\conf{C}}=\tm{\res{x_1}{x'_1}{\dots\res{x_n}{x'_n}{(\child\;M_1\parallel\dots\parallel\child\;M_m\parallel\main\;N)}}}$.
   \begin{description}[labelwidth=8ex]
   \item[Done]
-    There exists some $\tm{V}$ such that $\tm{\conf{C}}=\tm{\phi\;V}$; or
+    $\tm{N}$ is a value, and for each $\tm{M_i}$, $\blocked\pair{\tm{M_i}}{\tm{z}}$ for some $\tm{z}\in\{\tm{x_1},\tm{x'_1},\dots,\tm{x_n},\tm{x'_n}\}$.
   \item[Step]

### 2020-09-26 prf/thm-pcp-closed-progress.tex
Added closed progress for PCP.

\begin{proof}
  \label{prf:thm-pcp-closed-progress}
  If $\tm{P}=\tm{\halt}$, we are done. Otherwise, $\tm{P}$ is of the form
  \[
    \seq{\res{x_1}{x'_1}{\dots\res{x_n}{x'_n}{(P_1 \parallel\dots\parallel P_m)}}}{\emptyenv}
  \]
  where $m>0$ and each $\seq{P_i}{\ty{\Gamma_i}}$ is an action.

  Pick the process $\tm{P_i}$ with the smallest priority $\minpr{(\ty{\Gamma_i})}$. $\tm{P_i}$ acts on some endpoint $\tmty{y}{A}\in\ty{\Gamma_i}$. We must have $\minpr{(\ty{\Gamma_i})}=\pr{(\ty{A})}$, since the other actions in $\tm{P_i}$ are guarded by the action on $\tmty{y}{A}$.

  If $\tm{P_j}$ is a link $\tm{\link{y}{z}}$ or $\tm{\link{z}{y}}$, we apply \LabTirName{E-Link}.

  Otherwise, there must be a process $\seq{P_j}{\ty{\Gamma_j}}$ where $\tmty{y'}{\co{A}}\in\ty{\Gamma_j}$ and $\tm{y}$ and $\tm{y'}$ are dual endpoints of the same channel, since $\tm{y}$ is bound. By duality, $\pr{(\ty{A})}=\pr{(\ty{\co{A}})}$. We must have $\pr{(\ty{\co{A}})}=\minpr{(\ty{\Gamma_j})}$, since we chose $\tm{P_i}$ to have the smallest priority $\minpr{(\Gamma_i)}$. We cannot have $\tm{P_i}=\tm{P_j}$, otherwise the action on $\tm{y'}$ would be guarded by the action on $\tm{y}$, requiring $\pr{(\ty{A})}<\pr{(\ty{\co{A}})}$. The process $\tm{P_j}$ must act on $\tm{y'}$, otherwise the action on $\tm{y'}$ would be guarded by another action with priority smaller than $\pr{(\ty{A})}$. We have two processes, acting on dual endpoints. We apply the appropriate reduction rule, \ie \LabTirName{E-Send}, \LabTirName{E-Close}, \LabTirName{E-Select-Inl}, or \LabTirName{E-Select-Inr}.
\end{proof}

%%% Local Variables:
%%% TeX-master: "../priorities"
%%% End:

### 2020-09-26 priority-cp.tex
Added section headers back in.
Added figure references in.
Minor changes.
Added closed progress for PCP.

@@
   \vspace{-4mm}
-  \caption{Cyclic Scheduler from \cite{DardhaG18}}
+  \caption{Cyclic Scheduler from \cite{dardha2018}}
   \label{fig:scheduler}
@@
 
-In this section we present an updated version of Priority CP~\cite{DardhaG18}, which is Wadler's Classical Processes~\cite{W12} with \emph{priorities}.
-Some details of the calculus are inspired by Carbone \emph{et al.}~\cite{CLMSW16} and Caires and P\'{e}rez~\cite{CP17}.
+\subsection{Syntax}
+
+In this section we present an updated version of Priority CP~\cite{dardha2018}, which is Wadler's Classical Processes~\cite{wadler2012} with \emph{priorities}.
+Some details of the calculus are inspired by Carbone \emph{et al.}~\cite{carbone2016} and Caires and P\'{e}rez~\cite{caires2017}.
 
@@
   & \coloneqq & \tm{\link{x}{y}}
-         \sep   \tm{\res{x}{y}{P}}
+         \sep   \tm{\res{x}{x'}{P}}
          \sep   \tm{(\ppar{P}{Q})}
@@
 \subsection{Typing}
+See \cref{fig:pcp-typing}.
 \input{fig/pcp-typing}
@@
 \subsection{Operational Semantics}
+See \cref{fig:pcp-operational-semantics}.
 \input{fig/pcp-operational-semantics}
+
+\subsection{Metatheory}
+
+\begin{restatablelemma}{lempcpsubjectcongruence}[Subject Congruence, $\equiv$]
+  \label{lem:pcp-subject-congruence}
+  \hfill\\%newline before theorem statement
+  If $\seq{P}{\ty{\Gamma}}$ and $\tm{P}\equiv\tm{Q}$,
+  then $\seq{Q}{\ty{\Gamma}}$.
+\end{restatablelemma}
+\begin{proof}
+  See~\cite{dardha2018}.
+\end{proof}
+
+\begin{restatabletheorem}{thmpcpsubjectreduction}[Subject Reduction, $\red$]
+  \label{thm:pcp-subject-reduction}
+  \hfill\\%newline before theorem statement
+  If $\seq{P}{\ty{\Gamma}}$ and $\tm{P}\red\tm{Q}$, then $\seq{Q}{\ty{\Gamma}}$.
+\end{restatabletheorem}
+\begin{proof}
+  See~\cite{dardha2018}.
+\end{proof}
+
+\begin{definition}[Action]
+  A~process acts on an endpoint $\tm{x}$ if it is in one of the following forms:
+  \begin{multicols}{3}
+    \begin{itemize}[noitemsep,topsep=0pt,parsep=0pt,partopsep=0pt]
+    \item $\tm{\link{x}{y}}$ 
+    \item $\tm{\link{y}{x}}$
+    \item $\tm{\send{x}{y}{P}}$
+    \item $\tm{\recv{x}{y}{P}}$
+    \item $\tm{\close{x}{P}}$
+    \item $\tm{\wait{x}{P}}$
+    \item $\tm{\inl{x}{P}}$
+    \item $\tm{\inr{x}{P}}$
+    \item $\tm{\offer{x}{P}{Q}}$
+    \item $\tm{\absurd{x}}$
+    \end{itemize}
+  \end{multicols}
+  \noindent
+  A~process is an action if it acts on on some endpoint $\tm{x}$. 
+\end{definition}
+
+\begin{definition}[Canonical Forms]
+  \label{def:pcp-canonical-forms}
+  A~process $\tm{P}$ is in canonical form if it is in one of the following forms:
+  \begin{itemize}[noitemsep,topsep=0pt,parsep=0pt,partopsep=0pt]
+  \item
+    $\tm{\halt}$
+  \item
+    $\tm{\res{x_1}{x'_1}{\dots\res{x_n}{x'_n}{(P_1 \parallel\dots\parallel P_m)}}}$
+    where $m>0$ and each $\tm{P_j}$ is an action.
+  \end{itemize}
+\end{definition}
+
+\begin{restatablelemma}{lempcpcanonicalforms}[Canonical Forms]
+  \label{lem:pcp-canonical-forms}
+  If $\seq{P}{\ty{\Gamma}}$, there exists some $\tm{Q}$ such that $\tm{P}\equiv\tm{Q}$ and $\tm{Q}$ is in canonical form.
+\end{restatablelemma}
+\begin{proof}
+  If $\tm{P}=\tm{\halt}$, we are done. Otherwise, we move any $\nu$-binders to the top using \LabTirName{SC-ResExt}, and discard any occurrences of $\tm{\halt}$ using \LabTirName{SC-ParNil}.
+\end{proof}
+
+\begin{restatabletheorem}{thmpcpclosedprogress}[Closed Progress, $\red$]
+  \label{thm:pcp-closed-progress}
+  \hfill\\%newline before theorem statement
+  If $\seq{P}{\emptyenv}$ and $\tm{P}$ is in canonical form, then either $\tm{P}=\tm{\halt}$ or there exists a $\tm{Q}$ such that $\tm{P}\red\tm{Q}$.
+\end{restatabletheorem}
+\begin{proof}
+  Details in~\cref{prf:thm-pcp-closed-progress}.
+\end{proof}
 

### 2020-09-27 priority-gv.tex
Added closed progress for PCP.
Started PGV progress. Created Makefile with watch command.

--- 2020-09-17 priority-gv.tex
@@
 
-\begin{restatablelemma}{lempgvopenprogressterms}[Open Progress, $\tred$]
-  \label{lem:pgv-open-progress-terms}
-  If $\tseq[\cs{o}]{\ty{\Gamma}}{M}{T}$, then either:
-  \begin{description}[labelwidth=8ex]
-  \item[Done]
-    $\tm{M}$ is a value;
-  \item[Step]
-    There exists some $\tm{N}$ such that $\tm{M}\tred\tm{N}$; or
-  \item[Blocked]
-    There exists some $\tm{E}$ and $\tm{R}$ such that $\tm{M}=\tm{\plug{E}{R}}$.
-  \end{description}
-\end{restatablelemma}
-\begin{proof}
-  \admit
-\end{proof}
-
-\begin{restatabletheorem}{thmpgvopenprogressconfs}[Open Progress, $\cred$]
-  \label{thm:pgv-open-progress-confs}
-  If $\cseq[\phi]{\ty{\Gamma}}{\conf{C}}$, then either:
-  \begin{description}[labelwidth=8ex]
-  \item[Done]
-    There exists some $\tm{V}$ such that $\tm{\conf{C}}=\tm{\phi\;V}$;
-  \item[Step]
-    There exists some $\tm{\conf{D}}$ such that $\tm{\conf{C}} \cred \tm{\conf{D}}$; or
-  \item[Blocked]
-    We have $\tm{\conf{C}}\equiv\tm{\res{x_1y_1\dots x_ny_n}{}{(\phi_1R_1 \parallel\dots\parallel \phi_mR_m)}}$,\\
-    and for some $\tm{R_j}$ and $\tm{a_j}\in\fn(\ty{\Gamma})$, $\blocked(\tm{a_j},\tm{R_j})$.
-  \end{description}
-\end{restatabletheorem}
-\begin{proof}
-  \admit
-\end{proof}
-
-\begin{corollary}[Closed Progress, $\cred$]
-  \label{thm:pgv-closed-progress-confs}
-  If $\cseq[\phi]{\emptyenv}{\conf{C}}$, then either:
-  \begin{description}[labelwidth=8ex]
-  \item[Done]
-    There exists some $\tm{V}$ such that $\tm{\conf{C}}=\tm{\phi\;V}$; or
-  \item[Step]
-    There exists some $\tm{\conf{D}}$ such that $\tm{\conf{C}} \cred \tm{\conf{D}}$.
-  \end{description}
-\end{corollary}
-
 \end{document}

### 2020-09-27 priority-cp.tex
Started PGV progress. Created Makefile with watch command.
Rewrote text.

@@
 \paragraph*{Priorities, types, and environments}
-Types are based on CLL and range over $\ty{A}, \ty{B}, \ty{C}$.
-Types are annotated with priorities which range over $\cs{o} \in \mathbb{N} \sqcup \{\omega\}$.
-Let $\omega$ be a special element such that $\cs{o}<\omega$ for all $\cs{o}\in \mathbb{N}$.
-Often, we will omit $\omega$ when not necessary.
-\todo{check the omega thing is also part of our presentation} Types $\ty{\tyone[\cs{o}]}$ and $\ty{\tybot[\cs{o}]}$ are used to type channel endpoints which are ready to be closed.
-Types $\ty{\tynil[\cs{o}]}$ and $\ty{\tytop[\cs{o}]}$ are used to...
-Type $\ty{\tytens[\cs{o}]{A}{B}}$ (respectively, $\ty{\typarr[\cs{o}]{A}{B}}$) is used to type a channel endpoint that first {outputs} (respectively, {inputs}) a channel of type $\ty A$ and then proceeds as $\ty B$.
-$\ty{\typlus[\cs{o}]{A}{B}}$ is used to type a channel endpoint over which we can {select} either the \emph{left} branch and proceed as $\ty A$ or the \emph{right} branch and proceed as $\ty B$.
-$\ty{\tywith[\cs{o}]{A}{B}}$ is used to type a channel endpoint that can offer a both $\ty A$ and $\ty B$.
+Types are based on classical linear logic formulas, and are annotated with priorities $\cs{o}\in\mathbb{N}$.
+
+The types $\ty{\tytens[\cs{o}]{A}{B}}$ and $\ty{\typarr[\cs{o}]{A}{B}}$ type the endpoints of a channel over which we send or receive a channel of type $\ty{A}$, and then proceed as $\ty{B}$.
+The types $\ty{\tyone[\cs{o}]}$ and $\ty{\tybot[\cs{o}]}$ type the endpoints of a channel over which we send or receive a ping, and then close the channel.
+These act as the units for $\ty{\tytens[\cs{o}]{A}{B}}$ and $\ty{\typarr[\cs{o}]{A}{B}}$.
+
+The types $\ty{\typlus[\cs{o}]{A}{B}}$ and $\ty{\tywith[\cs{o}]{A}{B}}$ type the endpoints of a channel over which we can make or offer a choice between two options $\ty{A}$ or $\ty{B}$.
+The types $\ty{\tynil[\cs{o}]}$ and $\ty{\tytop[\cs{o}]}$ type the endpoints of a channel over which we can make offer a choice between \emph{no options}.
+These act as the units for $\ty{\typlus[\cs{o}]{A}{B}}$ and $\ty{\tywith[\cs{o}]{A}{B}}$.
+\[
+\begin{array}{lcl}
+  \ty{A}, \ty{B}, \ty{C}
+  & \coloneqq & \ty{\tytens[\cs{o}]{A}{B}}
+    \sep        \ty{\typarr[\cs{o}]{A}{B}}
+    \sep        \ty{\tyone[\cs{o}]}
+    \sep        \ty{\tybot[\cs{o}]}
+    \sep        \ty{\typlus[\cs{o}]{A}{B}}
+    \sep        \ty{\tywith[\cs{o}]{A}{B}}
+    \sep        \ty{\tynil[\cs{o}]}
+    \sep        \ty{\tytop[\cs{o}]}
+\end{array}
+\]
 A typing context $\ty\Gamma$ is a partial function associating channel names to types.
@@
 \begin{array}{lcl}
-  \ty{A}, \ty{B}, \ty{C}
-  & \coloneqq & \ty{\tyone[\cs{o}]}
-    \sep        \ty{\tybot[\cs{o}]}
-    \sep        \ty{\tynil[\cs{o}]}
-    \sep        \ty{\tytop[\cs{o}]}
-    \sep        \ty{\tytens[\cs{o}]{A}{B}}
-    \sep        \ty{\typarr[\cs{o}]{A}{B}}
-    \sep        \ty{\typlus[\cs{o}]{A}{B}}
-    \sep        \ty{\tywith[\cs{o}]{A}{B}}
-  \\
   \ty{\Gamma}, \ty{\Delta}
@@
 \]
-
-Duality on types is a total function and is defined below. It preserves priorities of types.
+Duality on types is an involutive function which preserves the priorities of types.
 \[
@@
 \]
-
-Finally, function $\pr(\cdot)$ defined over types returns the priority of that type; function $\minpr(\cdot)$ defined over typing contexts returns the \emph{minimum} priority of all types in that context.
-Both functions are defined as expected below.
+The function $\pr(\cdot)$ returns the priority of a type. The function $\minpr(\cdot)$ returns the \emph{minimum} priority of all types a typing context.
 \[
@@
 
-\begin{definition}[Action]
+\begin{definition}[Actions]
   A~process acts on an endpoint $\tm{x}$ if it is in one of the following forms:
@@
 \begin{proof}
-  If $\tm{P}=\tm{\halt}$, we are done. Otherwise, we move any $\nu$-binders to the top using \LabTirName{SC-ResExt}, and discard any occurrences of $\tm{\halt}$ using \LabTirName{SC-ParNil}.
+  If $\tm{P}=\tm{\halt}$, we are done.
+Otherwise, we move any $\nu$-binders to the top using \LabTirName{SC-ResExt}, and discard any superfluous occurrences of $\tm{\halt}$ using \LabTirName{SC-ParNil}.
 \end{proof}

### 2020-09-27 priority-cp.tex
Updated text.
Added text to PGV. Outlined lemmas for PGV closed progress.

@@
 
-In this section we present an updated version of Priority CP~\cite{dardha2018}, which is Wadler's Classical Processes~\cite{wadler2012} with \emph{priorities}.
-Some details of the calculus are inspired by Carbone \emph{et al.}~\cite{carbone2016} and Caires and P\'{e}rez~\cite{caires2017}.
+We present an updated version of Priority CP~\cite[PCP]{dardha2018}, which is Wadler's Classical Processes~\cite[CP]{wadler2012} with \emph{priorities}.
+The main change is the move away from a reduction strategy based on cut elimination using commuting conversions, and towards a reduction strategy using the structural congruence.
 
 \paragraph*{Priorities, types, and environments}
-Types are based on classical linear logic formulas, and are annotated with priorities $\cs{o}\in\mathbb{N}$.
+Types are based on classical linear logic connectives, and are annotated with priorities $\cs{o}\in\mathbb{N}$.
 
 The types $\ty{\tytens[\cs{o}]{A}{B}}$ and $\ty{\typarr[\cs{o}]{A}{B}}$ type the endpoints of a channel over which we send or receive a channel of type $\ty{A}$, and then proceed as $\ty{B}$.
-The types $\ty{\tyone[\cs{o}]}$ and $\ty{\tybot[\cs{o}]}$ type the endpoints of a channel over which we send or receive a ping, and then close the channel.
+The types $\ty{\tyone[\cs{o}]}$ and $\ty{\tybot[\cs{o}]}$ type the endpoints of a channel whose session has finished, and over which we send or receive a ping before closing the channel.
 These act as the units for $\ty{\tytens[\cs{o}]{A}{B}}$ and $\ty{\typarr[\cs{o}]{A}{B}}$.

### 2020-09-27 priority-gv.tex
Added text to PGV. Outlined lemmas for PGV closed progress.

@@
 \paragraph*{Priorities, types, and environments}
-\[
-  \cs{o}, \cs{p}, \cs{q}, \cs{r} \in \mathbb{N}\cup\{\cs{\pbot},\cs{\ptop}\}
-\]
+Types are divided into \emph{session types} $\ty{S}$ and \emph{value types} $\ty{T}$, $\ty{U}$.
+
+Session types are annotated with \emph{priorities} $\cs{o}\in\mathbb{N}$.
+The types $\ty{\tysend[\cs{o}]{T}{S}}$ and $\ty{\tyrecv[\cs{o}]{T}{S}}$ type the endpoints of a channel over which we send or receive a value of type $\ty{T}$, and then proceed as $\ty{S}$.
+The types $\ty{\tyends[\cs{o}]}$ and $\ty{\tyendr[\cs{o}]}$ type endpoints of a channel whose session has finished, and over which we send or receive a ping before closing the channel.
 \[
@@
     \sep        \ty{\tyendr[\cs{o}]}
-  \\
+\end{array}
+\]
+Value types are based on the types of the linear $\lambda$-calculus.
+The types $\ty{\typrod{T}{U}}$, $\ty{\tyunit}$, $\ty{\tysum{T}{U}}$, and $\ty{\tyvoid}$ are the standard linear $\lambda$-calculus product type, unit type, sum type, and empty type.
+The type $\ty{\tylolli[\cs{p},\cs{q}]{T}{U}}$ is the standard linear function type, annotated with \emph{priority bounds} $\cs{p},\cs{q}\in\mathbb{N}\cup\{\cs{\pbot},\cs{\ptop}\}$.
+Session types are value types, as endpoints are values.
+\[
+\begin{array}{lcl}
   \ty{T}, \ty{U}
-  & \coloneqq & \ty{\tyunit}
+  & \coloneqq & \ty{\typrod{T}{U}}
+    \sep        \ty{\tyunit}
+    \sep        \ty{\tysum{T}{U}}
     \sep        \ty{\tyvoid}
-    \sep        \ty{\typrod{T}{U}}
-    \sep        \ty{\tysum{T}{U}}
     \sep        \ty{\tylolli[\cs{p},\cs{q}]{T}{U}}
     \sep        \ty{S}
-  \\
+\end{array}
+\]
+Given a function with type $\ty{\tylolli[\cs{p},\cs{q}]{T}{U}}$, the priority bound $\cs{p}$ is a \emph{lower bound} on the priorities of the channel names captured by the body of the function, and the priority bound $\cs{q}$ is an \emph{upper bound} on the priority of the communications that take place as a result of applying the function.
+We discuss this in more detail when we give the typing rules.
+We define syntactic sugar for the type of \emph{pure functions}.
+\[
+  \ty{\tylolli{T}{U}} \elabarrow \ty{\tylolli[\cs{\ptop},\cs{\pbot}]{T}{U}}
+\]
+A typing context $\ty\Gamma$ is a partial function associating names to types.
+\[
+\begin{array}{lcl}
   \ty{\Gamma}, \ty{\Delta}
@@
   \pr(\ty{\tyrecv[\cs{o}]{T}{S}})            & = & \cs{o} \\
-  \pr(\ty{\typrod{T}{U}}) & = & \pr({\ty{T}})\sqcap\pr({\ty{U}}) \\
-  \pr(\ty{\tysum{T}{U}})  & = & \pr({\ty{T}})\sqcap\pr({\ty{U}}) \\
-  \pr(\ty{\tylolli[\cs{p},\cs{q}]{T}{U}}) & = & \cs{p} \\
-  \pr(\ty{\Gamma}, \tmty{x}{A}) & = & \pr(\ty{\Gamma})\sqcap\pr(\ty{A})
 \end{array}
@@
   \pr(\ty{\tyendr[\cs{o}]})                  & = & \cs{o} \\
-  \pr(\ty{\tyunit})         & = & \ptop \\
-  \pr(\ty{\tyvoid})         & = & \ptop \\
-  \pr(\ty{S})               & = & \pr(\ty{S}) \\
-  \pr(\ty{\emptyenv})       & = & \ptop
 \end{array}
 \]
-
-\paragraph*{Syntactic Sugar for Types}
 \[
-  \ty{\tylolli{T}{U}} \elabarrow \ty{\tylolli[\cs{\ptop},\cs{\pbot}]{T}{U}}
+\begin{array}{lcl}
+  \minpr(\ty{\typrod{T}{U}})                 & = & \minpr({\ty{T}})\sqcap\minpr({\ty{U}}) \\
+  \minpr(\ty{\tysum{T}{U}})                  & = & \minpr({\ty{T}})\sqcap\minpr({\ty{U}}) \\
+  \minpr(\ty{\tylolli[\cs{p},\cs{q}]{T}{U}}) & = & \cs{p} \\
+  \\
+  \minpr(\ty{\Gamma}, \tmty{x}{A})           & = & \minpr(\ty{\Gamma})\sqcap\pr(\ty{A})
+\end{array}
+\qquad
+\begin{array}{lcl}
+  \minpr(\ty{\tyunit})                       & = & \ptop \\
+  \minpr(\ty{\tyvoid})                       & = & \ptop \\
+  \minpr(\ty{S})                             & = & \pr(\ty{S}) \\
+  \\
+  \minpr(\ty{\emptyenv})                     & = & \ptop
+\end{array}
 \]
@@
   \\
-  \tm{R}
-  & \coloneqq & \tm{K\;V}
-  \\
   \tm{E}
@@
   \label{lem:pgv-value-done}
-  If $\tseq[\cs{o}]{\ty{\Gamma}}{V}{T}$, then $\cs{o}=\cs{\pbot}$, and $\pr(\ty{\Gamma})=\pr(\ty{T})$.
+  If $\tseq[\cs{p}]{\ty{\Gamma}}{V}{T}$, then $\cs{p}=\cs{\pbot}$, and $\minpr(\ty{\Gamma})=\minpr(\ty{T})$.
 \end{restatablelemma}
 \begin{proof}
-  By induction on the derivation of $\tseq[\cs{o}]{\ty{\Gamma}}{V}{T}$.
+  By induction on the derivation of $\tseq[\cs{p}]{\ty{\Gamma}}{V}{T}$.
   (Details in \cref{prf:lem-pgv-value-done}.)
@@
   \hfill\\%newline before theorem statement
-  If $\tseq[\cs{o}]{\ty{\Gamma}}{M}{T}$ and $\tm{M}\tred\tm{M'}$,
-  then $\tseq[\cs{o}]{\ty{\Gamma}}{M'}{T}$.
+  If $\tseq[\cs{p}]{\ty{\Gamma}}{M}{T}$ and $\tm{M}\tred\tm{M'}$,
+  then $\tseq[\cs{p}]{\ty{\Gamma}}{M'}{T}$.
 \end{restatablelemma}
@@
 
+\begin{definition}[Actions]
+  \label{def:pgv-actions}
+  A~term acts on an endpoint $\tm{x}$ if it is in one of the following forms:
+  \begin{multicols}{3}
+    \begin{itemize}[noitemsep,topsep=0pt,parsep=0pt,partopsep=0pt]
+    \item $\tm{\link\;\pair{x}{y}}$ 
+    \item $\tm{\link\;\pair{y}{x}}$
+    \item $\tm{\send\;\pair{x}{y}}$
+    \item $\tm{\recv\;{x}}$
+    \item $\tm{\close\;{x}}$
+    \item $\tm{\wait\;{x}}$
+    \end{itemize}
+  \end{multicols}
+  \noindent
+  A~term is an action if it acts on some endpoint $\tm{x}$.
+\end{definition}
+
+\begin{definition}[Ready Actions]
+  \label{def:pgv-ready-actions}
+  A~term is ready to act on an endpoint $\tm{x}$ if it is of the form $\tm{\plug{E}{M}}$, where $\tm{M}$ acts on $\tm{x}$. A~term is ready if it is ready to act on some endpoint $\tm{x}$.
+\end{definition}
+
+\begin{restatablelemma}{lempgvprogressterms}[Open Progress, $\tred$]
+
+\end{restatablelemma}
+
+\begin{definition}[Canonical Forms]
+  \label{def:pgv-canonical-forms}
+  A~configuration $\tm{\conf{C}}$ is in canonical form if it is of the following form:
+  \[
+    \tm{\res{x_1}{x'_1}{\dots\res{x_n}{x'_n}{(\child\;M_1\parallel\dots\parallel\child\;M_m\parallel\main\;N)}}}
+  \]
+\end{definition}
+
+\begin{restatablelemma}{lempgvcanonicalforms}[Canonical Forms]
+  \label{lem:pgv-canonical-forms}
+  If $\cseq[\main]{\ty{\Gamma}}{\conf{C}}$, there exists some $\tm{\conf{D}}$ such that $\tm{\conf{C}}\equiv\tm{\conf{D}}$ and $\tm{\conf{D}}$ is in canonical form.
+\end{restatablelemma}
+\begin{proof}
+  We move any $\nu$-binders to the top using \LabTirName{SC-ResExt}, discard any superfluous occurrences of $\tm{\child\;\unit}$ using \LabTirName{SC-ParNil}, and move the main thread to the rightmost position using \LabTirName{SC-ParComm} and \LabTirName{SC-ParAssoc}.
+\end{proof}
+
+\begin{restatabletheorem}{thmpgvclosedprogressconfs}[Closed Progress, $\cred$]
+  \label{lem:pgv-closed-progress-confs}
+  If $\cseq[\main]{\emptyenv}{\conf{C}}$ and $\tm{\conf{C}}$ is in canonical form, then either $\tm{\conf{C}}=\tm{\main\;V}$ or there exists a $\tm{\conf{D}}$ such that $\tm{\conf{C}}\cred\tm{\conf{D}}$.
+\end{restatabletheorem}
+\begin{proof}
+  \label{prf:thm-pgv-closed-progress}
+  Let $\tm{\conf{C}}=\tm{\res{x_1}{x'_1}{\dots\res{x_n}{x'_n}{(\child\;M_1\parallel\dots\parallel\child\;M_m\parallel\main\;N)}}}$. If any term $\tm{M_i}$ or $\tm{N}$ is \emph{not} ready, 
+\end{proof}
+
+  % Pick the process $\tm{P_i}$ with the smallest priority $\minpr{(\ty{\Gamma_i})}$.
+$\tm{P_i}$ acts on some endpoint $\tmty{y}{A}\in\ty{\Gamma_i}$.
+We must have $\minpr{(\ty{\Gamma_i})}=\pr{(\ty{A})}$, since the other actions in $\tm{P_i}$ are guarded by the action on $\tmty{y}{A}$.
+
+  % If $\tm{P_j}$ is a link $\tm{\link{y}{z}}$ or $\tm{\link{z}{y}}$, we apply \LabTirName{E-Link}.
+
+  % Otherwise, there must be a process $\seq{P_j}{\ty{\Gamma_j}}$ where $\tmty{y'}{\co{A}}\in\ty{\Gamma_j}$ and $\tm{y}$ and $\tm{y'}$ are dual endpoints of the same channel, since $\tm{y}$ is bound.
+By duality, $\pr{(\ty{A})}=\pr{(\ty{\co{A}})}$.
+We must have $\pr{(\ty{\co{A}})}=\minpr{(\ty{\Gamma_j})}$, since we chose $\tm{P_i}$ to have the smallest priority $\minpr{(\Gamma_i)}$.
+We cannot have $\tm{P_i}=\tm{P_j}$, otherwise the action on $\tm{y'}$ would be guarded by the action on $\tm{y}$, requiring $\pr{(\ty{A})}<\pr{(\ty{\co{A}})}$.
+The process $\tm{P_j}$ must act on $\tm{y'}$, otherwise the action on $\tm{y'}$ would be guarded by another action with priority smaller than $\pr{(\ty{A})}$.
+We have two processes, acting on dual endpoints.
+We apply the appropriate reduction rule, \ie \LabTirName{E-Send}, \LabTirName{E-Close}, \LabTirName{E-Select-Inl}, or \LabTirName{E-Select-Inr}.
+
 \end{document}

### 2020-09-28 priority-cp.tex
a bit of text...

@@
 We present an updated version of Priority CP~\cite[PCP]{dardha2018}, which is Wadler's Classical Processes~\cite[CP]{wadler2012} with \emph{priorities}.
-The main change is the move away from a reduction strategy based on cut elimination using commuting conversions, and towards a reduction strategy using the structural congruence.
+The main difference with respect to PCP by Dardha and Gay \cite{dardha2018} is moving away from a reduction strategy based on cycle elimination using commuting conversions, and towards a reduction strategy using structural congruence.
 
 \paragraph*{Priorities, types, and environments}
-Types are based on classical linear logic connectives, and are annotated with priorities $\cs{o}\in\mathbb{N}$.
+Types are based on classical linear logic propositions, and are annotated with priorities ranging over $\cs{o}\in\mathbb{N}$.
+We let $\ty{A}, \ty{B}, \ty{C}$ range over types, produced by the following grammar.
 
-The types $\ty{\tytens[\cs{o}]{A}{B}}$ and $\ty{\typarr[\cs{o}]{A}{B}}$ type the endpoints of a channel over which we send or receive a channel of type $\ty{A}$, and then proceed as $\ty{B}$.
-The types $\ty{\tyone[\cs{o}]}$ and $\ty{\tybot[\cs{o}]}$ type the endpoints of a channel whose session has finished, and over which we send or receive a ping before closing the channel.
-These act as the units for $\ty{\tytens[\cs{o}]{A}{B}}$ and $\ty{\typarr[\cs{o}]{A}{B}}$.
+The types $\ty{\tytens[\cs{o}]{A}{B}}$ and $\ty{\typarr[\cs{o}]{A}{B}}$ type the endpoints of a channel over which we send or receive a channel of type $\ty{A}$, and then proceed as type $\ty{B}$.
+The types $\ty{\tyone[\cs{o}]}$ and $\ty{\tybot[\cs{o}]}$ type the endpoints of a channel whose session has terminated, and over which we send or receive a \emph{ping} before closing the channel.
+These two types act as units for $\ty{\tytens[\cs{o}]{A}{B}}$ and $\ty{\typarr[\cs{o}]{A}{B}}$, respectively.
 
-The types $\ty{\typlus[\cs{o}]{A}{B}}$ and $\ty{\tywith[\cs{o}]{A}{B}}$ type the endpoints of a channel over which we can make or offer a choice between two options $\ty{A}$ or $\ty{B}$.
-The types $\ty{\tynil[\cs{o}]}$ and $\ty{\tytop[\cs{o}]}$ type the endpoints of a channel over which we can make offer a choice between \emph{no options}.
-These act as the units for $\ty{\typlus[\cs{o}]{A}{B}}$ and $\ty{\tywith[\cs{o}]{A}{B}}$.
+The types $\ty{\typlus[\cs{o}]{A}{B}}$ and $\ty{\tywith[\cs{o}]{A}{B}}$ type the endpoints of a channel over which we can receive or send a choice between two branches $\ty{A}$ or $\ty{B}$.
+We have opted for a simplified version of choice and followed the original Wadler's CP \cite{wadler2012}, however types $\oplus$ and $\with$ can be trivially generalised to $\ty{\oplus^{\cs{o}}\{l_i:A_i\}_{i\in I}}$ and $\ty{\with^{\cs{o}}\{l_i:A_i\}_{i\in I}}$, respectively as in the original PCP \cite{dardha2018}.
+The types $\ty{\tynil[\cs{o}]}$ and $\ty{\tytop[\cs{o}]}$ type the endpoints of a channel over which we can send or receive a choice between \emph{no options}.
+These two types act as units for $\ty{\typlus[\cs{o}]{A}{B}}$ and $\ty{\tywith[\cs{o}]{A}{B}}$, respectively.
 \[
@@
 \paragraph*{Terms}
+We let $\tm{P}, \tm{Q}$ range over processes, produced by the following grammar.
+
+Process $\tm{\link{x}{y}}$ links channels $x$ and $y$ and forwards communication from one to the other.
+Processes $\tm{\res{x}{y}{P}}$, $\tm{(\ppar{P}{Q})}$ and $\tm{\halt}$ denote respectively the restriction processes where channel endpoints $\tm x$ and $\tm y$ are bound together and with scope $\tm P$, the parallel composition of processes $\tm{P}$ and $\tm{Q}$ and the terminated process.
+
+Processes $\tm{\send{x}{y}{P}}$ and $\tm{\recv{x}{y}{P}}$ send or receive over channel $\tm x$ a value $\tm y$ and proceed as process $\tm P$.
+Processes $\tm{\close{x}{P}}$ and $\tm{\wait{x}{P}}$ send and receive an empty value -- denoting the closure of channel $\tm x$, and continue as $\tm P$.
+
+Processes $\tm{\inl{x}{P}}$ and $\tm{\inr{x}{P}}$ make a left and right choice, respectively and proceed as process $\tm P$.
+Process $\tm{\offer{x}{P}{Q}}$ dually offers both left and right branches, with continuations $\tm P$ and $\tm Q$.
+Process $\tm{\absurd{x}}$ denotes the empty offer.
+
 \[
@@
   & \coloneqq & \tm{\link{x}{y}}
-         \sep   \tm{\res{x}{x'}{P}}
+         \sep   \tm{\res{x}{y}{P}}
          \sep   \tm{(\ppar{P}{Q})}
@@
 \]
-
 \paragraph*{Syntactic Sugar}

### 2020-09-28 prf/thm-pgv-closed-progress-confs.tex

\begin{proof}
  \label{prf:thm-pgv-closed-progress-confs}
  Let $\tm{\conf{C}}=\tm{\res{x_1}{x'_1}{\dots\res{x_n}{x'_n}{(\child\;M_1\parallel\dots\parallel\child\;M_m\parallel\main\;N)}}}$.
  We apply \cref{lem:pgv-open-progress-terms} to each $\tm{M_i}$ and $\tm{N}$. If for any $\tm{M_i}$ or $\tm{N}$ we obtain a reduction $\tm{M_i}\tred\tm{M'_i}$ or $\tm{N}\tred\tm{N'}$, we apply \LabTirName{E-LiftM} and \LabTirName{E-LiftC} to obtain a reduction on $\tm{\conf{C}}$.
  Otherwise, each term $\tm{M_i}$ is ready, and $\tm{N}$ is either ready or a value.

  Pick the \emph{ready} term $\tm{L}\in\{\tm{M_1},\dots,\tm{M_m},\tm{N}\}$ with the smallest priority bound. There are four cases:
  \begin{itemize}
  \item
    If $\tm{L}$ is a new $\tm{\plug{E}{\new}}$, we apply \LabTirName{E-New}.
  \item
    If $\tm{L}$ is a spawn $\tm{\plug{E}{\spawn\;M}}$, we apply \LabTirName{E-Spawn}.
  \item
    If $\tm{L}$ is a link $\tm{\plug{E}{\link\;\pair{y}{z}}}$ or $\tm{\plug{E}{\link\;\pair{z}{y}}}$, we apply \LabTirName{E-Link}.
  \end{itemize}
  Otherwise, $\tm{L}$ is ready to act on some endpoint $\tmty{y}{S}$. Let $\tmty{y'}{\co{S}}$ be the dual endpoint of $\tm{y}$. There must be a term $\tm{L'}\in\{\tm{M_1},\dots,\tm{M_m},\tm{N}\}$ which uses $\tm{y'}$. There are two cases:
  \begin{itemize}
  \item
    $\tm{L'}$ is ready. By~\cref{lem:pgv-ready-priority}, the priority of $\tm{L}$ is $\pr(\ty{S})$. By duality, $\pr(\ty{\co{S}})=\pr(\ty{S})$.

    We cannot have $\tm{L}=\tm{L'}$, otherwise the action on $\tm{y'}$ would be guarded by the action on $\tm{y}$, requiring $\pr(\ty{\co{S}})<\pr(\ty{S})$. The term $\tm{L'}$ must be ready to act on $\tm{y'}$, otherwise the action $\tm{y'}$ would be guarded by another action with priority smaller than $\pr{(\ty{S})}$, which contradicts our choice of $\tm{L}$ as having the smallest priority. Therefore, we have two terms ready to act on dual endpoints. We apply the appropriate reduction rule, \ie \LabTirName{E-Send} or \LabTirName{E-Close}.
  \item
    $\tm{L'}=\tm{N}$ and is a value. We rewrite $\tm{\conf{C}}$ to put $\tm{L}$ in the position corresponding to the endpoint it is blocked on, using \LabTirName{SC-ParComm}, \LabTirName{SC-ParAssoc}, and optionally \LabTirName{SC-ResSwap}. We then repeat the steps above with the term with the next smallest priority, until either we find a reduction, or the configuration has reached the desired normal form. (The argument based on the priority being the smallest continues to hold, since we know that neither $\tm{L}$ nor $\tm{L'}$ will be chosen, and no other term uses $\tm{y}$ or $\tm{y'}$.)
  \end{itemize} 
\end{proof}

%%% Local Variables:
%%% TeX-master: "../priorities"
%%% End:

### 2020-09-28 priority-gv.tex

@@
 
+\todo{Define the ``priority bound of a term $\tm{M}$'' as $\cs{p}$ where $\tseq[\cs{p}]{\ty{\Gamma}}{M}{T}$.}
+
 \subsection{Operational Semantics}
@@
     \item $\tm{\link\;\pair{y}{x}}$
-    \item $\tm{\send\;\pair{x}{y}}$
+    \item $\tm{\send\;\pair{V}{x}}$
     \item $\tm{\recv\;{x}}$
@@
 
-\begin{definition}[Ready Actions]
+\begin{definition}[Ready Terms]
   \label{def:pgv-ready-actions}
-  A~term is ready to act on an endpoint $\tm{x}$ if it is of the form $\tm{\plug{E}{M}}$, where $\tm{M}$ acts on $\tm{x}$. A~term is ready if it is ready to act on some endpoint $\tm{x}$.
+  A~term $\tm{L}$ is ready if it is of the form $\tm{\plug{E}{M}}$, where $\tm{M}=\tm{\new}$, $\tm{M}=\tm{\spawn\;N}$, or $\tm{M}$ acts on $\tm{x}$.
+In the latter case, we say that $\tm{L}$ is ready to act on $\tm{x}$.
 \end{definition}
 
-\begin{restatablelemma}{lempgvprogressterms}[Open Progress, $\tred$]
+\begin{restatablelemma}{lempgvreadypriority}
+  \label{lem:pgv-ready-priority}
+  If $\tseq[\cs{p}]{\ty{\Gamma}}{M}{T}$ is ready, then $\cs{p}=\cs{o}$ for some priority $\cs{o}$.
+\end{restatablelemma}
+\begin{proof}
+  \admit
+\end{proof}
 
+\begin{restatablelemma}{lempgvopenprogressterms}[Open Progress, $\tred$]
+  \label{lem:pgv-open-progress-terms}
+  If $\tseq[\cs{p}]{\ty{\Gamma}}{M}{T}$, then:
+  \begin{itemize}
+  \item $\tm{M}$ is a value;
+  \item $\tm{M}\tred\tm{N}$ for some term $\tm{N}$; or
+  \item $\tm{M}$ is ready.
+  \end{itemize}
 \end{restatablelemma}
+\begin{proof}
+  \admit
+\end{proof}
 
@@
   \label{def:pgv-canonical-forms}
-  A~configuration $\tm{\conf{C}}$ is in canonical form if it is of the following form:
+  A~configuration $\tm{\conf{C}}$ is in canonical form if it is of the following form, where no term $\tm{M_i}$ is a value:
   \[
@@
   \label{lem:pgv-closed-progress-confs}
-  If $\cseq[\main]{\emptyenv}{\conf{C}}$ and $\tm{\conf{C}}$ is in canonical form, then either $\tm{\conf{C}}=\tm{\main\;V}$ or there exists a $\tm{\conf{D}}$ such that $\tm{\conf{C}}\cred\tm{\conf{D}}$.
+  If $\cseq[\main]{\emptyenv}{\conf{C}}$ and $\tm{\conf{C}}$ is in canonical form, then either $\tm{\conf{C}}\cred\tm{\conf{D}}$ for some $\tm{\conf{D}}$; or
+  \[
+    \tm{\conf{C}}
+    =
+    \tm{\res{x_1}{x'_1}{\dots\res{x_n}{x'_n}{(\child\;M_1\parallel\dots\parallel\child\;M_m\parallel\main\;V)}}},
+  \]
+  where each $\tm{M_i}$ is ready to act on $\tm{x_i}$.
 \end{restatabletheorem}
 \begin{proof}
-  \label{prf:thm-pgv-closed-progress}
-  Let $\tm{\conf{C}}=\tm{\res{x_1}{x'_1}{\dots\res{x_n}{x'_n}{(\child\;M_1\parallel\dots\parallel\child\;M_m\parallel\main\;N)}}}$. If any term $\tm{M_i}$ or $\tm{N}$ is \emph{not} ready, 
+  Details in \cref{prf:thm-pgv-closed-progress-confs}.
 \end{proof}
-
-  % Pick the process $\tm{P_i}$ with the smallest priority $\minpr{(\ty{\Gamma_i})}$.
-$\tm{P_i}$ acts on some endpoint $\tmty{y}{A}\in\ty{\Gamma_i}$.
-We must have $\minpr{(\ty{\Gamma_i})}=\pr{(\ty{A})}$, since the other actions in $\tm{P_i}$ are guarded by the action on $\tmty{y}{A}$.
-
-  % If $\tm{P_j}$ is a link $\tm{\link{y}{z}}$ or $\tm{\link{z}{y}}$, we apply \LabTirName{E-Link}.
-
-  % Otherwise, there must be a process $\seq{P_j}{\ty{\Gamma_j}}$ where $\tmty{y'}{\co{A}}\in\ty{\Gamma_j}$ and $\tm{y}$ and $\tm{y'}$ are dual endpoints of the same channel, since $\tm{y}$ is bound.
-By duality, $\pr{(\ty{A})}=\pr{(\ty{\co{A}})}$.
-We must have $\pr{(\ty{\co{A}})}=\minpr{(\ty{\Gamma_j})}$, since we chose $\tm{P_i}$ to have the smallest priority $\minpr{(\Gamma_i)}$.
-We cannot have $\tm{P_i}=\tm{P_j}$, otherwise the action on $\tm{y'}$ would be guarded by the action on $\tm{y}$, requiring $\pr{(\ty{A})}<\pr{(\ty{\co{A}})}$.
-The process $\tm{P_j}$ must act on $\tm{y'}$, otherwise the action on $\tm{y'}$ would be guarded by another action with priority smaller than $\pr{(\ty{A})}$.
-We have two processes, acting on dual endpoints.
-We apply the appropriate reduction rule, \ie \LabTirName{E-Send}, \LabTirName{E-Close}, \LabTirName{E-Select-Inl}, or \LabTirName{E-Select-Inr}.
 

### 2020-09-28 priority-cp.tex
typing rules for PCP

@@
+% !TeX root = priorities.tex
 \documentclass[main.tex]{subfiles}
@@
 Process $\tm{\absurd{x}}$ denotes the empty offer.
-
 \[
@@
 \subsection{Typing}
-See \cref{fig:pcp-typing}.
+Typing rules for our version of PCP are given in \cref{fig:pcp-typing}. A typing judgement $\seq{P}{\Gamma}$ states that "process $\tm P$ is well typed under the typing context $\ty \Gamma$".
+
+\textsc{T-Link} states that the link process $\tm{\link{x}{y}}$ is well typed under channels $\tm x$ and $\tm y$ having dual types, respectively $\ty A$ and $\ty {\co A}$.
+\textsc{T-Res} sates that the restriction process $\tm{\res{x}{y}{P}}$ is well typed under typing context $\ty\Gamma$ if process $\tm P$ is well typed in $\ty\Gamma$ augmented with channel endpoints $\tm x$ and $\tm y$ having dual types, respectively $\ty A$ and $\ty {\co A}$.
+\textsc{T-Par} states that the parallel composition of processes $\tm P$ and $\tm Q$ is well typed under the disjoint union of their respective typing contexts.
+\textsc{T-Halt} states that the terminated process $\tm{\halt}$ is well typed in the empty context.
+
+\textsc{T-Send} and \textsc{T-Recv} state that the sending and receiving of a bound name $\tm y$ over a channel $\tm x$ is well typed under $\ty\Gamma$ and $\tm x$ of type $\tytens[\cs{o}]{A}{B}$, respectively ${\typarr[\cs{o}]{A}{B}}$.
+Priority $\cs o$ is the smallest among all priorities of the types used by the output or input process, captured by the side condition $\cs{o}<\minpr(\ty{\Gamma},\ty{A},\ty{B})$.
+
+Rules \textsc{T-Close} and \textsc{T-Wait} type the closure of channel $\tm x$ and are in the same lines as the previous two rules, requiring that the priority of channel $\tm x$ is the smallest among all priorites in $\tm\Gamma$.
+
+\textsc{T-Select-Inl} and \textsc{T-Select-Inr} type respectively the left $\tm{\inl{x}{P}}$ and right $\tm{\inr{x}{P}}$ choice performed on channel $\tm x$.
+\textsc{T-Offer} and \textsc{T-Offer-Absurd} type the offering of a choice, or empty choice, on channel $\tm x$.
+In all the above rules the priority $\cs o$ of channel $\tm x$ is the smallest with respect to the typing context $\cs{o}<\minpr(\ty{\Gamma})$ and types involved in the choice $\cs{o}<\minpr(\ty{\Gamma},\ty{A},\ty{B})$.
+
 \input{fig/pcp-typing}

### 2020-09-28 prf/thm-pgv-closed-progress-confs.tex
Finished proofs. Now time for a consistency pass.

@@
 We then repeat the steps above with the term with the next smallest priority, until either we find a reduction, or the configuration has reached the desired normal form.
-(The argument based on the priority being the smallest continues to hold, since we know that neither $\tm{L}$ nor $\tm{L'}$ will be chosen, and no other term uses $\tm{y}$ or $\tm{y'}$.)
+(The argument based on the priority being the smallest continues to hold, since we know that neither $\tm{L}$ nor $\tm{L'}$ will be picked, and no other term uses $\tm{y}$ or $\tm{y'}$.)
   \end{itemize} 

### 2020-09-28 priority-gv.tex
Finished proofs. Now time for a consistency pass.

@@
   \label{def:pgv-actions}
-  A~term acts on an endpoint $\tm{x}$ if it is in one of the following forms:
-  \begin{multicols}{3}
-    \begin{itemize}[noitemsep,topsep=0pt,parsep=0pt,partopsep=0pt]
-    \item $\tm{\link\;\pair{x}{y}}$ 
-    \item $\tm{\link\;\pair{y}{x}}$
-    \item $\tm{\send\;\pair{V}{x}}$
-    \item $\tm{\recv\;{x}}$
-    \item $\tm{\close\;{x}}$
-    \item $\tm{\wait\;{x}}$
-    \end{itemize}
-  \end{multicols}
-  \noindent
+  A~term acts on an endpoint $\tm{x}$ if it is of the form $\tm{\send\;\pair{V}{x}}$, $\tm{\recv\;{x}}$, $\tm{\close\;{x}}$, or $\tm{\wait\;{x}}$.
 A~term is an action if it acts on some endpoint $\tm{x}$.
@@
   \label{def:pgv-ready-actions}
-  A~term $\tm{L}$ is ready if it is of the form $\tm{\plug{E}{M}}$, where $\tm{M}=\tm{\new}$, $\tm{M}=\tm{\spawn\;N}$, or $\tm{M}$ acts on $\tm{x}$.
-In the latter case, we say that $\tm{L}$ is ready to act on $\tm{x}$.
+  A~term $\tm{L}$ is ready if it is of the form $\tm{\plug{E}{M}}$, where $\tm{M}$ is of the form $\tm{\new}$, $\tm{\spawn\;N}$, $\tm{\link\;\pair{x}{y}}$ or $\tm{\link\;\pair{y}{x}}$, or $\tm{M}$ acts on $\tm{x}$.
+In the last case, we say that $\tm{L}$ is ready to act on $\tm{x}$.
 \end{definition}
@@
   \label{lem:pgv-ready-priority}
-  If $\tseq[\cs{p}]{\ty{\Gamma}}{M}{T}$ is ready, then $\cs{p}=\cs{o}$ for some priority $\cs{o}$.
+  If $\tseq[\cs{p}]{\ty{\Gamma}}{L}{T}$ is ready to act on $\tmty{x}{S}\in\ty{\Gamma}$, then the priority bound $\cs{p}$ is some priority $\cs{o}$, \ie not $\cs{\pbot}$ or $\cs{\ptop}$.
 \end{restatablelemma}
 \begin{proof}
-  \admit
+  Let $\tm{L}=\tm{\plug{E}{M}}$.
+By induction on the structure of $\tm{E}$.
+$\tm{M}$ has priority $\pr({\ty{S}})$, and each constructor of the evaluation context $\tm{E}$ passes on the \emph{maximum} of the priorities of its premises.
+No rule introduces the priority bound $\cs{\ptop}$ on the sequent.
 \end{proof}

### 2020-09-28 prf/lem-pgv-value-done.tex
Minor changes.
Finished review.

@@
 \begin{proof}
+  \label{prf:lem-pgv-value-done}
   By induction on the derivation of $\tseq[\cs{o}]{\ty{\Gamma}}{V}{T}$.
+
   \begin{case*}[\LabTirName{T-Lam}]
@@
       \inferrule*{
-        \tseq[\cs{o}]{\ty{\Gamma},\tmty{x}{T}}{M}{U}
-      }{\tseq[\cs{\pbot}]{\ty{\Gamma}}{\lambda x.M}{\tylolli[\cs{\pr(\ty{\Gamma})},\cs{o}]{T}{U}}}
+        \tseq[\cs{q}]{\ty{\Gamma},\tmty{x}{T}}{M}{U}
+      }{\tseq[\cs{\pbot}]{\ty{\Gamma}}{\lambda x.M}{\tylolli[\cs{\pr(\ty{\Gamma})},\cs{q}]{T}{U}}}
     \end{mathpar}
@@
   \begin{case*}[\LabTirName{T-Inl}]
-    The induction hypothesis gives us $\cs{o}=\cs{\pbot}$, and $\pr(\ty{\Gamma})=\pr{\ty{T}}$. We know $\pr(\ty{T})=\pr({\ty{U}})$, hence $\pr(\ty{\Gamma})=\pr(\ty{\tysum{T}{U}})$.
+    The induction hypothesis gives us $\cs{p}=\cs{\pbot}$, and $\pr(\ty{\Gamma})=\pr{\ty{T}}$. We know $\pr(\ty{T})=\pr({\ty{U}})$, hence $\pr(\ty{\Gamma})=\pr(\ty{\tysum{T}{U}})$.
     \begin{mathpar}
       \inferrule*{
-        \tseq[o]{\ty{\Gamma}}{V}{T}
+        \tseq[\cs{p}]{\ty{\Gamma}}{V}{T}
         \\
         \pr(\ty{T})=\pr(\ty{U})
-      }{\tseq[o]{\ty{\Gamma}}{\inl{V}}{\tysum{T}{U}}}
+      }{\tseq[\cs{p}]{\ty{\Gamma}}{\inl{V}}{\tysum{T}{U}}}
     \end{mathpar}
@@
   \begin{case*}[\LabTirName{T-Inr}]
-    The induction hypothesis gives us $\cs{o}=\cs{\pbot}$, and $\pr(\ty{\Gamma})=\pr{\ty{U}}$. We know $\pr(\ty{T})=\pr({\ty{U}})$, hence $\pr(\ty{\Gamma})=\pr(\ty{\tysum{T}{U}})$.
+    The induction hypothesis gives us $\cs{p}=\cs{\pbot}$, and $\pr(\ty{\Gamma})=\pr{\ty{U}}$. We know $\pr(\ty{T})=\pr({\ty{U}})$, hence $\pr(\ty{\Gamma})=\pr(\ty{\tysum{T}{U}})$.
     \begin{mathpar}
       \inferrule*{
-        \tseq[o]{\ty{\Gamma}}{V}{U}
+        \tseq[\cs{p}]{\ty{\Gamma}}{V}{U}
         \\
         \pr(\ty{T})=\pr(\ty{U})
-      }{\tseq[o]{\ty{\Gamma}}{\inr{V}}{\tysum{T}{U}}}
+      }{\tseq[\cs{p}]{\ty{\Gamma}}{\inr{V}}{\tysum{T}{U}}}
     \end{mathpar}

### 2020-09-28 priority-gv.tex
Added small bits of text.

@@
 \]
+Duality plays a crucial role in session types.
+The two endpoints of a channel are assigned dual types, which ensures that, for instances, whenever one program \emph{sends} a value on a channel, the program on the other end is waiting to \emph{receive}.
+Each \emph{session type} $\ty{T}$ has a dual, written $\ty{\co{T}}$.
+Duality is an involutive function which preserves priorities.
 \[
@@
 \]
+The function $\pr$ returns the smallest priority of a session type.
+The type system guarantees that the top-most connective always holds the smallest priority, so we simply return the priority of the top-most connective.
 \[
@@
 \]
+We extend the function $\pr$ to value types and typing contexts by returning the smallest priority in the type or context, or $\cs{\top}$ if there is no priority.
+\todo{explain function type.}
 \[
@@
 \paragraph*{Static Term Syntax}
-We use $\tm{x}$, $\tm{y}$, $\tm{z}$, and $\tm{w}$ to range over variable names. Occasionally, we use $\tm{a}$, $\tm{b}$, and $\tm{c}$ to range over \emph{free} channel endpoints.
+Let $\tm{x}$, $\tm{y}$, $\tm{z}$, and $\tm{w}$ to range over variable names.
+Occasionally, we use $\tm{a}$, $\tm{b}$, $\tm{c}$, and $\tm{d}$ to range over \emph{free} channel endpoints.
+The term language is the linear $\lambda$-calculus, extended with constants for the communication primitives.
 \[

### 2020-10-01 priority-cp.tex
checked proofs and added notes

@@
 We let $\ty{A}, \ty{B}, \ty{C}$ range over types, produced by the following grammar.
-
-The types $\ty{\tytens[\cs{o}]{A}{B}}$ and $\ty{\typarr[\cs{o}]{A}{B}}$ type the endpoints of a channel over which we send or receive a channel of type $\ty{A}$, and then proceed as type $\ty{B}$.
-The types $\ty{\tyone[\cs{o}]}$ and $\ty{\tybot[\cs{o}]}$ type the endpoints of a channel whose session has terminated, and over which we send or receive a \emph{ping} before closing the channel.
-These two types act as units for $\ty{\tytens[\cs{o}]{A}{B}}$ and $\ty{\typarr[\cs{o}]{A}{B}}$, respectively.
-
-The types $\ty{\typlus[\cs{o}]{A}{B}}$ and $\ty{\tywith[\cs{o}]{A}{B}}$ type the endpoints of a channel over which we can receive or send a choice between two branches $\ty{A}$ or $\ty{B}$.
-We have opted for a simplified version of choice and followed the original Wadler's CP \cite{wadler2012}, however types $\oplus$ and $\with$ can be trivially generalised to $\ty{\oplus^{\cs{o}}\{l_i:A_i\}_{i\in I}}$ and $\ty{\with^{\cs{o}}\{l_i:A_i\}_{i\in I}}$, respectively as in the original PCP \cite{dardha2018}.
-The types $\ty{\tynil[\cs{o}]}$ and $\ty{\tytop[\cs{o}]}$ type the endpoints of a channel over which we can send or receive a choice between \emph{no options}.
-These two types act as units for $\ty{\typlus[\cs{o}]{A}{B}}$ and $\ty{\tywith[\cs{o}]{A}{B}}$, respectively.
 \[
@@
 \]
+The types $\ty{\tytens[\cs{o}]{A}{B}}$ and $\ty{\typarr[\cs{o}]{A}{B}}$ type the endpoints of a channel over which we send or receive a channel of type $\ty{A}$, and then proceed as type $\ty{B}$.
+The types $\ty{\tyone[\cs{o}]}$ and $\ty{\tybot[\cs{o}]}$ type the endpoints of a channel whose session has terminated, and over which we send or receive a \emph{ping} before closing the channel.
+These two types act as units for $\ty{\tytens[\cs{o}]{A}{B}}$ and $\ty{\typarr[\cs{o}]{A}{B}}$, respectively.
+
+The types $\ty{\typlus[\cs{o}]{A}{B}}$ and $\ty{\tywith[\cs{o}]{A}{B}}$ type the endpoints of a channel over which we can receive or send a choice between two branches $\ty{A}$ or $\ty{B}$.
+We have opted for a simplified version of choice and followed the original Wadler's CP \cite{wadler2012}, however types $\oplus$ and $\with$ can be trivially generalised to $\ty{\oplus^{\cs{o}}\{l_i:A_i\}_{i\in I}}$ and $\ty{\with^{\cs{o}}\{l_i:A_i\}_{i\in I}}$, respectively as in the original PCP \cite{dardha2018}.
+The types $\ty{\tynil[\cs{o}]}$ and $\ty{\tytop[\cs{o}]}$ type the endpoints of a channel over which we can send or receive a choice between \emph{no options}.
+These two types act as units for $\ty{\typlus[\cs{o}]{A}{B}}$ and $\ty{\tywith[\cs{o}]{A}{B}}$, respectively.
+
 A typing context $\ty\Gamma$ is a partial function associating channel names to types.
@@
 We let $\tm{P}, \tm{Q}$ range over processes, produced by the following grammar.
-
-Process $\tm{\link{x}{y}}$ links channels $x$ and $y$ and forwards communication from one to the other.
-Processes $\tm{\res{x}{y}{P}}$, $\tm{(\ppar{P}{Q})}$ and $\tm{\halt}$ denote respectively the restriction processes where channel endpoints $\tm x$ and $\tm y$ are bound together and with scope $\tm P$, the parallel composition of processes $\tm{P}$ and $\tm{Q}$ and the terminated process.
-
-Processes $\tm{\send{x}{y}{P}}$ and $\tm{\recv{x}{y}{P}}$ send or receive over channel $\tm x$ a value $\tm y$ and proceed as process $\tm P$.
-Processes $\tm{\close{x}{P}}$ and $\tm{\wait{x}{P}}$ send and receive an empty value -- denoting the closure of channel $\tm x$, and continue as $\tm P$.
-
-Processes $\tm{\inl{x}{P}}$ and $\tm{\inr{x}{P}}$ make a left and right choice, respectively and proceed as process $\tm P$.
-Process $\tm{\offer{x}{P}{Q}}$ dually offers both left and right branches, with continuations $\tm P$ and $\tm Q$.
-Process $\tm{\absurd{x}}$ denotes the empty offer.
 \[
@@
 \]
-\paragraph*{Syntactic Sugar}
-We define \emph{unbound} send:
+Process $\tm{\link{x}{y}}$ links channels $x$ and $y$ and forwards communication from one to the other.
+$\tm{\res{x}{y}{P}}$, $\tm{(\ppar{P}{Q})}$ and $\tm{\halt}$ denote respectively the restriction processes where channel endpoints $\tm x$ and $\tm y$ are bound together and with scope $\tm P$, the parallel composition of processes $\tm{P}$ and $\tm{Q}$ and the terminated process.
+
+Processes $\tm{\send{x}{y}{P}}$ and $\tm{\recv{x}{y}{P}}$ send or receive over channel $\tm x$ a value $\tm y$ and proceed as process $\tm P$.
+$\tm{\close{x}{P}}$ and $\tm{\wait{x}{P}}$ send and receive an empty value -- denoting the closure of channel $\tm x$, and continue as $\tm P$.
+
+Processes $\tm{\inl{x}{P}}$ and $\tm{\inr{x}{P}}$ make a left and right choice, respectively and proceed as process $\tm P$.
+$\tm{\offer{x}{P}{Q}}$ dually offers both left and right branches, with continuations $\tm P$ and $\tm Q$.
+$\tm{\absurd{x}}$ denotes the empty offer.
+
+Finally, we define \emph{unbound} send as syntactic sugar:
 \[
@@
 
+Typing rules make use of priorities, defined in the previous section.
+Priorities are based on \emph{obligations/capabilities} used by Kobayashi \cite{kobayashi2006}, and simplified to single priorities following Padovani \cite{padovani2014}.
+Recalling PCP \cite{dardha2018}, priorities obey the following two laws:
+\begin{enumerate}
+  \item [(i)] an action of priority $\cs o$ must be prefixed only by actions of priorities \emph{strictly smaller} than $\cs o$.
+  \item [(ii)] communication requires \emph{equal} priorities for the dual actions.
+\end{enumerate}
 \textsc{T-Link} states that the link process $\tm{\link{x}{y}}$ is well typed under channels $\tm x$ and $\tm y$ having dual types, respectively $\ty A$ and $\ty {\co A}$.
@@
 \begin{proof}
-  See~\cite{dardha2018}.
+  See~\cite{dardha2018}. \todo{SR in \cite{dardha2018} uses commuting conversions. I guess here the result follows as we are simply considering a subset of reductions?}
 \end{proof}
@@
   \noindent
-  A~process is an action if it acts on on some endpoint $\tm{x}$.
+  A~process is an action if it acts on some endpoint $\tm{x}$.
+  \todo{Can't we simplify this def and say that every process is an action except restriction, par and zero.}
 \end{definition}
@@
 \end{proof}
-
+\todo{We need to comment here why it is that we presented canonical forms, to start with...}
 \begin{restatabletheorem}{thmpcpclosedprogress}[Closed Progress, $\red$]

### 2020-10-01 prf/thm-pcp-closed-progress.tex
checked proofs and added notes
Update thm-pcp-closed-progress.tex

@@
 
-  Pick the process $\tm{P_i}$ with the smallest priority $\minpr{(\ty{\Gamma_i})}$.
+  Our proof follows the same reasoning by Kobayashi \cite{kobayashi2006} used in the proof of deadlock freedom for closed processes (Theorem 2).
+
+  Consider processes $\tm {P_1 \parallel\dots\parallel P_m}$.
+Among them, we pick the process with the smallest priority $\minpr{(\ty{\Gamma_i})}$ for all $i$.
+Let this process be let this be $\tm{P_i}$ and the priority of the top prefix be $\cs o$.
 $\tm{P_i}$ acts on some endpoint $\tmty{y}{A}\in\ty{\Gamma_i}$.
-We must have $\minpr{(\ty{\Gamma_i})}=\pr{(\ty{A})}$, since the other actions in $\tm{P_i}$ are guarded by the action on $\tmty{y}{A}$.
+We must have $\minpr{(\ty{\Gamma_i})}=\pr{(\ty{A})} = \cs o$, since the other actions in $\tm{P_i}$ are guarded by the action on $\tmty{y}{A}$, thus satisfying law (i) of priorities.
 
-  If $\tm{P_j}$ is a link $\tm{\link{y}{z}}$ or $\tm{\link{z}{y}}$, we apply \LabTirName{E-Link}.
+  If $\tm{P_i}$ is a link $\tm{\link{y}{z}}$ or $\tm{\link{z}{y}}$, we apply \LabTirName{E-Link}.
 
-  Otherwise, there must be a process $\seq{P_j}{\ty{\Gamma_j}}$ where $\tmty{y'}{\co{A}}\in\ty{\Gamma_j}$ and $\tm{y}$ and $\tm{y'}$ are dual endpoints of the same channel, since $\tm{y}$ is bound.
-By duality, $\pr{(\ty{A})}=\pr{(\ty{\co{A}})}$.
-We must have $\pr{(\ty{\co{A}})}=\minpr{(\ty{\Gamma_j})}$, since we chose $\tm{P_i}$ to have the smallest priority $\minpr{(\Gamma_i)}$.
-We cannot have $\tm{P_i}=\tm{P_j}$, otherwise the action on $\tm{y'}$ would be guarded by the action on $\tm{y}$, requiring $\pr{(\ty{A})}<\pr{(\ty{\co{A}})}$.
-The process $\tm{P_j}$ must act on $\tm{y'}$, otherwise the action on $\tm{y'}$ would be guarded by another action with priority smaller than $\pr{(\ty{A})}$.
-We have two processes, acting on dual endpoints.
-We apply the appropriate reduction rule, \ie \LabTirName{E-Send}, \LabTirName{E-Close}, \LabTirName{E-Select-Inl}, or \LabTirName{E-Select-Inr}.
+  Otherwise, $\tm{P_i}$ is an input/branching or output/selection action on endpoint $\tm y$ of type $\ty A$ with priority $\cs o$.
+Since process $\tm P$ is closed and consequently it respects law (ii) of priorities, there must be a co-action $y'$ of type $\ty{\co{A}}$  where $\tm{y}$ and $\tm{y'}$ are dual endpoints of the same channel (by application of rule \textsc{T-Res}).
+By duality, $\pr{(\ty{A})}=\pr{(\ty{\co{A}})}= \cs o$.
+In the following we show that: $y'$ is the subject of a top level action of a process $P_j$ with $i\neq j$.
+This allows for the communication among $\tm{P_i}$ and $\tm{P_j}$ to happen immediately over channel endpoints $\tm{y}$ and $\tm{y'}$.
+
+  Suppose that $\tm{y'}$ is an action not in a different parallel process $P_j$ but rather of $P_i$ itself.
+That means that the action on $\tm{y'}$ must be prefixed by the action on $\tm{y}$, which is top level in $\tm{P_i}$.
+To respect law (i) of priorities we must have $\cs o < \cs o$, which is absurd.
+This means that $\tm{y'}$ is in another parallel process $\tm{P_j}$ for $i\neq j$.
+
+  Suppose that $\tm{y'}$ in $\tm{P_j}$ is not at top level.
+In order to respect law (i) of priorities, it means that $\tm{y'}$ is prefixed by actions that are smaller than its priority $\cs o$.
+This leads to a contradiction because stated that $\cs o$ is the smallest priority.
+Hence, $y'$ must be the subject of a top level action.
+
+  We have two processes, acting on dual endpoints. We apply the appropriate reduction rule, \ie \LabTirName{E-Send}, \LabTirName{E-Close}, \LabTirName{E-Select-Inl}, or \LabTirName{E-Select-Inr}.
 \end{proof}

### 2020-10-07 priority-cp.tex
on operational semantics

@@
 \subsection{Typing}
-Typing rules for our version of PCP are given in \cref{fig:pcp-typing}. A typing judgement $\seq{P}{\Gamma}$ states that "process $\tm P$ is well typed under the typing context $\ty \Gamma$".
+\input{fig/pcp-typing}
+
+Typing rules for our version of PCP are given in \Cref{fig:pcp-typing}. A typing judgement $\seq{P}{\Gamma}$ states that "process $\tm P$ is well typed under the typing context $\ty \Gamma$".
 
@@
 
-\input{fig/pcp-typing}
+\subsection{Operational Semantics}
+\input{fig/pcp-operational-semantics}
 
-\subsection{Operational Semantics}
-See \cref{fig:pcp-operational-semantics}.
-\input{fig/pcp-operational-semantics}
+The operational semantics for PCP given in \Cref{fig:pcp-operational-semantics} is defined as a reduction relation $\red$ on processes (bottom) and uses structural congruence (top).
+
+Structural congruence satisfies the following axioms.
+$\LabTirName{SC-LinkSwap}$ allows swapping channels in the link process.
+$\LabTirName{SC-ResLink}$ allows restriction to applied to link which is structurally equivalent to the terminated process, thus allowing elimination of unnecessary restrictions.
+$\LabTirName{SC-ResSwap}$ allows swapping channels and $\LabTirName{SC-ResComm}$ states that restriction is commutative.
+$\LabTirName{SC-ResExt}$ is the standard scope extrusion rule.
+Rules $\LabTirName{SC-ParNil}$, $\LabTirName{SC-ParComm}$ and $\LabTirName{SC-ParAssoc}$ state that parallel composition uses the terminated proceess as the neutral element; it is commutative and associative.
+
+The reduction relation is given by a set of axioms and inference rules for context closure. Reduction occurs under restriction following \cite{dardha2018} where channel endpoints are identified.
+$\LabTirName{E-Link}$ reduces a parallel composition with a link into a substitution.
+$\LabTirName{E-Send}$ is the main communication rule, where send and receive processes sychronise and reduce to the corresponding continuations.
+$\LabTirName{E-Close}$ follows the previous rule and it closes the channel identified by endpoints $\tm x$ and $\tm y$.
+$\LabTirName{E-Select-Inl}$ and $\LabTirName{E-Select-Inr}$ are generalised versions of $\LabTirName{E-Send}$.
+They state respectively that a left and right selection synchronises with a choice offering and reduces to the corresponding continuations.
+The last three rules state that reduction is closed under restriction, parallel composition and structural congruence, respectively.
 
@@
 \begin{proof}
-  Details in~\cref{prf:thm-pcp-closed-progress}.
+  Details in~\Cref{prf:thm-pcp-closed-progress}.
 \end{proof}

### 2020-10-07 priority-cp.tex
adding scheduler in PCP

@@
 
-Rules \textsc{T-Close} and \textsc{T-Wait} type the closure of channel $\tm x$ and are in the same lines as the previous two rules, requiring that the priority of channel $\tm x$ is the smallest among all priorites in $\tm\Gamma$.
+Rules \textsc{T-Close} and \textsc{T-Wait} type the closure of channel $\tm x$ and are in the same lines as the previous two rules, requiring that the priority of channel $\tm x$ is the smallest among all priorities in $\tm\Gamma$.
 
@@
 In all the above rules the priority $\cs o$ of channel $\tm x$ is the smallest with respect to the typing context $\cs{o}<\minpr(\ty{\Gamma})$ and types involved in the choice $\cs{o}<\minpr(\ty{\Gamma},\ty{A},\ty{B})$.
+
+\begin{example}[Milner's Cyclic Scheduler \Cref{fig:scheduler}]
+  % 
+  We will illustrate PCP with the cyclic scheduler \cite{milner99}. The full typing for $\tm {Sched}$ is given in \cite{dardha2018}.
+
+  We will consider a simplified version: a set of 3 agents $\tm {A_0}, \tm {A_1}, \tm {A_2}$ is scheduled to perform a certain task in cyclic order, starting with agent $\tm {A_0}$.
+Agents $\tm {A_1}$ and $\tm {A_2}$ will send the result of some computation a collector process, respectively $\tm {P_1}$ and $\tm {P_2}$, before transmitting data further to the next agent in the cycle, being $\tm {A_{(i+1)\texttt{ mod } 3}}$.
+At the end of the round, agent $\tm {A_0}$ will send the final result of the computation to its collector $\tm {P_0}$.
+
+  In line with the logical consistency requirements of the Curry-Howard correspondence, we define a finite version of Milner's scheduler, which executes one round of communication.
+  $$
+    \begin{array}{rcll}
+      \tm {Sched} & \defeq & ..\tm {(\nu {a_i}{b_i})}..\tm {(\nu {c_i}{d_{{(i+1)\texttt{ mod } n}}})} \tm {\big(A_0 \parallel A_1 \parallel A_2 \parallel P_0 \parallel P_1 \parallel P_2 \big)}\\
+      \tm{A_0} & \defeq & \tm{\send{c_0}{n_0}{\recv{d_0}{x_0}{\send{a_0}{m_0}{close_0}}}}\\
+      \tm{A_i} & \defeq & \tm{\recv{d_i}{x_i}{\send{a_i}{m_i}{\send{c_i}{n_i}{close_i}}}} &~~ i \in \{1,2\}\\
+      \tm {P_i} & \defeq & \tm{\recv{b_i}{y_i}{Q_i}} &~~ i \in \{0,1,2\}
+    \end{array}
+  $$
+
+  In the above process, channels $n_i$ and $m_i$ for all $i$, denote data being transmitted, and for simplicity we let them be channels of a terminated type.
+Process $\tm{close_i}$ for all $i$ closes all channels in $\tm{A_i}$.
+In particular, they are a sequence of $\tm{\close{x}{P}}$ and $\tm{\wait{x}{P}}$ for all channels in the agent and terminating in $\tm{\halt}$.
+\end{example}
 

### 2020-10-07 conclusion.tex
new files for intro and conclusion

--- 2020-10-01 priority-cp.tex
@@
 \begin{document}
-
-\section{Priority CP}
-\usingnamespace{pcp}
-
-\begin{figure}[t]
-  \centering
-  \includegraphics[width=0.8\columnwidth]{scheduler}
-  \vspace{-4mm}
-  \caption{Cyclic Scheduler from \cite{dardha2018}}
-  \label{fig:scheduler}
-  \vspace{-4mm}
-  \end{figure}
-
-\subsection{Syntax}
-
-We present an updated version of Priority CP~\cite[PCP]{dardha2018}, which is Wadler's Classical Processes~\cite[CP]{wadler2012} with \emph{priorities}.
-The main difference with respect to PCP by Dardha and Gay \cite{dardha2018} is moving away from a reduction strategy based on cycle elimination using commuting conversions, and towards a reduction strategy using structural congruence.
-
-\paragraph*{Priorities, types, and environments}
-Types are based on classical linear logic propositions, and are annotated with priorities ranging over $\cs{o}\in\mathbb{N}$.
-We let $\ty{A}, \ty{B}, \ty{C}$ range over types, produced by the following grammar.
-\[
-\begin{array}{lcl}
-  \ty{A}, \ty{B}, \ty{C}
-  & \coloneqq & \ty{\tytens[\cs{o}]{A}{B}}
-    \sep        \ty{\typarr[\cs{o}]{A}{B}}
-    \sep        \ty{\tyone[\cs{o}]}
-    \sep        \ty{\tybot[\cs{o}]}
-    \sep        \ty{\typlus[\cs{o}]{A}{B}}
-    \sep        \ty{\tywith[\cs{o}]{A}{B}}
-    \sep        \ty{\tynil[\cs{o}]}
-    \sep        \ty{\tytop[\cs{o}]}
-\end{array}
-\]
-The types $\ty{\tytens[\cs{o}]{A}{B}}$ and $\ty{\typarr[\cs{o}]{A}{B}}$ type the endpoints of a channel over which we send or receive a channel of type $\ty{A}$, and then proceed as type $\ty{B}$.
-The types $\ty{\tyone[\cs{o}]}$ and $\ty{\tybot[\cs{o}]}$ type the endpoints of a channel whose session has terminated, and over which we send or receive a \emph{ping} before closing the channel.
-These two types act as units for $\ty{\tytens[\cs{o}]{A}{B}}$ and $\ty{\typarr[\cs{o}]{A}{B}}$, respectively.
-
-The types $\ty{\typlus[\cs{o}]{A}{B}}$ and $\ty{\tywith[\cs{o}]{A}{B}}$ type the endpoints of a channel over which we can receive or send a choice between two branches $\ty{A}$ or $\ty{B}$.
-We have opted for a simplified version of choice and followed the original Wadler's CP \cite{wadler2012}, however types $\oplus$ and $\with$ can be trivially generalised to $\ty{\oplus^{\cs{o}}\{l_i:A_i\}_{i\in I}}$ and $\ty{\with^{\cs{o}}\{l_i:A_i\}_{i\in I}}$, respectively as in the original PCP \cite{dardha2018}.
-The types $\ty{\tynil[\cs{o}]}$ and $\ty{\tytop[\cs{o}]}$ type the endpoints of a channel over which we can send or receive a choice between \emph{no options}.
-These two types act as units for $\ty{\typlus[\cs{o}]{A}{B}}$ and $\ty{\tywith[\cs{o}]{A}{B}}$, respectively.
-
-A typing context $\ty\Gamma$ is a partial function associating channel names to types.
-\[
-\begin{array}{lcl}
-  \ty{\Gamma}, \ty{\Delta}
-  & \coloneqq & \ty{\emptyenv}
-    \sep        \ty{\Gamma}, \tmty{x}{A}
-\end{array}
-\]
-Duality on types is an involutive function which preserves the priorities of types.
-\[
-\begin{array}{lcl}
-  \ty{\co{(\tyone[\cs{o}])}} & = & \ty{\tybot[\cs{o}]} \\
-  \ty{\co{(\tybot[\cs{o}])}} & = & \ty{\tyone[\cs{o}]} \\
-  \ty{\co{(\tynil[\cs{o}])}} & = & \ty{\tytop[\cs{o}]} \\
-  \ty{\co{(\tytop[\cs{o}])}} & = & \ty{\tynil[\cs{o}]}
-\end{array}
-\qquad
-\begin{array}{lcl}
-  \ty{\co{(\tytens[\cs{o}]{A}{B})}} & = & \ty{\typarr[\cs{o}]{\co{A}}{\co{B}}} \\
-  \ty{\co{(\typarr[\cs{o}]{A}{B})}} & = & \ty{\tytens[\cs{o}]{\co{A}}{\co{B}}} \\
-  \ty{\co{(\typlus[\cs{o}]{A}{B})}} & = & \ty{\tywith[\cs{o}]{\co{A}}{\co{B}}} \\
-  \ty{\co{(\tywith[\cs{o}]{A}{B})}} & = & \ty{\typlus[\cs{o}]{\co{A}}{\co{B}}}
-\end{array}
-\]
-The function $\pr(\cdot)$ returns the priority of a type. The function $\minpr(\cdot)$ returns the \emph{minimum} priority of all types a typing context.
-\[
-\begin{array}{lclclcl}
-  \pr(\ty{\tyone[\cs{o}]})        & = & \cs{o}  \\
-  \pr(\ty{\tybot[\cs{o}]})        & = & \cs{o}  \\
-  \pr(\ty{\tynil[\cs{o}]})        & = & \cs{o}  \\
-  \pr(\ty{\tytop[\cs{o}]})        & = & \cs{o}  \\
-  \\
-  \minpr(\ty{\emptyenv})          & = & \varnothing
-\end{array}
-\qquad
-\begin{array}{lclclcl}
-  \pr(\ty{\tytens[\cs{o}]{A}{B}}) & = & \cs{o}  \\
-  \pr(\ty{\typarr[\cs{o}]{A}{B}}) & = & \cs{o}  \\
-  \pr(\ty{\typlus[\cs{o}]{A}{B}}) & = & \cs{o}  \\
-  \pr(\ty{\tywith[\cs{o}]{A}{B}}) & = & \cs{o}  \\
-  \\
-  \minpr(\ty{\Gamma},\tmty{x}{A}) & = & \minpr(\ty{\Gamma})\sqcap\minpr(\ty{A})
-\end{array}
-\]
-
-\paragraph*{Terms}
-We let $\tm{P}, \tm{Q}$ range over processes, produced by the following grammar.
-\[
-\begin{array}[t]{lcl}
-  \tm{P}, \tm{Q}
-  & \coloneqq & \tm{\link{x}{y}}
-         \sep   \tm{\res{x}{y}{P}}
-         \sep   \tm{(\ppar{P}{Q})}
-         \sep   \tm{\halt}
-  \\   & \sep & \tm{\send{x}{y}{P}}
-         \sep   \tm{\close{x}{P}}
-         \sep   \tm{\recv{x}{y}{P}}
-         \sep   \tm{\wait{x}{P}}
-  \\   & \sep & \tm{\inl{x}{P}}
-         \sep   \tm{\inr{x}{P}}
-         \sep   \tm{\offer{x}{P}{Q}}
-         \sep   \tm{\absurd{x}}
-\end{array}
-\]
-Process $\tm{\link{x}{y}}$ links channels $x$ and $y$ and forwards communication from one to the other.
-$\tm{\res{x}{y}{P}}$, $\tm{(\ppar{P}{Q})}$ and $\tm{\halt}$ denote respectively the restriction processes where channel endpoints $\tm x$ and $\tm y$ are bound together and with scope $\tm P$, the parallel composition of processes $\tm{P}$ and $\tm{Q}$ and the terminated process.
-
-Processes $\tm{\send{x}{y}{P}}$ and $\tm{\recv{x}{y}{P}}$ send or receive over channel $\tm x$ a value $\tm y$ and proceed as process $\tm P$.
-$\tm{\close{x}{P}}$ and $\tm{\wait{x}{P}}$ send and receive an empty value -- denoting the closure of channel $\tm x$, and continue as $\tm P$.
-
-Processes $\tm{\inl{x}{P}}$ and $\tm{\inr{x}{P}}$ make a left and right choice, respectively and proceed as process $\tm P$.
-$\tm{\offer{x}{P}{Q}}$ dually offers both left and right branches, with continuations $\tm P$ and $\tm Q$.
-$\tm{\absurd{x}}$ denotes the empty offer.
-
-Finally, we define \emph{unbound} send as syntactic sugar:
-\[
-  \tm{\usend{x}{y}{P}}\elabarrow\tm{\send{x}{z}{(\ppar{\link{y}{z}}{P})}}
-\]
-
-\subsection{Typing}
-Typing rules for our version of PCP are given in \cref{fig:pcp-typing}. A typing judgement $\seq{P}{\Gamma}$ states that "process $\tm P$ is well typed under the typing context $\ty \Gamma$".
-
-Typing rules make use of priorities, defined in the previous section.
-Priorities are based on \emph{obligations/capabilities} used by Kobayashi \cite{kobayashi2006}, and simplified to single priorities following Padovani \cite{padovani2014}.
-Recalling PCP \cite{dardha2018}, priorities obey the following two laws:
-\begin{enumerate}
-  \item [(i)] an action of priority $\cs o$ must be prefixed only by actions of priorities \emph{strictly smaller} than $\cs o$.
-  \item [(ii)] communication requires \emph{equal} priorities for the dual actions.
-\end{enumerate}
-\textsc{T-Link} states that the link process $\tm{\link{x}{y}}$ is well typed under channels $\tm x$ and $\tm y$ having dual types, respectively $\ty A$ and $\ty {\co A}$.
-\textsc{T-Res} sates that the restriction process $\tm{\res{x}{y}{P}}$ is well typed under typing context $\ty\Gamma$ if process $\tm P$ is well typed in $\ty\Gamma$ augmented with channel endpoints $\tm x$ and $\tm y$ having dual types, respectively $\ty A$ and $\ty {\co A}$.
-\textsc{T-Par} states that the parallel composition of processes $\tm P$ and $\tm Q$ is well typed under the disjoint union of their respective typing contexts.
-\textsc{T-Halt} states that the terminated process $\tm{\halt}$ is well typed in the empty context.
-
-\textsc{T-Send} and \textsc{T-Recv} state that the sending and receiving of a bound name $\tm y$ over a channel $\tm x$ is well typed under $\ty\Gamma$ and $\tm x$ of type $\tytens[\cs{o}]{A}{B}$, respectively ${\typarr[\cs{o}]{A}{B}}$.
-Priority $\cs o$ is the smallest among all priorities of the types used by the output or input process, captured by the side condition $\cs{o}<\minpr(\ty{\Gamma},\ty{A},\ty{B})$.
-
-Rules \textsc{T-Close} and \textsc{T-Wait} type the closure of channel $\tm x$ and are in the same lines as the previous two rules, requiring that the priority of channel $\tm x$ is the smallest among all priorites in $\tm\Gamma$.
-
-\textsc{T-Select-Inl} and \textsc{T-Select-Inr} type respectively the left $\tm{\inl{x}{P}}$ and right $\tm{\inr{x}{P}}$ choice performed on channel $\tm x$.
-\textsc{T-Offer} and \textsc{T-Offer-Absurd} type the offering of a choice, or empty choice, on channel $\tm x$.
-In all the above rules the priority $\cs o$ of channel $\tm x$ is the smallest with respect to the typing context $\cs{o}<\minpr(\ty{\Gamma})$ and types involved in the choice $\cs{o}<\minpr(\ty{\Gamma},\ty{A},\ty{B})$.
-
-\input{fig/pcp-typing}
-
-\subsection{Operational Semantics}
-See \cref{fig:pcp-operational-semantics}.
-\input{fig/pcp-operational-semantics}
-
-\subsection{Metatheory}
-
-\begin{restatablelemma}{lempcpsubjectcongruence}[Subject Congruence, $\equiv$]
-  \label{lem:pcp-subject-congruence}
-  \hfill\\%newline before theorem statement
-  If $\seq{P}{\ty{\Gamma}}$ and $\tm{P}\equiv\tm{Q}$,
-  then $\seq{Q}{\ty{\Gamma}}$.
-\end{restatablelemma}
-\begin{proof}
-  See~\cite{dardha2018}.
-\end{proof}
-
-\begin{restatabletheorem}{thmpcpsubjectreduction}[Subject Reduction, $\red$]
-  \label{thm:pcp-subject-reduction}
-  \hfill\\%newline before theorem statement
-  If $\seq{P}{\ty{\Gamma}}$ and $\tm{P}\red\tm{Q}$, then $\seq{Q}{\ty{\Gamma}}$.
-\end{restatabletheorem}
-\begin{proof}
-  See~\cite{dardha2018}. \todo{SR in \cite{dardha2018} uses commuting conversions. I guess here the result follows as we are simply considering a subset of reductions?}
-\end{proof}
-
-\begin{definition}[Actions]
-  A~process acts on an endpoint $\tm{x}$ if it is in one of the following forms:
-  \begin{multicols}{3}
-    \begin{itemize}[noitemsep,topsep=0pt,parsep=0pt,partopsep=0pt]
-    \item $\tm{\link{x}{y}}$ 
-    \item $\tm{\link{y}{x}}$
-    \item $\tm{\send{x}{y}{P}}$
-    \item $\tm{\recv{x}{y}{P}}$
-    \item $\tm{\close{x}{P}}$
-    \item $\tm{\wait{x}{P}}$
-    \item $\tm{\inl{x}{P}}$
-    \item $\tm{\inr{x}{P}}$
-    \item $\tm{\offer{x}{P}{Q}}$
-    \item $\tm{\absurd{x}}$
-    \end{itemize}
-  \end{multicols}
-  \noindent
-  A~process is an action if it acts on some endpoint $\tm{x}$.
-  \todo{Can't we simplify this def and say that every process is an action except restriction, par and zero.}
-\end{definition}
-
-\begin{definition}[Canonical Forms]
-  \label{def:pcp-canonical-forms}
-  A~process $\tm{P}$ is in canonical form if it is in one of the following forms:
-  \begin{itemize}[noitemsep,topsep=0pt,parsep=0pt,partopsep=0pt]
-  \item
-    $\tm{\halt}$
-  \item
-    $\tm{\res{x_1}{x'_1}{\dots\res{x_n}{x'_n}{(P_1 \parallel\dots\parallel P_m)}}}$
-    where $m>0$ and each $\tm{P_j}$ is an action.
-  \end{itemize}
-\end{definition}
-
-\begin{restatablelemma}{lempcpcanonicalforms}[Canonical Forms]
-  \label{lem:pcp-canonical-forms}
-  If $\seq{P}{\ty{\Gamma}}$, there exists some $\tm{Q}$ such that $\tm{P}\equiv\tm{Q}$ and $\tm{Q}$ is in canonical form.
-\end{restatablelemma}
-\begin{proof}
-  If $\tm{P}=\tm{\halt}$, we are done.
-Otherwise, we move any $\nu$-binders to the top using \LabTirName{SC-ResExt}, and discard any superfluous occurrences of $\tm{\halt}$ using \LabTirName{SC-ParNil}.
-\end{proof}
-\todo{We need to comment here why it is that we presented canonical forms, to start with...}
-\begin{restatabletheorem}{thmpcpclosedprogress}[Closed Progress, $\red$]
-  \label{thm:pcp-closed-progress}
-  \hfill\\%newline before theorem statement
-  If $\seq{P}{\emptyenv}$ and $\tm{P}$ is in canonical form, then either $\tm{P}=\tm{\halt}$ or there exists a $\tm{Q}$ such that $\tm{P}\red\tm{Q}$.
-\end{restatabletheorem}
-\begin{proof}
-  Details in~\cref{prf:thm-pcp-closed-progress}.
-\end{proof}
-
+\section{Related Work and Conclusion}
 \end{document}
-
-%%% Local Variables:
-%%% TeX-master: "priorities"
-%%% End:

### 2020-10-07 introduction.tex
new files for intro and conclusion
to discuss in the intro

--- 2020-10-07 conclusion.tex
@@
 \begin{document}
-\section{Related Work and Conclusion}
+\section{Introduction}
+
+Things to discuss in the intro:
+
+-- talking about what is GV and its "benefits"
+-- next, talking about deadlock freedom property, the benefits and the fact that the original GVs do not satisfy DF
+-- in the meantime, in a concurrent setting based on pi-calculus there is a curry-howard correspondence of logic and session typed pi, where DF holds
+-- talk about CP here
+-- \cite{dardha2018} brought the research fwd and added priorities a la Kobayashi and Padovani to CP, producing PCP
+-- following that lead, we present PGV
+-- running example of the scheduler in PGV (hopefully)
+-- translation of PCP to PGV and what it means to have that.
+
+a contribution paragraph with the following
+-- PGV and what are the benefits
+-- updated PCP
+-- encoding PCP to PGV
+-- technical results to be referred to explicitly
+
 \end{document}

### 2020-10-07 conclusion.tex
a start on related work, with copy-paste

@@
 \section{Related Work and Conclusion}
+
+\paragraph{On GV}
+
+\paragraph{On priorities}
+In Kobayashi's original type-theoretic approach to deadlock-freedom \cite{K98}, priorities were abstract tags from a partially ordered set.
+In later work abstract tags were simplified to natural numbers, and priorities were replaced by pairs of obligations and capabilities \cite{K02,K06}.
+The latter change allows more processes to be typed, at the expense of a more complex type system.
+Padovani \cite{P13} adapted Kobayashi's approach to session types, and later on he simplified it to a single priority for linear $\pi$-calculus \cite{P14}.
+Then, the single priority technique can be transferred to session types by the encoding of session types into linear types \cite{Koba07,DGS12,D14,Dardha16}.
+For simplicity, we have opted for single priorities, as Padovani \cite{P14}.
+
+\paragraph{On progress in session types}
+The first work on progress for session types, by Dezani-Ciancaglini \emph{et al.} \cite{DMYD06}, guaranteed the property by allowing only one active session at a time.
+Later work \cite{DLY07} introduced a partial order on channels in Kobayashi-style \cite{K98}.
+Bettini \emph{et al.} \cite{BCDDDY08} applied similar ideas to multiparty session types.
+The main difference with our work is that we associate priorities with individual communication operations, rather than with entire channels.
+Carbone \emph{et al.} \cite{CDM14} proved that progress is a compositional form of lock-freedom and introduced a new technique for progress in session types by adopting Kobayashi's type system and the encoding of session types \cite{DGS12}.
+Vieira and Vasconcelos \cite{VieiraV13} used single priorities and an abstract partial order in session types to guarantee deadlock-freedom.
+
+\paragraph{On Curry-Howard correspondence}
+The linear logic approach to deadlock-free session types started with Caires and Pfenning \cite{CP10}, based on dual intuitionistic linear logic, and was later formulated for classical linear logic by Wadler \cite{wadler2012}.
+All subsequent work on linear logic and session types enforces deadlock-freedom by forbidding cyclic connections.
+In their original work, Caires and Pfenning commented that it would be interesting to compare process typability in their system with other approaches including Kobayashi's and Dezani-Ciancaglini's.
+However, we are aware of only one comparative study of the expressivity of type systems for deadlock-freedom, by Dardha and P\'{e}rez \cite{DardhaP15}.
+They compared Kobayashi-style typing and CLL typing, and proved that CLL corresponds to Kobayashi's system with the restriction that only single cuts, not multicuts, are allowed.
+
+\paragraph{Conclusion}
+In this paper...
 \end{document}

### 2020-10-08 conclusion.tex
more work on related work

--- 2020-10-07 conclusion.tex
@@
 \section{Related Work and Conclusion}
+
+\paragraph{On GV}
+
+\paragraph{On deadlock freedom and progress in the $\pi$-calculus}
+Deadlock freedom and progress have been studied thoroughly in the $\pi$-calculus with or without session types.
+
+In the standard typed $\pi$-calculus a seminal line of work is by Kobayashi and his original type-theoretic approach to deadlock-freedom \cite{K98}, where priorities are abstract tags defined over a partially ordered set.
+Later on these abstract tags were simplified to natural numbers, and pairs of obligations and capabilities were used in the type system for deadlock freedom \cite{K02,K06}.
+This allowed more $\pi$-calculus processes to be typed.
+Following, Kobayashi's line of work, Padovani \cite{P13} adapted obligation/capability pairs to session types, and later on he further simplified them to a single priority for linear $\pi$-calculus \cite{P14}.
+As discussed by the author, by using the encoding of session types into linear types \cite{Koba07,DGS12,D14,Dardha16}, the deadlock freedom technique of the linear $\pi$-calculus with priorities can be transferred onto the $\pi$-calculus with session types.
+
+The seminal work on progress for session-typed $\pi$-calculus, by Dezani \emph{et al.} \cite{DMYD06}, guarantees progress by allowing only one active session at a time.
+In \cite{DLY07} the authors introduce a partial order on channels in line with Kobayashi's work \cite{K98}.
+Carbone and Debois \cite{CD10} define progress for session typed $\pi$-calculus in terms of a \emph{catalyser} used to provide a missing counterpart to a process, thus guaranteeing progress.
+Carbone \emph{et al.} \cite{CDM14} studied further the use of catalysers and showed that progress is a compositional form of lock-freedom.
+As in \cite{P14} the authors show that by using the encoding of session types \cite{DGS12} and Kobayashi's obligations/capabilities, we can obtain progress in for session types.
+Vieira and Vasconcelos \cite{VieiraV13} used single priorities and an abstract partial order to guarantee deadlock freedom in session-typed $\pi$-calculus.
+
+\paragraph{On Curry-Howard correspondences}
+The Curry-Howard correspondence between logic and types laid the foundation for functional programming \cite{Wadler15}.
+With the rise of linear logic \cite{Girard87} a Curry-Howard correspondence was established between linear logic and the linear $\pi$-calculus \cite{Abramsky94,BellinS94}.
+With the rise of session types, a new correspondence between linear logic and the $\pi$-calculus was proposed first by Caires and Pfenning \cite{CP10}, for dual intuitionistic linear logic (DILL), and later by Wadler \cite{wadler2012} for classical linear logic (CLL).
+The underlying language in both correspondences satisfy deadlock freedom by design, by forbidding cyclic connections via typing.
+Dardha and Gay follow Padovani's work \cite{P14} and use priorities in CP to define a new calculus which is deadlock-free by design, based on classical linear logic \cite{Girard87}  and the Curry-Howard correspondence with session types \cite{wadler2012}.
+Dardha and P\'{e}rez \cite{DardhaP15} compare Kobayashi-style typing and CLL typing, and prove that CLL corresponds to Kobayashi's system with the restriction that only single cuts are allowed.
+Balzer \emph{et al.} study sharing \cite{BalzerP17} and later on deadlock freedom \cite{BalzerTP19} in session-typed $\pi$-calculus based on DILL.
+In \cite{BalzerP17} types are either linear or shared with modal operators connecting them, which allows sharing, but not deadlock freedom.
+Later on in \cite{BalzerTP19} types are enhanced with information regarding the order of acquiring resources, which guarantees deadlock freedom.
+
+\paragraph{Conclusion}
+In this paper...
 \end{document}

### 2020-10-08 conclusion.tex
some conclusions

@@
 
-\paragraph{Conclusion}
-In this paper...
+\paragraph{Conclusion and Future Work}
+In this paper we presented Priority GV, which is a functional language with session types.
+We used priorities in session types, following Priority CP \cite{dardha2018} to guarantee deadlock freedom statically via our priority-based type system.
+We proved the progress property of PGV in \Cref{lem:pgv-closed-progress-confs}.
+We also presented an updated version of PCP, where we moved from commuting conversions to structural congruence in the reduction relation, thus remaining faithful to the reduction relation of the $\pi$-calculus.
+Finally, we presented an encoding of Priority CP to Pririty GV in \Cref{fig:pcp-to-pgv}.
+For this translation we proved that it is sound with respect to typing (\Cref{{lem:pcp-to-pgv-terms-preservation}}, \Cref{thm:pcp-to-pgv-confs-preservation}) and operational semantics, given by operational correspondence soundness (\Cref{thm:pcp-to-pgv-operational-correspondence-soundness}) and completeness (\Cref{thm:pcp-to-pgv-operational-correspondence-completeness}).
+
+As future work we intend to define the inverse encoding from Priority GV to Priority CP...
 \end{document}

### 2020-10-08 prf/thm-pgv-closed-progress-confs.tex
some notes on proofs

@@
 \begin{proof}
+  \todo{Shouldn't the statement be such that the def of C comes right after stating C is in canonical form, rather than in the or branch?}
+
   \label{prf:thm-pgv-closed-progress-confs}
@@
 
-  Pick the \emph{ready} term $\tm{L}\in\{\tm{M_1},\dots,\tm{M_m},\tm{N}\}$ with the smallest priority bound. There are four cases:
+  Pick the \emph{ready} term $\tm{L}\in\{\tm{M_1},\dots,\tm{M_m},\tm{N}\}$ with the smallest priority bound. There are four cases: \todo{the below is presented as if there are only 3 cases, not 4...}
   \begin{itemize}
@@
 There must be a term $\tm{L'}\in\{\tm{M_1},\dots,\tm{M_m},\tm{N}\}$ which uses $\tm{y'}$.
+  \todo{in the same lines as DF for PCP, justify here why it is true that "there must be a term..."}
   There are two cases:

### 2020-10-09 introduction.tex
Responded to several TODOs, updated bibliography. Minor additions to introduction.

--- 2020-10-07 conclusion.tex
@@
 \begin{document}
-\section{Related Work and Conclusion}
+\section{Introduction}
+
+Things to discuss in the intro:
+
+\begin{itemize}
+\item GV is a session-typed functional language.
+\item
+  The benefits of using a functional language as opposed to a process calculus are that:
+  \begin{itemize}
+  \item 
+    it supports higher-order functions (and therefore abstraction) not usually present in process calculi; and
+  \item
+    it allows us to derive extensions of the communication fragment of the language via well-understood extensions of the functional fragment, i.e., deriving internal/external choice from sum types.
+  \end{itemize}
+\item
+  The benefits of using GV over other session-typed functional languages are that:
+  \begin{itemize}
+  \item
+    GV has strong ties to linear logic~\cite{wadler15}.
+  \end{itemize}
+\item
+  Historically, calculi in the GV family~\cite{wadler15,lindley15} have achieved deadlock freedom via a syntactic restriction, i.e., by combining channel creation and thread creation into a single operation, called fork.
+Unfortunately, this is overly restrictive, as it limits the possible communication structures to trees.
+The calculus GV is based on, on the other hand, did not satisfy deadlock freedom~\cite{GAY2019}.
+\end{itemize}
+
+\begin{itemize}
+\item talking about what is GV and its "benefits"
+\item next, talking about deadlock freedom property, the benefits and the fact that the original GVs do not satisfy DF
+\item in the meantime, in a concurrent setting based on pi-calculus there is a curry-howard correspondence of logic and session typed pi, where DF holds
+\item talk about CP here
+\item \cite{dardha2018} brought the research fwd and added priorities a la Kobayashi and Padovani to CP, producing PCP
+\item following that lead, we present PGV
+\item running example of the scheduler in PGV (hopefully)
+\item translation of PCP to PGV and what it means to have that.
+\item a contribution paragraph with the following
+  \begin{itemize}
+  \item PGV and what are the benefits
+  \item updated PCP
+  \item encoding PCP to PGV
+  \item technical results to be referred to explicitly
+  \end{itemize}
+\end{itemize}
+
 \end{document}

### 2020-10-09 conclusion.tex
Cleaned up BibTeX files. Removed 'seminal'.

@@
 
-In the standard typed $\pi$-calculus a seminal line of work is by Kobayashi and his original type-theoretic approach to deadlock-freedom \cite{K98}, where priorities are abstract tags defined over a partially ordered set.
-Later on these abstract tags were simplified to natural numbers, and pairs of obligations and capabilities were used in the type system for deadlock freedom \cite{K02,K06}.
+In the standard typed $\pi$-calculus a foundational line of work is by Kobayashi and his original type-theoretic approach to deadlock-freedom~\cite{kobayashi98}, where priorities are abstract tags defined over a partially ordered set.
+Later on these abstract tags were simplified to natural numbers, and pairs of obligations and capabilities were used in the type system for deadlock freedom~\cite{kobayashi02,kobayashi06}.
 This allowed more $\pi$-calculus processes to be typed.
-Following, Kobayashi's line of work, Padovani \cite{P13} adapted obligation/capability pairs to session types, and later on he further simplified them to a single priority for linear $\pi$-calculus \cite{P14}.
-As discussed by the author, by using the encoding of session types into linear types \cite{Koba07,DGS12,D14,Dardha16}, the deadlock freedom technique of the linear $\pi$-calculus with priorities can be transferred onto the $\pi$-calculus with session types.
+Following, Kobayashi's line of work, Padovani~\cite{padovani13} adapted obligation/capability pairs to session types, and later on he further simplified them to a single priority for linear $\pi$-calculus~\cite{padovani14}.
+As discussed by the author, by using the encoding of session types into linear types~\cite{kobayashi07,dardhagiachino12,dardha14beat,dardha16}, the deadlock freedom technique of the linear $\pi$-calculus with priorities can be transferred onto the $\pi$-calculus with session types.
 
-The seminal work on progress for session-typed $\pi$-calculus, by Dezani \emph{et al.} \cite{DMYD06}, guarantees progress by allowing only one active session at a time.
-In \cite{DLY07} the authors introduce a partial order on channels in line with Kobayashi's work \cite{K98}.
-Carbone and Debois \cite{CD10} define progress for session typed $\pi$-calculus in terms of a \emph{catalyser} used to provide a missing counterpart to a process, thus guaranteeing progress.
-Carbone \emph{et al.} \cite{CDM14} studied further the use of catalysers and showed that progress is a compositional form of lock-freedom.
-As in \cite{P14} the authors show that by using the encoding of session types \cite{DGS12} and Kobayashi's obligations/capabilities, we can obtain progress in for session types.
-Vieira and Vasconcelos \cite{VieiraV13} used single priorities and an abstract partial order to guarantee deadlock freedom in session-typed $\pi$-calculus.
+The groundbreaking work on progress for session-typed $\pi$-calculus, by Dezani \emph{et al.}~\cite{dezani-ciancaglinimostrous06}, guarantees progress by allowing only one active session at a time.
+In~\cite{dezani-ciancagliniliguoro09progress} the authors introduce a partial order on channels in line with Kobayashi's work~\cite{kobayashi98}.
+Carbone and Debois~\cite{carbonedebois10} define progress for session typed $\pi$-calculus in terms of a \emph{catalyser} used to provide a missing counterpart to a process, thus guaranteeing progress.
+Carbone \emph{et al.}~\cite{carbonedardha14} studied further the use of catalysers and showed that progress is a compositional form of lock-freedom.
+As in~\cite{padovani14} the authors show that by using the encoding of session types~\cite{dardhagiachino12} and Kobayashi's obligations/capabilities, we can obtain progress in for session types.
+Vieira and Vasconcelos~\cite{vieiravasconcelos13} used single priorities and an abstract partial order to guarantee deadlock freedom in session-typed $\pi$-calculus.
 
 \paragraph{On Curry-Howard correspondences}
-The Curry-Howard correspondence between logic and types laid the foundation for functional programming \cite{Wadler15}.
-With the rise of linear logic \cite{Girard87} a Curry-Howard correspondence was established between linear logic and the linear $\pi$-calculus \cite{Abramsky94,BellinS94}.
-With the rise of session types, a new correspondence between linear logic and the $\pi$-calculus was proposed first by Caires and Pfenning \cite{CP10}, for dual intuitionistic linear logic (DILL), and later by Wadler \cite{wadler2012} for classical linear logic (CLL).
+The Curry-Howard correspondence between logic and types laid the foundation for functional programming~\cite{wadler12}.
+With the rise of linear logic~\cite{girard87} a Curry-Howard correspondence was established between linear logic and the linear $\pi$-calculus~\cite{abramsky94,bellinscott94}.
+With the rise of session types, a new correspondence between linear logic and the $\pi$-calculus was proposed first by Caires and Pfenning~\cite{cairespfenning10}, for dual intuitionistic linear logic (DILL), and later by Wadler~\cite{wadler12} for classical linear logic (CLL).
 The underlying language in both correspondences satisfy deadlock freedom by design, by forbidding cyclic connections via typing.
-Dardha and Gay follow Padovani's work \cite{P14} and use priorities in CP to define a new calculus which is deadlock-free by design, based on classical linear logic \cite{Girard87}  and the Curry-Howard correspondence with session types \cite{wadler2012}.
-Dardha and P\'{e}rez \cite{DardhaP15} compare Kobayashi-style typing and CLL typing, and prove that CLL corresponds to Kobayashi's system with the restriction that only single cuts are allowed.
-Balzer \emph{et al.} study sharing \cite{BalzerP17} and later on deadlock freedom \cite{BalzerTP19} in session-typed $\pi$-calculus based on DILL.
-In \cite{BalzerP17} types are either linear or shared with modal operators connecting them, which allows sharing, but not deadlock freedom.
-Later on in \cite{BalzerTP19} types are enhanced with information regarding the order of acquiring resources, which guarantees deadlock freedom.
+Dardha and Gay follow Padovani's work~\cite{padovani14} and use priorities in CP to define a new calculus which is deadlock-free by design, based on classical linear logic~\cite{girard87}  and the Curry-Howard correspondence with session types~\cite{wadler12}.
+Dardha and P\'{e}rez~\cite{dardhaperez15} compare Kobayashi-style typing and CLL typing, and prove that CLL corresponds to Kobayashi's system with the restriction that only single cuts are allowed.
+Balzer \emph{et al.} study sharing~\cite{balzerpfenning17} and later on deadlock freedom~\cite{balzertoninho19} in session-typed $\pi$-calculus based on DILL.
+In~\cite{balzerpfenning17} types are either linear or shared with modal operators connecting them, which allows sharing, but not deadlock freedom.
+Later on in~\cite{balzertoninho19} types are enhanced with information regarding the order of acquiring resources, which guarantees deadlock freedom.
 
@@
 In this paper we presented Priority GV, which is a functional language with session types.
-We used priorities in session types, following Priority CP \cite{dardha2018} to guarantee deadlock freedom statically via our priority-based type system.
+We used priorities in session types, following Priority CP~\cite{dardhagay18} to guarantee deadlock freedom statically via our priority-based type system.
 We proved the progress property of PGV in \Cref{lem:pgv-closed-progress-confs}.

### 2020-10-09 introduction.tex
Cleaned up BibTeX files. Removed 'seminal'.

@@
   \item
-    GV has strong ties to linear logic~\cite{wadler15}.
+    GV has strong ties to linear logic~\cite{wadler12}.
   \end{itemize}
 \item
-  Historically, calculi in the GV family~\cite{wadler15,lindley15} have achieved deadlock freedom via a syntactic restriction, i.e., by combining channel creation and thread creation into a single operation, called fork.
+  Historically, calculi in the GV family~\cite{wadler12,lindleymorris15} have achieved deadlock freedom via a syntactic restriction, i.e., by combining channel creation and thread creation into a single operation, called fork.
 Unfortunately, this is overly restrictive, as it limits the possible communication structures to trees.
-The calculus GV is based on, on the other hand, did not satisfy deadlock freedom~\cite{GAY2019}.
+The calculus GV is based on, on the other hand, did not satisfy deadlock freedom~\cite{gayvasconcelos10}.
 \end{itemize}
@@
 \item talk about CP here
-\item \cite{dardha2018} brought the research fwd and added priorities a la Kobayashi and Padovani to CP, producing PCP
+\item \cite{dardhagay18} brought the research fwd and added priorities a la Kobayashi and Padovani to CP, producing PCP
 \item following that lead, we present PGV

### 2020-10-09 priority-cp.tex
Responded to several TODOs, updated bibliography. Minor additions to introduction.
Cleaned up BibTeX files. Removed 'seminal'.

@@
   \vspace{-4mm}
-  \caption{Cyclic Scheduler from \cite{dardha2018}}
+  \caption{Cyclic Scheduler from \cite{dardhagay18}}
   \label{fig:scheduler}
@@
 
-We present an updated version of Priority CP~\cite[PCP]{dardha2018}, which is Wadler's Classical Processes~\cite[CP]{wadler2012} with \emph{priorities}.
-The main difference with respect to PCP by Dardha and Gay \cite{dardha2018} is moving away from a reduction strategy based on cycle elimination using commuting conversions, and towards a reduction strategy using structural congruence.
+We present an updated version of Priority CP~\cite[PCP]{dardhagay18}, which is Wadler's Classical Processes~\cite[CP]{wadler12} with \emph{priorities}.
+The main difference with respect to PCP by Dardha and Gay \cite{dardhagay18} is moving away from a reduction strategy based on cycle elimination using commuting conversions, and towards a reduction strategy using structural congruence.
 
@@
 The types $\ty{\typlus[\cs{o}]{A}{B}}$ and $\ty{\tywith[\cs{o}]{A}{B}}$ type the endpoints of a channel over which we can receive or send a choice between two branches $\ty{A}$ or $\ty{B}$.
-We have opted for a simplified version of choice and followed the original Wadler's CP \cite{wadler2012}, however types $\oplus$ and $\with$ can be trivially generalised to $\ty{\oplus^{\cs{o}}\{l_i:A_i\}_{i\in I}}$ and $\ty{\with^{\cs{o}}\{l_i:A_i\}_{i\in I}}$, respectively as in the original PCP \cite{dardha2018}.
+We have opted for a simplified version of choice and followed the original Wadler's CP \cite{wadler12}, however types $\oplus$ and $\with$ can be trivially generalised to $\ty{\oplus^{\cs{o}}\{l_i:A_i\}_{i\in I}}$ and $\ty{\with^{\cs{o}}\{l_i:A_i\}_{i\in I}}$, respectively as in the original PCP \cite{dardhagay18}.
 The types $\ty{\tynil[\cs{o}]}$ and $\ty{\tytop[\cs{o}]}$ type the endpoints of a channel over which we can send or receive a choice between \emph{no options}.
@@
 Typing rules make use of priorities, defined in the previous section.
-Priorities are based on \emph{obligations/capabilities} used by Kobayashi \cite{kobayashi2006}, and simplified to single priorities following Padovani \cite{padovani2014}.
-Recalling PCP \cite{dardha2018}, priorities obey the following two laws:
+Priorities are based on \emph{obligations/capabilities} used by Kobayashi~\cite{kobayashi06}, and simplified to single priorities following Padovani~\cite{padovani14}.
+Recalling PCP \cite{dardhagay18}, priorities obey the following two laws:
 \begin{enumerate}
@@
   % 
-  We will illustrate PCP with the cyclic scheduler \cite{milner99}. The full typing for $\tm {Sched}$ is given in \cite{dardha2018}.
+  We will illustrate PCP with the cyclic scheduler~\cite{milner89}. The full typing for $\tm {Sched}$ is given in \cite{dardhagay18}.
 
@@
 
-The reduction relation is given by a set of axioms and inference rules for context closure. Reduction occurs under restriction following \cite{dardha2018} where channel endpoints are identified.
+The reduction relation is given by a set of axioms and inference rules for context closure. Reduction occurs under restriction following \cite{dardhagay18} where channel endpoints are identified.
 $\LabTirName{E-Link}$ reduces a parallel composition with a link into a substitution.
@@
 \begin{proof}
-  See~\cite{dardha2018}.
+  See~\cite{dardhagay18}.
 \end{proof}
@@
 \begin{proof}
-  See~\cite{dardha2018}. \todo{SR in \cite{dardha2018} uses commuting conversions. I guess here the result follows as we are simply considering a subset of reductions?}
+  See~\cite{dardhagay18}.
 \end{proof}
@@
   A~process is an action if it acts on some endpoint $\tm{x}$.
-  \todo{Can't we simplify this def and say that every process is an action except restriction, par and zero.}
 \end{definition}

### 2020-10-09 prf/thm-pgv-closed-progress-confs.tex
Updated closed progress.

--- 2020-09-28 prf/thm-pgv-closed-progress-confs.tex
@@
   Otherwise, each term $\tm{M_i}$ is ready, and $\tm{N}$ is either ready or a value.
-
   Pick the \emph{ready} term $\tm{L}\in\{\tm{M_1},\dots,\tm{M_m},\tm{N}\}$ with the smallest priority bound. There are four cases:
-  \begin{itemize}
+  \begin{enumerate}
   \item
@@
     If $\tm{L}$ is a link $\tm{\plug{E}{\link\;\pair{y}{z}}}$ or $\tm{\plug{E}{\link\;\pair{z}{y}}}$, we apply \LabTirName{E-Link}.
-  \end{itemize}
+  \item
     Otherwise, $\tm{L}$ is ready to act on some endpoint $\tmty{y}{S}$.
 Let $\tmty{y'}{\co{S}}$ be the dual endpoint of $\tm{y}$.
-There must be a term $\tm{L'}\in\{\tm{M_1},\dots,\tm{M_m},\tm{N}\}$ which uses $\tm{y'}$.
+The typing rules enforce the linear use of endpoints, so there must be a term $\tm{L'}\in\{\tm{M_1},\dots,\tm{M_m},\tm{N}\}$ which uses $\tm{y'}$.
 There are two cases:
-  \begin{itemize}
+    \begin{enumerate}
     \item
@@
 (The argument based on the priority being the smallest continues to hold, since we know that neither $\tm{L}$ nor $\tm{L'}$ will be picked, and no other term uses $\tm{y}$ or $\tm{y'}$.)
-  \end{itemize} 
+    \end{enumerate} 
+  \end{enumerate} 
 \end{proof}

### 2020-10-09 introduction.tex
Updated lists for intro.

@@
 \item
-  The benefits of using GV over other session-typed functional languages are that:
-  \begin{itemize}
+  The benefits of using GV over other session-typed functional languages are that GV has strong ties to linear logic, via its relation to CP~\cite{wadler12}, and consequently it has strong formal properties, e.g., deadlock freedom.
 \item
-    GV has strong ties to linear logic~\cite{wadler12}.
-  \end{itemize}
-\item
-  Historically, calculi in the GV family~\cite{wadler12,lindleymorris15} have achieved deadlock freedom via a syntactic restriction, i.e., by combining channel creation and thread creation into a single operation, called fork.
+  Historically, calculi in the GV family~\cite{wadler12,lindleymorris15} have achieved deadlock freedom via a syntactic restriction, i.e., by combining channel creation and thread creation into a single operation, called fork, corresponding to the logical cut in CP, itself the combination of a name restriction and a parallel composition.
   Unfortunately, this is overly restrictive, as it limits the possible communication structures to trees.
 The calculus GV is based on, on the other hand, did not satisfy deadlock freedom~\cite{gayvasconcelos10}.
+\item
+  Recent developments in CP have led to various approaches to decoupling these constructs, either while maintaining the strong correspondence to logic, as in Hypersequent CP~\cite[HCP]{kokkemontesi19popl,kokkemontesi19tlla}, or by weakening the correspondence to logic in exchange for a more expressive language, as in Priority CP~\cite[PCP]{dardhagay18}.
+\item
+  PCP decouples CP's cut into separate constructs for name restriction and parallel composition, and restores deadlock freedom by adding priorities \`{a} la~\cite{kobayashi06}.
+\item
+  Following the example set by PCP, we present Priority GV (PGV), a~variant of GV which decouples channel creation from thread creation, and restores deadlock freedom by adding priorities.
+\item
+  We demonstrate the correspondence between PGV and PCP by proving operational correspondence for a translation from PCP to PGV.
 \end{itemize}
 
+\paragraph*{Contributions}
 \begin{itemize}
-\item talking about what is GV and its "benefits"
-\item next, talking about deadlock freedom property, the benefits and the fact that the original GVs do not satisfy DF
-\item in the meantime, in a concurrent setting based on pi-calculus there is a curry-howard correspondence of logic and session typed pi, where DF holds
-\item talk about CP here
-\item \cite{dardhagay18} brought the research fwd and added priorities a la Kobayashi and Padovani to CP, producing PCP
-\item following that lead, we present PGV
-\item running example of the scheduler in PGV (hopefully)
-\item translation of PCP to PGV and what it means to have that.
-\item a contribution paragraph with the following
-  \begin{itemize}
-  \item PGV and what are the benefits
-  \item updated PCP
-  \item encoding PCP to PGV
-  \item technical results to be referred to explicitly
-  \end{itemize}
+\item PGV (typing rules, operational semantics, subject reduction, closed progress).
+\item An updated version of PCP, in which we remove the commuting conversions, moving away from reduction as cut elimination, and towards reduction as one would expect for a process calculus.
+\item A~translation from PCP into PGV (type preservation, operational correspondence).
 \end{itemize}

### 2020-10-11 conclusion.tex
Messing around with schedulers.

@@
 \paragraph{On Curry-Howard correspondences}
-The Curry-Howard correspondence between logic and types laid the foundation for functional programming~\cite{wadler12}.
+The Curry-Howard correspondence between logic and types laid the foundation for functional programming~\cite{wadler15}.
 With the rise of linear logic~\cite{girard87} a Curry-Howard correspondence was established between linear logic and the linear $\pi$-calculus~\cite{abramsky94,bellinscott94}.
-With the rise of session types, a new correspondence between linear logic and the $\pi$-calculus was proposed first by Caires and Pfenning~\cite{cairespfenning10}, for dual intuitionistic linear logic (DILL), and later by Wadler~\cite{wadler12} for classical linear logic (CLL).
+With the rise of session types, a new correspondence between linear logic and the $\pi$-calculus was proposed first by Caires and Pfenning~\cite{cairespfenning10}, for dual intuitionistic linear logic (DILL), and later by Wadler~\cite{wadler15} for classical linear logic (CLL).
 The underlying language in both correspondences satisfy deadlock freedom by design, by forbidding cyclic connections via typing.
-Dardha and Gay follow Padovani's work~\cite{padovani14} and use priorities in CP to define a new calculus which is deadlock-free by design, based on classical linear logic~\cite{girard87}  and the Curry-Howard correspondence with session types~\cite{wadler12}.
+Dardha and Gay follow Padovani's work~\cite{padovani14} and use priorities in CP to define a new calculus which is deadlock-free by design, based on classical linear logic~\cite{girard87}  and the Curry-Howard correspondence with session types~\cite{wadler15}.
 Dardha and P\'{e}rez~\cite{dardhaperez15} compare Kobayashi-style typing and CLL typing, and prove that CLL corresponds to Kobayashi's system with the restriction that only single cuts are allowed.

### 2020-10-11 introduction.tex
Messing around with schedulers.

@@
 
-Things to discuss in the intro:
+We introduce Priority GV (PGV), a~session-typed concurrent $\lambda$-calculus.
 
 \begin{itemize}
-\item GV is a session-typed functional language.
+\item
+  Historically, calculi in the GV family~\cite{wadler15,lindleymorris15} have achieved deadlock freedom via a syntactic restriction, i.e., by combining channel creation and thread spawning into a single operation, called fork, corresponding to the logical cut in CP, itself the combination of a name restriction and a parallel composition.
+  Unfortunately, this is overly restrictive, as it limits the possible communication structures to trees.
+The calculus GV is based on, on the other hand, did not satisfy deadlock freedom~\cite{gayvasconcelos12}.
+\item
+  Recent developments in CP have led to various approaches to decoupling these constructs, either while maintaining the strong correspondence to logic, as in Hypersequent CP~\cite[HCP]{kokkemontesi19popl,kokkemontesi19tlla}, or by weakening the correspondence to logic in exchange for a more expressive language, as in Priority CP~\cite[PCP]{dardhagay18}.
+\item
+  PCP decouples CP's cut into separate constructs for name restriction and parallel composition, and restores deadlock freedom by adding priorities \`{a} la~\cite{kobayashi06}.
+\item
+  Following the example set by PCP, we present Priority GV (PGV), a~variant of GV which decouples channel creation from thread creation, and restores deadlock freedom by adding priorities.
+\item
+  We demonstrate the correspondence between PGV and PCP by proving operational correspondence for a translation from PCP to PGV.
 \item
@@
 \item
-  The benefits of using GV over other session-typed functional languages are that GV has strong ties to linear logic, via its relation to CP~\cite{wadler12}, and consequently it has strong formal properties, e.g., deadlock freedom.
-\item
-  Historically, calculi in the GV family~\cite{wadler12,lindleymorris15} have achieved deadlock freedom via a syntactic restriction, i.e., by combining channel creation and thread creation into a single operation, called fork, corresponding to the logical cut in CP, itself the combination of a name restriction and a parallel composition.
-  Unfortunately, this is overly restrictive, as it limits the possible communication structures to trees.
-The calculus GV is based on, on the other hand, did not satisfy deadlock freedom~\cite{gayvasconcelos10}.
-\item
-  Recent developments in CP have led to various approaches to decoupling these constructs, either while maintaining the strong correspondence to logic, as in Hypersequent CP~\cite[HCP]{kokkemontesi19popl,kokkemontesi19tlla}, or by weakening the correspondence to logic in exchange for a more expressive language, as in Priority CP~\cite[PCP]{dardhagay18}.
-\item
-  PCP decouples CP's cut into separate constructs for name restriction and parallel composition, and restores deadlock freedom by adding priorities \`{a} la~\cite{kobayashi06}.
-\item
-  Following the example set by PCP, we present Priority GV (PGV), a~variant of GV which decouples channel creation from thread creation, and restores deadlock freedom by adding priorities.
-\item
-  We demonstrate the correspondence between PGV and PCP by proving operational correspondence for a translation from PCP to PGV.
+  The benefits of using GV over other session-typed functional languages are that GV has strong ties to linear logic, via its relation to CP~\cite{wadler15}, and consequently it has strong formal properties, e.g., deadlock freedom.
 \end{itemize}

### 2020-10-11 priority-cp.tex
Messing around with schedulers.

@@
 Process $\tm{\link{x}{y}}$ links channels $x$ and $y$ and forwards communication from one to the other.
-$\tm{\res{x}{y}{P}}$, $\tm{(\ppar{P}{Q})}$ and $\tm{\halt}$ denote respectively the restriction processes where channel endpoints $\tm x$ and $\tm y$ are bound together and with scope $\tm P$, the parallel composition of processes $\tm{P}$ and $\tm{Q}$ and the terminated process.
+$\tm{\res{x}{y}{P}}$, $\tm{(\ppar{P}{Q})}$ and $\tm{\halt}$ denote respectively the restriction processes where channel endpoints $\tm{x}$ and $\tm{y}$ are bound together and with scope $\tm{P}$, the parallel composition of processes $\tm{P}$ and $\tm{Q}$ and the terminated process.
 
-Processes $\tm{\send{x}{y}{P}}$ and $\tm{\recv{x}{y}{P}}$ send or receive over channel $\tm x$ a value $\tm y$ and proceed as process $\tm P$.
-$\tm{\close{x}{P}}$ and $\tm{\wait{x}{P}}$ send and receive an empty value -- denoting the closure of channel $\tm x$, and continue as $\tm P$.
+Processes $\tm{\send{x}{y}{P}}$ and $\tm{\recv{x}{y}{P}}$ send or receive over channel $\tm{x}$ a value $\tm{y}$ and proceed as process $\tm{P}$.
+$\tm{\close{x}{P}}$ and $\tm{\wait{x}{P}}$ send and receive an empty value -- denoting the closure of channel $\tm{x}$, and continue as $\tm{P}$.
 
-Processes $\tm{\inl{x}{P}}$ and $\tm{\inr{x}{P}}$ make a left and right choice, respectively and proceed as process $\tm P$.
-$\tm{\offer{x}{P}{Q}}$ dually offers both left and right branches, with continuations $\tm P$ and $\tm Q$.
+Processes $\tm{\inl{x}{P}}$ and $\tm{\inr{x}{P}}$ make a left and right choice, respectively and proceed as process $\tm{P}$.
+$\tm{\offer{x}{P}{Q}}$ dually offers both left and right branches, with continuations $\tm{P}$ and $\tm{Q}$.
 $\tm{\absurd{x}}$ denotes the empty offer.
 
-Finally, we define \emph{unbound} send as syntactic sugar:
-\[
-  \tm{\usend{x}{y}{P}}\elabarrow\tm{\send{x}{z}{(\ppar{\link{y}{z}}{P})}}
-\]
+We write \emph{unbound} send as $\tm{\usend{x}{y}{P}}$, which is syntactic sugar for $\tm{\send{x}{z}{(\ppar{\link{y}{z}}{P})}}$.
+Alternatively, we could take $\tm{\usend{x}{y}{P}}$ as primitive, and let $\tm{\send{x}{y}{P}}$ be syntactic sugar for $\tm{\res{y}{z}{(\usend{x}{z}{P})}}$.
+CP takes \emph{bound} sending as primitive, as it is impossible to eliminate the top-level cut in terms such as $\tm{\res{y}{z}{(\usend{x}{z}{P})}}$, even with communting conversions.
+However, in our setting without commuting conversions and with more permissive normal forms, this is no longer an issue.
 
@@
 
-Typing rules for our version of PCP are given in \Cref{fig:pcp-typing}. A typing judgement $\seq{P}{\Gamma}$ states that "process $\tm P$ is well typed under the typing context $\ty \Gamma$".
+Typing rules for our version of PCP are given in \Cref{fig:pcp-typing}. A typing judgement $\seq{P}{\Gamma}$ states that "process $\tm{P}$ is well typed under the typing context $\ty \Gamma$".
 
@@
 \end{enumerate}
-\textsc{T-Link} states that the link process $\tm{\link{x}{y}}$ is well typed under channels $\tm x$ and $\tm y$ having dual types, respectively $\ty A$ and $\ty {\co A}$.
-\textsc{T-Res} sates that the restriction process $\tm{\res{x}{y}{P}}$ is well typed under typing context $\ty\Gamma$ if process $\tm P$ is well typed in $\ty\Gamma$ augmented with channel endpoints $\tm x$ and $\tm y$ having dual types, respectively $\ty A$ and $\ty {\co A}$.
-\textsc{T-Par} states that the parallel composition of processes $\tm P$ and $\tm Q$ is well typed under the disjoint union of their respective typing contexts.
+\textsc{T-Link} states that the link process $\tm{\link{x}{y}}$ is well typed under channels $\tm{x}$ and $\tm{y}$ having dual types, respectively $\ty A$ and $\ty {\co A}$.
+\textsc{T-Res} sates that the restriction process $\tm{\res{x}{y}{P}}$ is well typed under typing context $\ty\Gamma$ if process $\tm{P}$ is well typed in $\ty\Gamma$ augmented with channel endpoints $\tm{x}$ and $\tm{y}$ having dual types, respectively $\ty A$ and $\ty {\co A}$.
+\textsc{T-Par} states that the parallel composition of processes $\tm{P}$ and $\tm{Q}$ is well typed under the disjoint union of their respective typing contexts.
 \textsc{T-Halt} states that the terminated process $\tm{\halt}$ is well typed in the empty context.
 
-\textsc{T-Send} and \textsc{T-Recv} state that the sending and receiving of a bound name $\tm y$ over a channel $\tm x$ is well typed under $\ty\Gamma$ and $\tm x$ of type $\tytens[\cs{o}]{A}{B}$, respectively ${\typarr[\cs{o}]{A}{B}}$.
+\textsc{T-Send} and \textsc{T-Recv} state that the sending and receiving of a bound name $\tm{y}$ over a channel $\tm{x}$ is well typed under $\ty\Gamma$ and $\tm{x}$ of type $\tytens[\cs{o}]{A}{B}$, respectively ${\typarr[\cs{o}]{A}{B}}$.
 Priority $\cs o$ is the smallest among all priorities of the types used by the output or input process, captured by the side condition $\cs{o}<\minpr(\ty{\Gamma},\ty{A},\ty{B})$.
 
-Rules \textsc{T-Close} and \textsc{T-Wait} type the closure of channel $\tm x$ and are in the same lines as the previous two rules, requiring that the priority of channel $\tm x$ is the smallest among all priorities in $\tm\Gamma$.
+Rules \textsc{T-Close} and \textsc{T-Wait} type the closure of channel $\tm{x}$ and are in the same lines as the previous two rules, requiring that the priority of channel $\tm{x}$ is the smallest among all priorities in $\tm\Gamma$.
 
-\textsc{T-Select-Inl} and \textsc{T-Select-Inr} type respectively the left $\tm{\inl{x}{P}}$ and right $\tm{\inr{x}{P}}$ choice performed on channel $\tm x$.
-\textsc{T-Offer} and \textsc{T-Offer-Absurd} type the offering of a choice, or empty choice, on channel $\tm x$.
-In all the above rules the priority $\cs o$ of channel $\tm x$ is the smallest with respect to the typing context $\cs{o}<\minpr(\ty{\Gamma})$ and types involved in the choice $\cs{o}<\minpr(\ty{\Gamma},\ty{A},\ty{B})$.
+\textsc{T-Select-Inl} and \textsc{T-Select-Inr} type respectively the left $\tm{\inl{x}{P}}$ and right $\tm{\inr{x}{P}}$ choice performed on channel $\tm{x}$.
+\textsc{T-Offer} and \textsc{T-Offer-Absurd} type the offering of a choice, or empty choice, on channel $\tm{x}$.
+In all the above rules the priority $\cs o$ of channel $\tm{x}$ is the smallest with respect to the typing context $\cs{o}<\minpr(\ty{\Gamma})$ and types involved in the choice $\cs{o}<\minpr(\ty{\Gamma},\ty{A},\ty{B})$.
 
-\begin{example}[Milner's Cyclic Scheduler \Cref{fig:scheduler}]
-  % 
-  We will illustrate PCP with the cyclic scheduler~\cite{milner89}. The full typing for $\tm {Sched}$ is given in \cite{dardhagay18}.
+\subsection{Example: Milner's Cyclic Scheduler}
+We illustrate PCP with the cyclic scheduler~\cite{milner89}. The full typing for $\tm{\sched}$ is given in \cite{dardhagay18}.
 
-  We will consider a simplified version: a set of 3 agents $\tm {A_0}, \tm {A_1}, \tm {A_2}$ is scheduled to perform a certain task in cyclic order, starting with agent $\tm {A_0}$.
-Agents $\tm {A_1}$ and $\tm {A_2}$ will send the result of some computation a collector process, respectively $\tm {P_1}$ and $\tm {P_2}$, before transmitting data further to the next agent in the cycle, being $\tm {A_{(i+1)\texttt{ mod } 3}}$.
-At the end of the round, agent $\tm {A_0}$ will send the final result of the computation to its collector $\tm {P_0}$.
+We consider a simplified version: a set of 3 agents $\tm{\agent0}, \tm{\agent1}, \tm{\agent2}$ is scheduled to perform a certain task in cyclic order, starting with $\tm{\agent0}$.
+$\tm{\agent1}$ and $\tm{\agent2}$ send the result of some computation a collector process, respectively $\tm{\collector1}$ and $\tm{\collector2}$, before transmitting data further to the next agent in the cycle, being $\tm{\agent{(i+1)\bmod 3}}$.
+At the end of the round, $\tm{\agent0}$ send the final result of the computation to its collector $\tm{\collector0}$.
 
@@
 \begin{array}{rcll}
-      \tm {Sched} & \defeq & ..\tm {(\nu {a_i}{b_i})}..\tm {(\nu {c_i}{d_{{(i+1)\texttt{ mod } n}}})} \tm {\big(A_0 \parallel A_1 \parallel A_2 \parallel P_0 \parallel P_1 \parallel P_2 \big)}\\
-      \tm{A_0} & \defeq & \tm{\send{c_0}{n_0}{\recv{d_0}{x_0}{\send{a_0}{m_0}{close_0}}}}\\
-      \tm{A_i} & \defeq & \tm{\recv{d_i}{x_i}{\send{a_i}{m_i}{\send{c_i}{n_i}{close_i}}}} &~~ i \in \{1,2\}\\
-      \tm {P_i} & \defeq & \tm{\recv{b_i}{y_i}{Q_i}} &~~ i \in \{0,1,2\}
+  \tm{\sched}
+  & \defeq
+  & \tm{
+    \res{x_0}{x'_0}{\res{x_1}{x'_1}{\res{x_2}{x'_2}{
+    \res{\overrightarrow{y_0}}{\overleftarrow{y_1}}{
+    \res{\overrightarrow{y_1}}{\overleftarrow{y_2}}{
+    \res{\overrightarrow{y_2}}{\overleftarrow{y_0}}{}}}}}}
+    }\\
+  &
+  & \tm{(\ppar{\agent0}{\ppar{\agent1}{\ppar{\agent2}{\ppar{\collector0}{\ppar{\collector1}{\collector2}}}}})}\\
+  \tm{\agent0}
+  & \defeq
+  & \tm{\usend{\overrightarrow{y_0}}{d_0}{\recv{\overleftarrow{y_0}}{d'_0}{\usend{x_0}{d'_0}{
+    \wait{\overleftarrow{y_0}}{\close{\overrightarrow{y_0}}{\close{x_0}{\halt}}}}}}} \\
+  \tm{\agent{i}}
+  & \defeq
+  & \tm{\recv{\overleftarrow{y_i}}{d_i}{\usend{x_i}{d_i}{\recv{x_i}{d'_i}{\usend{\overrightarrow{y_i}}{d'_i}{
+    \wait{\overleftarrow{y_i}}{\close{\overrightarrow{y_i}}{\close{x_i}{\halt}}}}}}}}
+  & i \in \{1,2\} \\
+  \tm{\collector{i}}
+  & \defeq & \tm{\recv{x'_i}{d_i}{Q_i}}
+  & i \in \{0,1,2\}
 \end{array}
 $$
-
-  In the above process, channels $n_i$ and $m_i$ for all $i$, denote data being transmitted, and for simplicity we let them be channels of a terminated type.
-Process $\tm{close_i}$ for all $i$ closes all channels in $\tm{A_i}$.
-In particular, they are a sequence of $\tm{\close{x}{P}}$ and $\tm{\wait{x}{P}}$ for all channels in the agent and terminating in $\tm{\halt}$.
-\end{example}
-
+In the above process, channels $\tm{d_i}$ and $\tm{d'_i}$ represent the data being transmitted.
+For simplicity, let them be channels of a terminated type.
+The processes $\tm{Q_i}$ represent the inner workings of the collector processes, each of which must eventually send back a response over $\tm{x'_i}$ and close the channel.
 \subsection{Operational Semantics}
@@
 $\LabTirName{E-Send}$ is the main communication rule, where send and receive processes sychronise and reduce to the corresponding continuations.
-$\LabTirName{E-Close}$ follows the previous rule and it closes the channel identified by endpoints $\tm x$ and $\tm y$.
+$\LabTirName{E-Close}$ follows the previous rule and it closes the channel identified by endpoints $\tm{x}$ and $\tm{y}$.
 $\LabTirName{E-Select-Inl}$ and $\LabTirName{E-Select-Inr}$ are generalised versions of $\LabTirName{E-Send}$.

### 2020-10-11 priority-gv.tex
Messing around with schedulers.

--- 2020-09-28 priority-gv.tex
@@
 
-\subsection{Syntax}
+This section presents Priority GV (PGV), a session-typed functional language based on GV~\cite{wadler15,lindleymorris15}, but with a twist: it enforces deadlock freedom through priorities \`{a} la Kobayashi~\cite{kobayashi06,padovaninovara15}.
 
-\paragraph*{Priorities, types, and environments}
-Types are divided into \emph{session types} $\ty{S}$ and \emph{value types} $\ty{T}$, $\ty{U}$.
+GV enforces deadlock freedom with a syntactic restriction: it fuses channel creation and thread spawning into a single function, called fork.
+This guarantees that if some term $\tm{L}$ has the means to communicate with both $\tm{M}$ and $\tm{N}$, then $\tm{M}$ and $\tm{N}$ cannot communicate amongst themselves.
+In other words, GV only allows tree-shaped communication structures, so it is trivially free from cycles, and therefore deadlocks.
+However, this is overly restrictive, as we rule out \emph{all} cyclic communication structures, not just those with deadlocks.
 
-Session types are annotated with \emph{priorities} $\cs{o}\in\mathbb{N}$.
-The types $\ty{\tysend[\cs{o}]{T}{S}}$ and $\ty{\tyrecv[\cs{o}]{T}{S}}$ type the endpoints of a channel over which we send or receive a value of type $\ty{T}$, and then proceed as $\ty{S}$.
-The types $\ty{\tyends[\cs{o}]}$ and $\ty{\tyendr[\cs{o}]}$ type endpoints of a channel whose session has finished, and over which we send or receive a ping before closing the channel.
+Priority GV offers a more fine-grained analysis of the communication structure.
+It allows cyclic communication structures,  and as a result, it is more expressive.
+To illustrate this, we use PGV to implement an instance of Milner's Cyclic Scheduler in~\todo{Add reference}.
+
+\subsubsection{Types}
+Types in PGV are divided into \emph{session types} $\ty{S}$ and \emph{types} $\ty{T}$, $\ty{U}$.
+%
+Session types are defined by the following grammar:
 \[
@@
 \]
-Value types are based on the types of the linear $\lambda$-calculus.
-The types $\ty{\typrod{T}{U}}$, $\ty{\tyunit}$, $\ty{\tysum{T}{U}}$, and $\ty{\tyvoid}$ are the standard linear $\lambda$-calculus product type, unit type, sum type, and empty type.
-The type $\ty{\tylolli[\cs{p},\cs{q}]{T}{U}}$ is the standard linear function type, annotated with \emph{priority bounds} $\cs{p},\cs{q}\in\mathbb{N}\cup\{\cs{\pbot},\cs{\ptop}\}$.
-Session types are value types, as endpoints are values.
+The session types $\ty{\tysend[\cs{o}]{T}{S}}$ and $\ty{\tyrecv[\cs{o}]{T}{S}}$ describe the endpoints of a channel over which we send or receive a value of type $\ty{T}$, and then proceed as $\ty{S}$.
+The types $\ty{\tyends[\cs{o}]}$ and $\ty{\tyendr[\cs{o}]}$ describe endpoints of a channel whose communication has finished, and over which we must synchronise before closing the channel.
+Each connective in a session type is annotated with a \emph{priority} $\cs{o}\in\mathbb{N}$.
+
+Types are defined by the following grammar:
 \[
@@
 \]
-Given a function with type $\ty{\tylolli[\cs{p},\cs{q}]{T}{U}}$, the priority bound $\cs{p}$ is a \emph{lower bound} on the priorities of the channel names captured by the body of the function, and the priority bound $\cs{q}$ is an \emph{upper bound} on the priority of the communications that take place as a result of applying the function.
-We discuss this in more detail when we give the typing rules.
-We define syntactic sugar for the type of \emph{pure functions}.
-\[
-  \ty{\tylolli{T}{U}} \elabarrow \ty{\tylolli[\cs{\ptop},\cs{\pbot}]{T}{U}}
-\]
-A typing context $\ty\Gamma$ is a partial function associating names to types.
+The types $\ty{\typrod{T}{U}}$, $\ty{\tyunit}$, $\ty{\tysum{T}{U}}$, and $\ty{\tyvoid}$ are the standard linear $\lambda$-calculus product type, unit type, sum type, and empty type.
+The type $\ty{\tylolli[\cs{p},\cs{q}]{T}{U}}$ is the standard linear function type, annotated with \emph{priority bounds} $\cs{p},\cs{q}\in\mathbb{N}\cup\{\cs{\pbot},\cs{\ptop}\}$.
+Every session type is also a type.
+
+Given a function with type $\ty{\tylolli[\cs{p},\cs{q}]{T}{U}}$, $\cs{p}$ is a \emph{lower bound} on the priorities of the endpoints captured by the body of the function, and $\cs{q}$ is an \emph{upper bound} on the priority of the communications that take place as a result of applying the function.
+The type of \emph{pure functions} $\ty{\tylolli{T}{U}}$, \ie those which perform no communications, is syntactic sugar for $\ty{\tylolli[\cs{\ptop},\cs{\pbot}]{T}{U}}$.
+
+\subsubsection{Environments}
+Typing environments $\ty{\Gamma}$, $\ty{\Delta}$ associate names to types.
+Environments are linear, so two environments can only be combined as $\ty{\Gamma}, \ty{\Delta}$ if their names are distinct, \ie $\fv(\ty{\Gamma})\cap\fv(\ty{\Delta})=\varnothing$.
 \[
@@
 \]
+
+\subsubsection{Duality}
+Duality plays a crucial role in session types.
+The two endpoints of a channel are assigned dual types, which ensures that, for instance, whenever one program \emph{sends} a value on a channel, the program on the other end is waiting to \emph{receive}.
+Each session type $\ty{S}$ has a dual, written $\ty{\co{S}}$.
+Duality is an involutive function which preserves priorities.
 \[
@@
 \]
+
+\subsubsection{Priorities}
+The function $\pr$ returns the smallest priority of a session type.
+The type system guarantees that the top-most connective always holds the smallest priority, so we simply return the priority of the top-most connective.
 \[
@@
 \]
+We extend the function $\pr$ to value types and typing contexts by returning the smallest priority in the type or context, or $\cs{\top}$ if there is no priority.
+\todo{What about the function type?}
 \[
@@
 
-\paragraph*{Static Term Syntax}
-We use $\tm{x}$, $\tm{y}$, $\tm{z}$, and $\tm{w}$ to range over variable names. Occasionally, we use $\tm{a}$, $\tm{b}$, and $\tm{c}$ to range over \emph{free} channel endpoints.
+\subsubsection{Terms}
+Terms are defined by the following grammar:
 \[
@@
     \sep        \tm{\recv}
+    \sep        \tm{\close}
     \sep        \tm{\wait}
-    \sep        \tm{\close}
 \end{array}
 \]
+Let $\tm{x}$, $\tm{y}$, $\tm{z}$, and $\tm{w}$ range over variable names.
+Occasionally, we use $\tm{a}$, $\tm{b}$, $\tm{c}$, and $\tm{d}$, which range over \emph{free} endpoints.
+The term language is the standard linear $\lambda$-calculus with products, sums, and their units, extended with constants $\tm{K}$ for the communication primitives.
 
-\paragraph*{Syntactic Sugar for Terms}
-\[
-\begin{array}{lcl}
-  \tm{\andthen{M}{N}}
-  & \elabarrow & \tm{\letunit{M}{N}} \\
-  \tm{\letbind{x}{M}{N}}
-  & \elabarrow & \tm{(\lambda x.N)\;M} \\
-  \tm{\lambda\unit.M}
-  & \elabarrow & \tm{\lambda z.\letunit{z}{M}} \\
-  \tm{\lambda\pair{x}{y}.M}
-  & \elabarrow & \tm{\lambda z.\letpair{x}{y}{z}{M}} \\
-  \tm{\fork}
-  & \elabarrow & \tm{\lambda x.\letpair{y}{z}{\new}{\andthen{\spawn\;{(\lambda\unit.x\;y)}}{z}}}
-\end{array}
-\]
+The constants are best understood in conjunction with their typing and reduction rules in~\cref{fig:pgv-typing,fig:pgv-operational-semantics}.
+Briefly, $\tm{\link}$ links two endpoints together, forwarding messages from each to the other, $\tm{\new}$ creates a new channel and returns its endpoints, and $\tm{\spawn}$ spawns off its argument as a new thread.
+The $\tm{send}$ and $\tm{recv}$ functions send and receive values on a channel.
+However, since the typing rules for PGV ensure the linear usage of endpoints, they also return a new copy of the endpoint, to continue the session.
+The $\tm{close}$ and $\tm{wait}$ functions close a channel.
 
-\paragraph*{Syntactic Sugar for Internal and External Choice}
+We frequently use syntactic sugar to make terms more readable: we write $\tm{\andthen{M}{N}}$ in place of $\tm{\letunit{M}{N}}$, $\tm{\letbind{x}{M}{N}}$ in place of $\tm{(\lambda x.N)\;M}$, and pattern matching lambdas $\tm{\lambda\unit.M}$ and $\tm{\lambda\pair{x}{y}.M}$ in place of $\tm{\lambda z.\letunit{z}{M}}$ and $\tm{\lambda z.\letpair{x}{y}{z}{M}}$.
+
+\subsubsection{Internal and External Choice}
+Typically, session-typed languages feature constructs for internal and external choice.
+In GV, these can be defined in terms of the core language, by sending or receiving a value of a sum type~\cite{lindleymorris15}.
+We use the following syntactic sugar for internal ($\ty{\tyselect[\cs{o}]{S}{S'}}$) and external ($\ty{\tyoffer[\cs{o}]{S}{S'}}$) choice and their units:
 \[
@@
 \]
+As the syntax for units suggests, these are the binary and nullary forms of the more common n-ary choice constructs $\ty{{\oplus}^{\cs{o}}\{l_i:S_i\}_{i\in{I}}}$ and $\ty{{\with}^{\cs{o}}\{l_i:S_i\}_{i\in{I}}}$, which one may obtain generalising the sum types to variant types.
+For simplicity, we present only the binary and nullary forms.
+
+Similarly, we use syntactic sugar for the term forms of choice, which combine sending and receiving with the introduction and elimination forms for the sum and empty types.
+There are two constructs for binary internal choice, expressed using the meta-variable $\tm{\ell}$ which ranges over $\{\tm{\labinl},\tm{\labinr}\}$.
+As there is no introduction for the empty type, there is no construct for nullary internal choice:
 \[
@@
 
+\subsection{Operational Semantics}
 \paragraph*{Runtime Term Syntax}
@@
 \]
+See \cref{fig:pgv-operational-semantics}.
+\input{fig/pgv-operational-semantics}
 
 \subsection{Typing}
-See \cref{fig:pgv-typing}.
+
 \input{fig/pgv-typing}
@@
 
-\subsection{Operational Semantics}
-See \cref{fig:pgv-operational-semantics}.
-\input{fig/pgv-operational-semantics}
+\subsection{Example: A~Cyclic Scheduler}
+We illustrate the newfound ability of PGV by encoding a variant of Milner's cyclic scheduler~\cite{milner89,dardhagay18}.
+Suppose three processes ($\tm{\collector0}$, $\tm{\collector1}$, and $\tm{\collector2}$) should be scheduled to perform their tasks in cyclic order, starting with $\tm{\collector1}$, and circling all the way around to $\tm{\collector0}$.
+We implement the scheduler with three agents ($\tm{\agent0}$, $\tm{\agent1}$, and $\tm{\agent2}$) which represent their respective processes in the scheduler.
+
+\begin{center}
+  \begin{tikzpicture}
+    \node[shape=circle,draw=black] (agent0) at ( 0.00, 0.00) {\tm{\agent0}};
+    \node[shape=circle,draw=black] (agent1) at ( 1.75,-1.75) {\tm{\agent1}};
+    \node[shape=circle,draw=black] (agent2) at (-1.75,-1.75) {\tm{\agent2}};
+    \node[shape=circle,draw=black] (proc0)  at ( 3.00, 0.00) {\tm{\collector0}};
+    \node[shape=circle,draw=black] (proc1)  at ( 4.50,-1.75) {\tm{\collector1}};
+    \node[shape=circle,draw=black] (proc2)  at (-4.50,-1.75) {\tm{\collector2}};
+    \node[draw=none]               (middle) at ( 0.00,-1.75) {};
+
+    \draw
+    (agent0) -- node[pos=0.35,above] {$c_0$}
+    node[pos=1.0,above] {$d_1$} ++ (agent1);
+    \draw
+    (agent1) -- node[pos=0.1,above] {$c_1$}
+    node[pos=0.9,above] {$d_2$} ++ (agent2);
+    \draw
+    (agent2) -- node[pos=0.0,above] {$c_2$}
+    node[pos=0.65,above] {$d_0$} ++ (agent0);
+
+    \draw
+    (agent0) -- node[pos=0.15,above] {$a_0$}
+    node[pos=0.85,above] {$b_0$} (proc0);
+    \draw
+    (agent1) -- node[pos=0.15,above] {$a_1$}
+    node[pos=0.85,above] {$b_1$} (proc1);
+    \draw
+    (agent2) -- node[pos=0.15,above] {$a_2$}
+    node[pos=0.85,above] {$b_2$} (proc2);
+  \end{tikzpicture}
+\end{center}
+\[
+  \begin{array}{lrlr}
+    \tm{\sched}
+    & \defeq & \tm{
+               \res{a_0}{b_0}{\res{a_1}{b_1}{\res{a_2}{b_2}{
+               \res{c_0}{d_1}{
+               \res{c_1}{d_2}{
+               \res{c_2}{d_0}{}}}}}}}
+    \\ &     & \multicolumn{2}{l}{%
+               \tm{(\ppar{\child\;\agent0}{\ppar{\child\;\agent1}{\child\;\ppar{\agent2}{
+               \ppar{\child\;\collector0}{\ppar{\child\;\collector1}{\child\;\collector2}}}}})}}
+    \\
+    \tm{\agent0}
+    & \defeq & \tm{\letbind{c_0}{\send\;\pair{\text{data}_0}{c_0}}{%
+               \letpair{\text{data}'_0}{d_0}{\recv\;d_0}{}}}
+    \\ &     & \tm{\letbind{a_0}{\send\;\pair{\text{data}_0}{a_0}}{}}
+    \\ &     & \tm{\andthen{\wait\;{d_i}}{\andthen{\close\;{c_i}}{\close\;{a_i}}}}
+    \\
+    \tm{\agent{i}}
+    & \defeq & \tm{\letpair{d_i}{d_i}{\recv\;d_i}{%
+               \letbind{a_i}{\send\;\pair{\text{data}_i}{a_i}}{}}}
+    \\ &     & \tm{\letpair{\text{data}'_i}{a_i}{\recv\;{a_i}}{%
+               \letbind{c_i}{\send\;\pair{\text{data}_i}{c_i}}{}}}
+    \\ &     & \tm{\andthen{\wait\;{d_i}}{\andthen{\close\;{c_i}}{\close\;{a_i}}}}
+       & i \in \{1,2\}
+    \\
+    \tm{\collector{i}}
+    & \defeq & \tm{\letpair{d'_i}{d_i}{\recv\;d_i}{M_i}}
+       & i \in \{0,1,2\}
+  \end{array}
+\]
 
@@
 \end{restatablelemma}
-\begin{proof}
-  \admit
-\end{proof}
 

### 2020-10-11 fig/pgv-scheduler.tex
Added scheduler. Made reasonable progress.

\begin{center}
  \begin{tikzpicture}
    \node[shape=circle,draw=black] (agent1) at ( 0.00, 0.00) {\tm{\agent1}};
    \node[shape=circle,draw=black] (agent2) at ( 1.75,-1.75) {\tm{\agent2}};
    \node[shape=circle,draw=black] (agent3) at (-1.75,-1.75) {\tm{\agent3}};
    \node[shape=circle,draw=black] (proc1)  at ( 3.00, 0.00) {\tm{\proc1}};
    \node[shape=circle,draw=black] (proc2)  at ( 4.50,-1.75) {\tm{\proc2}};
    \node[shape=circle,draw=black] (proc3)  at (-4.50,-1.75) {\tm{\proc3}};

    \draw
    (agent1) -- node[pos=0.35,above] {$c_1$}
    node[pos=1.0,above] {$c'_1$} ++ (agent2);
    \draw
    (agent2) -- node[pos=0.1,above] {$c_2$}
    node[pos=0.9,above] {$c'_2$} ++ (agent3);
    \draw
    (agent3) -- node[pos=0.0,above] {$c_3$}
    node[pos=0.65,above] {$c'_3$} ++ (agent1);

    \draw[double]
    (agent1) -- node[pos=0.15,above] {$a_1$} node[pos=0.85,above] {$a'_1$}
    node[pos=0.15,below] {$b_1$} node[pos=0.85,below] {$b'_1$} ++ (proc1);
    \draw[double]
    (agent2) -- node[pos=0.15,above] {$a_2$} node[pos=0.85,above] {$a'_2$}
    node[pos=0.15,below] {$b_2$} node[pos=0.85,below] {$b'_2$} ++ (proc2);
    \draw[double]
    (agent3) -- node[pos=0.15,above] {$a_3$} node[pos=0.85,above] {$a'_3$}
    node[pos=0.15,below] {$b_3$} node[pos=0.85,below] {$b'_3$} ++ (proc3);

    \draw (agent1) to[bend left] node[pos=0.15,above] {$d$} node[pos=0.85,above] {$d'$} (proc1);

    \draw[->,dashed] (proc1) -- node[above right] {data transfer} (proc2);
    \node[draw=none] (proc2toproc3data1) at ( 3.25,-2.75) {};
    \node[draw=none] (proc2toproc3data2) at (-3.25,-2.75) {};
    \draw[->,dashed] (proc2) -- (proc2toproc3data1.center) -- (proc2toproc3data2.center) -- (proc3);
    \node[draw=none] (proc3toproc1data1) at (-2.00, 1.25) {};
    \node[draw=none] (proc3toproc1data2) at ( 2.00, 1.25) {};
    \draw[->,dashed] (proc3) -- (proc3toproc1data1.center) -- (proc3toproc1data2.center) -- (proc1);
  \end{tikzpicture}
\end{center}

### 2020-10-11 priority-cp.tex
Added scheduler. Made reasonable progress.

@@
 \usingnamespace{pcp}
-
-\begin{figure}[t]
-  \centering
-  \includegraphics[width=0.8\columnwidth]{scheduler}
-  \vspace{-4mm}
-  \caption{Cyclic Scheduler from \cite{dardhagay18}}
-  \label{fig:scheduler}
-  \vspace{-4mm}
-  \end{figure}
-
-\subsection{Syntax}
 
@@
 \]
-Process $\tm{\link{x}{y}}$ links channels $x$ and $y$ and forwards communication from one to the other.
+The process $\tm{\link{x}{y}}$ links endpoints $x$ and $y$ and forwards communication from one to the other.
 $\tm{\res{x}{y}{P}}$, $\tm{(\ppar{P}{Q})}$ and $\tm{\halt}$ denote respectively the restriction processes where channel endpoints $\tm{x}$ and $\tm{y}$ are bound together and with scope $\tm{P}$, the parallel composition of processes $\tm{P}$ and $\tm{Q}$ and the terminated process.
@@
 
-\subsection{Typing}
+\paragraph{Typing}
 \input{fig/pcp-typing}
@@
 
-\subsection{Example: Milner's Cyclic Scheduler}
-We illustrate PCP with the cyclic scheduler~\cite{milner89}. The full typing for $\tm{\sched}$ is given in \cite{dardhagay18}.
+\paragraph{Example: Milner's Cyclic Scheduler}
+\[
+  \begin{array}{lrlr}
+    \tm{\sched}
+    & \defeq & \tm{\res{a_1}{a'_1}{\dots{\res{a_n}{a'_n}{}}}}
+               \tm{\res{b_1}{b'_1}{\dots{\res{b_n}{b'_n}{}}}}
+               \tm{\res{c_1}{c'_1}{\dots{\res{c_n}{c'_n}{}}}}
+               \tm{\res{d}{d'}{}}
+    \\ &     & \tm{(
+               \ppar{\proc1}{\agent1}
+               \parallel
+               \prod_{2\leq{i}\leq{n}}(\ppar{\proc{i}}{\wait{c'_{i-1}}{\agent{i}}})
+               )}
+    \\
+    \tm{\agent1}
+    & \defeq & \tm{\close{a_i}{\wait{b_i}{\close{c_i}{\wait{c'_n}{\close{d}{\halt}}}}}}
+    \\
+    \tm{\agent{i}}
+    & \defeq & \tm{\close{a_i}{\wait{b_i}{\close{c_i}{\halt}}}}
+    \\
+    \tm{\proc{i}}
+    & \defeq & \tm{\wait{a'_i}{\plug{P_i}{\close{b'_i}{Q_i}}}}
+  \end{array}
+\]
 
-We consider a simplified version: a set of 3 agents $\tm{\agent0}, \tm{\agent1}, \tm{\agent2}$ is scheduled to perform a certain task in cyclic order, starting with $\tm{\agent0}$.
-$\tm{\agent1}$ and $\tm{\agent2}$ send the result of some computation a collector process, respectively $\tm{\collector1}$ and $\tm{\collector2}$, before transmitting data further to the next agent in the cycle, being $\tm{\agent{(i+1)\bmod 3}}$.
-At the end of the round, $\tm{\agent0}$ send the final result of the computation to its collector $\tm{\collector0}$.
-
-In line with the logical consistency requirements of the Curry-Howard correspondence, we define a finite version of Milner's scheduler, which executes one round of communication.
-$$
-\begin{array}{rcll}
-  \tm{\sched}
-  & \defeq
-  & \tm{
-    \res{x_0}{x'_0}{\res{x_1}{x'_1}{\res{x_2}{x'_2}{
-    \res{\overrightarrow{y_0}}{\overleftarrow{y_1}}{
-    \res{\overrightarrow{y_1}}{\overleftarrow{y_2}}{
-    \res{\overrightarrow{y_2}}{\overleftarrow{y_0}}{}}}}}}
-    }\\
-  &
-  & \tm{(\ppar{\agent0}{\ppar{\agent1}{\ppar{\agent2}{\ppar{\collector0}{\ppar{\collector1}{\collector2}}}}})}\\
-  \tm{\agent0}
-  & \defeq
-  & \tm{\usend{\overrightarrow{y_0}}{d_0}{\recv{\overleftarrow{y_0}}{d'_0}{\usend{x_0}{d'_0}{
-    \wait{\overleftarrow{y_0}}{\close{\overrightarrow{y_0}}{\close{x_0}{\halt}}}}}}} \\
-  \tm{\agent{i}}
-  & \defeq
-  & \tm{\recv{\overleftarrow{y_i}}{d_i}{\usend{x_i}{d_i}{\recv{x_i}{d'_i}{\usend{\overrightarrow{y_i}}{d'_i}{
-    \wait{\overleftarrow{y_i}}{\close{\overrightarrow{y_i}}{\close{x_i}{\halt}}}}}}}}
-  & i \in \{1,2\} \\
-  \tm{\collector{i}}
-  & \defeq & \tm{\recv{x'_i}{d_i}{Q_i}}
-  & i \in \{0,1,2\}
-\end{array}
-$$
-In the above process, channels $\tm{d_i}$ and $\tm{d'_i}$ represent the data being transmitted.
-For simplicity, let them be channels of a terminated type.
-The processes $\tm{Q_i}$ represent the inner workings of the collector processes, each of which must eventually send back a response over $\tm{x'_i}$ and close the channel.
-\subsection{Operational Semantics}
+\paragraph{Operational Semantics}
 \input{fig/pcp-operational-semantics}
@@
 
-\subsection{Metatheory}
+\paragraph{Metatheory}
 
@@
 
-\begin{definition}[Actions]
-  A~process acts on an endpoint $\tm{x}$ if it is in one of the following forms:
-  \begin{multicols}{3}
-    \begin{itemize}[noitemsep,topsep=0pt,parsep=0pt,partopsep=0pt]
-    \item $\tm{\link{x}{y}}$ 
-    \item $\tm{\link{y}{x}}$
-    \item $\tm{\send{x}{y}{P}}$
-    \item $\tm{\recv{x}{y}{P}}$
-    \item $\tm{\close{x}{P}}$
-    \item $\tm{\wait{x}{P}}$
-    \item $\tm{\inl{x}{P}}$
-    \item $\tm{\inr{x}{P}}$
-    \item $\tm{\offer{x}{P}{Q}}$
-    \item $\tm{\absurd{x}}$
-    \end{itemize}
-  \end{multicols}
-  \noindent
-  A~process is an action if it acts on some endpoint $\tm{x}$.
-\end{definition}
-
-\begin{definition}[Canonical Forms]
-  \label{def:pcp-canonical-forms}
-  A~process $\tm{P}$ is in canonical form if it is in one of the following forms:
-  \begin{itemize}[noitemsep,topsep=0pt,parsep=0pt,partopsep=0pt]
-  \item
-    $\tm{\halt}$
-  \item
-    $\tm{\res{x_1}{x'_1}{\dots\res{x_n}{x'_n}{(P_1 \parallel\dots\parallel P_m)}}}$
-    where $m>0$ and each $\tm{P_j}$ is an action.
-  \end{itemize}
-\end{definition}
-
-\begin{restatablelemma}{lempcpcanonicalforms}[Canonical Forms]
-  \label{lem:pcp-canonical-forms}
-  If $\seq{P}{\ty{\Gamma}}$, there exists some $\tm{Q}$ such that $\tm{P}\equiv\tm{Q}$ and $\tm{Q}$ is in canonical form.
-\end{restatablelemma}
-\begin{proof}
-  If $\tm{P}=\tm{\halt}$, we are done.
-Otherwise, we move any $\nu$-binders to the top using \LabTirName{SC-ResExt}, and discard any superfluous occurrences of $\tm{\halt}$ using \LabTirName{SC-ParNil}.
-\end{proof}
-\todo{We need to comment here why it is that we presented canonical forms, to start with...}
 \begin{restatabletheorem}{thmpcpclosedprogress}[Closed Progress, $\red$]
@@
   \hfill\\%newline before theorem statement
-  If $\seq{P}{\emptyenv}$ and $\tm{P}$ is in canonical form, then either $\tm{P}=\tm{\halt}$ or there exists a $\tm{Q}$ such that $\tm{P}\red\tm{Q}$.
+  If $\seq{P}{\emptyenv}$, then either $\tm{P}=\tm{\halt}$ or there exists a $\tm{Q}$ such that $\tm{P}\red\tm{Q}$.
 \end{restatabletheorem}

### 2020-10-11 priority-gv.tex
Added scheduler. Made reasonable progress.

@@
 In other words, GV only allows tree-shaped communication structures, so it is trivially free from cycles, and therefore deadlocks.
-However, this is overly restrictive, as we rule out \emph{all} cyclic communication structures, not just those with deadlocks.
+However, this is overly restrictive, as it rules out \emph{all} cyclic communication structures, not just those with deadlocks.
 
-Priority GV offers a more fine-grained analysis of the communication structure.
-It allows cyclic communication structures,  and as a result, it is more expressive.
-To illustrate this, we use PGV to implement an instance of Milner's Cyclic Scheduler in~\todo{Add reference}.
+Priority GV offers a more fine-grained analysis of the communication structure, and allows cyclic communication structures.
+To illustrate this, we use PGV to implement a variant of Milner's cyclic scheduler in~\cref{sec:pgv-example}.
 
-\subsubsection{Types}
-Types in PGV are divided into \emph{session types} $\ty{S}$ and \emph{types} $\ty{T}$, $\ty{U}$.
+\paragraph{Types}
+Types in PGV are divided into session types ($\ty{S}$) and types ($\ty{T}$, $\ty{U}$).
 %
@@
 
-\subsubsection{Environments}
+\paragraph{Environments}
 Typing environments $\ty{\Gamma}$, $\ty{\Delta}$ associate names to types.
@@
 
-\subsubsection{Duality}
+\paragraph{Duality}
 Duality plays a crucial role in session types.
@@
 
-\subsubsection{Priorities}
+\paragraph{Priorities}
 The function $\pr$ returns the smallest priority of a session type.
-The type system guarantees that the top-most connective always holds the smallest priority, so we simply return the priority of the top-most connective.
+The type system guarantees that the top-most connective always holds the smallest priority, so we simply return the priority of the top-most connective:
 \[
@@
 \]
-We extend the function $\pr$ to value types and typing contexts by returning the smallest priority in the type or context, or $\cs{\top}$ if there is no priority.
-\todo{What about the function type?}
+We extend the function $\pr$ to types and typing contexts by returning the smallest priority in the type or context, or $\cs{\top}$ if there is no priority:
 \[
@@
   \minpr(\ty{\tylolli[\cs{p},\cs{q}]{T}{U}}) & = & \cs{p} \\
-  \\
   \minpr(\ty{\Gamma}, \tmty{x}{A})           & = & \minpr(\ty{\Gamma})\sqcap\pr(\ty{A})
@@
   \minpr(\ty{S})                             & = & \pr(\ty{S}) \\
-  \\
   \minpr(\ty{\emptyenv})                     & = & \ptop
@@
 
-\subsubsection{Terms}
-Terms are defined by the following grammar:
+\paragraph{Terms}
+Terms ($\tm{L}$, $\tm{M}$, $\tm{N}$) are defined by the following grammar:
 \[
@@
 
-\subsubsection{Internal and External Choice}
+\paragraph{Internal and External Choice}
 Typically, session-typed languages feature constructs for internal and external choice.
@@
 
-\subsection{Operational Semantics}
-\paragraph*{Runtime Term Syntax}
+\paragraph{Operational Semantics}
+Priority GV terms are evaluated as part of a configuration of processes. Configurations are defined by the following grammar:
 \[
@@
     \sep        \tm{\ppar{\conf{C}}{\conf{D}}}
-    \sep        \tm{\res{x}{y}{\conf{C}}}
+    \sep        \tm{\res{x}{x'}{\conf{C}}}
   \\
@@
     \sep        \tm{\child}
+\end{array}
+\]
+Configurations ($\tm{\conf{C}}$, $\tm{\conf{D}}$, $\tm{\conf{E}}$) consist of threads $\tm{\phi\;M}$, parallel compositions $\tm{\ppar{\conf{C}}{\conf{D}}}$, and name restrictions $\tm{\res{x}{x'}{\conf{C}}}$.
+To preserve the functional nature of PGV, where programs return a single value, we use flags ($\tm{\phi}$) to differentiate between the main thread, marked $\tm{\main}$, and child threads created by $\tm{\spawn}$, marked $\tm{\child}$.
+Only the main thread returns a value.
+We determine the flag of a configuration by combining the flags of all threads in that configuration:
+\[
+\begin{array}{ll}
+  \tm{\main}  + \tm{\child} = \tm{\main}
   \\
+  \tm{\child} + \tm{\main}  = \tm{\main}
+\end{array}
+\quad
+\begin{array}{ll}
+  \tm{\child} + \tm{\child} = \tm{\child}
+  \\
+  \tm{\main}  + \tm{\main} \; \text{undefined}
+\end{array}
+\]
+Values ($\tm{V}$, $\tm{W}$), ready terms ($\tm{R}$) evaluation contexts ($\tm{E}$), thread evaluation contexts ($\tm{\conf{F}}$), and configuration contexts ($\tm{\conf{G}}$) are defined by the following grammars:
+\[
+\begin{array}[t]{lcl}
   \tm{V}, \tm{W}
@@
     \sep        \tm{\casesum{E}{x}{M}{y}{N}} \\
-  & \sep      & \tm{\absurd{E}} \\
+  & \sep      & \tm{\absurd{E}}
   \\
@@
   \tm{\conf{G}}
-  & \coloneqq & \tm{\hole} \\
-  & \sep      & \tm{\ppar{\conf{G}}{\conf{C}}}
+  & \coloneqq & \tm{\hole}
+    \sep        \tm{\ppar{\conf{G}}{\conf{C}}}
     \sep        \tm{\res{x}{y}{\conf{G}}}
@@
 \]
-\[
-\begin{array}{ll}
-  \tm{\main}  + \tm{\child} = \tm{\main}
-  \\
-  \tm{\child} + \tm{\main}  = \tm{\main}
-\end{array}
-\quad
-\begin{array}{ll}
-  \tm{\child} + \tm{\child} = \tm{\child}
-  \\
-  \tm{\main}  + \tm{\main} \; \text{undefined}
-\end{array}
-\]
-See \cref{fig:pgv-operational-semantics}.
+We factor the reduction relation of PGV into a deterministic reduction on terms ($\tred$) and a non-deterministic reduction on configurations ($\cred$), see \cref{fig:pgv-operational-semantics}.
+We write $\tred^+$ and $\cred^+$ for the transitive closures, and $\tred^\star$ and $\cred^\star$ for the reflexive-transitive closures.
+
+Term reduction is the standard call-by-value, left-to-right evaluation for GV, and only deviates from reduction for the linear $\lambda$-calculus in that it reduces terms to values \emph{or} ready terms waiting to perform a communication action.
+
+Configuration reduction resembles evaluation for a process calculus: \LabTirName{E-Link}, \LabTirName{E-Send}, and \LabTirName{E-Close} perform communications, \LabTirName{E-LiftC} allows reduction under configuration contexts, and \LabTirName{E-LiftSC} embeds a structural congruence $\equiv$.
+The remaining rules mediate between the process calculus and the functional language: \LabTirName{E-New} and \LabTirName{E-Spawn} evaluate the $\tm{\new}$ and $\tm{\spawn}$ constructs, creating the equivalent configuration constructs, and \LabTirName{E-LiftM} embeds term reduction.
+
+While our configuration reduction is based on the standard evaluation for GV, the increased expressiveness of PGV allows us to simplify the relation on two counts:
+
+Firstly, we decompose the $\tm{\fork}$ construct.
+In GV, $\tm{\fork}$ creates a new channel, spawns a child thread, and, when the child thread finishes, it closes the channel to its parent.
+In PGV, these are three separate operations: $\tm{\new}$, $\tm{\spawn}$, and $\tm{\close}$.
+We no longer require that every child thread finishes by returning a terminated channel.
+Consequently, we also simplify the evaluation of the $\tm{\link}$ construct.
+In GV, $\tm{\link}$ needs to return a terminated channel after evaluating the link.
+In PGV, the behaviour matches that of link in CP.
+
+Secondly, our structural congruence is type preserving.
+Consequently, we can embed it directly into the reduction relation.
+In GV, this is not the case, and subject reduction relies on proving that if $\equiv\cred$ ends up in an ill-typed configuration, we can rewrite it to a well-typed configuration using $\equiv$.
+
 \input{fig/pgv-operational-semantics}
 
-\subsection{Typing}
+\paragraph{Typing}
+\Cref{fig:pgv-typing} gives the typing rules for PGV.
+
+Terms are typed by a judgement $\tseq[\cs{p}]{\ty{\Gamma}}{M}{T}$.
+Typing is standard for the linear $\lambda$-calculus.
+Linearity is ensured by splitting environments on branching rules, requiring that the environment in the variable rule consists of just the variable, and the environment in the constant and unit rules are empty.
+The constants $\tm{K}$ are typed using type schemas, and embedded using \LabTirName{T-Const}.
+
+The only non-standard feature of the typing rules is the priority annotations.
+Priorities are based on \emph{obligations/capabilities} used by Kobayashi~\cite{kobayashi06}, and simplified to single priorities following Padovani~\cite{padovani14}.
+The integration of priorities into GV is adapted from Padovani and Novara~\cite{padovaninovara15}.
+Paraphrasing from Dardha and Gay~\cite{dardhagay18}, priorities obey the following two laws:
+\begin{enumerate}
+  \item an action with priority $\cs{o}$ happens before an action with priority $\cs{o+n}$;
+  \item communication requires \emph{equal} priorities for the dual actions.
+\end{enumerate}
+
+In PGV, Sequents are annotated by a priority bound $\cs{p}$, which is the \emph{upper bound} on the priority of the communication actions performed while evaluating the term.
+The \emph{lower bound} is obtained from the typing environment.
+Terms with sequential evaluation, such as \LabTirName{T-LetUnit}, have side conditions to ensure deadlock freedom.
+For instance, the side condition for typing $\tm{\letunit{M}{N}}$ requires that $\tm{M}$ has finished communicating before $\tm{N}$ starts communicating.
+
+Configurations are typed by a judgement $\cseq[\phi]{\ty{\Gamma}}{\conf{C}}$.
+Configuration typing is based on the standard typing for GV.
+Terms are embedded either as main or as child threads.
+The priority bound from the term typing is discarded, as configurations contain no further sequential actions.
+Main threads are allowed to return a value, whereas child threads are required to return the unit value.
+Sequents are annotated with a flag $\phi$, which ensure that there is at most one main thread.
+
+While our configuration typing is based on the standard typing for GV, it differs on two counts:
+
+Firstly, we require that child threads return the unit value, as opposed to a terminated channel.
+
+Secondly, we simplify typing for parallel composition.
+In GV, each parallel composition must split \emph{exactly one} channel of the channel pseudo-type $\ty{S^\sharp}$ into two endpoints of type $\ty{S}$ and $\ty{\co{S}}$ across the parallel composition.
+Consequently, associativity of parallel composition violates type preservation.
+Unfortunately, the practice plays a crucial role in guaranteeing deadlock freedom.
+In PGV, we guarantee deadlock freedom through different means, which removes the channel pseudo-type $\ty{S^\sharp}$, simplifies typing for parallel composition, and restores type preservation for the structural congruence.
 
@@
 
-\todo{Define the ``priority bound of a term $\tm{M}$'' as $\cs{p}$ where $\tseq[\cs{p}]{\ty{\Gamma}}{M}{T}$.}
+\paragraph{Example: Milner's Cyclic Scheduler}\label{sec:pgv-example}
+We illustrate the newfound ability of PGV by encoding a variant of Milner's cyclic scheduler~\cite{milner89,dardhagay18}.
+Suppose a set of processes $\tm{\proc{i}}$, $1\leq{i}\leq{n}$, should be scheduled to perform some tasks in cyclic order, starting with $\tm{\proc0}$, and circling all the way around to $\tm{\proc{n}}$, notifying $\tm{\proc0}$ when every other process has finished their task.
 
-\subsection{Example: A~Cyclic Scheduler}
-We illustrate the newfound ability of PGV by encoding a variant of Milner's cyclic scheduler~\cite{milner89,dardhagay18}.
-Suppose three processes ($\tm{\collector0}$, $\tm{\collector1}$, and $\tm{\collector2}$) should be scheduled to perform their tasks in cyclic order, starting with $\tm{\collector1}$, and circling all the way around to $\tm{\collector0}$.
-We implement the scheduler with three agents ($\tm{\agent0}$, $\tm{\agent1}$, and $\tm{\agent2}$) which represent their respective processes in the scheduler.
-
-\begin{center}
-  \begin{tikzpicture}
-    \node[shape=circle,draw=black] (agent0) at ( 0.00, 0.00) {\tm{\agent0}};
-    \node[shape=circle,draw=black] (agent1) at ( 1.75,-1.75) {\tm{\agent1}};
-    \node[shape=circle,draw=black] (agent2) at (-1.75,-1.75) {\tm{\agent2}};
-    \node[shape=circle,draw=black] (proc0)  at ( 3.00, 0.00) {\tm{\collector0}};
-    \node[shape=circle,draw=black] (proc1)  at ( 4.50,-1.75) {\tm{\collector1}};
-    \node[shape=circle,draw=black] (proc2)  at (-4.50,-1.75) {\tm{\collector2}};
-    \node[draw=none]               (middle) at ( 0.00,-1.75) {};
-
-    \draw
-    (agent0) -- node[pos=0.35,above] {$c_0$}
-    node[pos=1.0,above] {$d_1$} ++ (agent1);
-    \draw
-    (agent1) -- node[pos=0.1,above] {$c_1$}
-    node[pos=0.9,above] {$d_2$} ++ (agent2);
-    \draw
-    (agent2) -- node[pos=0.0,above] {$c_2$}
-    node[pos=0.65,above] {$d_0$} ++ (agent0);
-
-    \draw
-    (agent0) -- node[pos=0.15,above] {$a_0$}
-    node[pos=0.85,above] {$b_0$} (proc0);
-    \draw
-    (agent1) -- node[pos=0.15,above] {$a_1$}
-    node[pos=0.85,above] {$b_1$} (proc1);
-    \draw
-    (agent2) -- node[pos=0.15,above] {$a_2$}
-    node[pos=0.85,above] {$b_2$} (proc2);
-  \end{tikzpicture}
-\end{center}
+Our scheduler consists of set of agents $\tm{\agent{i}}$, $1\leq{i}\leq{n}$, each representing their respective process.
+Each process $\tm{\proc{i}}$ waits for the signal to start their task on $\tm{a'_i}$, and signals completion on $\tm{b'_i}$.
+Each agent signals their process to start on $\tm{a_i}$, waits for their process to finish on $\tm{b_i}$, and then signals for the next agent to continue on $\tm{c_i}$.
+The agent $\tm{\agent0}$ initiates, then waits for each other process to finish, and signals $\tm{\proc0}$ on $\tm{d}$.
+Every other agent $\tm{\agent{i}}$, $2\leq{i}\leq{n}$ waits on $\tm{c'_{i-1}}$ for the signal to start.
+The dotted lines represent the optional transfer of data between processes.
+With three processes, the scheduler looks as follows:
+\input{fig/pgv-scheduler.tex}
+We implement the scheduler as follows, using $\tm{\prod_{i}\conf{C}_i}$ to denote the parallel composition of the processes $\tm{\conf{C}_i}$:
 \[
@@
     \tm{\sched}
-    & \defeq & \tm{
-               \res{a_0}{b_0}{\res{a_1}{b_1}{\res{a_2}{b_2}{
-               \res{c_0}{d_1}{
-               \res{c_1}{d_2}{
-               \res{c_2}{d_0}{}}}}}}}
-    \\ &     & \multicolumn{2}{l}{%
-               \tm{(\ppar{\child\;\agent0}{\ppar{\child\;\agent1}{\child\;\ppar{\agent2}{
-               \ppar{\child\;\collector0}{\ppar{\child\;\collector1}{\child\;\collector2}}}}})}}
+    & \defeq & \tm{\res{a_1}{a'_1}{\dots{\res{a_n}{a'_n}{}}}}
+               \tm{\res{b_1}{b'_1}{\dots{\res{b_n}{b'_n}{}}}}
+               \tm{\res{c_1}{c'_1}{\dots{\res{c_n}{c'_n}{}}}}
+               \tm{\res{d}{d'}{}}
+    \\ &     & \tm{\begin{array}{lll}
+                     (
+                     & \phi\;\proc1
+                     \parallel
+                     \child\;\andthen{\agent1}{\andthen{\wait\;{c'_3}}{\close\;{d}}}
                      \\
-    \tm{\agent0}
-    & \defeq & \tm{\letbind{c_0}{\send\;\pair{\text{data}_0}{c_0}}{%
-               \letpair{\text{data}'_0}{d_0}{\recv\;d_0}{}}}
-    \\ &     & \tm{\letbind{a_0}{\send\;\pair{\text{data}_0}{a_0}}{}}
-    \\ &     & \tm{\andthen{\wait\;{d_i}}{\andthen{\close\;{c_i}}{\close\;{a_i}}}}
+                     \parallel
+                     &
+                     \prod_{2\leq{i}\leq{n}}
+                     (\child\;\proc{i} \parallel \child\;\andthen{\wait\;{c'_{i-1}}}{\agent{i}})
+                     & )
+                   \end{array}}
     \\
     \tm{\agent{i}}
-    & \defeq & \tm{\letpair{d_i}{d_i}{\recv\;d_i}{%
-               \letbind{a_i}{\send\;\pair{\text{data}_i}{a_i}}{}}}
-    \\ &     & \tm{\letpair{\text{data}'_i}{a_i}{\recv\;{a_i}}{%
-               \letbind{c_i}{\send\;\pair{\text{data}_i}{c_i}}{}}}
-    \\ &     & \tm{\andthen{\wait\;{d_i}}{\andthen{\close\;{c_i}}{\close\;{a_i}}}}
-       & i \in \{1,2\}
+    & \defeq & \tm{\andthen{\close\;{a_i}}{\andthen{\wait\;{b_i}}{\close\;{c_i}}}}
     \\
-    \tm{\collector{i}}
-    & \defeq & \tm{\letpair{d'_i}{d_i}{\recv\;d_i}{M_i}}
-       & i \in \{0,1,2\}
+    \tm{\proc{i}}
+    & \defeq & \tm{\andthen{\wait\;{a'_i}}{\plug{M_i}{\andthen{\close\;{b'_i}}{N_i}}}}
   \end{array}
 \]
+Each of the channels in the scheduler is has the terminated type, and is merely used to synchronise.
+For simplicity, we let $\tm{\sched}$ be a configuration. However, we could equivalently have represented it as a function which takes three processes as arguments, and creates this configuration.
 
@@
   By induction on the derivation of $\tseq[\cs{p}]{\ty{\Gamma}}{V}{T}$.
-  (Details in \cref{prf:lem-pgv-value-done}.)
+  Details in \cref{prf:lem-pgv-value-done}.
 \end{proof}
@@
 Whenever the contexts in the premises and conclusion differ, we use \cref{lem:pgv-value-done} to prove that the priorities are preserved.
-  (Details in \cref{prf:lem-pgv-substitution}.)
+  Details in \cref{prf:lem-pgv-substitution}.
 \end{proof}
@@
   By induction on the derivation of $\tm{M}\tred\tm{M'}$.
-  (Details in \cref{prf:lem-pgv-subject-reduction-terms}.)
+  Details in \cref{prf:lem-pgv-subject-reduction-terms}.
 \end{proof}
@@
   By induction on the derivation of $\tm{\conf{C}}\equiv\tm{\conf{C'}}$.
-  (Details in \cref{prf:lem-pgv-subject-congruence}.)
+  Details in \cref{prf:lem-pgv-subject-congruence}.
 \end{proof}
@@
   By induction on the derivation of $\tm{\conf{C}}\cred\tm{\conf{C'}}$.
-  (Details in \cref{prf:thm-pgv-subject-reduction-confs}.)
+  Details in \cref{prf:thm-pgv-subject-reduction-confs}.
 \end{proof}
@@
   A~term $\tm{L}$ is ready if it is of the form $\tm{\plug{E}{M}}$, where $\tm{M}$ is of the form $\tm{\new}$, $\tm{\spawn\;N}$, $\tm{\link\;\pair{x}{y}}$ or $\tm{\link\;\pair{y}{x}}$, or $\tm{M}$ acts on $\tm{x}$.
-In the last case, we say that $\tm{L}$ is ready to act on $\tm{x}$.
+In the latter case, we say that $\tm{L}$ is ready to act on $\tm{x}$.
 \end{definition}
-
-\begin{restatablelemma}{lempgvreadypriority}
-  \label{lem:pgv-ready-priority}
-  If $\tseq[\cs{p}]{\ty{\Gamma}}{L}{T}$ is ready to act on $\tmty{x}{S}\in\ty{\Gamma}$, then the priority bound $\cs{p}$ is some priority $\cs{o}$, \ie not $\cs{\pbot}$ or $\cs{\ptop}$.
-\end{restatablelemma}
-\begin{proof}
-  Let $\tm{L}=\tm{\plug{E}{M}}$.
-By induction on the structure of $\tm{E}$.
-$\tm{M}$ has priority $\pr({\ty{S}})$, and each constructor of the evaluation context $\tm{E}$ passes on the \emph{maximum} of the priorities of its premises.
-No rule introduces the priority bound $\cs{\ptop}$ on the sequent.
-\end{proof}
 
@@
 \end{restatabletheorem}
-\begin{proof}
-  Details in \cref{prf:thm-pgv-closed-progress-confs}.
-\end{proof}
+\input{prf/thm-pgv-closed-progress-confs}
 

### 2020-10-12 introduction.tex
work on introduction

--- 2020-10-09 introduction.tex
@@
 \section{Introduction}
+Communication and concurrency are key in our modern society.
+Every service or device we interact with, from smart speakers, to health monitors or any other IoT device, relies on communication.
+Hence, specification and verification of communication in concurrent and distributed systems is paramount for our own safety.
+Session types are a type formalism which can be used to specify and verify communication protocols between two or more communicating agents in concurrent scenarios \cite{honda93,takeuchihonda94,hondavasconcelos98,carbonehonda07}.
 
-Things to discuss in the intro:
+Session types have been integrated in several programming paradigms, but to date the most successful settings remain the $\pi$-calculus \cite{sangiorgiwalker01} ---  a calculus for communication and concurrency, and GV (Good Variation) \cite{gayvasconcelos10,gayvasconcelos12} ---  a concurrent $\lambda$ calculus.
+Session-typed concurrent $\lambda$-calculus or $\pi$-calculus satisfy safe communication and session fidelity, namely that communication proceeds without any mismatch and follows the predefined protocol specified by a session type.
+Another key property in a concurrent setting is deadlock freedom, meaning that agents communicate without getting stuck in cyclic dependencies.
+Unfortunately, deadlock freedom is a more involving property and session types alone are not enough to guarantee it.
 
+Deadlock freedom in the $\pi$-calculus with or without session types has been extensively studied.
+A growing line of work is based on the Curry-Howard correspondences between linear logic and session-typed $\pi$-calculus \cite{cairespfenning10,wadler12}, which guarantee deadlock freedom by design.
+This is achieved by the cut rule, which merges together constructs for parallel composition and restriction thus forcing processes to share only one channel for communication and eliminates cyclic dependencies.
+Recent developments in the Classical Process \cite[CP]{wadler12} line of work have led to various approaches to decoupling these constructs, either while maintaining the strong correspondence to logic, as in Hypersequent CP~\cite[HCP]{kokkemontesi19popl,kokkemontesi19tlla}, or by weakening the correspondence to logic in exchange for a more expressive language, as in Priority CP~\cite[PCP]{dardhagay18}.
+PCP decouples CP's cut into separate constructs for name parallel composition --- by  introducing a mix rule, and restriction --- by introducing a cycle rule.
+Differently from previous work, PCP allows for cyclic structures and restores deadlock freedom by adding priorities following Kobayashi~\cite{kobayashi06} and Padovani \cite{padovani14}.
+
+With regards to GV, the original work on session types in GV \cite{gayvasconcelos10,gayvasconcelos12} does not satisfy deadlock freedom.
+Later on, calculi in the GV family~\cite{wadler15,lindleymorris15} have achieved deadlock freedom via a syntactic restriction, i.e., by combining channel creation and thread spawning into a single operation, called fork, corresponding to the logical cut in CP, itself the combination of a name restriction and a parallel composition.
+Unfortunately, this is overly restrictive, as it limits the possible communication structures to trees.
+There are several benefits of working with a functional language as opposed to process calculi has its benefits: it supports higher-order functions (and therefore abstraction) not usually present in process calculi; and it allows us to derive extensions of the communication fragment of the language via well-understood extensions of the functional fragment, i.e., deriving internal/external choice from sum types.
+Moreover, the benefit of working with GV over other session-typed functional languages is that GV has strong ties to linear logic, via its relation to CP~\cite{wadler12}, and consequently it has strong formal properties, e.g., deadlock freedom, albeit in a restrictive way.
+
+The above observations on GV together with the work on PCP led us to our research question:
+
+\emph{Can we design a more expressive GV where deadlock freedom is guaranteed by design and communications structures are not limited to trees?}
+
+We address this research question and by following the transition from CP to Priority CP, we presents 
+Priority GV (PGV), a~variant of GV which decouples channel creation from thread creation, and restores deadlock freedom by adding priorities.
+This work closes the circle of CP and GV, their relation and their priority versions PCP and PGV.
+
+\paragraph*{Contributions and Structure of the Paper}
 \begin{itemize}
-\item GV is a session-typed functional language.
-\item
-  The benefits of using a functional language as opposed to a process calculus are that:
-  \begin{itemize}
-  \item 
-    it supports higher-order functions (and therefore abstraction) not usually present in process calculi; and
-  \item
-    it allows us to derive extensions of the communication fragment of the language via well-understood extensions of the functional fragment, i.e., deriving internal/external choice from sum types.
+\item \textbf{PGV}: we present Priority GV in sec ??, the syntax of types and terms, operational semantics and typing rules.
+We prove subject reduction (\Cref{thm:pgv-subject-reduction-confs}) and closed progress (\Cref{lem:pgv-closed-progress-confs})
+\item \textbf{Milner's Cyclic Scheduler}: We present Milner's Cyclic Scheduler \cite{milner89} \cref{sec:pgv-example} for the first time in GV.
+\item \textbf{Updated PCP}: we present an updated version of PCP \cite{dardhagay18} where we remove the commuting conversions, moving away from reduction as cycle elimination, and towards reduction as one would expect for a process calculus.
+We report the Cyclic Scheduler from \cite{dardhagay18}.
+\item \textbf{PCP into PGV}: We present an encoding of PCP into PGV and show that it is sound and complete with respect to typing (\Cref{thm:pcp-to-pgv-confs-preservation}) and semantics, by proving operational correspondence (\Cref{thm:pcp-to-pgv-operational-correspondence-soundness} and \Cref{thm:pcp-to-pgv-operational-correspondence-completeness}).
 \end{itemize}
-\item
-  The benefits of using GV over other session-typed functional languages are that:
-  \begin{itemize}
-  \item
-    GV has strong ties to linear logic~\cite{wadler12}.
-  \end{itemize}
-\item
-  Historically, calculi in the GV family~\cite{wadler12,lindleymorris15} have achieved deadlock freedom via a syntactic restriction, i.e., by combining channel creation and thread creation into a single operation, called fork.
-Unfortunately, this is overly restrictive, as it limits the possible communication structures to trees.
-The calculus GV is based on, on the other hand, did not satisfy deadlock freedom~\cite{gayvasconcelos10}.
-\end{itemize}
-
-\begin{itemize}
-\item talking about what is GV and its "benefits"
-\item next, talking about deadlock freedom property, the benefits and the fact that the original GVs do not satisfy DF
-\item in the meantime, in a concurrent setting based on pi-calculus there is a curry-howard correspondence of logic and session typed pi, where DF holds
-\item talk about CP here
-\item \cite{dardhagay18} brought the research fwd and added priorities a la Kobayashi and Padovani to CP, producing PCP
-\item following that lead, we present PGV
-\item running example of the scheduler in PGV (hopefully)
-\item translation of PCP to PGV and what it means to have that.
-\item a contribution paragraph with the following
-  \begin{itemize}
-  \item PGV and what are the benefits
-  \item updated PCP
-  \item encoding PCP to PGV
-  \item technical results to be referred to explicitly
-  \end{itemize}
-\end{itemize}
-
 \end{document}

### 2020-10-13 introduction.tex
Minor update.
another pas to intro

--- 2020-10-11 introduction.tex
@@
 \section{Introduction}
+Communication and concurrency are at the basis of our modern society.
+Every service or device we interact with, from smart phones, to health monitors or any other IoT devices, relies on communication.
+Hence, verification of communication in such concurrent scenarios is paramount for our own safety.
+Session types are a type formalism which can be used to specify and verify communication protocols between two or more communicating agents \cite{honda93,takeuchihonda94,hondavasconcelos98,carbonehonda07}.
 
-We introduce Priority GV (PGV), a~session-typed concurrent $\lambda$-calculus.
+Session types have been studied for many programming paradigms, but to date the most successful settings remain the $\pi$-calculus \cite{sangiorgiwalker01} ---  a calculus for communication and concurrency, and Good Variation \cite[GV]{wadler15,lindleymorris15} ---  a concurrent $\lambda$-calculus.
+Session-typed concurrent $\lambda$-calculus and $\pi$-calculus satisfy communication safety and session fidelity, namely that communication proceeds without any type mismatch of the data being transmitted and follows the predefined protocol specified by a session type.
+Another key property in a concurrent setting is deadlock freedom, stating that agents communicate without getting stuck in cyclic dependencies.
+Unfortunately, deadlock freedom is a more involving property and session types alone are not enough to guarantee it.
+Several techniques have been developed to address deadlock freedom in the $\pi$-calculus and concurrent $\lambda$-calculus with session types.
 
+With regards to $\pi$-calculus, a growing line of work on deadlock freedom leverages the Curry-Howard correspondences between (intuitionistic or classical) linear logic and session-typed $\pi$-calculus \cite{cairespfenning10,wadler12}, which guarantee deadlock freedom by design.
+This is achieved by the cut rule, which merges together constructs for parallel composition and restriction thus forcing processes to share only one channel for communication and eliminating cyclic dependencies.
+Recent developments in the Classical Process \cite[CP]{wadler12} line of work have led to various approaches to decoupling these constructs, either by maintaining a strong correspondence to logic, as in Hypersequent CP~\cite[HCP]{kokkemontesi19popl,kokkemontesi19tlla}, or by weakening the correspondence to logic in exchange for a more expressive language, as in Priority CP~\cite[PCP]{dardhagay18}.
+PCP decouples CP's cut rule into two separate constructs, for parallel composition --- by  introducing a mix rule, and for restriction --- by introducing a cycle rule.
+Differently from previous work, PCP allows for cyclic structures and restores deadlock freedom by adding priorities to types following Kobayashi~\cite{kobayashi06} and Padovani \cite{padovani14}.
+Priorities capture the readiness of an action to fire, thus they establish an order of actions in a $\pi$-calculus process.
+
+With regards to GV, the original work on session types in GV \cite{gayvasconcelos10,gayvasconcelos12} did not satisfy deadlock freedom.
+Later on, calculi in the GV family~\cite{wadler15,lindleymorris15} have achieved deadlock freedom via a syntactic restriction, i.e., by combining channel creation and thread spawning into a single operation, called fork.
+This corresponds to the logical cut rule in CP, which respectively is the combination of a name restriction and a parallel composition.
+Unfortunately, as with CP, this is overly restrictive, as it limits the possible communication structures to trees.
+There are several benefits of working with a functional language as opposed to process calculi: a functional language supports higher-order functions (and therefore abstraction) not usually present in process calculi; and it allows us to derive extensions of the communication fragment of the language via well-understood extensions of the functional fragment, i.e., deriving internal/external choice from sum types.
+Moreover, the benefit of working with GV over other session-typed functional languages is that GV has strong ties to linear logic, via its relation to CP~\cite{wadler12}, and consequently it has strong formal properties, e.g., deadlock freedom, albeit in a restrictive way.
+Following the above observations, we aim to further investigate GV and deadlock freedom and thus pose our research question:
+
+\emph{Can we design a more expressive GV where deadlock freedom is guaranteed by design and communication structures are not limited to trees?}
+
+We address this research question by following the line of work from CP to Priority CP, and presents 
+Priority GV (PGV), a~variant of GV which decouples channel creation from thread creation, and restores deadlock freedom by adding priorities.
+This work closes the circle of the strong connection of CP and GV, together with their priority versions, PCP and PGV.
+
+\noindent
+\textbf{Contributions and Structure of the Paper}
 \begin{itemize}
-\item
-  Historically, calculi in the GV family~\cite{wadler15,lindleymorris15} have achieved deadlock freedom via a syntactic restriction, i.e., by combining channel creation and thread spawning into a single operation, called fork, corresponding to the logical cut in CP, itself the combination of a name restriction and a parallel composition.
-  Unfortunately, this is overly restrictive, as it limits the possible communication structures to trees.
-The calculus GV is based on, on the other hand, did not satisfy deadlock freedom~\cite{gayvasconcelos12}.
-\item
-  Recent developments in CP have led to various approaches to decoupling these constructs, either while maintaining the strong correspondence to logic, as in Hypersequent CP~\cite[HCP]{kokkemontesi19popl,kokkemontesi19tlla}, or by weakening the correspondence to logic in exchange for a more expressive language, as in Priority CP~\cite[PCP]{dardhagay18}.
-\item
-  PCP decouples CP's cut into separate constructs for name restriction and parallel composition, and restores deadlock freedom by adding priorities \`{a} la~\cite{kobayashi06}.
-\item
-  Following the example set by PCP, we present Priority GV (PGV), a~variant of GV which decouples channel creation from thread creation, and restores deadlock freedom by adding priorities.
-\item
-  We demonstrate the correspondence between PGV and PCP by proving operational correspondence for a translation from PCP to PGV.
-\item
-  The benefits of using a functional language as opposed to a process calculus are that:
-  \begin{itemize}
-  \item 
-    it supports higher-order functions (and therefore abstraction) not usually present in process calculi; and
-  \item
-    it allows us to derive extensions of the communication fragment of the language via well-understood extensions of the functional fragment, i.e., deriving internal/external choice from sum types.
+\item \textbf{Priority GV}: we present Priority GV (PGV) in~\Cref{sec:pgv}; we present the syntax of types and terms, operational semantics and typing rules.
+We show our type system for PGV is sound by proving subject reduction (\Cref{thm:pgv-subject-reduction-confs}) and closed progress (\Cref{lem:pgv-closed-progress-confs}).
+\item \textbf{Milner's Cyclic Scheduler in PGV}: to illustrate PGV, and following previous work \cite{dardhagay18} we present for the first time Milner's Cyclic Scheduler \cite{milner89} (in \Cref{sec:pgv-example}) as a term in GV.
+This example illustrates how a cyclic structure can now be expressed in GV thanks to the separation of channel creation and thread creation, and priorities.
+\item \textbf{Updated Priority CP}: we present an updated version of Priority CP (PCP) \cite{dardhagay18} in \Cref{sec:pcp} where we remove the commuting conversions and move away from a reduction relation as cycle elimination, and towards a reduction as one would expect for a process calculus.
+We then show Milner's Cyclic Scheduler in PCP (in \Cref{sec:pcp-example}) originally presented in \cite{dardhagay18}.
+\item \textbf{Encoding PCP into PGV}: we present an encoding of PCP into PGV in \Cref{sec:pcp-to-pgv}.
+We prove the encoding is sound and complete with respect to typing (\Cref{thm:pcp-to-pgv-confs-preservation}) and operational semantics (\Cref{thm:pcp-to-pgv-operational-correspondence-soundness} and \Cref{thm:pcp-to-pgv-operational-correspondence-completeness}).
+Finally, we can encode Milner's Cyclic Scheduler in PCP (\Cref{sec:pcp-example}) resulting in the Scheduler in PGV (\Cref{sec:pgv-example}).
 \end{itemize}
-\item
-  The benefits of using GV over other session-typed functional languages are that GV has strong ties to linear logic, via its relation to CP~\cite{wadler15}, and consequently it has strong formal properties, e.g., deadlock freedom.
-\end{itemize}
-
-\paragraph*{Contributions}
-\begin{itemize}
-\item PGV (typing rules, operational semantics, subject reduction, closed progress).
-\item An updated version of PCP, in which we remove the commuting conversions, moving away from reduction as cut elimination, and towards reduction as one would expect for a process calculus.
-\item A~translation from PCP into PGV (type preservation, operational correspondence).
-\end{itemize}
-
 \end{document}

### 2020-10-13 introduction.tex
Update introduction.tex

@@
 
-\noindent
-\textbf{Contributions and Structure of the Paper}
+\paragraph*{Contributions and Structure of the Paper}
 \begin{itemize}
@@
 We show our type system for PGV is sound by proving subject reduction (\Cref{thm:pgv-subject-reduction-confs}) and closed progress (\Cref{lem:pgv-closed-progress-confs}).
-\item \textbf{Milner's Cyclic Scheduler in PGV}: to illustrate PGV, and following previous work \cite{dardhagay18} we present for the first time Milner's Cyclic Scheduler \cite{milner89} (in \Cref{sec:pgv-example}) as a term in GV.
-This example illustrates how a cyclic structure can now be expressed in GV thanks to the separation of channel creation and thread creation, and priorities.
+\item \textbf{Milner's Cyclic Scheduler in PGV}: we present for the first time Milner's Cyclic Scheduler \cite{milner89} (\Cref{sec:pgv-example}) as a GV term.
+This example illustrates how a cyclic structure can now be expressed in PGV thanks to the separation of channel creation and thread creation, and priorities.
 \item \textbf{Updated Priority CP}: we present an updated version of Priority CP (PCP) \cite{dardhagay18} in \Cref{sec:pcp} where we remove the commuting conversions and move away from a reduction relation as cycle elimination, and towards a reduction as one would expect for a process calculus.

### 2020-10-13 introduction.tex
work on abstract

@@
 
-We address this research question by following the line of work from CP to Priority CP, and presents 
-Priority GV (PGV), a~variant of GV which decouples channel creation from thread creation, and restores deadlock freedom by adding priorities.
+We address this research question by following the line of work from CP to Priority CP, and present
+Priority GV (PGV), a~variant of GV which decouples channel creation from thread spawning, and restores deadlock freedom by adding priorities.
 This work closes the circle of the strong connection of CP and GV, together with their priority versions, PCP and PGV.

### 2020-10-13 introduction.tex
logic -> CLL

@@
 This is achieved by the cut rule, which merges together constructs for parallel composition and restriction thus forcing processes to share only one channel for communication and eliminating cyclic dependencies.
-Recent developments in the Classical Process \cite[CP]{wadler12} line of work have led to various approaches to decoupling these constructs, either by maintaining a strong correspondence to logic, as in Hypersequent CP~\cite[HCP]{kokkemontesi19popl,kokkemontesi19tlla}, or by weakening the correspondence to logic in exchange for a more expressive language, as in Priority CP~\cite[PCP]{dardhagay18}.
+Recent developments in the Classical Process \cite[CP]{wadler12} line of work have led to various approaches to decoupling these constructs, either by maintaining a strong correspondence to classical linear logic, as in Hypersequent CP~\cite[HCP]{kokkemontesi19popl,kokkemontesi19tlla}, or by weakening the correspondence to classical linear logic in exchange for a more expressive language, as in Priority CP~\cite[PCP]{dardhagay18}.
 PCP decouples CP's cut rule into two separate constructs, for parallel composition --- by  introducing a mix rule, and for restriction --- by introducing a cycle rule.

### 2020-10-13 conclusion.tex
work on introduction
a pass to related work

@@
 
-\paragraph{On GV}
-
-\paragraph{On deadlock freedom and progress in the $\pi$-calculus}
+\paragraph{On deadlock freedom and progress}
 Deadlock freedom and progress have been studied thoroughly in the $\pi$-calculus with or without session types.
@@
 In the standard typed $\pi$-calculus a foundational line of work is by Kobayashi and his original type-theoretic approach to deadlock-freedom~\cite{kobayashi98}, where priorities are abstract tags defined over a partially ordered set.
-Later on these abstract tags were simplified to natural numbers, and pairs of obligations and capabilities were used in the type system for deadlock freedom~\cite{kobayashi02,kobayashi06}.
+Later on these abstract tags were simplified to natural numbers, and pairs of obligations and capabilities were used in the type system for deadlock or livelock freedom~\cite{kobayashi02,kobayashi06}.
 This allowed more $\pi$-calculus processes to be typed.
-Following, Kobayashi's line of work, Padovani~\cite{padovani13} adapted obligation/capability pairs to session types, and later on he further simplified them to a single priority for linear $\pi$-calculus~\cite{padovani14}.
-As discussed by the author, by using the encoding of session types into linear types~\cite{kobayashi07,dardhagiachino12,dardha14beat,dardha16}, the deadlock freedom technique of the linear $\pi$-calculus with priorities can be transferred onto the $\pi$-calculus with session types.
+Following, Kobayashi's line of work, Padovani~\cite{padovani13} adapted the obligation and capability pairs to session types, and later on he further simplified them to a single priority for linear $\pi$-calculus~\cite{padovani14}.
+As discussed by the author, by using the encoding of session types into linear types~\cite{kobayashi07,dardhagiachino12,dardha14beat,dardha16}, the priority-based technique for deadlock freedom can be transferred onto the $\pi$-calculus with session types.
 
@@
 Carbone \emph{et al.}~\cite{carbonedardha14} studied further the use of catalysers and showed that progress is a compositional form of lock-freedom.
-As in~\cite{padovani14} the authors show that by using the encoding of session types~\cite{dardhagiachino12} and Kobayashi's obligations/capabilities, we can obtain progress in for session types.
-Vieira and Vasconcelos~\cite{vieiravasconcelos13} used single priorities and an abstract partial order to guarantee deadlock freedom in session-typed $\pi$-calculus.
+As in~\cite{padovani14} the authors show that by using the encoding of session types~\cite{dardhagiachino12} and Kobayashi's obligations/capabilities, we can obtain progress for session types.
+Vieira and Vasconcelos~\cite{vieiravasconcelos13} used single priorities and an abstract partial order to guarantee deadlock freedom in a session-typed $\pi$-calculus.
+
+Gay \emph{et al.} \cite{gaynagarajan03} and Vasconcelos \emph{et al.} \cite{vasconcelosravara04,vasconcelosgay06} were the first to introduce a functional language with session types.
+However, such works and including the GV line \cite{gayvasconcelos10,gayvasconcelos12} did not guarantee deadlock freedom.
+This property was later on addressed \cite{lindleymorris15,wadler15} via syntactic restrictions allowing only communication in tree structures, as discussed in the introduction.
+Toninho \emph{et al.} \cite{toninhocaires12} present an encoding of simply-typed $\lambda$-calculus into session-typed $\pi$-calculus investigating the concurrency features of $\lambda$-calculus and providing a new logical explanation.
+Again, their focus is not on deadlock freedom.
 
@@
 With the rise of session types, a new correspondence between linear logic and the $\pi$-calculus was proposed first by Caires and Pfenning~\cite{cairespfenning10}, for dual intuitionistic linear logic (DILL), and later by Wadler~\cite{wadler15} for classical linear logic (CLL).
-The underlying language in both correspondences satisfy deadlock freedom by design, by forbidding cyclic connections via typing.
-Dardha and Gay follow Padovani's work~\cite{padovani14} and use priorities in CP to define a new calculus which is deadlock-free by design, based on classical linear logic~\cite{girard87}  and the Curry-Howard correspondence with session types~\cite{wadler15}.
+The underlying language in both correspondences satisfy deadlock freedom by design, by forbidding cyclic connections via the logical cut rule.
+Dardha and Gay follow Padovani's work~\cite{padovani14} and use priorities in CP to define a new calculus which is deadlock-free by design, based on classical linear logic correspondence with session types~\cite{wadler12}.
 Dardha and P\'{e}rez~\cite{dardhaperez15} compare Kobayashi-style typing and CLL typing, and prove that CLL corresponds to Kobayashi's system with the restriction that only single cuts are allowed.
-Balzer \emph{et al.} study sharing~\cite{balzerpfenning17} and later on deadlock freedom~\cite{balzertoninho19} in session-typed $\pi$-calculus based on DILL.
+Balzer \emph{et al.} study sharing~\cite{balzerpfenning17} and later on deadlock freedom~\cite{balzertoninho19} in a session-typed $\pi$-calculus based on the correspondence with DILL.
 In~\cite{balzerpfenning17} types are either linear or shared with modal operators connecting them, which allows sharing, but not deadlock freedom.
-Later on in~\cite{balzertoninho19} types are enhanced with information regarding the order of acquiring resources, which guarantees deadlock freedom.
+In~\cite{balzertoninho19} types are enhanced with information regarding the order of acquiring resources, which guarantees deadlock freedom.
 
@@
 We also presented an updated version of PCP, where we moved from commuting conversions to structural congruence in the reduction relation, thus remaining faithful to the reduction relation of the $\pi$-calculus.
-Finally, we presented an encoding of Priority CP to Pririty GV in \Cref{fig:pcp-to-pgv}.
-For this translation we proved that it is sound with respect to typing (\Cref{{lem:pcp-to-pgv-terms-preservation}}, \Cref{thm:pcp-to-pgv-confs-preservation}) and operational semantics, given by operational correspondence soundness (\Cref{thm:pcp-to-pgv-operational-correspondence-soundness}) and completeness (\Cref{thm:pcp-to-pgv-operational-correspondence-completeness}).
-
-As future work we intend to define the inverse encoding from Priority GV to Priority CP...
+Finally, we presented an encoding of Priority CP to Pririty GV in \Cref{sec:pcp-to-pgv}.
+For this translation we proved that it is sound with respect to typing (\Cref{thm:pcp-to-pgv-confs-preservation}) and operational semantics, given by operational correspondence soundness (\Cref{thm:pcp-to-pgv-operational-correspondence-soundness}) and completeness (\Cref{thm:pcp-to-pgv-operational-correspondence-completeness}).
+We identify the following main directions for future work: explore recursion in PGV; define the inverse encoding from Priority GV to Priority CP; add sharing to PGV and PCP following the lines of \cite{balzerpfenning17}.
 \end{document}

### 2020-10-13 introduction.tex

@@
 \begin{itemize}
-\item \textbf{Priority GV}: we present Priority GV (PGV) in~\Cref{sec:pgv}; we present the syntax of types and terms, operational semantics and typing rules.
+\item \emph{Priority GV} We present Priority GV (PGV) in~\Cref{sec:pgv}; we present the syntax of types and terms, operational semantics and typing rules.
 We show our type system for PGV is sound by proving subject reduction (\Cref{thm:pgv-subject-reduction-confs}) and closed progress (\Cref{lem:pgv-closed-progress-confs}).
-\item \textbf{Milner's Cyclic Scheduler in PGV}: we present for the first time Milner's Cyclic Scheduler \cite{milner89} (\Cref{sec:pgv-example}) as a GV term.
+\item \emph{Milner's Cyclic Scheduler in PGV} We present, for the first time, Milner's Cyclic Scheduler \cite{milner89} (\Cref{sec:pgv-example}) as a GV term.
 This example illustrates how a cyclic structure can now be expressed in PGV thanks to the separation of channel creation and thread creation, and priorities.
-\item \textbf{Updated Priority CP}: we present an updated version of Priority CP (PCP) \cite{dardhagay18} in \Cref{sec:pcp} where we remove the commuting conversions and move away from a reduction relation as cycle elimination, and towards a reduction as one would expect for a process calculus.
+\item \emph{Updated Priority CP} We present an updated version of Priority CP (PCP) \cite{dardhagay18} in \Cref{sec:pcp} where we remove the commuting conversions and move away from a reduction relation as cycle elimination, and towards a reduction as one would expect for a process calculus.
 We then show Milner's Cyclic Scheduler in PCP (in \Cref{sec:pcp-example}) originally presented in \cite{dardhagay18}.
-\item \textbf{Encoding PCP into PGV}: we present an encoding of PCP into PGV in \Cref{sec:pcp-to-pgv}.
+\item \emph{Encoding PCP into PGV} We present an encoding of PCP into PGV in \Cref{sec:pcp-to-pgv}.
 We prove the encoding is sound and complete with respect to typing (\Cref{thm:pcp-to-pgv-confs-preservation}) and operational semantics (\Cref{thm:pcp-to-pgv-operational-correspondence-soundness} and \Cref{thm:pcp-to-pgv-operational-correspondence-completeness}).

### 2020-10-13 conclusion.tex
just minor...

@@
 
-\paragraph{On deadlock freedom and progress}
+\paragraph{On deadlock freedom and progress in $\pi$-calculus and $\lambda$-calculus}
 Deadlock freedom and progress have been studied thoroughly in the $\pi$-calculus with or without session types.
-
 In the standard typed $\pi$-calculus a foundational line of work is by Kobayashi and his original type-theoretic approach to deadlock-freedom~\cite{kobayashi98}, where priorities are abstract tags defined over a partially ordered set.

### 2020-10-13 priority-gv.tex
Minor update.
Changed float position.
Finished PGV section.

@@
 
-\section{Priority GV}
+\section{Priority GV}\label{sec:pgv}
 \usingnamespace{pgv}
@@
 
-\paragraph{Types}
+\subsubsection*{Types}
 Types in PGV are divided into session types ($\ty{S}$) and types ($\ty{T}$, $\ty{U}$).
@@
 
-\paragraph{Environments}
+\subsubsection*{Environments}
 Typing environments $\ty{\Gamma}$, $\ty{\Delta}$ associate names to types.
@@
 
-\paragraph{Duality}
+\subsubsection*{Duality}
 Duality plays a crucial role in session types.
@@
 
-\paragraph{Priorities}
+\subsubsection*{Priorities}
 The function $\pr$ returns the smallest priority of a session type.
@@
 
-\paragraph{Terms}
+\subsubsection*{Terms}
 Terms ($\tm{L}$, $\tm{M}$, $\tm{N}$) are defined by the following grammar:
@@
 
-\paragraph{Internal and External Choice}
+\subsubsection*{Internal and External Choice}
 Typically, session-typed languages feature constructs for internal and external choice.
@@
 
-\paragraph{Operational Semantics}
+\subsubsection*{Operational Semantics}
 Priority GV terms are evaluated as part of a configuration of processes. Configurations are defined by the following grammar:
@@
 
-\paragraph{Typing}
+\subsubsection*{Typing}
 \Cref{fig:pgv-typing} gives the typing rules for PGV.
@@
 
-\paragraph{Example: Milner's Cyclic Scheduler}\label{sec:pgv-example}
+\subsubsection*{Example: Milner's Cyclic Scheduler}\label{sec:pgv-example}
 We illustrate the newfound ability of PGV by encoding a variant of Milner's cyclic scheduler~\cite{milner89,dardhagay18}.
@@
 
-\subsection{Metatheory}
+\subsubsection*{Subject reduction}
+Unlike with previous versions of GV, The structural congruence, term reduction, and configuration reduction are all type preserving, though it is not entirely trivial to show.
 
+Specifically, we need to show that substitution preserves priority constraints.
+For this, it helps to prove~\cref{lem:pgv-value-done}, which shows that values have finished all their communication actions, and that any priorities present in the type of the value come from the typing environment.
 \begin{restatablelemma}{lempgvvaluedone}
@@
   By induction on the derivation of $\tseq[\cs{p}]{\ty{\Gamma}}{V}{T}$.
-  Details in \cref{prf:lem-pgv-value-done}.
+We demonstrate the case for \LabTirName{T-Inl}, as it illustrates the need for the side conditions on \LabTirName{T-Inl} and \LabTirName{T-Inr}:
+  \begin{case*}[\LabTirName{T-Inl}]
+    The induction hypothesis gives us $\cs{p}=\cs{\pbot}$, and $\pr(\ty{\Gamma})=\pr{(\ty{T})}$. We know $\pr(\ty{T})=\pr({\ty{U}})$, hence $\pr(\ty{\Gamma})=\pr(\ty{\tysum{T}{U}})$.
+    \begin{mathpar}
+      \small
+      \inferrule*{
+        \tseq[\cs{p}]{\ty{\Gamma}}{V}{T}
+        \\
+        \pr(\ty{T})=\pr(\ty{U})
+      }{\tseq[\cs{p}]{\ty{\Gamma}}{\inl{V}}{\tysum{T}{U}}}
+    \end{mathpar}
+  \end{case*}
+  The full proof can be found in~\cref{prf:lem-pgv-value-done}.
 \end{proof}
-
 \begin{restatablelemma}{lempgvsubstitution}[Substitution]
@@
 Whenever the contexts in the premises and conclusion differ, we use \cref{lem:pgv-value-done} to prove that the priorities are preserved.
-  Details in \cref{prf:lem-pgv-substitution}.
+We demonstrate the case for \LabTirName{T-Lam}:
+  \begin{case*}[\LabTirName{T-Lam}]
+    By \cref{lem:pgv-value-done}, $\pr(\ty{\Theta})=\pr(\ty{U'})$, hence $\pr(\ty{\Gamma},\ty{\Theta})=\pr(\ty{\Gamma},\ty{U'})$.
+    \begin{mathpar}
+      \small
+      \inferrule*{
+        \tseq[\cs{q}]{\ty{\Gamma},\tmty{x}{U'},\tmty{y}{T}}{M}{U}
+      }{\tseq[\cs{\pbot}]{\ty{\Gamma},\tmty{x}{U'}}
+        {\lambda y.M}
+        {\tylolli[\cs{\pr(\ty{\Gamma},\ty{U'})},\cs{q}]{T}{U}}}
+      \substarrow{V}{x}
+      \inferrule*{
+        \tseq[\cs{q}]{\ty{\Gamma},\ty{\Theta},\tmty{y}{T}}{\subst{M}{V}{x}}{U}
+      }{\tseq[\cs{\pbot}]{\ty{\Gamma},\ty{\Theta}}
+        {\lambda y.\subst{M}{V}{x}}
+        {\tylolli[\cs{\pr(\ty{\Gamma},\ty{\Theta})},\cs{q}]{T}{U}}}
+    \end{mathpar}
+  \end{case*}
+  The complete proof can be found in~\cref{prf:lem-pgv-substitution}.
 \end{proof}
-
+The proofs for subject congruence and subject reduction are standard.
 \begin{restatablelemma}{lempgvsubjectreductionterms}[Subject Reduction, $\tred$]
@@
 \end{proof}
-
 \begin{restatablelemma}{lempgvsubjectcongruence}[Subject Congruence, $\equiv$]
@@
 \end{proof}
-
 \begin{restatabletheorem}{thmpgvsubjectreductionconfs}[Subject Reduction, $\cred$]
@@
 
+\subsubsection*{Progress and deadlock freedom}
+PGV satisfies progress, as PGV configurations either reduce or are in normal form.
+However, the normal forms may seem surprising at first, as evaluating a well-typed PGV term does not necessarily produce \emph{just} a value.
+If a term returns an endpoint, then its normal form contains a thread which is ready to communicate on the dual of that endpoint.
+For instance, let us consider an example, adapted from Lindley and Morris~\cite{lindleymorris15}, in which a term returns an endpoint linked to an echo server.
+The echo server receives a value and sends it back unchanged:
+\[
+  \tm{\echo_x} \defeq \tm{\letpair{y}{x}{\recv\;{x}}{\letbind{x}{\send\;\pair{y}{x}}{\close\;x}}}
+\]
+Consider the program which creates a new channel, with endpoints $\tm{x}$ and $\tm{x'}$, spawns off an echo server listening on $\tm{x'}$, and then returns $\tm{x}$:
+\[
+  \tm{\main\;\letpair{x}{x'}{\new}{\andthen{\spawn\;(\lambda\unit.\echo_{x'})}}{x}}
+\]
+If we reduce the above program, we get the following normal form.
+Clearly, no more evaluation is possible, even though the configuration contains the thread $\tm{\child\;\echo_{x'}}$, which is blocked on $\tm{x'}$:
+\[
+  \tm{\res{x}{x'}{(\ppar{\main\;x}{\child\;\echo_{x'}})}}
+\]
+In~\cref{cor:pgv-closed-progress}, we show that if a term does not return an endpoint, then it produces \emph{only} a value.
+
+\emph{Actions} are terms which perform communication actions which synchronisation between two threads:
 \begin{definition}[Actions]
@@
 
+\emph{Ready terms} are terms which are ready to perform a communication action, either by themselves, \eg creating a new channel or thread, or with another thread, \eg sending or receiving:
 \begin{definition}[Ready Terms]
@@
 
+Progress for the term language is standard for GV, and deviates from progress for linear $\lambda$-calculus only in that terms may reduce to values or \emph{ready terms}:
 \begin{restatablelemma}{lempgvopenprogressterms}[Open Progress, $\tred$]
@@
 
+Canonical forms deviate from those for GV, in that we opt to move all $\nu$-binders to the top.
+The standard GV normal forms, alternating $\nu$-binders and their corresponding parallel compositions, do not work for PGV, since multiple channels may be split across a single parallel composition:
 \begin{definition}[Canonical Forms]
@@
 \end{definition}
-
 \begin{restatablelemma}{lempgvcanonicalforms}[Canonical Forms]
@@
 
+For simplicity, we only present progress for closed configurations.
+A~suitable notion of progress for open configurations does exist, but its normal forms are significantly more complex, as they may also contain threads which are blocked on free endpoints, threads blocked on endpoints for which the duals are still to be used in those threads blocked on free endpoints, \etc.
+
+The statement of closed progress is rather subtle.
+A~configuration either reduces, or it is equivalent to configuration in the stated normal form.
+Crucial to the normal form is that each term $\tm{M_i}$ is blocked on the corresponding channel $\tm{x_i}$, and hence no two terms act on dual endpoints.
+Furthermore, no term $\tm{M_i}$ can perform a communication action by itself, since those are excluded by the definition of actions:
 \begin{restatabletheorem}{thmpgvclosedprogressconfs}[Closed Progress, $\cred$]
@@
     \tm{\conf{C}}
-    =
+    \equiv
     \tm{\res{x_1}{x'_1}{\dots\res{x_n}{x'_n}{(\child\;M_1\parallel\dots\parallel\child\;M_m\parallel\main\;V)}}},
@@
 
+Finally, as a corollary from closed progress, we get that well-typed terms which do not return endpoints must return \emph{only} a value:
+\begin{corollary}
+  \label{cor:pgv-closed-progress}
+  If $\cseq[\main]{\emptyenv}{\conf{C}}$, $\tm{\conf{C}}\centernot\cred$, and $\tm{\conf{C}}$ contains no endpoints, then $\tm{\conf{C}}\equiv\tm{\phi\;V}$ for some value V.
+\end{corollary}
+
 \end{document}

### 2020-10-14 priority-cp.tex
another pas to intro
Draft version! :D

@@
 
-\section{Priority CP}
+\section{Relation to Priority CP}\label{sec:pcp}
+
+This section presents an updated version of Priority CP~\cite[PCP]{dardhagay18}, which is Wadler's Classical Processes~\cite[CP]{wadler12} with \emph{priorities}, and a translation from PCP to PGV.
+
+The main difference with respect to PCP by Dardha and Gay~\cite{dardhagay18} is the removal of commuting conversions, in favour of using the structural congruence.
+Commuting conversions are necessary if we want our reduction strategy to correspond \emph{exactly} to standard cut elimination.
+However, as Lindley and Morris~\cite{lindleymorris15} show, all $\beta$-reductions that can be performed \emph{with} the use of commuting conversions, can also be performed \emph{without} them.
+They show this by decomposing the reduction relation for CP into two phases: first, all $\beta$-reductions are performed, and second, commuting conversions are used to bring one particular action to the front, to eliminate the top-level cut.
+
+From the perspective of process calculus, commuting conversions behave strangely. Let us consider the commuting conversion for $\pcp{\tm{\recv{x}{y}{P}}}$:
+\[
   \usingnamespace{pcp}
+  (\kappa_{\parr})
+  \quad
+  \tm{\res{z}{z'}{(\ppar{\recv{x}{y}{P}}{Q})}}
+  \red
+  \tm{\recv{x}{y}{\res{z}{z'}{(\ppar{P}{Q})}}}
+\]
+As a result of $(\kappa_{\parr})$, $\pcp{\tm{Q}}$ becomes blocked on $\pcp{\tm{\labrecv{x}{y}}}$, and any actions $\pcp{\tm{Q}}$ was able to perform become unavailable.
+Consequently, CP is non-confluent, for instance:
+\begin{mathpar}
+  \usingnamespace{pcp}
+  \setlength{\arraycolsep}{2em}
+  \begin{array}{cc}
+    \multicolumn{2}{c}{%
+    \hspace*{10ex}
+    {\tm{\res{x}{x'}{(\ppar{\recv{a}{y}{P}}{\res{z}{z'}{(\ppar{\close{z}{\halt}}{\wait{z'}{Q}})}})}}}}
+    \\
+    \qquad\rotatebox[origin=c]{215}{$\red$}
+    &
+    \rotatebox[origin=c]{325}{$\red$}\qquad
+    \\
+    {\tm{\recv{a}{y}{\res{x}{x'}{(\ppar{P}{\res{z}{z'}{(\ppar{\close{z}{\halt}}{\wait{z'}{Q}})}})}}}}
+    &
+    {\tm{\res{x}{x'}{(\ppar{\recv{a}{y}{P}}{Q})}}}
+    \\ 
+    &
+    \rotatebox[origin=c]{270}{$\red$}
+    \\
+    &
+    \tm{\recv{a}{y}{\res{x}{x'}{(\ppar{P}{Q})}}}
+  \end{array}
+\end{mathpar}
+Commuting conversions break our intuition for priorities, since ``if a prefix on a channel endpoint $\pcp{\tm{x}}$ with priority $\cs{o}$ is pulled out at top level, then to preserve priority conditions in the typing rules [..], it is necessary to increase priorities of all actions after the prefix on $\pcp{\tm{x}}$'' by $\cs{o+1}$~\cite{dardhagay18}.
+Consequently, the intuition that an action with lower priority occurs before an action with higher priority is shattered.
 
-We present an updated version of Priority CP~\cite[PCP]{dardhagay18}, which is Wadler's Classical Processes~\cite[CP]{wadler12} with \emph{priorities}.
-The main difference with respect to PCP by Dardha and Gay \cite{dardhagay18} is moving away from a reduction strategy based on cycle elimination using commuting conversions, and towards a reduction strategy using structural congruence.
-
-\paragraph*{Priorities, types, and environments}
-Types are based on classical linear logic propositions, and are annotated with priorities ranging over $\cs{o}\in\mathbb{N}$.
-We let $\ty{A}, \ty{B}, \ty{C}$ range over types, produced by the following grammar.
+\subsubsection*{Types}
+Types $\pcp{\ty{A}}$, $\pcp{\ty{B}}$, $\pcp{\ty{C}}$ in PCP are based on classical linear logic propositions, and are defined by the following grammar:
 \[
+  \usingnamespace{pcp}
   \begin{array}{lcl}
@@
 \]
-The types $\ty{\tytens[\cs{o}]{A}{B}}$ and $\ty{\typarr[\cs{o}]{A}{B}}$ type the endpoints of a channel over which we send or receive a channel of type $\ty{A}$, and then proceed as type $\ty{B}$.
-The types $\ty{\tyone[\cs{o}]}$ and $\ty{\tybot[\cs{o}]}$ type the endpoints of a channel whose session has terminated, and over which we send or receive a \emph{ping} before closing the channel.
-These two types act as units for $\ty{\tytens[\cs{o}]{A}{B}}$ and $\ty{\typarr[\cs{o}]{A}{B}}$, respectively.
+Each connective is annotated with a priority $\cs{o}\in\mathbb{N}$.
+The types $\pcp{\ty{\tytens[\cs{o}]{A}{B}}}$ and $\pcp{\ty{\typarr[\cs{o}]{A}{B}}}$ type the endpoints of a channel over which we send or receive a channel of type $\pcp{\ty{A}}$, and then proceed as type $\pcp{\ty{B}}$.
+The types $\pcp{\ty{\tyone[\cs{o}]}}$ and $\pcp{\ty{{\tybot}[\cs{o}]}}$ type the endpoints of a channel whose session has terminated, and over which we send or receive a \emph{ping} before closing the channel.
+These two types act as units for $\pcp{\ty{\tytens[\cs{o}]{A}{B}}}$ and $\pcp{\ty{\typarr[\cs{o}]{A}{B}}}$, respectively.
 
-The types $\ty{\typlus[\cs{o}]{A}{B}}$ and $\ty{\tywith[\cs{o}]{A}{B}}$ type the endpoints of a channel over which we can receive or send a choice between two branches $\ty{A}$ or $\ty{B}$.
-We have opted for a simplified version of choice and followed the original Wadler's CP \cite{wadler12}, however types $\oplus$ and $\with$ can be trivially generalised to $\ty{\oplus^{\cs{o}}\{l_i:A_i\}_{i\in I}}$ and $\ty{\with^{\cs{o}}\{l_i:A_i\}_{i\in I}}$, respectively as in the original PCP \cite{dardhagay18}.
-The types $\ty{\tynil[\cs{o}]}$ and $\ty{\tytop[\cs{o}]}$ type the endpoints of a channel over which we can send or receive a choice between \emph{no options}.
-These two types act as units for $\ty{\typlus[\cs{o}]{A}{B}}$ and $\ty{\tywith[\cs{o}]{A}{B}}$, respectively.
+The types $\pcp{\ty{\typlus[\cs{o}]{A}{B}}}$ and $\pcp{\ty{\tywith[\cs{o}]{A}{B}}}$ type the endpoints of a channel over which we can receive or send a choice between two branches $\pcp{\ty{A}}$ or $\pcp{\ty{B}}$.
+We have opted for a simplified version of choice and followed the original Wadler's CP \cite{wadler12}, however types $\ty{\oplus}$ and $\ty{\with}$ can be trivially generalised to $\pcp{\ty{\oplus^{\cs{o}}\{l_i:A_i\}_{i\in I}}}$ and $\pcp{\ty{\with^{\cs{o}}\{l_i:A_i\}_{i\in I}}}$, respectively, as in the original PCP \cite{dardhagay18}.
+The types $\pcp{\ty{\tynil[\cs{o}]}}$ and $\pcp{\ty{\tytop[\cs{o}]}}$ type the endpoints of a channel over which we can send or receive a choice between \emph{no options}.
+These two types act as units for $\pcp{\ty{\typlus[\cs{o}]{A}{B}}}$ and $\pcp{\ty{\tywith[\cs{o}]{A}{B}}}$, respectively.
 
-A typing context $\ty\Gamma$ is a partial function associating channel names to types.
+\subsubsection*{Environments}
+Typing environments $\pcp{\ty{\Gamma}}$, $\pcp{\ty{\Delta}}$ associate names to types.
+Environments are linear, so two environments can only be combined as $\pcp{\ty{\Gamma}}, \pcp{\ty{\Delta}}$ if their names are distinct, \ie $\pcp{\fv(\ty{\Gamma})\cap\fv(\ty{\Delta})=\varnothing}$.
 \[
+  \usingnamespace{pcp}
   \begin{array}{lcl}
@@
 \]
-Duality on types is an involutive function which preserves the priorities of types.
+
+\subsubsection*{Duality}
+Duality on types is an involutive function which preserves the priorities of types:
 \[
+  \usingnamespace{pcp}
   \begin{array}{lcl}
@@
 \]
-The function $\pr(\cdot)$ returns the priority of a type. The function $\minpr(\cdot)$ returns the \emph{minimum} priority of all types a typing context.
+
+\subsubsection*{Priorities}
+The function $\pr(\cdot)$ returns smallest priority of a type.
+As with PGV, the type system guarantees that the top-most connective always holds the smallest priority.
+The function $\minpr(\cdot)$ returns the \emph{minimum} priority of all types a typing context, or $\cs{\ptop}$ if the context is empty:
 \[
+  \usingnamespace{pcp}
   \begin{array}{lclclcl}
@@
     \\
-  \minpr(\ty{\emptyenv})          & = & \varnothing
+    \minpr(\ty{\emptyenv})          & = & \cs{\ptop}
   \end{array}
@@
 
-\paragraph*{Terms}
-We let $\tm{P}, \tm{Q}$ range over processes, produced by the following grammar.
+\subsubsection*{Terms}
+We let $\pcp{\tm{P}}$, $\pcp{\tm{Q}}$ range over processes, produced by the following grammar.
 \[
+  \usingnamespace{pcp}
   \begin{array}[t]{lcl}
@@
 \]
-The process $\tm{\link{x}{y}}$ links endpoints $x$ and $y$ and forwards communication from one to the other.
-$\tm{\res{x}{y}{P}}$, $\tm{(\ppar{P}{Q})}$ and $\tm{\halt}$ denote respectively the restriction processes where channel endpoints $\tm{x}$ and $\tm{y}$ are bound together and with scope $\tm{P}$, the parallel composition of processes $\tm{P}$ and $\tm{Q}$ and the terminated process.
+The process $\pcp{\tm{\link{x}{y}}}$ links endpoints $\pcp{\tm{x}}$ and $\pcp{\tm{y}}$ and forwards communication from one to the other.
+$\pcp{\tm{\res{x}{y}{P}}}$, $\pcp{\tm{(\ppar{P}{Q})}}$ and $\pcp{\tm{\halt}}$ denote respectively the restriction processes where channel endpoints $\pcp{\tm{x}}$ and $\pcp{\tm{y}}$ are bound together and with scope $\pcp{\tm{P}}$, the parallel composition of processes $\pcp{\tm{P}}$ and $\pcp{\tm{Q}}$ and the terminated process.
 
-Processes $\tm{\send{x}{y}{P}}$ and $\tm{\recv{x}{y}{P}}$ send or receive over channel $\tm{x}$ a value $\tm{y}$ and proceed as process $\tm{P}$.
-$\tm{\close{x}{P}}$ and $\tm{\wait{x}{P}}$ send and receive an empty value -- denoting the closure of channel $\tm{x}$, and continue as $\tm{P}$.
+Processes $\pcp{\tm{\send{x}{y}{P}}}$ and $\pcp{\tm{\recv{x}{y}{P}}}$ send or receive over channel $\pcp{\tm{x}}$ a value $\pcp{\tm{y}}$ and proceed as process $\pcp{\tm{P}}$.
+$\pcp{\tm{\close{x}{P}}}$ and $\pcp{\tm{\wait{x}{P}}}$ send and receive an empty value -- denoting the closure of channel $\pcp{\tm{x}}$, and continue as $\pcp{\tm{P}}$.
 
-Processes $\tm{\inl{x}{P}}$ and $\tm{\inr{x}{P}}$ make a left and right choice, respectively and proceed as process $\tm{P}$.
-$\tm{\offer{x}{P}{Q}}$ dually offers both left and right branches, with continuations $\tm{P}$ and $\tm{Q}$.
-$\tm{\absurd{x}}$ denotes the empty offer.
+Processes $\pcp{\tm{\inl{x}{P}}}$ and $\pcp{\tm{\inr{x}{P}}}$ make a left and right choice, respectively and proceed as process $\pcp{\tm{P}}$.
+$\pcp{\tm{\offer{x}{P}{Q}}}$ dually offers both left and right branches, with continuations $\pcp{\tm{P}}$ and $\pcp{\tm{Q}}$.
+$\pcp{\tm{\absurd{x}}}$ denotes the empty offer.
 
-We write \emph{unbound} send as $\tm{\usend{x}{y}{P}}$, which is syntactic sugar for $\tm{\send{x}{z}{(\ppar{\link{y}{z}}{P})}}$.
-Alternatively, we could take $\tm{\usend{x}{y}{P}}$ as primitive, and let $\tm{\send{x}{y}{P}}$ be syntactic sugar for $\tm{\res{y}{z}{(\usend{x}{z}{P})}}$.
-CP takes \emph{bound} sending as primitive, as it is impossible to eliminate the top-level cut in terms such as $\tm{\res{y}{z}{(\usend{x}{z}{P})}}$, even with communting conversions.
-However, in our setting without commuting conversions and with more permissive normal forms, this is no longer an issue.
+We write \emph{unbound} send as $\pcp{\tm{\usend{x}{y}{P}}}$, which is syntactic sugar for $\pcp{\tm{\send{x}{z}{(\ppar{\link{y}{z}}{P})}}}$.
+Alternatively, we could take $\pcp{\tm{\usend{x}{y}{P}}}$ as primitive, and let $\pcp{\tm{\send{x}{y}{P}}}$ be syntactic sugar for $\pcp{\tm{\res{y}{z}{(\usend{x}{z}{P})}}}$.
+CP takes \emph{bound} sending as primitive, as it is impossible to eliminate the top-level cut in terms such as $\pcp{\tm{\res{y}{z}{(\usend{x}{z}{P})}}}$, even with communting conversions.
+In our setting without commuting conversions and with more permissive normal forms, this is no longer an issue, but, for simplicity, we keep bound sending as primitive.
 
-\paragraph{Typing}
+\subsubsection*{Operational Semantics}
+The operational semantics for PCP, given in \cref{fig:pcp-operational-semantics}, is defined as a reduction relation $\pcp{\red}$ on processes (bottom) and uses structural congruence (top).
+Each of the axioms of structural congruence corresponds to the axiom of the same name for PGV.
+
+The reduction relation is given by a set of axioms and inference rules for context closure.
+Reduction occurs under restriction.
+$\LabTirName{E-Link}$ reduces a parallel composition with a link into a substitution.
+$\LabTirName{E-Send}$ is the main communication rule, where send and receive processes sychronise and reduce to the corresponding continuations.
+$\LabTirName{E-Close}$ follows the previous rule and it closes the channel identified by endpoints $\pcp{\tm{x}}$ and $\pcp{\tm{y}}$.
+$\LabTirName{E-Select-Inl}$ and $\LabTirName{E-Select-Inr}$ are generalised versions of $\LabTirName{E-Send}$.
+They state respectively that a left and right selection synchronises with a choice offering and reduces to the corresponding continuations.
+The last three rules state that reduction is closed under restriction, parallel composition and structural congruence, respectively.
+
+\input{fig/pcp-operational-semantics}
+
+\subsubsection*{Typing}
+\Cref{fig:pcp-typing} gives the typing rules for our version of PCP.
+A typing judgement $\pcp{\seq{P}{\Gamma}}$ states that ``process $\pcp{\tm{P}}$ is well typed under the typing context $\pcp{\ty{\Gamma}}$''.
+
+Typing rules make use of priorities, defined in the previous section.
+\textsc{T-Link} states that the link process $\pcp{\tm{\link{x}{y}}}$ is well typed under channels $\pcp{\tm{x}}$ and $\pcp{\tm{y}}$ having dual types, respectively $\pcp{\ty{A}}$ and $\pcp{\ty{\co{A}}}$.
+\textsc{T-Res} sates that the restriction process $\pcp{\tm{\res{x}{y}{P}}}$ is well typed under typing context $\pcp{\ty{\Gamma}}$ if process $\pcp{\tm{P}}$ is well typed in $\pcp{\ty{\Gamma}}$ augmented with channel endpoints $\pcp{\tm{x}}$ and $\pcp{\tm{y}}$ having dual types, respectively $\pcp{\ty{A}}$ and $\pcp{\ty{\co{A}}}$.
+\textsc{T-Par} states that the parallel composition of processes $\pcp{\tm{P}}$ and $\pcp{\tm{Q}}$ is well typed under the disjoint union of their respective typing contexts.
+\textsc{T-Halt} states that the terminated process $\pcp{\tm{\halt}}$ is well typed in the empty context.
+
+\textsc{T-Send} and \textsc{T-Recv} state that the sending and receiving of a bound name $\pcp{\tm{y}}$ over a channel $\pcp{\tm{x}}$ is well typed under $\pcp{\ty{\Gamma}}$ and $\pcp{\tm{x}}$ of type $\pcp{\ty{\tytens[\cs{o}]{A}{B}}}$, respectively $\pcp{\typarr[\cs{o}]{A}{B}}$.
+Priority $\cs{o}$ is the smallest among all priorities of the types used by the output or input process, captured by the side condition $\pcp{\cs{o}<\minpr(\ty{\Gamma},\ty{A},\ty{B})}$.
+
+Rules \textsc{T-Close} and \textsc{T-Wait} type the closure of channel $\pcp{\tm{x}}$ and are in the same lines as the previous two rules, requiring that the priority of channel $\pcp{\tm{x}}$ is the smallest among all priorities in $\pcp{\ty{\Gamma}}$.
+
+\textsc{T-Select-Inl} and \textsc{T-Select-Inr} type respectively the left $\pcp{\tm{\inl{x}{P}}}$ and right $\pcp{\tm{\inr{x}{P}}}$ choice performed on channel $\pcp{\tm{x}}$.
+\textsc{T-Offer} and \textsc{T-Offer-Absurd} type the offering of a choice, or empty choice, on channel $\pcp{\tm{x}}$.
+In all the above rules the priority $\cs{o}$ of channel $\pcp{\tm{x}}$ is the smallest with respect to the typing context $\pcp{\cs{o}<\minpr(\ty{\Gamma})}$ and types involved in the choice $\pcp{\cs{o}<\minpr(\ty{\Gamma},\ty{A},\ty{B})}$.
+
 \input{fig/pcp-typing}
 
-Typing rules for our version of PCP are given in \Cref{fig:pcp-typing}. A typing judgement $\seq{P}{\Gamma}$ states that "process $\tm{P}$ is well typed under the typing context $\ty \Gamma$".
+\subsubsection*{Subject reduction}
+Reduction in our version of PCP preserve types. Since our reduction relation is a strict subset of the reduction relation in the original~\cite{dardhagay18}, we defer to their proofs:
+\begin{restatablelemma}{lempcpsubjectcongruence}[Subject Congruence, $\pcp{\equiv}$]
+  \label{lem:pcp-subject-congruence}
+  \hfill\\%newline before theorem statement
+  If $\pcp{\seq{P}{\ty{\Gamma}}}$ and $\pcp{\tm{P}\equiv\tm{Q}}$,
+  then $\pcp{\seq{Q}{\ty{\Gamma}}}$.
+\end{restatablelemma}
+\begin{restatabletheorem}{thmpcpsubjectreduction}[Subject Reduction, $\pcp{\red}$]
+  \label{thm:pcp-subject-reduction}
+  \hfill\\%newline before theorem statement
+  If $\pcp{\seq{P}{\ty{\Gamma}}}$ and $\pcp{\tm{P}\red\tm{Q}}$, then $\pcp{\seq{Q}{\ty{\Gamma}}}$.
+\end{restatabletheorem}
 
-Typing rules make use of priorities, defined in the previous section.
-Priorities are based on \emph{obligations/capabilities} used by Kobayashi~\cite{kobayashi06}, and simplified to single priorities following Padovani~\cite{padovani14}.
-Recalling PCP \cite{dardhagay18}, priorities obey the following two laws:
-\begin{enumerate}
-  \item [(i)] an action of priority $\cs o$ must be prefixed only by actions of priorities \emph{strictly smaller} than $\cs o$.
-  \item [(ii)] communication requires \emph{equal} priorities for the dual actions.
-\end{enumerate}
-\textsc{T-Link} states that the link process $\tm{\link{x}{y}}$ is well typed under channels $\tm{x}$ and $\tm{y}$ having dual types, respectively $\ty A$ and $\ty {\co A}$.
-\textsc{T-Res} sates that the restriction process $\tm{\res{x}{y}{P}}$ is well typed under typing context $\ty\Gamma$ if process $\tm{P}$ is well typed in $\ty\Gamma$ augmented with channel endpoints $\tm{x}$ and $\tm{y}$ having dual types, respectively $\ty A$ and $\ty {\co A}$.
-\textsc{T-Par} states that the parallel composition of processes $\tm{P}$ and $\tm{Q}$ is well typed under the disjoint union of their respective typing contexts.
-\textsc{T-Halt} states that the terminated process $\tm{\halt}$ is well typed in the empty context.
+\subsubsection*{Progress}
+Reduction in our version of PCP satisfies progress. For simplicity, we only present closed progress:
+\begin{restatabletheorem}{thmpcpclosedprogress}[Closed Progress, $\pcp{\red}$]
+  \label{thm:pcp-closed-progress}
+  \hfill\\%newline before theorem statement
+  If $\pcp{\seq{P}{\emptyenv}}$, then either $\pcp{\tm{P}=\tm{\halt}}$ or there exists a $\pcp{\tm{Q}}$ such that $\pcp{\tm{P}\red\tm{Q}}$.
+\end{restatabletheorem}
+\begin{proof}
+    Our proof follows the same reasoning by Kobayashi~\cite{kobayashi06} used in the proof of deadlock freedom for closed processes (Theorem 2).
+The details can be found in~\Cref{prf:thm-pcp-closed-progress}.
+\end{proof}
 
-\textsc{T-Send} and \textsc{T-Recv} state that the sending and receiving of a bound name $\tm{y}$ over a channel $\tm{x}$ is well typed under $\ty\Gamma$ and $\tm{x}$ of type $\tytens[\cs{o}]{A}{B}$, respectively ${\typarr[\cs{o}]{A}{B}}$.
-Priority $\cs o$ is the smallest among all priorities of the types used by the output or input process, captured by the side condition $\cs{o}<\minpr(\ty{\Gamma},\ty{A},\ty{B})$.
+\subsubsection*{Translation}
+The translation on types from PCP to PGV, $\ty{\cpgvT{\cdot}}$, is:
+\[
+  \begin{array}{lcl}
+    \ty{\cpgvT{\pcp{\tytens[\cs{o}]{A}{B}}}}
+    & = & \pgv{\ty{\tysend[\cs{o}]{\co{\cpgvT{A}}}{\cpgvT{B}}}}
+    \\
+    \ty{\cpgvT{\pcp{\typlus[\cs{o}]{A}{B}}}}
+    & = & \pgv{\ty{\tyselect[\cs{o}]{\cpgvT{A}}{\cpgvT{B}}}}
+  \end{array}
+  \;
+  \begin{array}{lcl}
+    \ty{\cpgvT{\pcp{\tyone[\cs{o}]}}}
+    & = & \ty{\pgv{\tyends[\cs{o}]}}
+    \\
+    \ty{\cpgvT{\pcp{\tynil[\cs{o}]}}}
+    & = & \pgv{\ty{\tyselectemp[\cs{o}]}}
+  \end{array}
+  \;
+  \begin{array}{lcl}
+    \ty{\cpgvT{\pcp{\typarr[\cs{o}]{A}{B}}}}
+    & = & \pgv{\ty{\tyrecv[\cs{o}]{\cpgvT{A}}{\cpgvT{B}}}}
+    \\
+    \ty{\cpgvT{\pcp{\tywith[\cs{o}]{A}{B}}}}
+    & = & \pgv{\ty{\tyoffer[\cs{o}]{\cpgvT{A}}{\cpgvT{B}}}}
+  \end{array}
+  \;
+  \begin{array}{lcl}
+    \ty{\cpgvT{\pcp{\tybot[\cs{o}]}}}
+    & = & \ty{\pgv{\tyendr[\cs{o}]}}
+    \\
+    \ty{\cpgvT{\pcp{\tytop[\cs{o}]}}}
+    & = & \pgv{\ty{\tyofferemp[\cs{o}]}}
+  \end{array}
+\]
+There are two separate translations on processes.
+The first, $\tm{\cpgvC{\cdot}}$, translates processes to \emph{configurations}.
+It translates as much as possible of the process to the corresponding configuration constructs.
+Once it hits the first action, it switches to the second translation, $\tm{\cpgvM{\cdot}}$, which translates processes to \emph{terms}:
+\[
+  \begin{array}{lcl}
+    \pcp{\tm{\cpgvC{\res{x}{y}{P}}}}
+    & = & \pgv{\tm{\res{x}{y}{\cpgvC{P}}}} \\
+    \pcp{\tm{\cpgvC{\ppar{P}{Q}}}}
+    & = & \pgv{\tm{\ppar{\cpgvC{P}}{\cpgvC{Q}}}} \\
+    \pcp{\tm{\cpgvC{\send{x}{y}{P}}}}
+    & = & \pgv{\tm{\res{y}{z}{(\child\;\letbind{x}{\send\;\pair{z}{x}}{\cpgvM{P}})}}} \\
+    \pcp{\tm{\cpgvC{\inl{x}{P}}}}
+    & = & \pgv{\tm{\res{y}{z}{(\child\;\letbind{x}{\andthen{\close\;(\send\;\pair{\inl{y}}{x})}{z}}{\cpgvM{P}})}}} \\
+    \pcp{\tm{\cpgvC{\inr{x}{P}}}}
+    & = & \pgv{\tm{\res{y}{z}{(\child\;\letbind{x}{\andthen{\close\;(\send\;\pair{\inr{y}}{x})}{z}}{\cpgvM{P}})}}} \\
+    \pcp{\tm{\cpgvC{P}}}
+    & = & \pgv{\tm{\child{\cpgvM{P}}}},\hfill\text{if none of the above apply}\\
+    \\
+    \pcp{\tm{\cpgvM{\link{x}{y}}}}
+    & = & \pgv{\tm{\link\;{\pair{x}{y}}}} \\
+    \pcp{\tm{\cpgvM{\res{x}{y}{P}}}}
+    & = & \pgv{\tm{\letpair{x}{y}{\new}{\cpgvM{P}}}} \\
+    \pcp{\tm{\cpgvM{\ppar{P}{Q}}}}
+    & = & \pgv{\tm{\andthen{\spawn\;{(\lambda\unit.\cpgvM{P})}}{\cpgvM{Q}}}} \\
+    \pcp{\tm{\cpgvM{\halt}}}
+    & = & \pgv{\tm{\unit}} \\
+    \pcp{\tm{\cpgvM{\close{x}{P}}}}
+    & = & \pgv{\tm{\andthen{\close\;{x}}{\cpgvM{P}}}} \\
+    \pcp{\tm{\cpgvM{\wait{x}{P}}}}
+    & = & \pgv{\tm{\andthen{\wait\;{x}}{\cpgvM{P}}}} \\
+    \pcp{\tm{\cpgvM{\send{x}{y}{P}}}}
+    & = & \pgv{\tm{\letpair{y}{z}{\new}{\letbind{x}{\send\;{\pair{z}{x}}}{\cpgvM{P}}}}} \\
+    \pcp{\tm{\cpgvM{\recv{x}{y}{P}}}}
+    & = & \pgv{\tm{\letpair{y}{x}{\recv\;{x}}{\cpgvM{P}}}} \\
+    \pcp{\tm{\cpgvM{\inl{x}{P}}}}
+    & = & \pgv{\tm{\letbind{x}{\select{\labinl}\;{x}}{\cpgvM{P}}}} \\
+    \pcp{\tm{\cpgvM{\inr{x}{P}}}}
+    & = & \pgv{\tm{\letbind{x}{\select{\labinr}\;{x}}{\cpgvM{P}}}} \\
+    \multicolumn{3}{l}{\pcp{\tm{\cpgvM{\offer{x}{\cpgvM{P}}{\cpgvM{Q}}}}}} \\
+    & = & \pgv{\tm{\offer{x}{x}{\cpgvM{P}}{x}{\cpgvM{Q}}}} \\
+    \pcp{\tm{\cpgvM{\absurd{x}}}}
+    & = & \pgv{\tm{\offeremp{x}}}
+  \end{array}
+\]
+The third, fourth, and fifth cases for $\tm{\cpgvC{\cdot}}$ translate the $\nu$-binders implicit in the encoding of bound sending to configuration constructs, which is required for completeness of the operational correspondence.
+Typing environments are translated pointwise, and sequents $\pcp{\seq{P}{\ty{\Gamma}}}$ are translated as $\pgv{\cseq[\child]{\ty{\cpgvT{\ty{\Gamma}}}}{\cpgvC{P}}}$.
 
-Rules \textsc{T-Close} and \textsc{T-Wait} type the closure of channel $\tm{x}$ and are in the same lines as the previous two rules, requiring that the priority of channel $\tm{x}$ is the smallest among all priorities in $\tm\Gamma$.
-
-\textsc{T-Select-Inl} and \textsc{T-Select-Inr} type respectively the left $\tm{\inl{x}{P}}$ and right $\tm{\inr{x}{P}}$ choice performed on channel $\tm{x}$.
-\textsc{T-Offer} and \textsc{T-Offer-Absurd} type the offering of a choice, or empty choice, on channel $\tm{x}$.
-In all the above rules the priority $\cs o$ of channel $\tm{x}$ is the smallest with respect to the typing context $\cs{o}<\minpr(\ty{\Gamma})$ and types involved in the choice $\cs{o}<\minpr(\ty{\Gamma},\ty{A},\ty{B})$.
-
-\paragraph{Example: Milner's Cyclic Scheduler}
+\subsubsection*{Example: Milner's Cyclic Scheduler}\label{sec:pcp-example}
+We illustrate our translation by giving the PCP version of the cyclic scheduler in~\cref{sec:pgv-example}.
+We implement the scheduler as follows, using $\pcp{\tm{\prod_{i}P_i}}$ to denote the parallel composition of the processes $\pcp{\tm{P_i}}$, and $\pcp{\plug{P}{Q}}$ to denote the plugging of $\pcp{\tm{Q}}$ in the one-hole process context $\pcp{\tm{P}}$:
 \[
+  \usingnamespace{pcp}
   \begin{array}{lrlr}
@@
 \]
+The translation of $\pcp{\sched}$ is exactly $\pgv{sched}$ where all processes are child processes.
 
-\paragraph{Operational Semantics}
-\input{fig/pcp-operational-semantics}
-
-The operational semantics for PCP given in \Cref{fig:pcp-operational-semantics} is defined as a reduction relation $\red$ on processes (bottom) and uses structural congruence (top).
-
-Structural congruence satisfies the following axioms.
-$\LabTirName{SC-LinkSwap}$ allows swapping channels in the link process.
-$\LabTirName{SC-ResLink}$ allows restriction to applied to link which is structurally equivalent to the terminated process, thus allowing elimination of unnecessary restrictions.
-$\LabTirName{SC-ResSwap}$ allows swapping channels and $\LabTirName{SC-ResComm}$ states that restriction is commutative.
-$\LabTirName{SC-ResExt}$ is the standard scope extrusion rule.
-Rules $\LabTirName{SC-ParNil}$, $\LabTirName{SC-ParComm}$ and $\LabTirName{SC-ParAssoc}$ state that parallel composition uses the terminated proceess as the neutral element; it is commutative and associative.
-
-The reduction relation is given by a set of axioms and inference rules for context closure. Reduction occurs under restriction following \cite{dardhagay18} where channel endpoints are identified.
-$\LabTirName{E-Link}$ reduces a parallel composition with a link into a substitution.
-$\LabTirName{E-Send}$ is the main communication rule, where send and receive processes sychronise and reduce to the corresponding continuations.
-$\LabTirName{E-Close}$ follows the previous rule and it closes the channel identified by endpoints $\tm{x}$ and $\tm{y}$.
-$\LabTirName{E-Select-Inl}$ and $\LabTirName{E-Select-Inr}$ are generalised versions of $\LabTirName{E-Send}$.
-They state respectively that a left and right selection synchronises with a choice offering and reduces to the corresponding continuations.
-The last three rules state that reduction is closed under restriction, parallel composition and structural congruence, respectively.
-
-\paragraph{Metatheory}
-
-\begin{restatablelemma}{lempcpsubjectcongruence}[Subject Congruence, $\equiv$]
-  \label{lem:pcp-subject-congruence}
+\subsubsection*{Type preservation}
+The translations $\tm{\cpgvC{\cdot}}$ and $\tm{\cpgvM{\cdot}}$ preserve typing along $\ty{\cpgvT{\cdot}}$:
+\begin{restatablelemma}{lempcptopgvtermspreservation}[Preservation, $\cpgvM{\cdot}$]
+  \label{lem:pcp-to-pgv-terms-preservation}
   \hfill\\%newline before theorem statement
-  If $\seq{P}{\ty{\Gamma}}$ and $\tm{P}\equiv\tm{Q}$,
-  then $\seq{Q}{\ty{\Gamma}}$.
+  If $\pcp{\seq{P}{\ty{\Gamma}}}$ in PCP,
+  then $\pgv{\tseq[\cs{p}]{\ty{\cpgvT{\Gamma}}}{\cpgvM{P}}{\tyunit}}$ in PGV.
 \end{restatablelemma}
 \begin{proof}
-  See~\cite{dardhagay18}.
+  By induction on the derivation of $\pcp{\seq{P}{\ty{\Gamma}}}$.
+  Details in \cref{prf:lem-pcp-to-pgv-terms-preservation}.
+\end{proof}
+\begin{restatabletheorem}{thmpcptopgvconfspreservation}[Preservation, $\cpgvC{\cdot}$]
+  \label{thm:pcp-to-pgv-confs-preservation}
+  \hfill\\%newline before theorem statement
+  If $\pcp{\seq{P}{\ty{\Gamma}}}$ in PCP,
+  then $\pgv{\cseq[\child]{\ty{\cpgvT{\Gamma}}}{\cpgvC{P}}}$ in PGV.
+\end{restatabletheorem}
+\begin{proof}
+  By induction on the derivation of $\pcp{\seq{P}{\ty{\Gamma}}}$.
+  Details in \cref{prf:lem-pcp-to-pgv-terms-preservation}.
 \end{proof}
 
-\begin{restatabletheorem}{thmpcpsubjectreduction}[Subject Reduction, $\red$]
-  \label{thm:pcp-subject-reduction}
+\subsubsection*{Operational correspondence}
+Finally, we show that PCP and PGV are in operational correspondence along our translation.
+First, we link our two translations, and show that $\tm{\cpgvM{P}}$ \emph{must} reduce to $\tm{\cpgvC{P}}$, \ie $\tm{\cpgvC{\cdot}}$ is equivalent to translating with $\tm{\cpgvM{\cdot}}$ and performing several reduction steps:
+\begin{restatablelemma}{lempcptopgvcpgvMtocpgvC}
+  \label{lem:pcp-to-pgv-cpgvM-to-cpgvC}
+  For any $\pcp{\tm{P}}$, either:
+  \begin{itemize}
+  \item $\pgv{\tm{\child\;\cpgvM{P}}=\tm{\cpgvC{P}}}$; or
+  \item   $\pgv{\tm{\child\;\cpgvM{P}}\cred^+\tm{\cpgvC{P}}}$, and for any $\pgv{\tm{\conf{C}}}$, if $\pgv{\tm{\child\;\cpgvM{P}}\cred\tm{\conf{C}}}$, then $\pgv{\tm{\conf{C}}\cred^\star\tm{\cpgvC{P}}}$.
+  \end{itemize}
+\end{restatablelemma}
+\begin{proof}
+  By induction on the structure of $\pcp{\tm{P}}$.
+  Details in \cref{prf:lem-pcp-to-pgv-cpgvM-to-cpgvC}.
+\end{proof}
+
+\begin{restatabletheorem}{thmpcptopgvoperationalcorrespondencesoundness}%
+  [Operational Correspondence, Soundness, $\cpgvC{\cdot}$]
+  \label{thm:pcp-to-pgv-operational-correspondence-soundness}
   \hfill\\%newline before theorem statement
-  If $\seq{P}{\ty{\Gamma}}$ and $\tm{P}\red\tm{Q}$, then $\seq{Q}{\ty{\Gamma}}$.
+  If $\pcp{\seq{P}{\ty{\Gamma}}}$ and $\pgv{\tm{\cpgvC{P}}\cred\tm{\conf{C}}}$,
+  there exists a $\tm{Q}$ such that $\pcp{\tm{P}\red^+\tm{Q}}$ and $\pgv{\tm{\conf{C}}\cred^\star\tm{\cpgvC{Q}}}$
 \end{restatabletheorem}
 \begin{proof}
-  See~\cite{dardhagay18}.
+  By induction on the derivation of $\pgv{\tm{\cpgvC{P}}\cred\tm{\conf{C}}}$.
+  Details in \cref{prf:thm-pcp-to-pgv-operational-correspondence-soundness}.
 \end{proof}
 
-\begin{restatabletheorem}{thmpcpclosedprogress}[Closed Progress, $\red$]
-  \label{thm:pcp-closed-progress}
+\begin{restatabletheorem}{thmpcptopgvoperationalcorrespondencecompleteness}%
+  [Operational Correspondence, Completeness, $\cpgvC{\cdot}$]
+  \label{thm:pcp-to-pgv-operational-correspondence-completeness}
   \hfill\\%newline before theorem statement
-  If $\seq{P}{\emptyenv}$, then either $\tm{P}=\tm{\halt}$ or there exists a $\tm{Q}$ such that $\tm{P}\red\tm{Q}$.
+  If $\pcp{\seq{P}{\ty{\Gamma}}}$ and $\pcp{\tm{P}\red\tm{Q}}$,
+  then $\pgv{\tm{\cpgvC{P}}\cred^+\tm{\cpgvC{Q}}}$.
 \end{restatabletheorem}
 \begin{proof}
-  Details in~\Cref{prf:thm-pcp-closed-progress}.
+  By induction on the derivation of $\pcp{\tm{P}\red\tm{Q}}$.
+  We apply \cref{lem:pcp-to-pgv-cpgvM-to-cpgvC} for the cases in which
+  the translation switches from $\tm{\cpgvC{\cdot}}$ to $\tm{\cpgvM{\cdot}}$,
+  \ie under actions.
+  We apply \cref{lem:pcp-to-pgv-confs-operational-correspondence-equiv} for
+  \LabTirName{E-LiftSC}. 
+  (Details in \cref{prf:thm-pcp-to-pgv-operational-correspondence-completeness}.)
 \end{proof}

### 2020-10-14 priority-gv.tex
Draft version! :D

@@
 
+Structural congruence satisfies the following axioms: $\LabTirName{SC-LinkSwap}$ allows swapping channels in the link process.
+$\LabTirName{SC-ResLink}$ allows restriction to applied to link which is structurally equivalent to the terminated process, thus allowing elimination of unnecessary restrictions.
+$\LabTirName{SC-ResSwap}$ allows swapping channels and $\LabTirName{SC-ResComm}$ states that restriction is commutative.
+$\LabTirName{SC-ResExt}$ is the standard scope extrusion rule.
+Rules $\LabTirName{SC-ParNil}$, $\LabTirName{SC-ParComm}$ and $\LabTirName{SC-ParAssoc}$ state that parallel composition uses the terminated proceess as the neutral element; it is commutative and associative.
+
 While our configuration reduction is based on the standard evaluation for GV, the increased expressiveness of PGV allows us to simplify the relation on two counts:
@@
 
-Terms are typed by a judgement $\tseq[\cs{p}]{\ty{\Gamma}}{M}{T}$.
+Terms are typed by a judgement $\tseq[\cs{p}]{\ty{\Gamma}}{M}{T}$, which states that ``a term $\tm{M}$ has type $\ty{T}$ and an upper bound on its priority $\cs{p}$ under the typing environment $\ty{\Gamma}$''.
 Typing is standard for the linear $\lambda$-calculus.
@@
 
-Configurations are typed by a judgement $\cseq[\phi]{\ty{\Gamma}}{\conf{C}}$.
+Configurations are typed by a judgement $\cseq[\phi]{\ty{\Gamma}}{\conf{C}}$, which states that ``a configuration $\tm{\conf{C}}$ with flag $\phi$ is well-typed under typing environment $\ty{\Gamma}$''.
 Configuration typing is based on the standard typing for GV.

### 2020-10-14 conclusion.tex
Simon's comments.

@@
 \section{Related Work and Conclusion}
+\simon{I think you really need some mention of MPSTs here, since a key motivation for MPSTs is ensuring deadlock-freedom between participants in a single MPST session.
+You can then talk about the global progress work by Coppo et al., too, which incorporates an 'interaction typing system' (that few people understand).
+How does that relate?}
 
@@
 Following, Kobayashi's line of work, Padovani~\cite{padovani13} adapted the obligation and capability pairs to session types, and later on he further simplified them to a single priority for linear $\pi$-calculus~\cite{padovani14}.
-As discussed by the author, by using the encoding of session types into linear types~\cite{kobayashi07,dardhagiachino12,dardha14beat,dardha16}, the priority-based technique for deadlock freedom can be transferred onto the $\pi$-calculus with session types.
+By using the encoding of session types into linear types~\cite{kobayashi07,dardhagiachino12,dardha14beat,dardha16}, the priority-based technique for deadlock freedom can be transferred onto the $\pi$-calculus with session types.
 
@@
 
-Gay \emph{et al.} \cite{gaynagarajan03} and Vasconcelos \emph{et al.} \cite{vasconcelosravara04,vasconcelosgay06} were the first to introduce a functional language with session types.
-However, such works and including the GV line \cite{gayvasconcelos10,gayvasconcelos12} did not guarantee deadlock freedom.
-This property was later on addressed \cite{lindleymorris15,wadler15} via syntactic restrictions allowing only communication in tree structures, as discussed in the introduction.
-Toninho \emph{et al.} \cite{toninhocaires12} present an encoding of simply-typed $\lambda$-calculus into session-typed $\pi$-calculus investigating the concurrency features of $\lambda$-calculus and providing a new logical explanation.
+Gay \emph{et al.}~\cite{gaynagarajan03} and Vasconcelos \emph{et al.}~\cite{vasconcelosravara04,vasconcelosgay06} were the first to introduce a functional language with session types.
+However, such works and including the GV line~\cite{gayvasconcelos10,gayvasconcelos12} did not guarantee deadlock freedom.
+This property was later on addressed~\cite{lindleymorris15,wadler15} via syntactic restrictions allowing only communication in tree structures, as discussed in the introduction.
+Toninho \emph{et al.}~\cite{toninhocaires12} present an encoding of simply-typed $\lambda$-calculus into session-typed $\pi$-calculus investigating the concurrency features of $\lambda$-calculus and providing a new logical explanation.
 Again, their focus is not on deadlock freedom.
@@
 We used priorities in session types, following Priority CP~\cite{dardhagay18} to guarantee deadlock freedom statically via our priority-based type system.
-We proved the progress property of PGV in \Cref{lem:pgv-closed-progress-confs}.
+We proved the progress property of PGV in \cref{lem:pgv-closed-progress-confs}.
 We also presented an updated version of PCP, where we moved from commuting conversions to structural congruence in the reduction relation, thus remaining faithful to the reduction relation of the $\pi$-calculus.
-Finally, we presented an encoding of Priority CP to Pririty GV in \Cref{sec:pcp-to-pgv}.
-For this translation we proved that it is sound with respect to typing (\Cref{thm:pcp-to-pgv-confs-preservation}) and operational semantics, given by operational correspondence soundness (\Cref{thm:pcp-to-pgv-operational-correspondence-soundness}) and completeness (\Cref{thm:pcp-to-pgv-operational-correspondence-completeness}).
-We identify the following main directions for future work: explore recursion in PGV; define the inverse encoding from Priority GV to Priority CP; add sharing to PGV and PCP following the lines of \cite{balzerpfenning17}.
+Finally, we presented an encoding of Priority CP to Priority GV in \cref{sec:priority-cp}.
+For this translation we proved that it is sound with respect to typing (\cref{thm:pcp-to-pgv-confs-preservation}) and operational semantics, given by operational correspondence soundness (\cref{thm:pcp-to-pgv-operational-correspondence-soundness}) and completeness (\cref{thm:pcp-to-pgv-operational-correspondence-completeness}).
+
+We identify the following main directions for future work: exploring recursion in PGV, by integrating the works of Lindley and Morris~\cite{lindleymorris16} and Padovani and Novara~\cite{padovaninovara15}; defining a translation from Priority GV to Priority CP; and adding sharing following Balzer and Pfenning~\cite{balzerpfenning17}.
 \end{document}

### 2020-10-14 introduction.tex
Finished PGV section.
Simon's comments.

@@
 Communication and concurrency are at the basis of our modern society.
-Every service or device we interact with, from smart phones, to health monitors or any other IoT devices, relies on communication.
+Every service or device we interact with, from smart phones, to health monitors or any other IoT device, relies on communication.
 Hence, verification of communication in such concurrent scenarios is paramount for our own safety.
-Session types are a type formalism which can be used to specify and verify communication protocols between two or more communicating agents \cite{honda93,takeuchihonda94,hondavasconcelos98,carbonehonda07}.
+Session types are a type formalism which can be used to specify and verify communication protocols between two or more communicating agents~\cite{honda93,takeuchihonda94,hondavasconcelos98,carbonehonda07}.
 
-Session types have been studied for many programming paradigms, but to date the most successful settings remain the $\pi$-calculus \cite{sangiorgiwalker01} ---  a calculus for communication and concurrency, and Good Variation \cite[GV]{wadler15,lindleymorris15} ---  a concurrent $\lambda$-calculus.
+Session types have been studied for many programming paradigms, but to date the most successful settings remain the $\pi$-calculus~\cite{sangiorgiwalker01}---a calculus for communication and concurrency, and Good Variation~\cite[GV]{wadler15,lindleymorris15}---a linear concurrent $\lambda$-calculus.
 Session-typed concurrent $\lambda$-calculus and $\pi$-calculus satisfy communication safety and session fidelity, namely that communication proceeds without any type mismatch of the data being transmitted and follows the predefined protocol specified by a session type.
 Another key property in a concurrent setting is deadlock freedom, stating that agents communicate without getting stuck in cyclic dependencies.
-Unfortunately, deadlock freedom is a more involving property and session types alone are not enough to guarantee it.
+Unfortunately, deadlock freedom is a more involving property and session types alone are not enough to rule out deadlock in the presence of multiple interleaved sessions.
 Several techniques have been developed to address deadlock freedom in the $\pi$-calculus and concurrent $\lambda$-calculus with session types.
 
-With regards to $\pi$-calculus, a growing line of work on deadlock freedom leverages the Curry-Howard correspondences between (intuitionistic or classical) linear logic and session-typed $\pi$-calculus \cite{cairespfenning10,wadler12}, which guarantee deadlock freedom by design.
-This is achieved by the cut rule, which merges together constructs for parallel composition and restriction thus forcing processes to share only one channel for communication and eliminating cyclic dependencies.
-Recent developments in the Classical Process \cite[CP]{wadler12} line of work have led to various approaches to decoupling these constructs, either by maintaining a strong correspondence to classical linear logic, as in Hypersequent CP~\cite[HCP]{kokkemontesi19popl,kokkemontesi19tlla}, or by weakening the correspondence to classical linear logic in exchange for a more expressive language, as in Priority CP~\cite[PCP]{dardhagay18}.
-PCP decouples CP's cut rule into two separate constructs, for parallel composition --- by  introducing a mix rule, and for restriction --- by introducing a cycle rule.
-Differently from previous work, PCP allows for cyclic structures and restores deadlock freedom by adding priorities to types following Kobayashi~\cite{kobayashi06} and Padovani \cite{padovani14}.
-Priorities capture the readiness of an action to fire, thus they establish an order of actions in a $\pi$-calculus process.
+With regards to $\pi$-calculus, a growing line of work on deadlock freedom leverages the Curry-Howard correspondences between intuitionistic and classical linear logic and session-typed $\pi$-calculi~\cite{cairespfenning10,wadler12}, which guarantee deadlock freedom by design.
+This is achieved by the cut rule, which merges together constructs for parallel composition and restriction and thus forces processes to share only one channel for communication and eliminating cyclic dependencies.
+Recent developments in the Classical Process~\cite[CP]{wadler12} line of work have led to various approaches to decoupling these constructs, either by maintaining a strong correspondence to classical linear logic, as in Hypersequent CP~\cite[HCP]{kokkemontesi19popl,kokkemontesi19tlla}, or by weakening the correspondence to classical linear logic in exchange for a more expressive language, as in Priority CP~\cite[PCP]{dardhagay18}.
+PCP decouples CP's cut rule into two separate constructs: a construct for parallel composition, by introducing a mix rule; and a construct for restriction, by introducing a cycle rule.
+Differently from previous work, PCP allows for cyclic structures and restores deadlock freedom by adding priorities to types following Kobayashi~\cite{kobayashi06} and Padovani~\cite{padovani14}.
+Priorities establish an order of actions in a $\pi$-calculus process.
 
-With regards to GV, the original work on session types in GV \cite{gayvasconcelos10,gayvasconcelos12} did not satisfy deadlock freedom.
-Later on, calculi in the GV family~\cite{wadler15,lindleymorris15} have achieved deadlock freedom via a syntactic restriction, i.e., by combining channel creation and thread spawning into a single operation, called fork.
-This corresponds to the logical cut rule in CP, which respectively is the combination of a name restriction and a parallel composition.
+With regards to GV, the original work on session types in GV~\cite{gayvasconcelos10,gayvasconcelos12} did not satisfy deadlock freedom.
+Later on, calculi in the GV family~\cite{wadler15,lindleymorris15} have achieved deadlock freedom via a syntactic restriction, \ie by combining channel creation and thread spawning into a single operation, called fork.
+The fork construct is the term representation of the cut construct in CP, which evaluates to a name restriction combined with a parallel composition.
 Unfortunately, as with CP, this is overly restrictive, as it limits the possible communication structures to trees.
-There are several benefits of working with a functional language as opposed to process calculi: a functional language supports higher-order functions (and therefore abstraction) not usually present in process calculi; and it allows us to derive extensions of the communication fragment of the language via well-understood extensions of the functional fragment, i.e., deriving internal/external choice from sum types.
-Moreover, the benefit of working with GV over other session-typed functional languages is that GV has strong ties to linear logic, via its relation to CP~\cite{wadler12}, and consequently it has strong formal properties, e.g., deadlock freedom, albeit in a restrictive way.
+There are several benefits of working with a functional language as opposed to process calculi: a functional language supports higher-order functions (and therefore abstraction) not usually present in process calculi; it allows us to derive extensions of the communication fragment of the language via well-understood extensions of the functional fragment, \ie deriving internal/external choice from sum types; and in concurrent functional calculi, there is a strict separation between the program that the user writes, and the state of a system as it evaluation.
+In the setting of process calculi, these are conflated.
+
+Moreover, the benefit of working with GV over other session-typed functional languages is that GV has strong ties to linear logic, via its relation to CP~\cite{wadler12}, and consequently it has strong formal properties, \eg deadlock freedom, albeit in a restrictive way.
 Following the above observations, we aim to further investigate GV and deadlock freedom and thus pose our research question:
@@
 We address this research question by following the line of work from CP to Priority CP, and present
-Priority GV (PGV), a~variant of GV which decouples channel creation from thread spawning, and restores deadlock freedom by adding priorities.
+Priority GV (PGV), a~variant of GV which decouples channel creation from thread creation, and restores deadlock freedom by adding priorities.
 This work closes the circle of the strong connection of CP and GV, together with their priority versions, PCP and PGV.
 
-\paragraph*{Contributions and Structure of the Paper}
+We make the following contributions:
 \begin{itemize}
@@
 We show our type system for PGV is sound by proving subject reduction (\Cref{thm:pgv-subject-reduction-confs}) and closed progress (\Cref{lem:pgv-closed-progress-confs}).
-\item \emph{Milner's Cyclic Scheduler in PGV} We present, for the first time, Milner's Cyclic Scheduler \cite{milner89} (\Cref{sec:pgv-example}) as a GV term.
+\item \emph{Milner's Cyclic Scheduler in PGV} We present, for the first time, Milner's Cyclic Scheduler~\cite{milner89} (\Cref{sec:pgv-example}) as a GV term.
 This example illustrates how a cyclic structure can now be expressed in PGV thanks to the separation of channel creation and thread creation, and priorities.
-\item \emph{Updated Priority CP} We present an updated version of Priority CP (PCP) \cite{dardhagay18} in \Cref{sec:pcp} where we remove the commuting conversions and move away from a reduction relation as cycle elimination, and towards a reduction as one would expect for a process calculus.
-We then show Milner's Cyclic Scheduler in PCP (in \Cref{sec:pcp-example}) originally presented in \cite{dardhagay18}.
+\item \emph{Updated Priority CP} We present an updated version of Priority CP (PCP)~\cite{dardhagay18} in \Cref{sec:pcp} where we remove the commuting conversions and move away from a reduction relation as cycle elimination, and towards a reduction as one would expect for a process calculus.
+We then show Milner's Cyclic Scheduler in PCP (in \Cref{sec:pcp-example}) originally presented in~\cite{dardhagay18}.
 \item \emph{Encoding PCP into PGV} We present an encoding of PCP into PGV in \Cref{sec:pcp-to-pgv}.

### 2020-10-14 priority-cp.tex
Simon's comments.

@@
 
+\subsubsection*{Example}\label{sec:pcp-example}
+We introduce PCP by revising Milner's cyclic scheduler~\cite{milner89,dardhagay18}.
+Our PCP scheduler has exactly the same behaviour as the version presented in~\cref{sec:pgv-example}.
+We use $\pcp{\tm{\prod_{i}P_i}}$ to denote the parallel composition of the processes $\pcp{\tm{P_i}}$, and $\pcp{\plug{P}{Q}}$ to denote the plugging of $\pcp{\tm{Q}}$ in the one-hole process context $\pcp{\tm{P}}$:
+\[
+  \usingnamespace{pcp}
+  \begin{array}{lrlr}
+    \tm{\sched}
+    & \defeq & \tm{\res{a_1}{a'_1}{\dots{\res{a_n}{a'_n}{}}}}
+               \tm{\res{b_1}{b'_1}{\dots{\res{b_n}{b'_n}{}}}}
+               \tm{\res{c_1}{c'_1}{\dots{\res{c_n}{c'_n}{}}}}
+               \tm{\res{d}{d'}{}}
+    \\ &     & \tm{(
+               \ppar{\proc1}{\agent1}
+               \parallel
+               \prod_{2\leq{i}\leq{n}}(\ppar{\proc{i}}{\wait{c'_{i-1}}{\agent{i}}})
+               )}
+    \\
+    \tm{\agent1}
+    & \defeq & \tm{\close{a_i}{\wait{b_i}{\close{c_i}{\wait{c'_n}{\close{d}{\halt}}}}}}
+    \\
+    \tm{\agent{i}}
+    & \defeq & \tm{\close{a_i}{\wait{b_i}{\close{c_i}{\halt}}}}
+    \\
+    \tm{\proc{i}}
+    & \defeq & \tm{\wait{a'_i}{\plug{P_i}{\close{b'_i}{Q_i}}}}
+  \end{array}
+\]
+
 \subsubsection*{Types}
@@
     \ty{A}, \ty{B}, \ty{C}
-    & \coloneqq & \ty{\tytens[\cs{o}]{A}{B}}
+    & \Coloneqq & \ty{\tytens[\cs{o}]{A}{B}}
       \sep        \ty{\typarr[\cs{o}]{A}{B}}
@@
     \ty{\Gamma}, \ty{\Delta}
-    & \coloneqq & \ty{\emptyenv}
+    & \Coloneqq & \ty{\emptyenv}
       \sep        \ty{\Gamma}, \tmty{x}{A}
@@
     \tm{P}, \tm{Q}
-    & \coloneqq & \tm{\link{x}{y}}
+    & \Coloneqq & \tm{\link{x}{y}}
            \sep   \tm{\res{x}{y}{P}}
@@
 Processes $\pcp{\tm{\send{x}{y}{P}}}$ and $\pcp{\tm{\recv{x}{y}{P}}}$ send or receive over channel $\pcp{\tm{x}}$ a value $\pcp{\tm{y}}$ and proceed as process $\pcp{\tm{P}}$.
-$\pcp{\tm{\close{x}{P}}}$ and $\pcp{\tm{\wait{x}{P}}}$ send and receive an empty value -- denoting the closure of channel $\pcp{\tm{x}}$, and continue as $\pcp{\tm{P}}$.
+The processes $\pcp{\tm{\close{x}{P}}}$ and $\pcp{\tm{\wait{x}{P}}}$ send and receive an empty value---denoting the closure of channel $\pcp{\tm{x}}$, and continue as $\pcp{\tm{P}}$.
 
 Processes $\pcp{\tm{\inl{x}{P}}}$ and $\pcp{\tm{\inr{x}{P}}}$ make a left and right choice, respectively and proceed as process $\pcp{\tm{P}}$.
-$\pcp{\tm{\offer{x}{P}{Q}}}$ dually offers both left and right branches, with continuations $\pcp{\tm{P}}$ and $\pcp{\tm{Q}}$.
-$\pcp{\tm{\absurd{x}}}$ denotes the empty offer.
+Dually, $\pcp{\tm{\offer{x}{P}{Q}}}$ offers both left and right branches, with continuations $\pcp{\tm{P}}$ and $\pcp{\tm{Q}}$, and $\pcp{\tm{\absurd{x}}}$ is the empty offer.
 
@@
 The operational semantics for PCP, given in \cref{fig:pcp-operational-semantics}, is defined as a reduction relation $\pcp{\red}$ on processes (bottom) and uses structural congruence (top).
-Each of the axioms of structural congruence corresponds to the axiom of the same name for PGV.
+Each of the axioms of structural congruence corresponds to the axiom of the same name for PGV.\simon{Maybe worth highlighting that it is an internal mobility calculus, for the pi calculus people.}
 
@@
     & = & \pgv{\tm{\letbind{x}{\select{\labinr}\;{x}}{\cpgvM{P}}}} \\
-    \multicolumn{3}{l}{\pcp{\tm{\cpgvM{\offer{x}{\cpgvM{P}}{\cpgvM{Q}}}}}} \\
+    \multicolumn{3}{l}{\pcp{\tm{\cpgvM{\offer{x}{P}{Q}}}}} \\
     & = & \pgv{\tm{\offer{x}{x}{\cpgvM{P}}{x}{\cpgvM{Q}}}} \\
@@
 
-\subsubsection*{Example: Milner's Cyclic Scheduler}\label{sec:pcp-example}
-We illustrate our translation by giving the PCP version of the cyclic scheduler in~\cref{sec:pgv-example}.
-We implement the scheduler as follows, using $\pcp{\tm{\prod_{i}P_i}}$ to denote the parallel composition of the processes $\pcp{\tm{P_i}}$, and $\pcp{\plug{P}{Q}}$ to denote the plugging of $\pcp{\tm{Q}}$ in the one-hole process context $\pcp{\tm{P}}$:
-\[
-  \usingnamespace{pcp}
-  \begin{array}{lrlr}
-    \tm{\sched}
-    & \defeq & \tm{\res{a_1}{a'_1}{\dots{\res{a_n}{a'_n}{}}}}
-               \tm{\res{b_1}{b'_1}{\dots{\res{b_n}{b'_n}{}}}}
-               \tm{\res{c_1}{c'_1}{\dots{\res{c_n}{c'_n}{}}}}
-               \tm{\res{d}{d'}{}}
-    \\ &     & \tm{(
-               \ppar{\proc1}{\agent1}
-               \parallel
-               \prod_{2\leq{i}\leq{n}}(\ppar{\proc{i}}{\wait{c'_{i-1}}{\agent{i}}})
-               )}
-    \\
-    \tm{\agent1}
-    & \defeq & \tm{\close{a_i}{\wait{b_i}{\close{c_i}{\wait{c'_n}{\close{d}{\halt}}}}}}
-    \\
-    \tm{\agent{i}}
-    & \defeq & \tm{\close{a_i}{\wait{b_i}{\close{c_i}{\halt}}}}
-    \\
-    \tm{\proc{i}}
-    & \defeq & \tm{\wait{a'_i}{\plug{P_i}{\close{b'_i}{Q_i}}}}
-  \end{array}
-\]
+Recall our implementations of Milner's cyclic scheduler in PCP ($\pcp{\sched}$, \cref{sec:pcp-example}) and PGV ($\pgv{\sched}$, \cref{sec:pgv-example}).
 The translation of $\pcp{\sched}$ is exactly $\pgv{sched}$ where all processes are child processes.
@@
 \subsubsection*{Type preservation}
-The translations $\tm{\cpgvC{\cdot}}$ and $\tm{\cpgvM{\cdot}}$ preserve typing along $\ty{\cpgvT{\cdot}}$:
+The translations $\tm{\cpgvC{\cdot}}$ and $\tm{\cpgvM{\cdot}}$ preserve typing along $\ty{\cpgvT{\cdot}}$:\simon{This reads weirdly---what's the intuition?}
 \begin{restatablelemma}{lempcptopgvtermspreservation}[Preservation, $\cpgvM{\cdot}$]
@@
 
-\begin{restatabletheorem}{thmpcptopgvoperationalcorrespondencesoundness}%
-  [Operational Correspondence, Soundness, $\cpgvC{\cdot}$]
-  \label{thm:pcp-to-pgv-operational-correspondence-soundness}
-  \hfill\\%newline before theorem statement
-  If $\pcp{\seq{P}{\ty{\Gamma}}}$ and $\pgv{\tm{\cpgvC{P}}\cred\tm{\conf{C}}}$,
-  there exists a $\tm{Q}$ such that $\pcp{\tm{P}\red^+\tm{Q}}$ and $\pgv{\tm{\conf{C}}\cred^\star\tm{\cpgvC{Q}}}$
-\end{restatabletheorem}
-\begin{proof}
-  By induction on the derivation of $\pgv{\tm{\cpgvC{P}}\cred\tm{\conf{C}}}$.
-  Details in \cref{prf:thm-pcp-to-pgv-operational-correspondence-soundness}.
-\end{proof}
-
 \begin{restatabletheorem}{thmpcptopgvoperationalcorrespondencecompleteness}%
@@
 
+\begin{restatabletheorem}{thmpcptopgvoperationalcorrespondencesoundness}%
+  [Operational Correspondence, Soundness, $\cpgvC{\cdot}$]
+  \label{thm:pcp-to-pgv-operational-correspondence-soundness}
+  \hfill\\%newline before theorem statement
+  If $\pcp{\seq{P}{\ty{\Gamma}}}$ and $\pgv{\tm{\cpgvC{P}}\cred\tm{\conf{C}}}$,
+  there exists a $\tm{Q}$ such that $\pcp{\tm{P}\red^+\tm{Q}}$ and $\pgv{\tm{\conf{C}}\cred^\star\tm{\cpgvC{Q}}}$
+\end{restatabletheorem}
+\begin{proof}
+  By induction on the derivation of $\pgv{\tm{\cpgvC{P}}\cred\tm{\conf{C}}}$.\simon{Surprising that you can get it by the derivation of the reduction relation---normally, 'soundness' would be proved by shunting into a canonical form and then proceeding by induction the structure of the term / induction on typing derivations.}
+  Details in \cref{prf:thm-pcp-to-pgv-operational-correspondence-soundness}.
+\end{proof}
+
 \end{document}

### 2020-10-14 priority-gv.tex
Simon's comments.

@@
 
-GV enforces deadlock freedom with a syntactic restriction: it fuses channel creation and thread spawning into a single function, called fork.
+GV enforces deadlock freedom with a syntactic restriction: it fuses channel creation and thread spawning into a single function, called fork\simon{It feels like you've said this a *lot* at this point}.
 This guarantees that if some term $\tm{L}$ has the means to communicate with both $\tm{M}$ and $\tm{N}$, then $\tm{M}$ and $\tm{N}$ cannot communicate amongst themselves.
@@
 Priority GV offers a more fine-grained analysis of the communication structure, and allows cyclic communication structures.
-To illustrate this, we use PGV to implement a variant of Milner's cyclic scheduler in~\cref{sec:pgv-example}.
+To illustrate this, we start by examining a variant of Milner's cyclic scheduler in PGV~\cite{milner89}.
+
+\subsubsection*{Example}\label{sec:pgv-example}
+Suppose a set of processes $\tm{\proc{i}}$, $1\leq{i}\leq{n}$, should be scheduled to perform some tasks in cyclic order, starting with $\tm{\proc0}$, and circling all the way around to $\tm{\proc{n}}$, notifying $\tm{\proc0}$ when every other process has finished their task.
+
+Our scheduler consists of set of agents $\tm{\agent{i}}$, $1\leq{i}\leq{n}$, each representing their respective process.
+Each process $\tm{\proc{i}}$ waits for the signal to start their task on $\tm{a'_i}$, and signals completion on $\tm{b'_i}$.
+Each agent signals their process to start on $\tm{a_i}$, waits for their process to finish on $\tm{b_i}$, and then signals for the next agent to continue on $\tm{c_i}$.
+The agent $\tm{\agent0}$ initiates, then waits for every other process to finish, and signals $\tm{\proc0}$ on $\tm{d}$.
+Every other agent $\tm{\agent{i}}$, $2\leq{i}\leq{n}$ waits on $\tm{c'_{i-1}}$ for the signal to start.
+The dotted lines represent the optional transfer of data between processes.
+With three processes, the scheduler looks as follows:\simon{Sorry, I found this diagram super confusing\dots Is there any way you could make it look more cyclic?
+Also, have the convention of a single annotation per channel, with the convention that the greater agent / proc gets the primed end of the previous channel?}
+\input{fig/pgv-scheduler.tex}
+We implement the scheduler as follows, using $\tm{\prod_{i}\conf{C}_i}$ to denote the parallel composition of the processes $\tm{\conf{C}_i}$, and $\tm{\plug{M}{N}}$ to denote the plugging of $\tm{N}$ in the one-hole term context $\tm{M}$:
+\[
+  \begin{array}{lrlr}
+    \tm{\sched}
+    & \defeq & \tm{\res{a_1}{a'_1}{\dots{\res{a_n}{a'_n}{}}}}
+               \tm{\res{b_1}{b'_1}{\dots{\res{b_n}{b'_n}{}}}}
+               \tm{\res{c_1}{c'_1}{\dots{\res{c_n}{c'_n}{}}}}
+               \tm{\res{d}{d'}{}}
+    \\ &     & \tm{\begin{array}{lll}
+                     (
+                     & \phi\;\proc1
+                     \parallel
+                     \child\;\andthen{\agent1}{\andthen{\wait\;{c'_3}}{\close\;{d}}}
+                     \\
+                     \parallel
+                     &
+                     \prod_{2\leq{i}\leq{n}}
+                     (\child\;\proc{i} \parallel \child\;\andthen{\wait\;{c'_{i-1}}}{\agent{i}})
+                     & )
+                   \end{array}}
+    \\
+    \tm{\agent{i}}
+    & \defeq & \tm{\andthen{\close\;{a_i}}{\andthen{\wait\;{b_i}}{\close\;{c_i}}}}
+    \\
+    \tm{\proc{i}}
+    & \defeq & \tm{\andthen{\wait\;{a'_i}}{\plug{M_i}{\andthen{\close\;{b'_i}}{N_i}}}}
+  \end{array}
+\]
+Each of the channels in the scheduler is has the terminated type, and is merely used to synchronise.
+For simplicity, we let $\tm{\sched}$ be a configuration. However, we could equivalently have represented it as a function which takes three processes as arguments, and creates this configuration.
 
@@
   \ty{S}
-  & \coloneqq & \ty{\tysend[\cs{o}]{T}{S}}
+  & \Coloneqq & \ty{\tysend[\cs{o}]{T}{S}}
     \sep        \ty{\tyrecv[\cs{o}]{T}{S}}
@@
   \ty{T}, \ty{U}
-  & \coloneqq & \ty{\typrod{T}{U}}
+  & \Coloneqq & \ty{\typrod{T}{U}}
     \sep        \ty{\tyunit}
@@
   \ty{\Gamma}, \ty{\Delta}
-  & \coloneqq & \ty{\emptyenv}
+  & \Coloneqq & \ty{\emptyenv}
     \sep        \ty{\Gamma}, \tmty{x}{T}
@@
 \]
-We extend the function $\pr$ to types and typing contexts by returning the smallest priority in the type or context, or $\cs{\top}$ if there is no priority:
+We extend the function $\pr$ to types and typing contexts by returning the smallest priority in the type or context, or $\cs{\top}$ if there is no priority.
+We use $\sqcap$ and $\sqcup$ to denote the minimum and maximum:
 \[
@@
   \tm{L}, \tm{M}, \tm{N}
-  & \coloneqq & \tm{x}
+  & \Coloneqq & \tm{x}
     \sep        \tm{K}
@@
   \tm{K}
-  & \coloneqq & \tm{\link}
+  & \Coloneqq & \tm{\link}
     \sep        \tm{\new}
@@
 Let $\tm{x}$, $\tm{y}$, $\tm{z}$, and $\tm{w}$ range over variable names.
-Occasionally, we use $\tm{a}$, $\tm{b}$, $\tm{c}$, and $\tm{d}$, which range over \emph{free} endpoints.
+Occasionally, we use $\tm{a}$, $\tm{b}$, $\tm{c}$, and $\tm{d}$.
 The term language is the standard linear $\lambda$-calculus with products, sums, and their units, extended with constants $\tm{K}$ for the communication primitives.
@@
 The constants are best understood in conjunction with their typing and reduction rules in~\cref{fig:pgv-typing,fig:pgv-operational-semantics}.
-Briefly, $\tm{\link}$ links two endpoints together, forwarding messages from each to the other, $\tm{\new}$ creates a new channel and returns its endpoints, and $\tm{\spawn}$ spawns off its argument as a new thread.
-The $\tm{send}$ and $\tm{recv}$ functions send and receive values on a channel.
+Briefly, $\tm{\link}$ links two endpoints together, forwarding messages from each to the other, $\tm{\new}$ creates a new channel and returns a pair of its endpoints, and $\tm{\spawn}$ spawns off its argument as a new thread.
+The $\tm{\send}$ and $\tm{\recv}$ functions send and receive values on a channel.
 However, since the typing rules for PGV ensure the linear usage of endpoints, they also return a new copy of the endpoint, to continue the session.
-The $\tm{close}$ and $\tm{wait}$ functions close a channel.
+The $\tm{\close}$ and $\tm{\wait}$ functions close a channel.
 
 We frequently use syntactic sugar to make terms more readable: we write $\tm{\andthen{M}{N}}$ in place of $\tm{\letunit{M}{N}}$, $\tm{\letbind{x}{M}{N}}$ in place of $\tm{(\lambda x.N)\;M}$, and pattern matching lambdas $\tm{\lambda\unit.M}$ and $\tm{\lambda\pair{x}{y}.M}$ in place of $\tm{\lambda z.\letunit{z}{M}}$ and $\tm{\lambda z.\letpair{x}{y}{z}{M}}$.
+Finally, we recover $\tm{\fork}$:
+\[
+  \tm{\fork} \defeq \tm{\lambda x.\letpair{y}{z}{\new\;\unit}{\andthen{\spawn\;{(\lambda\unit.x\;y)}}{z}}}
+\]
 
@@
     \tm{\conf{C}}, \tm{\conf{D}}, \tm{\conf{E}}
-  & \coloneqq & \tm{\phi\;M}
+  & \Coloneqq & \tm{\phi\;M}
     \sep        \tm{\ppar{\conf{C}}{\conf{D}}}
@@
   \tm{\phi}
-  & \coloneqq & \tm{\main}
+  & \Coloneqq & \tm{\main}
     \sep        \tm{\child}
@@
 \]
-Values ($\tm{V}$, $\tm{W}$), ready terms ($\tm{R}$) evaluation contexts ($\tm{E}$), thread evaluation contexts ($\tm{\conf{F}}$), and configuration contexts ($\tm{\conf{G}}$) are defined by the following grammars:
+Values ($\tm{V}$, $\tm{W}$), evaluation contexts ($\tm{E}$), thread evaluation contexts ($\tm{\conf{F}}$), and configuration contexts ($\tm{\conf{G}}$) are defined by the following grammars:
 \[
@@
   \tm{V}, \tm{W}
-  & \coloneqq & \tm{x}
+  & \Coloneqq & \tm{x}
     \sep        \tm{K}
@@
     \sep        \tm{\unit}
-    \sep        \tm{\pair{M}{N}}
-    \sep        \tm{\inl{M}}
-    \sep        \tm{\inr{M}}
+    \sep        \tm{\pair{V}{W}}
+    \sep        \tm{\inl{V}}
+    \sep        \tm{\inr{V}}
   \\
   \tm{E}
-  & \coloneqq & \tm{\hole} \\
+  & \Coloneqq & \tm{\hole} \\
   & \sep      & \tm{E\;M}
@@
   \tm{\conf{F}}
-  & \coloneqq & \tm{\phi\;E}
+  & \Coloneqq & \tm{\phi\;E}
   \\
   \tm{\conf{G}}
-  & \coloneqq & \tm{\hole}
+  & \Coloneqq & \tm{\hole}
     \sep        \tm{\ppar{\conf{G}}{\conf{C}}}
@@
 The integration of priorities into GV is adapted from Padovani and Novara~\cite{padovaninovara15}.
-Paraphrasing from Dardha and Gay~\cite{dardhagay18}, priorities obey the following two laws:
+Paraphrasing Dardha and Gay~\cite{dardhagay18}, priorities obey the following two laws:
 \begin{enumerate}
@@
 
-In PGV, Sequents are annotated by a priority bound $\cs{p}$, which is the \emph{upper bound} on the priority of the communication actions performed while evaluating the term.
+In PGV\simon{I think the rules need more explanation than this---as someone who knows the basics of priorities but is not an expert, I found it difficult to understand the T-Lam rule for example}, sequents are annotated by a priority bound $\cs{p}$, which is the \emph{upper bound} on the priority of the communication actions performed while evaluating the term.
 The \emph{lower bound} is obtained from the typing environment.
@@
 
-Configurations are typed by a judgement $\cseq[\phi]{\ty{\Gamma}}{\conf{C}}$, which states that ``a configuration $\tm{\conf{C}}$ with flag $\phi$ is well-typed under typing environment $\ty{\Gamma}$''.
+Configurations are typed by a judgement $\cseq[\phi]{\ty{\Gamma}}{\conf{C}}$, which states that ``a configuration $\tm{\conf{C}}$ with flag $\phi$ is well-typed under typing environment $\ty{\Gamma}$''\simon{I found it surprising that configuration typing doesn't need priority annotations.
+Can you give an intuitition why?}.
 Configuration typing is based on the standard typing for GV.
@@
 In GV, each parallel composition must split \emph{exactly one} channel of the channel pseudo-type $\ty{S^\sharp}$ into two endpoints of type $\ty{S}$ and $\ty{\co{S}}$ across the parallel composition.
-Consequently, associativity of parallel composition violates type preservation.
-Unfortunately, the practice plays a crucial role in guaranteeing deadlock freedom.
-In PGV, we guarantee deadlock freedom through different means, which removes the channel pseudo-type $\ty{S^\sharp}$, simplifies typing for parallel composition, and restores type preservation for the structural congruence.
+Consequently, associativity of parallel composition does not preserve typing.
+Unfortunately, splitting exactly one channel across a parallel composition in GV plays a crucial role in guaranteeing deadlock freedom.
+In PGV, we guarantee deadlock freedom using priorities, which removes the channel pseudo-type $\ty{S^\sharp}$, simplifies typing for parallel composition, and restores type preservation for the structural congruence.
 
@@
 
-\subsubsection*{Example: Milner's Cyclic Scheduler}\label{sec:pgv-example}
-We illustrate the newfound ability of PGV by encoding a variant of Milner's cyclic scheduler~\cite{milner89,dardhagay18}.
-Suppose a set of processes $\tm{\proc{i}}$, $1\leq{i}\leq{n}$, should be scheduled to perform some tasks in cyclic order, starting with $\tm{\proc0}$, and circling all the way around to $\tm{\proc{n}}$, notifying $\tm{\proc0}$ when every other process has finished their task.
-
-Our scheduler consists of set of agents $\tm{\agent{i}}$, $1\leq{i}\leq{n}$, each representing their respective process.
-Each process $\tm{\proc{i}}$ waits for the signal to start their task on $\tm{a'_i}$, and signals completion on $\tm{b'_i}$.
-Each agent signals their process to start on $\tm{a_i}$, waits for their process to finish on $\tm{b_i}$, and then signals for the next agent to continue on $\tm{c_i}$.
-The agent $\tm{\agent0}$ initiates, then waits for each other process to finish, and signals $\tm{\proc0}$ on $\tm{d}$.
-Every other agent $\tm{\agent{i}}$, $2\leq{i}\leq{n}$ waits on $\tm{c'_{i-1}}$ for the signal to start.
-The dotted lines represent the optional transfer of data between processes.
-With three processes, the scheduler looks as follows:
-\input{fig/pgv-scheduler.tex}
-We implement the scheduler as follows, using $\tm{\prod_{i}\conf{C}_i}$ to denote the parallel composition of the processes $\tm{\conf{C}_i}$:
-\[
-  \begin{array}{lrlr}
-    \tm{\sched}
-    & \defeq & \tm{\res{a_1}{a'_1}{\dots{\res{a_n}{a'_n}{}}}}
-               \tm{\res{b_1}{b'_1}{\dots{\res{b_n}{b'_n}{}}}}
-               \tm{\res{c_1}{c'_1}{\dots{\res{c_n}{c'_n}{}}}}
-               \tm{\res{d}{d'}{}}
-    \\ &     & \tm{\begin{array}{lll}
-                     (
-                     & \phi\;\proc1
-                     \parallel
-                     \child\;\andthen{\agent1}{\andthen{\wait\;{c'_3}}{\close\;{d}}}
-                     \\
-                     \parallel
-                     &
-                     \prod_{2\leq{i}\leq{n}}
-                     (\child\;\proc{i} \parallel \child\;\andthen{\wait\;{c'_{i-1}}}{\agent{i}})
-                     & )
-                   \end{array}}
-    \\
-    \tm{\agent{i}}
-    & \defeq & \tm{\andthen{\close\;{a_i}}{\andthen{\wait\;{b_i}}{\close\;{c_i}}}}
-    \\
-    \tm{\proc{i}}
-    & \defeq & \tm{\andthen{\wait\;{a'_i}}{\plug{M_i}{\andthen{\close\;{b'_i}}{N_i}}}}
-  \end{array}
-\]
-Each of the channels in the scheduler is has the terminated type, and is merely used to synchronise.
-For simplicity, we let $\tm{\sched}$ be a configuration. However, we could equivalently have represented it as a function which takes three processes as arguments, and creates this configuration.
-
 \subsubsection*{Subject reduction}
-Unlike with previous versions of GV, The structural congruence, term reduction, and configuration reduction are all type preserving, though it is not entirely trivial to show.
+Unlike with previous versions of GV, structural congruence, term reduction, and configuration reduction are all type preserving, though it is not entirely trivial to show.
 
@@
 \end{proof}
-The proofs for subject congruence and subject reduction are standard.
+The proofs for subject congruence and subject reduction are standard inductions.
+The details can be found in \cref{prf:lem-pgv-subject-reduction-terms,prf:lem-pgv-subject-congruence,prf:thm-pgv-subject-reduction-confs}.
 \begin{restatablelemma}{lempgvsubjectreductionterms}[Subject Reduction, $\tred$]
@@
 \end{restatablelemma}
-\begin{proof}
-  By induction on the derivation of $\tm{M}\tred\tm{M'}$.
-  Details in \cref{prf:lem-pgv-subject-reduction-terms}.
-\end{proof}
 \begin{restatablelemma}{lempgvsubjectcongruence}[Subject Congruence, $\equiv$]
@@
 \end{restatablelemma}
-\begin{proof}
-  By induction on the derivation of $\tm{\conf{C}}\equiv\tm{\conf{C'}}$.
-  Details in \cref{prf:lem-pgv-subject-congruence}.
-\end{proof}
 \begin{restatabletheorem}{thmpgvsubjectreductionconfs}[Subject Reduction, $\cred$]
@@
 \end{restatabletheorem}
-\begin{proof}
-  By induction on the derivation of $\tm{\conf{C}}\cred\tm{\conf{C'}}$.
-  Details in \cref{prf:thm-pgv-subject-reduction-confs}.
-\end{proof}
 
@@
 If a term returns an endpoint, then its normal form contains a thread which is ready to communicate on the dual of that endpoint.
-For instance, let us consider an example, adapted from Lindley and Morris~\cite{lindleymorris15}, in which a term returns an endpoint linked to an echo server.
+This behaviour is not new to PGV.
+Let us consider an example, adapted from Lindley and Morris~\cite{lindleymorris15}, in which a term returns an endpoint linked to an echo server.
 The echo server receives a value and sends it back unchanged:
@@
 \]
-In~\cref{cor:pgv-closed-progress}, we show that if a term does not return an endpoint, then it produces \emph{only} a value.
+In~\cref{cor:pgv-closed-progress} we show that if a term does not return an endpoint, it must produce \emph{only} a value.
 
@@
   \label{def:pgv-actions}
-  A~term acts on an endpoint $\tm{x}$ if it is of the form $\tm{\send\;\pair{V}{x}}$, $\tm{\recv\;{x}}$, $\tm{\close\;{x}}$, or $\tm{\wait\;{x}}$.
-A~term is an action if it acts on some endpoint $\tm{x}$.
+  A~term \emph{acts on} an endpoint $\tm{x}$ if it is of the form $\tm{\send\;\pair{V}{x}}$, $\tm{\recv\;{x}}$, $\tm{\close\;{x}}$, or $\tm{\wait\;{x}}$.
+A~term is an \emph{action} if it acts on some endpoint $\tm{x}$.
 \end{definition}
@@
   \label{def:pgv-ready-actions}
-  A~term $\tm{L}$ is ready if it is of the form $\tm{\plug{E}{M}}$, where $\tm{M}$ is of the form $\tm{\new}$, $\tm{\spawn\;N}$, $\tm{\link\;\pair{x}{y}}$ or $\tm{\link\;\pair{y}{x}}$, or $\tm{M}$ acts on $\tm{x}$.
-In the latter case, we say that $\tm{L}$ is ready to act on $\tm{x}$.
+  A~term $\tm{L}$ is \emph{ready} if it is of the form $\tm{\plug{E}{M}}$, where $\tm{M}$ is of the form $\tm{\new}$, $\tm{\spawn\;N}$, $\tm{\link\;\pair{x}{y}}$ or $\tm{\link\;\pair{y}{x}}$, or $\tm{M}$ acts on $\tm{x}$.
+In the latter case, we say that $\tm{L}$ is \emph{ready to act on} $\tm{x}$.
 \end{definition}
@@
   \label{lem:pgv-open-progress-terms}
-  If $\tseq[\cs{p}]{\ty{\Gamma}}{M}{T}$, then:
+  If $\tseq[\cs{p}]{\ty{\Gamma}}{M}{T}$ and $\ty{\Gamma}$ contains only session types, then:
   \begin{itemize}
@@
 
-For simplicity, we only present progress for closed configurations.
+% For simplicity, we only present progress for closed configurations.
 A~suitable notion of progress for open configurations does exist, but its normal forms are significantly more complex, as they may also contain threads which are blocked on free endpoints, threads blocked on endpoints for which the duals are still to be used in those threads blocked on free endpoints, \etc.
@@
 The statement of closed progress is rather subtle.
-A~configuration either reduces, or it is equivalent to configuration in the stated normal form.
+A~configuration either reduces, or it is equivalent to configuration in normal form.
 Crucial to the normal form is that each term $\tm{M_i}$ is blocked on the corresponding channel $\tm{x_i}$, and hence no two terms act on dual endpoints.
 Furthermore, no term $\tm{M_i}$ can perform a communication action by itself, since those are excluded by the definition of actions:
-\begin{restatabletheorem}{thmpgvclosedprogressconfs}[Closed Progress, $\cred$]
-  \label{lem:pgv-closed-progress-confs}
-  If $\cseq[\main]{\emptyenv}{\conf{C}}$ and $\tm{\conf{C}}$ is in canonical form, then either $\tm{\conf{C}}\cred\tm{\conf{D}}$ for some $\tm{\conf{D}}$; or
+\begin{definition}[Normal Forms]
+  A~configuration $\tm{\conf{C}}$ is in normal form if it is of the following form, where each $\tm{M_i}$ is ready to act on $\tm{x_i}$:
   \[
@@
   \]
-  where each $\tm{M_i}$ is ready to act on $\tm{x_i}$.
+\end{definition}
+\begin{restatabletheorem}{thmpgvclosedprogressconfs}[Closed Progress, $\cred$]
+  \label{lem:pgv-closed-progress-confs}
+  If $\cseq[\main]{\emptyenv}{\conf{C}}$ and $\tm{\conf{C}}$ is in canonical form, then either $\tm{\conf{C}}\cred\tm{\conf{D}}$ for some $\tm{\conf{D}}$; or $\tm{\conf{C}}$ is equivalent to a configuration in normal form.
 \end{restatabletheorem}

### 2020-10-14 conclusion.tex
Minor changes.

@@
 \section{Related Work and Conclusion}
-\simon{I think you really need some mention of MPSTs here, since a key motivation for MPSTs is ensuring deadlock-freedom between participants in a single MPST session.
-You can then talk about the global progress work by Coppo et al., too, which incorporates an 'interaction typing system' (that few people understand).
+\textit{Simon: I think you really need some mention of MPSTs here, since a key motivation for MPSTs is ensuring deadlock-freedom between participants in a single MPST session.
+You can then talk about the global progress work by Coppo \etal, too, which incorporates an 'interaction typing system' (that few people understand).
 How does that relate?}
 
-\paragraph{On deadlock freedom and progress in $\pi$-calculus and $\lambda$-calculus}
+\subsubsection*{On deadlock freedom and progress in $\pi$-calculus and $\lambda$-calculus}
 Deadlock freedom and progress have been studied thoroughly in the $\pi$-calculus with or without session types.
@@
 
-The groundbreaking work on progress for session-typed $\pi$-calculus, by Dezani \emph{et al.}~\cite{dezani-ciancaglinimostrous06}, guarantees progress by allowing only one active session at a time.
+The groundbreaking work on progress for session-typed $\pi$-calculus, by Dezani \etal~\cite{dezani-ciancaglinimostrous06}, guarantees progress by allowing only one active session at a time.
 In~\cite{dezani-ciancagliniliguoro09progress} the authors introduce a partial order on channels in line with Kobayashi's work~\cite{kobayashi98}.
 Carbone and Debois~\cite{carbonedebois10} define progress for session typed $\pi$-calculus in terms of a \emph{catalyser} used to provide a missing counterpart to a process, thus guaranteeing progress.
-Carbone \emph{et al.}~\cite{carbonedardha14} studied further the use of catalysers and showed that progress is a compositional form of lock-freedom.
+Carbone \etal~\cite{carbonedardha14} studied further the use of catalysers and showed that progress is a compositional form of lock-freedom.
 As in~\cite{padovani14} the authors show that by using the encoding of session types~\cite{dardhagiachino12} and Kobayashi's obligations/capabilities, we can obtain progress for session types.
@@
 
-Gay \emph{et al.}~\cite{gaynagarajan03} and Vasconcelos \emph{et al.}~\cite{vasconcelosravara04,vasconcelosgay06} were the first to introduce a functional language with session types.
+Gay~\etal~\cite{gaynagarajan03} and Vasconcelos \etal~\cite{vasconcelosravara04,vasconcelosgay06} were the first to introduce a functional language with session types.
 However, such works and including the GV line~\cite{gayvasconcelos10,gayvasconcelos12} did not guarantee deadlock freedom.
 This property was later on addressed~\cite{lindleymorris15,wadler15} via syntactic restrictions allowing only communication in tree structures, as discussed in the introduction.
-Toninho \emph{et al.}~\cite{toninhocaires12} present an encoding of simply-typed $\lambda$-calculus into session-typed $\pi$-calculus investigating the concurrency features of $\lambda$-calculus and providing a new logical explanation.
+Toninho \etal~\cite{toninhocaires12} present an encoding of simply-typed $\lambda$-calculus into session-typed $\pi$-calculus investigating the concurrency features of $\lambda$-calculus and providing a new logical explanation.
 Again, their focus is not on deadlock freedom.
 
-\paragraph{On Curry-Howard correspondences}
+\subsubsection*{On Curry-Howard correspondences}
 The Curry-Howard correspondence between logic and types laid the foundation for functional programming~\cite{wadler15}.
@@
 Dardha and P\'{e}rez~\cite{dardhaperez15} compare Kobayashi-style typing and CLL typing, and prove that CLL corresponds to Kobayashi's system with the restriction that only single cuts are allowed.
-Balzer \emph{et al.} study sharing~\cite{balzerpfenning17} and later on deadlock freedom~\cite{balzertoninho19} in a session-typed $\pi$-calculus based on the correspondence with DILL.
+Balzer \etal study sharing~\cite{balzerpfenning17} and later on deadlock freedom~\cite{balzertoninho19} in a session-typed $\pi$-calculus based on the correspondence with DILL.
 In~\cite{balzerpfenning17} types are either linear or shared with modal operators connecting them, which allows sharing, but not deadlock freedom.
@@
 
-\paragraph{Conclusion and Future Work}
+\subsubsection*{Conclusion and Future Work}
 In this paper we presented Priority GV, which is a functional language with session types.

### 2020-10-14 fig/pgv-scheduler.tex
Changed drawing of scheduler.

@@
 
-    \draw
+    \draw[->]
     (agent1) -- node[pos=0.35,above] {$c_1$}
     node[pos=1.0,above] {$c'_1$} ++ (agent2);
-    \draw
+    \draw[->]
     (agent2) -- node[pos=0.1,above] {$c_2$}
     node[pos=0.9,above] {$c'_2$} ++ (agent3);
-    \draw
+    \draw[->]
     (agent3) -- node[pos=0.0,above] {$c_3$}
@@
 
-    \draw[double]
+    \draw[implies-implies,double]
     (agent1) -- node[pos=0.15,above] {$a_1$} node[pos=0.85,above] {$a'_1$}
     node[pos=0.15,below] {$b_1$} node[pos=0.85,below] {$b'_1$} ++ (proc1);
-    \draw[double]
+    \draw[implies-implies,double]
     (agent2) -- node[pos=0.15,above] {$a_2$} node[pos=0.85,above] {$a'_2$}
     node[pos=0.15,below] {$b_2$} node[pos=0.85,below] {$b'_2$} ++ (proc2);
-    \draw[double]
+    \draw[implies-implies,double]
     (agent3) -- node[pos=0.15,above] {$a_3$} node[pos=0.85,above] {$a'_3$}

### 2020-10-15 conclusion.tex
Processed more feedback.

@@
 Carbone \etal~\cite{carbonedardha14} studied further the use of catalysers and showed that progress is a compositional form of lock-freedom.
-As in~\cite{padovani14} the authors show that by using the encoding of session types~\cite{dardhagiachino12} and Kobayashi's obligations/capabilities, we can obtain progress for session types.
+Following Padovani~\cite{padovani14}, the authors show that by using the encoding of session types~\cite{dardhagiachino12} and Kobayashi's obligations/capabilities, we can obtain progress for session types.
 Vieira and Vasconcelos~\cite{vieiravasconcelos13} used single priorities and an abstract partial order to guarantee deadlock freedom in a session-typed $\pi$-calculus.
@@
 Dardha and P\'{e}rez~\cite{dardhaperez15} compare Kobayashi-style typing and CLL typing, and prove that CLL corresponds to Kobayashi's system with the restriction that only single cuts are allowed.
-Balzer \etal study sharing~\cite{balzerpfenning17} and later on deadlock freedom~\cite{balzertoninho19} in a session-typed $\pi$-calculus based on the correspondence with DILL.
-In~\cite{balzerpfenning17} types are either linear or shared with modal operators connecting them, which allows sharing, but not deadlock freedom.
-In~\cite{balzertoninho19} types are enhanced with information regarding the order of acquiring resources, which guarantees deadlock freedom.
+Balzer \etal study sharing~\cite{balzerpfenning17} and later on deadlock freedom~\cite{balzertoninho19} in $\text{SILL}_S$ and $\text{SILL}_{S+}$, session-typed $\pi$-calculi based on the correspondence with DILL.
+In $\text{SILL}_S$~\cite{balzerpfenning17}, types are either linear or shared with modal operators connecting them, which allows sharing, but not deadlock freedom.
+In $\text{SILL}_{S+}$~\cite{balzertoninho19} types are enhanced with information regarding the order of acquiring resources, which guarantees deadlock freedom.
 

### 2020-10-15 introduction.tex
Processed more feedback.

@@
 Unfortunately, as with CP, this is overly restrictive, as it limits the possible communication structures to trees.
-There are several benefits of working with a functional language as opposed to process calculi: a functional language supports higher-order functions (and therefore abstraction) not usually present in process calculi; it allows us to derive extensions of the communication fragment of the language via well-understood extensions of the functional fragment, \ie deriving internal/external choice from sum types; and in concurrent functional calculi, there is a strict separation between the program that the user writes, and the state of a system as it evaluation.
+There are several benefits of working with a functional language as opposed to process calculi.
+Functional languages support higher-order functions, and have a capability for abstraction not usually present in process calculi.
+Working within a functional language allows us to derive extensions of the communication capabilities of the language via well-understood extensions of the functional fragment, \ie deriving internal/external choice via sum types.
+Finally, in concurrent functional calculi, there is a strict separation between the program that the user writes, and the state of a system as it evaluation.
 In the setting of process calculi, these are conflated.
 
-Moreover, the benefit of working with GV over other session-typed functional languages is that GV has strong ties to linear logic, via its relation to CP~\cite{wadler12}, and consequently it has strong formal properties, \eg deadlock freedom, albeit in a restrictive way.
+Moreover, the benefit of working with GV over other session-typed functional languages is that GV has strong ties to linear logic, via its relation to CP~\cite{wadler12}, and consequently it has strong formal properties, \eg deadlock freedom in a restrictive way.
 Following the above observations, we aim to further investigate GV and deadlock freedom and thus pose our research question:
@@
 We then show Milner's Cyclic Scheduler in PCP (in \Cref{sec:pcp-example}) originally presented in~\cite{dardhagay18}.
-\item \emph{Encoding PCP into PGV} We present an encoding of PCP into PGV in \Cref{sec:pcp-to-pgv}.
+\item \emph{Encoding PCP into PGV} We present an encoding of PCP into PGV in \Cref{sec:pcp}.
 We prove the encoding is sound and complete with respect to typing (\Cref{thm:pcp-to-pgv-confs-preservation}) and operational semantics (\Cref{thm:pcp-to-pgv-operational-correspondence-soundness} and \Cref{thm:pcp-to-pgv-operational-correspondence-completeness}).

### 2020-10-15 priority-gv.tex
Processed more feedback.

@@
 \subsubsection*{Subject reduction}
-Unlike with previous versions of GV, structural congruence, term reduction, and configuration reduction are all type preserving, though it is not entirely trivial to show.
+Unlike with previous versions of GV, structural congruence, term reduction, and configuration reduction are all type preserving.
 
-Specifically, we need to show that substitution preserves priority constraints.
+Most of the subject reduction proof is standard.
+Novel is that we must show that substitution preserves priority constraints.
 For this, it helps to prove~\cref{lem:pgv-value-done}, which shows that values have finished all their communication actions, and that any priorities present in the type of the value come from the typing environment.

### 2020-10-15 priority-cp.tex
Compressed lists and subsubsections, minor rewrites in text.

@@
 As a result of $(\kappa_{\parr})$, $\pcp{\tm{Q}}$ becomes blocked on $\pcp{\tm{\labrecv{x}{y}}}$, and any actions $\pcp{\tm{Q}}$ was able to perform become unavailable.
-Consequently, CP is non-confluent, for instance:
+Consequently, CP is non-confluent.
+For instance,
 \begin{mathpar}
@@
     &
-    \rotatebox[origin=c]{325}{$\red$}\qquad
+    \rotatebox[origin=c]{325}{$\red^2$}\qquad
     \\
@@
     &
-    {\tm{\res{x}{x'}{(\ppar{\recv{a}{y}{P}}{Q})}}}
-    \\ 
-    &
-    \rotatebox[origin=c]{270}{$\red$}
-    \\
-    &
-    \tm{\recv{a}{y}{\res{x}{x'}{(\ppar{P}{Q})}}}
+    {\tm{\recv{a}{y}{\res{x}{x'}{(\ppar{P}{Q})}}}}
   \end{array}
 \end{mathpar}
-Commuting conversions break our intuition for priorities, since ``if a prefix on a channel endpoint $\pcp{\tm{x}}$ with priority $\cs{o}$ is pulled out at top level, then to preserve priority conditions in the typing rules [..], it is necessary to increase priorities of all actions after the prefix on $\pcp{\tm{x}}$'' by $\cs{o+1}$~\cite{dardhagay18}.
-Consequently, the intuition that an action with lower priority occurs before an action with higher priority is shattered.
+Commuting conversions break our intuition that an action with lower priority occurs before an action with higher priority, since ``if a prefix on a channel endpoint $\pcp{\tm{x}}$ with priority $\cs{o}$ is pulled out at top level, then to preserve priority conditions in the typing rules [..], it is necessary to increase priorities of all actions after the prefix on $\pcp{\tm{x}}$'' by $\cs{o+1}$~\cite{dardhagay18}.
 

### 2020-10-15 priority-gv.tex
Compressed lists and subsubsections, minor rewrites in text.

@@
 \[
-\begin{array}{ll}
   \tm{\main}  + \tm{\child} = \tm{\main}
-  \\
+  \quad
   \tm{\child} + \tm{\main}  = \tm{\main}
-\end{array}
   \quad
-\begin{array}{ll}
   \tm{\child} + \tm{\child} = \tm{\child}
-  \\
-  \tm{\main}  + \tm{\main} \; \text{undefined}
-\end{array}
+  \quad
+  (\tm{\main}  + \tm{\main} \; \text{is undefined})
 \]
@@
 
-While our configuration reduction is based on the standard evaluation for GV, the increased expressiveness of PGV allows us to simplify the relation on two counts:
-
+While our configuration reduction is based on the standard evaluation for GV, the increased expressiveness of PGV allows us to simplify the relation on two counts.
 Firstly, we decompose the $\tm{\fork}$ construct.
@@
 In PGV, the behaviour matches that of link in CP.
-
 Secondly, our structural congruence is type preserving.
@@
 
-While our configuration typing is based on the standard typing for GV, it differs on two counts:
-
+While our configuration typing is based on the standard typing for GV, it differs on two counts.
 Firstly, we require that child threads return the unit value, as opposed to a terminated channel.
-
 Secondly, we simplify typing for parallel composition.
@@
 \]
-If we reduce the above program, we get the following normal form.
-Clearly, no more evaluation is possible, even though the configuration contains the thread $\tm{\child\;\echo_{x'}}$, which is blocked on $\tm{x'}$:
-\[
-  \tm{\res{x}{x'}{(\ppar{\main\;x}{\child\;\echo_{x'}})}}
-\]
+If we reduce the above program, we get $\tm{\res{x}{x'}{(\ppar{\main\;x}{\child\;\echo_{x'}})}}$.
+Clearly, no more evaluation is possible, even though the configuration contains the thread $\tm{\child\;\echo_{x'}}$, which is blocked on $\tm{x'}$.
 In~\cref{cor:pgv-closed-progress} we show that if a term does not return an endpoint, it must produce \emph{only} a value.
@@
 Canonical forms deviate from those for GV, in that we opt to move all $\nu$-binders to the top.
-The standard GV normal forms, alternating $\nu$-binders and their corresponding parallel compositions, do not work for PGV, since multiple channels may be split across a single parallel composition:
+The standard GV canonical form, alternating $\nu$-binders and their corresponding parallel compositions, does not work for PGV, since multiple channels may be split across a single parallel composition:
 \begin{definition}[Canonical Forms]
@@
 
-% For simplicity, we only present progress for closed configurations.
-A~suitable notion of progress for open configurations does exist, but its normal forms are significantly more complex, as they may also contain threads which are blocked on free endpoints, threads blocked on endpoints for which the duals are still to be used in those threads blocked on free endpoints, \etc.
-
-The statement of closed progress is rather subtle.
 A~configuration either reduces, or it is equivalent to configuration in normal form.

### 2020-10-15 introduction.tex
Compressed lists and subsubsections, minor rewrites in text.
pass on introduction

@@
 Every service or device we interact with, from smart phones, to health monitors or any other IoT device, relies on communication.
-Hence, verification of communication in such concurrent scenarios is paramount for our own safety.
+Hence, verification of communication in concurrent settings is important for our own safety.
 Session types are a type formalism which can be used to specify and verify communication protocols between two or more communicating agents~\cite{honda93,takeuchihonda94,hondavasconcelos98,carbonehonda07}.
-
-Session types have been studied for many programming paradigms, but to date the most successful settings remain the $\pi$-calculus~\cite{sangiorgiwalker01}---a calculus for communication and concurrency, and Good Variation~\cite[GV]{wadler15,lindleymorris15}---a linear concurrent $\lambda$-calculus.
+Session types have been studied for many programming paradigms, but to date the most successful settings remain the $\pi$-calculus~\cite{sangiorgiwalker01}---a process calculus for communication and concurrency, and Good Variation (GV)~\cite{wadler15,lindleymorris15}---a linear concurrent $\lambda$-calculus.
 Session-typed concurrent $\lambda$-calculus and $\pi$-calculus satisfy communication safety and session fidelity, namely that communication proceeds without any type mismatch of the data being transmitted and follows the predefined protocol specified by a session type.
 Another key property in a concurrent setting is deadlock freedom, stating that agents communicate without getting stuck in cyclic dependencies.
-Unfortunately, deadlock freedom is a more involving property and session types alone are not enough to rule out deadlock in the presence of multiple interleaved sessions.
+Unfortunately, deadlock freedom is a more involving property and session types alone are not enough to rule out deadlocks in cyclic structures of multiple interleaved sessions.
 Several techniques have been developed to address deadlock freedom in the $\pi$-calculus and concurrent $\lambda$-calculus with session types.
 
-With regards to $\pi$-calculus, a growing line of work on deadlock freedom leverages the Curry-Howard correspondences between intuitionistic and classical linear logic and session-typed $\pi$-calculi~\cite{cairespfenning10,wadler12}, which guarantee deadlock freedom by design.
-This is achieved by the cut rule, which merges together constructs for parallel composition and restriction and thus forces processes to share only one channel for communication and eliminating cyclic dependencies.
-Recent developments in the Classical Process~\cite[CP]{wadler12} line of work have led to various approaches to decoupling these constructs, either by maintaining a strong correspondence to classical linear logic, as in Hypersequent CP~\cite[HCP]{kokkemontesi19popl,kokkemontesi19tlla}, or by weakening the correspondence to classical linear logic in exchange for a more expressive language, as in Priority CP~\cite[PCP]{dardhagay18}.
+With regards to $\pi$-calculus, a growing line of work on deadlock freedom leverages the Curry-Howard correspondences between intuitionistic or classical linear logic and session-typed $\pi$-calculus~\cite{cairespfenning10,wadler12}, which guarantee deadlock freedom by design.
+This is achieved by merging constructs for parallel composition and channel restriction under the logical cut rule, which forces processes to share only one channel for communication and eliminates cyclic dependencies.
+Recent developments in the Classical Process (CP)~\cite{wadler12} line of work have led to various approaches to decoupling these constructs, either by maintaining a strong correspondence to classical linear logic, as in Hypersequent CP~\cite{kokkemontesi19popl,kokkemontesi19tlla}, or by weakening the correspondence to classical linear logic in exchange for a more expressive language, as in Priority CP (PCP)~\cite{dardhagay18}.
 PCP decouples CP's cut rule into two separate constructs: a construct for parallel composition, by introducing a mix rule; and a construct for restriction, by introducing a cycle rule.
-Differently from previous work, PCP allows for cyclic structures and restores deadlock freedom by adding priorities to types following Kobayashi~\cite{kobayashi06} and Padovani~\cite{padovani14}.
-Priorities establish an order of actions in a $\pi$-calculus process.
+Differently from previous work, PCP allows cyclic structures and restores deadlock freedom by adding priorities to types following Kobayashi~\cite{kobayashi06} and Padovani~\cite{padovani14}.
+Priorities establish an order of actions in a $\pi$-calculus process and are used to rule out bad interleaving of channels, guaranteeing deadlock freedom.
 
@@
 Later on, calculi in the GV family~\cite{wadler15,lindleymorris15} have achieved deadlock freedom via a syntactic restriction, \ie by combining channel creation and thread spawning into a single operation, called fork.
-The fork construct is the term representation of the cut construct in CP, which evaluates to a name restriction combined with a parallel composition.
-Unfortunately, as with CP, this is overly restrictive, as it limits the possible communication structures to trees.
+The fork construct is the term representation of the cut construct in CP, which evaluates to a channel restriction combined with a parallel composition.
+Unfortunately, as with CP, this is restrictive as it limits communication structures to trees.
 There are several benefits of working with a functional language as opposed to process calculi.
 Functional languages support higher-order functions, and have a capability for abstraction not usually present in process calculi.
-Working within a functional language allows us to derive extensions of the communication capabilities of the language via well-understood extensions of the functional fragment, \ie deriving internal/external choice via sum types.
-Finally, in concurrent functional calculi, there is a strict separation between the program that the user writes, and the state of a system as it evaluation.
-In the setting of process calculi, these are conflated.
-
-Moreover, the benefit of working with GV over other session-typed functional languages is that GV has strong ties to linear logic, via its relation to CP~\cite{wadler12}, and consequently it has strong formal properties, \eg deadlock freedom in a restrictive way.
-Following the above observations, we aim to further investigate GV and deadlock freedom and thus pose our research question:
+Working within a functional language allows to derive extensions of the communication capabilities of the language via well-understood extensions of the functional fragment, \ie deriving internal/external choice via sum types.
+Finally, in concurrent functional calculi, there is a clear separation between a program that the user writes, and a configuration, which is the state of a system as it evaluates.
+In process calculi, these are conflated.
+Moreover, the benefit of working with GV over other session-typed functional languages is that GV has strong ties to linear logic, via its relation to CP~\cite{wadler12}, and consequently it has strong formal properties, \eg deadlock freedom, albeit in a restrictive way.
+Following the above observations, we aim to further investigate GV and deadlock freedom, thus we pose our research question:
 
@@
 We address this research question by following the line of work from CP to Priority CP, and present
-Priority GV (PGV), a~variant of GV which decouples channel creation from thread creation, and restores deadlock freedom by adding priorities.
+Priority GV (PGV), a~variant of GV which decouples channel creation from thread spawning and restores deadlock freedom by adding priorities.
 This work closes the circle of the strong connection of CP and GV, together with their priority versions, PCP and PGV.
@@
 \item \emph{Milner's Cyclic Scheduler in PGV} We present, for the first time, Milner's Cyclic Scheduler~\cite{milner89} (\Cref{sec:pgv-example}) as a GV term.
-This example illustrates how a cyclic structure can now be expressed in PGV thanks to the separation of channel creation and thread creation, and priorities.
-\item \emph{Updated Priority CP} We present an updated version of Priority CP (PCP)~\cite{dardhagay18} in \Cref{sec:pcp} where we remove the commuting conversions and move away from a reduction relation as cycle elimination, and towards a reduction as one would expect for a process calculus.
-We then show Milner's Cyclic Scheduler in PCP (in \Cref{sec:pcp-example}) originally presented in~\cite{dardhagay18}.
+This example illustrates how a cyclic structure can be expressed now in GV thanks to the separation of channel creation and thread spawning.
+\item \emph{Updated Priority CP} We present an updated version of Priority CP (PCP)~\cite{dardhagay18} in \Cref{sec:pcp}; we remove commuting conversions and move away from a reduction relation as cycle elimination, and towards a reduction relation based on structural congruence, as one would expect in a process calculus.
+We show Milner's Cyclic Scheduler in PCP (in \Cref{sec:pcp-example}) originally presented in~\cite{dardhagay18}.
 \item \emph{Encoding PCP into PGV} We present an encoding of PCP into PGV in \Cref{sec:pcp}.
 We prove the encoding is sound and complete with respect to typing (\Cref{thm:pcp-to-pgv-confs-preservation}) and operational semantics (\Cref{thm:pcp-to-pgv-operational-correspondence-soundness} and \Cref{thm:pcp-to-pgv-operational-correspondence-completeness}).
-Finally, we can encode Milner's Cyclic Scheduler in PCP (\Cref{sec:pcp-example}) resulting in the Scheduler in PGV (\Cref{sec:pgv-example}).
+Finally, using our encoding, we can encode Milner's Cyclic Scheduler in PCP (\Cref{sec:pcp-example}) resulting in the Scheduler in PGV (\Cref{sec:pgv-example}).
 \end{itemize}

### 2020-10-15 priority-cp.tex

@@
     &
-    \rotatebox[origin=c]{325}{$\red^2$}\qquad
+    \rotatebox[origin=c]{325}{$\red^+$}\qquad
     \\
@@
 The operational semantics for PCP, given in \cref{fig:pcp-operational-semantics}, is defined as a reduction relation $\pcp{\red}$ on processes (bottom) and uses structural congruence (top).
-Each of the axioms of structural congruence corresponds to the axiom of the same name for PGV.\simon{Maybe worth highlighting that it is an internal mobility calculus, for the pi calculus people.}
+Each of the axioms of structural congruence corresponds to the axiom of the same name for PGV.
+We write $\pcp{\red^+}$ for the transitive closures, and $\pcp{\tred^\star}$ for the reflexive-transitive closures.\simon{Maybe worth highlighting that it is an internal mobility calculus, for the pi calculus people.}
 

### 2020-10-15 priority-gv.tex

@@
 \subsubsection*{Example}\label{sec:pgv-example}
-Suppose a set of processes $\tm{\proc{i}}$, $1\leq{i}\leq{n}$, should be scheduled to perform some tasks in cyclic order, starting with $\tm{\proc0}$, and circling all the way around to $\tm{\proc{n}}$, notifying $\tm{\proc0}$ when every other process has finished their task.
+Suppose a set of processes $\tm{\proc{i}}$, $1\leq{i}\leq{n}$, should be scheduled to perform some tasks in cyclic order, starting with $\tm{\proc0}$, ending with $\tm{\proc{n}}$, and notifying $\tm{\proc0}$ when all processes have finished.
 

### 2020-10-15 priority-cp.tex
Created compacttheorems environment for compact theorems in sequence.

@@
 Each of the axioms of structural congruence corresponds to the axiom of the same name for PGV.
-We write $\pcp{\red^+}$ for the transitive closures, and $\pcp{\tred^\star}$ for the reflexive-transitive closures.\simon{Maybe worth highlighting that it is an internal mobility calculus, for the pi calculus people.}
+We write $\pcp{\red^+}$ for the transitive closures, and $\pcp{\red^\star}$ for the reflexive-transitive closures.\simon{Maybe worth highlighting that it is an internal mobility calculus, for the pi calculus people.}
 
@@
 Reduction in our version of PCP preserve types. Since our reduction relation is a strict subset of the reduction relation in the original~\cite{dardhagay18}, we defer to their proofs:
+\begin{compacttheorems}
   \begin{restatablelemma}{lempcpsubjectcongruence}[Subject Congruence, $\pcp{\equiv}$]
@@
   \end{restatabletheorem}
+\end{compacttheorems}
 
@@
 
-Recall our implementations of Milner's cyclic scheduler in PCP ($\pcp{\sched}$, \cref{sec:pcp-example}) and PGV ($\pgv{\sched}$, \cref{sec:pgv-example}).
-The translation of $\pcp{\sched}$ is exactly $\pgv{sched}$ where all processes are child processes.
+Recall our implementations of Milner's cyclic scheduler in PCP ($\pcp{\tm{\sched}}$, \cref{sec:pcp-example}) and PGV ($\pgv{\tm{\sched}}$, \cref{sec:pgv-example}).
+The translation of $\pcp{\tm{\sched}}$ is exactly $\pgv{\tm{\sched}}$ where all processes are child processes.
 
 \subsubsection*{Type preservation}
-The translations $\tm{\cpgvC{\cdot}}$ and $\tm{\cpgvM{\cdot}}$ preserve typing along $\ty{\cpgvT{\cdot}}$:\simon{This reads weirdly---what's the intuition?}
-\begin{restatablelemma}{lempcptopgvtermspreservation}[Preservation, $\cpgvM{\cdot}$]
+The translations $\tm{\cpgvC{\cdot}}$ and $\tm{\cpgvM{\cdot}}$ are preserve typing.
+The proofs are standard inductions.
+The details can be found in~\cref{prf:lem-pcp-to-pgv-terms-preservation,prf:lem-pcp-to-pgv-terms-preservation}:
+\begin{compacttheorems}
+  \begin{restatablelemma}{lempcptopgvtermspreservation}[Preservation, ${\tm{\cpgvM{\cdot}}}$]
     \label{lem:pcp-to-pgv-terms-preservation}
-  \hfill\\%newline before theorem statement
-  If $\pcp{\seq{P}{\ty{\Gamma}}}$ in PCP,
-  then $\pgv{\tseq[\cs{p}]{\ty{\cpgvT{\Gamma}}}{\cpgvM{P}}{\tyunit}}$ in PGV.
+    If $\pcp{\seq{P}{\ty{\Gamma}}}$, then $\pgv{\tseq[\cs{p}]{\ty{\cpgvT{\Gamma}}}{\cpgvM{P}}{\tyunit}}$.
   \end{restatablelemma}
-\begin{proof}
-  By induction on the derivation of $\pcp{\seq{P}{\ty{\Gamma}}}$.
-  Details in \cref{prf:lem-pcp-to-pgv-terms-preservation}.
-\end{proof}
-\begin{restatabletheorem}{thmpcptopgvconfspreservation}[Preservation, $\cpgvC{\cdot}$]
+  \begin{restatabletheorem}{thmpcptopgvconfspreservation}[Preservation, ${\tm{\cpgvC{\cdot}}}$]
     \label{thm:pcp-to-pgv-confs-preservation}
-  \hfill\\%newline before theorem statement
-  If $\pcp{\seq{P}{\ty{\Gamma}}}$ in PCP,
-  then $\pgv{\cseq[\child]{\ty{\cpgvT{\Gamma}}}{\cpgvC{P}}}$ in PGV.
+    If $\pcp{\seq{P}{\ty{\Gamma}}}$, then $\pgv{\cseq[\child]{\ty{\cpgvT{\Gamma}}}{\cpgvC{P}}}$.
   \end{restatabletheorem}
-\begin{proof}
-  By induction on the derivation of $\pcp{\seq{P}{\ty{\Gamma}}}$.
-  Details in \cref{prf:lem-pcp-to-pgv-terms-preservation}.
-\end{proof}
+\end{compacttheorems}
 
@@
 First, we link our two translations, and show that $\tm{\cpgvM{P}}$ \emph{must} reduce to $\tm{\cpgvC{P}}$, \ie $\tm{\cpgvC{\cdot}}$ is equivalent to translating with $\tm{\cpgvM{\cdot}}$ and performing several reduction steps:
+\begin{compacttheorems}
   \begin{restatablelemma}{lempcptopgvcpgvMtocpgvC}
@@
   \end{proof}
-
   \begin{restatabletheorem}{thmpcptopgvoperationalcorrespondencecompleteness}%
-  [Operational Correspondence, Completeness, $\cpgvC{\cdot}$]
+    [Operational Correspondence, Completeness, ${\tm{\cpgvC{\cdot}}}$]
     \label{thm:pcp-to-pgv-operational-correspondence-completeness}
@@
     \LabTirName{E-LiftSC}. 
-  (Details in \cref{prf:thm-pcp-to-pgv-operational-correspondence-completeness}.)
+    Details in \cref{prf:thm-pcp-to-pgv-operational-correspondence-completeness}.
   \end{proof}
-
   \begin{restatabletheorem}{thmpcptopgvoperationalcorrespondencesoundness}%
-  [Operational Correspondence, Soundness, $\cpgvC{\cdot}$]
+    [Operational Correspondence, Soundness, ${\tm{\cpgvC{\cdot}}}$]
     \label{thm:pcp-to-pgv-operational-correspondence-soundness}
@@
   \end{proof}
+\end{compacttheorems}
 

### 2020-10-15 priority-gv.tex
Created compacttheorems environment for compact theorems in sequence.

@@
 Every other agent $\tm{\agent{i}}$, $2\leq{i}\leq{n}$ waits on $\tm{c'_{i-1}}$ for the signal to start.
-The dotted lines represent the optional transfer of data between processes.
+% The dotted lines represent the optional transfer of data between processes.
 With three processes, the scheduler looks as follows:\simon{Sorry, I found this diagram super confusing\dots Is there any way you could make it look more cyclic?
 Also, have the convention of a single annotation per channel, with the convention that the greater agent / proc gets the primed end of the previous channel?}
-\input{fig/pgv-scheduler.tex}
+% \input{fig/pgv-scheduler.tex}
 We implement the scheduler as follows, using $\tm{\prod_{i}\conf{C}_i}$ to denote the parallel composition of the processes $\tm{\conf{C}_i}$, and $\tm{\plug{M}{N}}$ to denote the plugging of $\tm{N}$ in the one-hole term context $\tm{M}$:
@@
 \[
-\begin{array}{lcl}
-  \ty{\co{\tysend[\cs{o}]{T}{S}}} & = & \ty{\tyrecv[\cs{o}]{T}{\co{S}}} \\
-  \ty{\co{\tyrecv[\cs{o}]{T}{S}}} & = & \ty{\tysend[\cs{o}]{T}{\co{S}}}
-\end{array}
+  \ty{\co{\tysend[\cs{o}]{T}{S}}} = \ty{\tyrecv[\cs{o}]{T}{\co{S}}}
   \qquad
-\begin{array}{lcl}
-  \ty{\co{\tyends[\cs{o}]}} & = & \ty{\co{\tyendr[\cs{o}]}} \\
-  \ty{\co{\tyendr[\cs{o}]}} & = & \ty{\co{\tyends[\cs{o}]}}
-\end{array}
+  \ty{\co{\tyrecv[\cs{o}]{T}{S}}} = \ty{\tysend[\cs{o}]{T}{\co{S}}}
+  \qquad
+  \ty{\co{\tyends[\cs{o}]}} = \ty{\co{\tyendr[\cs{o}]}}
+  \qquad
+  \ty{\co{\tyendr[\cs{o}]}} = \ty{\co{\tyends[\cs{o}]}}
 \]
@@
 \[
-\begin{array}{lcl}
-  \pr(\ty{\tysend[\cs{o}]{T}{S}})            & = & \cs{o} \\
-  \pr(\ty{\tyrecv[\cs{o}]{T}{S}})            & = & \cs{o} \\
-\end{array}
+  \pr(\ty{\tysend[\cs{o}]{T}{S}}) = \cs{o}
   \qquad
-\begin{array}{lcl}
-  \pr(\ty{\tyends[\cs{o}]})                  & = & \cs{o} \\
-  \pr(\ty{\tyendr[\cs{o}]})                  & = & \cs{o} \\
-\end{array}
+  \pr(\ty{\tyrecv[\cs{o}]{T}{S}}) = \cs{o}
+  \qquad
+  \pr(\ty{\tyends[\cs{o}]})       = \cs{o}
+  \qquad
+  \pr(\ty{\tyendr[\cs{o}]})       = \cs{o}
 \]
@@
   & \sep      & \tm{\unit}
-    \sep        \tm{\letunit{M}{N}} \\
-  & \sep      & \tm{\pair{M}{N}}
+    \sep        \tm{\letunit{M}{N}}
+    \sep        \tm{\pair{M}{N}}
     \sep        \tm{\letpair{x}{y}{M}{N}} \\
@@
     \sep        \tm{\inr{M}}
-    \sep        \tm{\casesum{L}{x}{M}{y}{N}} \\
-  & \sep      & \tm{\absurd{M}} \\
+    \sep        \tm{\casesum{L}{x}{M}{y}{N}}
+    \sep        \tm{\absurd{M}} \\
   \tm{K}
@@
 
-We frequently use syntactic sugar to make terms more readable: we write $\tm{\andthen{M}{N}}$ in place of $\tm{\letunit{M}{N}}$, $\tm{\letbind{x}{M}{N}}$ in place of $\tm{(\lambda x.N)\;M}$, and pattern matching lambdas $\tm{\lambda\unit.M}$ and $\tm{\lambda\pair{x}{y}.M}$ in place of $\tm{\lambda z.\letunit{z}{M}}$ and $\tm{\lambda z.\letpair{x}{y}{z}{M}}$.
-Finally, we recover $\tm{\fork}$:
-\[
-  \tm{\fork} \defeq \tm{\lambda x.\letpair{y}{z}{\new\;\unit}{\andthen{\spawn\;{(\lambda\unit.x\;y)}}{z}}}
-\]
+We use syntactic sugar to make terms more readable: we write $\tm{\andthen{M}{N}}$ in place of $\tm{\letunit{M}{N}}$, $\tm{\letbind{x}{M}{N}}$ in place of $\tm{(\lambda x.N)\;M}$, and pattern matching functions $\tm{\lambda\unit.M}$ in place of $\tm{\lambda z.\letunit{z}{M}}$ and $\tm{\lambda\pair{x}{y}.M}$ in place of $\tm{\lambda z.\letpair{x}{y}{z}{M}}$.
+% We recover $\tm{\fork}$ as $\tm{\lambda x.\letpair{y}{z}{\new\;\unit}{\andthen{\spawn\;{(\lambda\unit.x\;y)}}{z}}}$.
 
@@
 
-Most of the subject reduction proof is standard.
-Novel is that we must show that substitution preserves priority constraints.
-For this, it helps to prove~\cref{lem:pgv-value-done}, which shows that values have finished all their communication actions, and that any priorities present in the type of the value come from the typing environment.
+We must show that substitution preserves priority constraints.
+For this, we prove~\cref{lem:pgv-value-done}, which shows that values have finished all their communication, and that any priorities in the type of the value come from the typing environment.
 \begin{restatablelemma}{lempgvvaluedone}
@@
 The details can be found in \cref{prf:lem-pgv-subject-reduction-terms,prf:lem-pgv-subject-congruence,prf:thm-pgv-subject-reduction-confs}.
+\begin{compacttheorems}
   \begin{restatablelemma}{lempgvsubjectreductionterms}[Subject Reduction, $\tred$]
@@
 The standard GV canonical form, alternating $\nu$-binders and their corresponding parallel compositions, does not work for PGV, since multiple channels may be split across a single parallel composition:
+\begin{compacttheorems}
   \begin{definition}[Canonical Forms]
@@
 Furthermore, no term $\tm{M_i}$ can perform a communication action by itself, since those are excluded by the definition of actions:
+\begin{compacttheorems}
   \begin{definition}[Normal Forms]
@@
   \input{prf/thm-pgv-closed-progress-confs}
+\end{compacttheorems}
 

### 2020-10-15 priority-cp.tex
Minor rewrites.

@@
 \subsubsection*{Duality}
-Duality on types is an involutive function which preserves the priorities of types:
+Duality is an involutive function on types which preserves priorities:
 \[
   \usingnamespace{pcp}
+  \setlength{\arraycolsep}{1pt}
   \begin{array}{lcl}
     \ty{\co{(\tyone[\cs{o}])}} & = & \ty{\tybot[\cs{o}]} \\
-    \ty{\co{(\tybot[\cs{o}])}} & = & \ty{\tyone[\cs{o}]} \\
+    \ty{\co{(\tybot[\cs{o}])}} & = & \ty{\tyone[\cs{o}]}
+  \end{array}
+  \quad
+  \begin{array}{lcl}
+    \ty{\co{(\tytens[\cs{o}]{A}{B})}} & = & \ty{\typarr[\cs{o}]{\co{A}}{\co{B}}} \\
+    \ty{\co{(\typarr[\cs{o}]{A}{B})}} & = & \ty{\tytens[\cs{o}]{\co{A}}{\co{B}}}
+  \end{array}
+  \quad
+  \begin{array}{lcl}
     \ty{\co{(\tynil[\cs{o}])}} & = & \ty{\tytop[\cs{o}]} \\
@@
   \end{array}
-  \qquad
+  \quad
   \begin{array}{lcl}
-    \ty{\co{(\tytens[\cs{o}]{A}{B})}} & = & \ty{\typarr[\cs{o}]{\co{A}}{\co{B}}} \\
-    \ty{\co{(\typarr[\cs{o}]{A}{B})}} & = & \ty{\tytens[\cs{o}]{\co{A}}{\co{B}}} \\
     \ty{\co{(\typlus[\cs{o}]{A}{B})}} & = & \ty{\tywith[\cs{o}]{\co{A}}{\co{B}}} \\
@@
   \usingnamespace{pcp}
+  \setlength{\arraycolsep}{1pt}
   \begin{array}{lclclcl}
     \pr(\ty{\tyone[\cs{o}]})        & = & \cs{o} \\
-    \pr(\ty{\tybot[\cs{o}]})        & = & \cs{o}  \\
-    \pr(\ty{\tynil[\cs{o}]})        & = & \cs{o}  \\
-    \pr(\ty{\tytop[\cs{o}]})        & = & \cs{o}  \\
-    \\
-    \minpr(\ty{\emptyenv})          & = & \cs{\ptop}
+    \pr(\ty{\tybot[\cs{o}]})        & = & \cs{o}
   \end{array}
@@
     \pr(\ty{\tytens[\cs{o}]{A}{B}}) & = & \cs{o} \\
-    \pr(\ty{\typarr[\cs{o}]{A}{B}}) & = & \cs{o}  \\
+    \pr(\ty{\typarr[\cs{o}]{A}{B}}) & = & \cs{o}
+  \end{array}
+  \qquad
+  \begin{array}{lclclcl}
+    \pr(\ty{\tynil[\cs{o}]})        & = & \cs{o} \\
+    \pr(\ty{\tytop[\cs{o}]})        & = & \cs{o}
+  \end{array}
+  \qquad
+  \begin{array}{lclclcl}
     \pr(\ty{\typlus[\cs{o}]{A}{B}}) & = & \cs{o} \\
-    \pr(\ty{\tywith[\cs{o}]{A}{B}}) & = & \cs{o}  \\
-    \\
-    \minpr(\ty{\Gamma},\tmty{x}{A}) & = & \minpr(\ty{\Gamma})\sqcap\minpr(\ty{A})
+    \pr(\ty{\tywith[\cs{o}]{A}{B}}) & = & \cs{o} 
   \end{array}
+\]
+\[
+    \minpr(\ty{\emptyenv})          = \cs{\ptop}
+    \qquad
+    \minpr(\ty{\Gamma},\tmty{x}{A}) = \minpr(\ty{\Gamma})\sqcap\minpr(\ty{A})
 \]
@@
     \label{lem:pcp-subject-congruence}
-    \hfill\\%newline before theorem statement
     If $\pcp{\seq{P}{\ty{\Gamma}}}$ and $\pcp{\tm{P}\equiv\tm{Q}}$,
@@
     \label{thm:pcp-subject-reduction}
-    \hfill\\%newline before theorem statement
     If $\pcp{\seq{P}{\ty{\Gamma}}}$ and $\pcp{\tm{P}\red\tm{Q}}$, then $\pcp{\seq{Q}{\ty{\Gamma}}}$.
@@
 \subsubsection*{Progress}
-Reduction in our version of PCP satisfies progress. For simplicity, we only present closed progress:
-\begin{restatabletheorem}{thmpcpclosedprogress}[Closed Progress, $\pcp{\red}$]
+Reduction in our version of PCP satisfies progress:
+\begin{restatabletheorem}{thmpcpclosedprogress}[Progress, $\pcp{\red}$]
   \label{thm:pcp-closed-progress}
@@
 \begin{proof}
-    Our proof follows the same reasoning by Kobayashi~\cite{kobayashi06} used in the proof of deadlock freedom for closed processes (Theorem 2).
+    Our proof follows the reasoning by Kobayashi~\cite{kobayashi06} used in the proof of deadlock freedom for closed processes (Theorem 2).
 The details can be found in~\Cref{prf:thm-pcp-closed-progress}.
@@
   \begin{proof}
-    By induction on the structure of $\pcp{\tm{P}}$.
-    Details in \cref{prf:lem-pcp-to-pgv-cpgvM-to-cpgvC}.
+    By induction on the structure of $\pcp{\tm{P}}$. Details in \cref{prf:lem-pcp-to-pgv-cpgvM-to-cpgvC}.
   \end{proof}
@@
     By induction on the derivation of $\pcp{\tm{P}\red\tm{Q}}$.
-    We apply \cref{lem:pcp-to-pgv-cpgvM-to-cpgvC} for the cases in which
-    the translation switches from $\tm{\cpgvC{\cdot}}$ to $\tm{\cpgvM{\cdot}}$,
-    \ie under actions.
-    We apply \cref{lem:pcp-to-pgv-confs-operational-correspondence-equiv} for
-    \LabTirName{E-LiftSC}. 
+We apply \cref{lem:pcp-to-pgv-cpgvM-to-cpgvC} for the cases in which the translation switches from $\tm{\cpgvC{\cdot}}$ to $\tm{\cpgvM{\cdot}}$, \ie under actions.
+We apply \cref{lem:pcp-to-pgv-confs-operational-correspondence-equiv} for \LabTirName{E-LiftSC}.
 Details in \cref{prf:thm-pcp-to-pgv-operational-correspondence-completeness}.
@@
     \hfill\\%newline before theorem statement
-    If $\pcp{\seq{P}{\ty{\Gamma}}}$ and $\pgv{\tm{\cpgvC{P}}\cred\tm{\conf{C}}}$,
-    there exists a $\tm{Q}$ such that $\pcp{\tm{P}\red^+\tm{Q}}$ and $\pgv{\tm{\conf{C}}\cred^\star\tm{\cpgvC{Q}}}$
+    If $\pcp{\seq{P}{\ty{\Gamma}}}$ and $\pgv{\tm{\cpgvC{P}}\cred\tm{\conf{C}}}$, there exists a $\tm{Q}$ such that $\pcp{\tm{P}\red^+\tm{Q}}$ and $\pgv{\tm{\conf{C}}\cred^\star\tm{\cpgvC{Q}}}$
   \end{restatabletheorem}

### 2020-10-15 priority-gv.tex
Minor rewrites.

@@
 For this, we prove~\cref{lem:pgv-value-done}, which shows that values have finished all their communication, and that any priorities in the type of the value come from the typing environment.
+\begin{compacttheorems}
   \begin{restatablelemma}{lempgvvaluedone}
@@
   \end{proof}
+\end{compacttheorems}
 The proofs for subject congruence and subject reduction are standard inductions.
@@
 Progress for the term language is standard for GV, and deviates from progress for linear $\lambda$-calculus only in that terms may reduce to values or \emph{ready terms}:
-\begin{restatablelemma}{lempgvopenprogressterms}[Open Progress, $\tred$]
+\begin{restatablelemma}{lempgvopenprogressterms}[Progress, $\tred$]
   \label{lem:pgv-open-progress-terms}
@@
   \end{definition}
-  \begin{restatabletheorem}{thmpgvclosedprogressconfs}[Closed Progress, $\cred$]
+  \begin{restatabletheorem}{thmpgvclosedprogressconfs}[Progress, $\cred$]
     \label{lem:pgv-closed-progress-confs}
-    If $\cseq[\main]{\emptyenv}{\conf{C}}$ and $\tm{\conf{C}}$ is in canonical form, then either $\tm{\conf{C}}\cred\tm{\conf{D}}$ for some $\tm{\conf{D}}$; or $\tm{\conf{C}}$ is equivalent to a configuration in normal form.
+    If $\cseq[\main]{\emptyenv}{\conf{C}}$ and $\tm{\conf{C}}$ is in canonical form, then either $\tm{\conf{C}}\cred\tm{\conf{D}}$ for some $\tm{\conf{D}}$; or $\tm{\conf{C}\equiv\conf{D}}$ for some $\tm{\conf{D}}$ in normal form.
   \end{restatabletheorem}

### 2020-10-15 priority-gv.tex
Moved closed progress to the appendices.

@@
 Canonical forms deviate from those for GV, in that we opt to move all $\nu$-binders to the top.
-The standard GV canonical form, alternating $\nu$-binders and their corresponding parallel compositions, does not work for PGV, since multiple channels may be split across a single parallel composition:
+The standard GV canonical form, alternating $\nu$-binders and their corresponding parallel compositions, does not work for PGV, since multiple channels may be split across a single parallel composition.
+
+A~configuration either reduces, or it is equivalent to configuration in normal form.
+Crucial to the normal form is that each term $\tm{M_i}$ is blocked on the corresponding channel $\tm{x_i}$, and hence no two terms act on dual endpoints.
+Furthermore, no term $\tm{M_i}$ can perform a communication action by itself, since those are excluded by the definition of actions.
+Finally, as a corollary, we get that well-typed terms which do not return endpoints return \emph{just} a value:
+
 \begin{compacttheorems}
@@
   \end{definition}
-  \begin{restatablelemma}{lempgvcanonicalforms}[Canonical Forms]
-    \label{lem:pgv-canonical-forms}
-    If $\cseq[\main]{\ty{\Gamma}}{\conf{C}}$, there exists some $\tm{\conf{D}}$ such that $\tm{\conf{C}}\equiv\tm{\conf{D}}$ and $\tm{\conf{D}}$ is in canonical form.
-  \end{restatablelemma}
-  \begin{proof}
-    We move any $\nu$-binders to the top using \LabTirName{SC-ResExt}, discard any superfluous occurrences of $\tm{\child\;\unit}$ using \LabTirName{SC-ParNil}, and move the main thread to the rightmost position using \LabTirName{SC-ParComm} and \LabTirName{SC-ParAssoc}.
-  \end{proof}
-\end{compacttheorems}
-
-A~configuration either reduces, or it is equivalent to configuration in normal form.
-Crucial to the normal form is that each term $\tm{M_i}$ is blocked on the corresponding channel $\tm{x_i}$, and hence no two terms act on dual endpoints.
-Furthermore, no term $\tm{M_i}$ can perform a communication action by itself, since those are excluded by the definition of actions:
-\begin{compacttheorems}
   \begin{definition}[Normal Forms]
@@
   \begin{restatabletheorem}{thmpgvclosedprogressconfs}[Progress, $\cred$]
-    \label{lem:pgv-closed-progress-confs}
+    \label{thm:pgv-closed-progress-confs}
     If $\cseq[\main]{\emptyenv}{\conf{C}}$ and $\tm{\conf{C}}$ is in canonical form, then either $\tm{\conf{C}}\cred\tm{\conf{D}}$ for some $\tm{\conf{D}}$; or $\tm{\conf{C}\equiv\conf{D}}$ for some $\tm{\conf{D}}$ in normal form.
   \end{restatabletheorem}
-  \input{prf/thm-pgv-closed-progress-confs}
-\end{compacttheorems}
-
-Finally, as a corollary from closed progress, we get that well-typed terms which do not return endpoints must return \emph{only} a value:
+  \begin{proof}
+    Our proof follows the reasoning by Kobayashi~\cite{kobayashi06} used in the proof of deadlock freedom for closed processes (Theorem 2). Details are in~\Cref{prf:thm-pgv-closed-progress-confs}.
+  \end{proof}
   \begin{corollary}
@@
   \end{corollary}
+\end{compacttheorems}
 

### 2020-10-15 introduction.tex
bad reference solved

@@
 \item \emph{Priority GV} We present Priority GV (PGV) in~\Cref{sec:pgv}; we present the syntax of types and terms, operational semantics and typing rules.
-We show our type system for PGV is sound by proving subject reduction (\Cref{thm:pgv-subject-reduction-confs}) and closed progress (\Cref{lem:pgv-closed-progress-confs}).
+We show our type system for PGV is sound by proving subject reduction (\Cref{thm:pgv-subject-reduction-confs}) and closed progress (\Cref{thm:pgv-closed-progress-confs}).
 \item \emph{Milner's Cyclic Scheduler in PGV} We present, for the first time, Milner's Cyclic Scheduler~\cite{milner89} (\Cref{sec:pgv-example}) as a GV term.

### 2020-10-15 priority-cp.tex
Compressed PCP metatheory.

@@
 
-\subsubsection*{Subject reduction}
-Reduction in our version of PCP preserve types. Since our reduction relation is a strict subset of the reduction relation in the original~\cite{dardhagay18}, we defer to their proofs:
-\begin{compacttheorems}
-  \begin{restatablelemma}{lempcpsubjectcongruence}[Subject Congruence, $\pcp{\equiv}$]
-    \label{lem:pcp-subject-congruence}
-    If $\pcp{\seq{P}{\ty{\Gamma}}}$ and $\pcp{\tm{P}\equiv\tm{Q}}$,
-    then $\pcp{\seq{Q}{\ty{\Gamma}}}$.
-  \end{restatablelemma}
-  \begin{restatabletheorem}{thmpcpsubjectreduction}[Subject Reduction, $\pcp{\red}$]
-    \label{thm:pcp-subject-reduction}
-    If $\pcp{\seq{P}{\ty{\Gamma}}}$ and $\pcp{\tm{P}\red\tm{Q}}$, then $\pcp{\seq{Q}{\ty{\Gamma}}}$.
-  \end{restatabletheorem}
-\end{compacttheorems}
-
-\subsubsection*{Progress}
-Reduction in our version of PCP satisfies progress:
-\begin{restatabletheorem}{thmpcpclosedprogress}[Progress, $\pcp{\red}$]
-  \label{thm:pcp-closed-progress}
-  \hfill\\%newline before theorem statement
-  If $\pcp{\seq{P}{\emptyenv}}$, then either $\pcp{\tm{P}=\tm{\halt}}$ or there exists a $\pcp{\tm{Q}}$ such that $\pcp{\tm{P}\red\tm{Q}}$.
-\end{restatabletheorem}
-\begin{proof}
-    Our proof follows the reasoning by Kobayashi~\cite{kobayashi06} used in the proof of deadlock freedom for closed processes (Theorem 2).
-The details can be found in~\Cref{prf:thm-pcp-closed-progress}.
-\end{proof}
+\subsubsection*{Metatheory}
+Reduction in our version of PCP preserve types.
+Since our reduction relation is a strict subset of the reduction relation in the original~\cite{dardhagay18}, we defer to their proofs.
+We prove progress for our version of PCP, see~\cref{prf:thm-pcp-closed-progress}.
 
 \subsubsection*{Translation}
-The translation on types from PCP to PGV, $\ty{\cpgvT{\cdot}}$, is:
+We illustrate the relation between PCP and PGV by giving a translation from PCP to PGV. The translation on types is defined as follows:
 \[
@@
 
-\subsubsection*{Type preservation}
 The translations $\tm{\cpgvC{\cdot}}$ and $\tm{\cpgvM{\cdot}}$ are preserve typing.
@@
 
-\subsubsection*{Operational correspondence}
-Finally, we show that PCP and PGV are in operational correspondence along our translation.
-First, we link our two translations, and show that $\tm{\cpgvM{P}}$ \emph{must} reduce to $\tm{\cpgvC{P}}$, \ie $\tm{\cpgvC{\cdot}}$ is equivalent to translating with $\tm{\cpgvM{\cdot}}$ and performing several reduction steps:
+Our translation satisfies operational correspondence.
+We link our translations, and show that $\tm{\cpgvM{P}}$ \emph{must} reduce to $\tm{\cpgvC{P}}$, \ie $\tm{\cpgvC{\cdot}}$ is equivalent to translating with $\tm{\cpgvM{\cdot}}$ and performing several reduction steps.
+We use this fact to prove operational correspondence.
+The details are in~\cref{prf:lem-pcp-to-pgv-cpgvM-to-cpgvC,prf:thm-pcp-to-pgv-operational-correspondence-completeness,prf:thm-pcp-to-pgv-operational-correspondence-soundness}:
 \begin{compacttheorems}
@@
   \end{restatablelemma}
-  \begin{proof}
-    By induction on the structure of $\pcp{\tm{P}}$. Details in \cref{prf:lem-pcp-to-pgv-cpgvM-to-cpgvC}.
-  \end{proof}
   \begin{restatabletheorem}{thmpcptopgvoperationalcorrespondencecompleteness}%
@@
     If $\pcp{\seq{P}{\ty{\Gamma}}}$ and $\pcp{\tm{P}\red\tm{Q}}$,
-    then $\pgv{\tm{\cpgvC{P}}\cred^+\tm{\cpgvC{Q}}}$.
+    then $\pgv{\tm{\cpgvC{P}}\cred^+\tm{\cpgvC{Q}}}$. Details in~\cref{prf:lem-pcp-to-pgv-cpgvM-to-cpgvC}.
   \end{restatabletheorem}
-  \begin{proof}
-    By induction on the derivation of $\pcp{\tm{P}\red\tm{Q}}$.
+  % \begin{proof}
+  %   By induction on the derivation of $\pcp{\tm{P}\red\tm{Q}}$.
 We apply \cref{lem:pcp-to-pgv-cpgvM-to-cpgvC} for the cases in which the translation switches from $\tm{\cpgvC{\cdot}}$ to $\tm{\cpgvM{\cdot}}$, \ie under actions.
@@
 Details in \cref{prf:thm-pcp-to-pgv-operational-correspondence-completeness}.
-  \end{proof}
+  % \end{proof}
   \begin{restatabletheorem}{thmpcptopgvoperationalcorrespondencesoundness}%
@@
   \end{restatabletheorem}
-  \begin{proof}
-    By induction on the derivation of $\pgv{\tm{\cpgvC{P}}\cred\tm{\conf{C}}}$.\simon{Surprising that you can get it by the derivation of the reduction relation---normally, 'soundness' would be proved by shunting into a canonical form and then proceeding by induction the structure of the term / induction on typing derivations.}
-    Details in \cref{prf:thm-pcp-to-pgv-operational-correspondence-soundness}.
-  \end{proof}
+  % \begin{proof}
+  %   By induction on the derivation of $\pgv{\tm{\cpgvC{P}}\cred\tm{\conf{C}}}$.\simon{Surprising that you can get it by the derivation of the reduction relation---normally, 'soundness' would be proved by shunting into a canonical form and then proceeding by induction the structure of the term / induction on typing derivations.}
+  %   Details in \cref{prf:thm-pcp-to-pgv-operational-correspondence-soundness}.
+  % \end{proof}
 \end{compacttheorems}

### 2020-10-15 introduction.tex
Changes to contributions.

@@
 \begin{itemize}
-\item \emph{Priority GV} We present Priority GV (PGV) in~\Cref{sec:pgv}; we present the syntax of types and terms, operational semantics and typing rules.
-We show our type system for PGV is sound by proving subject reduction (\Cref{thm:pgv-subject-reduction-confs}) and closed progress (\Cref{thm:pgv-closed-progress-confs}).
-\item \emph{Milner's Cyclic Scheduler in PGV} We present, for the first time, Milner's Cyclic Scheduler~\cite{milner89} (\Cref{sec:pgv-example}) as a GV term.
-This example illustrates how a cyclic structure can be expressed now in GV thanks to the separation of channel creation and thread spawning.
-\item \emph{Updated Priority CP} We present an updated version of Priority CP (PCP)~\cite{dardhagay18} in \Cref{sec:pcp}; we remove commuting conversions and move away from a reduction relation as cycle elimination, and towards a reduction relation based on structural congruence, as one would expect in a process calculus.
-We show Milner's Cyclic Scheduler in PCP (in \Cref{sec:pcp-example}) originally presented in~\cite{dardhagay18}.
-\item \emph{Encoding PCP into PGV} We present an encoding of PCP into PGV in \Cref{sec:pcp}.
-We prove the encoding is sound and complete with respect to typing (\Cref{thm:pcp-to-pgv-confs-preservation}) and operational semantics (\Cref{thm:pcp-to-pgv-operational-correspondence-soundness} and \Cref{thm:pcp-to-pgv-operational-correspondence-completeness}).
-Finally, using our encoding, we can encode Milner's Cyclic Scheduler in PCP (\Cref{sec:pcp-example}) resulting in the Scheduler in PGV (\Cref{sec:pgv-example}).
+\item \emph{Priority GV} We present Priority GV (PGV) in~\cref{sec:pgv}, and prove subject reduction (\cref{thm:pgv-subject-reduction-confs}) and progress (\cref{thm:pgv-closed-progress-confs}).
+\item \emph{Milner's Cyclic Scheduler} To illustrate newfound expressiveness of PGV, we present Milner's Cyclic Scheduler~\cite{milner89} (\cref{sec:pgv-example}).
+\item \emph{Updated Priority CP} We present an updated version of Priority CP (PCP)~\cite{dardhagay18} in \cref{sec:pcp}.
+We remove commuting conversions and move away from reduction as cut elimination, towards reduction based on structural congruence, as one expects in a process calculus.
+We present Milner's Cyclic Scheduler in PCP (\cref{sec:pcp-example}).
+\item \emph{Translation from PCP to PGV} We present a translation from PCP in PGV in~\cref{sec:pcp}.
+We prove the translation preserves typing (\cref{thm:pcp-to-pgv-confs-preservation}) and satisfies operational correspondence (\cref{thm:pcp-to-pgv-operational-correspondence-soundness,thm:pcp-to-pgv-operational-correspondence-completeness}), and show that the scheduler in PCP (\cref{sec:pcp-example}) translates to the scheduler in PGV (\cref{sec:pgv-example}).
 \end{itemize}

### 2020-10-15 priority-gv.tex
Tiny change.

@@
 % \input{fig/pgv-scheduler.tex}
-We implement the scheduler as follows, using $\tm{\prod_{i}\conf{C}_i}$ to denote the parallel composition of the processes $\tm{\conf{C}_i}$, and $\tm{\plug{M}{N}}$ to denote the plugging of $\tm{N}$ in the one-hole term context $\tm{M}$:
+We implement the scheduler as follows, using $\tm{\prod_{i}\conf{C}_i}$ to denote the parallel composition of the processes $\tm{\conf{C}_i}$, and $\tm{\plug{M}{N}}$ to denote the plugging of $\tm{N}$ in the one-hole term-context $\tm{M}$:
 \[

### 2020-10-15 priority-gv.tex
Tiny change around PGV scheduler.
Tiny change around session types/types.

@@
 \]
-Each of the channels in the scheduler is has the terminated type, and is merely used to synchronise.
-For simplicity, we let $\tm{\sched}$ be a configuration. However, we could equivalently have represented it as a function which takes three processes as arguments, and creates this configuration.
+The channels in the scheduler all have the terminated type, and are merely used to synchronise.
+For simplicity, we let $\tm{\sched}$ be a configuration.
+However, we could equivalently have represented it as a function which takes three processes as arguments, and creates this configuration.
 
-\subsubsection*{Types}
-Types in PGV are divided into session types ($\ty{S}$) and types ($\ty{T}$, $\ty{U}$).
-%
-Session types are defined by the following grammar:
+\subsubsection*{Session types}
+Session ($\ty{S}$) types are defined by the following grammar:
 \[
@@
 
-Types are defined by the following grammar:
+\subsubsection*{Types}
+Types ($\ty{T}$, $\ty{U}$) are defined by the following grammar:
 \[

### 2020-10-15 priority-gv.tex
Moved ref to figure in PGV typing.

@@
 \subsubsection*{Typing}
-\Cref{fig:pgv-typing} gives the typing rules for PGV.
-
 Terms are typed by a judgement $\tseq[\cs{p}]{\ty{\Gamma}}{M}{T}$, which states that ``a term $\tm{M}$ has type $\ty{T}$ and an upper bound on its priority $\cs{p}$ under the typing environment $\ty{\Gamma}$''.
@@
 The constants $\tm{K}$ are typed using type schemas, and embedded using \LabTirName{T-Const}.
+\Cref{fig:pgv-typing} gives the typing rules for PGV.
 

### 2020-10-15 priority-gv.tex
Update priority-gv.tex

--- 2020-10-15 priority-gv.tex
@@
 
-This section presents Priority GV (PGV), a session-typed functional language based on GV~\cite{wadler15,lindleymorris15}, but with a twist: it enforces deadlock freedom through priorities \`{a} la Kobayashi~\cite{kobayashi06,padovaninovara15}.
+We present Priority GV (PGV), a session-typed functional language based on GV~\cite{wadler15,lindleymorris15} which uses priorities \`{a} la Kobayashi~\cite{kobayashi06,padovaninovara15} to enforce deadlock freedom.
+%The separation of channel creation and thread spawning in GV only allows tree-shaped communication structures, so it is trivially free from cycles, and therefore deadlocks.
+For e.g., if some term $\tm{L}$ has the means to communicate with both $\tm{M}$ and $\tm{N}$, then $\tm{M}$ and $\tm{N}$ cannot communicate amongst themselves.
+However, as with CP, this rules out \emph{all} cyclic communication structures, not just those with deadlocks.
+Priority GV offers a more fine-grained analysis of communication structures, and by separating channel creation form thread spawning it allows cyclic structures.
+We illustrate PGV with Milner's cyclic scheduler~\cite{milner89}.
 
-GV enforces deadlock freedom with a syntactic restriction: it fuses channel creation and thread spawning into a single function, called fork\simon{It feels like you've said this a *lot* at this point}.
-This guarantees that if some term $\tm{L}$ has the means to communicate with both $\tm{M}$ and $\tm{N}$, then $\tm{M}$ and $\tm{N}$ cannot communicate amongst themselves.
-In other words, GV only allows tree-shaped communication structures, so it is trivially free from cycles, and therefore deadlocks.
-However, this is overly restrictive, as it rules out \emph{all} cyclic communication structures, not just those with deadlocks.
-
-Priority GV offers a more fine-grained analysis of the communication structure, and allows cyclic communication structures.
-To illustrate this, we start by examining a variant of Milner's cyclic scheduler in PGV~\cite{milner89}.
-
-\subsubsection*{Example}\label{sec:pgv-example}
-Suppose a set of processes $\tm{\proc{i}}$, $1\leq{i}\leq{n}$, should be scheduled to perform some tasks in cyclic order, starting with $\tm{\proc0}$, ending with $\tm{\proc{n}}$, and notifying $\tm{\proc0}$ when all processes have finished.
-
-Our scheduler consists of set of agents $\tm{\agent{i}}$, $1\leq{i}\leq{n}$, each representing their respective process.
-Each process $\tm{\proc{i}}$ waits for the signal to start their task on $\tm{a'_i}$, and signals completion on $\tm{b'_i}$.
-Each agent signals their process to start on $\tm{a_i}$, waits for their process to finish on $\tm{b_i}$, and then signals for the next agent to continue on $\tm{c_i}$.
-The agent $\tm{\agent0}$ initiates, then waits for every other process to finish, and signals $\tm{\proc0}$ on $\tm{d}$.
-Every other agent $\tm{\agent{i}}$, $2\leq{i}\leq{n}$ waits on $\tm{c'_{i-1}}$ for the signal to start.
+\begin{example}[Milner's Cyclic Scheduler in PGV]\label{sec:pgv-example}
+Suppose a set of processes $\tm{\proc{i}}$, $1\leq{i}\leq{n}$, is scheduled to perform some tasks in cyclic order, starting with $\tm{\proc0}$, ending with $\tm{\proc{n}}$, and notifying $\tm{\proc0}$ when all processes have finished.
 % The dotted lines represent the optional transfer of data between processes.
@@
 % \input{fig/pgv-scheduler.tex}
-We implement the scheduler as follows, using $\tm{\prod_{i}\conf{C}_i}$ to denote the parallel composition of the processes $\tm{\conf{C}_i}$, and $\tm{\plug{M}{N}}$ to denote the plugging of $\tm{N}$ in the one-hole term-context $\tm{M}$:
+We implement the scheduler as follows, using $\tm{\prod_{i}\conf{C}_i}$ to denote the parallel composition of the processes $\tm{\conf{C}_i}$, and $\tm{\plug{M}{N}}$ to denote the plugging of $\tm{N}$ in the one-hole term context $\tm{M}$:
 \[
@@
 \]
-Each of the channels in the scheduler is has the terminated type, and is merely used to synchronise.
+Our scheduler $\tm{\sched}$ consists of set of agents $\tm{\agent{i}}$, $1\leq{i}\leq{n}$, each representing their respective process.
+Each process $\tm{\proc{i}}$ waits for the signal to start their task on $\tm{a'_i}$, and signals completion on $\tm{b'_i}$.
+Each agent signals their process to start on $\tm{a_i}$, waits for their process to finish on $\tm{b_i}$, and then signals for the next agent to continue on $\tm{c_i}$.
+The agent $\tm{\agent0}$ initiates, then waits for every other process to finish, and signals $\tm{\proc0}$ on $\tm{d}$.
+Every other agent $\tm{\agent{i}}$, $2\leq{i}\leq{n}$ waits on $\tm{c'_{i-1}}$ for the signal to start.
+Each of the channels in the scheduler is of a terminated type, and is merely used to synchronise.
 For simplicity, we let $\tm{\sched}$ be a configuration. However, we could equivalently have represented it as a function which takes three processes as arguments, and creates this configuration.
+\end{example}
 

### 2020-10-15 priority-cp.tex

@@
 Our PCP scheduler has exactly the same behaviour as the version presented in~\cref{sec:pgv-example}.
-We use $\pcp{\tm{\prod_{i}P_i}}$ to denote the parallel composition of the processes $\pcp{\tm{P_i}}$, and $\pcp{\plug{P}{Q}}$ to denote the plugging of $\pcp{\tm{Q}}$ in the one-hole process context $\pcp{\tm{P}}$:
+We use $\pcp{\tm{\prod_{i}P_i}}$ to denote the parallel composition of the processes $\pcp{\tm{P_i}}$, and $\pcp{\plug{P}{Q}}$ to denote the plugging of $\pcp{\tm{Q}}$ in the one-hole process-context $\pcp{\tm{P}}$:
 \[

### 2020-10-15 priority-gv.tex

@@
 
-\subsubsection*{Types}
-Types in PGV are divided into session types ($\ty{S}$) and types ($\ty{T}$, $\ty{U}$).
-%
-Session types are defined by the following grammar:
+\subsubsection*{Session types}
+Session ($\ty{S}$) types are defined by the following grammar:
 \[

### 2020-10-15 priority-gv.tex
Handled one of Simon's worries.

@@
 
-Configurations are typed by a judgement $\cseq[\phi]{\ty{\Gamma}}{\conf{C}}$, which states that ``a configuration $\tm{\conf{C}}$ with flag $\phi$ is well-typed under typing environment $\ty{\Gamma}$''\simon{I found it surprising that configuration typing doesn't need priority annotations.
-Can you give an intuitition why?}.
+Configurations are typed by a judgement $\cseq[\phi]{\ty{\Gamma}}{\conf{C}}$, which states that ``a configuration $\tm{\conf{C}}$ with flag $\phi$ is well-typed under typing environment $\ty{\Gamma}$''.
 Configuration typing is based on the standard typing for GV.
 Terms are embedded either as main or as child threads.
-The priority bound from the term typing is discarded, as configurations contain no further sequential actions.
+The priority bound from the term typing is discarded, as configurations contain no further blocking actions.
 Main threads are allowed to return a value, whereas child threads are required to return the unit value.

### 2020-10-15 priority-gv.tex
Update priority-gv.tex

@@
 Priority GV offers a more fine-grained analysis of communication structures, and by separating channel creation form thread spawning it allows cyclic structures.
-We illustrate PGV with Milner's cyclic scheduler~\cite{milner89}.
+We illustrate PGV with Milner's Cyclic Scheduler~\cite{milner89}.
 
@@
 \]
-The session types $\ty{\tysend[\cs{o}]{T}{S}}$ and $\ty{\tyrecv[\cs{o}]{T}{S}}$ describe the endpoints of a channel over which we send or receive a value of type $\ty{T}$, and then proceed as $\ty{S}$.
-The types $\ty{\tyends[\cs{o}]}$ and $\ty{\tyendr[\cs{o}]}$ describe endpoints of a channel whose communication has finished, and over which we must synchronise before closing the channel.
+Session types $\ty{\tysend[\cs{o}]{T}{S}}$ and $\ty{\tyrecv[\cs{o}]{T}{S}}$ describe the endpoints of a channel over which we send or receive a value of type $\ty{T}$, and then proceed as $\ty{S}$.
+Types $\ty{\tyends[\cs{o}]}$ and $\ty{\tyendr[\cs{o}]}$ describe endpoints of a channel whose communication has finished, and over which we must synchronise before closing the channel.
 Each connective in a session type is annotated with a \emph{priority} $\cs{o}\in\mathbb{N}$.
@@
 \]
-The types $\ty{\typrod{T}{U}}$, $\ty{\tyunit}$, $\ty{\tysum{T}{U}}$, and $\ty{\tyvoid}$ are the standard linear $\lambda$-calculus product type, unit type, sum type, and empty type.
-The type $\ty{\tylolli[\cs{p},\cs{q}]{T}{U}}$ is the standard linear function type, annotated with \emph{priority bounds} $\cs{p},\cs{q}\in\mathbb{N}\cup\{\cs{\pbot},\cs{\ptop}\}$.
+Types $\ty{\typrod{T}{U}}$, $\ty{\tyunit}$, $\ty{\tysum{T}{U}}$, and $\ty{\tyvoid}$ are the standard linear $\lambda$-calculus product type, unit type, sum type, and empty type.
+Type $\ty{\tylolli[\cs{p},\cs{q}]{T}{U}}$ is the standard linear function type, annotated with \emph{priority bounds} $\cs{p},\cs{q}\in\mathbb{N}\cup\{\cs{\pbot},\cs{\ptop}\}$.
 Every session type is also a type.
-
 Given a function with type $\ty{\tylolli[\cs{p},\cs{q}]{T}{U}}$, $\cs{p}$ is a \emph{lower bound} on the priorities of the endpoints captured by the body of the function, and $\cs{q}$ is an \emph{upper bound} on the priority of the communications that take place as a result of applying the function.
@@
 \subsubsection*{Priorities}
-The function $\pr$ returns the smallest priority of a session type.
+Function $\pr(\cdot)$ returns the smallest priority of a session type.
 The type system guarantees that the top-most connective always holds the smallest priority, so we simply return the priority of the top-most connective:
@@
 \]
-We extend the function $\pr$ to types and typing contexts by returning the smallest priority in the type or context, or $\cs{\top}$ if there is no priority.
+We extend the function $\pr(\cdot)$ to types and typing contexts by returning the smallest priority in the type or context, or $\cs{\top}$ if there is no priority.
 We use $\sqcap$ and $\sqcup$ to denote the minimum and maximum:
@@
 
-The constants are best understood in conjunction with their typing and reduction rules in~\cref{fig:pgv-typing,fig:pgv-operational-semantics}.
-Briefly, $\tm{\link}$ links two endpoints together, forwarding messages from each to the other, $\tm{\new}$ creates a new channel and returns a pair of its endpoints, and $\tm{\spawn}$ spawns off its argument as a new thread.
+Constants are best understood in conjunction with their typing and reduction rules in~\cref{fig:pgv-typing,fig:pgv-operational-semantics}.
+Briefly, $\tm{\link}$ links two endpoints together, forwarding messages from one to the other, $\tm{\new}$ creates a new channel and returns a pair of its endpoints, and $\tm{\spawn}$ spawns off its argument as a new thread.
 The $\tm{\send}$ and $\tm{\recv}$ functions send and receive values on a channel.
-However, since the typing rules for PGV ensure the linear usage of endpoints, they also return a new copy of the endpoint, to continue the session.
+However, since the typing rules for PGV ensure the linear usage of endpoints, they also return a new copy of the endpoint to continue the session.
 The $\tm{\close}$ and $\tm{\wait}$ functions close a channel.
-
 We use syntactic sugar to make terms more readable: we write $\tm{\andthen{M}{N}}$ in place of $\tm{\letunit{M}{N}}$, $\tm{\letbind{x}{M}{N}}$ in place of $\tm{(\lambda x.N)\;M}$, and pattern matching functions $\tm{\lambda\unit.M}$ in place of $\tm{\lambda z.\letunit{z}{M}}$ and $\tm{\lambda\pair{x}{y}.M}$ in place of $\tm{\lambda z.\letpair{x}{y}{z}{M}}$.

### 2020-10-15 priority-gv.tex
Update priority-gv.tex

@@
 \subsubsection*{Operational Semantics}
+
 Priority GV terms are evaluated as part of a configuration of processes. Configurations are defined by the following grammar:
@@
 \]
+
+\input{fig/pgv-operational-semantics}
+
 We factor the reduction relation of PGV into a deterministic reduction on terms ($\tred$) and a non-deterministic reduction on configurations ($\cred$), see \cref{fig:pgv-operational-semantics}.
@@
 
-\input{fig/pgv-operational-semantics}
-
 \subsubsection*{Typing}
@@
 \Cref{fig:pgv-typing} gives the typing rules for PGV.
+
+\input{fig/pgv-typing}
 
@@
 In PGV, we guarantee deadlock freedom using priorities, which removes the channel pseudo-type $\ty{S^\sharp}$, simplifies typing for parallel composition, and restores type preservation for the structural congruence.
-
-\input{fig/pgv-typing}
 

### 2020-10-15 priority-cp.tex
Minor rewrite. Took down another of Simon's worries.

@@
 Our PCP scheduler has exactly the same behaviour as the version presented in~\cref{sec:pgv-example}.
-We use $\pcp{\tm{\prod_{i}P_i}}$ to denote the parallel composition of the processes $\pcp{\tm{P_i}}$, and $\pcp{\plug{P}{Q}}$ to denote the plugging of $\pcp{\tm{Q}}$ in the one-hole process-context $\pcp{\tm{P}}$:
+We use $\pcp{\tm{\prod_{I}P_i}}$ to denote the parallel composition of the processes $\pcp{\tm{P_i}}$, $\tm{i}\in\tm{I}$, and $\pcp{\tm{\plug{P}{Q}}}$ to denote the plugging of $\pcp{\tm{Q}}$ in the one-hole process-context $\pcp{\tm{P}}$:
 \[

### 2020-10-15 priority-gv.tex
Minor rewrite. Took down another of Simon's worries.

@@
 % \input{fig/pgv-scheduler.tex}
-We implement the scheduler as follows, using $\tm{\prod_{i}\conf{C}_i}$ to denote the parallel composition of the processes $\tm{\conf{C}_i}$, and $\tm{\plug{M}{N}}$ to denote the plugging of $\tm{N}$ in the one-hole term-context $\tm{M}$:
+We implement the scheduler as follows, using $\tm{\prod_{I}\conf{C}_i}$ to denote the parallel composition of the processes $\tm{\conf{C}_i}$, $\tm{i}\in\tm{I}$, and $\tm{\plug{M}{N}}$ to denote the plugging of $\tm{N}$ in the one-hole term-context $\tm{M}$:
 \[
@@
 
-In PGV\simon{I think the rules need more explanation than this---as someone who knows the basics of priorities but is not an expert, I found it difficult to understand the T-Lam rule for example}, sequents are annotated by a priority bound $\cs{p}$, which is the \emph{upper bound} on the priority of the communication actions performed while evaluating the term.
+In PGV, sequents are annotated by a priority bound $\cs{p}$, which is the \emph{upper bound} on the priority of the communication actions performed while evaluating the term.
 The \emph{lower bound} is obtained from the typing environment.
@@
 For instance, the side condition for typing $\tm{\letunit{M}{N}}$ requires that $\tm{M}$ has finished communicating before $\tm{N}$ starts communicating.
+The rule \LabTirName{T-Lam} stores the lower and upper priority bound of the function body in the type.
 
@@
     \[
-      \tm{\res{x_1}{x'_1}{\dots\res{x_n}{x'_n}{(\child\;M_1\parallel\dots\parallel\child\;M_m\parallel\main\;N)}}}
+      \tm{\res{x_1}{x'_1}{\dots\res{x_n}{x'_n}{(\child\;M_1\parallel\dots\parallel\child\;M_m\parallel\main\;N)}}}.
     \]
@@
     \[
-      \tm{\conf{C}}
-      \equiv
-      \tm{\res{x_1}{x'_1}{\dots\res{x_n}{x'_n}{(\child\;M_1\parallel\dots\parallel\child\;M_m\parallel\main\;V)}}},
+      \tm{\res{x_1}{x'_1}{\dots\res{x_n}{x'_n}{(\child\;M_1\parallel\dots\parallel\child\;M_m\parallel\main\;V)}}}.
     \]

### 2020-10-15 priority-cp.tex
Moved figs.

@@
 Each of the axioms of structural congruence corresponds to the axiom of the same name for PGV.
-We write $\pcp{\red^+}$ for the transitive closures, and $\pcp{\red^\star}$ for the reflexive-transitive closures.\simon{Maybe worth highlighting that it is an internal mobility calculus, for the pi calculus people.}
+We write $\pcp{\red^+}$ for the transitive closures, and $\pcp{\red^\star}$ for the reflexive-transitive closures.
 

### 2020-10-15 conclusion.tex
Compressed lists and subsubsections, minor rewrites in text.
Compressed PCP metatheory.
Minor changes.

@@
 \subsubsection*{Conclusion and Future Work}
-In this paper we presented Priority GV, which is a functional language with session types.
-We used priorities in session types, following Priority CP~\cite{dardhagay18} to guarantee deadlock freedom statically via our priority-based type system.
-We proved the progress property of PGV in \cref{lem:pgv-closed-progress-confs}.
-We also presented an updated version of PCP, where we moved from commuting conversions to structural congruence in the reduction relation, thus remaining faithful to the reduction relation of the $\pi$-calculus.
-Finally, we presented an encoding of Priority CP to Priority GV in \cref{sec:priority-cp}.
-For this translation we proved that it is sound with respect to typing (\cref{thm:pcp-to-pgv-confs-preservation}) and operational semantics, given by operational correspondence soundness (\cref{thm:pcp-to-pgv-operational-correspondence-soundness}) and completeness (\cref{thm:pcp-to-pgv-operational-correspondence-completeness}).
+We presented Priority GV, a~session-typed functional language which uses priorities to ensure deadlock freedom, and showed its relation to Priority CP~\cite{dardhagay18} via an operational correspondence.
 
-We identify the following main directions for future work: exploring recursion in PGV, by integrating the works of Lindley and Morris~\cite{lindleymorris16} and Padovani and Novara~\cite{padovaninovara15}; defining a translation from Priority GV to Priority CP; and adding sharing following Balzer and Pfenning~\cite{balzerpfenning17}.
+Our formalism so far only captures the core of GV.
+In future work, we plan to explore: recursion in PGV, by integrating the works of Lindley and Morris~\cite{lindleymorris16} and Padovani and Novara~\cite{padovaninovara15}; a~translation from Priority GV to Priority CP; and sharing, following Balzer and Pfenning~\cite{balzerpfenning17}.
 \end{document}

### 2020-10-15 conclusion.tex
Proc0->Proc1, started compressing related work.o

@@
 \subsubsection*{On deadlock freedom and progress in $\pi$-calculus and $\lambda$-calculus}
-Deadlock freedom and progress have been studied thoroughly in the $\pi$-calculus with or without session types.
-In the standard typed $\pi$-calculus a foundational line of work is by Kobayashi and his original type-theoretic approach to deadlock-freedom~\cite{kobayashi98}, where priorities are abstract tags defined over a partially ordered set.
-Later on these abstract tags were simplified to natural numbers, and pairs of obligations and capabilities were used in the type system for deadlock or livelock freedom~\cite{kobayashi02,kobayashi06}.
-This allowed more $\pi$-calculus processes to be typed.
+
+Deadlock freedom and progress are well-studied in the $\pi$-calculus.
+In the standard typed $\pi$-calculus, one line of work is Kobayashi's approach to deadlock-freedom~\cite{kobayashi98}, where priorities are abstract tags defined over a partially ordered set.
+These abstract tags were later simplified to natural numbers.
+Pairs of obligations and capabilities were used in the type system for lock freedom~\cite{kobayashi02,kobayashi06}, which allowed more $\pi$-calculus processes to be typed.
 Following, Kobayashi's line of work, Padovani~\cite{padovani13} adapted the obligation and capability pairs to session types, and later on he further simplified them to a single priority for linear $\pi$-calculus~\cite{padovani14}.

### 2020-10-15 priority-gv.tex
Fix minor bug in PGV example, compressed E defn.
Proc0->Proc1, started compressing related work.o

@@
 \begin{example}[Milner's Cyclic Scheduler in PGV]\label{sec:pgv-example}
-Suppose a set of processes $\tm{\proc{i}}$, $1\leq{i}\leq{n}$, is scheduled to perform some tasks in cyclic order, starting with $\tm{\proc0}$, ending with $\tm{\proc{n}}$, and notifying $\tm{\proc0}$ when all processes have finished.
+Suppose a set of processes $\tm{\proc{i}}$, $1\leq{i}\leq{n}$, is scheduled to perform some tasks in cyclic order, starting with $\tm{\proc1}$, ending with $\tm{\proc{n}}$, and notifying $\tm{\proc1}$ when all processes have finished.
 % The dotted lines represent the optional transfer of data between processes.
@@
                      \parallel
-                     \child\;\andthen{\agent1}{\andthen{\wait\;{c'_3}}{\close\;{d}}}
+                     \child\;\andthen{\agent1}{\andthen{\wait\;{c'_n}}{\close\;{d}}}
                      \\
@@
 Each agent signals their process to start on $\tm{a_i}$, waits for their process to finish on $\tm{b_i}$, and then signals for the next agent to continue on $\tm{c_i}$.
-The agent $\tm{\agent0}$ initiates, then waits for every other process to finish, and signals $\tm{\proc0}$ on $\tm{d}$.
+The agent $\tm{\agent0}$ initiates, then waits for every other process to finish, and signals $\tm{\proc1}$ on $\tm{d}$.
 Every other agent $\tm{\agent{i}}$, $2\leq{i}\leq{n}$ waits on $\tm{c'_{i-1}}$ for the signal to start.
@@
     \sep        \tm{\inl{V}}
-    \sep        \tm{\inr{V}}
-  \\
+    \sep        \tm{\inr{V}} \\
   \tm{E}
-  & \Coloneqq & \tm{\hole} \\
-  & \sep      & \tm{E\;M}
+  & \Coloneqq & \tm{\hole}
+    \sep        \tm{E\;M}
     \sep        \tm{V\;E} \\
-  & \sep      & \tm{\letunit{E}{N}} \\
-  & \sep      & \tm{\pair{E}{M}}
+  & \sep      & \tm{\letunit{E}{N}}
+    \sep        \tm{\pair{E}{M}}
     \sep        \tm{\pair{V}{E}}
@@
     \sep        \tm{\inr{E}}
-    \sep        \tm{\casesum{E}{x}{M}{y}{N}} \\
-  & \sep      & \tm{\absurd{E}}
-  \\
+    \sep        \tm{\casesum{E}{x}{M}{y}{N}}
+    \sep        \tm{\absurd{E}} \\
   \tm{\conf{F}}

### 2020-10-15 priority-gv.tex
on typing in pgv

--- 2020-10-15 priority-gv.tex
@@
 \subsubsection*{Typing}
-Terms are typed by a judgement $\tseq[\cs{p}]{\ty{\Gamma}}{M}{T}$, which states that ``a term $\tm{M}$ has type $\ty{T}$ and an upper bound on its priority $\cs{p}$ under the typing environment $\ty{\Gamma}$''.
-Typing is standard for the linear $\lambda$-calculus.
+\Cref{fig:pgv-typing} gives the typing rules for PGV.
+Typing rules for terms are at the top of \cref{fig:pgv-typing}.
+Terms are typed by a judgement $\tseq[\cs{p}]{\ty{\Gamma}}{M}{T}$ stating that ``a term $\tm{M}$ has type $\ty{T}$ and an upper bound on its priority $\cs{p}$ under the typing environment $\ty{\Gamma}$''.
+Typing for the linear $\lambda$-calculus is standard.
 Linearity is ensured by splitting environments on branching rules, requiring that the environment in the variable rule consists of just the variable, and the environment in the constant and unit rules are empty.
-The constants $\tm{K}$ are typed using type schemas, and embedded using \LabTirName{T-Const}.
-\Cref{fig:pgv-typing} gives the typing rules for PGV.
+Constants $\tm{K}$ are typed using type schemas, and embedded using \LabTirName{T-Const} (mid of \cref{fig:pgv-typing}).
 
@@
 
-In PGV, sequents are annotated by a priority bound $\cs{p}$, which is the \emph{upper bound} on the priority of the communication actions performed while evaluating the term.
+In PGV, sequents are annotated with a priority bound $\cs{p}$, which is the \emph{upper bound} on the priority of actions performed while evaluating the term.
 The \emph{lower bound} is obtained from the typing environment.
@@
 
-Configurations are typed by a judgement $\cseq[\phi]{\ty{\Gamma}}{\conf{C}}$, which states that ``a configuration $\tm{\conf{C}}$ with flag $\phi$ is well-typed under typing environment $\ty{\Gamma}$''.
+Typing rules for configurations are at the bottom of \cref{fig:pgv-typing}.
+Configurations are typed by a judgement $\cseq[\phi]{\ty{\Gamma}}{\conf{C}}$ stating that ``a configuration $\tm{\conf{C}}$ with flag $\phi$ is well typed under typing environment $\ty{\Gamma}$''.
 Configuration typing is based on the standard typing for GV.
@@
 
-While our configuration typing is based on the standard typing for GV, it differs on two counts.
-Firstly, we require that child threads return the unit value, as opposed to a terminated channel.
-Secondly, we simplify typing for parallel composition.
-In GV, each parallel composition must split \emph{exactly one} channel of the channel pseudo-type $\ty{S^\sharp}$ into two endpoints of type $\ty{S}$ and $\ty{\co{S}}$ across the parallel composition.
+While our configuration typing is based on the standard typing for GV, it differs on two counts
+$(i)$ we require that child threads return the unit value, as opposed to a terminated channel;
+$(ii)$ we simplify typing for parallel composition.
+In order to guarantee deadlock freedom, in GV each parallel composition must split \emph{exactly one} channel of the channel pseudo-type $\ty{S^\sharp}$ into two endpoints of type $\ty{S}$ and $\ty{\co{S}}$.
 Consequently, associativity of parallel composition does not preserve typing.
-Unfortunately, splitting exactly one channel across a parallel composition in GV plays a crucial role in guaranteeing deadlock freedom.
 In PGV, we guarantee deadlock freedom using priorities, which removes the channel pseudo-type $\ty{S^\sharp}$, simplifies typing for parallel composition, and restores type preservation for the structural congruence.

### 2020-10-15 conclusion.tex
I messed with the related work and it became LONGER :'(

@@
 \textit{Simon: I think you really need some mention of MPSTs here, since a key motivation for MPSTs is ensuring deadlock-freedom between participants in a single MPST session.
-You can then talk about the global progress work by Coppo \etal, too, which incorporates an 'interaction typing system' (that few people understand).
+You can then talk about the global progress work by Coppo~\etal, too, which incorporates an 'interaction typing system' (that few people understand).
 How does that relate?}
 
-\subsubsection*{On deadlock freedom and progress in $\pi$-calculus and $\lambda$-calculus}
+\subsubsection*{Deadlock freedom}
 
 Deadlock freedom and progress are well-studied in the $\pi$-calculus.
-In the standard typed $\pi$-calculus, one line of work is Kobayashi's approach to deadlock-freedom~\cite{kobayashi98}, where priorities are abstract tags defined over a partially ordered set.
+For the standard typed $\pi$-calculus, one line of work is Kobayashi's approach to deadlock-freedom~\cite{kobayashi98}, where priorities are abstract tags defined over a partially ordered set.
 These abstract tags were later simplified to natural numbers.
 Pairs of obligations and capabilities were used in the type system for lock freedom~\cite{kobayashi02,kobayashi06}, which allowed more $\pi$-calculus processes to be typed.
-Following, Kobayashi's line of work, Padovani~\cite{padovani13} adapted the obligation and capability pairs to session types, and later on he further simplified them to a single priority for linear $\pi$-calculus~\cite{padovani14}.
-By using the encoding of session types into linear types~\cite{kobayashi07,dardhagiachino12,dardha14beat,dardha16}, the priority-based technique for deadlock freedom can be transferred onto the $\pi$-calculus with session types.
+Padovani~\cite{padovani13} adapted the obligation and capability pairs to session types, and later simplified them to a single priority for linear $\pi$-calculus~\cite{padovani14}.
+Using the encoding of session types into linear types~\cite{kobayashi07,dardhagiachino12,dardha14beat,dardha16}, the priority-based technique for deadlock freedom can be transferred onto the $\pi$-calculus with session types.
 
-The groundbreaking work on progress for session-typed $\pi$-calculus, by Dezani \etal~\cite{dezani-ciancaglinimostrous06}, guarantees progress by allowing only one active session at a time.
-In~\cite{dezani-ciancagliniliguoro09progress} the authors introduce a partial order on channels in line with Kobayashi's work~\cite{kobayashi98}.
-Carbone and Debois~\cite{carbonedebois10} define progress for session typed $\pi$-calculus in terms of a \emph{catalyser} used to provide a missing counterpart to a process, thus guaranteeing progress.
-Carbone \etal~\cite{carbonedardha14} studied further the use of catalysers and showed that progress is a compositional form of lock-freedom.
-Following Padovani~\cite{padovani14}, the authors show that by using the encoding of session types~\cite{dardhagiachino12} and Kobayashi's obligations/capabilities, we can obtain progress for session types.
-Vieira and Vasconcelos~\cite{vieiravasconcelos13} used single priorities and an abstract partial order to guarantee deadlock freedom in a session-typed $\pi$-calculus.
+For the session-typed $\pi$-calculus, the foundational work by Dezani~\etal~\cite{dezani-ciancaglinimostrous06} guarantees progress by allowing only one active session at a time.
+Dezani later~\cite{dezani-ciancagliniliguoro09progress} introduces a partial order on channels in line with Kobayashi's work~\cite{kobayashi98}.
+Carbone and Debois~\cite{carbonedebois10} define progress for session typed $\pi$-calculus in terms of a \emph{catalyser} used to provide a missing counterpart to a process, thus guaranteeing deadlock freedom.
+Carbone~\etal~\cite{carbonedardha14} studied the use of catalysers and showed that progress is a compositional form of lock-freedom.
+Following Padovani~\cite{padovani14}, Dardha~\etal~\cite{dardhagiachino12} show that by using the encoding of session types and Kobayashi's obligations/capabilities, we can obtain progress for session types.\wen{I'm not sure what this means.} Vieira and Vasconcelos~\cite{vieiravasconcelos13} used single priorities and an abstract partial order to guarantee deadlock freedom in a session-typed $\pi$-calculus.
 
-Gay~\etal~\cite{gaynagarajan03} and Vasconcelos \etal~\cite{vasconcelosravara04,vasconcelosgay06} were the first to introduce a functional language with session types.
-However, such works and including the GV line~\cite{gayvasconcelos10,gayvasconcelos12} did not guarantee deadlock freedom.
-This property was later on addressed~\cite{lindleymorris15,wadler15} via syntactic restrictions allowing only communication in tree structures, as discussed in the introduction.
-Toninho \etal~\cite{toninhocaires12} present an encoding of simply-typed $\lambda$-calculus into session-typed $\pi$-calculus investigating the concurrency features of $\lambda$-calculus and providing a new logical explanation.
-Again, their focus is not on deadlock freedom.
+Gay~\etal~\cite{gaynagarajan03} and Vasconcelos~\etal~\cite{vasconcelosravara04,vasconcelosgay06} were the first to introduce a functional language with session types.
+However, such works, including early GV~\cite{gayvasconcelos10,gayvasconcelos12} did not guarantee deadlock freedom, until it was addressed via syntactic restrictions~\cite{lindleymorris15,wadler15}.
+Toninho~\etal~\cite{toninhocaires12} present a translation of simply-typed $\lambda$-calculus into session-typed $\pi$-calculus.
+However, their focus is not on deadlock freedom.
 
-\subsubsection*{On Curry-Howard correspondences}
-The Curry-Howard correspondence between logic and types laid the foundation for functional programming~\cite{wadler15}.
-With the rise of linear logic~\cite{girard87} a Curry-Howard correspondence was established between linear logic and the linear $\pi$-calculus~\cite{abramsky94,bellinscott94}.
-With the rise of session types, a new correspondence between linear logic and the $\pi$-calculus was proposed first by Caires and Pfenning~\cite{cairespfenning10}, for dual intuitionistic linear logic (DILL), and later by Wadler~\cite{wadler15} for classical linear logic (CLL).
-The underlying language in both correspondences satisfy deadlock freedom by design, by forbidding cyclic connections via the logical cut rule.
-Dardha and Gay follow Padovani's work~\cite{padovani14} and use priorities in CP to define a new calculus which is deadlock-free by design, based on classical linear logic correspondence with session types~\cite{wadler12}.
-Dardha and P\'{e}rez~\cite{dardhaperez15} compare Kobayashi-style typing and CLL typing, and prove that CLL corresponds to Kobayashi's system with the restriction that only single cuts are allowed.
-Balzer \etal study sharing~\cite{balzerpfenning17} and later on deadlock freedom~\cite{balzertoninho19} in $\text{SILL}_S$ and $\text{SILL}_{S+}$, session-typed $\pi$-calculi based on the correspondence with DILL.
-In $\text{SILL}_S$~\cite{balzerpfenning17}, types are either linear or shared with modal operators connecting them, which allows sharing, but not deadlock freedom.
-In $\text{SILL}_{S+}$~\cite{balzertoninho19} types are enhanced with information regarding the order of acquiring resources, which guarantees deadlock freedom.
+\subsubsection*{Ties with logic}
+The correspondence between logic and types lays the foundation for functional programming~\cite{wadler15}.
+Since its inception by Girard~\cite{girard87}, linear logic has been a candidate for a foundational correspondence for concurrent programs.
+A~correspondence with linear $\pi$-calculus was established early on by Abramsky~\cite{abramsky94,bellinscott94}.
+A~correspondence between session-typed $\pi$-calculus and dual intuitionistic linear logic was developed by Caires and Pfenning~\cite[$\pi\text{DILL}$]{cairespfenning10}, and with classical linear logic by Wadler~\cite[CP]{wadler15}, both guaranteeing deadlock freedom as a result of their connection to logic.
+Dardha and Gay~\cite[PCP]{dardhagay18} integrate Padovani's work on priorities~\cite{padovani14} with CP, creating the first calculus which combines priorities and strong ties with logic.
+Dardha and P\'{e}rez~\cite{dardhaperez15} compare Kobayashi-style typing and CLL typing, and show that CLL corresponds to Kobayashi's system with the restriction that only single cuts are allowed.\wen{What do you mean by ``single cuts''?} Balzer~\etal~\cite[$\text{SILL}_S$]{balzerpfenning17} introduce shared state, which breaks deadlock freedom.
+They later restore deadlock freedom using priorities~\cite[$\text{SILL}_{S+}$]{balzertoninho19}.
 

### 2020-10-15 priority-gv.tex
Update priority-gv.tex

@@
 
-\emph{Actions} are terms which perform communication actions which synchronisation between two threads:
+\emph{Actions} are terms which perform communication actions and which synchronise between two threads.
 \begin{definition}[Actions]
@@
 
-\emph{Ready terms} are terms which are ready to perform a communication action, either by themselves, \eg creating a new channel or thread, or with another thread, \eg sending or receiving:
+\emph{Ready terms} are terms which are ready to perform a communication action, either by themselves, \eg creating a new channel or thread, or with another thread, \eg sending or receiving.
 \begin{definition}[Ready Terms]

### 2020-10-15 introduction.tex
important contribution
Update introduction.tex

@@
 \item \emph{Priority GV} We present Priority GV (PGV) in~\cref{sec:pgv}, and prove subject reduction (\cref{thm:pgv-subject-reduction-confs}) and progress (\cref{thm:pgv-closed-progress-confs}).
+Unlike  previous work on GV, structural congruence and reduction are type preserving in PGV.
 \item \emph{Milner's Cyclic Scheduler} To illustrate newfound expressiveness of PGV, we present Milner's Cyclic Scheduler~\cite{milner89} (\cref{sec:pgv-example}).
@@
 We present Milner's Cyclic Scheduler in PCP (\cref{sec:pcp-example}).
-\item \emph{Translation from PCP to PGV} We present a translation from PCP in PGV in~\cref{sec:pcp}.
-We prove the translation preserves typing (\cref{thm:pcp-to-pgv-confs-preservation}) and satisfies operational correspondence (\cref{thm:pcp-to-pgv-operational-correspondence-soundness,thm:pcp-to-pgv-operational-correspondence-completeness}), and show that the scheduler in PCP (\cref{sec:pcp-example}) translates to the scheduler in PGV (\cref{sec:pgv-example}).
+\item \emph{Translation from PCP to PGV} We present a translation from PCP to PGV in~\cref{sec:pcp} and prove that it preserves typing (\cref{thm:pcp-to-pgv-confs-preservation}) and satisfies operational correspondence (\cref{thm:pcp-to-pgv-operational-correspondence-soundness,thm:pcp-to-pgv-operational-correspondence-completeness}).
+We show that the scheduler in PCP (\cref{sec:pcp-example}) translates to the scheduler in PGV (\cref{sec:pgv-example}).
 \end{itemize}

### 2020-10-15 priority-cp.tex
Undo ugliest space gain.
Update priority-cp.tex

@@
 
-This section presents an updated version of Priority CP~\cite[PCP]{dardhagay18}, which is Wadler's Classical Processes~\cite[CP]{wadler12} with \emph{priorities}, and a translation from PCP to PGV.
-
-The main difference with respect to PCP by Dardha and Gay~\cite{dardhagay18} is the removal of commuting conversions, in favour of using the structural congruence.
-Commuting conversions are necessary if we want our reduction strategy to correspond \emph{exactly} to standard cut elimination.
+We present an updated version of Priority CP (PCP)~\cite{dardhagay18}, which is Wadler's CP~\cite{wadler12} with {priorities}, and a translation from PCP to PGV.
+The main difference with respect to PCP~\cite{dardhagay18} is removing commuting conversions, in favour of using structural congruence.
+Commuting conversions are necessary if we want our reduction strategy to correspond \emph{exactly} to cut (or cycle as in \cite{dardhagay18}) elimination.
 However, as Lindley and Morris~\cite{lindleymorris15} show, all $\beta$-reductions that can be performed \emph{with} the use of commuting conversions, can also be performed \emph{without} them.
@@
 
-From the perspective of process calculus, commuting conversions behave strangely. Let us consider the commuting conversion for $\pcp{\tm{\recv{x}{y}{P}}}$:
+From the perspective of process calculi, commuting conversions behave strangely. Let us consider the commuting conversion $(\kappa_{\parr})$ for $\pcp{\tm{\recv{x}{y}{P}}}$:
 \[
@@
 \end{mathpar}
-Commuting conversions break our intuition that an action with lower priority occurs before an action with higher priority, since ``if a prefix on a channel endpoint $\pcp{\tm{x}}$ with priority $\cs{o}$ is pulled out at top level, then to preserve priority conditions in the typing rules [..], it is necessary to increase priorities of all actions after the prefix on $\pcp{\tm{x}}$'' by $\cs{o+1}$~\cite{dardhagay18}.
+Commuting conversions break our intuition that an action with lower priority occurs before an action with higher priority.
+To cite Dardha and Gay~\cite{dardhagay18} ``\emph{if a prefix on a channel endpoint $\pcp{\tm{x}}$ with priority $\cs{o}$ is pulled out at top level, then to preserve priority constraints in the typing rules [..], it is necessary to increase priorities of all actions after the prefix on $\pcp{\tm{x}}$}'' by $\cs{o+1}$.
 
-\subsubsection*{Example}\label{sec:pcp-example}
+\begin{example}[Milner's Cyclic Scheduler in PCP]\label{sec:pcp-example}
 We introduce PCP by revising Milner's cyclic scheduler~\cite{milner89,dardhagay18}.
-Our PCP scheduler has exactly the same behaviour as the version presented in~\cref{sec:pgv-example}.
+Our PCP scheduler has exactly the same behaviour as the PGV version in~\cref{sec:pgv-example}.
 We use $\pcp{\tm{\prod_{I}P_i}}$ to denote the parallel composition of the processes $\pcp{\tm{P_i}}$, $\tm{i}\in\tm{I}$, and $\pcp{\tm{\plug{P}{Q}}}$ to denote the plugging of $\pcp{\tm{Q}}$ in the one-hole process-context $\pcp{\tm{P}}$:
@@
 \]
+\end{example}
 
 \subsubsection*{Types}
-Types $\pcp{\ty{A}}$, $\pcp{\ty{B}}$, $\pcp{\ty{C}}$ in PCP are based on classical linear logic propositions, and are defined by the following grammar:
+Types ($\pcp{\ty{A}}, \pcp{\ty{B}}$) in PCP are based on classical linear logic propositions, and are defined by the following grammar:
 \[
@@
   \begin{array}{lcl}
-    \ty{A}, \ty{B}, \ty{C}
+    \ty{A}, \ty{B}
     & \Coloneqq & \ty{\tytens[\cs{o}]{A}{B}}
@@
 Each connective is annotated with a priority $\cs{o}\in\mathbb{N}$.
-The types $\pcp{\ty{\tytens[\cs{o}]{A}{B}}}$ and $\pcp{\ty{\typarr[\cs{o}]{A}{B}}}$ type the endpoints of a channel over which we send or receive a channel of type $\pcp{\ty{A}}$, and then proceed as type $\pcp{\ty{B}}$.
-The types $\pcp{\ty{\tyone[\cs{o}]}}$ and $\pcp{\ty{{\tybot}[\cs{o}]}}$ type the endpoints of a channel whose session has terminated, and over which we send or receive a \emph{ping} before closing the channel.
+Types $\pcp{\ty{\tytens[\cs{o}]{A}{B}}}$ and $\pcp{\ty{\typarr[\cs{o}]{A}{B}}}$ type the endpoints of a channel over which we send or receive a channel of type $\pcp{\ty{A}}$, and then proceed as type $\pcp{\ty{B}}$.
+Types $\pcp{\ty{\tyone[\cs{o}]}}$ and $\pcp{\ty{{\tybot}[\cs{o}]}}$ type the endpoints of a channel whose session has terminated, and over which we send or receive a \emph{ping} before closing the channel.
 These two types act as units for $\pcp{\ty{\tytens[\cs{o}]{A}{B}}}$ and $\pcp{\ty{\typarr[\cs{o}]{A}{B}}}$, respectively.
-
-The types $\pcp{\ty{\typlus[\cs{o}]{A}{B}}}$ and $\pcp{\ty{\tywith[\cs{o}]{A}{B}}}$ type the endpoints of a channel over which we can receive or send a choice between two branches $\pcp{\ty{A}}$ or $\pcp{\ty{B}}$.
+Types $\pcp{\ty{\typlus[\cs{o}]{A}{B}}}$ and $\pcp{\ty{\tywith[\cs{o}]{A}{B}}}$ type the endpoints of a channel over which we can receive or send a choice between two branches $\pcp{\ty{A}}$ or $\pcp{\ty{B}}$.
 We have opted for a simplified version of choice and followed the original Wadler's CP \cite{wadler12}, however types $\ty{\oplus}$ and $\ty{\with}$ can be trivially generalised to $\pcp{\ty{\oplus^{\cs{o}}\{l_i:A_i\}_{i\in I}}}$ and $\pcp{\ty{\with^{\cs{o}}\{l_i:A_i\}_{i\in I}}}$, respectively, as in the original PCP \cite{dardhagay18}.
-The types $\pcp{\ty{\tynil[\cs{o}]}}$ and $\pcp{\ty{\tytop[\cs{o}]}}$ type the endpoints of a channel over which we can send or receive a choice between \emph{no options}.
+Types $\pcp{\ty{\tynil[\cs{o}]}}$ and $\pcp{\ty{\tytop[\cs{o}]}}$ type the endpoints of a channel over which we can send or receive a choice between \emph{no options}.
 These two types act as units for $\pcp{\ty{\typlus[\cs{o}]{A}{B}}}$ and $\pcp{\ty{\tywith[\cs{o}]{A}{B}}}$, respectively.
@@
 \[
+  \begin{array}{l}
     \minpr(\ty{\emptyenv})          = \cs{\ptop}
-    \qquad
+    \\
     \minpr(\ty{\Gamma},\tmty{x}{A}) = \minpr(\ty{\Gamma})\sqcap\minpr(\ty{A})
+  \end{array}
 \]
@@
 \subsubsection*{Terms}
-We let $\pcp{\tm{P}}$, $\pcp{\tm{Q}}$ range over processes, produced by the following grammar.
+Processes ($\pcp{\tm{P}}$, $\pcp{\tm{Q}}$) in PCP are defined by the following grammar.
 \[
@@
 \]
-The process $\pcp{\tm{\link{x}{y}}}$ links endpoints $\pcp{\tm{x}}$ and $\pcp{\tm{y}}$ and forwards communication from one to the other.
+Process $\pcp{\tm{\link{x}{y}}}$ links endpoints $\pcp{\tm{x}}$ and $\pcp{\tm{y}}$ and forwards communication from one to the other.
 $\pcp{\tm{\res{x}{y}{P}}}$, $\pcp{\tm{(\ppar{P}{Q})}}$ and $\pcp{\tm{\halt}}$ denote respectively the restriction processes where channel endpoints $\pcp{\tm{x}}$ and $\pcp{\tm{y}}$ are bound together and with scope $\pcp{\tm{P}}$, the parallel composition of processes $\pcp{\tm{P}}$ and $\pcp{\tm{Q}}$ and the terminated process.
-
 Processes $\pcp{\tm{\send{x}{y}{P}}}$ and $\pcp{\tm{\recv{x}{y}{P}}}$ send or receive over channel $\pcp{\tm{x}}$ a value $\pcp{\tm{y}}$ and proceed as process $\pcp{\tm{P}}$.
-The processes $\pcp{\tm{\close{x}{P}}}$ and $\pcp{\tm{\wait{x}{P}}}$ send and receive an empty value---denoting the closure of channel $\pcp{\tm{x}}$, and continue as $\pcp{\tm{P}}$.
-
+Processes $\pcp{\tm{\close{x}{P}}}$ and $\pcp{\tm{\wait{x}{P}}}$ send and receive an empty value---denoting the closure of channel $\pcp{\tm{x}}$, and continue as $\pcp{\tm{P}}$.
 Processes $\pcp{\tm{\inl{x}{P}}}$ and $\pcp{\tm{\inr{x}{P}}}$ make a left and right choice, respectively and proceed as process $\pcp{\tm{P}}$.
 Dually, $\pcp{\tm{\offer{x}{P}{Q}}}$ offers both left and right branches, with continuations $\pcp{\tm{P}}$ and $\pcp{\tm{Q}}$, and $\pcp{\tm{\absurd{x}}}$ is the empty offer.
-
 We write \emph{unbound} send as $\pcp{\tm{\usend{x}{y}{P}}}$, which is syntactic sugar for $\pcp{\tm{\send{x}{z}{(\ppar{\link{y}{z}}{P})}}}$.
 Alternatively, we could take $\pcp{\tm{\usend{x}{y}{P}}}$ as primitive, and let $\pcp{\tm{\send{x}{y}{P}}}$ be syntactic sugar for $\pcp{\tm{\res{y}{z}{(\usend{x}{z}{P})}}}$.
-CP takes \emph{bound} sending as primitive, as it is impossible to eliminate the top-level cut in terms such as $\pcp{\tm{\res{y}{z}{(\usend{x}{z}{P})}}}$, even with communting conversions.
+CP takes \emph{bound} sending as primitive, as it is impossible to eliminate the top-level cut in terms such as $\pcp{\tm{\res{y}{z}{(\usend{x}{z}{P})}}}$, even with commuting conversions.
 In our setting without commuting conversions and with more permissive normal forms, this is no longer an issue, but, for simplicity, we keep bound sending as primitive.
@@
 We write $\pcp{\red^+}$ for the transitive closures, and $\pcp{\red^\star}$ for the reflexive-transitive closures.
-
 The reduction relation is given by a set of axioms and inference rules for context closure.
@@
 
-Typing rules make use of priorities, defined in the previous section.
 \textsc{T-Link} states that the link process $\pcp{\tm{\link{x}{y}}}$ is well typed under channels $\pcp{\tm{x}}$ and $\pcp{\tm{y}}$ having dual types, respectively $\pcp{\ty{A}}$ and $\pcp{\ty{\co{A}}}$.
@@
 \textsc{T-Halt} states that the terminated process $\pcp{\tm{\halt}}$ is well typed in the empty context.
-
 \textsc{T-Send} and \textsc{T-Recv} state that the sending and receiving of a bound name $\pcp{\tm{y}}$ over a channel $\pcp{\tm{x}}$ is well typed under $\pcp{\ty{\Gamma}}$ and $\pcp{\tm{x}}$ of type $\pcp{\ty{\tytens[\cs{o}]{A}{B}}}$, respectively $\pcp{\typarr[\cs{o}]{A}{B}}$.
 Priority $\cs{o}$ is the smallest among all priorities of the types used by the output or input process, captured by the side condition $\pcp{\cs{o}<\minpr(\ty{\Gamma},\ty{A},\ty{B})}$.
-
 Rules \textsc{T-Close} and \textsc{T-Wait} type the closure of channel $\pcp{\tm{x}}$ and are in the same lines as the previous two rules, requiring that the priority of channel $\pcp{\tm{x}}$ is the smallest among all priorities in $\pcp{\ty{\Gamma}}$.
-
 \textsc{T-Select-Inl} and \textsc{T-Select-Inr} type respectively the left $\pcp{\tm{\inl{x}{P}}}$ and right $\pcp{\tm{\inr{x}{P}}}$ choice performed on channel $\pcp{\tm{x}}$.
@@
 
-\subsubsection*{Metatheory}
-Reduction in our version of PCP preserve types.
-Since our reduction relation is a strict subset of the reduction relation in the original~\cite{dardhagay18}, we defer to their proofs.
+Finally, since our reduction relation is a strict subset of the reduction relation in the original~\cite{dardhagay18}, we defer to their proofs.
 We prove progress for our version of PCP, see~\cref{prf:thm-pcp-closed-progress}.

### 2020-10-15 priority-cp.tex
Merge semantics and typing for PCP, translation using align* so it can break pages.

@@
 To cite Dardha and Gay~\cite{dardhagay18} ``\emph{if a prefix on a channel endpoint $\pcp{\tm{x}}$ with priority $\cs{o}$ is pulled out at top level, then to preserve priority constraints in the typing rules [..], it is necessary to increase priorities of all actions after the prefix on $\pcp{\tm{x}}$}'' by $\cs{o+1}$.
+
+\input{fig/pcp-operational-semantics-and-typing}
 
@@
 \subsubsection*{Operational Semantics}
-The operational semantics for PCP, given in \cref{fig:pcp-operational-semantics}, is defined as a reduction relation $\pcp{\red}$ on processes (bottom) and uses structural congruence (top).
+The operational semantics for PCP, given in \cref{fig:pcp-operational-semantics-and-typing}, is defined as a reduction relation $\pcp{\red}$ on processes (bottom) and uses structural congruence (top).
 Each of the axioms of structural congruence corresponds to the axiom of the same name for PGV.
 We write $\pcp{\red^+}$ for the transitive closures, and $\pcp{\red^\star}$ for the reflexive-transitive closures.
+
 The reduction relation is given by a set of axioms and inference rules for context closure.
@@
 
-\input{fig/pcp-operational-semantics}
-
 \subsubsection*{Typing}
-\Cref{fig:pcp-typing} gives the typing rules for our version of PCP.
+\Cref{fig:pcp-operational-semantics-and-typing} gives the typing rules for our version of PCP.
 A typing judgement $\pcp{\seq{P}{\Gamma}}$ states that ``process $\pcp{\tm{P}}$ is well typed under the typing context $\pcp{\ty{\Gamma}}$''.
@@
 In all the above rules the priority $\cs{o}$ of channel $\pcp{\tm{x}}$ is the smallest with respect to the typing context $\pcp{\cs{o}<\minpr(\ty{\Gamma})}$ and types involved in the choice $\pcp{\cs{o}<\minpr(\ty{\Gamma},\ty{A},\ty{B})}$.
-
-\input{fig/pcp-typing}
 
@@
 \]
-There are two separate translations on processes.
-The first, $\tm{\cpgvC{\cdot}}$, translates processes to \emph{configurations}.
+There are two separate translations on processes. The first, $\tm{\cpgvC{\cdot}}$, translates processes to \emph{configurations}:
+\begin{align*}
+  &\pcp{\tm{\cpgvC{\res{x}{y}{P}}}}
+  &&= \pgv{\tm{\res{x}{y}{\cpgvC{P}}}}
+  \\
+  &\pcp{\tm{\cpgvC{\ppar{P}{Q}}}}
+  &&= \pgv{\tm{\ppar{\cpgvC{P}}{\cpgvC{Q}}}}
+  \\
+  &\pcp{\tm{\cpgvC{\send{x}{y}{P}}}}
+  &&= \pgv{\tm{\res{y}{z}{(\child\;\letbind{x}{\send\;\pair{z}{x}}{\cpgvM{P}})}}}
+  \\
+  &\pcp{\tm{\cpgvC{\inl{x}{P}}}}
+  &&= \pgv{\tm{\res{y}{z}{(\child\;\letbind{x}{\andthen{\close\;(\send\;\pair{\inl{y}}{x})}{z}}{\cpgvM{P}})}}}
+  \\
+  &\pcp{\tm{\cpgvC{\inr{x}{P}}}}
+  &&= \pgv{\tm{\res{y}{z}{(\child\;\letbind{x}{\andthen{\close\;(\send\;\pair{\inr{y}}{x})}{z}}{\cpgvM{P}})}}}
+  \\
+  &\pcp{\tm{\cpgvC{P}}}
+  &&= \pgv{\tm{\child{\cpgvM{P}}}},\quad\text{if none of the above apply}
+\end{align*}
+The third, fourth, and fifth cases translate the $\nu$-binders implicit in the encoding of bound sending to configuration constructs, which is required for completeness of the operational correspondence.
 It translates as much as possible of the process to the corresponding configuration constructs.
 Once it hits the first action, it switches to the second translation, $\tm{\cpgvM{\cdot}}$, which translates processes to \emph{terms}:
-\[
-  \begin{array}{lcl}
-    \pcp{\tm{\cpgvC{\res{x}{y}{P}}}}
-    & = & \pgv{\tm{\res{x}{y}{\cpgvC{P}}}} \\
-    \pcp{\tm{\cpgvC{\ppar{P}{Q}}}}
-    & = & \pgv{\tm{\ppar{\cpgvC{P}}{\cpgvC{Q}}}} \\
-    \pcp{\tm{\cpgvC{\send{x}{y}{P}}}}
-    & = & \pgv{\tm{\res{y}{z}{(\child\;\letbind{x}{\send\;\pair{z}{x}}{\cpgvM{P}})}}} \\
-    \pcp{\tm{\cpgvC{\inl{x}{P}}}}
-    & = & \pgv{\tm{\res{y}{z}{(\child\;\letbind{x}{\andthen{\close\;(\send\;\pair{\inl{y}}{x})}{z}}{\cpgvM{P}})}}} \\
-    \pcp{\tm{\cpgvC{\inr{x}{P}}}}
-    & = & \pgv{\tm{\res{y}{z}{(\child\;\letbind{x}{\andthen{\close\;(\send\;\pair{\inr{y}}{x})}{z}}{\cpgvM{P}})}}} \\
-    \pcp{\tm{\cpgvC{P}}}
-    & = & \pgv{\tm{\child{\cpgvM{P}}}},\hfill\text{if none of the above apply}\\
-    \\
-    \pcp{\tm{\cpgvM{\link{x}{y}}}}
-    & = & \pgv{\tm{\link\;{\pair{x}{y}}}} \\
-    \pcp{\tm{\cpgvM{\res{x}{y}{P}}}}
-    & = & \pgv{\tm{\letpair{x}{y}{\new}{\cpgvM{P}}}} \\
-    \pcp{\tm{\cpgvM{\ppar{P}{Q}}}}
-    & = & \pgv{\tm{\andthen{\spawn\;{(\lambda\unit.\cpgvM{P})}}{\cpgvM{Q}}}} \\
-    \pcp{\tm{\cpgvM{\halt}}}
-    & = & \pgv{\tm{\unit}} \\
-    \pcp{\tm{\cpgvM{\close{x}{P}}}}
-    & = & \pgv{\tm{\andthen{\close\;{x}}{\cpgvM{P}}}} \\
-    \pcp{\tm{\cpgvM{\wait{x}{P}}}}
-    & = & \pgv{\tm{\andthen{\wait\;{x}}{\cpgvM{P}}}} \\
-    \pcp{\tm{\cpgvM{\send{x}{y}{P}}}}
-    & = & \pgv{\tm{\letpair{y}{z}{\new}{\letbind{x}{\send\;{\pair{z}{x}}}{\cpgvM{P}}}}} \\
-    \pcp{\tm{\cpgvM{\recv{x}{y}{P}}}}
-    & = & \pgv{\tm{\letpair{y}{x}{\recv\;{x}}{\cpgvM{P}}}} \\
-    \pcp{\tm{\cpgvM{\inl{x}{P}}}}
-    & = & \pgv{\tm{\letbind{x}{\select{\labinl}\;{x}}{\cpgvM{P}}}} \\
-    \pcp{\tm{\cpgvM{\inr{x}{P}}}}
-    & = & \pgv{\tm{\letbind{x}{\select{\labinr}\;{x}}{\cpgvM{P}}}} \\
-    \multicolumn{3}{l}{\pcp{\tm{\cpgvM{\offer{x}{P}{Q}}}}} \\
-    & = & \pgv{\tm{\offer{x}{x}{\cpgvM{P}}{x}{\cpgvM{Q}}}} \\
-    \pcp{\tm{\cpgvM{\absurd{x}}}}
-    & = & \pgv{\tm{\offeremp{x}}}
-  \end{array}
-\]
-The third, fourth, and fifth cases for $\tm{\cpgvC{\cdot}}$ translate the $\nu$-binders implicit in the encoding of bound sending to configuration constructs, which is required for completeness of the operational correspondence.
+\begin{align*}
+  &\pcp{\tm{\cpgvM{\link{x}{y}}}}
+  &&= \pgv{\tm{\link\;{\pair{x}{y}}}} \\
+  &\pcp{\tm{\cpgvM{\res{x}{y}{P}}}}
+  &&= \pgv{\tm{\letpair{x}{y}{\new}{\cpgvM{P}}}} \\
+  &\pcp{\tm{\cpgvM{\ppar{P}{Q}}}}
+  &&= \pgv{\tm{\andthen{\spawn\;{(\lambda\unit.\cpgvM{P})}}{\cpgvM{Q}}}} \\
+  &\pcp{\tm{\cpgvM{\halt}}}
+  &&= \pgv{\tm{\unit}} \\
+  &\pcp{\tm{\cpgvM{\close{x}{P}}}}
+  &&= \pgv{\tm{\andthen{\close\;{x}}{\cpgvM{P}}}} \\
+  &\pcp{\tm{\cpgvM{\wait{x}{P}}}}
+  &&= \pgv{\tm{\andthen{\wait\;{x}}{\cpgvM{P}}}} \\
+  &\pcp{\tm{\cpgvM{\send{x}{y}{P}}}}
+  &&= \pgv{\tm{\letpair{y}{z}{\new}{\letbind{x}{\send\;{\pair{z}{x}}}{\cpgvM{P}}}}} \\
+  &\pcp{\tm{\cpgvM{\recv{x}{y}{P}}}}
+  &&= \pgv{\tm{\letpair{y}{x}{\recv\;{x}}{\cpgvM{P}}}} \\
+  &\pcp{\tm{\cpgvM{\inl{x}{P}}}}
+  &&= \pgv{\tm{\letbind{x}{\select{\labinl}\;{x}}{\cpgvM{P}}}} \\
+  &\pcp{\tm{\cpgvM{\inr{x}{P}}}}
+  &&= \pgv{\tm{\letbind{x}{\select{\labinr}\;{x}}{\cpgvM{P}}}} \\
+  &\pcp{\tm{\cpgvM{\offer{x}{P}{Q}}}}
+  &&= \pgv{\tm{\offer{x}{x}{\cpgvM{P}}{x}{\cpgvM{Q}}}} \\
+  &\pcp{\tm{\cpgvM{\absurd{x}}}}
+  &&= \pgv{\tm{\offeremp{x}}}
+\end{align*}
 Typing environments are translated pointwise, and sequents $\pcp{\seq{P}{\ty{\Gamma}}}$ are translated as $\pgv{\cseq[\child]{\ty{\cpgvT{\ty{\Gamma}}}}{\cpgvC{P}}}$.

### 2020-10-15 appendix.tex
Included scheduler in appendix. Moved appendices to file.

% !TeX root = priorities.tex
\section{Priority GV}
\begingroup
\usingnamespace{pgv}

% Scheduler example
\input{fig/pgv-scheduler}

% Proofs that the syntactic sugar is well-typed.
\input{fig/pgv-typing-sugar}

\restatelemma{lempgvvaluedone}
\input{prf/lem-pgv-value-done}

\restatelemma{lempgvsubstitution}
\input{prf/lem-pgv-substitution}

\restatelemma{lempgvsubjectreductionterms}
\input{prf/lem-pgv-subject-reduction-terms}

\restatelemma{lempgvsubjectcongruence}
\input{prf/lem-pgv-subject-congruence}

\restatetheorem{thmpgvsubjectreductionconfs}
\input{prf/thm-pgv-subject-reduction-confs}

\begin{restatablelemma}{lempgvreadypriority}
  \label{lem:pgv-ready-priority}
  If $\tseq[\cs{p}]{\ty{\Gamma}}{L}{T}$ is ready to act on $\tmty{x}{S}\in\ty{\Gamma}$, then the priority bound $\cs{p}$ is some priority $\cs{o}$, \ie not $\cs{\pbot}$ or $\cs{\ptop}$.
\end{restatablelemma}
\begin{proof}
  Let $\tm{L}=\tm{\plug{E}{M}}$. By induction on the structure of $\tm{E}$. $\tm{M}$ has priority $\pr({\ty{S}})$, and each constructor of the evaluation context $\tm{E}$ passes on the \emph{maximum} of the priorities of its premises. No rule introduces the priority bound $\cs{\ptop}$ on the sequent.
\end{proof}

\begin{restatablelemma}{lempgvcanonicalforms}[Canonical Forms]
  \label{lem:pgv-canonical-forms}
  If $\cseq[\main]{\ty{\Gamma}}{\conf{C}}$, there exists some $\tm{\conf{D}}$ such that $\tm{\conf{C}}\equiv\tm{\conf{D}}$ and $\tm{\conf{D}}$ is in canonical form.
\end{restatablelemma}
\begin{proof}
  We move any $\nu$-binders to the top using \LabTirName{SC-ResExt}, discard any superfluous occurrences of $\tm{\child\;\unit}$ using \LabTirName{SC-ParNil}, and move the main thread to the rightmost position using \LabTirName{SC-ParComm} and \LabTirName{SC-ParAssoc}.
\end{proof}

\restatetheorem{thmpgvclosedprogressconfs}
\input{prf/thm-pgv-closed-progress-confs}
\endgroup

\section{Relation to Priority CP}
\begingroup
\usingnamespace{pcp}

% Proofs that the syntactic sugar is well-typed.
\input{fig/pcp-typing-sugar}

\begin{definition}[Actions]
  A~process acts on an endpoint $\tm{x}$ if it is in one of the following forms:
  \begin{multicols}{3}
    \begin{itemize}[noitemsep,topsep=0pt,parsep=0pt,partopsep=0pt]
    \item $\tm{\link{x}{y}}$ 
    \item $\tm{\link{y}{x}}$
    \item $\tm{\send{x}{y}{P}}$
    \item $\tm{\recv{x}{y}{P}}$
    \item $\tm{\close{x}{P}}$
    \item $\tm{\wait{x}{P}}$
    \item $\tm{\inl{x}{P}}$
    \item $\tm{\inr{x}{P}}$
    \item $\tm{\offer{x}{P}{Q}}$
    \item $\tm{\absurd{x}}$
    \end{itemize}
  \end{multicols}
  \noindent
  A~process is an action if it acts on some endpoint $\tm{x}$.
\end{definition}

\begin{definition}[Canonical Forms]
  \label{def:pcp-canonical-forms}
  A~process $\tm{P}$ is in canonical form if it is in one of the following forms:
  \begin{itemize}[noitemsep,topsep=0pt,parsep=0pt,partopsep=0pt]
  \item
    $\tm{\halt}$
  \item
    $\tm{\res{x_1}{x'_1}{\dots\res{x_n}{x'_n}{(P_1 \parallel\dots\parallel P_m)}}}$
    where $m>0$ and each $\tm{P_j}$ is an action.
  \end{itemize}
\end{definition}

\begin{restatablelemma}{lempcpcanonicalforms}[Canonical Forms]
  \label{lem:pcp-canonical-forms}
  If $\seq{P}{\ty{\Gamma}}$, there exists some $\tm{Q}$ such that $\tm{P}\equiv\tm{Q}$ and $\tm{Q}$ is in canonical form.
\end{restatablelemma}
\begin{proof}
  If $\tm{P}=\tm{\halt}$, we are done. Otherwise, we move any $\nu$-binders to the top using \LabTirName{SC-ResExt}, and discard any superfluous occurrences of $\tm{\halt}$ using \LabTirName{SC-ParNil}.
\end{proof}

\begin{restatabletheorem}{thmpcpclosedprogress}[Progress, $\pcp{\red}$]
  \label{thm:pcp-closed-progress}
  \hfill\\%newline before theorem statement
  If $\pcp{\seq{P}{\emptyenv}}$, then either $\pcp{\tm{P}=\tm{\halt}}$ or there exists a $\pcp{\tm{Q}}$ such that $\pcp{\tm{P}\red\tm{Q}}$.
\end{restatabletheorem}
\input{prf/thm-pcp-closed-progress}
\endgroup

\begingroup
\restatelemma{lempcptopgvtermspreservation}
\input{prf/lem-pcp-to-pgv-terms-preservation}

\restatetheorem{thmpcptopgvconfspreservation}
\input{prf/thm-pcp-to-pgv-confs-preservation}

\restatetheorem{thmpcptopgvoperationalcorrespondencesoundness}
\input{prf/thm-pcp-to-pgv-operational-correspondence-soundness}

\restatelemma{lempcptopgvcpgvMtocpgvC}
\input{prf/lem-pcp-to-pgv-cpgvM-to-cpgvC}

\begin{restatablelemma}{lempcptopgvoperationalcorrespondenceequiv}%
  \label{lem:pcp-to-pgv-confs-operational-correspondence-equiv}
  If $\pcp{\seq{P}{\ty{\Gamma}}}$ and $\pcp{\tm{P}\equiv\tm{Q}}$,
  then $\pgv{\tm{\cpgvC{P}}\equiv\tm{\cpgvC{Q}}}$.
\end{restatablelemma}
\begin{proof}
  Every axiom of the structural congruence in PCP maps directly to the axiom of the same name in PGV.
\end{proof}

\restatetheorem{thmpcptopgvoperationalcorrespondencecompleteness}
\input{prf/thm-pcp-to-pgv-operational-correspondence-completeness}
\endgroup

%%% Local Variables:
%%% TeX-master: "priorities"
%%% End:

### 2020-10-15 conclusion.tex
done with related work

@@
 \section{Related Work and Conclusion}
-\textit{Simon: I think you really need some mention of MPSTs here, since a key motivation for MPSTs is ensuring deadlock-freedom between participants in a single MPST session.
-You can then talk about the global progress work by Coppo~\etal, too, which incorporates an 'interaction typing system' (that few people understand).
-How does that relate?}
-
-\subsubsection*{Deadlock freedom}
-
-Deadlock freedom and progress are well-studied in the $\pi$-calculus.
-For the standard typed $\pi$-calculus, one line of work is Kobayashi's approach to deadlock-freedom~\cite{kobayashi98}, where priorities are abstract tags defined over a partially ordered set.
+\subsubsection*{Deadlock freedom and progress}
+Deadlock freedom and progress are well studied properties in the $\pi$-calculus.
+For the standard typed $\pi$-calculus, one line of work is Kobayashi's approach to deadlock freedom~\cite{kobayashi98}, where priorities are abstract tags defined over a partially ordered set.
 These abstract tags were later simplified to natural numbers.
-Pairs of obligations and capabilities were used in the type system for lock freedom~\cite{kobayashi02,kobayashi06}, which allowed more $\pi$-calculus processes to be typed.
+Pairs of obligations and capabilities were used in the type system for deadlock freedom~\cite{kobayashi06}, which allowed more $\pi$-calculus processes to be typed.
 Padovani~\cite{padovani13} adapted the obligation and capability pairs to session types, and later simplified them to a single priority for linear $\pi$-calculus~\cite{padovani14}.
-Using the encoding of session types into linear types~\cite{kobayashi07,dardhagiachino12,dardha14beat,dardha16}, the priority-based technique for deadlock freedom can be transferred onto the $\pi$-calculus with session types.
+As discussed by the author, by combining the encoding of session types into linear types~\cite{kobayashi07,dardhagiachino12,dardha14beat,dardha16} and the priority-based linear $\pi$-calculus, we can obtain deadlock freedom for the $\pi$-calculus with session types.
 
@@
 Carbone~\etal~\cite{carbonedardha14} studied the use of catalysers and showed that progress is a compositional form of lock-freedom.
-Following Padovani~\cite{padovani14}, Dardha~\etal~\cite{dardhagiachino12} show that by using the encoding of session types and Kobayashi's obligations/capabilities, we can obtain progress for session types.\wen{I'm not sure what this means.} Vieira and Vasconcelos~\cite{vieiravasconcelos13} used single priorities and an abstract partial order to guarantee deadlock freedom in a session-typed $\pi$-calculus.
+In the same lines of Padovani~\cite{padovani14}, Carbone~\etal~\cite{carbonedardha14} show that progress can be lifted to session types via the encoding of session types to linear types and Kobayashi's obligations/capabilities.
+Vieira and Vasconcelos~\cite{vieiravasconcelos13} use single priorities and an abstract partial order to guarantee deadlock freedom in a session-typed $\pi$-calculus.
+
+Multiparty Session Types (MPST) \cite{hondayoshida08} on the other hand guarantee deadlock freedom within a single session, but more advanced techniques are required for session interleaving.
+Bettini~\etal~\cite{bettinicoppo08} follow a similar technique to Kobayashi for multiparty session types.
+The main
+difference with our work is that we associate priorities to communication actions, and not to channels.
+Carbone and Montesi~\cite{carbonemontesi13} explore and combine MPST and choreographies to obtain a formalism that satisfies deadlock freedom by design.
 
@@
 A~correspondence between session-typed $\pi$-calculus and dual intuitionistic linear logic was developed by Caires and Pfenning~\cite[$\pi\text{DILL}$]{cairespfenning10}, and with classical linear logic by Wadler~\cite[CP]{wadler15}, both guaranteeing deadlock freedom as a result of their connection to logic.
-Dardha and Gay~\cite[PCP]{dardhagay18} integrate Padovani's work on priorities~\cite{padovani14} with CP, creating the first calculus which combines priorities and strong ties with logic.
-Dardha and P\'{e}rez~\cite{dardhaperez15} compare Kobayashi-style typing and CLL typing, and show that CLL corresponds to Kobayashi's system with the restriction that only single cuts are allowed.\wen{What do you mean by ``single cuts''?} Balzer~\etal~\cite[$\text{SILL}_S$]{balzerpfenning17} introduce shared state, which breaks deadlock freedom.
+Dardha and Gay~\cite[PCP]{dardhagay18} integrate Kobayashi's and Padovani's work on priorities~\cite{kobayashi06,padovani14} with CP, creating the first calculus which combines priorities and strong ties with logic.
+Dardha and P\'{e}rez~\cite{dardhaperez15} compare Kobayashi-style typing and CLL typing, and show that CLL corresponds to a subsystem of Kobayashi's where restriction is applied once.
+Balzer~\etal~\cite[$\text{SILL}_S$]{balzerpfenning17} introduce shared state, which breaks deadlock freedom.
 They later restore deadlock freedom using priorities~\cite[$\text{SILL}_{S+}$]{balzertoninho19}.
+Carbone~\etal~\cite{carbonelindley16} give a logical view of MPST with a generalised duality, called coherence.
 

### 2020-10-15 priority-cp.tex
Minor changes.

@@
 \end{align*}
-The third, fourth, and fifth cases translate the $\nu$-binders implicit in the encoding of bound sending to configuration constructs, which is required for completeness of the operational correspondence.
+The third, fourth, and fifth cases translate the $\nu$-binders in the translation of bound sending to configuration constructs, which is required for completeness of the operational correspondence.
 It translates as much as possible of the process to the corresponding configuration constructs.
@@
 Recall our implementations of Milner's cyclic scheduler in PCP ($\pcp{\tm{\sched}}$, \cref{sec:pcp-example}) and PGV ($\pgv{\tm{\sched}}$, \cref{sec:pgv-example}).
-The translation of $\pcp{\tm{\sched}}$ is exactly $\pgv{\tm{\sched}}$ where all processes are child processes.
+It follows immediately from the definition of our translation that the translation of $\pcp{\tm{\sched}}$ is exactly $\pgv{\tm{\sched}}$, where all processes are child processes.
 

### 2020-10-16 introduction.tex
final pass, hopefully

@@
 Moreover, the benefit of working with GV over other session-typed functional languages is that GV has strong ties to linear logic, via its relation to CP~\cite{wadler12}, and consequently it has strong formal properties, \eg deadlock freedom, albeit in a restrictive way.
-Following the above observations, we aim to further investigate GV and deadlock freedom, thus we pose our research question:
+Following the above observations, we aim to further investigate GV and deadlock freedom, and thus we pose our research question:
 
@@
 
-We make the following contributions:
+In this paper we make the following contributions:
 \begin{itemize}
-\item \emph{Priority GV} We present Priority GV (PGV) in~\cref{sec:pgv}, and prove subject reduction (\cref{thm:pgv-subject-reduction-confs}) and progress (\cref{thm:pgv-closed-progress-confs}).
+\item \emph{Priority GV.} We present Priority GV (PGV) in~\cref{sec:pgv}, and prove subject reduction (\cref{thm:pgv-subject-reduction-confs}) and progress (\cref{thm:pgv-closed-progress-confs}).
 Unlike  previous work on GV, structural congruence and reduction are type preserving in PGV.
-\item \emph{Milner's Cyclic Scheduler} To illustrate newfound expressiveness of PGV, we present Milner's Cyclic Scheduler~\cite{milner89} (\cref{sec:pgv-example}).
-\item \emph{Updated Priority CP} We present an updated version of Priority CP (PCP)~\cite{dardhagay18} in \cref{sec:pcp}.
-We remove commuting conversions and move away from reduction as cut elimination, towards reduction based on structural congruence, as one expects in a process calculus.
+\item \emph{Milner's Cyclic Scheduler.} To illustrate newfound expressiveness of PGV, we present Milner's Cyclic Scheduler~\cite{milner89} (\cref{sec:pgv-example}).
+\item \emph{Updated Priority CP.} We present an updated version of Priority CP~\cite{dardhagay18} in \cref{sec:pcp}.
+We remove commuting conversions and move away from reduction as cut elimination, towards reduction based on structural congruence, as in process calculi.
 We present Milner's Cyclic Scheduler in PCP (\cref{sec:pcp-example}).
-\item \emph{Translation from PCP to PGV} We present a translation from PCP to PGV in~\cref{sec:pcp} and prove that it preserves typing (\cref{thm:pcp-to-pgv-confs-preservation}) and satisfies operational correspondence (\cref{thm:pcp-to-pgv-operational-correspondence-soundness,thm:pcp-to-pgv-operational-correspondence-completeness}).
-We show that the scheduler in PCP (\cref{sec:pcp-example}) translates to the scheduler in PGV (\cref{sec:pgv-example}).
+\item \emph{Translation from PCP to PGV.} We present a translation from PCP to PGV in~\cref{sec:pcp} and prove that it preserves typing (\cref{thm:pcp-to-pgv-confs-preservation}) and satisfies operational correspondence (\cref{thm:pcp-to-pgv-operational-correspondence-soundness,thm:pcp-to-pgv-operational-correspondence-completeness}).
+Finally, the scheduler in PCP (\cref{sec:pcp-example}) translates to the scheduler in PGV (\cref{sec:pgv-example}).
 \end{itemize}

### 2020-10-16 priority-cp.tex
Fix typo.
final pass, hopefully

@@
 We present an updated version of Priority CP (PCP)~\cite{dardhagay18}, which is Wadler's CP~\cite{wadler12} with {priorities}, and a translation from PCP to PGV.
-The main difference with respect to PCP~\cite{dardhagay18} is removing commuting conversions, in favour of using structural congruence.
-Commuting conversions are necessary if we want our reduction strategy to correspond \emph{exactly} to cut (or cycle as in \cite{dardhagay18}) elimination.
+The main difference with respect to PCP~\cite{dardhagay18} is removing commuting conversions and using structural congruence.
+Commuting conversions are necessary if we want our reduction strategy to correspond \emph{exactly} to cut (or cycle in \cite{dardhagay18}) elimination.
 However, as Lindley and Morris~\cite{lindleymorris15} show, all $\beta$-reductions that can be performed \emph{with} the use of commuting conversions, can also be performed \emph{without} them.
-They show this by decomposing the reduction relation for CP into two phases: first, all $\beta$-reductions are performed, and second, commuting conversions are used to bring one particular action to the front, to eliminate the top-level cut.
+%They show this by decomposing the reduction relation for CP into two phases: first, all $\beta$-reductions are performed, and second, commuting conversions are used to bring one particular action to the front, to eliminate the top-level cut.
 
@@
 
-Recall our implementations of Milner's cyclic scheduler in PCP ($\pcp{\tm{\sched}}$, \cref{sec:pcp-example}) and PGV ($\pgv{\tm{\sched}}$, \cref{sec:pgv-example}).
+Recall our implementations of Milner's Cyclic Scheduler in PCP ($\pcp{\tm{\sched}}$, \cref{sec:pcp-example}) and PGV ($\pgv{\tm{\sched}}$, \cref{sec:pgv-example}).
 It follows immediately from the definition of our translation that the translation of $\pcp{\tm{\sched}}$ is exactly $\pgv{\tm{\sched}}$, where all processes are child processes.
 
-The translations $\tm{\cpgvC{\cdot}}$ and $\tm{\cpgvM{\cdot}}$ are preserve typing.
+The translations $\tm{\cpgvC{\cdot}}$ and $\tm{\cpgvM{\cdot}}$ preserve typing.
 The proofs are standard inductions.
@@
   \end{restatablelemma}
+  \begin{restatabletheorem}{thmpcptopgvoperationalcorrespondencesoundness}%
+    [Operational Correspondence, Soundness, ${\tm{\cpgvC{\cdot}}}$]
+    \label{thm:pcp-to-pgv-operational-correspondence-soundness}
+    \hfill\\%newline before theorem statement
+    If $\pcp{\seq{P}{\ty{\Gamma}}}$ and $\pgv{\tm{\cpgvC{P}}\cred\tm{\conf{C}}}$, there exists a $\tm{Q}$ such that $\pcp{\tm{P}\red^+\tm{Q}}$ and $\pgv{\tm{\conf{C}}\cred^\star\tm{\cpgvC{Q}}}$
+  \end{restatabletheorem}
+  % \begin{proof}
+  %   By induction on the derivation of $\pgv{\tm{\cpgvC{P}}\cred\tm{\conf{C}}}$.\simon{Surprising that you can get it by the derivation of the reduction relation---normally, 'soundness' would be proved by shunting into a canonical form and then proceeding by induction the structure of the term / induction on typing derivations.}
+  %   Details in \cref{prf:thm-pcp-to-pgv-operational-correspondence-soundness}.
+  % \end{proof}
+\end{compacttheorems}
 \begin{restatabletheorem}{thmpcptopgvoperationalcorrespondencecompleteness}%
@@
 % \end{proof}
-  \begin{restatabletheorem}{thmpcptopgvoperationalcorrespondencesoundness}%
-    [Operational Correspondence, Soundness, ${\tm{\cpgvC{\cdot}}}$]
-    \label{thm:pcp-to-pgv-operational-correspondence-soundness}
-    \hfill\\%newline before theorem statement
-    If $\pcp{\seq{P}{\ty{\Gamma}}}$ and $\pgv{\tm{\cpgvC{P}}\cred\tm{\conf{C}}}$, there exists a $\tm{Q}$ such that $\pcp{\tm{P}\red^+\tm{Q}}$ and $\pgv{\tm{\conf{C}}\cred^\star\tm{\cpgvC{Q}}}$
-  \end{restatabletheorem}
-  % \begin{proof}
-  %   By induction on the derivation of $\pgv{\tm{\cpgvC{P}}\cred\tm{\conf{C}}}$.\simon{Surprising that you can get it by the derivation of the reduction relation---normally, 'soundness' would be proved by shunting into a canonical form and then proceeding by induction the structure of the term / induction on typing derivations.}
-  %   Details in \cref{prf:thm-pcp-to-pgv-operational-correspondence-soundness}.
-  % \end{proof}
-\end{compacttheorems}
-
 \end{document}

### 2020-10-16 priority-gv.tex
final pass, hopefully

@@
 We present Priority GV (PGV), a session-typed functional language based on GV~\cite{wadler15,lindleymorris15} which uses priorities \`{a} la Kobayashi~\cite{kobayashi06,padovaninovara15} to enforce deadlock freedom.
-%The separation of channel creation and thread spawning in GV only allows tree-shaped communication structures, so it is trivially free from cycles, and therefore deadlocks.
-For e.g., if some term $\tm{L}$ has the means to communicate with both $\tm{M}$ and $\tm{N}$, then $\tm{M}$ and $\tm{N}$ cannot communicate amongst themselves.
-However, as with CP, this rules out \emph{all} cyclic communication structures, not just those with deadlocks.
 Priority GV offers a more fine-grained analysis of communication structures, and by separating channel creation form thread spawning it allows cyclic structures.
-We illustrate PGV with Milner's Cyclic Scheduler~\cite{milner89}.
+We start by illustrating PGV with Milner's Cyclic Scheduler~\cite{milner89}.
 
@@
 Suppose a set of processes $\tm{\proc{i}}$, $1\leq{i}\leq{n}$, is scheduled to perform some tasks in cyclic order, starting with $\tm{\proc1}$, ending with $\tm{\proc{n}}$, and notifying $\tm{\proc1}$ when all processes have finished.
-% The dotted lines represent the optional transfer of data between processes.
-With three processes, the scheduler looks as follows:\simon{Sorry, I found this diagram super confusing\dots Is there any way you could make it look more cyclic?
-Also, have the convention of a single annotation per channel, with the convention that the greater agent / proc gets the primed end of the previous channel?}
-% \input{fig/pgv-scheduler.tex}
 We implement the scheduler as follows, using $\tm{\prod_{I}\conf{C}_i}$ to denote the parallel composition of the processes $\tm{\conf{C}_i}$, $\tm{i}\in\tm{I}$, and $\tm{\plug{M}{N}}$ to denote the plugging of $\tm{N}$ in the one-hole term-context $\tm{M}$:
@@
 Each of the channels in the scheduler is of a terminated type, and is merely used to synchronise.
-For simplicity, we let $\tm{\sched}$ be a configuration. However, we could equivalently have represented it as a function which takes three processes as arguments, and creates this configuration.
+For simplicity, we let $\tm{\sched}$ be a configuration.
+However, we could equivalently have represented it as a function which takes three processes as arguments, and creates this configuration.
+A~diagram of our scheduler with three processes can be found in \Cref{fig:pgv-scheduler}.
 \end{example}
@@
   \label{def:pgv-actions}
-  A~term \emph{acts on} an endpoint $\tm{x}$ if it is of the form $\tm{\send\;\pair{V}{x}}$, $\tm{\recv\;{x}}$, $\tm{\close\;{x}}$, or $\tm{\wait\;{x}}$.
+  A~term \emph{acts on} an endpoint $\tm{x}$ if it is $\tm{\send\;\pair{V}{x}}$, $\tm{\recv\;{x}}$, $\tm{\close\;{x}}$, or $\tm{\wait\;{x}}$.
 A~term is an \emph{action} if it acts on some endpoint $\tm{x}$.
@@
 \end{compacttheorems}
-
+\Cref{thm:pgv-closed-progress-confs} and \cref{cor:pgv-closed-progress} state that a configuration in PGV will complete all its communication actions, thus satisfying deadlock freedom or reduce to a term which is a value in PGV and contains no channel endpoints.
 \end{document}

### 2020-10-16 conclusion.tex
Final version.

@@
 
-\subsubsection*{Conclusion and Future Work}
+\subsubsection*{Conclusion and future work}
 We presented Priority GV, a~session-typed functional language which uses priorities to ensure deadlock freedom, and showed its relation to Priority CP~\cite{dardhagay18} via an operational correspondence.

### 2020-10-16 priority-gv.tex
Final version.

@@
 Paraphrasing Dardha and Gay~\cite{dardhagay18}, priorities obey the following two laws:
-\begin{enumerate}
-  \item an action with priority $\cs{o}$ happens before an action with priority $\cs{o+n}$;
+\begin{enumerate*}[label=(\roman*)]
+\item an action with priority $\cs{o}$ happens before an action with priority $\cs{o+n}$; and
 \item communication requires \emph{equal} priorities for the dual actions.
-\end{enumerate}
+\end{enumerate*}
 
@@
 
-While our configuration typing is based on the standard typing for GV, it differs on two counts
-$(i)$ we require that child threads return the unit value, as opposed to a terminated channel;
-$(ii)$ we simplify typing for parallel composition.
+While our configuration typing is based on the standard typing for GV, it differs on two counts:
+\begin{enumerate*}[label=(\roman*)]
+\item we require that child threads return the unit value, as opposed to a terminated channel; and
+\item we simplify typing for parallel composition.
+\end{enumerate*}
 In order to guarantee deadlock freedom, in GV each parallel composition must split \emph{exactly one} channel of the channel pseudo-type $\ty{S^\sharp}$ into two endpoints of type $\ty{S}$ and $\ty{\co{S}}$.
@@
 \end{compacttheorems}
-\Cref{thm:pgv-closed-progress-confs} and \cref{cor:pgv-closed-progress} state that a configuration in PGV will complete all its communication actions, thus satisfying deadlock freedom or reduce to a term which is a value in PGV and contains no channel endpoints.
+It follows from \cref{thm:pgv-closed-progress-confs} and \cref{cor:pgv-closed-progress} that a term which does not return an endpoint will complete all its communication actions, thus satisfying deadlock freedom.
 \end{document}

### 2020-10-16 priority-cp.tex
give space to results

@@
 
-Our translation satisfies operational correspondence.
-We link our translations, and show that $\tm{\cpgvM{P}}$ \emph{must} reduce to $\tm{\cpgvC{P}}$, \ie $\tm{\cpgvC{\cdot}}$ is equivalent to translating with $\tm{\cpgvM{\cdot}}$ and performing several reduction steps.
-We use this fact to prove operational correspondence.
+Our translation satisfies soundness and completeness of operational correspondence.
+We link ${\tm{\cpgvM{\cdot}}}$ and ${\tm{\cpgvC{\cdot}}}$ and show that $\tm{\cpgvM{P}}$ \emph{must} reduce to $\tm{\cpgvC{P}}$, \ie $\tm{\cpgvC{\cdot}}$ is equivalent to translating with $\tm{\cpgvM{\cdot}}$ and performing several reduction steps.
+We use this fact to prove the results below.
 The details are in~\cref{prf:lem-pcp-to-pgv-cpgvM-to-cpgvC,prf:thm-pcp-to-pgv-operational-correspondence-completeness,prf:thm-pcp-to-pgv-operational-correspondence-soundness}:
-\begin{compacttheorems}
   \begin{restatablelemma}{lempcptopgvcpgvMtocpgvC}
@@
   % \end{proof}
-\end{compacttheorems}
 \begin{restatabletheorem}{thmpcptopgvoperationalcorrespondencecompleteness}%
@@
   If $\pcp{\seq{P}{\ty{\Gamma}}}$ and $\pcp{\tm{P}\red\tm{Q}}$,
-  then $\pgv{\tm{\cpgvC{P}}\cred^+\tm{\cpgvC{Q}}}$. Details in~\cref{prf:lem-pcp-to-pgv-cpgvM-to-cpgvC}.
+  then $\pgv{\tm{\cpgvC{P}}\cred^+\tm{\cpgvC{Q}}}$.
 \end{restatabletheorem}

### 2020-12-08 rebuttal.txt
Added comments.

----------------------- REVIEW 1 ---------------------

> The key elements of the type system, particularly the annotations in arrow types, the shape of typing judgments, the typing of communication primitives, are basically taken from [37]. The authors also move away from a calculus strictly justified by logic, e.g. by resorting to traditional process-algebraic structural congruence and reduction relations, and they work with a hybrid form of (partially encoded) session types. Overall, I see a substantial overlap between this paper and [37], which somehow lessens the depth of the contribution.

> I appreciate the effort made by the authors of tackling an example from the bibliography (Milner's cyclic scheduler), but I'm not convinced that presenting that particular example alone, without even trying to showcase a typing derivation, is an effective strategy. I gained more insights about the key aspects of the typing discipline by looking at [37], in which many (small) examples are used to illustrate the type system.

* Otherwise, the paper is technically sound as far as I checked, it is well-written and quite accessible.

In summary, this is a fine paper that continues an established investigation on the use of priorities for enabling a more flexible use of session channels, but it lacks a particularly compelling message to convey.

## On the choice of the example(s)

As presented, Example 1 (also the PCP version) is ill-typed and, I think, not self contained. Note that d is used while d' is not and there is a restriction for d and d'. I think a wait d' is missing somewhere, possibly next to proc‚ÇÅ if I interpret the diagram in the Appendix correctly.

> Also, I don't understand what M_i and N_i are (similarly for P_i and Q_i in the PCP version of the example). Maybe the authors imply that the actual system is more complex that the one shown here, where M_i and N_i contain unspecified parts of the system. However, I think that the example is already more complex than necessary, and that it would be fine to simply state that all that matters is the communication structure being modeled.

That’s exactly what they’re meant to convey, yes.

> Speaking of complexity, the case for PGV can be made by considering
>
>  	(aa')(bb') (close a; wait b' || wait a'; close b)
>
> which is a streamlined version of the example in which there are just two processes connected by two sessions. This is the essential configuration that cannot be modeled by the calculi strictly based on linear logic, since the two threads share two channels. Yet, this configuration is well typed in the presented type system and therefore is deadlock free.
>
> This example is simple enough that a full typing derivation can be shown in the paper, possibly at the expense of the few proof sketched which can be moved in the appendix anyway.

That’s probably a good call.

> At the same time, the example given by the authors does not make use of any interesting function (apart from the constants wait and close) and therefore it fails to illustrate the relevance of annotations on arrow types.

We should probably add an example which uses closures.

> I presume there are side conditions in SC-RESCOMM requiring that {z,w} \cap {x,y} = \emptyset.
> I presume there are side conditions in E-NEW requiring that x,y do not occur free in F

Yes, to both.

> The encoding from PCP to PGV is routine and does not reveal particularly challenging or surprising aspects given the expressiveness of PGV. In fact, I suspect it can be further simplified (at least in its presentation) by considering only the encoding into the functional fragment of PGV. Clearly the translation to configurations deals with a subset of cases compared to the translation to terms. I wonder if it's possible to *present* the second translation only, which alone is complete as far as the encoding goes.

Such a translation would be complete, but not sound.

----------------------- REVIEW 2 ---------------------

The paper fits well the scope of FOSSACS.

> Counter-examples, i.e., deadlocking processes that cannot be typed, such as dining philosophers.

Yes.

> At least one example demonstrating the interaction between higher order and deadlock-freedom, i.e., an example in which a function is sent and applied and where the actions present in the function's body are controlled by the latent effect of the arrow type.

Yes.

----------------------- REVIEW 3 ---------------------

> As a non-expert reviewing the paper, my main concern is the amount of novel technical content: the technique in use has been developed by one of the authors in the context of a process calculus and is adapted here to a known linear concurrent λ-calculus. This would not be an issue if the attempt exposed some technical difficulties in adapting the technique to the setting of GV or uncovered some important insights; however, this does not seem to be the case. Thus, from the perspective of a non-expert, this paper seems to be too closely related to prior work to warrant an acceptance recommendation. Still, particularly as a non-expert, I would welcome any comments that would highlight the significance of technical contributions of the paper, and if so, would be happy to adjust my assessment.

### 2020-12-08 rebuttal.txt
Minor changes to rebuttal.

@@
 ----------------------- REVIEW 1 ---------------------
-
-> The key elements of the type system, particularly the annotations in arrow types, the shape of typing judgments, the typing of communication primitives, are basically taken from [37].
-The authors also move away from a calculus strictly justified by logic, e.g.
-by resorting to traditional process-algebraic structural congruence and reduction relations, and they work with a hybrid form of (partially encoded) session types.
-Overall, I see a substantial overlap between this paper and [37], which somehow lessens the depth of the contribution.
-
-> I appreciate the effort made by the authors of tackling an example from the bibliography (Milner's cyclic scheduler), but I'm not convinced that presenting that particular example alone, without even trying to showcase a typing derivation, is an effective strategy.
-I gained more insights about the key aspects of the typing discipline by looking at [37], in which many (small) examples are used to illustrate the type system.
-
-* Otherwise, the paper is technically sound as far as I checked, it is well-written and quite accessible.
-
-In summary, this is a fine paper that continues an established investigation on the use of priorities for enabling a more flexible use of session channels, but it lacks a particularly compelling message to convey.
-
-## On the choice of the example(s)
-
-As presented, Example 1 (also the PCP version) is ill-typed and, I think, not self contained.
-Note that d is used while d' is not and there is a restriction for d and d'.
-I think a wait d' is missing somewhere, possibly next to proc‚ÇÅ if I interpret the diagram in the Appendix correctly.
 

### 2020-12-08 rebuttal.txt
Added todos.

@@
+TODO:
+
+- a word of thanks;
+- acknowledge the lack of examples, and commit to addressing that;
+- clarify our contribution: priority gv is the first session-typed functional language which with priorities and strong ties to linear logic via its correspondence with PCP and its relation to the original CP and GV languages;
+
 ----------------------- REVIEW 1 ---------------------

### 2020-12-09 rebuttal.txt
Update rebuttal.txt

--- 2020-12-08 rebuttal.txt
@@
+We thank the reviewers for their thorough reviews and very useful feedback.
+
+First of all, we acknowledge the lack of examples and we commit to adding them.
+We will illustrate PGV with further well-typed programs together with type derivation; and deadlocked programs (including the example by Reviewer 1).
+
+The reason for including cyclic scheduler was deliberately to connect our work to the previous FoSSaCS work on PCP.
+We wanted to show how the well-known Milner's scheduler can be written in PGV (for the first time) and use it to illustrate our encoding of PGV to PCP, thus closing this circle.
+
+We will clarify our contributions in the introduction.
+Summing them up here: PGV is the first session-typed functional language with priorities *and* strong ties to linear logic via its correspondence with PCP and its relation to the CP and GV languages.
+
 ----------------------- REVIEW 1 ---------------------

### 2020-12-09 rebuttal.txt
Update rebuttal.txt

--- 2020-12-08 rebuttal.txt
@@
+We thank the reviewers for their thorough reviews and very useful feedback, which we'll address as follows.
+
+First of all, we acknowledge the lack of examples and we commit to adding them to the (potential) final version.
+We will illustrate PGV by adding examples for each of the following:
+-- deadlock-free processes (including the streamlined version with two processes connected by two sessions, suggested by Reviewer 1) together with their typing derivations;
+-- deadlocked processes that cannot be typed, thus illustrating how priorities work, suggested by Reviewers 1 and 2;
+-- processes which use closures, thus illustrating the relevance of annotations on arrow types, suggested by Reviewer 1;
+-- processes demonstrating the interaction between higher order and deadlock freedom, suggested by Reviewer 2.
+
+Regarding the cyclic scheduler, we included it to deliberately connect our work to the previous FoSSaCS work on PCP.
+We wanted to show how the well-known Milner's scheduler can be written in PGV (for the first time) and use it to illustrate our encoding of PCP to PGV, thus closing this circle on the cyclic scheduler.
+We will make this clearer in the paper.
+
+Regarding our contributions, we will highlight them better in the introduction.
+Summing them up here: PGV is the first session-typed functional language with priorities *and* strong ties to linear logic via its correspondence with PCP and its relation to the CP and GV languages.
+
+Lastly, we will implement all minor comments by all reviewers, thanks for pointing them out.
+
 ----------------------- REVIEW 1 ---------------------
+Regarding the overlap with [37] we would like to mention that our work was developed completely independently of [37] as the authors came to know this work only during the writing up of the paper while working on related works.
+We still want to emphasise the differences between our work and [37]: i) our work has strong ties to logic, hence why we also present the encoding PGV->PCP; ii) COMPLETE... 
+We will state the above clearly in the paper.
 
-> Also, I don't understand what M_i and N_i are (similarly for P_i and Q_i in the PCP version of the example).
-Maybe the authors imply that the actual system is more complex that the one shown here, where M_i and N_i contain unspecified parts of the system.
-However, I think that the example is already more complex than necessary, and that it would be fine to simply state that all that matters is the communication structure being modeled.
+> As presented, Example 1 (also the PCP version) is ill-typed and, I think, not self contained.
+Note that d is used while d' is not and there is a restriction for d and d'.
+I think a wait d' is missing somewhere, possibly next to proc₁ if I interpret the diagram in the Appendix correctly.
 
-That’s exactly what they’re meant to convey, yes.
+Regarding M_i and N_i in the PCP cyclic scheduler, the reviewer understood this correctly and we will simply state that what matters is the communication structure being modelled.
 
-> Speaking of complexity, the case for PGV can be made by considering
->
->  	(aa')(bb') (close a; wait b' || wait a'; close b)
->
-> which is a streamlined version of the example in which there are just two processes connected by two sessions.
-This is the essential configuration that cannot be modeled by the calculi strictly based on linear logic, since the two threads share two channels.
-Yet, this configuration is well typed in the presented type system and therefore is deadlock free.
->
-> This example is simple enough that a full typing derivation can be shown in the paper, possibly at the expense of the few proof sketched which can be moved in the appendix anyway.
+Regarding the encoding of PCP to PGV and presenting only the encoding to the functional fragment, we have thought about this too. However, such encoding is complete but not sound. The reason being...
 
-That’s probably a good call.
+Minor:
 
-> At the same time, the example given by the authors does not make use of any interesting function (apart from the constants wait and close) and therefore it fails to illustrate the relevance of annotations on arrow types.
+Regarding side conditions in SC-RESCOMM and E-NEW, the reviewer is right and we will add them.
 
-We should probably add an example which uses closures.
-
-> I presume there are side conditions in SC-RESCOMM requiring that {z,w} \cap {x,y} = \emptyset.
-> I presume there are side conditions in E-NEW requiring that x,y do not occur free in F
-
-Yes, to both.
-
-> The encoding from PCP to PGV is routine and does not reveal particularly challenging or surprising aspects given the expressiveness of PGV.
-In fact, I suspect it can be further simplified (at least in its presentation) by considering only the encoding into the functional fragment of PGV.
-Clearly the translation to configurations deals with a subset of cases compared to the translation to terms.
-I wonder if it's possible to *present* the second translation only, which alone is complete as far as the encoding goes.
-
-Such a translation would be complete, but not sound.
+Regarding T-VAR and T-CONST we will mention your suggestion.
 
@@
 
-The paper fits well the scope of FOSSACS.
+As mentioned at the top, we will add more examples to the paper.
 
-> Counter-examples, i.e., deadlocking processes that cannot be typed, such as dining philosophers.
+Regarding the presentation of the main sections: we will explain more in detail how priorities work, especially with the aid of the aforementioned examples.
+We will systematise the the prevention of CP, PCP, GV and PGV, by dedicating subsections to each of them before diving into the details of our work.
 
-Yes.
-
-> At least one example demonstrating the interaction between higher order and deadlock-freedom, i.e., an example in which a function is sent and applied and where the actions present in the function's body are controlled by the latent effect of the arrow type.
-
-Yes.
+Regarding our contribution, we will make them clearer for non-experts by first illustrating the use and benefits of GV and then the benefits of splitting new and spawn in PGV and its ties to logic (this will relate to the CP,PCP,GV,PGV explanations above).
 
@@
 
-> As a non-expert reviewing the paper, my main concern is the amount of novel technical content: the technique in use has been developed by one of the authors in the context of a process calculus and is adapted here to a known linear concurrent λ-calculus.
-This would not be an issue if the attempt exposed some technical difficulties in adapting the technique to the setting of GV or uncovered some important insights; however, this does not seem to be the case.
-Thus, from the perspective of a non-expert, this paper seems to be too closely related to prior work to warrant an acceptance recommendation.
-Still, particularly as a non-expert, I would welcome any comments that would highlight the significance of technical contributions of the paper, and if so, would be happy to adjust my assessment.
+We will clarify the contributions better, this was also suggested by other reviewers.
+
+It is true that priorities are not new and they were already used in PCP, which is a process calculus ties to logic.
+However, PGV is the first functional calculus with priorities that also ties to logic as PCP.
+
+We are in fact making this connection to linear logic explicit by defining a sound encoding of PCP to PGV.
+Furthermore, as a case study for PGV and the encoding, we deliberately chose Milner's cyclic scheduler presented in PCP previously.
+The reason for this choice is to show how a process in the literature of PCP can be translated using our encoding into PGV.
+
+However, in addition to the cyclic scheduler, we will add the examples mentioned at the top of this letter to better illustrate PGV and its type system.

### 2020-12-09 rebuttal.txt
Editted the rebuttal

@@
 
-First of all, we acknowledge the lack of examples and we commit to adding them to the (potential) final version.
+First of all, we acknowledge the lack of examples and we commit to adding them to the potential final version.
 We will illustrate PGV by adding examples for each of the following:
--- deadlock-free processes (including the streamlined version with two processes connected by two sessions, suggested by Reviewer 1) together with their typing derivations;
--- deadlocked processes that cannot be typed, thus illustrating how priorities work, suggested by Reviewers 1 and 2;
--- processes which use closures, thus illustrating the relevance of annotations on arrow types, suggested by Reviewer 1;
--- processes demonstrating the interaction between higher order and deadlock freedom, suggested by Reviewer 2.
+
+  - deadlock-free processes—including the streamlined version with two processes connected by two sessions, suggested by Reviewer 1—together with their typing derivations;
+  - deadlocked processes that cannot be typed, thus illustrating how priorities work, suggested by Reviewers 1 and 2;
+  - processes which use closures, thus illustrating the relevance of annotations on arrow types, suggested by Reviewer 1;
+  - processes demonstrating the interaction between higher order and deadlock freedom, suggested by Reviewer 2.
 
@@
 
-Regarding our contributions, we will highlight them better in the introduction.
-Summing them up here: PGV is the first session-typed functional language with priorities *and* strong ties to linear logic via its correspondence with PCP and its relation to the CP and GV languages.
+Regarding our contributions, we will highlight them better in the introduction. Summing them up here:
+  - PGV is the first session-typed functional language with priorities *and* strong ties to linear logic via its correspondence with PCP and its relation to the CP and GV languages.
+  - PGV addresses several problems in the original GV language, most notably:
+    * our version does not have the pseudo-type S^#; and
+    * our structural congruence is type preserving.
 
@@
 Regarding the overlap with [37] we would like to mention that our work was developed completely independently of [37] as the authors came to know this work only during the writing up of the paper while working on related works.
-We still want to emphasise the differences between our work and [37]: i) our work has strong ties to logic, hence why we also present the encoding PGV->PCP; ii) COMPLETE... 
+We still want to emphasise the differences between our work and [37]:
+
+  - PGV has a clear correspondence with a process calculus, PCP;
+  - PGV has strong ties to linear logic, via its correspondence to PCP and its transitive relation to CP;
+
 We will state the above clearly in the paper.
@@
 > As presented, Example 1 (also the PCP version) is ill-typed and, I think, not self contained.
-Note that d is used while d' is not and there is a restriction for d and d'.
-I think a wait d' is missing somewhere, possibly next to proc₁ if I interpret the diagram in the Appendix correctly.
+> Note that d is used while d' is not and there is a restriction for d and d'.
+> I think a `wait d'` is missing somewhere, possibly next to proc_1 if I interpret the diagram in the Appendix correctly.
 
 Regarding M_i and N_i in the PCP cyclic scheduler, the reviewer understood this correctly and we will simply state that what matters is the communication structure being modelled.
+The missing `wait d'` should be in N_1, as the reviewer correctly inferred.
+We will mention this clearly in the paper.
 
-Regarding the encoding of PCP to PGV and presenting only the encoding to the functional fragment, we have thought about this too. However, such encoding is complete but not sound. The reason being...
+Regarding the encoding of PCP to PGV and presenting only the encoding to the functional fragment, we have thought about this too.
+Such an encoding is complete but not sound, as the translation to configurations does some evaluation of top-level new/spawn constructs.
+For instance, the process ‘x[y].P’ has no reductions.
+Similarly, its translation to a configuration has no reductions.
+However, its translation to a term, ‘o let (y,z) = new in let x = send (z,x) in ((P))_M’ can reduce to ‘(vyz)(o let x = send (z,x) in ((P))_M)’.
 

### 2020-12-09 rebuttal.txt
Update rebuttal.txt

@@
 
-Regarding the cyclic scheduler, we included it to deliberately connect our work to the previous FoSSaCS work on PCP.
-We wanted to show how the well-known Milner's scheduler can be written in PGV (for the first time) and use it to illustrate our encoding of PCP to PGV, thus closing this circle on the cyclic scheduler.
-We will make this clearer in the paper.
+Regarding the cyclic scheduler, we included it to deliberately connect our work to the previous FoSSaCS work on PCP where the cyclic scheduler was originally presented as an example of a deadlock-free process not typable in CP.
+In our work, we wanted to show how the well-known Milner's scheduler can be written in PGV-for the first time-and use it to illustrate our encoding from PCP to PGV, thus closing the circle on the cyclic scheduler example.
+We will make this more explicit in the paper.
 
 Regarding our contributions, we will highlight them better in the introduction. Summing them up here:
-  - PGV is the first session-typed functional language with priorities *and* strong ties to linear logic via its correspondence with PCP and its relation to the CP and GV languages.
-  - PGV addresses several problems in the original GV language, most notably:
-    * our version does not have the pseudo-type S^#; and
-    * our structural congruence is type preserving.
+  1. PGV is the first session-typed functional language with priorities *and* strong ties to linear logic via its correspondence with PCP and its relation to the CP and GV languages.
+  2. PGV addresses several problems in the original GV language, most notably:
+    a. our version does not have the pseudo-type S^#; and
+    b. our structural congruence is type preserving.
+  3. The connection of PGV to logic is given via a sound and complete encoding of PCP to PGV and showcase it with the encoding of cyclic scheduler to PGV.
 
-Lastly, we will implement all minor comments by all reviewers, thanks for pointing them out.
+We would like to draw attention to a. and b. above.
+Previous work on GV does not satisfy a.
+and the pseudo-type S^# is explicitly included in the type language, but then explicitly excluded in all the typing rules where only "concrete" types, like (input) ?S or (output) !S can be used.
+Consequently, the technical presentation is cumbersome in original GV.
+Another major issue with GV is that structural congruence is not type preserving.
+Consequently, in order to prove the main results in GV, one would first need to prove that: any ill-typed term which is obtained via structural congruence (or reduction) starting from a well-typed term, can be rewritten into a well-typed term.
+In our work, the type preserving structural congruence makes the presentation cleaner and more elegant and aligns for example with structural congruence in process calculi.
+
+Lastly, we will implement all "minor comments" by all reviewers, thanks for pointing them out.
 
 ----------------------- REVIEW 1 ---------------------
-Regarding the overlap with [37] we would like to mention that our work was developed completely independently of [37] as the authors came to know this work only during the writing up of the paper while working on related works.
-We still want to emphasise the differences between our work and [37]:
+Regarding the overlap with [37] we would like to point out that our work was developed completely independently of [37].
+The authors were not aware of[37] until working on the related work of this paper.
+We still want to emphasise the differences between our work and [37], which we will also state in the paper:
 
@@
 
-We will state the above clearly in the paper.
-
-> As presented, Example 1 (also the PCP version) is ill-typed and, I think, not self contained.
-> Note that d is used while d' is not and there is a restriction for d and d'.
-> I think a `wait d'` is missing somewhere, possibly next to proc_1 if I interpret the diagram in the Appendix correctly.
-
 Regarding M_i and N_i in the PCP cyclic scheduler, the reviewer understood this correctly and we will simply state that what matters is the communication structure being modelled.
-The missing `wait d'` should be in N_1, as the reviewer correctly inferred.
-We will mention this clearly in the paper.
+The missing `wait d'` should be in N_1, as the reviewer correctly inferred. Moreover, with this change the process is well typed. We will clarify this in the paper.
 
@@
 
-Regarding T-VAR and T-CONST we will mention your suggestion.
+Regarding T-VAR and T-CONST we will mention your suggestion. Thanks
 
@@
 
-As mentioned at the top, we will add more examples to the paper.
+As mentioned at the top of this letter, we will add more examples in the paper.
 
 Regarding the presentation of the main sections: we will explain more in detail how priorities work, especially with the aid of the aforementioned examples.
-We will systematise the the prevention of CP, PCP, GV and PGV, by dedicating subsections to each of them before diving into the details of our work.
+We will systematise the prevention of CP, PCP, GV and PGV, by dedicating subsections to each of them before diving into the details of our work.
 
-Regarding our contribution, we will make them clearer for non-experts by first illustrating the use and benefits of GV and then the benefits of splitting new and spawn in PGV and its ties to logic (this will relate to the CP,PCP,GV,PGV explanations above).
+Regarding our contribution, we will make them clearer for non-experts, which we summarised at the top of the letter.
+We will illustrate the use and benefits of GV and then the benefits of splitting new and spawn in PGV and its ties to logic (this will relate to the CP,PCP,GV,PGV explanations that we will add).
 
@@
 
-We will clarify the contributions better, this was also suggested by other reviewers.
+We will clarify the contributions better, this was also suggested by other reviewers and we summarised them at the top.
 
-It is true that priorities are not new and they were already used in PCP, which is a process calculus ties to logic.
-However, PGV is the first functional calculus with priorities that also ties to logic as PCP.
+A key contribution of our work is the connection of PGV with logic, which we make explicit by defining a sound and complete encoding of PCP to PGV.
+Furthermore, as a case study for PGV and the encoding, we deliberately chose Milner's cyclic scheduler presented in PCP.
 
-We are in fact making this connection to linear logic explicit by defining a sound encoding of PCP to PGV.
-Furthermore, as a case study for PGV and the encoding, we deliberately chose Milner's cyclic scheduler presented in PCP previously.
-The reason for this choice is to show how a process in the literature of PCP can be translated using our encoding into PGV.
-
-However, in addition to the cyclic scheduler, we will add the examples mentioned at the top of this letter to better illustrate PGV and its type system.
+As mentioned, in addition to the cyclic scheduler, we will add the examples at the top of this letter to better illustrate PGV and its type system.

### 2020-12-09 rebuttal.txt
Update rebuttal.txt

@@
 
-  - PGV has a clear correspondence with a process calculus, PCP;
-  - PGV has strong ties to linear logic, via its correspondence to PCP and its transitive relation to CP;
+  - PGV has a clear correspondence with a *process calculus*, PCP;
+  - PGV has strong ties to *linear logic*, via its correspondence to PCP and its transitive relation to CP;
 

### 2020-12-10 rebuttal.txt
Spacing.
Update rebuttal.txt

@@
 ----------------------- REVIEW 1 ---------------------
+
 Regarding the overlap with [37] we would like to point out that our work was developed completely independently of [37].
 The authors were not aware of[37] until working on the related work of this paper.
-We still want to emphasise the differences between our work and [37], which we will also state in the paper:
+We still want to emphasise the differences between our work and [37], which we will also state in the paper. PGV stems from GV differently from [37] and:
 

### 2021-01-05 introduction.tex
Processed minor comments and inserted unresolved comments as todo notes.

@@
 \section{Introduction}
+\todo{The first three sentences of the introduction are too general to motivate the paper and feel a bit out of place. Consider removing them.}
 Communication and concurrency are at the basis of our modern society.
@@
 Session types are a type formalism which can be used to specify and verify communication protocols between two or more communicating agents~\cite{honda93,takeuchihonda94,hondavasconcelos98,carbonehonda07}.
+\todo{These two settings are very different.
+The $\pi$-calculus is a mobile process algebra preceding inside which session types were developed and GV is an ad-hoc calculus made to demonstrate how propositions can be used as types.}
 Session types have been studied for many programming paradigms, but to date the most successful settings remain the $\pi$-calculus~\cite{sangiorgiwalker01}---a process calculus for communication and concurrency, and Good Variation (GV)~\cite{wadler15,lindleymorris15}---a linear concurrent $\lambda$-calculus.
@@
 Unfortunately, as with CP, this is restrictive as it limits communication structures to trees.
-There are several benefits of working with a functional language as opposed to process calculi.
+There are several benefits of working with a functional language as opposed to name-passing calculi.
 Functional languages support higher-order functions, and have a capability for abstraction not usually present in process calculi.

### 2021-01-05 priority-cp.tex
Processed minor comments and inserted unresolved comments as todo notes.

@@
 We present an updated version of Priority CP (PCP)~\cite{dardhagay18}, which is Wadler's CP~\cite{wadler12} with {priorities}, and a translation from PCP to PGV.
+\todo{This one is a bit difficult to understand without context: in [14] it appears inside a typed semantics context, where changing dynamically priorities make sense.
+This is quite different from what is presented here.
+({It would certainly be prudent of us that we do not need to dynamically change priorities in the updated version.})}
 The main difference with respect to PCP~\cite{dardhagay18} is removing commuting conversions and using structural congruence.

### 2021-01-05 priority-gv.tex
give space to results
Processed minor comments and inserted unresolved comments as todo notes.

@@
 
-We present Priority GV (PGV), a session-typed functional language based on GV~\cite{wadler15,lindleymorris15} which uses priorities \`{a} la Kobayashi~\cite{kobayashi06,padovaninovara15} to enforce deadlock freedom.
+We present Priority GV (PGV), a session-typed functional language based on GV~\cite{wadler15,lindleymorris15} which uses priorities \`{a} la Padovani~\cite{kobayashi06,padovaninovara15} to enforce deadlock freedom.
 Priority GV offers a more fine-grained analysis of communication structures, and by separating channel creation form thread spawning it allows cyclic structures.
 We start by illustrating PGV with Milner's Cyclic Scheduler~\cite{milner89}.
+
+\todo{The $\phi$ and $\child$ symbols are not explained at this stage and become clear only later. Either warn the reader or remove them altogether.}
+
+\todo{An explanation is needed about what M and N represents and how the hole-construction models this "signal b when finished" behaviour.}
+
+\todo{It is difficult to understand how a "close d" operation is supposed to work inside a (nu d d') construct without a corresponding wait d.}
 
@@
 Each agent signals their process to start on $\tm{a_i}$, waits for their process to finish on $\tm{b_i}$, and then signals for the next agent to continue on $\tm{c_i}$.
-The agent $\tm{\agent0}$ initiates, then waits for every other process to finish, and signals $\tm{\proc1}$ on $\tm{d}$.
+The agent $\tm{\agent1}$ initiates, then waits for every other process to finish, and signals $\tm{\proc1}$ on $\tm{d}$.
 Every other agent $\tm{\agent{i}}$, $2\leq{i}\leq{n}$ waits on $\tm{c'_{i-1}}$ for the signal to start.
-Each of the channels in the scheduler is of a terminated type, and is merely used to synchronise.
-For simplicity, we let $\tm{\sched}$ be a configuration.
-However, we could equivalently have represented it as a function which takes three processes as arguments, and creates this configuration.
+Each of the channels in the scheduler is of a terminated type, and is merely used to synchronise. For simplicity, we let $\tm{\sched}$ be a configuration.
 A~diagram of our scheduler with three processes can be found in \Cref{fig:pgv-scheduler}.
@@
 \subsubsection*{Environments}
-Typing environments $\ty{\Gamma}$, $\ty{\Delta}$ associate names to types.
+Typing environments $\ty{\Gamma}$, $\ty{\Delta}$ associate types to names.
 Environments are linear, so two environments can only be combined as $\ty{\Gamma}, \ty{\Delta}$ if their names are distinct, \ie $\fv(\ty{\Gamma})\cap\fv(\ty{\Delta})=\varnothing$.
@@
 Duality plays a crucial role in session types.
-The two endpoints of a channel are assigned dual types, which ensures that, for instance, whenever one program \emph{sends} a value on a channel, the program on the other end is waiting to \emph{receive}.
+The two endpoints of a channel are assigned dual types, ensuring that, for instance, whenever one program \emph{sends} a value on a channel, the program on the other end is waiting to \emph{receive}.
 Each session type $\ty{S}$ has a dual, written $\ty{\co{S}}$.
@@
   \minpr(\ty{\tylolli[\cs{p},\cs{q}]{T}{U}}) & = & \cs{p} \\
-  \minpr(\ty{\Gamma}, \tmty{x}{A})           & = & \minpr(\ty{\Gamma})\sqcap\pr(\ty{A})
+  \minpr(\ty{\Gamma}, \tmty{x}{A})           & = & \minpr(\ty{\Gamma})\sqcap\minpr(\ty{A})
 \end{array}
@@
 We use syntactic sugar to make terms more readable: we write $\tm{\andthen{M}{N}}$ in place of $\tm{\letunit{M}{N}}$, $\tm{\letbind{x}{M}{N}}$ in place of $\tm{(\lambda x.N)\;M}$, and pattern matching functions $\tm{\lambda\unit.M}$ in place of $\tm{\lambda z.\letunit{z}{M}}$ and $\tm{\lambda\pair{x}{y}.M}$ in place of $\tm{\lambda z.\letpair{x}{y}{z}{M}}$.
+\todo{Confusion arises from the difference between what is considered a native operator and a shortcut.
+For instance, "unit let" and "pair let" are native (representing sequence and new channel decomposition) while "singleton let" is a shortcut.
+The initial GV syntax was clear, this one is a bit confusing and might need some cleaning.
+At least explain why M;N cannot be a primitive construct.}
 % We recover $\tm{\fork}$ as $\tm{\lambda x.\letpair{y}{z}{\new\;\unit}{\andthen{\spawn\;{(\lambda\unit.x\;y)}}{z}}}$.
@@
 \subsubsection*{Internal and External Choice}
+\todo{You might want to clarify (repeat) that external choice can be modeled by sum type communication only because the language uses linear channels.}
+\todo{We also might want to emphasise that the modelling of choice in GV by~\cite{lindleymorris15} is unsound.}
 Typically, session-typed languages feature constructs for internal and external choice.
@@
 In GV, $\tm{\link}$ needs to return a terminated channel after evaluating the link.
-In PGV, the behaviour matches that of link in CP.
+In PGV, the behaviour matches that of link in CP.\todo{Please clarify in what way it matches the behaviour of link in CP.}
 Secondly, our structural congruence is type preserving.
@@
 \subsubsection*{Typing}
+\todo{This is a type-and-effect system and should probably be presented as such, as in [37].}
 \Cref{fig:pgv-typing} gives the typing rules for PGV.
@@
 Constants $\tm{K}$ are typed using type schemas, and embedded using \LabTirName{T-Const} (mid of \cref{fig:pgv-typing}).
+
+\todo{From the way the typing rules \LabTirName{T-Var} and \LabTirName{T-Const} are formulated, it is clear that all variables are treated as linear resources, even if they happen to have type 1 or another non-linear type.
+This could be worth mentioning in the text, saying that the type system can be easily extended to ignore non-linear resources.}
 
@@
   \end{restatablelemma}
-  \begin{proof}
-    By induction on the derivation of $\tseq[\cs{p}]{\ty{\Gamma}}{V}{T}$.
-We demonstrate the case for \LabTirName{T-Inl}, as it illustrates the need for the side conditions on \LabTirName{T-Inl} and \LabTirName{T-Inr}:
-    \begin{case*}[\LabTirName{T-Inl}]
-      The induction hypothesis gives us $\cs{p}=\cs{\pbot}$, and $\pr(\ty{\Gamma})=\pr{(\ty{T})}$. We know $\pr(\ty{T})=\pr({\ty{U}})$, hence $\pr(\ty{\Gamma})=\pr(\ty{\tysum{T}{U}})$.
-      \begin{mathpar}
-        \small
-        \inferrule*{
-          \tseq[\cs{p}]{\ty{\Gamma}}{V}{T}
-          \\
-          \pr(\ty{T})=\pr(\ty{U})
-        }{\tseq[\cs{p}]{\ty{\Gamma}}{\inl{V}}{\tysum{T}{U}}}
-      \end{mathpar}
-    \end{case*}
-    The full proof can be found in~\cref{prf:lem-pgv-value-done}.
-  \end{proof}
   \begin{restatablelemma}{lempgvsubstitution}[Substitution]
@@
   \end{restatablelemma}
-  \begin{proof}
-    By induction on the derivation of $\tseq[\cs{p}]{\ty{\Gamma},\tmty{x}{U'}}{M}{T}$.
-Whenever the contexts in the premises and conclusion differ, we use \cref{lem:pgv-value-done} to prove that the priorities are preserved.
-We demonstrate the case for \LabTirName{T-Lam}:
-    \begin{case*}[\LabTirName{T-Lam}]
-      By \cref{lem:pgv-value-done}, $\pr(\ty{\Theta})=\pr(\ty{U'})$, hence $\pr(\ty{\Gamma},\ty{\Theta})=\pr(\ty{\Gamma},\ty{U'})$.
-      \begin{mathpar}
-        \small
-        \inferrule*{
-          \tseq[\cs{q}]{\ty{\Gamma},\tmty{x}{U'},\tmty{y}{T}}{M}{U}
-        }{\tseq[\cs{\pbot}]{\ty{\Gamma},\tmty{x}{U'}}
-          {\lambda y.M}
-          {\tylolli[\cs{\pr(\ty{\Gamma},\ty{U'})},\cs{q}]{T}{U}}}
-        \substarrow{V}{x}
-        \inferrule*{
-          \tseq[\cs{q}]{\ty{\Gamma},\ty{\Theta},\tmty{y}{T}}{\subst{M}{V}{x}}{U}
-        }{\tseq[\cs{\pbot}]{\ty{\Gamma},\ty{\Theta}}
-          {\lambda y.\subst{M}{V}{x}}
-          {\tylolli[\cs{\pr(\ty{\Gamma},\ty{\Theta})},\cs{q}]{T}{U}}}
-      \end{mathpar}
-    \end{case*}
-    The complete proof can be found in~\cref{prf:lem-pgv-substitution}.
-  \end{proof}
-\end{compacttheorems}
-The proofs for subject congruence and subject reduction are standard inductions.
-The details can be found in \cref{prf:lem-pgv-subject-reduction-terms,prf:lem-pgv-subject-congruence,prf:thm-pgv-subject-reduction-confs}.
-\begin{compacttheorems}
   \begin{restatablelemma}{lempgvsubjectreductionterms}[Subject Reduction, $\tred$]
@@
 \end{compacttheorems}
+The proofs are standard inductions.
+The details can be found in \cref{prf:lem-pgv-value-done,prf:lem-pgv-substitution,prf:lem-pgv-subject-reduction-terms,prf:lem-pgv-subject-congruence,prf:thm-pgv-subject-reduction-confs}.
 
@@
   \end{definition}
+  \begin{restatablelemma}{lempgvcanonicalforms}[Canonical Forms]
+    \label{lem:pgv-canonical-forms}
+    If $\cseq[\main]{\ty{\Gamma}}{\conf{C}}$, there exists some $\tm{\conf{D}}$ such that $\tm{\conf{C}}\equiv\tm{\conf{D}}$ and $\tm{\conf{D}}$ is in canonical form.
+  \end{restatablelemma}
   \begin{restatabletheorem}{thmpgvclosedprogressconfs}[Progress, $\cred$]

### 2021-01-06 priority-gv.tex
Minor change.c

@@
 \subsubsection*{Internal and External Choice}
-\todo{You might want to clarify (repeat) that external choice can be modeled by sum type communication only because the language uses linear channels.}
-\todo{We also might want to emphasise that the modelling of choice in GV by~\cite{lindleymorris15} is unsound.}
+\todo{You might want to clarify (repeat) that external choice can be modeled by sum type communication only because the language uses linear channels.
+(We also might want to emphasise that the modelling of choice in GV by~\cite{lindleymorris15} is unsound.)}
 Typically, session-typed languages feature constructs for internal and external choice.

### 2021-01-13 introduction.tex
a pass on intro

--- 2020-10-16 introduction.tex
@@
 \section{Introduction}
-Communication and concurrency are at the basis of our modern society.
-Every service or device we interact with, from smart phones, to health monitors or any other IoT device, relies on communication.
-Hence, verification of communication in concurrent settings is important for our own safety.
 Session types are a type formalism which can be used to specify and verify communication protocols between two or more communicating agents~\cite{honda93,takeuchihonda94,hondavasconcelos98,carbonehonda07}.
-Session types have been studied for many programming paradigms, but to date the most successful settings remain the $\pi$-calculus~\cite{sangiorgiwalker01}---a process calculus for communication and concurrency, and Good Variation (GV)~\cite{wadler15,lindleymorris15}---a linear concurrent $\lambda$-calculus.
+\todo{These two settings are very different.
+The $\pi$-calculus is a mobile process algebra preceding inside which session types were developed and GV is an ad-hoc calculus made to demonstrate how propositions can be used as types.}
+Session types have been studied for many programming paradigms, including concurrent and functional paradigms.
+Most notably, they have been defined for the $\pi$-calculus~\cite{sangiorgiwalker01}---a process calculus for communication and concurrency, and for Good Variation (GV)~\cite{wadler15,lindleymorris15}---a linear concurrent $\lambda$-calculus.
+
 Session-typed concurrent $\lambda$-calculus and $\pi$-calculus satisfy communication safety and session fidelity, namely that communication proceeds without any type mismatch of the data being transmitted and follows the predefined protocol specified by a session type.
@@
 
-\emph{Can we design a more expressive GV where deadlock freedom is guaranteed by design and communication structures are not limited to trees?}
+\textbf{RQ:} \emph{Can we design a more expressive GV where deadlock freedom is guaranteed by design and communication structures are not limited to trees?}
 
@@
 In this paper we make the following contributions:
-\begin{itemize}
-\item \emph{Priority GV.} We present Priority GV (PGV) in~\cref{sec:pgv}, and prove subject reduction (\cref{thm:pgv-subject-reduction-confs}) and progress (\cref{thm:pgv-closed-progress-confs}).
-Unlike  previous work on GV, structural congruence and reduction are type preserving in PGV.
-\item \emph{Milner's Cyclic Scheduler.} To illustrate newfound expressiveness of PGV, we present Milner's Cyclic Scheduler~\cite{milner89} (\cref{sec:pgv-example}).
-\item \emph{Updated Priority CP.} We present an updated version of Priority CP~\cite{dardhagay18} in \cref{sec:pcp}.
+\begin{enumerate}
+\item
+\textbf{Priority GV}. We present Priority GV (PGV) in~\cref{sec:pgv}, and prove subject reduction (\cref{thm:pgv-subject-reduction-confs}) and progress (\cref{thm:pgv-closed-progress-confs}).
+PGV is the \emph{first} session-typed functional language with priorities \emph{and} strong ties to linear logic via its correspondence with PCP and its relation to the CP and GV languages.
+To illustrate newfound expressiveness of PGV, we present Milner's Cyclic Scheduler~\cite{milner89} (\cref{sec:pgv-example}) as well as other examples.
+
+ \item
+ \textbf{Solving GV problems}. PGV addresses several problems in the original GV language, most notably:
+ 	\begin{enumerate}
+	\item our version does not have the pseudo-type $S^\sharp$;
+	\item our structural congruence is type preserving.
+	\end{enumerate}
+\item
+\textbf{Updated Priority CP.} We present an updated version of Priority CP~\cite{dardhagay18} in \cref{sec:pcp}.
 We remove commuting conversions and move away from reduction as cut elimination, towards reduction based on structural congruence, as in process calculi.
 We present Milner's Cyclic Scheduler in PCP (\cref{sec:pcp-example}).
-\item \emph{Translation from PCP to PGV.} We present a translation from PCP to PGV in~\cref{sec:pcp} and prove that it preserves typing (\cref{thm:pcp-to-pgv-confs-preservation}) and satisfies operational correspondence (\cref{thm:pcp-to-pgv-operational-correspondence-soundness,thm:pcp-to-pgv-operational-correspondence-completeness}).
-Finally, the scheduler in PCP (\cref{sec:pcp-example}) translates to the scheduler in PGV (\cref{sec:pgv-example}).
-\end{itemize}
+
+\item
+\textbf{Connection to linear logic}.
+The connection of PGV to logic is given via a sound and complete encoding of PCP to PGV in~\cref{sec:pcp} and prove that it preserves typing (\cref{thm:pcp-to-pgv-confs-preservation}) and satisfies operational correspondence (\cref{thm:pcp-to-pgv-operational-correspondence-soundness,thm:pcp-to-pgv-operational-correspondence-completeness}).
+We showcase it with the encoding of cyclic scheduler to PGV (\cref{sec:pgv-example}).
+\end{enumerate}
 \end{document}

### 2021-01-15 introduction.tex
Handled remaining comments; removed cyclic scheduler.

@@
 This is achieved by merging constructs for parallel composition and channel restriction under the logical cut rule, which forces processes to share only one channel for communication and eliminates cyclic dependencies.
-Recent developments in the Classical Process (CP)~\cite{wadler12} line of work have led to various approaches to decoupling these constructs, either by maintaining a strong correspondence to classical linear logic, as in Hypersequent CP~\cite{kokkemontesi19popl,kokkemontesi19tlla}, or by weakening the correspondence to classical linear logic in exchange for a more expressive language, as in Priority CP (PCP)~\cite{dardhagay18}.
+Recent developments in the Classical Process (CP)~\cite{wadler14} line of work have led to various approaches to decoupling these constructs, either by maintaining a strong correspondence to classical linear logic, as in Hypersequent CP~\cite{kokkemontesi19popl,kokkemontesi19tlla}, or by weakening the correspondence to classical linear logic in exchange for a more expressive language, as in Priority CP (PCP)~\cite{dardhagay18}.
 PCP decouples CP's cut rule into two separate constructs: a construct for parallel composition, by introducing a mix rule; and a construct for restriction, by introducing a cycle rule.
@@
 In process calculi, these are conflated.
-Moreover, the benefit of working with GV over other session-typed functional languages is that GV has strong ties to linear logic, via its relation to CP~\cite{wadler12}, and consequently it has strong formal properties, \eg deadlock freedom, albeit in a restrictive way.
+Moreover, the benefit of working with GV over other session-typed functional languages is that GV has strong ties to linear logic, via its relation to CP~\cite{wadler14}, and consequently it has strong formal properties, \eg deadlock freedom, albeit in a restrictive way.
 Following the above observations, we aim to further investigate GV and deadlock freedom, and thus we pose our research question:

### 2021-01-15 priority-cp.tex
Handled remaining comments; removed cyclic scheduler.

--- 2020-10-16 priority-cp.tex
@@
 
-We present an updated version of Priority CP (PCP)~\cite{dardhagay18}, which is Wadler's CP~\cite{wadler12} with {priorities}, and a translation from PCP to PGV.
-The main difference with respect to PCP~\cite{dardhagay18} is removing commuting conversions and using structural congruence.
+We present an updated version of Priority CP (PCP)~\cite{dardhagay18}, which is Wadler's CP~\cite{wadler14} with {priorities}, and a translation from PCP to PGV.
+The main difference with respect to PCP~\cite{dardhagay18} is the removal of commuting conversions in favour of using structural congruence.
 Commuting conversions are necessary if we want our reduction strategy to correspond \emph{exactly} to cut (or cycle in \cite{dardhagay18}) elimination.
 However, as Lindley and Morris~\cite{lindleymorris15} show, all $\beta$-reductions that can be performed \emph{with} the use of commuting conversions, can also be performed \emph{without} them.
-%They show this by decomposing the reduction relation for CP into two phases: first, all $\beta$-reductions are performed, and second, commuting conversions are used to bring one particular action to the front, to eliminate the top-level cut.
 
@@
 \end{mathpar}
+
 Commuting conversions break our intuition that an action with lower priority occurs before an action with higher priority.
@@
 
+As a consequence of removing commuting conversions, we no longer need to dynamically change the priorities during reduction, which means that the intuition for priorities holds true in our updated version of PCP!
+
 \input{fig/pcp-operational-semantics-and-typing}
-
-\begin{example}[Milner's Cyclic Scheduler in PCP]\label{sec:pcp-example}
-We introduce PCP by revising Milner's cyclic scheduler~\cite{milner89,dardhagay18}.
-Our PCP scheduler has exactly the same behaviour as the PGV version in~\cref{sec:pgv-example}.
-We use $\pcp{\tm{\prod_{I}P_i}}$ to denote the parallel composition of the processes $\pcp{\tm{P_i}}$, $\tm{i}\in\tm{I}$, and $\pcp{\tm{\plug{P}{Q}}}$ to denote the plugging of $\pcp{\tm{Q}}$ in the one-hole process-context $\pcp{\tm{P}}$:
-\[
-  \usingnamespace{pcp}
-  \begin{array}{lrlr}
-    \tm{\sched}
-    & \defeq & \tm{\res{a_1}{a'_1}{\dots{\res{a_n}{a'_n}{}}}}
-               \tm{\res{b_1}{b'_1}{\dots{\res{b_n}{b'_n}{}}}}
-               \tm{\res{c_1}{c'_1}{\dots{\res{c_n}{c'_n}{}}}}
-               \tm{\res{d}{d'}{}}
-    \\ &     & \tm{(
-               \ppar{\proc1}{\agent1}
-               \parallel
-               \prod_{2\leq{i}\leq{n}}(\ppar{\proc{i}}{\wait{c'_{i-1}}{\agent{i}}})
-               )}
-    \\
-    \tm{\agent1}
-    & \defeq & \tm{\close{a_i}{\wait{b_i}{\close{c_i}{\wait{c'_n}{\close{d}{\halt}}}}}}
-    \\
-    \tm{\agent{i}}
-    & \defeq & \tm{\close{a_i}{\wait{b_i}{\close{c_i}{\halt}}}}
-    \\
-    \tm{\proc{i}}
-    & \defeq & \tm{\wait{a'_i}{\plug{P_i}{\close{b'_i}{Q_i}}}}
-  \end{array}
-\]
-\end{example}
 
@@
 Types $\pcp{\ty{\typlus[\cs{o}]{A}{B}}}$ and $\pcp{\ty{\tywith[\cs{o}]{A}{B}}}$ type the endpoints of a channel over which we can receive or send a choice between two branches $\pcp{\ty{A}}$ or $\pcp{\ty{B}}$.
-We have opted for a simplified version of choice and followed the original Wadler's CP \cite{wadler12}, however types $\ty{\oplus}$ and $\ty{\with}$ can be trivially generalised to $\pcp{\ty{\oplus^{\cs{o}}\{l_i:A_i\}_{i\in I}}}$ and $\pcp{\ty{\with^{\cs{o}}\{l_i:A_i\}_{i\in I}}}$, respectively, as in the original PCP \cite{dardhagay18}.
+We have opted for a simplified version of choice and followed the original Wadler's CP \cite{wadler14}, however types $\ty{\oplus}$ and $\ty{\with}$ can be trivially generalised to $\pcp{\ty{\oplus^{\cs{o}}\{l_i:A_i\}_{i\in I}}}$ and $\pcp{\ty{\with^{\cs{o}}\{l_i:A_i\}_{i\in I}}}$, respectively, as in the original PCP \cite{dardhagay18}.
 Types $\pcp{\ty{\tynil[\cs{o}]}}$ and $\pcp{\ty{\tytop[\cs{o}]}}$ type the endpoints of a channel over which we can send or receive a choice between \emph{no options}.
@@
 
-Recall our implementations of Milner's Cyclic Scheduler in PCP ($\pcp{\tm{\sched}}$, \cref{sec:pcp-example}) and PGV ($\pgv{\tm{\sched}}$, \cref{sec:pgv-example}).
-It follows immediately from the definition of our translation that the translation of $\pcp{\tm{\sched}}$ is exactly $\pgv{\tm{\sched}}$, where all processes are child processes.
-
 The translations $\tm{\cpgvC{\cdot}}$ and $\tm{\cpgvM{\cdot}}$ preserve typing.

### 2021-01-15 priority-gv.tex
Handled remaining comments; removed cyclic scheduler.

@@
 Priority GV offers a more fine-grained analysis of communication structures, and by separating channel creation form thread spawning it allows cyclic structures.
-We start by illustrating PGV with Milner's Cyclic Scheduler~\cite{milner89}.
-
-\todo{The $\phi$ and $\child$ symbols are not explained at this stage and become clear only later. Either warn the reader or remove them altogether.}
-
-\todo{An explanation is needed about what M and N represents and how the hole-construction models this "signal b when finished" behaviour.}
-
-\todo{It is difficult to understand how a "close d" operation is supposed to work inside a (nu d d') construct without a corresponding wait d.}
-
-\begin{example}[Milner's Cyclic Scheduler in PGV]\label{sec:pgv-example}
-Suppose a set of processes $\tm{\proc{i}}$, $1\leq{i}\leq{n}$, is scheduled to perform some tasks in cyclic order, starting with $\tm{\proc1}$, ending with $\tm{\proc{n}}$, and notifying $\tm{\proc1}$ when all processes have finished.
-We implement the scheduler as follows, using $\tm{\prod_{I}\conf{C}_i}$ to denote the parallel composition of the processes $\tm{\conf{C}_i}$, $\tm{i}\in\tm{I}$, and $\tm{\plug{M}{N}}$ to denote the plugging of $\tm{N}$ in the one-hole term-context $\tm{M}$:
-\[
-  \begin{array}{lrlr}
-    \tm{\sched}
-    & \defeq & \tm{\res{a_1}{a'_1}{\dots{\res{a_n}{a'_n}{}}}}
-               \tm{\res{b_1}{b'_1}{\dots{\res{b_n}{b'_n}{}}}}
-               \tm{\res{c_1}{c'_1}{\dots{\res{c_n}{c'_n}{}}}}
-               \tm{\res{d}{d'}{}}
-    \\ &     & \tm{\begin{array}{lll}
-                     (
-                     & \phi\;\proc1
-                     \parallel
-                     \child\;\andthen{\agent1}{\andthen{\wait\;{c'_n}}{\close\;{d}}}
-                     \\
-                     \parallel
-                     &
-                     \prod_{2\leq{i}\leq{n}}
-                     (\child\;\proc{i} \parallel \child\;\andthen{\wait\;{c'_{i-1}}}{\agent{i}})
-                     & )
-                   \end{array}}
-    \\
-    \tm{\agent{i}}
-    & \defeq & \tm{\andthen{\close\;{a_i}}{\andthen{\wait\;{b_i}}{\close\;{c_i}}}}
-    \\
-    \tm{\proc{i}}
-    & \defeq & \tm{\andthen{\wait\;{a'_i}}{\plug{M_i}{\andthen{\close\;{b'_i}}{N_i}}}}
-  \end{array}
-\]
-Our scheduler $\tm{\sched}$ consists of set of agents $\tm{\agent{i}}$, $1\leq{i}\leq{n}$, each representing their respective process.
-Each process $\tm{\proc{i}}$ waits for the signal to start their task on $\tm{a'_i}$, and signals completion on $\tm{b'_i}$.
-Each agent signals their process to start on $\tm{a_i}$, waits for their process to finish on $\tm{b_i}$, and then signals for the next agent to continue on $\tm{c_i}$.
-The agent $\tm{\agent1}$ initiates, then waits for every other process to finish, and signals $\tm{\proc1}$ on $\tm{d}$.
-Every other agent $\tm{\agent{i}}$, $2\leq{i}\leq{n}$ waits on $\tm{c'_{i-1}}$ for the signal to start.
-Each of the channels in the scheduler is of a terminated type, and is merely used to synchronise. For simplicity, we let $\tm{\sched}$ be a configuration.
-A~diagram of our scheduler with three processes can be found in \Cref{fig:pgv-scheduler}.
-\end{example}
 
@@
 Each session type $\ty{S}$ has a dual, written $\ty{\co{S}}$.
-Duality is an involutive function which preserves priorities.
+Duality is an involutive function which \emph{preserves priorities}:
 \[
@@
   & \sep      & \tm{\unit}
-    \sep        \tm{\letunit{M}{N}}
+    \sep        \tm{\andthen{M}{N}}
     \sep        \tm{\pair{M}{N}}
@@
 The $\tm{\close}$ and $\tm{\wait}$ functions close a channel.
-We use syntactic sugar to make terms more readable: we write $\tm{\andthen{M}{N}}$ in place of $\tm{\letunit{M}{N}}$, $\tm{\letbind{x}{M}{N}}$ in place of $\tm{(\lambda x.N)\;M}$, and pattern matching functions $\tm{\lambda\unit.M}$ in place of $\tm{\lambda z.\letunit{z}{M}}$ and $\tm{\lambda\pair{x}{y}.M}$ in place of $\tm{\lambda z.\letpair{x}{y}{z}{M}}$.
-\todo{Confusion arises from the difference between what is considered a native operator and a shortcut.
-For instance, "unit let" and "pair let" are native (representing sequence and new channel decomposition) while "singleton let" is a shortcut.
-The initial GV syntax was clear, this one is a bit confusing and might need some cleaning.
-At least explain why M;N cannot be a primitive construct.}
-% We recover $\tm{\fork}$ as $\tm{\lambda x.\letpair{y}{z}{\new\;\unit}{\andthen{\spawn\;{(\lambda\unit.x\;y)}}{z}}}$.
+We use syntactic sugar to make terms more readable: we write $\tm{\letbind{x}{M}{N}}$ in place of $\tm{(\lambda x.N)\;M}$, $\tm{\lambda\unit.M}$ in place of $\tm{\lambda z.\andthen{z}{M}}$, and $\tm{\lambda\pair{x}{y}.M}$ in place of $\tm{\lambda z.\letpair{x}{y}{z}{M}}$.
+We recover $\tm{\fork}$ as $\tm{\lambda x.\letpair{y}{z}{\new\;\unit}{\andthen{\spawn\;{(\lambda\unit.x\;y)}}{z}}}$.
 
 \subsubsection*{Internal and External Choice}
-\todo{You might want to clarify (repeat) that external choice can be modeled by sum type communication only because the language uses linear channels.
-(We also might want to emphasise that the modelling of choice in GV by~\cite{lindleymorris15} is unsound.)}
 Typically, session-typed languages feature constructs for internal and external choice.
@@
 \[
-\begin{array}{lcl}
-  \tm{\select{\ell}}
-  & \elabarrow & \tm{\lambda x.\letpair{y}{z}{\new}{\andthen{\close\;(\send\;{\pair{\ell\;y}{x}})}{z}}} \\
-  \multicolumn{3}{l}{\tm{\offer{L}{x}{M}{y}{N}}\elabarrow} \\
-  \multicolumn{3}{l}{
-  \qquad\tm{\letpair{z}{w}{\recv\;{L}}{\andthen{\wait\;{w}}{\casesum{z}{x}{M}{y}{N}}}}} \\
-  \tm{\offeremp{L}}
-  & \elabarrow & \tm{\letpair{z}{w}{\recv\;L}{\andthen{\wait\;{w}}{\absurd{z}}}}
+\begin{array}{l}
+  \tm{\select{\ell}}\elabarrow \\
+  \qquad\tm{\lambda x.\letpair{y}{z}{\new}{\andthen{\close\;(\send\;{\pair{\ell\;y}{x}})}{z}}} \\
+  \tm{\offer{L}{x}{M}{y}{N}}\elabarrow \\
+  \qquad\tm{\letpair{z}{w}{\recv\;{L}}{\andthen{\wait\;{w}}{\casesum{z}{x}{M}{y}{N}}}} \\
+  \tm{\offeremp{L}}\elabarrow \\
+  \qquad\tm{\letpair{z}{w}{\recv\;L}{\andthen{\wait\;{w}}{\absurd{z}}}}
 \end{array}
@@
 \begin{array}[t]{lcl}
+  \tm{\phi}
+  & \Coloneqq & \tm{\main}
+    \sep        \tm{\child}
+  \\
   \tm{\conf{C}}, \tm{\conf{D}}, \tm{\conf{E}}
@@
     \sep        \tm{\res{x}{x'}{\conf{C}}}
-  \\
-  \tm{\phi}
-  & \Coloneqq & \tm{\main}
-    \sep        \tm{\child}
 \end{array}
@@
     \sep        \tm{V\;E} \\
-  & \sep      & \tm{\letunit{E}{N}}
+  & \sep      & \tm{\andthen{E}{N}}
     \sep        \tm{\pair{E}{M}}
@@
 Consequently, we also simplify the evaluation of the $\tm{\link}$ construct.
-In GV, $\tm{\link}$ needs to return a terminated channel after evaluating the link.
-In PGV, the behaviour matches that of link in CP.\todo{Please clarify in what way it matches the behaviour of link in CP.}
+Intuitively, evaluating $\tm{\link}$ causes a substitution: if we have a channel bound as $\tm{\res{x}{y}{}}$, then $\tm{\link\;\pair{w}{x}}$ replaces all occurrences of $\tm{y}$ by $\tm{w}$.
+However, in GV, $\tm{\link}$ is required to return a terminated channel, which means that the semantics for $\tm{link}$ must \emph{create} a fresh channel of type $\ty{\tyends}/\ty{\tyendr}$.
+The endpoint of type $\ty{\tyends}$ is returned by the $\tm{link}$ construct, and a $\tm{\wait}$ on the other endpoint guards the \emph{actual} substitution.
+In PGV, evaluating $\tm{\link}$ simply causes a substitution.
+
 Secondly, our structural congruence is type preserving.
@@
 Constants $\tm{K}$ are typed using type schemas, and embedded using \LabTirName{T-Const} (mid of \cref{fig:pgv-typing}).
-
-\todo{From the way the typing rules \LabTirName{T-Var} and \LabTirName{T-Const} are formulated, it is clear that all variables are treated as linear resources, even if they happen to have type 1 or another non-linear type.
-This could be worth mentioning in the text, saying that the type system can be easily extended to ignore non-linear resources.}
+The typing rules treat \emph{all variables} as linear resources, even those of non-linear types such as $\ty{\tyunit}$.
+However, the rules can easily be extended to allow values with unrestricted usage~\cite{wadler14}.
 
@@
 The \emph{lower bound} is obtained from the typing environment.
-Terms with sequential evaluation, such as \LabTirName{T-LetUnit}, have side conditions to ensure deadlock freedom.
-For instance, the side condition for typing $\tm{\letunit{M}{N}}$ requires that $\tm{M}$ has finished communicating before $\tm{N}$ starts communicating.
+Terms with sequential evaluation, such as \LabTirName{T-Andthen}, have side conditions to ensure deadlock freedom.
+For instance, the side condition for typing $\tm{\andthen{M}{N}}$ requires that $\tm{M}$ has finished communicating before $\tm{N}$ starts communicating.
 The rule \LabTirName{T-Lam} stores the lower and upper priority bound of the function body in the type.

### 2021-01-16 priority-cp.tex
Minor rewrites; reformatted typing sugar in appendices.

@@
 \[
-  \begin{array}{l}
   \minpr(\ty{\emptyenv})          = \cs{\ptop}
-    \\
+  \quad
   \minpr(\ty{\Gamma},\tmty{x}{A}) = \minpr(\ty{\Gamma})\sqcap\minpr(\ty{A})
-  \end{array}
 \]
@@
 \]
-There are two separate translations on processes. The first, $\tm{\cpgvC{\cdot}}$, translates processes to \emph{configurations}:
-\begin{align*}
-  &\pcp{\tm{\cpgvC{\res{x}{y}{P}}}}
-  &&= \pgv{\tm{\res{x}{y}{\cpgvC{P}}}}
-  \\
-  &\pcp{\tm{\cpgvC{\ppar{P}{Q}}}}
-  &&= \pgv{\tm{\ppar{\cpgvC{P}}{\cpgvC{Q}}}}
-  \\
-  &\pcp{\tm{\cpgvC{\send{x}{y}{P}}}}
-  &&= \pgv{\tm{\res{y}{z}{(\child\;\letbind{x}{\send\;\pair{z}{x}}{\cpgvM{P}})}}}
-  \\
-  &\pcp{\tm{\cpgvC{\inl{x}{P}}}}
-  &&= \pgv{\tm{\res{y}{z}{(\child\;\letbind{x}{\andthen{\close\;(\send\;\pair{\inl{y}}{x})}{z}}{\cpgvM{P}})}}}
-  \\
-  &\pcp{\tm{\cpgvC{\inr{x}{P}}}}
-  &&= \pgv{\tm{\res{y}{z}{(\child\;\letbind{x}{\andthen{\close\;(\send\;\pair{\inr{y}}{x})}{z}}{\cpgvM{P}})}}}
-  \\
-  &\pcp{\tm{\cpgvC{P}}}
-  &&= \pgv{\tm{\child{\cpgvM{P}}}},\quad\text{if none of the above apply}
-\end{align*}
-The third, fourth, and fifth cases translate the $\nu$-binders in the translation of bound sending to configuration constructs, which is required for completeness of the operational correspondence.
-It translates as much as possible of the process to the corresponding configuration constructs.
-Once it hits the first action, it switches to the second translation, $\tm{\cpgvM{\cdot}}$, which translates processes to \emph{terms}:
+
+There are two separate translations on processes. The main translation, $\tm{\cpgvM{\cdot}}$, translates processes to \emph{terms}:
 \begin{align*}
@@
 \end{align*}
+
+Unfortunately, the operational correspondence along $\tm{\cpgvM{\cdot}}$ is incomplete, as it translates $\nu$-binders and parallel compositions to $\pgv{\tm{\new}}$ and $\pgv{\tm{\spawn}}$, which reduce back to their equivalent configuration constructs using \LabTirName{E-New} and \LabTirName{E-Spawn}.
+The same goes for $\nu$-binders which are inserted as a result of translating bound send to unbound send.
+For instance, the process $\pcp{\tm{\send{x}{y}{P}}}$ is blocked, but its translation uses $\pgv{\tm{\new}}$ and can reduce.
+To address this issue, we use a second translation, $\tm{\cpgvC{\cdot}}$, which is equivalent to $\tm{\cpgvM{\cdot}}$ followed by reduction using \LabTirName{E-New} and \LabTirName{E-Spawn}:
+\begin{align*}
+  &\pcp{\tm{\cpgvC{\res{x}{y}{P}}}}
+  &&= \pgv{\tm{\res{x}{y}{\cpgvC{P}}}}
+  \\
+  &\pcp{\tm{\cpgvC{\ppar{P}{Q}}}}
+  &&= \pgv{\tm{\ppar{\cpgvC{P}}{\cpgvC{Q}}}}
+  \\
+  &\pcp{\tm{\cpgvC{\send{x}{y}{P}}}}
+  &&= \pgv{\tm{\res{y}{z}{(\child\;\letbind{x}{\send\;\pair{z}{x}}{\cpgvM{P}})}}}
+  \\
+  &\pcp{\tm{\cpgvC{\inl{x}{P}}}}
+  &&= \pgv{\tm{\res{y}{z}{(\child\;\letbind{x}{\andthen{\close\;(\send\;\pair{\inl{y}}{x})}{z}}{\cpgvM{P}})}}}
+  \\
+  &\pcp{\tm{\cpgvC{\inr{x}{P}}}}
+  &&= \pgv{\tm{\res{y}{z}{(\child\;\letbind{x}{\andthen{\close\;(\send\;\pair{\inr{y}}{x})}{z}}{\cpgvM{P}})}}}
+  \\
+  &\pcp{\tm{\cpgvC{P}}}
+  &&= \pgv{\tm{\child{\cpgvM{P}}}},\quad\text{if none of the above apply}
+\end{align*}
+
 Typing environments are translated pointwise, and sequents $\pcp{\seq{P}{\ty{\Gamma}}}$ are translated as $\pgv{\cseq[\child]{\ty{\cpgvT{\ty{\Gamma}}}}{\cpgvC{P}}}$.
-
-The translations $\tm{\cpgvC{\cdot}}$ and $\tm{\cpgvM{\cdot}}$ preserve typing.
+The translations $\tm{\cpgvM{\cdot}}$ and $\tm{\cpgvC{\cdot}}$ preserve typing, and the latter induces a sound and complete operational correspondence.
 The proofs are standard inductions.
-The details can be found in~\cref{prf:lem-pcp-to-pgv-terms-preservation,prf:lem-pcp-to-pgv-terms-preservation}:
+The details can be found in~\cref{prf:lem-pcp-to-pgv-terms-preservation,prf:lem-pcp-to-pgv-terms-preservation,prf:lem-pcp-to-pgv-cpgvM-to-cpgvC,prf:thm-pcp-to-pgv-operational-correspondence-soundness,prf:thm-pcp-to-pgv-operational-correspondence-completeness}.
 \begin{compacttheorems}
@@
   \end{restatabletheorem}
-\end{compacttheorems}
-
-Our translation satisfies soundness and completeness of operational correspondence.
-We link ${\tm{\cpgvM{\cdot}}}$ and ${\tm{\cpgvC{\cdot}}}$ and show that $\tm{\cpgvM{P}}$ \emph{must} reduce to $\tm{\cpgvC{P}}$, \ie $\tm{\cpgvC{\cdot}}$ is equivalent to translating with $\tm{\cpgvM{\cdot}}$ and performing several reduction steps.
-We use this fact to prove the results below.
-The details are in~\cref{prf:lem-pcp-to-pgv-cpgvM-to-cpgvC,prf:thm-pcp-to-pgv-operational-correspondence-completeness,prf:thm-pcp-to-pgv-operational-correspondence-soundness}:
   \begin{restatablelemma}{lempcptopgvcpgvMtocpgvC}
@@
   \end{restatabletheorem}
-  % \begin{proof}
-  %   By induction on the derivation of $\pgv{\tm{\cpgvC{P}}\cred\tm{\conf{C}}}$.\simon{Surprising that you can get it by the derivation of the reduction relation---normally, 'soundness' would be proved by shunting into a canonical form and then proceeding by induction the structure of the term / induction on typing derivations.}
-  %   Details in \cref{prf:thm-pcp-to-pgv-operational-correspondence-soundness}.
-  % \end{proof}
   \begin{restatabletheorem}{thmpcptopgvoperationalcorrespondencecompleteness}%
@@
   \end{restatabletheorem}
-% \begin{proof}
-%   By induction on the derivation of $\pcp{\tm{P}\red\tm{Q}}$.
-We apply \cref{lem:pcp-to-pgv-cpgvM-to-cpgvC} for the cases in which the translation switches from $\tm{\cpgvC{\cdot}}$ to $\tm{\cpgvM{\cdot}}$, \ie under actions.
-We apply \cref{lem:pcp-to-pgv-confs-operational-correspondence-equiv} for \LabTirName{E-LiftSC}.
-Details in \cref{prf:thm-pcp-to-pgv-operational-correspondence-completeness}.
-% \end{proof}
+\end{compacttheorems}
 \end{document}

### 2021-01-16 appendix.tex
Processed minor comments and inserted unresolved comments as todo notes.
Rewrote PCP section.

@@
 
-\begin{restatablelemma}{lempgvcanonicalforms}[Canonical Forms]
-  \label{lem:pgv-canonical-forms}
-  If $\cseq[\main]{\ty{\Gamma}}{\conf{C}}$, there exists some $\tm{\conf{D}}$ such that $\tm{\conf{C}}\equiv\tm{\conf{D}}$ and $\tm{\conf{D}}$ is in canonical form.
-\end{restatablelemma}
+\restatelemma{lempgvcanonicalforms}
 \begin{proof}
@@
 
+\subsubsection*{Types}
+Types ($\pcp{\ty{A}}, \pcp{\ty{B}}$) in PCP are based on classical linear logic propositions, and are defined by the following grammar:
+\[
+  \usingnamespace{pcp}
+  \begin{array}{lcl}
+    \ty{A}, \ty{B}
+    & \Coloneqq & \ty{\tytens[\cs{o}]{A}{B}}
+      \sep        \ty{\typarr[\cs{o}]{A}{B}}
+      \sep        \ty{\tyone[\cs{o}]}
+      \sep        \ty{\tybot[\cs{o}]}
+      \sep        \ty{\typlus[\cs{o}]{A}{B}}
+      \sep        \ty{\tywith[\cs{o}]{A}{B}}
+      \sep        \ty{\tynil[\cs{o}]}
+      \sep        \ty{\tytop[\cs{o}]}
+  \end{array}
+\]
+Each connective is annotated with a priority $\cs{o}\in\mathbb{N}$.
+Types $\pcp{\ty{\tytens[\cs{o}]{A}{B}}}$ and $\pcp{\ty{\typarr[\cs{o}]{A}{B}}}$ type the endpoints of a channel over which we send or receive a channel of type $\pcp{\ty{A}}$, and then proceed as type $\pcp{\ty{B}}$.
+Types $\pcp{\ty{\tyone[\cs{o}]}}$ and $\pcp{\ty{{\tybot}[\cs{o}]}}$ type the endpoints of a channel whose session has terminated, and over which we send or receive a \emph{ping} before closing the channel.
+These two types act as units for $\pcp{\ty{\tytens[\cs{o}]{A}{B}}}$ and $\pcp{\ty{\typarr[\cs{o}]{A}{B}}}$, respectively.
+Types $\pcp{\ty{\typlus[\cs{o}]{A}{B}}}$ and $\pcp{\ty{\tywith[\cs{o}]{A}{B}}}$ type the endpoints of a channel over which we can receive or send a choice between two branches $\pcp{\ty{A}}$ or $\pcp{\ty{B}}$.
+We have opted for a simplified version of choice and followed the original Wadler's CP \cite{wadler14}, however types $\ty{\oplus}$ and $\ty{\with}$ can be trivially generalised to $\pcp{\ty{\oplus^{\cs{o}}\{l_i:A_i\}_{i\in I}}}$ and $\pcp{\ty{\with^{\cs{o}}\{l_i:A_i\}_{i\in I}}}$, respectively, as in the original PCP \cite{dardhagay18}.
+Types $\pcp{\ty{\tynil[\cs{o}]}}$ and $\pcp{\ty{\tytop[\cs{o}]}}$ type the endpoints of a channel over which we can send or receive a choice between \emph{no options}.
+These two types act as units for $\pcp{\ty{\typlus[\cs{o}]{A}{B}}}$ and $\pcp{\ty{\tywith[\cs{o}]{A}{B}}}$, respectively.
+
+\subsubsection*{Environments}
+\label{sec:pcp-environments}
+Typing environments $\pcp{\ty{\Gamma}}$, $\pcp{\ty{\Delta}}$ associate names to types.
+Environments are linear, so two environments can only be combined as $\pcp{\ty{\Gamma}}, \pcp{\ty{\Delta}}$ if their names are distinct, \ie $\pcp{\fv(\ty{\Gamma})\cap\fv(\ty{\Delta})=\varnothing}$.
+\[
+  \usingnamespace{pcp}
+  \begin{array}{lcl}
+    \ty{\Gamma}, \ty{\Delta}
+    & \Coloneqq & \ty{\emptyenv}
+      \sep        \ty{\Gamma}, \tmty{x}{A}
+  \end{array}
+\]
+
+\subsubsection*{Duality}
+\label{sec:pcp-duality}
+Duality is an involutive function on types which preserves priorities:
+\[
+  \usingnamespace{pcp}
+  \setlength{\arraycolsep}{1pt}
+  \begin{array}{lcl}
+    \ty{\co{(\tyone[\cs{o}])}} & = & \ty{\tybot[\cs{o}]} \\
+    \ty{\co{(\tybot[\cs{o}])}} & = & \ty{\tyone[\cs{o}]}
+  \end{array}
+  \quad
+  \begin{array}{lcl}
+    \ty{\co{(\tytens[\cs{o}]{A}{B})}} & = & \ty{\typarr[\cs{o}]{\co{A}}{\co{B}}} \\
+    \ty{\co{(\typarr[\cs{o}]{A}{B})}} & = & \ty{\tytens[\cs{o}]{\co{A}}{\co{B}}}
+  \end{array}
+  \quad
+  \begin{array}{lcl}
+    \ty{\co{(\tynil[\cs{o}])}} & = & \ty{\tytop[\cs{o}]} \\
+    \ty{\co{(\tytop[\cs{o}])}} & = & \ty{\tynil[\cs{o}]}
+  \end{array}
+  \quad
+  \begin{array}{lcl}
+    \ty{\co{(\typlus[\cs{o}]{A}{B})}} & = & \ty{\tywith[\cs{o}]{\co{A}}{\co{B}}} \\
+    \ty{\co{(\tywith[\cs{o}]{A}{B})}} & = & \ty{\typlus[\cs{o}]{\co{A}}{\co{B}}}
+  \end{array}
+\]
+
+\subsubsection*{Priorities}
+\label{sec:pcp-priorities}
+The function $\pr(\cdot)$ returns smallest priority of a type.
+As with PGV, the type system guarantees that the top-most connective always holds the smallest priority.
+The function $\minpr(\cdot)$ returns the \emph{minimum} priority of all types a typing context, or $\cs{\ptop}$ if the context is empty:
+\[
+  \usingnamespace{pcp}
+  \setlength{\arraycolsep}{1pt}
+  \begin{array}{lclclcl}
+    \pr(\ty{\tyone[\cs{o}]})        & = & \cs{o} \\
+    \pr(\ty{\tybot[\cs{o}]})        & = & \cs{o}
+  \end{array}
+  \qquad
+  \begin{array}{lclclcl}
+    \pr(\ty{\tytens[\cs{o}]{A}{B}}) & = & \cs{o} \\
+    \pr(\ty{\typarr[\cs{o}]{A}{B}}) & = & \cs{o}
+  \end{array}
+  \qquad
+  \begin{array}{lclclcl}
+    \pr(\ty{\tynil[\cs{o}]})        & = & \cs{o} \\
+    \pr(\ty{\tytop[\cs{o}]})        & = & \cs{o}
+  \end{array}
+  \qquad
+  \begin{array}{lclclcl}
+    \pr(\ty{\typlus[\cs{o}]{A}{B}}) & = & \cs{o} \\
+    \pr(\ty{\tywith[\cs{o}]{A}{B}}) & = & \cs{o} 
+  \end{array}
+\]
+\[
+  \minpr(\ty{\emptyenv})          = \cs{\ptop}
+  \quad
+  \minpr(\ty{\Gamma},\tmty{x}{A}) = \minpr(\ty{\Gamma})\sqcap\minpr(\ty{A})
+\]
+
+\subsubsection*{Terms}
+Processes ($\pcp{\tm{P}}$, $\pcp{\tm{Q}}$) in PCP are defined by the following grammar.
+\[
+  \usingnamespace{pcp}
+  \begin{array}[t]{lcl}
+    \tm{P}, \tm{Q}
+    & \Coloneqq & \tm{\link{x}{y}}
+           \sep   \tm{\res{x}{y}{P}}
+           \sep   \tm{(\ppar{P}{Q})}
+           \sep   \tm{\halt}
+    \\   & \sep & \tm{\send{x}{y}{P}}
+           \sep   \tm{\close{x}{P}}
+           \sep   \tm{\recv{x}{y}{P}}
+           \sep   \tm{\wait{x}{P}}
+    \\   & \sep & \tm{\inl{x}{P}}
+           \sep   \tm{\inr{x}{P}}
+           \sep   \tm{\offer{x}{P}{Q}}
+           \sep   \tm{\absurd{x}}
+  \end{array}
+\]
+Process $\pcp{\tm{\link{x}{y}}}$ links endpoints $\pcp{\tm{x}}$ and $\pcp{\tm{y}}$ and forwards communication from one to the other.
+$\pcp{\tm{\res{x}{y}{P}}}$, $\pcp{\tm{(\ppar{P}{Q})}}$ and $\pcp{\tm{\halt}}$ denote respectively the restriction processes where channel endpoints $\pcp{\tm{x}}$ and $\pcp{\tm{y}}$ are bound together and with scope $\pcp{\tm{P}}$, the parallel composition of processes $\pcp{\tm{P}}$ and $\pcp{\tm{Q}}$ and the terminated process.
+Processes $\pcp{\tm{\send{x}{y}{P}}}$ and $\pcp{\tm{\recv{x}{y}{P}}}$ send or receive over channel $\pcp{\tm{x}}$ a value $\pcp{\tm{y}}$ and proceed as process $\pcp{\tm{P}}$.
+Processes $\pcp{\tm{\close{x}{P}}}$ and $\pcp{\tm{\wait{x}{P}}}$ send and receive an empty value---denoting the closure of channel $\pcp{\tm{x}}$, and continue as $\pcp{\tm{P}}$.
+Processes $\pcp{\tm{\inl{x}{P}}}$ and $\pcp{\tm{\inr{x}{P}}}$ make a left and right choice, respectively and proceed as process $\pcp{\tm{P}}$.
+Dually, $\pcp{\tm{\offer{x}{P}{Q}}}$ offers both left and right branches, with continuations $\pcp{\tm{P}}$ and $\pcp{\tm{Q}}$, and $\pcp{\tm{\absurd{x}}}$ is the empty offer.
+We write \emph{unbound} send as $\pcp{\tm{\usend{x}{y}{P}}}$, which is syntactic sugar for $\pcp{\tm{\send{x}{z}{(\ppar{\link{y}{z}}{P})}}}$.
+Alternatively, we could take $\pcp{\tm{\usend{x}{y}{P}}}$ as primitive, and let $\pcp{\tm{\send{x}{y}{P}}}$ be syntactic sugar for $\pcp{\tm{\res{y}{z}{(\usend{x}{z}{P})}}}$.
+CP takes \emph{bound} sending as primitive, as it is impossible to eliminate the top-level cut in terms such as $\pcp{\tm{\res{y}{z}{(\usend{x}{z}{P})}}}$, even with commuting conversions.
+In our setting without commuting conversions and with more permissive normal forms, this is no longer an issue, but, for simplicity, we keep bound sending as primitive.
+
+\input{fig/pcp-operational-semantics-and-typing}
+
+\subsubsection*{Operational Semantics}
+The operational semantics for PCP, given in \cref{fig:pcp-operational-semantics-and-typing}, is defined as a reduction relation $\pcp{\red}$ on processes (bottom) and uses structural congruence (top).
+Each of the axioms of structural congruence corresponds to the axiom of the same name for PGV.
+We write $\pcp{\red^+}$ for the transitive closures, and $\pcp{\red^\star}$ for the reflexive-transitive closures.
+
+The reduction relation is given by a set of axioms and inference rules for context closure.
+Reduction occurs under restriction.
+$\LabTirName{E-Link}$ reduces a parallel composition with a link into a substitution.
+$\LabTirName{E-Send}$ is the main communication rule, where send and receive processes sychronise and reduce to the corresponding continuations.
+$\LabTirName{E-Close}$ follows the previous rule and it closes the channel identified by endpoints $\pcp{\tm{x}}$ and $\pcp{\tm{y}}$.
+$\LabTirName{E-Select-Inl}$ and $\LabTirName{E-Select-Inr}$ are generalised versions of $\LabTirName{E-Send}$.
+They state respectively that a left and right selection synchronises with a choice offering and reduces to the corresponding continuations.
+The last three rules state that reduction is closed under restriction, parallel composition and structural congruence, respectively.
+
+\subsubsection*{Typing}
+\Cref{fig:pcp-operational-semantics-and-typing} gives the typing rules for our version of PCP.
+A typing judgement $\pcp{\seq{P}{\Gamma}}$ states that ``process $\pcp{\tm{P}}$ is well typed under the typing context $\pcp{\ty{\Gamma}}$''.
+
+\textsc{T-Link} states that the link process $\pcp{\tm{\link{x}{y}}}$ is well typed under channels $\pcp{\tm{x}}$ and $\pcp{\tm{y}}$ having dual types, respectively $\pcp{\ty{A}}$ and $\pcp{\ty{\co{A}}}$.
+\textsc{T-Res} sates that the restriction process $\pcp{\tm{\res{x}{y}{P}}}$ is well typed under typing context $\pcp{\ty{\Gamma}}$ if process $\pcp{\tm{P}}$ is well typed in $\pcp{\ty{\Gamma}}$ augmented with channel endpoints $\pcp{\tm{x}}$ and $\pcp{\tm{y}}$ having dual types, respectively $\pcp{\ty{A}}$ and $\pcp{\ty{\co{A}}}$.
+\textsc{T-Par} states that the parallel composition of processes $\pcp{\tm{P}}$ and $\pcp{\tm{Q}}$ is well typed under the disjoint union of their respective typing contexts.
+\textsc{T-Halt} states that the terminated process $\pcp{\tm{\halt}}$ is well typed in the empty context.
+\textsc{T-Send} and \textsc{T-Recv} state that the sending and receiving of a bound name $\pcp{\tm{y}}$ over a channel $\pcp{\tm{x}}$ is well typed under $\pcp{\ty{\Gamma}}$ and $\pcp{\tm{x}}$ of type $\pcp{\ty{\tytens[\cs{o}]{A}{B}}}$, respectively $\pcp{\typarr[\cs{o}]{A}{B}}$.
+Priority $\cs{o}$ is the smallest among all priorities of the types used by the output or input process, captured by the side condition $\pcp{\cs{o}<\minpr(\ty{\Gamma},\ty{A},\ty{B})}$.
+Rules \textsc{T-Close} and \textsc{T-Wait} type the closure of channel $\pcp{\tm{x}}$ and are in the same lines as the previous two rules, requiring that the priority of channel $\pcp{\tm{x}}$ is the smallest among all priorities in $\pcp{\ty{\Gamma}}$.
+\textsc{T-Select-Inl} and \textsc{T-Select-Inr} type respectively the left $\pcp{\tm{\inl{x}{P}}}$ and right $\pcp{\tm{\inr{x}{P}}}$ choice performed on channel $\pcp{\tm{x}}$.
+\textsc{T-Offer} and \textsc{T-Offer-Absurd} type the offering of a choice, or empty choice, on channel $\pcp{\tm{x}}$.
+In all the above rules the priority $\cs{o}$ of channel $\pcp{\tm{x}}$ is the smallest with respect to the typing context $\pcp{\cs{o}<\minpr(\ty{\Gamma})}$ and types involved in the choice $\pcp{\cs{o}<\minpr(\ty{\Gamma},\ty{A},\ty{B})}$.
+
+Finally, since our reduction relation is a strict subset of the reduction relation in the original~\cite{dardhagay18}, we defer to their proofs.
+We prove progress for our version of PCP, see~\cref{prf:thm-pcp-closed-progress}.
+
 % Proofs that the syntactic sugar is well-typed.
@@
 
-\begin{definition}[Actions]
-  A~process acts on an endpoint $\tm{x}$ if it is in one of the following forms:
-  \begin{multicols}{3}
-    \begin{itemize}[noitemsep,topsep=0pt,parsep=0pt,partopsep=0pt]
-    \item $\tm{\link{x}{y}}$ 
-    \item $\tm{\link{y}{x}}$
-    \item $\tm{\send{x}{y}{P}}$
-    \item $\tm{\recv{x}{y}{P}}$
-    \item $\tm{\close{x}{P}}$
-    \item $\tm{\wait{x}{P}}$
-    \item $\tm{\inl{x}{P}}$
-    \item $\tm{\inr{x}{P}}$
-    \item $\tm{\offer{x}{P}{Q}}$
-    \item $\tm{\absurd{x}}$
-    \end{itemize}
-  \end{multicols}
-  \noindent
+\begin{compacttheorems}
+    \begin{restatabledefinition}{defpcpactions}[Actions]
+    A~process acts on an endpoint $\tm{x}$ if it is of the form $\tm{\link{x}{y}}$, $\tm{\link{y}{x}}$, $\tm{\send{x}{y}{P}}$, $\tm{\recv{x}{y}{P}}$, $\tm{\close{x}{P}}$, $\tm{\wait{x}{P}}$, $\tm{\inl{x}{P}}$, $\tm{\inr{x}{P}}$, $\tm{\offer{x}{P}{Q}}$, or $\tm{\absurd{x}}$.
 A~process is an action if it acts on some endpoint $\tm{x}$.
-\end{definition}
-
-\begin{definition}[Canonical Forms]
+  \end{restatabledefinition}
+  \begin{restatabledefinition}{defpcpcanonicalforms}[Canonical Forms]
     \label{def:pcp-canonical-forms}
-  A~process $\tm{P}$ is in canonical form if it is in one of the following forms:
-  \begin{itemize}[noitemsep,topsep=0pt,parsep=0pt,partopsep=0pt]
-  \item
-    $\tm{\halt}$
-  \item
-    $\tm{\res{x_1}{x'_1}{\dots\res{x_n}{x'_n}{(P_1 \parallel\dots\parallel P_m)}}}$
-    where $m>0$ and each $\tm{P_j}$ is an action.
-  \end{itemize}
-\end{definition}
-
+    A~process $\tm{P}$ is in canonical form if it is of the form $\tm{\halt}$ or $\tm{\res{x_1}{x'_1}{\dots\res{x_n}{x'_n}{(P_1 \parallel\dots\parallel P_m)}}}$ where $m>0$ and each $\tm{P_j}$ is an action.
+  \end{restatabledefinition}
   \begin{restatablelemma}{lempcpcanonicalforms}[Canonical Forms]
@@
   \end{proof}
-
-\begin{restatabletheorem}{thmpcpclosedprogress}[Progress, $\pcp{\red}$]
-  \label{thm:pcp-closed-progress}
-  \hfill\\%newline before theorem statement
-  If $\pcp{\seq{P}{\emptyenv}}$, then either $\pcp{\tm{P}=\tm{\halt}}$ or there exists a $\pcp{\tm{Q}}$ such that $\pcp{\tm{P}\red\tm{Q}}$.
-\end{restatabletheorem}
+  \restatetheorem{thmpcpclosedprogress}
   \input{prf/thm-pcp-closed-progress}
+\end{compacttheorems}
 \endgroup
 
+\subsubsection*{Translation}
 \begingroup

### 2021-01-16 priority-cp.tex
Rewrote PCP section.

@@
 
-We present an updated version of Priority CP (PCP)~\cite{dardhagay18}, which is Wadler's CP~\cite{wadler14} with {priorities}, and a translation from PCP to PGV.
-The main difference with respect to PCP~\cite{dardhagay18} is the removal of commuting conversions in favour of using structural congruence.
-Commuting conversions are necessary if we want our reduction strategy to correspond \emph{exactly} to cut (or cycle in \cite{dardhagay18}) elimination.
-However, as Lindley and Morris~\cite{lindleymorris15} show, all $\beta$-reductions that can be performed \emph{with} the use of commuting conversions, can also be performed \emph{without} them.
+We present a correspondence between Priority GV and an updated version of Priority CP~\cite[PCP]{dardhagay18}, which is Wadler's CP~\cite{wadler14} with priorities.
+The correspondence connects PGV to (a~relaxed variant of) classical linear logic.
 
-From the perspective of process calculi, commuting conversions behave strangely. Let us consider the commuting conversion $(\kappa_{\parr})$ for $\pcp{\tm{\recv{x}{y}{P}}}$:
+\subsubsection*{Revisiting Priority CP}
+\begingroup
+\usingnamespace{pcp}
+Types ($\ty{A}$, $\ty{B}$) in PCP correspond to the connectives of linear logic annotated with priorities $\cs{o}\in\mathbb{N}$.
+Typing environments, duality, and the priority function $\pr$ are defined as expected (see~\cref{sec:pcp-environments,sec:pcp-duality,sec:pcp-priorities}).
 \[
-  \usingnamespace{pcp}
-  (\kappa_{\parr})
-  \quad
-  \tm{\res{z}{z'}{(\ppar{\recv{x}{y}{P}}{Q})}}
-  \red
-  \tm{\recv{x}{y}{\res{z}{z'}{(\ppar{P}{Q})}}}
-\]
-As a result of $(\kappa_{\parr})$, $\pcp{\tm{Q}}$ becomes blocked on $\pcp{\tm{\labrecv{x}{y}}}$, and any actions $\pcp{\tm{Q}}$ was able to perform become unavailable.
-Consequently, CP is non-confluent.
-For instance,
-\begin{mathpar}
-  \usingnamespace{pcp}
-  \setlength{\arraycolsep}{2em}
-  \begin{array}{cc}
-    \multicolumn{2}{c}{%
-    \hspace*{10ex}
-    {\tm{\res{x}{x'}{(\ppar{\recv{a}{y}{P}}{\res{z}{z'}{(\ppar{\close{z}{\halt}}{\wait{z'}{Q}})}})}}}}
-    \\
-    \qquad\rotatebox[origin=c]{215}{$\red$}
-    &
-    \rotatebox[origin=c]{325}{$\red^+$}\qquad
-    \\
-    {\tm{\recv{a}{y}{\res{x}{x'}{(\ppar{P}{\res{z}{z'}{(\ppar{\close{z}{\halt}}{\wait{z'}{Q}})}})}}}}
-    &
-    {\tm{\recv{a}{y}{\res{x}{x'}{(\ppar{P}{Q})}}}}
-  \end{array}
-\end{mathpar}
-
-Commuting conversions break our intuition that an action with lower priority occurs before an action with higher priority.
-To cite Dardha and Gay~\cite{dardhagay18} ``\emph{if a prefix on a channel endpoint $\pcp{\tm{x}}$ with priority $\cs{o}$ is pulled out at top level, then to preserve priority constraints in the typing rules [..], it is necessary to increase priorities of all actions after the prefix on $\pcp{\tm{x}}$}'' by $\cs{o+1}$.
-
-As a consequence of removing commuting conversions, we no longer need to dynamically change the priorities during reduction, which means that the intuition for priorities holds true in our updated version of PCP!
-
-\input{fig/pcp-operational-semantics-and-typing}
-
-\subsubsection*{Types}
-Types ($\pcp{\ty{A}}, \pcp{\ty{B}}$) in PCP are based on classical linear logic propositions, and are defined by the following grammar:
-\[
-  \usingnamespace{pcp}
   \begin{array}{lcl}
@@
 \]
-Each connective is annotated with a priority $\cs{o}\in\mathbb{N}$.
-Types $\pcp{\ty{\tytens[\cs{o}]{A}{B}}}$ and $\pcp{\ty{\typarr[\cs{o}]{A}{B}}}$ type the endpoints of a channel over which we send or receive a channel of type $\pcp{\ty{A}}$, and then proceed as type $\pcp{\ty{B}}$.
-Types $\pcp{\ty{\tyone[\cs{o}]}}$ and $\pcp{\ty{{\tybot}[\cs{o}]}}$ type the endpoints of a channel whose session has terminated, and over which we send or receive a \emph{ping} before closing the channel.
-These two types act as units for $\pcp{\ty{\tytens[\cs{o}]{A}{B}}}$ and $\pcp{\ty{\typarr[\cs{o}]{A}{B}}}$, respectively.
-Types $\pcp{\ty{\typlus[\cs{o}]{A}{B}}}$ and $\pcp{\ty{\tywith[\cs{o}]{A}{B}}}$ type the endpoints of a channel over which we can receive or send a choice between two branches $\pcp{\ty{A}}$ or $\pcp{\ty{B}}$.
-We have opted for a simplified version of choice and followed the original Wadler's CP \cite{wadler14}, however types $\ty{\oplus}$ and $\ty{\with}$ can be trivially generalised to $\pcp{\ty{\oplus^{\cs{o}}\{l_i:A_i\}_{i\in I}}}$ and $\pcp{\ty{\with^{\cs{o}}\{l_i:A_i\}_{i\in I}}}$, respectively, as in the original PCP \cite{dardhagay18}.
-Types $\pcp{\ty{\tynil[\cs{o}]}}$ and $\pcp{\ty{\tytop[\cs{o}]}}$ type the endpoints of a channel over which we can send or receive a choice between \emph{no options}.
-These two types act as units for $\pcp{\ty{\typlus[\cs{o}]{A}{B}}}$ and $\pcp{\ty{\tywith[\cs{o}]{A}{B}}}$, respectively.
-
-\subsubsection*{Environments}
-Typing environments $\pcp{\ty{\Gamma}}$, $\pcp{\ty{\Delta}}$ associate names to types.
-Environments are linear, so two environments can only be combined as $\pcp{\ty{\Gamma}}, \pcp{\ty{\Delta}}$ if their names are distinct, \ie $\pcp{\fv(\ty{\Gamma})\cap\fv(\ty{\Delta})=\varnothing}$.
+Processes ($\tm{P}$, $\tm{Q}$) in PCP are defined by the following grammar.
 \[
-  \usingnamespace{pcp}
-  \begin{array}{lcl}
-    \ty{\Gamma}, \ty{\Delta}
-    & \Coloneqq & \ty{\emptyenv}
-      \sep        \ty{\Gamma}, \tmty{x}{A}
-  \end{array}
-\]
-
-\subsubsection*{Duality}
-Duality is an involutive function on types which preserves priorities:
-\[
-  \usingnamespace{pcp}
-  \setlength{\arraycolsep}{1pt}
-  \begin{array}{lcl}
-    \ty{\co{(\tyone[\cs{o}])}} & = & \ty{\tybot[\cs{o}]} \\
-    \ty{\co{(\tybot[\cs{o}])}} & = & \ty{\tyone[\cs{o}]}
-  \end{array}
-  \quad
-  \begin{array}{lcl}
-    \ty{\co{(\tytens[\cs{o}]{A}{B})}} & = & \ty{\typarr[\cs{o}]{\co{A}}{\co{B}}} \\
-    \ty{\co{(\typarr[\cs{o}]{A}{B})}} & = & \ty{\tytens[\cs{o}]{\co{A}}{\co{B}}}
-  \end{array}
-  \quad
-  \begin{array}{lcl}
-    \ty{\co{(\tynil[\cs{o}])}} & = & \ty{\tytop[\cs{o}]} \\
-    \ty{\co{(\tytop[\cs{o}])}} & = & \ty{\tynil[\cs{o}]}
-  \end{array}
-  \quad
-  \begin{array}{lcl}
-    \ty{\co{(\typlus[\cs{o}]{A}{B})}} & = & \ty{\tywith[\cs{o}]{\co{A}}{\co{B}}} \\
-    \ty{\co{(\tywith[\cs{o}]{A}{B})}} & = & \ty{\typlus[\cs{o}]{\co{A}}{\co{B}}}
-  \end{array}
-\]
-
-\subsubsection*{Priorities}
-The function $\pr(\cdot)$ returns smallest priority of a type.
-As with PGV, the type system guarantees that the top-most connective always holds the smallest priority.
-The function $\minpr(\cdot)$ returns the \emph{minimum} priority of all types a typing context, or $\cs{\ptop}$ if the context is empty:
-\[
-  \usingnamespace{pcp}
-  \setlength{\arraycolsep}{1pt}
-  \begin{array}{lclclcl}
-    \pr(\ty{\tyone[\cs{o}]})        & = & \cs{o} \\
-    \pr(\ty{\tybot[\cs{o}]})        & = & \cs{o}
-  \end{array}
-  \qquad
-  \begin{array}{lclclcl}
-    \pr(\ty{\tytens[\cs{o}]{A}{B}}) & = & \cs{o} \\
-    \pr(\ty{\typarr[\cs{o}]{A}{B}}) & = & \cs{o}
-  \end{array}
-  \qquad
-  \begin{array}{lclclcl}
-    \pr(\ty{\tynil[\cs{o}]})        & = & \cs{o} \\
-    \pr(\ty{\tytop[\cs{o}]})        & = & \cs{o}
-  \end{array}
-  \qquad
-  \begin{array}{lclclcl}
-    \pr(\ty{\typlus[\cs{o}]{A}{B}}) & = & \cs{o} \\
-    \pr(\ty{\tywith[\cs{o}]{A}{B}}) & = & \cs{o} 
-  \end{array}
-\]
-\[
-  \minpr(\ty{\emptyenv})          = \cs{\ptop}
-  \quad
-  \minpr(\ty{\Gamma},\tmty{x}{A}) = \minpr(\ty{\Gamma})\sqcap\minpr(\ty{A})
-\]
-
-\subsubsection*{Terms}
-Processes ($\pcp{\tm{P}}$, $\pcp{\tm{Q}}$) in PCP are defined by the following grammar.
-\[
-  \usingnamespace{pcp}
   \begin{array}[t]{lcl}
@@
 \]
-Process $\pcp{\tm{\link{x}{y}}}$ links endpoints $\pcp{\tm{x}}$ and $\pcp{\tm{y}}$ and forwards communication from one to the other.
-$\pcp{\tm{\res{x}{y}{P}}}$, $\pcp{\tm{(\ppar{P}{Q})}}$ and $\pcp{\tm{\halt}}$ denote respectively the restriction processes where channel endpoints $\pcp{\tm{x}}$ and $\pcp{\tm{y}}$ are bound together and with scope $\pcp{\tm{P}}$, the parallel composition of processes $\pcp{\tm{P}}$ and $\pcp{\tm{Q}}$ and the terminated process.
-Processes $\pcp{\tm{\send{x}{y}{P}}}$ and $\pcp{\tm{\recv{x}{y}{P}}}$ send or receive over channel $\pcp{\tm{x}}$ a value $\pcp{\tm{y}}$ and proceed as process $\pcp{\tm{P}}$.
-Processes $\pcp{\tm{\close{x}{P}}}$ and $\pcp{\tm{\wait{x}{P}}}$ send and receive an empty value---denoting the closure of channel $\pcp{\tm{x}}$, and continue as $\pcp{\tm{P}}$.
-Processes $\pcp{\tm{\inl{x}{P}}}$ and $\pcp{\tm{\inr{x}{P}}}$ make a left and right choice, respectively and proceed as process $\pcp{\tm{P}}$.
-Dually, $\pcp{\tm{\offer{x}{P}{Q}}}$ offers both left and right branches, with continuations $\pcp{\tm{P}}$ and $\pcp{\tm{Q}}$, and $\pcp{\tm{\absurd{x}}}$ is the empty offer.
-We write \emph{unbound} send as $\pcp{\tm{\usend{x}{y}{P}}}$, which is syntactic sugar for $\pcp{\tm{\send{x}{z}{(\ppar{\link{y}{z}}{P})}}}$.
-Alternatively, we could take $\pcp{\tm{\usend{x}{y}{P}}}$ as primitive, and let $\pcp{\tm{\send{x}{y}{P}}}$ be syntactic sugar for $\pcp{\tm{\res{y}{z}{(\usend{x}{z}{P})}}}$.
-CP takes \emph{bound} sending as primitive, as it is impossible to eliminate the top-level cut in terms such as $\pcp{\tm{\res{y}{z}{(\usend{x}{z}{P})}}}$, even with commuting conversions.
-In our setting without commuting conversions and with more permissive normal forms, this is no longer an issue, but, for simplicity, we keep bound sending as primitive.
+Processes are typed by sequents $\seq{P}{\ty{\Gamma}}$, which correspond to the one-sided sequents classical linear logic.
+Different from PGV, we do not need to store the greatest priority on the sequent, as, due to the absence of higher-order functions, we cannot compose processes \emph{sequentially}.
 
-\subsubsection*{Operational Semantics}
-The operational semantics for PCP, given in \cref{fig:pcp-operational-semantics-and-typing}, is defined as a reduction relation $\pcp{\red}$ on processes (bottom) and uses structural congruence (top).
-Each of the axioms of structural congruence corresponds to the axiom of the same name for PGV.
-We write $\pcp{\red^+}$ for the transitive closures, and $\pcp{\red^\star}$ for the reflexive-transitive closures.
+PCP decomposes cut into \LabTirName{T-Res} and \LabTirName{T-Par}, and guarantees deadlock freedom using priority constraints, \eg, in \LabTirName{T-Send}.
+The full typing rules and operational semantics are available in \cref{fig:pcp-operational-semantics-and-typing} and in \cite{dardhagay18}:
+\begin{mathpar}
+  \inferrule*[lab=T-Res]{
+    \seq{P}{\ty{\Gamma},\tmty{x}{A},\tmty{y}{\co{A}}}
+  }{\seq{\res{x}{y}{P}}{\ty{\Gamma}}}
 
-The reduction relation is given by a set of axioms and inference rules for context closure.
-Reduction occurs under restriction.
-$\LabTirName{E-Link}$ reduces a parallel composition with a link into a substitution.
-$\LabTirName{E-Send}$ is the main communication rule, where send and receive processes sychronise and reduce to the corresponding continuations.
-$\LabTirName{E-Close}$ follows the previous rule and it closes the channel identified by endpoints $\pcp{\tm{x}}$ and $\pcp{\tm{y}}$.
-$\LabTirName{E-Select-Inl}$ and $\LabTirName{E-Select-Inr}$ are generalised versions of $\LabTirName{E-Send}$.
-They state respectively that a left and right selection synchronises with a choice offering and reduces to the corresponding continuations.
-The last three rules state that reduction is closed under restriction, parallel composition and structural congruence, respectively.
+  \inferrule*[lab=T-Par]{
+    \seq{P}{\ty{\Gamma}}
+    \\
+    \seq{Q}{\ty{\Delta}}
+  }{\seq{\ppar{P}{Q}}{\ty{\Gamma},\ty{\Delta}}}
 
-\subsubsection*{Typing}
-\Cref{fig:pcp-operational-semantics-and-typing} gives the typing rules for our version of PCP.
-A typing judgement $\pcp{\seq{P}{\Gamma}}$ states that ``process $\pcp{\tm{P}}$ is well typed under the typing context $\pcp{\ty{\Gamma}}$''.
+  \inferrule*[lab=T-Send]{
+    \seq{P}{\ty{\Gamma},\tmty{y}{A},\tmty{x}{B}}
+    \\
+    \cs{o}<\minpr(\ty{\Gamma},\ty{A},\ty{B})
+  }{\seq{\send{x}{y}{P}}{\ty{\Gamma},\tmty{x}{\tytens[\cs{o}]{A}{B}}}}
+\end{mathpar}
 
-\textsc{T-Link} states that the link process $\pcp{\tm{\link{x}{y}}}$ is well typed under channels $\pcp{\tm{x}}$ and $\pcp{\tm{y}}$ having dual types, respectively $\pcp{\ty{A}}$ and $\pcp{\ty{\co{A}}}$.
-\textsc{T-Res} sates that the restriction process $\pcp{\tm{\res{x}{y}{P}}}$ is well typed under typing context $\pcp{\ty{\Gamma}}$ if process $\pcp{\tm{P}}$ is well typed in $\pcp{\ty{\Gamma}}$ augmented with channel endpoints $\pcp{\tm{x}}$ and $\pcp{\tm{y}}$ having dual types, respectively $\pcp{\ty{A}}$ and $\pcp{\ty{\co{A}}}$.
-\textsc{T-Par} states that the parallel composition of processes $\pcp{\tm{P}}$ and $\pcp{\tm{Q}}$ is well typed under the disjoint union of their respective typing contexts.
-\textsc{T-Halt} states that the terminated process $\pcp{\tm{\halt}}$ is well typed in the empty context.
-\textsc{T-Send} and \textsc{T-Recv} state that the sending and receiving of a bound name $\pcp{\tm{y}}$ over a channel $\pcp{\tm{x}}$ is well typed under $\pcp{\ty{\Gamma}}$ and $\pcp{\tm{x}}$ of type $\pcp{\ty{\tytens[\cs{o}]{A}{B}}}$, respectively $\pcp{\typarr[\cs{o}]{A}{B}}$.
-Priority $\cs{o}$ is the smallest among all priorities of the types used by the output or input process, captured by the side condition $\pcp{\cs{o}<\minpr(\ty{\Gamma},\ty{A},\ty{B})}$.
-Rules \textsc{T-Close} and \textsc{T-Wait} type the closure of channel $\pcp{\tm{x}}$ and are in the same lines as the previous two rules, requiring that the priority of channel $\pcp{\tm{x}}$ is the smallest among all priorities in $\pcp{\ty{\Gamma}}$.
-\textsc{T-Select-Inl} and \textsc{T-Select-Inr} type respectively the left $\pcp{\tm{\inl{x}{P}}}$ and right $\pcp{\tm{\inr{x}{P}}}$ choice performed on channel $\pcp{\tm{x}}$.
-\textsc{T-Offer} and \textsc{T-Offer-Absurd} type the offering of a choice, or empty choice, on channel $\pcp{\tm{x}}$.
-In all the above rules the priority $\cs{o}$ of channel $\pcp{\tm{x}}$ is the smallest with respect to the typing context $\pcp{\cs{o}<\minpr(\ty{\Gamma})}$ and types involved in the choice $\pcp{\cs{o}<\minpr(\ty{\Gamma},\ty{A},\ty{B})}$.
+The main change we make to PCP is \emph{removing commuting conversions}.
+Commuting conversions are necessary if we want our reduction strategy to correspond \emph{exactly} to cut (or cycle in \cite{dardhagay18}) elimination.
+However, as Lindley and Morris~\cite{lindleymorris15} show, all communications that can be performed \emph{with} the use of commuting conversions, can also be performed \emph{without} them, using structural congruence.
 
-Finally, since our reduction relation is a strict subset of the reduction relation in the original~\cite{dardhagay18}, we defer to their proofs.
-We prove progress for our version of PCP, see~\cref{prf:thm-pcp-closed-progress}.
+From the perspective of process calculi, commuting conversions behave strangely. Consider the commuting conversion $(\kappa_{\parr})$ for $\pcp{\tm{\recv{x}{y}{P}}}$:
+\begin{mathpar}
+  (\kappa_{\parr})
+  \quad
+  \tm{\res{z}{z'}{(\ppar{\recv{x}{y}{P}}{Q})}}
+  \red
+  \tm{\recv{x}{y}{\res{z}{z'}{(\ppar{P}{Q})}}}
+\end{mathpar}
+As a result of $(\kappa_{\parr})$, $\pcp{\tm{Q}}$ becomes blocked on $\pcp{\tm{\labrecv{x}{y}}}$, and any actions $\pcp{\tm{Q}}$ was able to perform become unavailable.
+Consequently, CP is non-confluent:
+\begin{mathpar}
+  \setlength{\arraycolsep}{2em}
+  \begin{array}{cc}
+    \multicolumn{2}{c}{%
+    \hspace*{10ex}
+    {\tm{\res{x}{x'}{(\ppar{\recv{a}{y}{P}}{\res{z}{z'}{(\ppar{\close{z}{\halt}}{\wait{z'}{Q}})}})}}}}
+    \\
+    \qquad\rotatebox[origin=c]{270}{$\red\hphantom{{}^+}$}
+    &
+    \rotatebox[origin=c]{270}{$\red^+$}\qquad
+    \\
+    {\tm{\recv{a}{y}{\res{x}{x'}{(\ppar{P}{\res{z}{z'}{(\ppar{\close{z}{\halt}}{\wait{z'}{Q}})}})}}}}
+    &
+    {\tm{\recv{a}{y}{\res{x}{x'}{(\ppar{P}{Q})}}}}
+  \end{array}
+\end{mathpar}
+
+In PCP, commuting conversions break our intuition that an action with lower priority occurs before an action with higher priority.
+To cite Dardha and Gay~\cite{dardhagay18} ``\emph{if a prefix on a channel endpoint $\pcp{\tm{x}}$ with priority $\cs{o}$ is pulled out at top level, then to preserve priority constraints in the typing rules [..], it is necessary to increase priorities of all actions after the prefix on $\pcp{\tm{x}}$}'' by $\cs{o+1}$.
+
+One benefit of removing commuting conversions is that we no longer need to dynamically change the priorities during reduction, which means that the intuition for priorities holds true in our updated version of PCP.
+Furthermore, we can safely define reduction on untyped processes, which means that type and priority information is erasable!
+
+We prove closed progress for our updated PCP. The proof is in \cref{prf:thm-pcp-closed-progress}.
+\begin{compacttheorems}
+  \begin{restatabletheorem}{thmpcpclosedprogress}[Progress, $\pcp{\red}$]
+    \label{thm:pcp-closed-progress}
+    \hfill\\%newline before theorem statement
+    If $\pcp{\seq{P}{\emptyenv}}$, then either $\pcp{\tm{P}=\tm{\halt}}$ or there exists a $\pcp{\tm{Q}}$ such that $\pcp{\tm{P}\red\tm{Q}}$.
+  \end{restatabletheorem}
+\end{compacttheorems}
+\endgroup
 
 \subsubsection*{Translation}
+\begingroup
 We illustrate the relation between PCP and PGV by giving a translation from PCP to PGV. The translation on types is defined as follows:
@@
 
-Unfortunately, the operational correspondence along $\tm{\cpgvM{\cdot}}$ is incomplete, as it translates $\nu$-binders and parallel compositions to $\pgv{\tm{\new}}$ and $\pgv{\tm{\spawn}}$, which reduce back to their equivalent configuration constructs using \LabTirName{E-New} and \LabTirName{E-Spawn}.
+Unfortunately, the operational correspondence along $\tm{\cpgvM{\cdot}}$ is unsound, as it translates $\nu$-binders and parallel compositions to $\pgv{\tm{\new}}$ and $\pgv{\tm{\spawn}}$, which reduce back to their equivalent configuration constructs using \LabTirName{E-New} and \LabTirName{E-Spawn}.
 The same goes for $\nu$-binders which are inserted as a result of translating bound send to unbound send.
 For instance, the process $\pcp{\tm{\send{x}{y}{P}}}$ is blocked, but its translation uses $\pgv{\tm{\new}}$ and can reduce.
-To address this issue, we use a second translation, $\tm{\cpgvC{\cdot}}$, which is equivalent to $\tm{\cpgvM{\cdot}}$ followed by reduction using \LabTirName{E-New} and \LabTirName{E-Spawn}:
+To address this issue, we use a second translation, $\tm{\cpgvC{\cdot}}$, which is equivalent to $\tm{\cpgvM{\cdot}}$ followed by reductions using \LabTirName{E-New} and \LabTirName{E-Spawn}:
 \begin{align*}
@@
 The translations $\tm{\cpgvM{\cdot}}$ and $\tm{\cpgvC{\cdot}}$ preserve typing, and the latter induces a sound and complete operational correspondence.
-The proofs are standard inductions.
-The details can be found in~\cref{prf:lem-pcp-to-pgv-terms-preservation,prf:lem-pcp-to-pgv-terms-preservation,prf:lem-pcp-to-pgv-cpgvM-to-cpgvC,prf:thm-pcp-to-pgv-operational-correspondence-soundness,prf:thm-pcp-to-pgv-operational-correspondence-completeness}.
+The proofs are standard inductions and can be found in~\cref{prf:lem-pcp-to-pgv-terms-preservation,prf:lem-pcp-to-pgv-terms-preservation,prf:lem-pcp-to-pgv-cpgvM-to-cpgvC,prf:thm-pcp-to-pgv-operational-correspondence-soundness,prf:thm-pcp-to-pgv-operational-correspondence-completeness}.
 \begin{compacttheorems}
@@
 \end{compacttheorems}
+\endgroup
 \end{document}

### 2021-01-16 priority-gv.tex
Rewrote PCP section.

@@
 \subsubsection*{Typing}
-\todo{This is a type-and-effect system and should probably be presented as such, as in [37].}
 \Cref{fig:pgv-typing} gives the typing rules for PGV.
@@
 For this, we prove~\cref{lem:pgv-value-done}, which shows that values have finished all their communication, and that any priorities in the type of the value come from the typing environment.
+The proofs are standard inductions and can be found in \cref{prf:lem-pgv-value-done,prf:lem-pgv-substitution,prf:lem-pgv-subject-reduction-terms,prf:lem-pgv-subject-congruence,prf:thm-pgv-subject-reduction-confs}.
 \begin{compacttheorems}
@@
 \end{compacttheorems}
-The proofs are standard inductions.
-The details can be found in \cref{prf:lem-pgv-value-done,prf:lem-pgv-substitution,prf:lem-pgv-subject-reduction-terms,prf:lem-pgv-subject-congruence,prf:thm-pgv-subject-reduction-confs}.
 

### 2021-01-16 appendix.tex
Minor changes.

@@
 \usingnamespace{pcp}
+\subsection{Revisiting Priority CP}
 
@@
     \begin{restatabledefinition}{defpcpactions}[Actions]
-    A~process acts on an endpoint $\tm{x}$ if it is of the form $\tm{\link{x}{y}}$, $\tm{\link{y}{x}}$, $\tm{\send{x}{y}{P}}$, $\tm{\recv{x}{y}{P}}$, $\tm{\close{x}{P}}$, $\tm{\wait{x}{P}}$, $\tm{\inl{x}{P}}$, $\tm{\inr{x}{P}}$, $\tm{\offer{x}{P}{Q}}$, or $\tm{\absurd{x}}$.
+    A~process acts on an endpoint $\tm{x}$ if it is $\tm{\link{x}{y}}$, $\tm{\link{y}{x}}$, $\tm{\send{x}{y}{P}}$, $\tm{\recv{x}{y}{P}}$, $\tm{\close{x}{P}}$, $\tm{\wait{x}{P}}$, $\tm{\inl{x}{P}}$, $\tm{\inr{x}{P}}$, $\tm{\offer{x}{P}{Q}}$, or $\tm{\absurd{x}}$.
 A~process is an action if it acts on some endpoint $\tm{x}$.
@@
     \label{def:pcp-canonical-forms}
-    A~process $\tm{P}$ is in canonical form if it is of the form $\tm{\halt}$ or $\tm{\res{x_1}{x'_1}{\dots\res{x_n}{x'_n}{(P_1 \parallel\dots\parallel P_m)}}}$ where $m>0$ and each $\tm{P_j}$ is an action.
+    A~process $\tm{P}$ is in canonical form if it is $\tm{\halt}$ or $\tm{\res{x_1}{x'_1}{\dots\res{x_n}{x'_n}{(P_1 \parallel\dots\parallel P_m)}}}$ where $m>0$ and each $\tm{P_j}$ is an action.
   \end{restatabledefinition}
@@
 
-\subsubsection*{Translation}
+\subsection{Translation}
 \begingroup

### 2021-01-16 priority-cp.tex
Minor changes.
Changed sec title

@@
 
-\subsubsection*{Revisiting Priority CP}
+\subsection{Revisiting Priority CP}
 \begingroup
@@
 
-\subsubsection*{Translation}
+\subsection{Correspondence between PGV and PCP}
 \begingroup
@@
 
-Unfortunately, the operational correspondence along $\tm{\cpgvM{\cdot}}$ is unsound, as it translates $\nu$-binders and parallel compositions to $\pgv{\tm{\new}}$ and $\pgv{\tm{\spawn}}$, which reduce back to their equivalent configuration constructs using \LabTirName{E-New} and \LabTirName{E-Spawn}.
-The same goes for $\nu$-binders which are inserted as a result of translating bound send to unbound send.
+Unfortunately, the operational correspondence along $\tm{\cpgvM{\cdot}}$ is unsound, as it translates $\nu$-binders and parallel compositions to $\pgv{\tm{\new}}$ and $\pgv{\tm{\spawn}}$, which can reduce to their equivalent configuration constructs using \LabTirName{E-New} and \LabTirName{E-Spawn}.
+The same goes for $\nu$-binders which are inserted when translating bound send to unbound send.
 For instance, the process $\pcp{\tm{\send{x}{y}{P}}}$ is blocked, but its translation uses $\pgv{\tm{\new}}$ and can reduce.

### 2021-01-20 introduction.tex
Update introduction.tex

@@
 Session types are a type formalism which can be used to specify and verify communication protocols between two or more communicating agents~\cite{honda93,takeuchihonda94,hondavasconcelos98,carbonehonda07}.
-\todo{These two settings are very different.
-The $\pi$-calculus is a mobile process algebra preceding inside which session types were developed and GV is an ad-hoc calculus made to demonstrate how propositions can be used as types.}
-Session types have been studied for many programming paradigms, including concurrent and functional paradigms.
-Most notably, they have been defined for the $\pi$-calculus~\cite{sangiorgiwalker01}---a process calculus for communication and concurrency, and for Good Variation (GV)~\cite{wadler15,lindleymorris15}---a linear concurrent $\lambda$-calculus.
+Session types have been studied for many programming paradigms, most notably for concurrent and functional paradigms.
+They have been defined for the $\pi$-calculus~\cite{sangiorgiwalker01}---a process calculus for communication and concurrency, and for Good Variation (GV)~\cite{wadler15,lindleymorris15}---a linear concurrent $\lambda$-calculus.
 
-Session-typed concurrent $\lambda$-calculus and $\pi$-calculus satisfy communication safety and session fidelity, namely that communication proceeds without any type mismatch of the data being transmitted and follows the predefined protocol specified by a session type.
+Session type systems guarantee communication safety and session fidelity, namely that communication proceeds without any type mismatch of the data being transmitted and follows the predefined protocol specified by a session type.
 Another key property in a concurrent setting is deadlock freedom, stating that agents communicate without getting stuck in cyclic dependencies.
@@
 Moreover, the benefit of working with GV over other session-typed functional languages is that GV has strong ties to linear logic, via its relation to CP~\cite{wadler14}, and consequently it has strong formal properties, \eg deadlock freedom, albeit in a restrictive way.
-Following the above observations, we aim to further investigate GV and deadlock freedom, and thus we pose our research question:
+
+Following the above observations, we aim to further investigate GV and deadlock freedom, and thus we pose our research question.
 
@@
 \textbf{Updated Priority CP.} We present an updated version of Priority CP~\cite{dardhagay18} in \cref{sec:pcp}.
-We remove commuting conversions and move away from reduction as cut elimination, towards reduction based on structural congruence, as in process calculi.
-We present Milner's Cyclic Scheduler in PCP (\cref{sec:pcp-example}).
+We remove commuting conversions and move away from reduction as cut elimination, towards reduction based on structural congruence, as it is standard in process calculi.
+%We present Milner's Cyclic Scheduler in PCP (\cref{sec:pcp-example}).
 
@@
 \textbf{Connection to linear logic}.
-The connection of PGV to logic is given via a sound and complete encoding of PCP to PGV in~\cref{sec:pcp} and prove that it preserves typing (\cref{thm:pcp-to-pgv-confs-preservation}) and satisfies operational correspondence (\cref{thm:pcp-to-pgv-operational-correspondence-soundness,thm:pcp-to-pgv-operational-correspondence-completeness}).
-We showcase it with the encoding of cyclic scheduler to PGV (\cref{sec:pgv-example}).
+The connection of PGV to logic is given via a \emph{sound and complete encoding} of PCP to PGV in~\cref{sec:pcp} and prove that it preserves typing (\cref{thm:pcp-to-pgv-confs-preservation}) and satisfies operational correspondence (\cref{thm:pcp-to-pgv-operational-correspondence-soundness,thm:pcp-to-pgv-operational-correspondence-completeness}).
+%We showcase it with the encoding of cyclic scheduler to PGV (\cref{sec:pgv-example}).
 \end{enumerate}

### 2021-01-20 conclusion.tex
Fixed citations.

@@
 Gay~\etal~\cite{gaynagarajan03} and Vasconcelos~\etal~\cite{vasconcelosravara04,vasconcelosgay06} were the first to introduce a functional language with session types.
-However, such works, including early GV~\cite{gayvasconcelos10,gayvasconcelos12} did not guarantee deadlock freedom, until it was addressed via syntactic restrictions~\cite{lindleymorris15,wadler15}.
+However, such works, including early GV~\cite{gayvasconcelos10,gayvasconcelos12} did not guarantee deadlock freedom, until it was addressed via syntactic restrictions~\cite{lindleymorris15,wadler14}.
 Toninho~\etal~\cite{toninhocaires12} present a translation of simply-typed $\lambda$-calculus into session-typed $\pi$-calculus.

### 2021-01-20 introduction.tex
Fixed citations.

@@
 Session types are a type formalism which can be used to specify and verify communication protocols between two or more communicating agents~\cite{honda93,takeuchihonda94,hondavasconcelos98,carbonehonda07}.
+
 Session types have been studied for many programming paradigms, most notably for concurrent and functional paradigms.
-They have been defined for the $\pi$-calculus~\cite{sangiorgiwalker01}---a process calculus for communication and concurrency, and for Good Variation (GV)~\cite{wadler15,lindleymorris15}---a linear concurrent $\lambda$-calculus.
+They have been defined for the $\pi$-calculus~\cite{sangiorgiwalker01}---a process calculus for communication and concurrency, and for Good Variation (GV)~\cite{wadler14,lindleymorris15}---a linear concurrent $\lambda$-calculus.
 
@@
 
-With regards to $\pi$-calculus, a growing line of work on deadlock freedom leverages the Curry-Howard correspondences between intuitionistic or classical linear logic and session-typed $\pi$-calculus~\cite{cairespfenning10,wadler12}, which guarantee deadlock freedom by design.
+With regards to $\pi$-calculus, a growing line of work on deadlock freedom leverages the Curry-Howard correspondences between intuitionistic or classical linear logic and session-typed $\pi$-calculus~\cite{cairespfenning10,wadler14}, which guarantee deadlock freedom by design.
 This is achieved by merging constructs for parallel composition and channel restriction under the logical cut rule, which forces processes to share only one channel for communication and eliminates cyclic dependencies.
@@
 With regards to GV, the original work on session types in GV~\cite{gayvasconcelos10,gayvasconcelos12} did not satisfy deadlock freedom.
-Later on, calculi in the GV family~\cite{wadler15,lindleymorris15} have achieved deadlock freedom via a syntactic restriction, \ie by combining channel creation and thread spawning into a single operation, called fork.
+Later on, calculi in the GV family~\cite{wadler14,lindleymorris15} have achieved deadlock freedom via a syntactic restriction, \ie by combining channel creation and thread spawning into a single operation, called fork.
 The fork construct is the term representation of the cut construct in CP, which evaluates to a channel restriction combined with a parallel composition.

### 2021-01-22 appendix.tex
Changed sec title
Made a bunch of minor changes?

@@
 \usingnamespace{pgv}
-
-% Scheduler example
-\input{fig/pgv-scheduler}
-
-% Proofs that the syntactic sugar is well-typed.
-\input{fig/pgv-typing-sugar}
-
 \restatelemma{lempgvvaluedone}
@@
 \input{prf/thm-pgv-closed-progress-confs}
+
+\input{fig/pgv-typing-sugar}
 \endgroup
@@
 
-\input{fig/pcp-operational-semantics-and-typing}
-
 \subsubsection*{Operational Semantics}
-The operational semantics for PCP, given in \cref{fig:pcp-operational-semantics-and-typing}, is defined as a reduction relation $\pcp{\red}$ on processes (bottom) and uses structural congruence (top).
+The operational semantics for PCP, given in \cref{fig:pcp-operational-semantics}, is defined as a reduction relation $\pcp{\red}$ on processes (bottom) and uses structural congruence (top).
 Each of the axioms of structural congruence corresponds to the axiom of the same name for PGV.
@@
 
+\input{fig/pcp-operational-semantics}
+
 \subsubsection*{Typing}
-\Cref{fig:pcp-operational-semantics-and-typing} gives the typing rules for our version of PCP.
+\Cref{fig:pcp-typing} gives the typing rules for our version of PCP.
 A typing judgement $\pcp{\seq{P}{\Gamma}}$ states that ``process $\pcp{\tm{P}}$ is well typed under the typing context $\pcp{\ty{\Gamma}}$''.
+
+\input{fig/pcp-typing}
 
@@
 
-\subsection{Translation}
+\subsection{Correspondence between PGV and PCP}
 \begingroup

### 2021-01-26 priority-gv.tex
Made a bunch of minor changes?
In the process of adding examples.

@@
 Priority GV offers a more fine-grained analysis of communication structures, and by separating channel creation form thread spawning it allows cyclic structures.
+
+\todo{Describe examples.}
+
+\begin{center}
+\begin{minipage}{0.5\linewidth}
+\begin{example}[Cyclic Structure]
+  \[\tm{%
+    \begin{array}{l}
+      \letpair{x}{x'}{\new}{}
+      \\
+      \letpair{y}{y'}{\new}{}
+      \\
+      \spawn\left(%
+      \begin{array}{l}
+        {\andthen{\recv\;{x'}}{\send\;{\pair{\unit}{y}}}};
+        \\
+        {\andthen{\wait\;{x'}}{\close\;{y}}}
+      \end{array}%
+      \right);
+      \\
+      \underline{\andthen{\send\;{\pair{\unit}{x}}}{\recv\;{y'}}};
+      \\
+      {\andthen{\close\;{x}}{\wait\;{y'}}}
+    \end{array}
+  }\]
+\end{example}
+\end{minipage}%
+\begin{minipage}{0.5\linewidth}
+\begin{example}[Deadlock]
+  \[\tm{%
+    \begin{array}{l}
+      \letpair{x}{x'}{\new}{}
+      \\
+      \letpair{y}{y'}{\new}{}
+      \\
+      \spawn\left(%
+      \begin{array}{l}
+        {\andthen{\recv\;{x'}}{\send\;{\pair{\unit}{y}}}};
+        \\
+        {\andthen{\wait\;{x'}}{\close\;{y}}}
+      \end{array}%
+      \right);
+      \\
+      \underline{\andthen{\recv\;{y'}}{\send\;{\pair{\unit}{x}}}}
+      \\
+      {\andthen{\close\;{x}}{\wait\;{y'}}}
+    \end{array}
+  }\]
+\end{example}
+\end{minipage}
+\end{center}
 
@@
 \begin{enumerate*}[label=(\roman*)]
-\item an action with priority $\cs{o}$ happens before an action with priority $\cs{o+n}$; and
-\item communication requires \emph{equal} priorities for the dual actions.
+\item an action with lower priority happens before an action with higher priority; and
+\item communication requires \emph{equal} priorities for dual actions.
 \end{enumerate*}
 
-In PGV, sequents are annotated with a priority bound $\cs{p}$, which is the \emph{upper bound} on the priority of actions performed while evaluating the term.
-The \emph{lower bound} is obtained from the typing environment.
-Terms with sequential evaluation, such as \LabTirName{T-Andthen}, have side conditions to ensure deadlock freedom.
-For instance, the side condition for typing $\tm{\andthen{M}{N}}$ requires that $\tm{M}$ has finished communicating before $\tm{N}$ starts communicating.
-The rule \LabTirName{T-Lam} stores the lower and upper priority bound of the function body in the type.
+In PGV, we keep track of a lower and upper bound on the priorities of a term, \ie, while evaluating the term, when does it start communicating, and when does it finish.
+The upper bound is written on the sequent, whereas the lower bound is approximated from the typing environment.
+Typing rules for sequential constructs enforce sequentially, \eg, the typing for $\tm{\andthen{M}{N}}$ has a side condition which requires that the upper bound of $\tm{M}$ is smaller than the lower bound of $\tm{N}$, \ie, $\tm{M}$ finishes before $\tm{N}$ starts.
+\begin{mathpar}
+  \inferrule*{
+    \inferrule*{
+      \inferrule*{
+      }{
+        \tseq{\tmty{y'}{\tyrecv{\tyunit}{\tyendr}}}{y'}{\tyrecv{\tyunit}{\tyendr}}
+      }
+    }{
+      \tseq{\tmty{y'}{\tyrecv{\tyunit}{\tyendr}}}{\recv\;{y'}}{\typrod{\tyunit}{\tyendr}}
+    }
+    \\
+    \inferrule*{
+    }{
+      \tseq{}{\send\;{\pair{\unit}{x}}}{}
+    }
+  }{
+    \tseq{}{\andthen{\send\;{\pair{\unit}{x}}}{\recv\;{y'}}}{}
+  }
+
+  {\andthen{\recv\;{y'}}{\send\;{\pair{\unit}{x}}}}
+\end{mathpar}
+Closures suspend communication, so \LabTirName{T-Lam} stores the priority bounds of the function body on the function type, and \LabTirName{T-App} restores them.
+For instance, $\tm{\lambda{x}.\send\;\pair{x}{y}}$ is assigned the type $\ty{\tylolli[\cs{o},\cs{o}]{A}{S}}$, \ie, a~function which, when applied, starts and finishes communicating at priority $\cs{o}$:
+\begin{mathpar}
+  \inferrule*{
+    \inferrule*{
+      \inferrule*{}{\tm{\send}}
+      \\
+      \inferrule*{
+        \inferrule*{
+        }{
+          \tseq[\pbot]{\tmty{x}{A}}{x}{A}
+        }
+        \\
+        \inferrule*{
+        }{
+          \tseq[\pbot]
+          {\tmty{x}{A},\tmty{y}{\tysend[\cs{o}]{A}{S}}}
+          {y}{\tysend[\cs{o}]{A}{S}}
+        }
+      }{
+        \tseq[\pbot]
+        {\tmty{x}{A},\tmty{y}{\tysend[\cs{o}]{A}{S}}}
+        {\pair{x}{y}}{\typrod{A}{\tysend[\cs{o}]{A}{S}}}
+      }
+    }{
+      \tseq[\cs{o}]
+      {\tmty{x}{A},\tmty{y}{\tysend[\cs{o}]{A}{S}}}
+      {\send\;\pair{x}{y}}{S}
+    }
+  }{
+    \tseq[\pbot]
+    {\tmty{y}{\tysend[\cs{o}]{A}{S}}}
+    {\lambda{x}.\send\;\pair{x}{y}}{\tylolli[\cs{o},\cs{o}]{A}{S}}
+  }
+\end{mathpar}
+The approximation of the lower bound is not entirely accurate, \eg, $\tm{\lambda{x}.\send\;\pair{x}{y}}$ has lower bound $\ptop$, not $\cs{o}$, but as we only ever under-approximate, it is safe.
 

### 2021-01-27 priority-gv.tex
Progress on adding examples.

@@
 \begin{example}[Cyclic Structure]
+  \label{ex:cycle}
   \[\tm{%
@@
       \begin{array}{l}
-        {\andthen{\recv\;{x'}}{\send\;{\pair{\unit}{y}}}};
+        {\letpair{\unit}{x'}{\recv\;{x'}}{}}
+        \\
+        {\letbind{y}{\send\;{\pair{\unit}{y}}}{}}
         \\
@@
       \\
-      \underline{\andthen{\send\;{\pair{\unit}{x}}}{\recv\;{y'}}};
+      \underline{\letbind{x}{\send\;{\pair{\unit}{x}}}{}}
+      \\
+      \underline{\letpair{\unit}{y'}{\recv\;{y'}}{}}
       \\
@@
 \begin{example}[Deadlock]
+  \label{ex:deadlock}
   \[\tm{%
@@
       \\
-      \underline{\andthen{\recv\;{y'}}{\send\;{\pair{\unit}{x}}}}
+      \underline{\letpair{\unit}{y'}{\recv\;{y'}}{}}
+      \\
+      \underline{\letbind{x}{\send\;{\pair{\unit}{x}}}{}}
       \\
@@
 Typing rules for sequential constructs enforce sequentially, \eg, the typing for $\tm{\andthen{M}{N}}$ has a side condition which requires that the upper bound of $\tm{M}$ is smaller than the lower bound of $\tm{N}$, \ie, $\tm{M}$ finishes before $\tm{N}$ starts.
+The typing rule for $\tm{\new}$ ensures that both endpoints of a channel share the same priorities.
+Together, these two constraints guarantee deadlock freedom.
+For instance, the program in \cref{ex:deadlock} is deadlocked.
+Crucially, it composes the terms below in parallel.
+While each of these terms itself is well-typed, they impose opposite conditions on the priorities, so their composition is ill-typed.
+(We omit the priorities on $\ty{\tyends}$ and $\ty{\tyendr}$.)
 \begin{mathpar}
+  \resizebox{\linewidth}{!}{\ensuremath{%
+      \inferrule*{
+        \tseq[\cs{o'}]
+        {\tmty{y'}{\tyrecv[\cs{o'}]{\tyunit}{\tyendr}}}
+        {\recv\;{y'}}
+        {\typrod{\tyunit}{\tyendr}}
+        {\hskip 1em}
+        \tseq[\cs{p}]
+        {\tmty{x}{\tysend[\cs{o}]{\tyunit}{\tyends}},\tmty{y'}{\tyendr}}
+        {\letbind{x}{\send\;{\pair{\unit}{x}}}{\dots}}
+        {\tyunit}
+        {\hskip 1em}
+        \cs{o'}<\cs{o}
+      }{%
+        \tseq[\cs{p}]
+        {%
+          \tmty{x}{\tysend[\cs{o}]{\tyunit}{\tyends}},
+          \tmty{y'}{\tyrecv[\cs{o'}]{\tyunit}{\tyendr}}
+        }
+        {\letpair{\unit}{y'}{\recv\;{y'}}{\letbind{x}{\send\;{\pair{\unit}{x}}}{\dots}}}
+        {\tyunit}
+      }
+    }}
+  \\
+  \resizebox{\linewidth}{!}{\ensuremath{%
+      \inferrule*{
+        \tseq[\cs{o}]
+        {\tmty{x'}{\tyrecv[\cs{o}]{\tyunit}{\tyendr}}}
+        {\recv\;{x'}}
+        {\typrod{\tyunit}{\tyendr}}
+        {\hskip 1em}
+        \tseq[\cs{q}]
+        {\tmty{y}{\tysend[\cs{o'}]{\tyunit}{\tyends}},\tmty{x'}{\tyendr}}
+        {\letbind{y}{\send\;{\pair{\unit}{y}}}{\dots}}
+        {\tyunit}
+        {\hskip 1em}
+        \cs{o}<\cs{o'}
+      }{%
+        \tseq[\cs{q}]
+        {%
+          \tmty{y}{\tysend[\cs{o'}]{\tyunit}{\tyends}},
+          \tmty{x'}{\tyrecv[\cs{o}]{\tyunit}{\tyendr}}
+        }
+        {\letpair{\unit}{x'}{\recv\;{x'}}{\letbind{y}{\send\;{\pair{\unit}{y}}}{\dots}}}
+        {\tyunit}
+      }
+    }}
+\end{mathpar}
+Closures suspend communication, so \LabTirName{T-Lam} stores the priority bounds of the function body on the function type, and \LabTirName{T-App} restores them.
+For instance, $\tm{\lambda{x}.\send\;\pair{x}{y}}$ is assigned the type $\ty{\tylolli[\cs{o},\cs{o}]{A}{S}}$, \ie, a~function which, when applied, starts and finishes communicating at priority $\cs{o}$:
+\begin{mathpar}
+  \resizebox{\linewidth}{!}{\ensuremath{%
       \inferrule*{
@@
           }{
-        \tseq{\tmty{y'}{\tyrecv{\tyunit}{\tyendr}}}{y'}{\tyrecv{\tyunit}{\tyendr}}
+            \tmty{\send}{\tylolli[\ptop,\cs{o}]{\typrod{A}{\tysend[\cs{o}]{A}{S}}}{S}}
           }
-    }{
-      \tseq{\tmty{y'}{\tyrecv{\tyunit}{\tyendr}}}{\recv\;{y'}}{\typrod{\tyunit}{\tyendr}}
-    }
-    \\
-    \inferrule*{
-    }{
-      \tseq{}{\send\;{\pair{\unit}{x}}}{}
-    }
-  }{
-    \tseq{}{\andthen{\send\;{\pair{\unit}{x}}}{\recv\;{y'}}}{}
-  }
-
-  {\andthen{\recv\;{y'}}{\send\;{\pair{\unit}{x}}}}
-\end{mathpar}
-Closures suspend communication, so \LabTirName{T-Lam} stores the priority bounds of the function body on the function type, and \LabTirName{T-App} restores them.
-For instance, $\tm{\lambda{x}.\send\;\pair{x}{y}}$ is assigned the type $\ty{\tylolli[\cs{o},\cs{o}]{A}{S}}$, \ie, a~function which, when applied, starts and finishes communicating at priority $\cs{o}$:
-\begin{mathpar}
-  \inferrule*{
-    \inferrule*{
-      \inferrule*{}{\tm{\send}}
-      \\
+          \and
           \inferrule*{
@@
             }
-        \\
+            \and
             \inferrule*{
@@
       }
+    }}
 \end{mathpar}

### 2021-01-27 introduction.tex
Minor layout changes.

@@
 \begin{enumerate}
-\item
-\textbf{Priority GV}. We present Priority GV (PGV) in~\cref{sec:pgv}, and prove subject reduction (\cref{thm:pgv-subject-reduction-confs}) and progress (\cref{thm:pgv-closed-progress-confs}).
+\item \textbf{Priority GV}. We present Priority GV (PGV) in~\cref{sec:pgv}, and prove subject reduction (\cref{thm:pgv-subject-reduction-confs}) and progress (\cref{thm:pgv-closed-progress-confs}).
 PGV is the \emph{first} session-typed functional language with priorities \emph{and} strong ties to linear logic via its correspondence with PCP and its relation to the CP and GV languages.
-To illustrate newfound expressiveness of PGV, we present Milner's Cyclic Scheduler~\cite{milner89} (\cref{sec:pgv-example}) as well as other examples.
-
- \item
- \textbf{Solving GV problems}. PGV addresses several problems in the original GV language, most notably:
- 	\begin{enumerate}
+\item \textbf{Solving GV problems}. PGV addresses several problems in the original GV language, most notably:
+ 	\begin{enumerate*}
 	\item our version does not have the pseudo-type $S^\sharp$;
 	\item our structural congruence is type preserving.
-	\end{enumerate}
-\item
-\textbf{Updated Priority CP.} We present an updated version of Priority CP~\cite{dardhagay18} in \cref{sec:pcp}.
+	\end{enumerate*}
+\item \textbf{Updated Priority CP.} We present an updated version of Priority CP~\cite{dardhagay18} in \cref{sec:pcp}.
 We remove commuting conversions and move away from reduction as cut elimination, towards reduction based on structural congruence, as it is standard in process calculi.
-%We present Milner's Cyclic Scheduler in PCP (\cref{sec:pcp-example}).
-
-\item
-\textbf{Connection to linear logic}.
+\item \textbf{Connection to linear logic}.
 The connection of PGV to logic is given via a \emph{sound and complete encoding} of PCP to PGV in~\cref{sec:pcp} and prove that it preserves typing (\cref{thm:pcp-to-pgv-confs-preservation}) and satisfies operational correspondence (\cref{thm:pcp-to-pgv-operational-correspondence-soundness,thm:pcp-to-pgv-operational-correspondence-completeness}).
-%We showcase it with the encoding of cyclic scheduler to PGV (\cref{sec:pgv-example}).
 \end{enumerate}

### 2021-01-27 priority-gv.tex
Minor layout changes.

--- 2021-01-26 priority-gv.tex
@@
 Typing rules for sequential constructs enforce sequentially, \eg, the typing for $\tm{\andthen{M}{N}}$ has a side condition which requires that the upper bound of $\tm{M}$ is smaller than the lower bound of $\tm{N}$, \ie, $\tm{M}$ finishes before $\tm{N}$ starts.
+The typing rule for $\tm{\new}$ ensures that both endpoints of a channel share the same priorities.
+Together, these two constraints guarantee deadlock freedom.
+For instance, the program in \cref{ex:deadlock} is deadlocked.
+Crucially, it composes the terms below in parallel.
+While each of these terms itself is well-typed, they impose opposite conditions on the priorities, so their composition is ill-typed.
+(We omit the priorities on $\ty{\tyends}$ and $\ty{\tyendr}$.)
 \begin{mathpar}
+  \small
+  \inferrule*{
+    \tseq[\cs{o'}]
+    {\tmty{y'}{\tyrecv[\cs{o'}]{\tyunit}{\tyendr}}}
+    {\recv\;{y'}}
+    {\typrod{\tyunit}{\tyendr}}
+    \\
+    \tseq[\cs{p}]
+    {\tmty{x}{\tysend[\cs{o}]{\tyunit}{\tyends}},\tmty{y'}{\tyendr}}
+    {\letbind{x}{\send\;{\pair{\unit}{x}}}{\dots}}
+    {\tyunit}
+    \and
+    \cs{o'}<\cs{o}
+  }{%
+    \tseq[\cs{p}]
+    {%
+      \tmty{x}{\tysend[\cs{o}]{\tyunit}{\tyends}},
+      \tmty{y'}{\tyrecv[\cs{o'}]{\tyunit}{\tyendr}}
+    }
+    {\letpair{\unit}{y'}{\recv\;{y'}}{\letbind{x}{\send\;{\pair{\unit}{x}}}{\dots}}}
+    {\tyunit}
+  }
+
+  \inferrule*{
+    \tseq[\cs{o}]
+    {\tmty{x'}{\tyrecv[\cs{o}]{\tyunit}{\tyendr}}}
+    {\recv\;{x'}}
+    {\typrod{\tyunit}{\tyendr}}
+    \\
+    \tseq[\cs{q}]
+    {\tmty{y}{\tysend[\cs{o'}]{\tyunit}{\tyends}},\tmty{x'}{\tyendr}}
+    {\letbind{y}{\send\;{\pair{\unit}{y}}}{\dots}}
+    {\tyunit}
+    \and
+    \cs{o}<\cs{o'}
+  }{%
+    \tseq[\cs{q}]
+    {%
+      \tmty{y}{\tysend[\cs{o'}]{\tyunit}{\tyends}},
+      \tmty{x'}{\tyrecv[\cs{o}]{\tyunit}{\tyendr}}
+    }
+    {\letpair{\unit}{x'}{\recv\;{x'}}{\letbind{y}{\send\;{\pair{\unit}{y}}}{\dots}}}
+    {\tyunit}
+  }
+\end{mathpar}
+Closures suspend communication, so \LabTirName{T-Lam} stores the priority bounds of the function body on the function type, and \LabTirName{T-App} restores them.
+For instance, $\tm{\lambda{x}.\send\;\pair{x}{y}}$ is assigned the type $\ty{\tylolli[\cs{o},\cs{o}]{A}{S}}$, \ie, a~function which, when applied, starts and finishes communicating at priority $\cs{o}$:
+\begin{mathpar}
+  \small
   \inferrule*{
@@
       }{
-        \tseq{\tmty{y'}{\tyrecv{\tyunit}{\tyendr}}}{y'}{\tyrecv{\tyunit}{\tyendr}}
+        \tmty{\send}{\tylolli[\ptop,\cs{o}]{\typrod{A}{\tysend[\cs{o}]{A}{S}}}{S}}
       }
-    }{
-      \tseq{\tmty{y'}{\tyrecv{\tyunit}{\tyendr}}}{\recv\;{y'}}{\typrod{\tyunit}{\tyendr}}
-    }
-    \\
-    \inferrule*{
-    }{
-      \tseq{}{\send\;{\pair{\unit}{x}}}{}
-    }
-  }{
-    \tseq{}{\andthen{\send\;{\pair{\unit}{x}}}{\recv\;{y'}}}{}
-  }
-
-  {\andthen{\recv\;{y'}}{\send\;{\pair{\unit}{x}}}}
-\end{mathpar}
-Closures suspend communication, so \LabTirName{T-Lam} stores the priority bounds of the function body on the function type, and \LabTirName{T-App} restores them.
-For instance, $\tm{\lambda{x}.\send\;\pair{x}{y}}$ is assigned the type $\ty{\tylolli[\cs{o},\cs{o}]{A}{S}}$, \ie, a~function which, when applied, starts and finishes communicating at priority $\cs{o}$:
-\begin{mathpar}
-  \inferrule*{
-    \inferrule*{
-      \inferrule*{}{\tm{\send}}
-      \\
+      \and
       \inferrule*{
@@
         }
-        \\
+        \and
         \inferrule*{

### 2021-01-27 priority-gv.tex
Removed part of the higher-order typing example.

@@
 Closures suspend communication, so \LabTirName{T-Lam} stores the priority bounds of the function body on the function type, and \LabTirName{T-App} restores them.
-For instance, $\tm{\lambda{x}.\send\;\pair{x}{y}}$ is assigned the type $\ty{\tylolli[\cs{o},\cs{o}]{A}{S}}$, \ie, a~function which, when applied, starts and finishes communicating at priority $\cs{o}$:
+For instance, $\tm{\lambda{x}.\send\;\pair{x}{y}}$ is assigned the type $\ty{\tylolli[\cs{o},\cs{o}]{A}{S}}$, \ie, a~function which, when applied, starts and finishes communicating at priority $\cs{o}$.
 \begin{mathpar}
@@
       \and
-      \inferrule*{
-        \inferrule*{
-        }{
-          \tseq[\pbot]{\tmty{x}{A}}{x}{A}
-        }
-        \and
-        \inferrule*{
-        }{
-          \tseq[\pbot]
-          {\tmty{x}{A},\tmty{y}{\tysend[\cs{o}]{A}{S}}}
-          {y}{\tysend[\cs{o}]{A}{S}}
-        }
+      \inferrule*{%
+        % \inferrule*{
+        % }{
+        %   \tseq[\pbot]{\tmty{x}{A}}{x}{A}
+        % }
+        % \and
+        % \inferrule*{
+        % }{
+        %   \tseq[\pbot]
+        %   {\tmty{x}{A},\tmty{y}{\tysend[\cs{o}]{A}{S}}}
+        %   {y}{\tysend[\cs{o}]{A}{S}}
+        % }
       }{
@@
 \end{mathpar}
-The approximation of the lower bound is not entirely accurate, \eg, $\tm{\lambda{x}.\send\;\pair{x}{y}}$ has lower bound $\ptop$, not $\cs{o}$, but as we only ever under-approximate, it is safe.
 

### 2021-01-27 priority-gv.tex
Wrote text introducing the examples.

@@
 
-\todo{Describe examples.}
+We illustrate this with two programs in PGV, \cref{ex:cycle,ex:deadlock}.
+Each program contains two processes---the main process, and the child process created by $\tm{\spawn}$---which communicate using \emph{two} channels.
+The child process receives a unit over the channel $\tm{x}/\tm{x'}$, and then sends a unit over the channel $\tm{y}/\tm{y'}$.
+The main process does one of two things:
+\begin{enumerate*}[label=(\alph*)]
+\item in \cref{ex:cycle}, it sends a unit over the channel $\tm{x}/\tm{x'}$, and then waits to receive a unit over the channel $\tm{y}/\tm{y'}$;
+\item in \cref{ex:deadlock}, it does these in the opposite order, which results in a deadlock.
+\end{enumerate*}
+We will show that only \cref{ex:cycle} is typeable in PGV, as opposed to neither in GV~\cite{wadler14} and both in its predecessor~\cite{gayvasconcelos10}.
 
@@
 \subsubsection*{Operational Semantics}
-
 Priority GV terms are evaluated as part of a configuration of processes. Configurations are defined by the following grammar:
@@
       \inferrule*{%
-        % \inferrule*{
-        % }{
-        %   \tseq[\pbot]{\tmty{x}{A}}{x}{A}
-        % }
-        % \and
-        % \inferrule*{
-        % }{
-        %   \tseq[\pbot]
-        %   {\tmty{x}{A},\tmty{y}{\tysend[\cs{o}]{A}{S}}}
-        %   {y}{\tysend[\cs{o}]{A}{S}}
-        % }
+        \inferrule*{
+        }{
+          \tseq[\pbot]{\tmty{x}{A}}{x}{A}
+        }
+        \and
+        \inferrule*{
+        }{
+          \tseq[\pbot]
+          {\tmty{x}{A},\tmty{y}{\tysend[\cs{o}]{A}{S}}}
+          {y}{\tysend[\cs{o}]{A}{S}}
+        }
       }{
@@
   If $\tseq[\cs{p}]{\ty{\Gamma}}{M}{T}$ and $\ty{\Gamma}$ contains only session types, then:
-  \begin{itemize}
+  \begin{enumerate*}[label=\emph{(\alph*)}]
   \item $\tm{M}$ is a value;
@@
   \item $\tm{M}$ is ready.
-  \end{itemize}
+  \end{enumerate*}
 \end{restatablelemma}

### 2021-01-28 appendix.tex
Added scheduler example in the appendix.

@@
 
+\section{Milner's Cyclic Scheduler}
+\label{sec:milner}
+As an example of a deadlock-free cyclic process, Dardha and Gay~\cite{dardhagay18} introduce an implementation of Milner's cyclic scheduler~\cite{milner89} in Priority CP.
+We reproduce that scheduler here, and show its translation to Priority GV.
+
+\begingroup
+\usingnamespace{pcp}
+\begin{example}[Milner's Cyclic Scheduler, PCP]
+  \label{ex:pcp-scheduler}
+  A set of processes $\tm{\proc{i}}$, $1\leq{i}\leq{n}$, is scheduled to perform some tasks in cyclic order, starting with $\tm{\proc1}$, ending with $\tm{\proc{n}}$, and notifying $\tm{\proc1}$ when all processes have finished.
+
+  Our scheduler $\tm{\sched}$ consists of set of agents $\tm{\agent{i}}$, $1\leq{i}\leq{n}$, each representing their respective process.
+Each process $\tm{\proc{i}}$ waits for the signal to start their task on $\tm{a'_i}$, and signals completion on $\tm{b'_i}$.
+Each agent signals their process to start on $\tm{a_i}$, waits for their process to finish on $\tm{b_i}$, and then signals for the next agent to continue on $\tm{c_i}$.
+The agent $\tm{\agent1}$ initiates, then waits for every other process to finish, and signals $\tm{\proc1}$ on $\tm{d}$.
+Every other agent $\tm{\agent{i}}$, $2\leq{i}\leq{n}$ waits on $\tm{c'_{i-1}}$ for the signal to start.
+Each of the channels in the scheduler is of a terminated type, and is merely used to synchronise.
+
+  Below is a diagram of our scheduler instantiated with three processes:
+  \begin{center}
+    \begin{tikzpicture}
+      \node[shape=circle,draw=black]                       (agent1) {\tm{\agent1}};
+      \node[shape=circle,draw=black,below right=of agent1] (agent2) {\tm{\agent2}};
+      \node[shape=circle,draw=black,below left =of agent1] (agent3) {\tm{\agent3}};
+
+      \node[shape=circle,draw=black,above      =of agent1] (proc1) {\tm{\proc1}};
+      \node[shape=circle,draw=black,below right=of agent2] (proc2) {\tm{\proc2}};
+      \node[shape=circle,draw=black,below left =of agent3] (proc3) {\tm{\proc3}};
+
+      \draw[->]
+      (agent1) -- node[pos=0.35,above] {$c_1$}
+      node[pos=1.0,above] {$c'_1$} ++ (agent2);
+      \draw[->]
+      (agent2) -- node[pos=0.1,above] {$c_2$}
+      node[pos=0.9,above] {$c'_2$} ++ (agent3);
+      \draw[->]
+      (agent3) -- node[pos=0.0,above] {$c_3$}
+      node[pos=0.65,above] {$c'_3$} ++ (agent1);
+
+      \draw[implies-implies,double]
+      (agent1) -- node[pos=0.15,left] {$a_1$} node[pos=0.85,left] {$a'_1$}
+      node[pos=0.15,right] {$b_1$} node[pos=0.85,right] {$b'_1$} ++ (proc1);
+      \draw[implies-implies,double]
+      (agent2) -- node[pos=0.15,above right] {$a_2$} node[pos=0.85,above right] {$a'_2$}
+      node[pos=0.15,below left] {$b_2$} node[pos=0.85,below left] {$b'_2$} ++ (proc2);
+      \draw[implies-implies,double]
+      (agent3) -- node[pos=0.15,above left] {$a_3$} node[pos=0.85,above left] {$a'_3$}
+      node[pos=0.15,below right] {$b_3$} node[pos=0.85,below right] {$b'_3$} ++ (proc3);
+
+      \draw[->]
+      (agent1) to[bend left=45] node[pos=0.15,left] {$d$} node[pos=0.85,left] {$d'$} (proc1);
+
+      \draw[->,dashed]
+      (proc1.east) to[bend left] node[pos=0.15,align=center,above right] {optional\\data transfer} (proc2.north);
+      \draw[->,dashed]
+      (proc2.west) to[bend left] (proc3.east);
+      \draw[->,dashed]
+      (proc3.north) to[bend left] (proc1.west);
+    \end{tikzpicture}
+  \end{center}
+
+    We implement the scheduler as follows, using $\tm{\prod_{I}P_i}$ to denote the parallel composition of the processes $\tm{P_i}$, $\tm{i}\in\tm{I}$, and $\tm{\plug{P}{Q}}$ to denote the plugging of $\tm{Q}$ in the one-hole process-context $\tm{P}$.
+The process-contexts $\tm{P_i}$ represent the computations performed by each process $\tm{\proc{i}}$.
+The process-contexts $\tm{Q_i}$ represent any post-processing, and any possible data transfer from $\tm{\proc{i}}$ to $\tm{\proc{i+1}}$.
+Finally, $\tm{Q_1}$ should contain $\tm{\labwait{d'}}$.
+\[
+  \begin{array}{lrlr}
+    \tm{\sched}
+    & \defeq & \tm{\res{a_1}{a'_1}{\dots{\res{a_n}{a'_n}{}}}}
+               \tm{\res{b_1}{b'_1}{\dots{\res{b_n}{b'_n}{}}}}
+               \tm{\res{c_1}{c'_1}{\dots{\res{c_n}{c'_n}{}}}}
+               \tm{\res{d}{d'}{}}
+    \\ &     & \tm{(
+               \ppar{\proc1}{\agent1}
+               \parallel
+               \prod_{2\leq{i}\leq{n}}(\ppar{\proc{i}}{\wait{c'_{i-1}}{\agent{i}}})
+               )}
+    \\
+    \tm{\agent1}
+    & \defeq & \tm{\close{a_i}{\wait{b_i}{\close{c_i}{\wait{c'_n}{\close{d}{\halt}}}}}}
+    \\
+    \tm{\agent{i}}
+    & \defeq & \tm{\close{a_i}{\wait{b_i}{\close{c_i}{\halt}}}}
+    \\
+    \tm{\proc{i}}
+    & \defeq & \tm{\wait{a'_i}{\plug{P_i}{\close{b'_i}{Q_i}}}}
+  \end{array}
+\]
+\end{example}
+\endgroup
+
+\begingroup
+\usingnamespace{pgv}
+\begin{example}[Milner's Cyclic Scheduler, PGV]
+  \label{ex:pgv-scheduler}
+  The PGV scheduler has exactly the same behaviour as the PCP version in~\cref{ex:pcp-scheduler}.
+It is implemented as follows, using $\tm{\prod_{I}\conf{C}_i}$ to denote the parallel composition of the processes $\tm{\conf{C}_i}$, $\tm{i}\in\tm{I}$, and $\tm{\plug{M}{N}}$ to denote the plugging of $\tm{N}$ in the one-hole term-context $\tm{M}$.
+For simplicity, we let $\tm{\sched}$ be a configuration.
+The terms $\tm{M_i}$ represent the computations performed by each process $\tm{\proc{i}}$.
+The terms $\tm{N_i}$ represent any post-processing, and any possible data transfer from $\tm{\proc{i}}$ to $\tm{\proc{i+1}}$.
+Finally, $\tm{N_1}$ should contain $\tm{\wait\;{d'}}$.
+  \[
+    \begin{array}{lrlr}
+      \tm{\sched}
+      & \defeq & \tm{\res{a_1}{a'_1}{\dots{\res{a_n}{a'_n}{}}}}
+                 \tm{\res{b_1}{b'_1}{\dots{\res{b_n}{b'_n}{}}}}
+                 \tm{\res{c_1}{c'_1}{\dots{\res{c_n}{c'_n}{}}}}
+                 \tm{\res{d}{d'}{}}
+      \\ &     & \tm{\begin{array}{lll}
+                       (
+                       & \phi\;\proc1
+                         \parallel
+                         \child\;\andthen{\agent1}{\andthen{\wait\;{c'_n}}{\close\;{d}}}
+                       \\
+                       \parallel
+                       &
+                         \prod_{2\leq{i}\leq{n}}
+                         (\child\;\proc{i} \parallel \child\;\andthen{\wait\;{c'_{i-1}}}{\agent{i}})
+                       & )
+                     \end{array}}
+      \\
+      \tm{\agent{i}}
+      & \defeq & \tm{\andthen{\close\;{a_i}}{\andthen{\wait\;{b_i}}{\close\;{c_i}}}}
+      \\
+      \tm{\proc{i}}
+      & \defeq & \tm{\andthen{\wait\;{a'_i}}{\plug{M_i}{\andthen{\close\;{b'_i}}{N_i}}}}
+    \end{array}
+  \]
+\end{example}
+\endgroup
+
+If $\pcp{\tm{\cpgvM{P_i}}}=\pgv{\tm{M_i}}$ and $\pcp{\tm{\cpgvM{Q_i}}}=\pgv{\tm{N_i}}$, then the translation of $\pcp{\tm{\sched}}$ (\cref{ex:pcp-scheduler}), $\pcp{\tm{\cpgvC{\sched\,}}}$, is exactly $\pgv{\tm{\sched}}$ (\cref{ex:pgv-scheduler}).
+
 %%% Local Variables:

### 2021-02-01 introduction.tex
on abstract and intro

@@
-% !TeX root = priorities.tex
 \documentclass[main.tex]{subfiles}
@@
 \section{Introduction}
-Session types are a type formalism which can be used to specify and verify communication protocols between two or more communicating agents~\cite{honda93,takeuchihonda94,hondavasconcelos98,carbonehonda07}.
+Session types are a type formalism used to specify and verify communication protocols between two or more communicating agents~\cite{honda93,takeuchihonda94,hondavasconcelos98,carbonehonda07}.
+They have been studied for many programming paradigms, including concurrent and functional paradigms.
+In particular, they have been defined for the $\pi$-calculus~\cite{sangiorgiwalker01}---a process calculus for communication and concurrency, and for Good Variation (GV)~\cite{wadler14,lindleymorris15}---a linear concurrent $\lambda$-calculus.
+Session type systems guarantee communication safety and session fidelity, namely that communication proceeds without any type mismatch and it follows the predefined protocol specified as a session type.
+Another key property in a concurrent setting is deadlock freedom, stating that agents communicate without getting stuck in cyclic dependencies.
+Deadlock freedom is a more involving property and session types alone are not enough to rule out deadlocks in cyclic structures of multiple interleaved sessions.
+Several techniques have been developed to address deadlock freedom in the $\pi$-calculus and concurrent $\lambda$-calculus with session types, as detailed below.
 
-Session types have been studied for many programming paradigms, most notably for concurrent and functional paradigms.
-They have been defined for the $\pi$-calculus~\cite{sangiorgiwalker01}---a process calculus for communication and concurrency, and for Good Variation (GV)~\cite{wadler14,lindleymorris15}---a linear concurrent $\lambda$-calculus.
-
-Session type systems guarantee communication safety and session fidelity, namely that communication proceeds without any type mismatch of the data being transmitted and follows the predefined protocol specified by a session type.
-Another key property in a concurrent setting is deadlock freedom, stating that agents communicate without getting stuck in cyclic dependencies.
-Unfortunately, deadlock freedom is a more involving property and session types alone are not enough to rule out deadlocks in cyclic structures of multiple interleaved sessions.
-Several techniques have been developed to address deadlock freedom in the $\pi$-calculus and concurrent $\lambda$-calculus with session types.
-
-With regards to $\pi$-calculus, a growing line of work on deadlock freedom leverages the Curry-Howard correspondences between intuitionistic or classical linear logic and session-typed $\pi$-calculus~\cite{cairespfenning10,wadler14}, which guarantee deadlock freedom by design.
+With regards to $\pi$-calculus, a growing line of work on deadlock freedom leverages the Curry-Howard correspondences between intuitionistic or classical linear logic and $\pi$-calculus with session types~\cite{cairespfenning10,wadler14}, which guarantee deadlock freedom by design.
 This is achieved by merging constructs for parallel composition and channel restriction under the logical cut rule, which forces processes to share only one channel for communication and eliminates cyclic dependencies.
@@
 Later on, calculi in the GV family~\cite{wadler14,lindleymorris15} have achieved deadlock freedom via a syntactic restriction, \ie by combining channel creation and thread spawning into a single operation, called fork.
-The fork construct is the term representation of the cut construct in CP, which evaluates to a channel restriction combined with a parallel composition.
-Unfortunately, as with CP, this is restrictive as it limits communication structures to trees.
-There are several benefits of working with a functional language as opposed to name-passing calculi.
+The fork construct is the term representation of the cut construct in CP, which as mentioned above, evaluates to a channel restriction combined with a parallel composition.
+As with CP, this is restrictive because it limits communication structures to only trees.
+
+We aim to further investigate GV and deadlock freedom, as there are several benefits of working with a functional language as opposed to name-passing calculi.
 Functional languages support higher-order functions, and have a capability for abstraction not usually present in process calculi.
@@
 Moreover, the benefit of working with GV over other session-typed functional languages is that GV has strong ties to linear logic, via its relation to CP~\cite{wadler14}, and consequently it has strong formal properties, \eg deadlock freedom, albeit in a restrictive way.
+Following the above observations, we thus pose our research question.
 
-Following the above observations, we aim to further investigate GV and deadlock freedom, and thus we pose our research question.
-
-\textbf{RQ:} \emph{Can we design a more expressive GV where deadlock freedom is guaranteed by design and communication structures are not limited to trees?}
+\textbf{RQ:} \emph{Can we design a more expressive GV where deadlock freedom is guaranteed by design and communication structures are not limited to only trees?}
 
@@
 Priority GV (PGV), a~variant of GV which decouples channel creation from thread spawning and restores deadlock freedom by adding priorities.
-This work closes the circle of the strong connection of CP and GV, together with their priority versions, PCP and PGV.
+This work closes the circle of the strong connection of CP and GV, together with their priority-based versions, PCP and PGV.
 
@@
 \item \textbf{Connection to linear logic}.
-The connection of PGV to logic is given via a \emph{sound and complete encoding} of PCP to PGV in~\cref{sec:pcp} and prove that it preserves typing (\cref{thm:pcp-to-pgv-confs-preservation}) and satisfies operational correspondence (\cref{thm:pcp-to-pgv-operational-correspondence-soundness,thm:pcp-to-pgv-operational-correspondence-completeness}).
+The connection of PGV to linear logic is given via a \emph{sound and complete encoding} of PCP to PGV in~\cref{sec:pcp}, where we prove the encoding preserves typing (\cref{thm:pcp-to-pgv-confs-preservation}) and satisfies operational correspondence (\cref{thm:pcp-to-pgv-operational-correspondence-soundness,thm:pcp-to-pgv-operational-correspondence-completeness}).
+We recall Milner's cyclic schedule from PCP~\cite{dardhagay18} and encode it in PGV.
 \end{enumerate}

### 2021-02-01 priority-gv.tex
some updates on pgv

@@
 
-We present Priority GV (PGV), a session-typed functional language based on GV~\cite{wadler15,lindleymorris15} which uses priorities \`{a} la Padovani~\cite{kobayashi06,padovaninovara15} to enforce deadlock freedom.
+We present Priority GV (PGV), a session-typed functional language based on GV~\cite{wadler15,lindleymorris15} which uses priorities \`{a} la Kobayashi and Padovani~\cite{kobayashi06,padovaninovara15} to enforce deadlock freedom.
 Priority GV offers a more fine-grained analysis of communication structures, and by separating channel creation form thread spawning it allows cyclic structures. 
-
 We illustrate this with two programs in PGV, \cref{ex:cycle,ex:deadlock}.
@@
 \end{enumerate*}
-We will show that only \cref{ex:cycle} is typeable in PGV, as opposed to neither in GV~\cite{wadler14} and both in its predecessor~\cite{gayvasconcelos10}.
+We will show that only \cref{ex:cycle} is typeable in PGV, while it is not typable in GV~\cite{wadler14} or its predecessor~\cite{gayvasconcelos10}.
 
@@
 \subsubsection*{Session types}
-Session ($\ty{S}$) types are defined by the following grammar:
+Session types ($\ty{S}$) are defined by the following grammar:
 \[
@@
 $\LabTirName{SC-ResExt}$ is the standard scope extrusion rule.
-Rules $\LabTirName{SC-ParNil}$, $\LabTirName{SC-ParComm}$ and $\LabTirName{SC-ParAssoc}$ state that parallel composition uses the terminated proceess as the neutral element; it is commutative and associative.
+Rules $\LabTirName{SC-ParNil}$, $\LabTirName{SC-ParComm}$ and $\LabTirName{SC-ParAssoc}$ state that parallel composition uses the terminated process as the neutral element; it is commutative and associative.
 
 While our configuration reduction is based on the standard evaluation for GV, the increased expressiveness of PGV allows us to simplify the relation on two counts.
-Firstly, we decompose the $\tm{\fork}$ construct.
+\begin{enumerate*}[label=(\alph*)]
+\item
+\emph{We decompose the $\tm{\fork}$ construct}.
 In GV, $\tm{\fork}$ creates a new channel, spawns a child thread, and, when the child thread finishes, it closes the channel to its parent.
@@
 In PGV, evaluating $\tm{\link}$ simply causes a substitution.
-
-Secondly, our structural congruence is type preserving.
+\item
+\emph{Our structural congruence is type preserving}.
 Consequently, we can embed it directly into the reduction relation.
 In GV, this is not the case, and subject reduction relies on proving that if $\equiv\cred$ ends up in an ill-typed configuration, we can rewrite it to a well-typed configuration using $\equiv$.
+\end{enumerate*}
 

### 2021-02-01 priority-gv.tex
pass on typing

@@
 Together, these two constraints guarantee deadlock freedom.
-For instance, the program in \cref{ex:deadlock} is deadlocked.
+
+To illustrate this, let's go back to the deadlocked program in \cref{ex:deadlock}.
 Crucially, it composes the terms below in parallel.
@@
 \begin{enumerate*}[label=(\roman*)]
-\item we require that child threads return the unit value, as opposed to a terminated channel; and
-\item we simplify typing for parallel composition.
+\item \emph{we require that child threads return the unit value}, as opposed to a terminated channel; and
+\item \emph{we simplify typing for parallel composition}.
 \end{enumerate*}
@@
 Consequently, associativity of parallel composition does not preserve typing.
-In PGV, we guarantee deadlock freedom using priorities, which removes the channel pseudo-type $\ty{S^\sharp}$, simplifies typing for parallel composition, and restores type preservation for the structural congruence.
+In PGV, we guarantee deadlock freedom using priorities, which removes the need for the channel pseudo-type $\ty{S^\sharp}$, and simplifies typing for parallel composition, while restoring type preservation for the structural congruence.
 
@@
 For this, we prove~\cref{lem:pgv-value-done}, which shows that values have finished all their communication, and that any priorities in the type of the value come from the typing environment.
-The proofs are standard inductions and can be found in \cref{prf:lem-pgv-value-done,prf:lem-pgv-substitution,prf:lem-pgv-subject-reduction-terms,prf:lem-pgv-subject-congruence,prf:thm-pgv-subject-reduction-confs}.
+The proofs are by structural induction and can be found in \cref{prf:lem-pgv-value-done,prf:lem-pgv-substitution,prf:lem-pgv-subject-reduction-terms,prf:lem-pgv-subject-congruence,prf:thm-pgv-subject-reduction-confs}.
 \begin{compacttheorems}
@@
   \item $\tm{M}$ is a value;
-  \item $\tm{M}\tred\tm{N}$ for some term $\tm{N}$; or
+  \item $\tm{M}\tred\tm{N}$ for some $\tm{N}$; or
   \item $\tm{M}$ is ready.
@@
 \end{compacttheorems}
-It follows from \cref{thm:pgv-closed-progress-confs} and \cref{cor:pgv-closed-progress} that a term which does not return an endpoint will complete all its communication actions, thus satisfying deadlock freedom.
+It follows immediately from \cref{thm:pgv-closed-progress-confs} and \cref{cor:pgv-closed-progress} that a term which does not return an endpoint will complete all its communication actions, thus satisfying deadlock freedom.
 \end{document}

### 2021-02-01 priority-gv.tex
polishing

@@
     \label{cor:pgv-closed-progress}
-    If $\cseq[\main]{\emptyenv}{\conf{C}}$, $\tm{\conf{C}}\centernot\cred$, and $\tm{\conf{C}}$ contains no endpoints, then $\tm{\conf{C}}\equiv\tm{\phi\;V}$ for some value V.
+    If $\cseq[\main]{\emptyenv}{\conf{C}}$, $\tm{\conf{C}}\centernot\cred$, and $\tm{\conf{C}}$ contains no endpoints, then $\tm{\conf{C}}\equiv\tm{\phi\;V}$ for some value $\tm{V}$.
   \end{corollary}

### 2021-02-02 priority-cp.tex
Minor layout changes.
started on PCP

@@
 We present a correspondence between Priority GV and an updated version of Priority CP~\cite[PCP]{dardhagay18}, which is Wadler's CP~\cite{wadler14} with priorities.
-The correspondence connects PGV to (a~relaxed variant of) classical linear logic.
+This correspondence connects PGV to (a~relaxed variant of) classical linear logic.
 
@@
 \usingnamespace{pcp}
-Types ($\ty{A}$, $\ty{B}$) in PCP correspond to the connectives of linear logic annotated with priorities $\cs{o}\in\mathbb{N}$.
-Typing environments, duality, and the priority function $\pr$ are defined as expected (see~\cref{sec:pcp-environments,sec:pcp-duality,sec:pcp-priorities}).
+\textbf{Types} ($\ty{A}$, $\ty{B}$) in PCP correspond to linear logic connectives annotated with priorities $\cs{o}\in\mathbb{N}$.
+Typing environments, duality, and the priority function $\pr(\cdot)$ are defined as expected (see~\cref{sec:pcp-environments,sec:pcp-duality,sec:pcp-priorities}).
 \[
@@
 \]
-Processes ($\tm{P}$, $\tm{Q}$) in PCP are defined by the following grammar.
+\textbf{Processes} ($\tm{P}$, $\tm{Q}$) in PCP are defined by the following grammar.
 \[
@@
 \]
-Processes are typed by sequents $\seq{P}{\ty{\Gamma}}$, which correspond to the one-sided sequents classical linear logic.
-Different from PGV, we do not need to store the greatest priority on the sequent, as, due to the absence of higher-order functions, we cannot compose processes \emph{sequentially}.
+Processes are typed by sequents $\seq{P}{\ty{\Gamma}}$, which correspond to the one-sided sequents in classical linear logic.
+Differently from PGV, in PCP we do not need to store the greatest priority on the sequent, as, due to the absence of higher-order functions, we cannot compose processes \emph{sequentially}.
 
-PCP decomposes cut into \LabTirName{T-Res} and \LabTirName{T-Par}, and guarantees deadlock freedom using priority constraints, \eg, in \LabTirName{T-Send}.
-The full typing rules and operational semantics are available in \cref{fig:pcp-operational-semantics-and-typing} and in \cite{dardhagay18}:
+PCP decomposes cut into \LabTirName{T-Res} and \LabTirName{T-Par} rules---logically corresponding to cycle and mix rules, respectively---and guarantees deadlock freedom by using priority constraints, \eg, as in \LabTirName{T-Send}.
+The full typing rules and operational semantics are available in \cref{fig:pcp-operational-semantics,fig:pcp-typing} in \cref{app:revisiting-PCP} and in \cite{dardhagay18}:
 \begin{mathpar}
+  \small
   \inferrule*[lab=T-Res]{
@@
   }{\seq{\res{x}{y}{P}}{\ty{\Gamma}}}
-
   \inferrule*[lab=T-Par]{
     \seq{P}{\ty{\Gamma}}
-    \\
+    \and
     \seq{Q}{\ty{\Delta}}
   }{\seq{\ppar{P}{Q}}{\ty{\Gamma},\ty{\Delta}}}
-
   \inferrule*[lab=T-Send]{
     \seq{P}{\ty{\Gamma},\tmty{y}{A},\tmty{x}{B}}
-    \\
+    \and
     \cs{o}<\minpr(\ty{\Gamma},\ty{A},\ty{B})
@@
 Commuting conversions are necessary if we want our reduction strategy to correspond \emph{exactly} to cut (or cycle in \cite{dardhagay18}) elimination.
-However, as Lindley and Morris~\cite{lindleymorris15} show, all communications that can be performed \emph{with} the use of commuting conversions, can also be performed \emph{without} them, using structural congruence.
+However, as Lindley and Morris~\cite{lindleymorris15} show, all communications that can be performed \emph{with} the use of commuting conversions, can also be performed \emph{without} them, but using structural congruence.
 

### 2021-02-02 appendix.tex
started on PCP
a pass on commuting conversions

@@
 \subsection{Revisiting Priority CP}
+\label{app:revisiting-PCP}
 
@@
 
+\subsubsection*{On Commuting Conversions}
+\label{app:commuting-conversions}
+
+The main change we make to PCP is \emph{removing commuting conversions}.
+Commuting conversions are necessary if we want our reduction strategy to correspond \emph{exactly} to cut (or cycle in \cite{dardhagay18}) elimination.
+However, as Lindley and Morris~\cite{lindleymorris15} show, all communications that can be performed \emph{with} the use of commuting conversions, can also be performed \emph{without} them, but using structural congruence.
+
+From the perspective of process calculi, commuting conversions behave strangely. 
+Consider the commuting conversion $(\kappa_{\parr})$ for $\pcp{\tm{\recv{x}{y}{P}}}$:
+\begin{mathpar}
+  (\kappa_{\parr})
+  \quad
+  \tm{\res{z}{z'}{(\ppar{\recv{x}{y}{P}}{Q})}}
+  \red
+  \tm{\recv{x}{y}{\res{z}{z'}{(\ppar{P}{Q})}}}
+\end{mathpar}
+As a result of $(\kappa_{\parr})$, $\pcp{\tm{Q}}$ becomes blocked on $\pcp{\tm{\labrecv{x}{y}}}$, and any actions $\pcp{\tm{Q}}$ was able to perform become unavailable.
+Consequently, CP is non-confluent:
+\begin{mathpar}
+  \setlength{\arraycolsep}{2em}
+  \begin{array}{cc}
+    \multicolumn{2}{c}{%
+    \hspace*{10ex}
+    {\tm{\res{x}{x'}{(\ppar{\recv{a}{y}{P}}{\res{z}{z'}{(\ppar{\close{z}{\halt}}{\wait{z'}{Q}})}})}}}}
+    \\
+    \qquad\rotatebox[origin=c]{270}{$\red\hphantom{{}^+}$}
+    &
+    \rotatebox[origin=c]{270}{$\red^+$}\qquad
+    \\
+    {\tm{\recv{a}{y}{\res{x}{x'}{(\ppar{P}{\res{z}{z'}{(\ppar{\close{z}{\halt}}{\wait{z'}{Q}})}})}}}}
+    &
+    {\tm{\recv{a}{y}{\res{x}{x'}{(\ppar{P}{Q})}}}}
+  \end{array}
+\end{mathpar}
+
+In PCP, commuting conversions break our intuition that an action with lower priority occurs before an action with higher priority.
+To cite Dardha and Gay~\cite{dardhagay18} ``\emph{if a prefix on a channel endpoint $\pcp{\tm{x}}$ with priority $\cs{o}$ is pulled out at top level, then to preserve priority constraints in the typing rules [..], it is necessary to increase priorities of all actions after the prefix on $\pcp{\tm{x}}$}'' by $\cs{o+1}$.
+
 \subsubsection*{Operational Semantics}
@@
     \label{def:pcp-canonical-forms}
-    A~process $\tm{P}$ is in canonical form if it is $\tm{\halt}$ or $\tm{\res{x_1}{x'_1}{\dots\res{x_n}{x'_n}{(P_1 \parallel\dots\parallel P_m)}}}$ where $m>0$ and each $\tm{P_j}$ is an action.
+    A~process $\tm{P}$ is in canonical form if it is either $\tm{\halt}$ or of the form $\tm{\res{x_1}{x'_1}{\dots\res{x_n}{x'_n}{(P_1 \parallel\dots\parallel P_m)}}}$ where $m>0$ and each $\tm{P_j}$ is an action.
   \end{restatabledefinition}

### 2021-02-02 priority-cp.tex
a pass on commuting conversions

@@
 PCP decomposes cut into \LabTirName{T-Res} and \LabTirName{T-Par} rules---logically corresponding to cycle and mix rules, respectively---and guarantees deadlock freedom by using priority constraints, \eg, as in \LabTirName{T-Send}.
-The full typing rules and operational semantics are available in \cref{fig:pcp-operational-semantics,fig:pcp-typing} in \cref{app:revisiting-PCP} and in \cite{dardhagay18}:
+The full typing rules and operational semantics are available in \cref{fig:pcp-operational-semantics,fig:pcp-typing} in \cref{app:revisiting-PCP}:
 \begin{mathpar}
@@
 
-The main change we make to PCP is \emph{removing commuting conversions}.
+The main change we make to PCP is \emph{removing commuting conversions} and defining an operational semantics based on structural congruence.
 Commuting conversions are necessary if we want our reduction strategy to correspond \emph{exactly} to cut (or cycle in \cite{dardhagay18}) elimination.
-However, as Lindley and Morris~\cite{lindleymorris15} show, all communications that can be performed \emph{with} the use of commuting conversions, can also be performed \emph{without} them, but using structural congruence.
+However, from the perspective of process calculi, commuting conversions behave strangely: they allow an input/output action to be moved to the top of a process, thus potentially blocking actions which were previously possible (an example is given in \cref{app:commuting-conversions}).
+This makes CP, and PCP in \cite{dardhagay18}, non-confluent.
+As Lindley and Morris~\cite{lindleymorris15} show, all communications that can be performed \emph{with} the use of commuting conversions, can also be performed \emph{without} them, if using structural congruence.
+%Consider the commuting conversion $(\kappa_{\parr})$ for $\pcp{\tm{\recv{x}{y}{P}}}$:
+%\begin{mathpar}
+%  (\kappa_{\parr})
+%  \quad
+%  \tm{\res{z}{z'}{(\ppar{\recv{x}{y}{P}}{Q})}}
+%  \red
+%  \tm{\recv{x}{y}{\res{z}{z'}{(\ppar{P}{Q})}}}
+%\end{mathpar}
+%As a result of $(\kappa_{\parr})$, $\pcp{\tm{Q}}$ becomes blocked on $\pcp{\tm{\labrecv{x}{y}}}$, and any actions $\pcp{\tm{Q}}$ was able to perform become unavailable.
+Consequently, CP is non-confluent:
+%\begin{mathpar}
+%  \setlength{\arraycolsep}{2em}
+%  \begin{array}{cc}
+%    \multicolumn{2}{c}{%
+%    \hspace*{10ex}
+%    {\tm{\res{x}{x'}{(\ppar{\recv{a}{y}{P}}{\res{z}{z'}{(\ppar{\close{z}{\halt}}{\wait{z'}{Q}})}})}}}}
+%    \\
+%    \qquad\rotatebox[origin=c]{270}{$\red\hphantom{{}^+}$}
+%    &
+%    \rotatebox[origin=c]{270}{$\red^+$}\qquad
+%    \\
+%    {\tm{\recv{a}{y}{\res{x}{x'}{(\ppar{P}{\res{z}{z'}{(\ppar{\close{z}{\halt}}{\wait{z'}{Q}})}})}}}}
+%    &
+%    {\tm{\recv{a}{y}{\res{x}{x'}{(\ppar{P}{Q})}}}}
+%  \end{array}
+%\end{mathpar}
+%
 
-From the perspective of process calculi, commuting conversions behave strangely. Consider the commuting conversion $(\kappa_{\parr})$ for $\pcp{\tm{\recv{x}{y}{P}}}$:
-\begin{mathpar}
-  (\kappa_{\parr})
-  \quad
-  \tm{\res{z}{z'}{(\ppar{\recv{x}{y}{P}}{Q})}}
-  \red
-  \tm{\recv{x}{y}{\res{z}{z'}{(\ppar{P}{Q})}}}
-\end{mathpar}
-As a result of $(\kappa_{\parr})$, $\pcp{\tm{Q}}$ becomes blocked on $\pcp{\tm{\labrecv{x}{y}}}$, and any actions $\pcp{\tm{Q}}$ was able to perform become unavailable.
-Consequently, CP is non-confluent:
-\begin{mathpar}
-  \setlength{\arraycolsep}{2em}
-  \begin{array}{cc}
-    \multicolumn{2}{c}{%
-    \hspace*{10ex}
-    {\tm{\res{x}{x'}{(\ppar{\recv{a}{y}{P}}{\res{z}{z'}{(\ppar{\close{z}{\halt}}{\wait{z'}{Q}})}})}}}}
-    \\
-    \qquad\rotatebox[origin=c]{270}{$\red\hphantom{{}^+}$}
-    &
-    \rotatebox[origin=c]{270}{$\red^+$}\qquad
-    \\
-    {\tm{\recv{a}{y}{\res{x}{x'}{(\ppar{P}{\res{z}{z'}{(\ppar{\close{z}{\halt}}{\wait{z'}{Q}})}})}}}}
-    &
-    {\tm{\recv{a}{y}{\res{x}{x'}{(\ppar{P}{Q})}}}}
-  \end{array}
-\end{mathpar}
-
-In PCP, commuting conversions break our intuition that an action with lower priority occurs before an action with higher priority.
+In  particular for PCP, commuting conversions break our intuition that an action with lower priority \emph{occurs before} an action with higher priority.
 To cite Dardha and Gay~\cite{dardhagay18} ``\emph{if a prefix on a channel endpoint $\pcp{\tm{x}}$ with priority $\cs{o}$ is pulled out at top level, then to preserve priority constraints in the typing rules [..], it is necessary to increase priorities of all actions after the prefix on $\pcp{\tm{x}}$}'' by $\cs{o+1}$.
-
 One benefit of removing commuting conversions is that we no longer need to dynamically change the priorities during reduction, which means that the intuition for priorities holds true in our updated version of PCP.

### 2021-02-02 priority-cp.tex
on the cyclic scheduler

@@
 \begingroup
-We illustrate the relation between PCP and PGV by giving a translation from PCP to PGV. The translation on types is defined as follows:
+We illustrate the relation between PCP and PGV by defining a translation from PCP to PGV. The translation on types is defined as follows:
 \[
@@
 The translations $\tm{\cpgvM{\cdot}}$ and $\tm{\cpgvC{\cdot}}$ preserve typing, and the latter induces a sound and complete operational correspondence.
-The proofs are standard inductions and can be found in~\cref{prf:lem-pcp-to-pgv-terms-preservation,prf:lem-pcp-to-pgv-terms-preservation,prf:lem-pcp-to-pgv-cpgvM-to-cpgvC,prf:thm-pcp-to-pgv-operational-correspondence-soundness,prf:thm-pcp-to-pgv-operational-correspondence-completeness}.
+The proofs are by structural induction and can be found in~\cref{prf:lem-pcp-to-pgv-terms-preservation,prf:lem-pcp-to-pgv-terms-preservation,prf:lem-pcp-to-pgv-cpgvM-to-cpgvC,prf:thm-pcp-to-pgv-operational-correspondence-soundness,prf:thm-pcp-to-pgv-operational-correspondence-completeness}.
 \begin{compacttheorems}
@@
 \endgroup
+
+\noindent
+\textbf{Milner's Cyclic Scheduler}.
+As an example of a deadlock-free cyclic process, Dardha and Gay~\cite{dardhagay18} introduce an implementation of Milner's cyclic scheduler~\cite{milner89} in Priority CP.
+We have reproduced the scheduler in our updated PCP, and showed its translation to PGV.
+Full details are given in \cref{sec:milner}.
+
 \end{document}

### 2021-02-02 conclusion.tex
conclusions and 15!

@@
 \subsubsection*{Deadlock freedom and progress}
-Deadlock freedom and progress are well studied properties in the $\pi$-calculus.
+%Deadlock freedom and progress are well studied properties in the $\pi$-calculus.
 For the standard typed $\pi$-calculus, one line of work is Kobayashi's approach to deadlock freedom~\cite{kobayashi98}, where priorities are abstract tags defined over a partially ordered set.
-These abstract tags were later simplified to natural numbers.
-Pairs of obligations and capabilities were used in the type system for deadlock freedom~\cite{kobayashi06}, which allowed more $\pi$-calculus processes to be typed.
-Padovani~\cite{padovani13} adapted the obligation and capability pairs to session types, and later simplified them to a single priority for linear $\pi$-calculus~\cite{padovani14}.
-As discussed by the author, by combining the encoding of session types into linear types~\cite{kobayashi07,dardhagiachino12,dardha14beat,dardha16} and the priority-based linear $\pi$-calculus, we can obtain deadlock freedom for the $\pi$-calculus with session types.
+These tags were later simplified to pairs of natural numbers, called obligations and capabilities~\cite{kobayashi06}, which allowed more $\pi$-calculus processes to be typed.
+Padovani~\cite{padovani13} adapted the obligation and capability pairs to session types, and later simplified them to a single priority for the linear $\pi$-calculus~\cite{padovani14}.
+Furthermore, by exploiting the encoding of session types into linear types~\cite{kobayashi07,dardhagiachino12}
+%,dardha14beat,dardha16 %
+and the priority-based linear $\pi$-calculus,
+we can obtain deadlock freedom for the $\pi$-calculus with session types.
 
-For the session-typed $\pi$-calculus, the foundational work by Dezani~\etal~\cite{dezani-ciancaglinimostrous06} guarantees progress by allowing only one active session at a time.
-Dezani later~\cite{dezani-ciancagliniliguoro09progress} introduces a partial order on channels in line with Kobayashi's work~\cite{kobayashi98}.
-Carbone and Debois~\cite{carbonedebois10} define progress for session typed $\pi$-calculus in terms of a \emph{catalyser} used to provide a missing counterpart to a process, thus guaranteeing deadlock freedom.
-Carbone~\etal~\cite{carbonedardha14} studied the use of catalysers and showed that progress is a compositional form of lock-freedom.
-In the same lines of Padovani~\cite{padovani14}, Carbone~\etal~\cite{carbonedardha14} show that progress can be lifted to session types via the encoding of session types to linear types and Kobayashi's obligations/capabilities.
+For the session-typed $\pi$-calculus, Dezani~\etal~\cite{dezani-ciancaglinimostrous06} guarantee progress by allowing only one active session at a time.
+Dezani later~\cite{dezani-ciancagliniliguoro09progress} introduces a partial order on channels similar to Kobayashi~\cite{kobayashi98}.
+Carbone and Debois~\cite{carbonedebois10} define progress for session typed $\pi$-calculus in terms of a \emph{catalyser}, which provides a missing counterpart to a process, thus guaranteeing deadlock freedom.
+Carbone~\etal~\cite{carbonedardha14} use such catalysers to show that progress is a compositional form of lock-freedom and that it can be lifted to session types via the encoding of session types to linear types.
 Vieira and Vasconcelos~\cite{vieiravasconcelos13} use single priorities and an abstract partial order to guarantee deadlock freedom in a session-typed $\pi$-calculus.
-
-Multiparty Session Types (MPST) \cite{hondayoshida08} on the other hand guarantee deadlock freedom within a single session, but more advanced techniques are required for session interleaving.
-Bettini~\etal~\cite{bettinicoppo08} follow a similar technique to Kobayashi for multiparty session types.
-The main
-difference with our work is that we associate priorities to communication actions, and not to channels.
+%
+Multiparty Session Types (MPST) \cite{hondayoshida08} guarantee deadlock freedom \emph{within a single} session, but not for session interleaving.
+Consequently, several techniques for deadlock freedom in MPST have been defined \cite{bettinicoppo08,carbonemontesi13}.
+%Bettini~\etal~\cite{bettinicoppo08} follow a similar technique to Kobayashi for multiparty session types. The main
+%difference with our work is that we associate priorities to communication actions, and not to channels.
 Carbone and Montesi~\cite{carbonemontesi13} explore and combine MPST and choreographies to obtain a formalism that satisfies deadlock freedom by design.
@@
 A~correspondence between session-typed $\pi$-calculus and dual intuitionistic linear logic was developed by Caires and Pfenning~\cite[$\pi\text{DILL}$]{cairespfenning10}, and with classical linear logic by Wadler~\cite[CP]{wadler15}, both guaranteeing deadlock freedom as a result of their connection to logic.
-Dardha and Gay~\cite[PCP]{dardhagay18} integrate Kobayashi's and Padovani's work on priorities~\cite{kobayashi06,padovani14} with CP, creating the first calculus which combines priorities and strong ties with logic.
+Dardha and Gay~\cite[PCP]{dardhagay18} integrate Kobayashi and Padovani's work on priorities~\cite{kobayashi06,padovani14} with CP, creating the first calculus which combines priorities and strong ties with logic.
 Dardha and P\'{e}rez~\cite{dardhaperez15} compare Kobayashi-style typing and CLL typing, and show that CLL corresponds to a subsystem of Kobayashi's where restriction is applied once.
@@
 \subsubsection*{Conclusion and future work}
-We presented Priority GV, a~session-typed functional language which uses priorities to ensure deadlock freedom, and showed its relation to Priority CP~\cite{dardhagay18} via an operational correspondence.
+We answered our research question by presenting Priority GV, a~session-typed functional language which allows cyclic communication structures and uses priorities to ensure deadlock freedom.
+We showed its relation to Priority CP~\cite{dardhagay18} via an operational correspondence.
 
 Our formalism so far only captures the core of GV.
-In future work, we plan to explore: recursion in PGV, by integrating the works of Lindley and Morris~\cite{lindleymorris16} and Padovani and Novara~\cite{padovaninovara15}; a~translation from Priority GV to Priority CP; and sharing, following Balzer and Pfenning~\cite{balzerpfenning17}.
+In future work, we plan to explore: recursion in PGV, by integrating the works of Lindley and Morris~\cite{lindleymorris16} and Padovani and Novara~\cite{padovaninovara15}; %a~translation from Priority GV to Priority CP;
+and sharing, following Balzer and Pfenning~\cite{balzerpfenning17}.
+
+\subsubsection*{Acknowledgement} The authors would like to thank Simon Fowler and Philip Wadler for they useful and insightful feedback.
+
 \end{document}

### 2021-04-07 introduction.tex
polishing
Fixed small bugs. Went over reviews.

@@
 This is achieved by merging constructs for parallel composition and channel restriction under the logical cut rule, which forces processes to share only one channel for communication and eliminates cyclic dependencies.
-Recent developments in the Classical Process (CP)~\cite{wadler14} line of work have led to various approaches to decoupling these constructs, either by maintaining a strong correspondence to classical linear logic, as in Hypersequent CP~\cite{kokkemontesi19popl,kokkemontesi19tlla}, or by weakening the correspondence to classical linear logic in exchange for a more expressive language, as in Priority CP (PCP)~\cite{dardhagay18}.
+Recent developments in the Classical Processes (CP)~\cite{wadler14} line of work have led to various approaches to decoupling these constructs, either by maintaining a strong correspondence to classical linear logic, as in Hypersequent CP~\cite{kokkemontesi19popl,kokkemontesi19tlla}, or by weakening the correspondence to classical linear logic in exchange for a more expressive language, as in Priority CP (PCP)~\cite{dardhagay18}.
 PCP decouples CP's cut rule into two separate constructs: a construct for parallel composition, by introducing a mix rule; and a construct for restriction, by introducing a cycle rule.
@@
 The connection of PGV to linear logic is given via a \emph{sound and complete encoding} of PCP to PGV in~\cref{sec:pcp}, where we prove the encoding preserves typing (\cref{thm:pcp-to-pgv-confs-preservation}) and satisfies operational correspondence (\cref{thm:pcp-to-pgv-operational-correspondence-soundness,thm:pcp-to-pgv-operational-correspondence-completeness}).
-We recall Milner's cyclic schedule from PCP~\cite{dardhagay18} and encode it in PGV.
+We recall Milner's cyclic scheduler~\cite{milner89} presented in PCP~\cite{dardhagay18} and encode it in PGV.
 \end{enumerate}

### 2021-04-07 priority-gv.tex
Fixed small bugs. Went over reviews.

@@
 \end{enumerate*}
-We will show that only \cref{ex:cycle} is typeable in PGV, while it is not typable in GV~\cite{wadler14} or its predecessor~\cite{gayvasconcelos10}.
+We will show that only \cref{ex:cycle} is typeable in PGV, while it is not typable in GV~\cite{wadler14} nor guaranteed to be deadlock-free by its predecessor~\cite{gayvasconcelos10}.
 
@@
   \qquad
-  \ty{\co{\tyends[\cs{o}]}} = \ty{\co{\tyendr[\cs{o}]}}
+  \ty{\co{\tyends[\cs{o}]}} = \ty{\tyendr[\cs{o}]}
   \qquad
-  \ty{\co{\tyendr[\cs{o}]}} = \ty{\co{\tyends[\cs{o}]}}
+  \ty{\co{\tyendr[\cs{o}]}} = \ty{\tyends[\cs{o}]}
 \]
@@
 Main threads are allowed to return a value, whereas child threads are required to return the unit value.
-Sequents are annotated with a flag $\phi$, which ensure that there is at most one main thread.
+Sequents are annotated with a flag $\phi$, which ensures that there is at most one main thread.
 
@@
 Clearly, no more evaluation is possible, even though the configuration contains the thread $\tm{\child\;\echo_{x'}}$, which is blocked on $\tm{x'}$.
-In~\cref{cor:pgv-closed-progress} we show that if a term does not return an endpoint, it must produce \emph{only} a value.
+In~\cref{cor:pgv-closed-progress} we will show that if a term does not return an endpoint, it must produce \emph{only} a value.
 
@@
   \label{def:pgv-ready-actions}
-  A~term $\tm{L}$ is \emph{ready} if it is of the form $\tm{\plug{E}{M}}$, where $\tm{M}$ is of the form $\tm{\new}$, $\tm{\spawn\;N}$, $\tm{\link\;\pair{x}{y}}$ or $\tm{\link\;\pair{y}{x}}$, or $\tm{M}$ acts on $\tm{x}$.
+  A~term $\tm{L}$ is \emph{ready} if it is of the form $\tm{\plug{E}{M}}$, where $\tm{M}$ is of the form $\tm{\new}$, $\tm{\spawn\;N}$, $\tm{\link\;\pair{x}{y}}$, or $\tm{M}$ acts on $\tm{x}$.
 In the latter case, we say that $\tm{L}$ is \emph{ready to act on} $\tm{x}$.

### 2021-04-08 priority-cp.tex
a quick pass to the paper

@@
 
-\noindent
-\textbf{Milner's Cyclic Scheduler}.
-As an example of a deadlock-free cyclic process, Dardha and Gay~\cite{dardhagay18} introduce an implementation of Milner's cyclic scheduler~\cite{milner89} in Priority CP.
+%\noindent
+%\textbf{Milner's Cyclic Scheduler}.
+%As an example of a deadlock-free cyclic process, Dardha and Gay~\cite{dardhagay18} introduce an implementation of Milner's cyclic scheduler~\cite{milner89} in Priority CP.
 We have reproduced the scheduler in our updated PCP, and showed its translation to PGV.

### 2021-04-08 conclusion.tex
Fixed small bugs. Went over reviews.
Update conclusion.tex

@@
 \subsubsection*{Deadlock freedom and progress}
-%Deadlock freedom and progress are well studied properties in the $\pi$-calculus.
-For the standard typed $\pi$-calculus, one line of work is Kobayashi's approach to deadlock freedom~\cite{kobayashi98}, where priorities are abstract tags defined over a partially ordered set.
+Deadlock freedom and progress are well studied properties in the $\pi$-calculus.
+For the `standard' typed $\pi$-calculus, one line of work is Kobayashi's approach to deadlock freedom~\cite{kobayashi98}, where priorities are abstract tags defined over a partially ordered set.
 These tags were later simplified to pairs of natural numbers, called obligations and capabilities~\cite{kobayashi06}, which allowed more $\pi$-calculus processes to be typed.
 Padovani~\cite{padovani13} adapted the obligation and capability pairs to session types, and later simplified them to a single priority for the linear $\pi$-calculus~\cite{padovani14}.
-Furthermore, by exploiting the encoding of session types into linear types~\cite{kobayashi07,dardhagiachino12}
-%,dardha14beat,dardha16 %
-and the priority-based linear $\pi$-calculus,
+Furthermore, by exploiting the encoding of session types into linear types~\cite{kobayashi07,dardhagiachino12,dardha14beat} and the priority-based linear $\pi$-calculus,
 we can obtain deadlock freedom for the $\pi$-calculus with session types.
@@
 Carbone~\etal~\cite{carbonedardha14} use such catalysers to show that progress is a compositional form of lock-freedom and that it can be lifted to session types via the encoding of session types to linear types.
-Vieira and Vasconcelos~\cite{vieiravasconcelos13} use single priorities and an abstract partial order to guarantee deadlock freedom in a session-typed $\pi$-calculus.
-%
-Multiparty Session Types (MPST) \cite{hondayoshida08} guarantee deadlock freedom \emph{within a single} session, but not for session interleaving.
-Consequently, several techniques for deadlock freedom in MPST have been defined \cite{bettinicoppo08,carbonemontesi13}.
-%Bettini~\etal~\cite{bettinicoppo08} follow a similar technique to Kobayashi for multiparty session types. The main
-%difference with our work is that we associate priorities to communication actions, and not to channels.
+Vieira and Vasconcelos~\cite{vieiravasconcelos13} use single priorities and an abstract partial order to guarantee deadlock freedom in a binary session-typed $\pi$-calculus and building on conservation types.
+
+While our work focuses on \emph{binary} session types, it is worth to discuss related work on Multiparty Session Types (MPST) in order to give a broader context.
+The original work on MPST by Honda~\etal~\cite{hondayoshida08} guarantees deadlock freedom \emph{within a single} session, but not for session interleaving.
+Later on, several techniques for deadlock freedom in MPST were defined.
+Bettini~\etal~\cite{bettinicoppo08} follow a similar technique to Kobayashi's for the MPST setting.
+The main difference with our work is that we associate priorities with communication actions, whether Bettini~\etal~\cite{bettinicoppo08} with channels.
 Carbone and Montesi~\cite{carbonemontesi13} explore and combine MPST and choreographies to obtain a formalism that satisfies deadlock freedom by design.
+Castellani~\etal~\cite{CastellaniDGH20} guarantee lock freedom--a stronger property than deadlock freedom--in the context of MPST with \emph{internal delegation}, where participants are allowed to delegate tasks to each other as long as they fall within the same multiparty session.
+This allows for the internal delegation to be captured by the global type.
 
@@
 
-\subsubsection*{Acknowledgement} The authors would like to thank Simon Fowler and Philip Wadler for they useful and insightful feedback.
+\subsubsection*{Acknowledgement}
+The authors would like to thank Simon Fowler, April Gon\c{c}alves, and Philip Wadler for their comments on the manuscript.
 

### 2021-04-08 conclusion.tex
some more related work

@@
 Carbone and Montesi~\cite{carbonemontesi13} explore and combine MPST and choreographies to obtain a formalism that satisfies deadlock freedom by design.
+Deni{\'{e}}lou and Yoshida \cite{DenielouY13} introduce the notion of \emph{multiparty compatibility} which generalises that of duality in binary session types and they synthesise safe and deadlock-free global types from local session types via the use of LTSs and communicating automata (CFSA).
 Castellani~\etal~\cite{CastellaniDGH20} guarantee lock freedom--a stronger property than deadlock freedom--in the context of MPST with \emph{internal delegation}, where participants are allowed to delegate tasks to each other as long as they fall within the same multiparty session.
 This allows for the internal delegation to be captured by the global type.
+Scalas and Yoshida \cite{scalasyoshida19} provide a revision of the theoretical foundations of MPST leading to a less complicated and more general theory, by completely removing the notion of duality/consistency.
+The type systems is parametric and ensures decidability of type checking, while allowing for a novel integration of model checking techniques in MPST.
+In this new theory, more protocols and processes can be typed that are guaranteed to be free of deadlocks.
 
@@
 They later restore deadlock freedom using priorities~\cite[$\text{SILL}_{S+}$]{balzertoninho19}.
-Carbone~\etal~\cite{carbonelindley16} give a logical view of MPST with a generalised duality, called coherence.
+Carbone~\etal~\cite{CarboneMSY15,carbonelindley16} give a logical view of MPST with a generalised duality, called \emph{coherence}.
 

### 2021-04-09 conclusion.tex
references on horne
more on related work

@@
 Since its inception by Girard~\cite{girard87}, linear logic has been a candidate for a foundational correspondence for concurrent programs.
-A~correspondence with linear $\pi$-calculus was established early on by Abramsky~\cite{abramsky94,bellinscott94}.
-A~correspondence between session-typed $\pi$-calculus and dual intuitionistic linear logic was developed by Caires and Pfenning~\cite[$\pi\text{DILL}$]{cairespfenning10}, and with classical linear logic by Wadler~\cite[CP]{wadler15}, both guaranteeing deadlock freedom as a result of their connection to logic.
-Dardha and Gay~\cite[PCP]{dardhagay18} integrate Kobayashi and Padovani's work on priorities~\cite{kobayashi06,padovani14} with CP, creating the first calculus which combines priorities and strong ties with logic.
+A~correspondence with linear $\pi$-calculus was established early on by Abramsky~\cite{abramsky94} and Bellin and Scott~\cite{bellinscott94}.
+Many years later, a correspondence between the $\pi$-calculus with binary session types and linear logic was proposed: Caires and Pfenning~\cite[$\pi\text{DILL}$]{cairespfenning10} proposed a correspondence with dual intuitionistic linear logic, while Wadler~\cite[CP]{wadler14} proposed a correspondence with classical linear logic.
+Both works guarantee deadlock freedom as result of their connection to logic and in particular due to the fact that the cut rule only allows tree-structured processes.
+Dardha and Gay~\cite[PCP]{dardhagay18} integrate Kobayashi and Padovani's work on priorities~\cite{kobayashi06,padovani14} with CP, creating the first calculus which combines priorities and logic.
+They prove cycle-elimination, in the same lines as cut-elimination for linear logic, from which the meta-theory and deadlock freedom follow.
 Dardha and P\'{e}rez~\cite{dardhaperez15} compare Kobayashi-style typing and CLL typing, and show that CLL corresponds to a subsystem of Kobayashi's where restriction is applied once.
-Balzer~\etal~\cite[$\text{SILL}_S$]{balzerpfenning17} introduce shared state, which breaks deadlock freedom.
+Balzer~\etal~\cite[$\text{SILL}_S$]{balzerpfenning17} introduce sharing, which breaks deadlock freedom.
 They later restore deadlock freedom using priorities~\cite[$\text{SILL}_{S+}$]{balzertoninho19}.
 Carbone~\etal~\cite{CarboneMSY15,carbonelindley16} give a logical view of MPST with a generalised duality, called \emph{coherence}.
+Caires and P\'{e}rez~\cite{CairesP16} give a novel presentation of MPST in terms of binary session types and the use of a \emph{medium process} which guarantee protocol fidelity and deadlock freedom.
+Their binary session types are rooted in linear logic following the proofs-as-processes correspondences.
+Ciobanu and Horne~\cite{CiobanuH15} give the first processes-as-formulae correspondence of MPST and \emph{the calculus of structures} by \cite{Guglielmi07}, and give a calculus where global and local session types are specified inspired by Scribble\footnote{\url{http://www.scribble.org/}}.
+The calculus is then used to determine whether a set of local session types can be composed safely and in a deadlock free manner.
+Horne \cite{Horne20} presents a proof system for subtyping and multiparty compatibility where compatible processes are race free and deadlock free.
+This work also gives a processes-as-formulae correspondence--similar to the approach in~\cite{CiobanuH15}--with a non-commutative extension of linear logic.
+As such, the meta-theoretical results of this work are obtained in terms of cut-elimination.
 

### 2021-04-09 introduction.tex
a pass on abstract and intro

@@
 \section{Introduction}
-Session types are a type formalism used to specify and verify communication protocols between two or more communicating agents~\cite{honda93,takeuchihonda94,hondavasconcelos98,carbonehonda07}.
-They have been studied for many programming paradigms, including concurrent and functional paradigms.
-In particular, they have been defined for the $\pi$-calculus~\cite{sangiorgiwalker01}---a process calculus for communication and concurrency, and for Good Variation (GV)~\cite{wadler14,lindleymorris15}---a linear concurrent $\lambda$-calculus.
-Session type systems guarantee communication safety and session fidelity, namely that communication proceeds without any type mismatch and it follows the predefined protocol specified as a session type.
-Another key property in a concurrent setting is deadlock freedom, stating that agents communicate without getting stuck in cyclic dependencies.
-Deadlock freedom is a more involving property and session types alone are not enough to rule out deadlocks in cyclic structures of multiple interleaved sessions.
+Session types have been studied for many programming paradigms, including concurrent and functional paradigms.
+In particular, \emph{binary} session types have been defined for the $\pi$-calculus~\cite{sangiorgiwalker01}---a process calculus for communication and concurrency---originally by Honda~\etal~\cite{honda93,takeuchihonda94,hondavasconcelos98}; and later on for Good Variation (GV)~\cite{wadler15,lindleymorris15}---a linear concurrent $\lambda$-calculus.
+Binary session types guarantee communication safety and session fidelity, namely that communication proceeds without any type mismatch and it follows the predefined session protocol.
+Another key property in a concurrent setting is \emph{deadlock freedom}, stating that agents communicate without getting stuck in cyclic dependencies.
+Deadlock freedom is a more involving property and binary session types \emph{alone} are not enough to rule out deadlocks in cyclic structures of different interleaved sessions.
 Several techniques have been developed to address deadlock freedom in the $\pi$-calculus and concurrent $\lambda$-calculus with session types, as detailed below.
 
-With regards to $\pi$-calculus, a growing line of work on deadlock freedom leverages the Curry-Howard correspondences between intuitionistic or classical linear logic and $\pi$-calculus with session types~\cite{cairespfenning10,wadler14}, which guarantee deadlock freedom by design.
-This is achieved by merging constructs for parallel composition and channel restriction under the logical cut rule, which forces processes to share only one channel for communication and eliminates cyclic dependencies.
+With regards to $\pi$-calculus, a growing line of work on deadlock freedom leverages the Curry-Howard correspondences between intuitionistic or classical linear logic and the $\pi$-calculus with binary session types~\cite{cairespfenning10,wadler14}, which guarantees deadlock freedom by restricting the allowed processes to only trees.
+This is achieved by merging constructs for parallel composition and channel restriction under the logical cut rule, which forces processes to share \emph{only one} channel for communication, thus eliminating cyclic dependencies and deadlocks.
 Recent developments in the Classical Processes (CP)~\cite{wadler14} line of work have led to various approaches to decoupling these constructs, either by maintaining a strong correspondence to classical linear logic, as in Hypersequent CP~\cite{kokkemontesi19popl,kokkemontesi19tlla}, or by weakening the correspondence to classical linear logic in exchange for a more expressive language, as in Priority CP (PCP)~\cite{dardhagay18}.
@@
 Differently from previous work, PCP allows cyclic structures and restores deadlock freedom by adding priorities to types following Kobayashi~\cite{kobayashi06} and Padovani~\cite{padovani14}.
-Priorities establish an order of actions in a $\pi$-calculus process and are used to rule out bad interleaving of channels, guaranteeing deadlock freedom.
-
-With regards to GV, the original work on session types in GV~\cite{gayvasconcelos10,gayvasconcelos12} did not satisfy deadlock freedom.
-Later on, calculi in the GV family~\cite{wadler14,lindleymorris15} have achieved deadlock freedom via a syntactic restriction, \ie by combining channel creation and thread spawning into a single operation, called fork.
+Priorities establish an order of actions in a $\pi$-calculus process and are used to rule out bad interleaving of channels.
+Dardha and Gay prove cycle-elimination of PCP (following the standard cut-elimination proof) by leveraging priorities.
+Then, deadlock freedom follows as a result of the cycle-elimination proof.
+%
+With regards to GV, the original work on session types in GV~\cite{gayvasconcelos12} did not satisfy deadlock freedom.
+Later on, calculi in the GV family~\cite{wadler15,lindleymorris15} have achieved deadlock freedom via a syntactic restriction, \ie by combining channel creation and thread spawning into a single operation, called fork.
 The fork construct is the term representation of the cut construct in CP, which as mentioned above, evaluates to a channel restriction combined with a parallel composition.
-As with CP, this is restrictive because it limits communication structures to only trees.
+As with CP, this is restrictive because it limits process structures to only trees.
 

### 2021-04-09 introduction.tex
adding LCC

@@
 This is achieved by merging constructs for parallel composition and channel restriction under the logical cut rule, which forces processes to share \emph{only one} channel for communication, thus eliminating cyclic dependencies and deadlocks.
-Recent developments in the Classical Processes (CP)~\cite{wadler14} line of work have led to various approaches to decoupling these constructs, either by maintaining a strong correspondence to classical linear logic, as in Hypersequent CP~\cite{kokkemontesi19popl,kokkemontesi19tlla}, or by weakening the correspondence to classical linear logic in exchange for a more expressive language, as in Priority CP (PCP)~\cite{dardhagay18}.
+Recent developments in the Classical Processes (CP)~\cite{wadler14} line of work have led to various approaches to decoupling these constructs, either by maintaining a strong correspondence to classical linear logic, as in Hypersequent CP (HCP)~\cite{kokkemontesi19popl,kokkemontesi19tlla} or Linear Compositional Choreographies (LCC)~\cite{CarboneMS18}; or by weakening the correspondence to classical linear logic in exchange for a more expressive language, as in Priority CP (PCP)~\cite{dardhagay18}.
 PCP decouples CP's cut rule into two separate constructs: a construct for parallel composition, by introducing a mix rule; and a construct for restriction, by introducing a cycle rule.

### 2021-04-09 introduction.tex
another pass to intro

--- 2021-04-09 introduction.tex
@@
 This is achieved by merging constructs for parallel composition and channel restriction under the logical cut rule, which forces processes to share \emph{only one} channel for communication, thus eliminating cyclic dependencies and deadlocks.
-Recent developments in the Classical Processes (CP)~\cite{wadler14} line of work have led to various approaches to decoupling these constructs, either by maintaining a strong correspondence to classical linear logic, as in Hypersequent CP~\cite{kokkemontesi19popl,kokkemontesi19tlla}, or by weakening the correspondence to classical linear logic in exchange for a more expressive language, as in Priority CP (PCP)~\cite{dardhagay18}.
+Recent developments in the Classical Processes (CP)~\cite{wadler14} line of work have led to various approaches to decoupling these constructs.
+One way is by maintaining a strong correspondence to classical linear logic and allowing only tree-structured processes, such as in Hypersequent CP (HCP)~\cite{kokkemontesi19popl,kokkemontesi19tlla}, Linear Compositional Choreographies (LCC)~\cite{CarboneMS18} and Classical Transitions (CT)~\cite{MP18}.
+Another way is by weakening the correspondence to classical linear logic in exchange for a more expressive language where cyclic-structured processes are allowed, such as in Priority CP (PCP)~\cite{dardhagay18}.
 PCP decouples CP's cut rule into two separate constructs: a construct for parallel composition, by introducing a mix rule; and a construct for restriction, by introducing a cycle rule.
-Differently from previous work, PCP allows cyclic structures and restores deadlock freedom by adding priorities to types following Kobayashi~\cite{kobayashi06} and Padovani~\cite{padovani14}.
-Priorities establish an order of actions in a $\pi$-calculus process and are used to rule out bad interleaving of channels.
+This decoupling allows cyclic processes.
+In order to restore deadlock freedom, PCP uses priorities following Kobayashi~\cite{kobayashi06} and Padovani~\cite{padovani14}.
+Priorities establish an order of actions in a $\pi$-calculus process and are used to rule out bad interleavings of channels.
 Dardha and Gay prove cycle-elimination of PCP (following the standard cut-elimination proof) by leveraging priorities.

### 2021-04-12 priority-cp.tex
Added clarification for child vs priority o.

@@
 
-Typing environments are translated pointwise, and sequents $\pcp{\seq{P}{\ty{\Gamma}}}$ are translated as $\pgv{\cseq[\child]{\ty{\cpgvT{\ty{\Gamma}}}}{\cpgvC{P}}}$.
+Typing environments are translated pointwise, and sequents $\pcp{\seq{P}{\ty{\Gamma}}}$ are translated as $\pgv{\cseq[\child]{\ty{\cpgvT{\ty{\Gamma}}}}{\cpgvC{P}}}$, where $\tm{\pgv{\child}}$ indicates a child thread.
+Translated processes do not have a main thread.
 The translations $\tm{\cpgvM{\cdot}}$ and $\tm{\cpgvC{\cdot}}$ preserve typing, and the latter induces a sound and complete operational correspondence.
@@
 \begin{compacttheorems}
-  \begin{restatablelemma}{lempcptopgvtermspreservation}[Preservation, ${\tm{\cpgvM{\cdot}}}$]
+  \begin{restatablelemma}{lempcptopgvtermspreservation}[Preservation, ${\tm{\cpgvM{\cdot}}}$]%
     \label{lem:pcp-to-pgv-terms-preservation}
@@
   \end{restatablelemma}
-  \begin{restatabletheorem}{thmpcptopgvconfspreservation}[Preservation, ${\tm{\cpgvC{\cdot}}}$]
+  \begin{restatabletheorem}{thmpcptopgvconfspreservation}[Preservation, ${\tm{\cpgvC{\cdot}}}$]%
     \label{thm:pcp-to-pgv-confs-preservation}
@@
   \end{restatabletheorem}
-  \begin{restatablelemma}{lempcptopgvcpgvMtocpgvC}
+  \begin{restatablelemma}{lempcptopgvcpgvMtocpgvC}%
     \label{lem:pcp-to-pgv-cpgvM-to-cpgvC}
@@
   \begin{restatabletheorem}{thmpcptopgvoperationalcorrespondencesoundness}%
-    [Operational Correspondence, Soundness, ${\tm{\cpgvC{\cdot}}}$]
+    [Operational Correspondence, Soundness, ${\tm{\cpgvC{\cdot}}}$]%
     \label{thm:pcp-to-pgv-operational-correspondence-soundness}
@@
   \begin{restatabletheorem}{thmpcptopgvoperationalcorrespondencecompleteness}%
-    [Operational Correspondence, Completeness, ${\tm{\cpgvC{\cdot}}}$]
+    [Operational Correspondence, Completeness, ${\tm{\cpgvC{\cdot}}}$]%
     \label{thm:pcp-to-pgv-operational-correspondence-completeness}

### 2021-04-12 priority-gv.tex
Added clarification for child vs priority o.

@@
 \]
+The use of $\tm{\child}$ for child threads~\cite{lindleymorris15} overlaps with the use of the meta-variable $\cs{o}$ for priorities~\cite{dardhagay18}.
+Both are used to annotate sequents: flags appear on the sequent in configuration typing, and priorities in term typing.
+To distinguish the two symbols, they are typeset in a different font and a different colour.
+
 Values ($\tm{V}$, $\tm{W}$), evaluation contexts ($\tm{E}$), thread evaluation contexts ($\tm{\conf{F}}$), and configuration contexts ($\tm{\conf{G}}$) are defined by the following grammars:
@@
 Paraphrasing Dardha and Gay~\cite{dardhagay18}, priorities obey the following two laws:
-\begin{enumerate*}[label=(\roman*)]
+\begin{enumerate*}[label= (\roman*) ]
 \item an action with lower priority happens before an action with higher priority; and
@@
 While our configuration typing is based on the standard typing for GV, it differs on two counts:
-\begin{enumerate*}[label=(\roman*)]
+\begin{enumerate*}[label= (\roman*) ]
 \item \emph{we require that child threads return the unit value}, as opposed to a terminated channel; and
@@
 \begin{compacttheorems}
-  \begin{restatablelemma}{lempgvvaluedone}
+  \begin{restatablelemma}{lempgvvaluedone}%
     \label{lem:pgv-value-done}
@@
   \end{restatablelemma}
-  \begin{restatablelemma}{lempgvsubstitution}[Substitution]
+  \begin{restatablelemma}{lempgvsubstitution}[Substitution]%
     \label{lem:pgv-substitution}
@@
   \end{restatablelemma}
-  \begin{restatablelemma}{lempgvsubjectreductionterms}[Subject Reduction, $\tred$]
+  \begin{restatablelemma}{lempgvsubjectreductionterms}[Subject Reduction, $\tred$]%
     \label{lem:pgv-subject-reduction-terms}
@@
   \end{restatablelemma}
-  \begin{restatablelemma}{lempgvsubjectcongruence}[Subject Congruence, $\equiv$]
+  \begin{restatablelemma}{lempgvsubjectcongruence}[Subject Congruence, $\equiv$]%
     \label{lem:pgv-subject-congruence}
@@
   \end{restatablelemma}
-  \begin{restatabletheorem}{thmpgvsubjectreductionconfs}[Subject Reduction, $\cred$]
+  \begin{restatabletheorem}{thmpgvsubjectreductionconfs}[Subject Reduction, $\cred$]%
     \label{thm:pgv-subject-reduction-confs}
@@
 \emph{Actions} are terms which perform communication actions and which synchronise between two threads.
-\begin{definition}[Actions]
+\begin{definition}[Actions]%
   \label{def:pgv-actions}
@@
 \emph{Ready terms} are terms which are ready to perform a communication action, either by themselves, \eg creating a new channel or thread, or with another thread, \eg sending or receiving.
-\begin{definition}[Ready Terms]
+\begin{definition}[Ready Terms]%
   \label{def:pgv-ready-actions}
@@
 Progress for the term language is standard for GV, and deviates from progress for linear $\lambda$-calculus only in that terms may reduce to values or \emph{ready terms}:
-\begin{restatablelemma}{lempgvopenprogressterms}[Progress, $\tred$]
+\begin{restatablelemma}{lempgvopenprogressterms}[Progress, $\tred$]%
   \label{lem:pgv-open-progress-terms}
@@
 \begin{compacttheorems}
-  \begin{definition}[Canonical Forms]
+  \begin{definition}[Canonical Forms]%
     \label{def:pgv-canonical-forms}
@@
   \end{definition}
-  \begin{definition}[Normal Forms]
+  \begin{definition}[Normal Forms]%
     A~configuration $\tm{\conf{C}}$ is in normal form if it is of the following form, where each $\tm{M_i}$ is ready to act on $\tm{x_i}$:
@@
   \end{definition}
-  \begin{restatablelemma}{lempgvcanonicalforms}[Canonical Forms]
+  \begin{restatablelemma}{lempgvcanonicalforms}[Canonical Forms]%
     \label{lem:pgv-canonical-forms}
@@
   \end{restatablelemma}
-  \begin{restatabletheorem}{thmpgvclosedprogressconfs}[Progress, $\cred$]
+  \begin{restatabletheorem}{thmpgvclosedprogressconfs}[Progress, $\cred$]%
     \label{thm:pgv-closed-progress-confs}
@@
   \end{proof}
-  \begin{corollary}
+  \begin{corollary}%
     \label{cor:pgv-closed-progress}

### 2021-04-12 priority-cp.tex
Remove instances of \cite as noun.

@@
 The main change we make to PCP is \emph{removing commuting conversions} and defining an operational semantics based on structural congruence.
-Commuting conversions are necessary if we want our reduction strategy to correspond \emph{exactly} to cut (or cycle in \cite{dardhagay18}) elimination.
+Commuting conversions are necessary if we want our reduction strategy to correspond \emph{exactly} to cut elimination.
 However, from the perspective of process calculi, commuting conversions behave strangely: they allow an input/output action to be moved to the top of a process, thus potentially blocking actions which were previously possible (an example is given in \cref{app:commuting-conversions}).
-This makes CP, and PCP in \cite{dardhagay18}, non-confluent.
+This makes CP, and Dardha and Gay's PCP~\cite{dardhagay18}, non-confluent.
 As Lindley and Morris~\cite{lindleymorris15} show, all communications that can be performed \emph{with} the use of commuting conversions, can also be performed \emph{without} them, if using structural congruence.
@@
 \begin{compacttheorems}
-  \begin{restatabletheorem}{thmpcpclosedprogress}[Progress, $\pcp{\red}$]
+  \begin{restatabletheorem}{thmpcpclosedprogress}[Progress, $\pcp{\red}$]%
     \label{thm:pcp-closed-progress}
@@
 
-%\noindent
-%\textbf{Milner's Cyclic Scheduler}.
-%As an example of a deadlock-free cyclic process, Dardha and Gay~\cite{dardhagay18} introduce an implementation of Milner's cyclic scheduler~\cite{milner89} in Priority CP.
-We have reproduced the scheduler in our updated PCP, and showed its translation to PGV.
-Full details are given in \cref{sec:milner}.
-
 \end{document}

### 2021-04-13 introduction.tex
work on abstract and intro

@@
 Session types have been studied for many programming paradigms, including concurrent and functional paradigms.
-In particular, \emph{binary} session types have been defined for the $\pi$-calculus~\cite{sangiorgiwalker01}---a process calculus for communication and concurrency---originally by Honda~\etal~\cite{honda93,takeuchihonda94,hondavasconcelos98}; and later on for Good Variation (GV)~\cite{wadler15,lindleymorris15}---a linear concurrent $\lambda$-calculus.
-Binary session types guarantee communication safety and session fidelity, namely that communication proceeds without any type mismatch and it follows the predefined session protocol.
+In particular, \emph{binary} session types have been defined for the $\pi$-calculus~\cite{honda93,takeuchihonda94,hondavasconcelos98}---a process calculus for communication and concurrency; and later on for Good Variation (GV)~\cite{wadler15,lindleymorris15}---a linear concurrent $\lambda$-calculus.
+Binary session types guarantee communication safety and session fidelity, namely that communication proceeds without any type mismatch and it follows the predefined protocol.
 Another key property in a concurrent setting is \emph{deadlock freedom}, stating that agents communicate without getting stuck in cyclic dependencies.
 Deadlock freedom is a more involving property and binary session types \emph{alone} are not enough to rule out deadlocks in cyclic structures of different interleaved sessions.
-Several techniques have been developed to address deadlock freedom in the $\pi$-calculus and concurrent $\lambda$-calculus with session types, as detailed below.
+Several techniques have been developed to address deadlock freedom in session-typed $\pi$-calculus and concurrent $\lambda$-calculus.
 
-With regards to $\pi$-calculus, a growing line of work on deadlock freedom leverages the Curry-Howard correspondences between intuitionistic or classical linear logic and the $\pi$-calculus with binary session types~\cite{cairespfenning10,wadler14}, which guarantees deadlock freedom by restricting the allowed processes to only trees.
-This is achieved by merging constructs for parallel composition and channel restriction under the logical cut rule, which forces processes to share \emph{only one} channel for communication, thus eliminating cyclic dependencies and deadlocks.
-Recent developments in the Classical Processes (CP)~\cite{wadler14} line of work have led to various approaches to decoupling these constructs.
-One way is by maintaining a strong correspondence to classical linear logic and allowing only tree-structured processes, such as in Hypersequent CP (HCP)~\cite{kokkemontesi19popl,kokkemontesi19tlla}, Linear Compositional Choreographies (LCC)~\cite{CarboneMS18} and Classical Transitions (CT)~\cite{MP18}.
-Another way is by weakening the correspondence to classical linear logic in exchange for a more expressive language where cyclic-structured processes are allowed, such as in Priority CP (PCP)~\cite{dardhagay18}.
+With regards to $\pi$-calculus, a growing line of work leverages the Curry-Howard correspondences between session-typed $\pi$-calculus and intuitionistic~\cite{cairespfenning10} or classical linear logic~\cite{wadler14}, which guarantees deadlock freedom by restricting processes to only trees.
+The tree structure is achieved by merging constructs for parallel composition and channel restriction under the logical cut rule, which forces processes to share \emph{only one} channel for communication.
+Recent developments in the CP~\cite{wadler14} line of work have led to various approaches to decoupling these constructs.
+One way is by maintaining a strong correspondence to classical linear logic and allowing only tree-structured processes, such as in Hypersequent CP (HCP)~\cite{kokkemontesi19popl,kokkemontesi19tlla}, Linear Compositional Choreographies (LCC)~\cite{CarboneMS18} or Classical Transitions (CT)~\cite{MP18}.
+Another way is by weakening the correspondence to classical linear logic in exchange for a more expressive language which allows cyclic-structured processes, such as in Priority CP (PCP)~\cite{dardhagay18}.
 PCP decouples CP's cut rule into two separate constructs: a construct for parallel composition, by introducing a mix rule; and a construct for restriction, by introducing a cycle rule.
-This decoupling allows cyclic processes.
 In order to restore deadlock freedom, PCP uses priorities following Kobayashi~\cite{kobayashi06} and Padovani~\cite{padovani14}.
-Priorities establish an order of actions in a $\pi$-calculus process and are used to rule out bad interleavings of channels.
-Dardha and Gay prove cycle-elimination of PCP (following the standard cut-elimination proof) by leveraging priorities.
-Then, deadlock freedom follows as a result of the cycle-elimination proof.
-%
+Priorities establish an \emph{order} of actions and are used to rule out bad interleavings of channels.
+By leveraging priorities, Dardha and Gay prove cycle-elimination of PCP (following the standard cut-elimination proof for classical linear logic).
+Deadlock freedom then follows as a corollary of the cycle-elimination result.
+
 With regards to GV, the original work on session types in GV~\cite{gayvasconcelos12} did not satisfy deadlock freedom.
 Later on, calculi in the GV family~\cite{wadler15,lindleymorris15} have achieved deadlock freedom via a syntactic restriction, \ie by combining channel creation and thread spawning into a single operation, called fork.
-The fork construct is the term representation of the cut construct in CP, which as mentioned above, evaluates to a channel restriction combined with a parallel composition.
+The fork construct is the term representation of the cut construct in CP, which as discussed above, evaluates to a channel restriction combined with a parallel composition.
 As with CP, this is restrictive because it limits process structures to only trees.
 
-We aim to further investigate GV and deadlock freedom, as there are several benefits of working with a functional language as opposed to name-passing calculi.
+We aim to further investigate GV and deadlock freedom, as there are several benefits of working with a functional language as opposed to name-passing process calculi.
 Functional languages support higher-order functions, and have a capability for abstraction not usually present in process calculi.
 Working within a functional language allows to derive extensions of the communication capabilities of the language via well-understood extensions of the functional fragment, \ie deriving internal/external choice via sum types.
-Finally, in concurrent functional calculi, there is a clear separation between a program that the user writes, and a configuration, which is the state of a system as it evaluates.
+In concurrent functional calculi, there is a clear separation between a program that the user writes, and a configuration, which is the state of a system as it evaluates.
 In process calculi, these are conflated.
-Moreover, the benefit of working with GV over other session-typed functional languages is that GV has strong ties to linear logic, via its relation to CP~\cite{wadler14}, and consequently it has strong formal properties, \eg deadlock freedom, albeit in a restrictive way.
-Following the above observations, we thus pose our research question.
+Finally, functional calculi bring us a step closer to transferring our work to real-world (functional) programming languages~\cite{KD21b}.
+Moreover, the benefit of working with GV as opposed to other session-typed functional languages is that GV can be embedded in CP~\cite{wadler14}, and consequently it guarantees the same formal properties as CP, \eg deadlock freedom, albeit only for tree-structured processes.
+Following the above observations, we thus pose our research question for GV.
 
-\textbf{RQ:} \emph{Can we design a more expressive GV where deadlock freedom is guaranteed by design and communication structures are not limited to only trees?}
+\textbf{RQ:} \emph{Can we design a more expressive GV which guarantees deadlock freedom for cyclic-structured processes?}
 
 We address this research question by following the line of work from CP to Priority CP, and present
-Priority GV (PGV), a~variant of GV which decouples channel creation from thread spawning and restores deadlock freedom by adding priorities.
-This work closes the circle of the strong connection of CP and GV, together with their priority-based versions, PCP and PGV.
-
+Priority GV (PGV), a~variant of GV which decouples channel creation from thread spawning, thus allowing cyclic-structured processes and restores deadlock freedom via priorities.
+This closes the circle of the connection between CP and GV~\cite{wadler14}, and their priority-based versions, PCP~\cite{dardhagay18} and PGV.
 In this paper we make the following contributions:

### 2021-04-13 introduction.tex
update contributions

@@
 Finally, functional calculi bring us a step closer to transferring our work to real-world (functional) programming languages~\cite{KD21b}.
-Moreover, the benefit of working with GV as opposed to other session-typed functional languages is that GV can be embedded in CP~\cite{wadler14}, and consequently it guarantees the same formal properties as CP, \eg deadlock freedom, albeit only for tree-structured processes.
+Moreover, the benefit of working with GV as opposed to other session-typed functional languages is that GV can be translated into CP~\cite{wadler14}, thus guaranteeing the same formal properties as CP, \eg deadlock freedom, albeit only for tree-structured processes.
 Following the above observations, we thus pose our research question for GV.
@@
 This closes the circle of the connection between CP and GV~\cite{wadler14}, and their priority-based versions, PCP~\cite{dardhagay18} and PGV.
-In this paper we make the following contributions:
+%
+In this paper we make the following main contributions:
 \begin{enumerate}
-\item \textbf{Priority GV}. We present Priority GV (PGV) in~\cref{sec:pgv}, and prove subject reduction (\cref{thm:pgv-subject-reduction-confs}) and progress (\cref{thm:pgv-closed-progress-confs}).
-PGV is the \emph{first} session-typed functional language with priorities \emph{and} strong ties to linear logic via its correspondence with PCP and its relation to the CP and GV languages.
-\item \textbf{Solving GV problems}. PGV addresses several problems in the original GV language, most notably:
+	\item \textbf{Priority GV}.
+We present Priority GV (PGV), a session-typed functional language with priorities (\cref{sec:pgv}) and prove subject reduction (\cref{thm:pgv-subject-reduction-confs}) and progress (\cref{thm:pgv-closed-progress-confs}).
+	PGV addresses several problems in the original GV language, most notably:
  	\begin{enumerate*}
-	\item our version does not have the pseudo-type $S^\sharp$;
+	\item we do not require the pseudo-type $S^\sharp$; and
 	\item our structural congruence is type preserving.
 	\end{enumerate*}
-\item \textbf{Updated Priority CP.} We present an updated version of Priority CP~\cite{dardhagay18} in \cref{sec:pcp}.
+	PGV answers our research question as it allows cyclic-structured binary session-typed processes that are deadlock free.
+	\item \textbf{Updated PCP}.
+We present an updated version of Priority CP~\cite{dardhagay18} in \cref{sec:pcp}.
+We remove commuting conversions and move away from reduction as cut elimination, and present a reduction relation based on structural congruence, as it is standard in process calculi.
+	\item \textbf{Translation of PGV to PCP}.
+We present a \emph{sound and complete encoding} of PCP to PGV in~\cref{sec:pcp}, where we prove the encoding preserves typing (\cref{thm:pcp-to-pgv-confs-preservation}) and satisfies operational correspondence (\cref{thm:pcp-to-pgv-operational-correspondence-soundness,thm:pcp-to-pgv-operational-correspondence-completeness}).
+\end{enumerate}
+
+%\begin{enumerate}
+%\item \textbf{Priority GV}. We present Priority GV (PGV) in~\cref{sec:pgv}, and prove subject reduction (\cref{thm:pgv-subject-reduction-confs}) and progress (\cref{thm:pgv-closed-progress-confs}).
+%PGV is the \emph{first} session-typed functional language with priorities \emph{and} strong ties to linear logic via its correspondence with PCP and its relation to the CP and GV languages.
+%\item \textbf{Solving GV problems}. PGV addresses several problems in the original GV language, most notably:
+% 	\begin{enumerate*}
+%	\item our version does not have the pseudo-type $S^\sharp$;
+%	\item our structural congruence is type preserving.
+%	\end{enumerate*}
+%\item \textbf{Updated Priority CP.} We present an updated version of Priority CP~\cite{dardhagay18} in \cref{sec:pcp}.
 We remove commuting conversions and move away from reduction as cut elimination, towards reduction based on structural congruence, as it is standard in process calculi.
-\item \textbf{Connection to linear logic}.
+%\item \textbf{Connection to linear logic}.
 The connection of PGV to linear logic is given via a \emph{sound and complete encoding} of PCP to PGV in~\cref{sec:pcp}, where we prove the encoding preserves typing (\cref{thm:pcp-to-pgv-confs-preservation}) and satisfies operational correspondence (\cref{thm:pcp-to-pgv-operational-correspondence-soundness,thm:pcp-to-pgv-operational-correspondence-completeness}).
 We recall Milner's cyclic scheduler~\cite{milner89} presented in PCP~\cite{dardhagay18} and encode it in PGV.
-\end{enumerate}
+%\end{enumerate}
 \end{document}

### 2021-04-13 conclusion.tex
adding more on future work

@@
 Our formalism so far only captures the core of GV.
-In future work, we plan to explore: recursion in PGV, by integrating the works of Lindley and Morris~\cite{lindleymorris16} and Padovani and Novara~\cite{padovaninovara15}; %a~translation from Priority GV to Priority CP;
-and sharing, following Balzer and Pfenning~\cite{balzerpfenning17}.
+In future work, we plan to explore: recursion in PGV, by integrating the works of Lindley and Morris~\cite{lindleymorris16} and Padovani and Novara~\cite{padovaninovara15}; or in the lines of Kobayashi and Laneve for unbounded networks~\cite{KL17}, which is a refinement of previous work by Kobayashi.
+%a~translation from Priority GV to Priority CP;
+Another interesting future work is on sharing, following Balzer and Pfenning~\cite{balzerpfenning17}, or Voinea \etal~\cite{VoineaDG19}.
 

### 2021-04-13 introduction.tex
Minor changes.

@@
 Session types have been studied for many programming paradigms, including concurrent and functional paradigms.
-In particular, \emph{binary} session types have been defined for the $\pi$-calculus~\cite{honda93,takeuchihonda94,hondavasconcelos98}---a process calculus for communication and concurrency; and later on for Good Variation (GV)~\cite{wadler15,lindleymorris15}---a linear concurrent $\lambda$-calculus.
+In particular, \emph{binary} session types have been defined for the $\pi$-calculus~\cite{honda93,takeuchihonda94,hondavasconcelos98}, a process calculus for communication and concurrency; and later on for Good Variation~\cite[GV]{wadler15,lindleymorris15}, a session-typed concurrent $\lambda$-calculus.
 Binary session types guarantee communication safety and session fidelity, namely that communication proceeds without any type mismatch and it follows the predefined protocol.
@@
 Recent developments in the CP~\cite{wadler14} line of work have led to various approaches to decoupling these constructs.
-One way is by maintaining a strong correspondence to classical linear logic and allowing only tree-structured processes, such as in Hypersequent CP (HCP)~\cite{kokkemontesi19popl,kokkemontesi19tlla}, Linear Compositional Choreographies (LCC)~\cite{CarboneMS18} or Classical Transitions (CT)~\cite{MP18}.
-Another way is by weakening the correspondence to classical linear logic in exchange for a more expressive language which allows cyclic-structured processes, such as in Priority CP (PCP)~\cite{dardhagay18}.
+One way is by maintaining a strong correspondence to classical linear logic and allowing only tree-structured processes, such as in Hypersequent CP~\cite[HCP]{kokkemontesi19popl,kokkemontesi19tlla}, Linear Compositional Choreographies~\cite[LCC]{CarboneMS18} or Classical Transitions~\cite[CT]{MP18}.
+Another way is by weakening the correspondence to classical linear logic in exchange for a more expressive language which allows cyclic-structured processes, such as in Priority CP~\cite[PCP]{dardhagay18}.
 PCP decouples CP's cut rule into two separate constructs: a construct for parallel composition, by introducing a mix rule; and a construct for restriction, by introducing a cycle rule.
@@
 With regards to GV, the original work on session types in GV~\cite{gayvasconcelos12} did not satisfy deadlock freedom.
-Later on, calculi in the GV family~\cite{wadler15,lindleymorris15} have achieved deadlock freedom via a syntactic restriction, \ie by combining channel creation and thread spawning into a single operation, called fork.
+Later on, calculi in the GV family~\cite{wadler15,lindleymorris15} have achieved deadlock freedom via a syntactic restriction, \ie, by combining channel creation and thread spawning into a single operation, called fork.
 The fork construct is the term representation of the cut construct in CP, which as discussed above, evaluates to a channel restriction combined with a parallel composition.
@@
 Functional languages support higher-order functions, and have a capability for abstraction not usually present in process calculi.
-Working within a functional language allows to derive extensions of the communication capabilities of the language via well-understood extensions of the functional fragment, \ie deriving internal/external choice via sum types.
+Working within a functional language allows to derive extensions of the communication capabilities of the language via well-understood extensions of the functional fragment, \ie, deriving internal/external choice via sum types.
 In concurrent functional calculi, there is a clear separation between a program that the user writes, and a configuration, which is the state of a system as it evaluates.
@@
 Finally, functional calculi bring us a step closer to transferring our work to real-world (functional) programming languages~\cite{KD21b}.
-Moreover, the benefit of working with GV as opposed to other session-typed functional languages is that GV can be translated into CP~\cite{wadler14}, thus guaranteeing the same formal properties as CP, \eg deadlock freedom, albeit only for tree-structured processes.
-Following the above observations, we thus pose our research question for GV.
+Moreover, the benefit of working with GV as opposed to other session-typed functional languages is that GV can be translated into CP~\cite{wadler14}, thus guaranteeing the same formal properties as CP, \eg, deadlock freedom, albeit only for tree-structured processes.
+Following the above observations, we thus pose our research question for GV.%
 
@@
 Priority GV (PGV), a~variant of GV which decouples channel creation from thread spawning, thus allowing cyclic-structured processes and restores deadlock freedom via priorities.
-This closes the circle of the connection between CP and GV~\cite{wadler14}, and their priority-based versions, PCP~\cite{dardhagay18} and PGV.
-%
+This closes the circle of the connection between CP and GV~\cite{wadler14}, and their priority-based versions, PCP~\cite{dardhagay18} and PGV.%
 In this paper we make the following main contributions:
@@
 We remove commuting conversions and move away from reduction as cut elimination, and present a reduction relation based on structural congruence, as it is standard in process calculi.
-	\item \textbf{Translation of PGV to PCP}.
+	\item \textbf{Translation from PCP to PGV}.
 We present a \emph{sound and complete encoding} of PCP to PGV in~\cref{sec:pcp}, where we prove the encoding preserves typing (\cref{thm:pcp-to-pgv-confs-preservation}) and satisfies operational correspondence (\cref{thm:pcp-to-pgv-operational-correspondence-soundness,thm:pcp-to-pgv-operational-correspondence-completeness}).
 \end{enumerate}
-
-%\begin{enumerate}
-%\item \textbf{Priority GV}. We present Priority GV (PGV) in~\cref{sec:pgv}, and prove subject reduction (\cref{thm:pgv-subject-reduction-confs}) and progress (\cref{thm:pgv-closed-progress-confs}).
-%PGV is the \emph{first} session-typed functional language with priorities \emph{and} strong ties to linear logic via its correspondence with PCP and its relation to the CP and GV languages.
-%\item \textbf{Solving GV problems}. PGV addresses several problems in the original GV language, most notably:
-% 	\begin{enumerate*}
-%	\item our version does not have the pseudo-type $S^\sharp$;
-%	\item our structural congruence is type preserving.
-%	\end{enumerate*}
-%\item \textbf{Updated Priority CP.} We present an updated version of Priority CP~\cite{dardhagay18} in \cref{sec:pcp}.
-We remove commuting conversions and move away from reduction as cut elimination, towards reduction based on structural congruence, as it is standard in process calculi.
-%\item \textbf{Connection to linear logic}.
-The connection of PGV to linear logic is given via a \emph{sound and complete encoding} of PCP to PGV in~\cref{sec:pcp}, where we prove the encoding preserves typing (\cref{thm:pcp-to-pgv-confs-preservation}) and satisfies operational correspondence (\cref{thm:pcp-to-pgv-operational-correspondence-soundness,thm:pcp-to-pgv-operational-correspondence-completeness}).
-We recall Milner's cyclic scheduler~\cite{milner89} presented in PCP~\cite{dardhagay18} and encode it in PGV.
-%\end{enumerate}
 \end{document}

### 2021-04-13 introduction.tex
Compacted references.

@@
 	\item \textbf{Translation from PCP to PGV}.
-We present a \emph{sound and complete encoding} of PCP to PGV in~\cref{sec:pcp}, where we prove the encoding preserves typing (\cref{thm:pcp-to-pgv-confs-preservation}) and satisfies operational correspondence (\cref{thm:pcp-to-pgv-operational-correspondence-soundness,thm:pcp-to-pgv-operational-correspondence-completeness}).
+We present a \emph{sound and complete encoding} of PCP in PGV in~\cref{sec:pcp}, where we prove the encoding preserves typing (\cref{thm:pcp-to-pgv-confs-preservation}) and satisfies operational correspondence (\cref{thm:pcp-to-pgv-operational-correspondence-soundness,thm:pcp-to-pgv-operational-correspondence-completeness}).
 \end{enumerate}

### 2021-04-13 priority-gv.tex
Compacted references.

@@
 \end{enumerate*}
-We will show that only \cref{ex:cycle} is typeable in PGV, while it is not typable in GV~\cite{wadler14} nor guaranteed to be deadlock-free by its predecessor~\cite{gayvasconcelos10}.
+We will show that only \cref{ex:cycle} is typeable in PGV, while it is not typable in GV~\cite{wadler14} nor guaranteed to be deadlock-free by its predecessor~\cite{gayvasconcelos12}.
 

### 2021-04-13 conclusion.tex
More citations appeared :'(

@@
 Gay~\etal~\cite{gaynagarajan03} and Vasconcelos~\etal~\cite{vasconcelosravara04,vasconcelosgay06} were the first to introduce a functional language with session types.
-However, such works, including early GV~\cite{gayvasconcelos10,gayvasconcelos12} did not guarantee deadlock freedom, until it was addressed via syntactic restrictions~\cite{lindleymorris15,wadler14}.
+However, such works, including early GV~\cite{gayvasconcelos12} did not guarantee deadlock freedom, until it was addressed via syntactic restrictions~\cite{lindleymorris15,wadler14}.
 Toninho~\etal~\cite{toninhocaires12} present a translation of simply-typed $\lambda$-calculus into session-typed $\pi$-calculus.

### 2021-04-13 conclusion.tex
Rewrote related work.

--- 2021-04-09 conclusion.tex
@@
 \begin{document}
-\section{Related Work and Conclusion}
+\section{Related Work}
 \subsubsection*{Deadlock freedom and progress}
 Deadlock freedom and progress are well studied properties in the $\pi$-calculus.
-For the `standard' typed $\pi$-calculus, one line of work is Kobayashi's approach to deadlock freedom~\cite{kobayashi98}, where priorities are abstract tags defined over a partially ordered set.
-These tags were later simplified to pairs of natural numbers, called obligations and capabilities~\cite{kobayashi06}, which allowed more $\pi$-calculus processes to be typed.
-Padovani~\cite{padovani13} adapted the obligation and capability pairs to session types, and later simplified them to a single priority for the linear $\pi$-calculus~\cite{padovani14}.
-Furthermore, by exploiting the encoding of session types into linear types~\cite{kobayashi07,dardhagiachino12,dardha14beat} and the priority-based linear $\pi$-calculus,
-we can obtain deadlock freedom for the $\pi$-calculus with session types.
+For the `standard' typed $\pi$-calculus, an important line of work starts from Kobayashi's approach to deadlock freedom~\cite{kobayashi98}, where priorities are values from an abstract poset.
+Kobayashi~\cite{kobayashi06} simplifies the abstract poset to pairs of naturals, called \emph{obligations} and\emph{capabilities}.
+Padovani simplifies these further to a single natural, called a \emph{priority}~\cite{padovani14}, and adapts obligations/capabilities to session types~\cite{padovani13}.
+
+% We can obtain deadlock freedom for the $\pi$-calculus with session types via the encoding of session types using linear types~\cite{kobayashi07,dardhagiachino12,dardha14beat} and linear $\pi$-calculus with priorities.
 
 For the session-typed $\pi$-calculus, Dezani~\etal~\cite{dezani-ciancaglinimostrous06} guarantee progress by allowing only one active session at a time.
-Dezani later~\cite{dezani-ciancagliniliguoro09progress} introduces a partial order on channels similar to Kobayashi~\cite{kobayashi98}.
-Carbone and Debois~\cite{carbonedebois10} define progress for session typed $\pi$-calculus in terms of a \emph{catalyser}, which provides a missing counterpart to a process, thus guaranteeing deadlock freedom.
-Carbone~\etal~\cite{carbonedardha14} use such catalysers to show that progress is a compositional form of lock-freedom and that it can be lifted to session types via the encoding of session types to linear types.
+Dezani~\cite{dezani-ciancagliniliguoro09progress} introduces a partial order on channels, similar to Kobayashi~\cite{kobayashi98}.
+Carbone and Debois~\cite{carbonedebois10} define progress for session typed $\pi$-calculus in terms of a \emph{catalyser} which provides the missing counterpart to a process.
+Carbone~\etal~\cite{carbonedardha14} use catalysers to show that progress is a compositional form of lock-freedom and can be lifted to session types via the encoding of session types to linear types~\cite{kobayashi07,dardhagiachino12,dardha14beat}.
 Vieira and Vasconcelos~\cite{vieiravasconcelos13} use single priorities and an abstract partial order to guarantee deadlock freedom in a binary session-typed $\pi$-calculus and building on conservation types.
 
-While our work focuses on \emph{binary} session types, it is worth to discuss related work on Multiparty Session Types (MPST) in order to give a broader context.
-The original work on MPST by Honda~\etal~\cite{hondayoshida08} guarantees deadlock freedom \emph{within a single} session, but not for session interleaving.
-Later on, several techniques for deadlock freedom in MPST were defined.
-Bettini~\etal~\cite{bettinicoppo08} follow a similar technique to Kobayashi's for the MPST setting.
-The main difference with our work is that we associate priorities with communication actions, whether Bettini~\etal~\cite{bettinicoppo08} with channels.
-Carbone and Montesi~\cite{carbonemontesi13} explore and combine MPST and choreographies to obtain a formalism that satisfies deadlock freedom by design.
-Deni{\'{e}}lou and Yoshida \cite{DenielouY13} introduce the notion of \emph{multiparty compatibility} which generalises that of duality in binary session types and they synthesise safe and deadlock-free global types from local session types via the use of LTSs and communicating automata (CFSA).
-Castellani~\etal~\cite{CastellaniDGH20} guarantee lock freedom--a stronger property than deadlock freedom--in the context of MPST with \emph{internal delegation}, where participants are allowed to delegate tasks to each other as long as they fall within the same multiparty session.
-This allows for the internal delegation to be captured by the global type.
-Scalas and Yoshida \cite{scalasyoshida19} provide a revision of the theoretical foundations of MPST leading to a less complicated and more general theory, by completely removing the notion of duality/consistency.
-The type systems is parametric and ensures decidability of type checking, while allowing for a novel integration of model checking techniques in MPST.
-In this new theory, more protocols and processes can be typed that are guaranteed to be free of deadlocks.
+While our work focuses on \emph{binary} session types, it is worth to discuss related work on Multiparty Session Types (MPST).
+The line of work on MPST starts with Honda~\etal~\cite{hondayoshida08}, which guarantees deadlock freedom \emph{within a single session}, but not for session interleaving.
+Bettini~\etal~\cite{bettinicoppo08} follow a technique similar to Kobayashi's for MPST.
+The main difference with our work is that we associate priorities with communication actions, where Bettini~\etal~\cite{bettinicoppo08} associate them with channels.
+Carbone and Montesi~\cite{carbonemontesi13} combine MPST with choreographies and obtain a formalism that satisfies deadlock freedom by design.
+Deni\'{e}lou and Yoshida~\cite{DenielouY13} introduce \emph{multiparty compatibility} which generalises duality in binary session types.
+They synthesise safe and deadlock-free global types from local types leveraging LTSs and communicating automata.
+Castellani~\etal~\cite{CastellaniDGH20} guarantee lock freedom, a stronger property than deadlock freedom, for MPST with \emph{internal delegation}, where participants in the same session are allowed to delegate tasks to each other, and internal delegation is captured by the global type.
+Scalas and Yoshida~\cite{scalasyoshida19} provide a revision of the foundations for MPST, and offer a less complicated and more general theory, by removing duality/consistency.
+The type systems is parametric and type checking is decidable, but allows for a novel integration of model checking techniques.
+More protocols and processes can be typed and are guaranteed to be free of deadlocks.
 
-Gay~\etal~\cite{gaynagarajan03} and Vasconcelos~\etal~\cite{vasconcelosravara04,vasconcelosgay06} were the first to introduce a functional language with session types.
-However, such works, including early GV~\cite{gayvasconcelos10,gayvasconcelos12} did not guarantee deadlock freedom, until it was addressed via syntactic restrictions~\cite{lindleymorris15,wadler14}.
-Toninho~\etal~\cite{toninhocaires12} present a translation of simply-typed $\lambda$-calculus into session-typed $\pi$-calculus.
-However, their focus is not on deadlock freedom.
+Gay~\etal~\cite{gaynagarajan03} and Vasconcelos~\etal~\cite{vasconcelosravara04,vasconcelosgay06} introduce the first functional language with session types.
+Such works did not guarantee deadlock freedom until GV~\cite{lindleymorris15,wadler14}.
+Toninho~\etal~\cite{toninhocaires12} present a translation of simply-typed $\lambda$-calculus into session-typed $\pi$-calculus, but their focus is not on deadlock freedom.
 
@@
 A~correspondence with linear $\pi$-calculus was established early on by Abramsky~\cite{abramsky94} and Bellin and Scott~\cite{bellinscott94}.
-Many years later, a correspondence between the $\pi$-calculus with binary session types and linear logic was proposed: Caires and Pfenning~\cite[$\pi\text{DILL}$]{cairespfenning10} proposed a correspondence with dual intuitionistic linear logic, while Wadler~\cite[CP]{wadler14} proposed a correspondence with classical linear logic.
-Both works guarantee deadlock freedom as result of their connection to logic and in particular due to the fact that the cut rule only allows tree-structured processes.
-Dardha and Gay~\cite[PCP]{dardhagay18} integrate Kobayashi and Padovani's work on priorities~\cite{kobayashi06,padovani14} with CP, creating the first calculus which combines priorities and logic.
-They prove cycle-elimination, in the same lines as cut-elimination for linear logic, from which the meta-theory and deadlock freedom follow.
+Many years later, correspondences with the $\pi$-calculus with binary session types were proposed.
+Caires and Pfenning~\cite{cairespfenning10} propose a correspondence with dual intuitionistic linear logic, while Wadler~\cite{wadler14} proposes a correspondence with classical linear logic.
+Both guarantee deadlock freedom via their connection to logic, as a consequence of cut elimination.
+Dardha and Gay~\cite{dardhagay18} integrate Kobayashi and Padovani's work on priorities~\cite{kobayashi06,padovani14} with CP, loosening its ties to linear logic in exchange for expressivity.
 Dardha and P\'{e}rez~\cite{dardhaperez15} compare Kobayashi-style typing and CLL typing, and show that CLL corresponds to a subsystem of Kobayashi's where restriction is applied once.
-Balzer~\etal~\cite[$\text{SILL}_S$]{balzerpfenning17} introduce sharing, which breaks deadlock freedom.
-They later restore deadlock freedom using priorities~\cite[$\text{SILL}_{S+}$]{balzertoninho19}.
-Carbone~\etal~\cite{CarboneMSY15,carbonelindley16} give a logical view of MPST with a generalised duality, called \emph{coherence}.
-Caires and P\'{e}rez~\cite{CairesP16} give a novel presentation of MPST in terms of binary session types and the use of a \emph{medium process} which guarantee protocol fidelity and deadlock freedom.
-Their binary session types are rooted in linear logic following the proofs-as-processes correspondences.
-Ciobanu and Horne~\cite{CiobanuH15} give the first processes-as-formulae correspondence of MPST and \emph{the calculus of structures} by \cite{Guglielmi07}, and give a calculus where global and local session types are specified inspired by Scribble\footnote{\url{http://www.scribble.org/}}.
-The calculus is then used to determine whether a set of local session types can be composed safely and in a deadlock free manner.
-Horne \cite{Horne20} presents a proof system for subtyping and multiparty compatibility where compatible processes are race free and deadlock free.
-This work also gives a processes-as-formulae correspondence--similar to the approach in~\cite{CiobanuH15}--with a non-commutative extension of linear logic.
-As such, the meta-theoretical results of this work are obtained in terms of cut-elimination.
+Balzer~\etal~\cite{balzerpfenning17} introduce sharing, which breaks deadlock freedom.
+They later restore deadlock freedom using priorities~\cite{balzertoninho19}.
+Carbone~\etal~\cite{CarboneMSY15,carbonelindley16} give a logical view of MPST with a generalised duality.
+Caires and P\'{e}rez~\cite{CairesP16} give a presentation of MPST in terms of binary session types and the use of a \emph{medium process} which guarantee protocol fidelity and deadlock freedom.
+Their binary session types are rooted in linear logic following the propositions-as-sessions correspondences.
+Ciobanu and Horne~\cite{CiobanuH15} gave the first propositions-as-sessions correspondence of MPST with the \emph{calculus of structures}~\cite{Guglielmi07} with global and local session types are inspired by Scribble, which determines whether local session types can be combined in a deadlock free manner.
+Horne~\cite{Horne20} presents a system for subtyping and multiparty compatibility where compatible processes are race free and deadlock free using a propositions-as-sessions correspondence, similar to the approach in~\cite{CiobanuH15}, with a non-commutative extension of linear logic.
 
@@
 Our formalism so far only captures the core of GV.
-In future work, we plan to explore: recursion in PGV, by integrating the works of Lindley and Morris~\cite{lindleymorris16} and Padovani and Novara~\cite{padovaninovara15}; %a~translation from Priority GV to Priority CP;
-and sharing, following Balzer and Pfenning~\cite{balzerpfenning17}.
+In future work, we plan to explore recursion, following Lindley and Morris~\cite{lindleymorris16} and Padovani and Novara~\cite{padovaninovara15}, and sharing, following Balzer and Pfenning~\cite{balzerpfenning17} or Voinea~\etal~\cite{VoineaDG19}.
+% or in the lines of Kobayashi and Laneve for unbounded networks~\cite{KL17}, which is a refinement of previous work by Kobayashi.
 

### 2021-04-13 conclusion.tex
Moved conclusion to its own section.

@@
 \begin{document}
+
 \section{Related Work}
-\subsubsection*{Deadlock freedom and progress}
+
+\paragraph*{Deadlock freedom and progress}
 Deadlock freedom and progress are well studied properties in the $\pi$-calculus.
@@
 
-\subsubsection*{Ties with logic}
+\paragraph*{Ties with logic}
 The correspondence between logic and types lays the foundation for functional programming~\cite{wadler15}.
@@
 
-\subsubsection*{Conclusion and future work}
+\section{Conclusion and Future Work}
 We answered our research question by presenting Priority GV, a~session-typed functional language which allows cyclic communication structures and uses priorities to ensure deadlock freedom.
@@
 
-\subsubsection*{Acknowledgement}
+\paragraph*{Acknowledgements}
 The authors would like to thank Simon Fowler, April Gon\c{c}alves, and Philip Wadler for their comments on the manuscript.

### 2021-04-13 conclusion.tex
Minor changes.

--- 2021-04-13 conclusion.tex
@@
 \begin{document}
-\section{Related Work}
+
+\section{Related Work and Discussion}
+
 \subsubsection*{Deadlock freedom and progress}
@@
 
-\subsubsection*{Conclusion and future work}
+\subsubsection*{Conclusion}
 We answered our research question by presenting Priority GV, a~session-typed functional language which allows cyclic communication structures and uses priorities to ensure deadlock freedom.
@@
 
+\subsubsection*{Future work}
 Our formalism so far only captures the core of GV.
@@
 
-\subsubsection*{Acknowledgement}
+\subsubsection*{Acknowledgements}
 The authors would like to thank Simon Fowler, April Gon\c{c}alves, and Philip Wadler for their comments on the manuscript.

### 2021-04-13 prf/thm-pgv-closed-progress-confs.tex
Included progress. Removed references to appendices.

@@
   Otherwise, each term $\tm{M_i}$ is ready, and $\tm{N}$ is either ready or a value.
-  Pick the \emph{ready} term $\tm{L}\in\{\tm{M_1},\dots,\tm{M_m},\tm{N}\}$ with the smallest priority bound. There are four cases:
+  Pick the \emph{ready} term $\tm{L}\in\{\tm{M_1},\dots,\tm{M_m},\tm{N}\}$ with the smallest priority bound.
+  % There are four cases:
   \begin{enumerate}
@@
 The typing rules enforce the linear use of endpoints, so there must be a term $\tm{L'}\in\{\tm{M_1},\dots,\tm{M_m},\tm{N}\}$ which uses $\tm{y'}$.
-There are two cases:
+    % There are two cases:
     \begin{enumerate}
@@
       $\tm{L'}$ is ready. By~\cref{lem:pgv-ready-priority}, the priority of $\tm{L}$ is $\pr(\ty{S})$. By duality, $\pr(\ty{\co{S}})=\pr(\ty{S})$.
-
       We cannot have $\tm{L}=\tm{L'}$, otherwise the action on $\tm{y'}$ would be guarded by the action on $\tm{y}$, requiring $\pr(\ty{\co{S}})<\pr(\ty{S})$.
@@
 We then repeat the steps above with the term with the next smallest priority, until either we find a reduction, or the configuration has reached the desired normal form.
-(The argument based on the priority being the smallest continues to hold, since we know that neither $\tm{L}$ nor $\tm{L'}$ will be picked, and no other term uses $\tm{y}$ or $\tm{y'}$.)
+The argument based on the priority being the smallest continues to hold, since we know that neither $\tm{L}$ nor $\tm{L'}$ will be picked, and no other term uses $\tm{y}$ or $\tm{y'}$.
     \end{enumerate}

### 2021-04-13 priority-cp.tex
Included progress. Removed references to appendices.

@@
 \textbf{Types} ($\ty{A}$, $\ty{B}$) in PCP correspond to linear logic connectives annotated with priorities $\cs{o}\in\mathbb{N}$.
-Typing environments, duality, and the priority function $\pr(\cdot)$ are defined as expected (see~\cref{sec:pcp-environments,sec:pcp-duality,sec:pcp-priorities}).
+Typing environments, duality, and the priority function $\pr(\cdot)$ are defined as expected.
+% (see~\cref{sec:pcp-environments,sec:pcp-duality,sec:pcp-priorities}).
 \[
@@
 
-We prove closed progress for our updated PCP. The proof is in \cref{prf:thm-pcp-closed-progress}.
+We prove closed progress for our updated PCP.
+% The proof is in \cref{prf:thm-pcp-closed-progress}.
 \begin{compacttheorems}
@@
 The translations $\tm{\cpgvM{\cdot}}$ and $\tm{\cpgvC{\cdot}}$ preserve typing, and the latter induces a sound and complete operational correspondence.
-The proofs are by structural induction and can be found in~\cref{prf:lem-pcp-to-pgv-terms-preservation,prf:lem-pcp-to-pgv-terms-preservation,prf:lem-pcp-to-pgv-cpgvM-to-cpgvC,prf:thm-pcp-to-pgv-operational-correspondence-soundness,prf:thm-pcp-to-pgv-operational-correspondence-completeness}.
+% The proofs are by structural induction and can be found in~\cref{prf:lem-pcp-to-pgv-terms-preservation,prf:lem-pcp-to-pgv-terms-preservation,prf:lem-pcp-to-pgv-cpgvM-to-cpgvC,prf:thm-pcp-to-pgv-operational-correspondence-soundness,prf:thm-pcp-to-pgv-operational-correspondence-completeness}.
 \begin{compacttheorems}

### 2021-04-13 priority-gv.tex
Examples in figure.
Minor changes.
Included progress. Removed references to appendices.

@@
 The main process does one of two things:
-\begin{enumerate*}[label=(\alph*)]
+\begin{enumerate*}[label= (\alph*) ]
 \item in \cref{ex:cycle}, it sends a unit over the channel $\tm{x}/\tm{x'}$, and then waits to receive a unit over the channel $\tm{y}/\tm{y'}$;
@@
 
-\begin{center}
+\begin{figure}
 \begin{minipage}{0.5\linewidth}
-\begin{example}[Cyclic Structure]
+\begin{example}[Cyclic Structure]%
   \label{ex:cycle}
@@
 \begin{minipage}{0.5\linewidth}
-\begin{example}[Deadlock]
+\begin{example}[Deadlock]%
   \label{ex:deadlock}
@@
 \end{minipage}
-\end{center}
+\end{figure}
 
@@
 \[
-\begin{array}{l}
-  \tm{\select{\ell}}\elabarrow \\
-  \qquad\tm{\lambda x.\letpair{y}{z}{\new}{\andthen{\close\;(\send\;{\pair{\ell\;y}{x}})}{z}}} \\
-  \tm{\offer{L}{x}{M}{y}{N}}\elabarrow \\
-  \qquad\tm{\letpair{z}{w}{\recv\;{L}}{\andthen{\wait\;{w}}{\casesum{z}{x}{M}{y}{N}}}} \\
-  \tm{\offeremp{L}}\elabarrow \\
-  \qquad\tm{\letpair{z}{w}{\recv\;L}{\andthen{\wait\;{w}}{\absurd{z}}}}
+\begin{array}{lcl}
+  \tm{\select{\ell}}
+  & \elabarrow
+  & \tm{\lambda x.\letpair{y}{z}{\new}{\andthen{\close\;(\send\;{\pair{\ell\;y}{x}})}{z}}}
+  \\
+  \multicolumn{3}{l}{%
+  \tm{\offer{L}{x}{M}{y}{N}}\elabarrow}
+  \\
+  \multicolumn{3}{l}{%
+  \qquad\tm{\letpair{z}{w}{\recv\;{L}}{\andthen{\wait\;{w}}{\casesum{z}{x}{M}{y}{N}}}}}
+  \\
+  \tm{\offeremp{L}}
+  & \elabarrow
+  & \tm{\letpair{z}{w}{\recv\;L}{\andthen{\wait\;{w}}{\absurd{z}}}}
 \end{array}
@@
 \[
-\begin{array}[t]{lcl}
   \tm{\phi}
-  & \Coloneqq & \tm{\main}
+  \Coloneqq \tm{\main}
   \sep      \tm{\child}
-  \\
+  \hfill
   \tm{\conf{C}}, \tm{\conf{D}}, \tm{\conf{E}}
-  & \Coloneqq & \tm{\phi\;M}
+  \Coloneqq \tm{\phi\;M}
   \sep      \tm{\ppar{\conf{C}}{\conf{D}}}
   \sep      \tm{\res{x}{x'}{\conf{C}}}
-\end{array}
+  \hfill
 \]
@@
   \tm{\main}  + \tm{\child} = \tm{\main}
-  \quad
+  \hfill
   \tm{\child} + \tm{\main}  = \tm{\main}
-  \quad
+  \hfill
   \tm{\child} + \tm{\child} = \tm{\child}
-  \quad
+  \hfill
   (\tm{\main}  + \tm{\main} \; \text{is undefined})
+  \hfill
 \]
@@
 For this, we prove~\cref{lem:pgv-value-done}, which shows that values have finished all their communication, and that any priorities in the type of the value come from the typing environment.
-The proofs are by structural induction and can be found in \cref{prf:lem-pgv-value-done,prf:lem-pgv-substitution,prf:lem-pgv-subject-reduction-terms,prf:lem-pgv-subject-congruence,prf:thm-pgv-subject-reduction-confs}.
+% The proofs are by structural induction and can be found in \cref{prf:lem-pgv-value-done,prf:lem-pgv-substitution,prf:lem-pgv-subject-reduction-terms,prf:lem-pgv-subject-congruence,prf:thm-pgv-subject-reduction-confs}.
 \begin{compacttheorems}
@@
     \label{def:pgv-canonical-forms}
-    A~configuration $\tm{\conf{C}}$ is in canonical form if it is of the following form, where no term $\tm{M_i}$ is a value:
-    \[
-      \tm{\res{x_1}{x'_1}{\dots\res{x_n}{x'_n}{(\child\;M_1\parallel\dots\parallel\child\;M_m\parallel\main\;N)}}}.
-    \]
-  \end{definition}
-  \begin{definition}[Normal Forms]%
-    A~configuration $\tm{\conf{C}}$ is in normal form if it is of the following form, where each $\tm{M_i}$ is ready to act on $\tm{x_i}$:
-    \[
-      \tm{\res{x_1}{x'_1}{\dots\res{x_n}{x'_n}{(\child\;M_1\parallel\dots\parallel\child\;M_m\parallel\main\;V)}}}.
-    \]
+    A~configuration $\tm{\conf{C}}$ is in canonical form if it is of the form
+    \(
+    \tm{\res{x_1}{x'_1}{\dots\res{x_n}{x'_n}{(\child\;M_1\parallel\dots\parallel\child\;M_m\parallel\main\;N)}}}
+    \)
+    where no term $\tm{M_i}$ is a value.
   \end{definition}
@@
   \end{restatablelemma}
+  \begin{definition}[Normal Forms]%
+    A~configuration $\tm{\conf{C}}$ is in normal form if it is of the form
+    \(
+    \tm{\res{x_1}{x'_1}{\dots\res{x_n}{x'_n}{(\child\;M_1\parallel\dots\parallel\child\;M_m\parallel\main\;V)}}}
+    \)
+    where each $\tm{M_i}$ is ready to act on $\tm{x_i}$.
+  \end{definition}
   \begin{restatabletheorem}{thmpgvclosedprogressconfs}[Progress, $\cred$]%
@@
   \end{restatabletheorem}
-  \begin{proof}
-    Our proof follows the reasoning by Kobayashi~\cite{kobayashi06} used in the proof of deadlock freedom for closed processes (Theorem 2). Details are in~\Cref{prf:thm-pgv-closed-progress-confs}.
+  \begin{proof}[Sketch]
+    Our proof follows the reasoning by Kobayashi~\cite{kobayashi06} used in the proof of deadlock freedom for closed processes (Theorem 2).
+We apply \cref{lem:pgv-open-progress-terms} to each thread.
+Either we obtain a reduction, or each child thread is ready and the main thread ready or a value.
+We pick the ready term $\tm{L}$ with the smallest priority bound.
+If $\tm{L}$ contains new, spawn, or a link, we apply \LabTirName{E-New}, \LabTirName{E-Spawn}, or \LabTirName{E-Link}.
+Otherwise, $\tm{L}$ must be ready on some $\tm{x_i}$.
+Linearity guarantees there is some thread  $\tm{L'}$ which acts on $\tm{x'_i}$.
+If $\tm{L'}$ is ready, priority typing guarantees it is ready on $\tm{x'_i}$, and we apply \LabTirName{E-Send} or \LabTirName{E-Close}.
+If $\tm{L'}$ is not ready, it must be the main thread returning a value.
+We move $\tm{L}$ into the $i^{\text{th}}$ position and repeat until we either find a reduction or reach normal form.
   \end{proof}

### 2021-04-13 introduction.tex
Rewrote intro.

@@
 \section{Introduction}
-Session types have been studied for many programming paradigms, including concurrent and functional paradigms.
-In particular, \emph{binary} session types have been defined for the $\pi$-calculus~\cite{honda93,takeuchihonda94,hondavasconcelos98}, a process calculus for communication and concurrency; and later on for Good Variation~\cite[GV]{wadler15,lindleymorris15}, a session-typed concurrent $\lambda$-calculus.
-Binary session types guarantee communication safety and session fidelity, namely that communication proceeds without any type mismatch and it follows the predefined protocol.
-Another key property in a concurrent setting is \emph{deadlock freedom}, stating that agents communicate without getting stuck in cyclic dependencies.
-Deadlock freedom is a more involving property and binary session types \emph{alone} are not enough to rule out deadlocks in cyclic structures of different interleaved sessions.
-Several techniques have been developed to address deadlock freedom in session-typed $\pi$-calculus and concurrent $\lambda$-calculus.
+Session types~\cite{honda93,takeuchihonda94,hondavasconcelos98} are types for protocols.
+Regular types ensure functions are used according to their specification.
+Session types ensure \emph{communication channels} are used according to their protocols.
+Session types have been studied in many settings.
+For instance, in the $\pi$-calculus~\cite{honda93,takeuchihonda94,hondavasconcelos98}, a foundational calculus for communication and concurrency, and in concurrent $\lambda$-calculi~\cite{gayvasconcelos10}, including the focus of our paper: Good Variation~\cite[GV]{wadler15,lindleymorris15}.
 
-With regards to $\pi$-calculus, a growing line of work leverages the Curry-Howard correspondences between session-typed $\pi$-calculus and intuitionistic~\cite{cairespfenning10} or classical linear logic~\cite{wadler14}, which guarantees deadlock freedom by restricting processes to only trees.
-The tree structure is achieved by merging constructs for parallel composition and channel restriction under the logical cut rule, which forces processes to share \emph{only one} channel for communication.
-Recent developments in the CP~\cite{wadler14} line of work have led to various approaches to decoupling these constructs.
-One way is by maintaining a strong correspondence to classical linear logic and allowing only tree-structured processes, such as in Hypersequent CP~\cite[HCP]{kokkemontesi19popl,kokkemontesi19tlla}, Linear Compositional Choreographies~\cite[LCC]{CarboneMS18} or Classical Transitions~\cite[CT]{MP18}.
-Another way is by weakening the correspondence to classical linear logic in exchange for a more expressive language which allows cyclic-structured processes, such as in Priority CP~\cite[PCP]{dardhagay18}.
-PCP decouples CP's cut rule into two separate constructs: a construct for parallel composition, by introducing a mix rule; and a construct for restriction, by introducing a cycle rule.
-In order to restore deadlock freedom, PCP uses priorities following Kobayashi~\cite{kobayashi06} and Padovani~\cite{padovani14}.
-Priorities establish an \emph{order} of actions and are used to rule out bad interleavings of channels.
-By leveraging priorities, Dardha and Gay prove cycle-elimination of PCP (following the standard cut-elimination proof for classical linear logic).
-Deadlock freedom then follows as a corollary of the cycle-elimination result.
+GV is a concurrent $\lambda$-calculus with \emph{binary} session types, where each channel is shared between exactly two processes.
+Binary session types guarantee two crucial properties \emph{communication safety}---\eg, if the protocol says to transmit an integer, you transmit an integer---and \emph{session fidelity}---\eg, if the protocol says send, you send.
+A third crucial property is \emph{deadlock freedom}, which ensures that processes do not have cyclic dependencies---\eg, when two processes wait for each other to send a value.
+Binary session types \emph{alone} are insufficient to rule out deadlocks arising from interleaved sessions, but several additional techniques have been developed to guarantee deadlock freedom in session-typed $\pi$-calculus and concurrent $\lambda$-calculus.
 
-With regards to GV, the original work on session types in GV~\cite{gayvasconcelos12} did not satisfy deadlock freedom.
-Later on, calculi in the GV family~\cite{wadler15,lindleymorris15} have achieved deadlock freedom via a syntactic restriction, \ie, by combining channel creation and thread spawning into a single operation, called fork.
-The fork construct is the term representation of the cut construct in CP, which as discussed above, evaluates to a channel restriction combined with a parallel composition.
-As with CP, this is restrictive because it limits process structures to only trees.
+In the $\pi$-calculus literature, a growing line of work leverages the Curry-Howard correspondences between session-typed $\pi$-calculus and linear logic~\cite{girard87}: Caires and Pfenning's $\pi$DILL~\cite{cairespfenning10} corresponds to dual intuitionistic linear logic~\cite{barber96}, and Wadler's Classical Processes~\cite[CP]{wadler14} corresponds to classical linear logic~\cite[CLL]{girard87}.
+Both calculi guarantee deadlock freedom, which they achieve by restricting structure of processes and shared channels to \emph{trees}, by combing name restriction and parallel composition into a single construct, corresponding to the logical cut.
+This ensures that two processes can only communicate via exactly one series of channels, which rules out interleavings of sessions, and guarantees deadlock freedom.
+There are many downsides to combining name restriction and parallel composition---lack of modularity, difficulty with typing structural congruence or with label-transition semantics, \etc---which have led to various approaches to decoupling these constructs.
+Hypersequent CP~\cite{MP18,kokkemontesi19popl,kokkemontesi19tlla} and Linear Compositional Choreographies~\cite{CarboneMS18} decouple them, but maintain the correspondence to CLL and allow only tree-structured processes.
+Priority CP~\cite[PCP]{dardhagay18} weakens the correspondence to CLL in exchange for a more expressive language which allows cyclic-structured processes.
+PCP decouples CP's cut rule into two separate constructs: one for parallel composition via a mix rule, and one for name restriction via a cycle rule.
+To restore deadlock freedom, PCP uses \emph{priorities}~\cite{kobayashi06,padovani14}.
+Priorities encode the \emph{order of actions} and rule out bad interleavings.
+Dardha and Gay~\cite{dardhagay18} prove cycle-elimination for PCP, adapting the cut-elimination proof for classical linear logic, and deadlock freedom follows as a corollary.
 
-We aim to further investigate GV and deadlock freedom, as there are several benefits of working with a functional language as opposed to name-passing process calculi.
-Functional languages support higher-order functions, and have a capability for abstraction not usually present in process calculi.
-Working within a functional language allows to derive extensions of the communication capabilities of the language via well-understood extensions of the functional fragment, \ie, deriving internal/external choice via sum types.
-In concurrent functional calculi, there is a clear separation between a program that the user writes, and a configuration, which is the state of a system as it evaluates.
-In process calculi, these are conflated.
-Finally, functional calculi bring us a step closer to transferring our work to real-world (functional) programming languages~\cite{KD21b}.
-Moreover, the benefit of working with GV as opposed to other session-typed functional languages is that GV can be translated into CP~\cite{wadler14}, thus guaranteeing the same formal properties as CP, \eg, deadlock freedom, albeit only for tree-structured processes.
-Following the above observations, we thus pose our research question for GV.%
+CP is related to GV via a pair of translations which satisfy simulation~\cite{lindleymorris16}, and which can be tweaked to satisfy reflection.
+Consequently, the two calculi share the same strong guarantees.
+GV achieves deadlock freedom via a similar syntactic restriction: it combines channel creation and thread spawning into a single operation, called ``fork'', which closely corresponds to the cut construct in CP.
+Unfortunately, as with CP, this syntactic restriction has many downsides.
 
+Our aims are to develop the expressiveness of GV but maintain deadlock freedom.
+We chose GV for several reasons, some of which apply more generally as benefits to working within a concurrent $\lambda$-calculus as opposed to a name-passing process calculus.
+Concurrent $\lambda$-calculi support higher-order functions, and have a capability for abstraction not usually present in process calculi.
+Within a concurrent $\lambda$-calculus, one can derive extensions of the communication capabilities of the language via well-understood extensions of the functional fragment, \eg, we can derive internal/external choice from sum types.
+Concurrent $\lambda$-calculi maintain a clear separation between the program which the user writes and the configurations which represent the state of the system as it evaluates the program.
+However, our main motivation is that results obtained for $\lambda$-calculi transfer more easily to real-world functional programming languages.
+\emph{Case in point}: we were able to embed the theory developed in this paper in Linear Haskell~\cite{bernardyboespflug18} without much effort~\cite{KD21b}.
+For us, these advantages outweigh the benefits of process calculi, \eg, their succinctness.
+
+The benefits of working with GV specifically, as opposed to other concurrent $\lambda$-calculi, is its relation to CP~\cite{wadler14}, and its strong formal properties.
+We thus pose our research question for GV:
+\begin{quotation}
 \textbf{RQ:} \emph{Can we design a more expressive GV which guarantees deadlock freedom for cyclic-structured processes?}
+\end{quotation}
 
-We address this research question by following the line of work from CP to Priority CP, and present
-Priority GV (PGV), a~variant of GV which decouples channel creation from thread spawning, thus allowing cyclic-structured processes and restores deadlock freedom via priorities.
-This closes the circle of the connection between CP and GV~\cite{wadler14}, and their priority-based versions, PCP~\cite{dardhagay18} and PGV.%
-In this paper we make the following main contributions:
+We follow the line of work from CP to Priority CP, and present Priority GV (PGV), a~variant of GV which decouples channel creation from thread spawning, thus allowing cyclic-structured processes, but which nonetheless guarantees deadlock freedom via priorities.
+This closes the circle of the connection between CP and GV~\cite{wadler14}, and their priority-based versions, PCP~\cite{dardhagay18} and PGV.
+We make the following main contributions:
 \begin{enumerate}
 \item \textbf{Priority GV}.
-We present Priority GV (PGV), a session-typed functional language with priorities (\cref{sec:pgv}) and prove subject reduction (\cref{thm:pgv-subject-reduction-confs}) and progress (\cref{thm:pgv-closed-progress-confs}).
-	PGV addresses several problems in the original GV language, most notably:
+We present Priority GV (\secref{sec:pgv}, PGV), a session-typed functional language with priorities, and prove subject reduction (\cref{thm:pgv-subject-reduction-confs}) and progress (\cref{thm:pgv-closed-progress-confs}).
+We addresses several problems in the original GV language, most notably:
   \begin{enumerate*}
-	\item we do not require the pseudo-type $S^\sharp$; and
-	\item our structural congruence is type preserving.
+  \item PGV does not require the pseudo-type $S^\sharp$; and
+  \item its structural congruence is type preserving.
   \end{enumerate*}
-	PGV answers our research question as it allows cyclic-structured binary session-typed processes that are deadlock free.
+  PGV answers our research question positively as it allows cyclic-structured binary session-typed processes that are deadlock free.
 \item \textbf{Updated PCP}.
-We present an updated version of Priority CP~\cite{dardhagay18} in \cref{sec:pcp}.
+We present an updated version of Priority CP~\cite{dardhagay18} (\secref{sec:pcp}).
 We remove commuting conversions and move away from reduction as cut elimination, and present a reduction relation based on structural congruence, as it is standard in process calculi.
 \item \textbf{Translation from PCP to PGV}.
-We present a \emph{sound and complete encoding} of PCP in PGV in~\cref{sec:pcp}, where we prove the encoding preserves typing (\cref{thm:pcp-to-pgv-confs-preservation}) and satisfies operational correspondence (\cref{thm:pcp-to-pgv-operational-correspondence-soundness,thm:pcp-to-pgv-operational-correspondence-completeness}).
+We present a \emph{sound and complete encoding} of PCP in PGV (\secref{sec:pcp}).
+We prove the encoding preserves typing (\cref{thm:pcp-to-pgv-confs-preservation}) and satisfies operational correspondence (\cref{thm:pcp-to-pgv-operational-correspondence-soundness,thm:pcp-to-pgv-operational-correspondence-completeness}).
 \end{enumerate}

### 2021-04-14 introduction.tex
Minor typographical updates.

@@
 \begin{quotation}
-\textbf{RQ:} \emph{Can we design a more expressive GV which guarantees deadlock freedom for cyclic-structured processes?}
+  \textbf{RQ:}
+  Can we design a more expressive GV which guarantees deadlock freedom for cyclic-structured processes?
 \end{quotation}
@@
 We make the following main contributions:
-\begin{enumerate}
-\item \textbf{Priority GV}.
+\begin{enumerate}[labelindent=0pt,labelwidth=1.333\parindent,align=left,labelsep*=0pt,leftmargin=!]
+\item[(\secref{sec:pgv})] \textbf{Priority GV}.
 We present Priority GV (\secref{sec:pgv}, PGV), a session-typed functional language with priorities, and prove subject reduction (\cref{thm:pgv-subject-reduction-confs}) and progress (\cref{thm:pgv-closed-progress-confs}).
@@
   PGV answers our research question positively as it allows cyclic-structured binary session-typed processes that are deadlock free.
-\item \textbf{Updated PCP}.
-We present an updated version of Priority CP~\cite{dardhagay18} (\secref{sec:pcp}).
-We remove commuting conversions and move away from reduction as cut elimination, and present a reduction relation based on structural congruence, as it is standard in process calculi.
-\item \textbf{Translation from PCP to PGV}.
-We present a \emph{sound and complete encoding} of PCP in PGV (\secref{sec:pcp}).
+\item[(\secref{sec:pcp})] \textbf{Translation from PCP to PGV}.
+  We present a \emph{sound and complete encoding} of Priority CP~\cite{dardhagay18} in PGV (\secref{sec:pcp}).
 We prove the encoding preserves typing (\cref{thm:pcp-to-pgv-confs-preservation}) and satisfies operational correspondence (\cref{thm:pcp-to-pgv-operational-correspondence-soundness,thm:pcp-to-pgv-operational-correspondence-completeness}).
+  To obtain a tight correspondence, we update PCP, moving away from commuting conversions and reduction as cut elimination towards reduction based on structural congruence, as it is standard in process calculi.
 \end{enumerate}

### 2021-04-14 appendix.tex
Consistently reference extended versions.

@@
 Types $\pcp{\ty{\typlus[\cs{o}]{A}{B}}}$ and $\pcp{\ty{\tywith[\cs{o}]{A}{B}}}$ type the endpoints of a channel over which we can receive or send a choice between two branches $\pcp{\ty{A}}$ or $\pcp{\ty{B}}$.
-We have opted for a simplified version of choice and followed the original Wadler's CP \cite{wadler14}, however types $\ty{\oplus}$ and $\ty{\with}$ can be trivially generalised to $\pcp{\ty{\oplus^{\cs{o}}\{l_i:A_i\}_{i\in I}}}$ and $\pcp{\ty{\with^{\cs{o}}\{l_i:A_i\}_{i\in I}}}$, respectively, as in the original PCP \cite{dardhagay18}.
+We have opted for a simplified version of choice and followed the original Wadler's CP \cite{wadler14}, however types $\ty{\oplus}$ and $\ty{\with}$ can be trivially generalised to $\pcp{\ty{\oplus^{\cs{o}}\{l_i:A_i\}_{i\in I}}}$ and $\pcp{\ty{\with^{\cs{o}}\{l_i:A_i\}_{i\in I}}}$, respectively, as in the original PCP \cite{dardhagay18extended}.
 Types $\pcp{\ty{\tynil[\cs{o}]}}$ and $\pcp{\ty{\tytop[\cs{o}]}}$ type the endpoints of a channel over which we can send or receive a choice between \emph{no options}.
@@
 The main change we make to PCP is \emph{removing commuting conversions}.
-Commuting conversions are necessary if we want our reduction strategy to correspond \emph{exactly} to cut (or cycle in \cite{dardhagay18}) elimination.
+Commuting conversions are necessary if we want our reduction strategy to correspond \emph{exactly} to cut (or cycle in \cite{dardhagay18extended}) elimination.
 However, as Lindley and Morris~\cite{lindleymorris15} show, all communications that can be performed \emph{with} the use of commuting conversions, can also be performed \emph{without} them, but using structural congruence.
@@
 In PCP, commuting conversions break our intuition that an action with lower priority occurs before an action with higher priority.
-To cite Dardha and Gay~\cite{dardhagay18} ``\emph{if a prefix on a channel endpoint $\pcp{\tm{x}}$ with priority $\cs{o}$ is pulled out at top level, then to preserve priority constraints in the typing rules [..], it is necessary to increase priorities of all actions after the prefix on $\pcp{\tm{x}}$}'' by $\cs{o+1}$.
+To cite Dardha and Gay~\cite{dardhagay18extended} ``\emph{if a prefix on a channel endpoint $\pcp{\tm{x}}$ with priority $\cs{o}$ is pulled out at top level, then to preserve priority constraints in the typing rules [..], it is necessary to increase priorities of all actions after the prefix on $\pcp{\tm{x}}$}'' by $\cs{o+1}$.
 
@@
 
-Finally, since our reduction relation is a strict subset of the reduction relation in the original~\cite{dardhagay18}, we defer to their proofs.
+Finally, since our reduction relation is a strict subset of the reduction relation in the original~\cite{dardhagay18extended}, we defer to their proofs.
 We prove progress for our version of PCP, see~\cref{prf:thm-pcp-closed-progress}.
@@
 \label{sec:milner}
-As an example of a deadlock-free cyclic process, Dardha and Gay~\cite{dardhagay18} introduce an implementation of Milner's cyclic scheduler~\cite{milner89} in Priority CP.
+As an example of a deadlock-free cyclic process, Dardha and Gay~\cite{dardhagay18extended} introduce an implementation of Milner's cyclic scheduler~\cite{milner89} in Priority CP.
 We reproduce that scheduler here, and show its translation to Priority GV.

### 2021-04-14 conclusion.tex
Consistently reference extended versions.

@@
 Both guarantee deadlock freedom via their connection to logic, as a consequence of cut elimination.
-Dardha and Gay~\cite{dardhagay18} integrate Kobayashi and Padovani's work on priorities~\cite{kobayashi06,padovani14} with CP, loosening its ties to linear logic in exchange for expressivity.
-Dardha and P\'{e}rez~\cite{dardhaperez15} compare Kobayashi-style typing and CLL typing, and show that CLL corresponds to a subsystem of Kobayashi's where restriction is applied once.
+Dardha and Gay~\cite{dardhagay18extended} integrate Kobayashi and Padovani's work on priorities~\cite{kobayashi06,padovani14} with CP, loosening its ties to linear logic in exchange for expressivity.
+Dardha and P\'{e}rez~\cite{dardhaperez15extended} compare Kobayashi-style typing and CLL typing, and show that CLL corresponds to a subsystem of Kobayashi's where restriction is applied once.
 Balzer~\etal~\cite{balzerpfenning17} introduce sharing, which breaks deadlock freedom.
@@
 We answered our research question by presenting Priority GV, a~session-typed functional language which allows cyclic communication structures and uses priorities to ensure deadlock freedom.
-We showed its relation to Priority CP~\cite{dardhagay18} via an operational correspondence.
+We showed its relation to Priority CP~\cite{dardhagay18extended} via an operational correspondence.
 

### 2021-04-14 introduction.tex
Consistently reference extended versions.

@@
 Session types have been studied in many settings.
-For instance, in the $\pi$-calculus~\cite{honda93,takeuchihonda94,hondavasconcelos98}, a foundational calculus for communication and concurrency, and in concurrent $\lambda$-calculi~\cite{gayvasconcelos10}, including the focus of our paper: Good Variation~\cite[GV]{wadler15,lindleymorris15}.
+For instance, in the $\pi$-calculus~\cite{honda93,takeuchihonda94,hondavasconcelos98}, a foundational calculus for communication and concurrency, and in concurrent $\lambda$-calculi~\cite{gayvasconcelos12}, including the focus of our paper: Good Variation~\cite[GV]{wadler15,lindleymorris15}.
 
@@
 Hypersequent CP~\cite{MP18,kokkemontesi19popl,kokkemontesi19tlla} and Linear Compositional Choreographies~\cite{CarboneMS18} decouple them, but maintain the correspondence to CLL and allow only tree-structured processes.
-Priority CP~\cite[PCP]{dardhagay18} weakens the correspondence to CLL in exchange for a more expressive language which allows cyclic-structured processes.
+Priority CP~\cite[PCP]{dardhagay18extended} weakens the correspondence to CLL in exchange for a more expressive language which allows cyclic-structured processes.
 PCP decouples CP's cut rule into two separate constructs: one for parallel composition via a mix rule, and one for name restriction via a cycle rule.
@@
 Priorities encode the \emph{order of actions} and rule out bad interleavings.
-Dardha and Gay~\cite{dardhagay18} prove cycle-elimination for PCP, adapting the cut-elimination proof for classical linear logic, and deadlock freedom follows as a corollary.
+Dardha and Gay~\cite{dardhagay18extended} prove cycle-elimination for PCP, adapting the cut-elimination proof for classical linear logic, and deadlock freedom follows as a corollary.
 
@@
 We follow the line of work from CP to Priority CP, and present Priority GV (PGV), a~variant of GV which decouples channel creation from thread spawning, thus allowing cyclic-structured processes, but which nonetheless guarantees deadlock freedom via priorities.
-This closes the circle of the connection between CP and GV~\cite{wadler14}, and their priority-based versions, PCP~\cite{dardhagay18} and PGV.
+This closes the circle of the connection between CP and GV~\cite{wadler14}, and their priority-based versions, PCP~\cite{dardhagay18extended} and PGV.
 We make the following main contributions:
@@
 \item[(\secref{sec:pcp})] \textbf{Translation from PCP to PGV}.
-  We present a \emph{sound and complete encoding} of Priority CP~\cite{dardhagay18} in PGV (\secref{sec:pcp}).
+  We present a \emph{sound and complete encoding} of Priority CP~\cite{dardhagay18extended} in PGV (\secref{sec:pcp}).
 We prove the encoding preserves typing (\cref{thm:pcp-to-pgv-confs-preservation}) and satisfies operational correspondence (\cref{thm:pcp-to-pgv-operational-correspondence-soundness,thm:pcp-to-pgv-operational-correspondence-completeness}).

### 2021-04-14 priority-cp.tex
Consistently reference extended versions.

@@
 
-We present a correspondence between Priority GV and an updated version of Priority CP~\cite[PCP]{dardhagay18}, which is Wadler's CP~\cite{wadler14} with priorities.
+We present a correspondence between Priority GV and an updated version of Priority CP~\cite[PCP]{dardhagay18extended}, which is Wadler's CP~\cite{wadler14} with priorities.
 This correspondence connects PGV to (a~relaxed variant of) classical linear logic.
@@
 
-PCP decomposes cut into \LabTirName{T-Res} and \LabTirName{T-Par} rules---logically corresponding to cycle and mix rules, respectively---and guarantees deadlock freedom by using priority constraints, \eg, as in \LabTirName{T-Send}.
-The full typing rules and operational semantics are available in \cref{fig:pcp-operational-semantics,fig:pcp-typing} in \cref{app:revisiting-PCP}:
+PCP decomposes cut into \LabTirName{T-Res} and \LabTirName{T-Par} rules---corresponding to cycle and mix rules, respectively---and guarantees deadlock freedom by using priority constraints, \eg, as in \LabTirName{T-Send}.
+% The full typing rules and operational semantics are available in \cref{fig:pcp-operational-semantics,fig:pcp-typing} in \cref{app:revisiting-PCP}:
 \begin{mathpar}
@@
 Commuting conversions are necessary if we want our reduction strategy to correspond \emph{exactly} to cut elimination.
-However, from the perspective of process calculi, commuting conversions behave strangely: they allow an input/output action to be moved to the top of a process, thus potentially blocking actions which were previously possible (an example is given in \cref{app:commuting-conversions}).
-This makes CP, and Dardha and Gay's PCP~\cite{dardhagay18}, non-confluent.
-As Lindley and Morris~\cite{lindleymorris15} show, all communications that can be performed \emph{with} the use of commuting conversions, can also be performed \emph{without} them, if using structural congruence.
+However, from the perspective of process calculi, commuting conversions behave strangely: they allow an input/output action to be moved to the top of a process, thus potentially blocking actions which were previously possible.
+This makes CP, and Dardha and Gay's PCP~\cite{dardhagay18extended}, non-confluent.
+As Lindley and Morris~\cite{lindleymorris15} show, all communications that can be performed \emph{with} the use of commuting conversions, can also be performed \emph{without} them, using structural congruence.
+
 %Consider the commuting conversion $(\kappa_{\parr})$ for $\pcp{\tm{\recv{x}{y}{P}}}$:
@@
 %\end{mathpar}
-%
 
 In  particular for PCP, commuting conversions break our intuition that an action with lower priority \emph{occurs before} an action with higher priority.
-To cite Dardha and Gay~\cite{dardhagay18} ``\emph{if a prefix on a channel endpoint $\pcp{\tm{x}}$ with priority $\cs{o}$ is pulled out at top level, then to preserve priority constraints in the typing rules [..], it is necessary to increase priorities of all actions after the prefix on $\pcp{\tm{x}}$}'' by $\cs{o+1}$.
+To cite Dardha and Gay~\cite{dardhagay18extended} ``\emph{if a prefix on a channel endpoint $\pcp{\tm{x}}$ with priority $\cs{o}$ is pulled out at top level, then to preserve priority constraints in the typing rules [..], it is necessary to increase priorities of all actions after the prefix on $\pcp{\tm{x}}$}'' by $\cs{o+1}$.
 One benefit of removing commuting conversions is that we no longer need to dynamically change the priorities during reduction, which means that the intuition for priorities holds true in our updated version of PCP.

### 2021-04-14 priority-gv.tex
Rewrote intro.
Consistently reference extended versions.

@@
 \]
-The use of $\tm{\child}$ for child threads~\cite{lindleymorris15} overlaps with the use of the meta-variable $\cs{o}$ for priorities~\cite{dardhagay18}.
+The use of $\tm{\child}$ for child threads~\cite{lindleymorris15} overlaps with the use of the meta-variable $\cs{o}$ for priorities~\cite{dardhagay18extended}.
 Both are used to annotate sequents: flags appear on the sequent in configuration typing, and priorities in term typing.
@@
 The integration of priorities into GV is adapted from Padovani and Novara~\cite{padovaninovara15}.
-Paraphrasing Dardha and Gay~\cite{dardhagay18}, priorities obey the following two laws:
+Paraphrasing Dardha and Gay~\cite{dardhagay18extended}, priorities obey the following two laws:
 \begin{enumerate*}[label= (\roman*) ]
@@
   If $\tseq[\cs{p}]{\ty{\Gamma}}{M}{T}$ and $\ty{\Gamma}$ contains only session types, then:
-  \begin{enumerate*}[label=\emph{(\alph*)}]
+  \begin{enumerate*}[label= (\alph*) ]
   \item $\tm{M}$ is a value;

### 2021-04-14 priority-gv.tex
Change layout of echo server.

@@
 Let us consider an example, adapted from Lindley and Morris~\cite{lindleymorris15}, in which a term returns an endpoint linked to an echo server.
-The echo server receives a value and sends it back unchanged:
+The echo server receives a value and sends it back unchanged.
+Consider the program which creates a new channel, with endpoints $\tm{x}$ and $\tm{x'}$, spawns off an echo server listening on $\tm{x}$, and then returns $\tm{x'}$:
 \[
-  \tm{\echo_x} \defeq \tm{\letpair{y}{x}{\recv\;{x}}{\letbind{x}{\send\;\pair{y}{x}}{\close\;x}}}
+  \begin{array}{ll}
+    \tm{\main}
+    & \tm{\letpair{x}{x'}{\new}{}}
+    \\
+    & \tm{\andthen{\spawn\;(\lambda\unit.\echo_{x})}{x'}}
+  \end{array}
+  \hfill
+  \begin{array}{lcl}
+    \tm{\echo_x}
+    & \defeq
+    & \tm{\letpair{y}{x}{\recv\;{x}}{}}
+    \\
+    &
+    & \tm{\letbind{x}{\send\;\pair{y}{x}}{\close\;x}}
+  \end{array}
+  \hfill
 \]
-Consider the program which creates a new channel, with endpoints $\tm{x}$ and $\tm{x'}$, spawns off an echo server listening on $\tm{x'}$, and then returns $\tm{x}$:
-\[
-  \tm{\main\;\letpair{x}{x'}{\new}{\andthen{\spawn\;(\lambda\unit.\echo_{x'})}}{x}}
-\]
-If we reduce the above program, we get $\tm{\res{x}{x'}{(\ppar{\main\;x}{\child\;\echo_{x'}})}}$.
-Clearly, no more evaluation is possible, even though the configuration contains the thread $\tm{\child\;\echo_{x'}}$, which is blocked on $\tm{x'}$.
+If we reduce the above program, we get $\tm{\res{x}{x'}{(\ppar{\child\;\echo_{x}}{\main\;x'})}}$.
+Clearly, no more evaluation is possible, even though the configuration contains the thread $\tm{\child\;\echo_{x}}$, which is blocked on $\tm{x}$.
 In~\cref{cor:pgv-closed-progress} we will show that if a term does not return an endpoint, it must produce \emph{only} a value.

### 2021-04-14 introduction.tex
Minor changes in introduction.

@@
 This ensures that two processes can only communicate via exactly one series of channels, which rules out interleavings of sessions, and guarantees deadlock freedom.
-There are many downsides to combining name restriction and parallel composition---lack of modularity, difficulty with typing structural congruence or with label-transition semantics, \etc---which have led to various approaches to decoupling these constructs.
+There are many downsides to combining name restriction and parallel composition---lack of modularity, difficulty typing structural congruence and formulating label-transition semantics, \etc---which have led to various approaches to decoupling these constructs.
 Hypersequent CP~\cite{MP18,kokkemontesi19popl,kokkemontesi19tlla} and Linear Compositional Choreographies~\cite{CarboneMS18} decouple them, but maintain the correspondence to CLL and allow only tree-structured processes.
@@
 
-CP is related to GV via a pair of translations which satisfy simulation~\cite{lindleymorris16}, and which can be tweaked to satisfy reflection.
-Consequently, the two calculi share the same strong guarantees.
-GV achieves deadlock freedom via a similar syntactic restriction: it combines channel creation and thread spawning into a single operation, called ``fork'', which closely corresponds to the cut construct in CP.
+CP and GV are related via a pair of translations which satisfy simulation~\cite{lindleymorris16}, and which can be tweaked to satisfy reflection.
+The two calculi share the same strong guarantees.
+GV achieves deadlock freedom via a similar syntactic restriction: it combines channel creation and thread spawning into a single operation, called ``fork'', which is related to the cut construct in CP.
 Unfortunately, as with CP, this syntactic restriction has many downsides.

### 2021-04-14 priority-gv.tex
Minor changes in introduction.

@@
 \emph{Actions} are terms which perform communication actions and which synchronise between two threads.
+\emph{Ready terms} are terms which perform communication actions, either by themselves, \eg creating a new channel or thread, or with another thread, \eg sending or receiving.
+Progress for the term language is standard for GV, and deviates from progress for linear $\lambda$-calculus only in that terms may reduce to values or \emph{ready terms}:
+\begin{compacttheorems}
   \begin{definition}[Actions]%
@@
   \end{definition}
-
-\emph{Ready terms} are terms which are ready to perform a communication action, either by themselves, \eg creating a new channel or thread, or with another thread, \eg sending or receiving.
   \begin{definition}[Ready Terms]%
@@
   \end{definition}
-
-Progress for the term language is standard for GV, and deviates from progress for linear $\lambda$-calculus only in that terms may reduce to values or \emph{ready terms}:
   \begin{restatablelemma}{lempgvopenprogressterms}[Progress, $\tred$]%
@@
   \end{restatablelemma}
+\end{compacttheorems}
 

### 2021-04-14 priority-gv.tex
Minor changes.

@@
   \begin{proof}[Sketch]
-    Our proof follows the reasoning by Kobayashi~\cite{kobayashi06} used in the proof of deadlock freedom for closed processes (Theorem 2).
+    Our proof follows that of Kobayashi~\cite[theorem 2]{kobayashi06}.
 We apply \cref{lem:pgv-open-progress-terms} to each thread.
@@
     \label{cor:pgv-closed-progress}
-    If $\cseq[\main]{\emptyenv}{\conf{C}}$, $\tm{\conf{C}}\centernot\cred$, and $\tm{\conf{C}}$ contains no endpoints, then $\tm{\conf{C}}\equiv\tm{\phi\;V}$ for some value $\tm{V}$.
+    If $\cseq[\phi]{\emptyenv}{\conf{C}}$, $\tm{\conf{C}}\centernot\cred$, and $\tm{\conf{C}}$ contains no endpoints, then $\tm{\conf{C}}\equiv\tm{\phi\;V}$ for some value $\tm{V}$.
   \end{corollary}

### 2021-04-14 conclusion.tex
Added remarks about GV->PGV.

@@
 A~correspondence with linear $\pi$-calculus was established early on by Abramsky~\cite{abramsky94} and Bellin and Scott~\cite{bellinscott94}.
-Many years later, correspondences with the $\pi$-calculus with binary session types were proposed.
+Many years later, several correspondences between linear logic and the $\pi$-calculus with binary session types were proposed.
 Caires and Pfenning~\cite{cairespfenning10} propose a correspondence with dual intuitionistic linear logic, while Wadler~\cite{wadler14} proposes a correspondence with classical linear logic.
-Both guarantee deadlock freedom via their connection to logic, as a consequence of cut elimination.
+Both guarantee deadlock freedom as a consequence of cut elimination.
 Dardha and Gay~\cite{dardhagay18extended} integrate Kobayashi and Padovani's work on priorities~\cite{kobayashi06,padovani14} with CP, loosening its ties to linear logic in exchange for expressivity.
-Dardha and P\'{e}rez~\cite{dardhaperez15extended} compare Kobayashi-style typing and CLL typing, and show that CLL corresponds to a subsystem of Kobayashi's where restriction is applied once.
-Balzer~\etal~\cite{balzerpfenning17} introduce sharing, which breaks deadlock freedom.
-They later restore deadlock freedom using priorities~\cite{balzertoninho19}.
+Dardha and P\'{e}rez~\cite{dardhaperez15extended} compare priorities \`a la Kobayashi with tree restrictions \`a la CP, and show that the latter is a subsystem of the former.
+Balzer~\etal~\cite{balzerpfenning17} introduce sharing at the cost of deadlock freedom, which they restore using an approach similar to priorities~\cite{balzertoninho19}.
 Carbone~\etal~\cite{CarboneMSY15,carbonelindley16} give a logical view of MPST with a generalised duality.
 Caires and P\'{e}rez~\cite{CairesP16} give a presentation of MPST in terms of binary session types and the use of a \emph{medium process} which guarantee protocol fidelity and deadlock freedom.
-Their binary session types are rooted in linear logic following the propositions-as-sessions correspondences.
-Ciobanu and Horne~\cite{CiobanuH15} gave the first propositions-as-sessions correspondence of MPST with the \emph{calculus of structures}~\cite{Guglielmi07} with global and local session types are inspired by Scribble, which determines whether local session types can be combined in a deadlock free manner.
-Horne~\cite{Horne20} presents a system for subtyping and multiparty compatibility where compatible processes are race free and deadlock free using a propositions-as-sessions correspondence, similar to the approach in~\cite{CiobanuH15}, with a non-commutative extension of linear logic.
+Their binary session types are rooted in linear logic.
+Ciobanu and Horne~\cite{CiobanuH15} give the first Curry-Howard correspondence between MPST and BV~\cite{Guglielmi07}, a conservative extension of linear logic with a non-commutative operator for sequencing.
+Horne~\cite{Horne20} give a system for subtyping and multiparty compatibility where compatible processes are race free and deadlock free using a Curry-Howard correspondence, similar to the approach in~\cite{CiobanuH15}.
 

### 2021-04-14 introduction.tex
Minor changes.
Added remarks about GV->PGV.

@@
 
-In the $\pi$-calculus literature, a growing line of work leverages the Curry-Howard correspondences between session-typed $\pi$-calculus and linear logic~\cite{girard87}: Caires and Pfenning's $\pi$DILL~\cite{cairespfenning10} corresponds to dual intuitionistic linear logic~\cite{barber96}, and Wadler's Classical Processes~\cite[CP]{wadler14} corresponds to classical linear logic~\cite[CLL]{girard87}.
+In the $\pi$-calculus literature, there is a growing line of work in developing Curry-Howard correspondences between session-typed $\pi$-calculus and linear logic~\cite{girard87}: Caires and Pfenning's $\pi$DILL~\cite{cairespfenning10} corresponds to dual intuitionistic linear logic~\cite{barber96}, and Wadler's Classical Processes~\cite[CP]{wadler14} corresponds to classical linear logic~\cite[CLL]{girard87}.
 Both calculi guarantee deadlock freedom, which they achieve by restricting structure of processes and shared channels to \emph{trees}, by combing name restriction and parallel composition into a single construct, corresponding to the logical cut.
@@
 For us, these advantages outweigh the benefits of process calculi, \eg, their succinctness.
-
 The benefits of working with GV specifically, as opposed to other concurrent $\lambda$-calculi, is its relation to CP~\cite{wadler14}, and its strong formal properties.

### 2021-04-14 priority-gv.tex
Added remarks about GV->PGV.

@@
 \end{enumerate*}
-We will show that only \cref{ex:cycle} is typeable in PGV, while it is not typable in GV~\cite{wadler14} nor guaranteed to be deadlock-free by its predecessor~\cite{gayvasconcelos12}.
+PGV is more expressive than GV: \cref{ex:cycle} is typeable and guaranteed deadlock-free in PGV, but is not typeable in GV~\cite{wadler14} and not guaranteed deadlock-free in GV's predecessor~\cite{gayvasconcelos12}.
+We believe PGV is a non-conservative \emph{extension} of GV, as CP can be embedded in a Kobayashi-style system~\cite{dardhaperez15extended}.
 

### 2021-04-16 introduction.tex
Fixed bib.

@@
 However, our main motivation is that results obtained for $\lambda$-calculi transfer more easily to real-world functional programming languages.
-\emph{Case in point}: we were able to embed the theory developed in this paper in Linear Haskell~\cite{bernardyboespflug18} without much effort~\cite{KD21b}.
+\emph{Case in point}: we were able to embed the theory developed in this paper in Linear Haskell~\cite{bernardyboespflug18} without much effort~\cite{kokkedardha21hs}.
 For us, these advantages outweigh the benefits of process calculi, \eg, their succinctness.

### 2021-04-20 introduction.tex
Update introduction.tex

--- 2021-04-14 introduction.tex
@@
 This ensures that two processes can only communicate via exactly one series of channels, which rules out interleavings of sessions, and guarantees deadlock freedom.
-There are many downsides to combining name restriction and parallel composition---lack of modularity, difficulty typing structural congruence and formulating label-transition semantics, \etc---which have led to various approaches to decoupling these constructs.
+There are many downsides to combining name restriction and parallel composition, such as lack of modularity, difficulty typing structural congruence and formulating label-transition semantics, which have led to various approaches to decoupling these constructs.
 Hypersequent CP~\cite{MP18,kokkemontesi19popl,kokkemontesi19tlla} and Linear Compositional Choreographies~\cite{CarboneMS18} decouple them, but maintain the correspondence to CLL and allow only tree-structured processes.
@@
 To restore deadlock freedom, PCP uses \emph{priorities}~\cite{kobayashi06,padovani14}.
-Priorities encode the \emph{order of actions} and rule out bad interleavings.
+Priorities encode the \emph{order of actions} and rule out bad cyclic interleavings.
 Dardha and Gay~\cite{dardhagay18extended} prove cycle-elimination for PCP, adapting the cut-elimination proof for classical linear logic, and deadlock freedom follows as a corollary.
@@
 GV achieves deadlock freedom via a similar syntactic restriction: it combines channel creation and thread spawning into a single operation, called ``fork'', which is related to the cut construct in CP.
-Unfortunately, as with CP, this syntactic restriction has many downsides.
+Unfortunately, as with CP, this syntactic restriction has its downsides.
 
-Our aims are to develop the expressiveness of GV but maintain deadlock freedom.
+Our aim is to develop a more expressive version of GV while maintaining deadlock freedom.
 We chose GV for several reasons, some of which apply more generally as benefits to working within a concurrent $\lambda$-calculus as opposed to a name-passing process calculus.
@@
 However, our main motivation is that results obtained for $\lambda$-calculi transfer more easily to real-world functional programming languages.
-\emph{Case in point}: we were able to embed the theory developed in this paper in Linear Haskell~\cite{bernardyboespflug18} without much effort~\cite{KD21b}.
-For us, these advantages outweigh the benefits of process calculi, \eg, their succinctness.
-The benefits of working with GV specifically, as opposed to other concurrent $\lambda$-calculi, is its relation to CP~\cite{wadler14}, and its strong formal properties.
+Case in point: we embed the theory developed in this paper in Linear Haskell~\cite{bernardyboespflug18} where PGV easily translates into a library for deadlock free session typed programming~\cite{kokkedardha21hs}.
+%For us, these advantages outweigh the benefits of process calculi, \eg, their succinctness.
+The benefit of working specifically with GV, as opposed to other concurrent $\lambda$-calculi, is its relation to CP~\cite{wadler14}, and its formal properties, including deadlock freedom.
 We thus pose our research question for GV:

### 2021-04-20 introduction.tex
Minor changes.

@@
 However, our main motivation is that results obtained for $\lambda$-calculi transfer more easily to real-world functional programming languages.
-Case in point: we embed the theory developed in this paper in Linear Haskell~\cite{bernardyboespflug18} where PGV easily translates into a library for deadlock free session typed programming~\cite{kokkedardha21hs}.
+Case in point: we easily adapted the type system of PGV to Linear Haskell~\cite{bernardyboespflug18}, which gives us a library for deadlock-free session-typed programming~\cite{kokkedardha21hs}.
 %For us, these advantages outweigh the benefits of process calculi, \eg, their succinctness.

### 2021-04-20 introduction.tex
Minor changes.

@@
 
-In the $\pi$-calculus literature, there is a growing line of work in developing Curry-Howard correspondences between session-typed $\pi$-calculus and linear logic~\cite{girard87}: Caires and Pfenning's $\pi$DILL~\cite{cairespfenning10} corresponds to dual intuitionistic linear logic~\cite{barber96}, and Wadler's Classical Processes~\cite[CP]{wadler14} corresponds to classical linear logic~\cite[CLL]{girard87}.
+In the $\pi$-calculus literature, there have been several attempts at developing Curry-Howard correspondences between session-typed $\pi$-calculus and linear logic~\cite{girard87}: Caires and Pfenning's $\pi$DILL~\cite{cairespfenning10} corresponds to dual intuitionistic linear logic~\cite{barber96}, and Wadler's Classical Processes~\cite[CP]{wadler14} corresponds to classical linear logic~\cite[CLL]{girard87}.
 Both calculi guarantee deadlock freedom, which they achieve by restricting structure of processes and shared channels to \emph{trees}, by combing name restriction and parallel composition into a single construct, corresponding to the logical cut.

### 2021-04-20 introduction.tex
Final changes?

@@
 Our aim is to develop a more expressive version of GV while maintaining deadlock freedom.
-We chose GV for several reasons, some of which apply more generally as benefits to working within a concurrent $\lambda$-calculus as opposed to a name-passing process calculus.
-Concurrent $\lambda$-calculi support higher-order functions, and have a capability for abstraction not usually present in process calculi.
+While process calculi have their advantages, \eg, their succinctness, we chose to work with GV for several reasons.
+In general, concurrent $\lambda$-calculi support higher-order functions, and have a capability for abstraction not usually present in process calculi.
 Within a concurrent $\lambda$-calculus, one can derive extensions of the communication capabilities of the language via well-understood extensions of the functional fragment, \eg, we can derive internal/external choice from sum types.
@@
 Case in point: we easily adapted the type system of PGV to Linear Haskell~\cite{bernardyboespflug18}, which gives us a library for deadlock-free session-typed programming~\cite{kokkedardha21hs}.
-%For us, these advantages outweigh the benefits of process calculi, \eg, their succinctness.
 The benefit of working specifically with GV, as opposed to other concurrent $\lambda$-calculi, is its relation to CP~\cite{wadler14}, and its formal properties, including deadlock freedom.

### 2021-04-26 conclusion.tex
Update conclusion.tex
Final version.

@@
 The main difference with our work is that we associate priorities with communication actions, where Bettini~\etal~\cite{bettinicoppo08} associate them with channels.
-Carbone and Montesi~\cite{carbonemontesi13} combine MPST with choreographies and obtain a formalism that satisfies deadlock freedom by design.
+Carbone and Montesi~\cite{carbonemontesi13} combine MPST with choreographies and obtain a formalism that satisfies deadlock freedom.
 Deni\'{e}lou and Yoshida~\cite{DenielouY13} introduce \emph{multiparty compatibility} which generalises duality in binary session types.
@@
 
-Gay~\etal~\cite{gaynagarajan03} and Vasconcelos~\etal~\cite{vasconcelosravara04,vasconcelosgay06} introduce the first functional language with session types.
+Neubauer and Thiemann~\cite{neubauert04} and Vasconcelos~\etal~\cite{vasconcelosravara04,vasconcelosgay06} introduce the first functional language with session types.
 Such works did not guarantee deadlock freedom until GV~\cite{lindleymorris15,wadler14}.
@@
 In future work, we plan to explore recursion, following Lindley and Morris~\cite{lindleymorris16} and Padovani and Novara~\cite{padovaninovara15}, and sharing, following Balzer and Pfenning~\cite{balzerpfenning17} or Voinea~\etal~\cite{VoineaDG19}.
-% or in the lines of Kobayashi and Laneve for unbounded networks~\cite{KL17}, which is a refinement of previous work by Kobayashi.
 

### 2021-04-26 introduction.tex
Final version.

@@
 Session types have been studied in many settings.
-For instance, in the $\pi$-calculus~\cite{honda93,takeuchihonda94,hondavasconcelos98}, a foundational calculus for communication and concurrency, and in concurrent $\lambda$-calculi~\cite{gayvasconcelos12}, including the focus of our paper: Good Variation~\cite[GV]{wadler15,lindleymorris15}.
+For instance, in the $\pi$-calculus~\cite{honda93,takeuchihonda94,hondavasconcelos98}, a foundational calculus for communication and concurrency, and in concurrent $\lambda$-calculi~\cite{gayvasconcelos10}, including the focus of our paper: Good Variation~\cite[GV]{wadler15,lindleymorris15}.
 

### 2021-04-26 priority-gv.tex
Final version.

@@
 \end{enumerate*}
-PGV is more expressive than GV: \cref{ex:cycle} is typeable and guaranteed deadlock-free in PGV, but is not typeable in GV~\cite{wadler14} and not guaranteed deadlock-free in GV's predecessor~\cite{gayvasconcelos12}.
+PGV is more expressive than GV: \cref{ex:cycle} is typeable and guaranteed deadlock-free in PGV, but is not typeable in GV~\cite{wadler14} and not guaranteed deadlock-free in GV's predecessor~\cite{gayvasconcelos10}.
 We believe PGV is a non-conservative \emph{extension} of GV, as CP can be embedded in a Kobayashi-style system~\cite{dardhaperez15extended}.

### 2021-11-17 coordination21/prf/thm-pcp-closed-progress.tex
restructuring before reworking

--- 2020-10-01 prf/thm-pcp-closed-progress.tex
@@
   \label{prf:thm-pcp-closed-progress}
-  If $\tm{P}=\tm{\halt}$, we are done. Otherwise, $\tm{P}$ is of the form
+  By~\cref{lem:pcp-canonical-forms}, we rewrite $\tm{P}$ to canonical form. If the resulting process is $\tm{\halt}$, we are done. Otherwise, it is of the form
   \[
@@
 
-  Our proof follows the same reasoning by Kobayashi \cite{kobayashi2006} used in the proof of deadlock freedom for closed processes (Theorem 2).
+  Our proof follows the same reasoning by Kobayashi~\cite{kobayashi06} used in the proof of deadlock freedom for closed processes (Theorem 2).
 

### 2021-11-17 coordination21/prf/thm-pcp-to-pgv-operational-correspondence-soundness.tex
restructuring before reworking

--- 2020-09-17 prf/thm-pcp-to-pgv-operational-correspondence-soundness.tex
@@
   By induction on the derivation of $\pgv{\tm{\cpgvC{P}}\cred\tm{\conf{C}}}$.
-  We omit the cases which cannot occur, \ie \LabTirName{E-New}, \LabTirName{E-Spawn}, and \LabTirName{LiftM}.
+  We omit the cases which cannot occur as their left-hand side term forms are not in the image of the translation function, \ie \LabTirName{E-New}, \LabTirName{E-Spawn}, and \LabTirName{E-LiftM}.
 

### 2021-11-17 lmcs/fig/pgv-scheduler.tex
restructuring before reworking

--- 2021-01-28 appendix.tex
@@
-% !TeX root = priorities.tex
-\section{Priority GV}
-\begingroup
-\usingnamespace{pgv}
-\restatelemma{lempgvvaluedone}
-\input{prf/lem-pgv-value-done}
-
-\restatelemma{lempgvsubstitution}
-\input{prf/lem-pgv-substitution}
-
-\restatelemma{lempgvsubjectreductionterms}
-\input{prf/lem-pgv-subject-reduction-terms}
-
-\restatelemma{lempgvsubjectcongruence}
-\input{prf/lem-pgv-subject-congruence}
-
-\restatetheorem{thmpgvsubjectreductionconfs}
-\input{prf/thm-pgv-subject-reduction-confs}
-
-\begin{restatablelemma}{lempgvreadypriority}
-  \label{lem:pgv-ready-priority}
-  If $\tseq[\cs{p}]{\ty{\Gamma}}{L}{T}$ is ready to act on $\tmty{x}{S}\in\ty{\Gamma}$, then the priority bound $\cs{p}$ is some priority $\cs{o}$, \ie not $\cs{\pbot}$ or $\cs{\ptop}$.
-\end{restatablelemma}
-\begin{proof}
-  Let $\tm{L}=\tm{\plug{E}{M}}$.
-By induction on the structure of $\tm{E}$.
-$\tm{M}$ has priority $\pr({\ty{S}})$, and each constructor of the evaluation context $\tm{E}$ passes on the \emph{maximum} of the priorities of its premises.
-No rule introduces the priority bound $\cs{\ptop}$ on the sequent.
-\end{proof}
-
-\restatelemma{lempgvcanonicalforms}
-\begin{proof}
-  We move any $\nu$-binders to the top using \LabTirName{SC-ResExt}, discard any superfluous occurrences of $\tm{\child\;\unit}$ using \LabTirName{SC-ParNil}, and move the main thread to the rightmost position using \LabTirName{SC-ParComm} and \LabTirName{SC-ParAssoc}.
-\end{proof}
-
-\restatetheorem{thmpgvclosedprogressconfs}
-\input{prf/thm-pgv-closed-progress-confs}
-
-\input{fig/pgv-typing-sugar}
-\endgroup
-
-\section{Relation to Priority CP}
-\begingroup
-\usingnamespace{pcp}
-\subsection{Revisiting Priority CP}
-
-\subsubsection*{Types}
-Types ($\pcp{\ty{A}}, \pcp{\ty{B}}$) in PCP are based on classical linear logic propositions, and are defined by the following grammar:
-\[
-  \usingnamespace{pcp}
-  \begin{array}{lcl}
-    \ty{A}, \ty{B}
-    & \Coloneqq & \ty{\tytens[\cs{o}]{A}{B}}
-      \sep        \ty{\typarr[\cs{o}]{A}{B}}
-      \sep        \ty{\tyone[\cs{o}]}
-      \sep        \ty{\tybot[\cs{o}]}
-      \sep        \ty{\typlus[\cs{o}]{A}{B}}
-      \sep        \ty{\tywith[\cs{o}]{A}{B}}
-      \sep        \ty{\tynil[\cs{o}]}
-      \sep        \ty{\tytop[\cs{o}]}
-  \end{array}
-\]
-Each connective is annotated with a priority $\cs{o}\in\mathbb{N}$.
-Types $\pcp{\ty{\tytens[\cs{o}]{A}{B}}}$ and $\pcp{\ty{\typarr[\cs{o}]{A}{B}}}$ type the endpoints of a channel over which we send or receive a channel of type $\pcp{\ty{A}}$, and then proceed as type $\pcp{\ty{B}}$.
-Types $\pcp{\ty{\tyone[\cs{o}]}}$ and $\pcp{\ty{{\tybot}[\cs{o}]}}$ type the endpoints of a channel whose session has terminated, and over which we send or receive a \emph{ping} before closing the channel.
-These two types act as units for $\pcp{\ty{\tytens[\cs{o}]{A}{B}}}$ and $\pcp{\ty{\typarr[\cs{o}]{A}{B}}}$, respectively.
-Types $\pcp{\ty{\typlus[\cs{o}]{A}{B}}}$ and $\pcp{\ty{\tywith[\cs{o}]{A}{B}}}$ type the endpoints of a channel over which we can receive or send a choice between two branches $\pcp{\ty{A}}$ or $\pcp{\ty{B}}$.
-We have opted for a simplified version of choice and followed the original Wadler's CP \cite{wadler14}, however types $\ty{\oplus}$ and $\ty{\with}$ can be trivially generalised to $\pcp{\ty{\oplus^{\cs{o}}\{l_i:A_i\}_{i\in I}}}$ and $\pcp{\ty{\with^{\cs{o}}\{l_i:A_i\}_{i\in I}}}$, respectively, as in the original PCP \cite{dardhagay18}.
-Types $\pcp{\ty{\tynil[\cs{o}]}}$ and $\pcp{\ty{\tytop[\cs{o}]}}$ type the endpoints of a channel over which we can send or receive a choice between \emph{no options}.
-These two types act as units for $\pcp{\ty{\typlus[\cs{o}]{A}{B}}}$ and $\pcp{\ty{\tywith[\cs{o}]{A}{B}}}$, respectively.
-
-\subsubsection*{Environments}
-\label{sec:pcp-environments}
-Typing environments $\pcp{\ty{\Gamma}}$, $\pcp{\ty{\Delta}}$ associate names to types.
-Environments are linear, so two environments can only be combined as $\pcp{\ty{\Gamma}}, \pcp{\ty{\Delta}}$ if their names are distinct, \ie $\pcp{\fv(\ty{\Gamma})\cap\fv(\ty{\Delta})=\varnothing}$.
-\[
-  \usingnamespace{pcp}
-  \begin{array}{lcl}
-    \ty{\Gamma}, \ty{\Delta}
-    & \Coloneqq & \ty{\emptyenv}
-      \sep        \ty{\Gamma}, \tmty{x}{A}
-  \end{array}
-\]
-
-\subsubsection*{Duality}
-\label{sec:pcp-duality}
-Duality is an involutive function on types which preserves priorities:
-\[
-  \usingnamespace{pcp}
-  \setlength{\arraycolsep}{1pt}
-  \begin{array}{lcl}
-    \ty{\co{(\tyone[\cs{o}])}} & = & \ty{\tybot[\cs{o}]} \\
-    \ty{\co{(\tybot[\cs{o}])}} & = & \ty{\tyone[\cs{o}]}
-  \end{array}
-  \quad
-  \begin{array}{lcl}
-    \ty{\co{(\tytens[\cs{o}]{A}{B})}} & = & \ty{\typarr[\cs{o}]{\co{A}}{\co{B}}} \\
-    \ty{\co{(\typarr[\cs{o}]{A}{B})}} & = & \ty{\tytens[\cs{o}]{\co{A}}{\co{B}}}
-  \end{array}
-  \quad
-  \begin{array}{lcl}
-    \ty{\co{(\tynil[\cs{o}])}} & = & \ty{\tytop[\cs{o}]} \\
-    \ty{\co{(\tytop[\cs{o}])}} & = & \ty{\tynil[\cs{o}]}
-  \end{array}
-  \quad
-  \begin{array}{lcl}
-    \ty{\co{(\typlus[\cs{o}]{A}{B})}} & = & \ty{\tywith[\cs{o}]{\co{A}}{\co{B}}} \\
-    \ty{\co{(\tywith[\cs{o}]{A}{B})}} & = & \ty{\typlus[\cs{o}]{\co{A}}{\co{B}}}
-  \end{array}
-\]
-
-\subsubsection*{Priorities}
-\label{sec:pcp-priorities}
-The function $\pr(\cdot)$ returns smallest priority of a type.
-As with PGV, the type system guarantees that the top-most connective always holds the smallest priority.
-The function $\minpr(\cdot)$ returns the \emph{minimum} priority of all types a typing context, or $\cs{\ptop}$ if the context is empty:
-\[
-  \usingnamespace{pcp}
-  \setlength{\arraycolsep}{1pt}
-  \begin{array}{lclclcl}
-    \pr(\ty{\tyone[\cs{o}]})        & = & \cs{o} \\
-    \pr(\ty{\tybot[\cs{o}]})        & = & \cs{o}
-  \end{array}
-  \qquad
-  \begin{array}{lclclcl}
-    \pr(\ty{\tytens[\cs{o}]{A}{B}}) & = & \cs{o} \\
-    \pr(\ty{\typarr[\cs{o}]{A}{B}}) & = & \cs{o}
-  \end{array}
-  \qquad
-  \begin{array}{lclclcl}
-    \pr(\ty{\tynil[\cs{o}]})        & = & \cs{o} \\
-    \pr(\ty{\tytop[\cs{o}]})        & = & \cs{o}
-  \end{array}
-  \qquad
-  \begin{array}{lclclcl}
-    \pr(\ty{\typlus[\cs{o}]{A}{B}}) & = & \cs{o} \\
-    \pr(\ty{\tywith[\cs{o}]{A}{B}}) & = & \cs{o}
-  \end{array}
-\]
-\[
-  \minpr(\ty{\emptyenv})          = \cs{\ptop}
-  \quad
-  \minpr(\ty{\Gamma},\tmty{x}{A}) = \minpr(\ty{\Gamma})\sqcap\minpr(\ty{A})
-\]
-
-\subsubsection*{Terms}
-Processes ($\pcp{\tm{P}}$, $\pcp{\tm{Q}}$) in PCP are defined by the following grammar.
-\[
-  \usingnamespace{pcp}
-  \begin{array}[t]{lcl}
-    \tm{P}, \tm{Q}
-    & \Coloneqq & \tm{\link{x}{y}}
-           \sep   \tm{\res{x}{y}{P}}
-           \sep   \tm{(\ppar{P}{Q})}
-           \sep   \tm{\halt}
-    \\   & \sep & \tm{\send{x}{y}{P}}
-           \sep   \tm{\close{x}{P}}
-           \sep   \tm{\recv{x}{y}{P}}
-           \sep   \tm{\wait{x}{P}}
-    \\   & \sep & \tm{\inl{x}{P}}
-           \sep   \tm{\inr{x}{P}}
-           \sep   \tm{\offer{x}{P}{Q}}
-           \sep   \tm{\absurd{x}}
-  \end{array}
-\]
-Process $\pcp{\tm{\link{x}{y}}}$ links endpoints $\pcp{\tm{x}}$ and $\pcp{\tm{y}}$ and forwards communication from one to the other.
-$\pcp{\tm{\res{x}{y}{P}}}$, $\pcp{\tm{(\ppar{P}{Q})}}$ and $\pcp{\tm{\halt}}$ denote respectively the restriction processes where channel endpoints $\pcp{\tm{x}}$ and $\pcp{\tm{y}}$ are bound together and with scope $\pcp{\tm{P}}$, the parallel composition of processes $\pcp{\tm{P}}$ and $\pcp{\tm{Q}}$ and the terminated process.
-Processes $\pcp{\tm{\send{x}{y}{P}}}$ and $\pcp{\tm{\recv{x}{y}{P}}}$ send or receive over channel $\pcp{\tm{x}}$ a value $\pcp{\tm{y}}$ and proceed as process $\pcp{\tm{P}}$.
-Processes $\pcp{\tm{\close{x}{P}}}$ and $\pcp{\tm{\wait{x}{P}}}$ send and receive an empty value---denoting the closure of channel $\pcp{\tm{x}}$, and continue as $\pcp{\tm{P}}$.
-Processes $\pcp{\tm{\inl{x}{P}}}$ and $\pcp{\tm{\inr{x}{P}}}$ make a left and right choice, respectively and proceed as process $\pcp{\tm{P}}$.
-Dually, $\pcp{\tm{\offer{x}{P}{Q}}}$ offers both left and right branches, with continuations $\pcp{\tm{P}}$ and $\pcp{\tm{Q}}$, and $\pcp{\tm{\absurd{x}}}$ is the empty offer.
-We write \emph{unbound} send as $\pcp{\tm{\usend{x}{y}{P}}}$, which is syntactic sugar for $\pcp{\tm{\send{x}{z}{(\ppar{\link{y}{z}}{P})}}}$.
-Alternatively, we could take $\pcp{\tm{\usend{x}{y}{P}}}$ as primitive, and let $\pcp{\tm{\send{x}{y}{P}}}$ be syntactic sugar for $\pcp{\tm{\res{y}{z}{(\usend{x}{z}{P})}}}$.
-CP takes \emph{bound} sending as primitive, as it is impossible to eliminate the top-level cut in terms such as $\pcp{\tm{\res{y}{z}{(\usend{x}{z}{P})}}}$, even with commuting conversions.
-In our setting without commuting conversions and with more permissive normal forms, this is no longer an issue, but, for simplicity, we keep bound sending as primitive.
-
-\subsubsection*{Operational Semantics}
-The operational semantics for PCP, given in \cref{fig:pcp-operational-semantics}, is defined as a reduction relation $\pcp{\red}$ on processes (bottom) and uses structural congruence (top).
-Each of the axioms of structural congruence corresponds to the axiom of the same name for PGV.
-We write $\pcp{\red^+}$ for the transitive closures, and $\pcp{\red^\star}$ for the reflexive-transitive closures.
-
-The reduction relation is given by a set of axioms and inference rules for context closure.
-Reduction occurs under restriction.
-$\LabTirName{E-Link}$ reduces a parallel composition with a link into a substitution.
-$\LabTirName{E-Send}$ is the main communication rule, where send and receive processes sychronise and reduce to the corresponding continuations.
-$\LabTirName{E-Close}$ follows the previous rule and it closes the channel identified by endpoints $\pcp{\tm{x}}$ and $\pcp{\tm{y}}$.
-$\LabTirName{E-Select-Inl}$ and $\LabTirName{E-Select-Inr}$ are generalised versions of $\LabTirName{E-Send}$.
-They state respectively that a left and right selection synchronises with a choice offering and reduces to the corresponding continuations.
-The last three rules state that reduction is closed under restriction, parallel composition and structural congruence, respectively.
-
-\input{fig/pcp-operational-semantics}
-
-\subsubsection*{Typing}
-\Cref{fig:pcp-typing} gives the typing rules for our version of PCP.
-A typing judgement $\pcp{\seq{P}{\Gamma}}$ states that ``process $\pcp{\tm{P}}$ is well typed under the typing context $\pcp{\ty{\Gamma}}$''.
-
-\input{fig/pcp-typing}
-
-\textsc{T-Link} states that the link process $\pcp{\tm{\link{x}{y}}}$ is well typed under channels $\pcp{\tm{x}}$ and $\pcp{\tm{y}}$ having dual types, respectively $\pcp{\ty{A}}$ and $\pcp{\ty{\co{A}}}$.
-\textsc{T-Res} sates that the restriction process $\pcp{\tm{\res{x}{y}{P}}}$ is well typed under typing context $\pcp{\ty{\Gamma}}$ if process $\pcp{\tm{P}}$ is well typed in $\pcp{\ty{\Gamma}}$ augmented with channel endpoints $\pcp{\tm{x}}$ and $\pcp{\tm{y}}$ having dual types, respectively $\pcp{\ty{A}}$ and $\pcp{\ty{\co{A}}}$.
-\textsc{T-Par} states that the parallel composition of processes $\pcp{\tm{P}}$ and $\pcp{\tm{Q}}$ is well typed under the disjoint union of their respective typing contexts.
-\textsc{T-Halt} states that the terminated process $\pcp{\tm{\halt}}$ is well typed in the empty context.
-\textsc{T-Send} and \textsc{T-Recv} state that the sending and receiving of a bound name $\pcp{\tm{y}}$ over a channel $\pcp{\tm{x}}$ is well typed under $\pcp{\ty{\Gamma}}$ and $\pcp{\tm{x}}$ of type $\pcp{\ty{\tytens[\cs{o}]{A}{B}}}$, respectively $\pcp{\typarr[\cs{o}]{A}{B}}$.
-Priority $\cs{o}$ is the smallest among all priorities of the types used by the output or input process, captured by the side condition $\pcp{\cs{o}<\minpr(\ty{\Gamma},\ty{A},\ty{B})}$.
-Rules \textsc{T-Close} and \textsc{T-Wait} type the closure of channel $\pcp{\tm{x}}$ and are in the same lines as the previous two rules, requiring that the priority of channel $\pcp{\tm{x}}$ is the smallest among all priorities in $\pcp{\ty{\Gamma}}$.
-\textsc{T-Select-Inl} and \textsc{T-Select-Inr} type respectively the left $\pcp{\tm{\inl{x}{P}}}$ and right $\pcp{\tm{\inr{x}{P}}}$ choice performed on channel $\pcp{\tm{x}}$.
-\textsc{T-Offer} and \textsc{T-Offer-Absurd} type the offering of a choice, or empty choice, on channel $\pcp{\tm{x}}$.
-In all the above rules the priority $\cs{o}$ of channel $\pcp{\tm{x}}$ is the smallest with respect to the typing context $\pcp{\cs{o}<\minpr(\ty{\Gamma})}$ and types involved in the choice $\pcp{\cs{o}<\minpr(\ty{\Gamma},\ty{A},\ty{B})}$.
-
-Finally, since our reduction relation is a strict subset of the reduction relation in the original~\cite{dardhagay18}, we defer to their proofs.
-We prove progress for our version of PCP, see~\cref{prf:thm-pcp-closed-progress}.
-
-% Proofs that the syntactic sugar is well-typed.
-\input{fig/pcp-typing-sugar}
-
-\begin{compacttheorems}
-    \begin{restatabledefinition}{defpcpactions}[Actions]
-    A~process acts on an endpoint $\tm{x}$ if it is $\tm{\link{x}{y}}$, $\tm{\link{y}{x}}$, $\tm{\send{x}{y}{P}}$, $\tm{\recv{x}{y}{P}}$, $\tm{\close{x}{P}}$, $\tm{\wait{x}{P}}$, $\tm{\inl{x}{P}}$, $\tm{\inr{x}{P}}$, $\tm{\offer{x}{P}{Q}}$, or $\tm{\absurd{x}}$.
-A~process is an action if it acts on some endpoint $\tm{x}$.
-  \end{restatabledefinition}
-  \begin{restatabledefinition}{defpcpcanonicalforms}[Canonical Forms]
-    \label{def:pcp-canonical-forms}
-    A~process $\tm{P}$ is in canonical form if it is $\tm{\halt}$ or $\tm{\res{x_1}{x'_1}{\dots\res{x_n}{x'_n}{(P_1 \parallel\dots\parallel P_m)}}}$ where $m>0$ and each $\tm{P_j}$ is an action.
-  \end{restatabledefinition}
-  \begin{restatablelemma}{lempcpcanonicalforms}[Canonical Forms]
-    \label{lem:pcp-canonical-forms}
-    If $\seq{P}{\ty{\Gamma}}$, there exists some $\tm{Q}$ such that $\tm{P}\equiv\tm{Q}$ and $\tm{Q}$ is in canonical form.
-  \end{restatablelemma}
-  \begin{proof}
-    If $\tm{P}=\tm{\halt}$, we are done.
-Otherwise, we move any $\nu$-binders to the top using \LabTirName{SC-ResExt}, and discard any superfluous occurrences of $\tm{\halt}$ using \LabTirName{SC-ParNil}.
-  \end{proof}
-  \restatetheorem{thmpcpclosedprogress}
-  \input{prf/thm-pcp-closed-progress}
-\end{compacttheorems}
-\endgroup
-
-\subsection{Correspondence between PGV and PCP}
-\begingroup
-\restatelemma{lempcptopgvtermspreservation}
-\input{prf/lem-pcp-to-pgv-terms-preservation}
-
-\restatetheorem{thmpcptopgvconfspreservation}
-\input{prf/thm-pcp-to-pgv-confs-preservation}
-
-\restatetheorem{thmpcptopgvoperationalcorrespondencesoundness}
-\input{prf/thm-pcp-to-pgv-operational-correspondence-soundness}
-
-\restatelemma{lempcptopgvcpgvMtocpgvC}
-\input{prf/lem-pcp-to-pgv-cpgvM-to-cpgvC}
-
-\begin{restatablelemma}{lempcptopgvoperationalcorrespondenceequiv}%
-  \label{lem:pcp-to-pgv-confs-operational-correspondence-equiv}
-  If $\pcp{\seq{P}{\ty{\Gamma}}}$ and $\pcp{\tm{P}\equiv\tm{Q}}$,
-  then $\pgv{\tm{\cpgvC{P}}\equiv\tm{\cpgvC{Q}}}$.
-\end{restatablelemma}
-\begin{proof}
-  Every axiom of the structural congruence in PCP maps directly to the axiom of the same name in PGV.
-\end{proof}
-
-\restatetheorem{thmpcptopgvoperationalcorrespondencecompleteness}
-\input{prf/thm-pcp-to-pgv-operational-correspondence-completeness}
-\endgroup
-
-\section{Milner's Cyclic Scheduler}
-\label{sec:milner}
-As an example of a deadlock-free cyclic process, Dardha and Gay~\cite{dardhagay18} introduce an implementation of Milner's cyclic scheduler~\cite{milner89} in Priority CP.
-We reproduce that scheduler here, and show its translation to Priority GV.
-
-\begingroup
-\usingnamespace{pcp}
-\begin{example}[Milner's Cyclic Scheduler, PCP]
-  \label{ex:pcp-scheduler}
-  A set of processes $\tm{\proc{i}}$, $1\leq{i}\leq{n}$, is scheduled to perform some tasks in cyclic order, starting with $\tm{\proc1}$, ending with $\tm{\proc{n}}$, and notifying $\tm{\proc1}$ when all processes have finished.
-
-  Our scheduler $\tm{\sched}$ consists of set of agents $\tm{\agent{i}}$, $1\leq{i}\leq{n}$, each representing their respective process.
-Each process $\tm{\proc{i}}$ waits for the signal to start their task on $\tm{a'_i}$, and signals completion on $\tm{b'_i}$.
-Each agent signals their process to start on $\tm{a_i}$, waits for their process to finish on $\tm{b_i}$, and then signals for the next agent to continue on $\tm{c_i}$.
-The agent $\tm{\agent1}$ initiates, then waits for every other process to finish, and signals $\tm{\proc1}$ on $\tm{d}$.
-Every other agent $\tm{\agent{i}}$, $2\leq{i}\leq{n}$ waits on $\tm{c'_{i-1}}$ for the signal to start.
-Each of the channels in the scheduler is of a terminated type, and is merely used to synchronise.
-
-  Below is a diagram of our scheduler instantiated with three processes:
-  \begin{center}
+\begin{figure}
+  \centering
   \begin{tikzpicture}
@@
   \end{tikzpicture}
-  \end{center}
-
-    We implement the scheduler as follows, using $\tm{\prod_{I}P_i}$ to denote the parallel composition of the processes $\tm{P_i}$, $\tm{i}\in\tm{I}$, and $\tm{\plug{P}{Q}}$ to denote the plugging of $\tm{Q}$ in the one-hole process-context $\tm{P}$.
-The process-contexts $\tm{P_i}$ represent the computations performed by each process $\tm{\proc{i}}$.
-The process-contexts $\tm{Q_i}$ represent any post-processing, and any possible data transfer from $\tm{\proc{i}}$ to $\tm{\proc{i+1}}$.
-Finally, $\tm{Q_1}$ should contain $\tm{\labwait{d'}}$.
-\[
-  \begin{array}{lrlr}
-    \tm{\sched}
-    & \defeq & \tm{\res{a_1}{a'_1}{\dots{\res{a_n}{a'_n}{}}}}
-               \tm{\res{b_1}{b'_1}{\dots{\res{b_n}{b'_n}{}}}}
-               \tm{\res{c_1}{c'_1}{\dots{\res{c_n}{c'_n}{}}}}
-               \tm{\res{d}{d'}{}}
-    \\ &     & \tm{(
-               \ppar{\proc1}{\agent1}
-               \parallel
-               \prod_{2\leq{i}\leq{n}}(\ppar{\proc{i}}{\wait{c'_{i-1}}{\agent{i}}})
-               )}
-    \\
-    \tm{\agent1}
-    & \defeq & \tm{\close{a_i}{\wait{b_i}{\close{c_i}{\wait{c'_n}{\close{d}{\halt}}}}}}
-    \\
-    \tm{\agent{i}}
-    & \defeq & \tm{\close{a_i}{\wait{b_i}{\close{c_i}{\halt}}}}
-    \\
-    \tm{\proc{i}}
-    & \defeq & \tm{\wait{a'_i}{\plug{P_i}{\close{b'_i}{Q_i}}}}
-  \end{array}
-\]
-\end{example}
-\endgroup
-
-\begingroup
-\usingnamespace{pgv}
-\begin{example}[Milner's Cyclic Scheduler, PGV]
-  \label{ex:pgv-scheduler}
-  The PGV scheduler has exactly the same behaviour as the PCP version in~\cref{ex:pcp-scheduler}.
-It is implemented as follows, using $\tm{\prod_{I}\conf{C}_i}$ to denote the parallel composition of the processes $\tm{\conf{C}_i}$, $\tm{i}\in\tm{I}$, and $\tm{\plug{M}{N}}$ to denote the plugging of $\tm{N}$ in the one-hole term-context $\tm{M}$.
-For simplicity, we let $\tm{\sched}$ be a configuration.
-The terms $\tm{M_i}$ represent the computations performed by each process $\tm{\proc{i}}$.
-The terms $\tm{N_i}$ represent any post-processing, and any possible data transfer from $\tm{\proc{i}}$ to $\tm{\proc{i+1}}$.
-Finally, $\tm{N_1}$ should contain $\tm{\wait\;{d'}}$.
-  \[
-    \begin{array}{lrlr}
-      \tm{\sched}
-      & \defeq & \tm{\res{a_1}{a'_1}{\dots{\res{a_n}{a'_n}{}}}}
-                 \tm{\res{b_1}{b'_1}{\dots{\res{b_n}{b'_n}{}}}}
-                 \tm{\res{c_1}{c'_1}{\dots{\res{c_n}{c'_n}{}}}}
-                 \tm{\res{d}{d'}{}}
-      \\ &     & \tm{\begin{array}{lll}
-                       (
-                       & \phi\;\proc1
-                         \parallel
-                         \child\;\andthen{\agent1}{\andthen{\wait\;{c'_n}}{\close\;{d}}}
-                       \\
-                       \parallel
-                       &
-                         \prod_{2\leq{i}\leq{n}}
-                         (\child\;\proc{i} \parallel \child\;\andthen{\wait\;{c'_{i-1}}}{\agent{i}})
-                       & )
-                     \end{array}}
-      \\
-      \tm{\agent{i}}
-      & \defeq & \tm{\andthen{\close\;{a_i}}{\andthen{\wait\;{b_i}}{\close\;{c_i}}}}
-      \\
-      \tm{\proc{i}}
-      & \defeq & \tm{\andthen{\wait\;{a'_i}}{\plug{M_i}{\andthen{\close\;{b'_i}}{N_i}}}}
-    \end{array}
-  \]
-\end{example}
-\endgroup
-
-If $\pcp{\tm{\cpgvM{P_i}}}=\pgv{\tm{M_i}}$ and $\pcp{\tm{\cpgvM{Q_i}}}=\pgv{\tm{N_i}}$, then the translation of $\pcp{\tm{\sched}}$ (\cref{ex:pcp-scheduler}), $\pcp{\tm{\cpgvC{\sched\,}}}$, is exactly $\pgv{\tm{\sched}}$ (\cref{ex:pgv-scheduler}).
+  \label{fig:pgv-scheduler}
+  \caption{Milner's Cyclic Scheduler}
+\end{figure}
 
 %%% Local Variables:
-%%% TeX-master: "priorities"
+%%% TeX-master: "../priorities"
 %%% End:

### 2021-11-17 lmcs/prf/lem-pgv-value-done.tex
restructuring before reworking

--- 2020-09-28 prf/lem-pgv-value-done.tex
@@
   \begin{case*}[\LabTirName{T-Inl}]
-    The induction hypothesis gives us $\cs{p}=\cs{\pbot}$, and $\pr(\ty{\Gamma})=\pr{\ty{T}}$. We know $\pr(\ty{T})=\pr({\ty{U}})$, hence $\pr(\ty{\Gamma})=\pr(\ty{\tysum{T}{U}})$.
+    The induction hypothesis gives us $\cs{p}=\cs{\pbot}$, and $\pr(\ty{\Gamma})=\pr{(\ty{T})}$. We know $\pr(\ty{T})=\pr({\ty{U}})$, hence $\pr(\ty{\Gamma})=\pr(\ty{\tysum{T}{U}})$.
     \begin{mathpar}
@@
   \begin{case*}[\LabTirName{T-Inr}]
-    The induction hypothesis gives us $\cs{p}=\cs{\pbot}$, and $\pr(\ty{\Gamma})=\pr{\ty{U}}$. We know $\pr(\ty{T})=\pr({\ty{U}})$, hence $\pr(\ty{\Gamma})=\pr(\ty{\tysum{T}{U}})$.
+    The induction hypothesis gives us $\cs{p}=\cs{\pbot}$, and $\pr(\ty{\Gamma})=\pr{(\ty{U})}$. We know $\pr(\ty{T})=\pr({\ty{U}})$, hence $\pr(\ty{\Gamma})=\pr(\ty{\tysum{T}{U}})$.
     \begin{mathpar}

### 2021-12-03 lmcs/priority-gv.tex
Document conversion to LMCS
working on thms

--- 2021-04-26 priority-gv.tex
@@
-\documentclass[main.tex]{subfiles}
-
-\begin{document}
-
 \section{Priority GV}\label{sec:pgv}
@@
 Priority GV offers a more fine-grained analysis of communication structures, and by separating channel creation form thread spawning it allows cyclic structures.
+
 We illustrate this with two programs in PGV, \cref{ex:cycle,ex:deadlock}.
@@
 The main process does one of two things:
-\begin{enumerate*}[label= (\alph*) ]
+\begin{enumerate}[label= (\alph*) ]
 \item in \cref{ex:cycle}, it sends a unit over the channel $\tm{x}/\tm{x'}$, and then waits to receive a unit over the channel $\tm{y}/\tm{y'}$;
 \item in \cref{ex:deadlock}, it does these in the opposite order, which results in a deadlock.
-\end{enumerate*}
+\end{enumerate}
 PGV is more expressive than GV: \cref{ex:cycle} is typeable and guaranteed deadlock-free in PGV, but is not typeable in GV~\cite{wadler14} and not guaranteed deadlock-free in GV's predecessor~\cite{gayvasconcelos10}.
@@
 \]
+
 We extend the function $\pr(\cdot)$ to types and typing contexts by returning the smallest priority in the type or context, or $\cs{\top}$ if there is no priority.
@@
 \]
+
 Let $\tm{x}$, $\tm{y}$, $\tm{z}$, and $\tm{w}$ range over variable names.
@@
 Paraphrasing Dardha and Gay~\cite{dardhagay18extended}, priorities obey the following two laws:
-\begin{enumerate*}[label= (\roman*) ]
+\begin{enumerate}[label= (\roman*) ]
 \item an action with lower priority happens before an action with higher priority; and
 \item communication requires \emph{equal} priorities for dual actions.
-\end{enumerate*}
+\end{enumerate}
 
@@
 
-\subsubsection*{Subject reduction}
+\section{Technical Developments}
+\subsection{Subject reduction}
 Unlike with previous versions of GV, structural congruence, term reduction, and configuration reduction are all type preserving.
@@
 For this, we prove~\cref{lem:pgv-value-done}, which shows that values have finished all their communication, and that any priorities in the type of the value come from the typing environment.
-% The proofs are by structural induction and can be found in \cref{prf:lem-pgv-value-done,prf:lem-pgv-substitution,prf:lem-pgv-subject-reduction-terms,prf:lem-pgv-subject-congruence,prf:thm-pgv-subject-reduction-confs}.
-\begin{compacttheorems}
-  \begin{restatablelemma}{lempgvvaluedone}%
+%The proofs are by structural induction and can be found in \cref{prf:lem-pgv-value-done,prf:lem-pgv-substitution,prf:lem-pgv-subject-reduction-terms,prf:lem-pgv-subject-congruence,prf:thm-pgv-subject-reduction-confs}.
+  \begin{lemma}%
     \label{lem:pgv-value-done}
     If $\tseq[\cs{p}]{\ty{\Gamma}}{V}{T}$, then $\cs{p}=\cs{\pbot}$, and $\minpr(\ty{\Gamma})=\minpr(\ty{T})$.
-  \end{restatablelemma}
-  \begin{restatablelemma}{lempgvsubstitution}[Substitution]%
+  \end{lemma}
+  \input{prf/lem-pgv-value-done}
+
+  \begin{lemma}[Substitution]%
     \label{lem:pgv-substitution}
@@
     If $\tseq[\cs{p}]{\ty{\Gamma},\tmty{x}{U'}}{M}{T}$ and $\tseq[\cs{q}]{\ty{\Theta}}{V}{U'}$, then $\tseq[\cs{p}]{\ty{\Gamma},\ty{\Theta}}{\subst{M}{V}{x}}{T}$.
-  \end{restatablelemma}
-  \begin{restatablelemma}{lempgvsubjectreductionterms}[Subject Reduction, $\tred$]%
+  \end{lemma}
+  \input{prf/lem-pgv-substitution}
+
+  \begin{lemma}[Subject Reduction, $\tred$]%
     \label{lem:pgv-subject-reduction-terms}
@@
     then $\tseq[\cs{p}]{\ty{\Gamma}}{M'}{T}$.
-  \end{restatablelemma}
-  \begin{restatablelemma}{lempgvsubjectcongruence}[Subject Congruence, $\equiv$]%
+  \end{lemma}
+  \input{prf/lem-pgv-subject-reduction-terms}
+
+  \begin{lemma}[Subject Congruence, $\equiv$]%
     \label{lem:pgv-subject-congruence}
@@
     then $\cseq[\phi]{\ty{\Gamma}}{\conf{C'}}$.
-  \end{restatablelemma}
-  \begin{restatabletheorem}{thmpgvsubjectreductionconfs}[Subject Reduction, $\cred$]%
+  \end{lemma}
+  \input{prf/lem-pgv-subject-congruence}
+
+  \begin{theorem}[Subject Reduction, $\cred$]%
     \label{thm:pgv-subject-reduction-confs}
@@
     then $\cseq[\phi]{\ty{\Gamma}}{\conf{C'}}$.
-  \end{restatabletheorem}
-\end{compacttheorems}
+  \end{theorem}
+  \input{prf/thm-pgv-subject-reduction-confs}
 
-\subsubsection*{Progress and deadlock freedom}
+\subsection{Progress and deadlock freedom}
 PGV satisfies progress, as PGV configurations either reduce or are in normal form.
@@
 
-\begin{compacttheorems}
   \begin{definition}[Canonical Forms]%
@@
     A~configuration $\tm{\conf{C}}$ is in canonical form if it is of the form
-    \(
-    \tm{\res{x_1}{x'_1}{\dots\res{x_n}{x'_n}{(\child\;M_1\parallel\dots\parallel\child\;M_m\parallel\main\;N)}}}
-    \)
+
+    $\tm{\res{x_1}{x'_1}{\dots\res{x_n}{x'_n}{(\child\;M_1\parallel\dots\parallel\child\;M_m\parallel\main\;N)}}}$
+
     where no term $\tm{M_i}$ is a value.
   \end{definition}
-  \begin{restatablelemma}{lempgvcanonicalforms}[Canonical Forms]%
+
+  \begin{lemma}[Canonical Forms]%
     \label{lem:pgv-canonical-forms}
     If $\cseq[\main]{\ty{\Gamma}}{\conf{C}}$, there exists some $\tm{\conf{D}}$ such that $\tm{\conf{C}}\equiv\tm{\conf{D}}$ and $\tm{\conf{D}}$ is in canonical form.
-  \end{restatablelemma}
+  \end{lemma}
+  \begin{proof}
+    We move any $\nu$-binders to the top using \LabTirName{SC-ResExt}, discard any superfluous occurrences of $\tm{\child\;\unit}$ using \LabTirName{SC-ParNil}, and move the main thread to the rightmost position using \LabTirName{SC-ParComm} and \LabTirName{SC-ParAssoc}.
+  \end{proof}
+
   \begin{definition}[Normal Forms]%
     A~configuration $\tm{\conf{C}}$ is in normal form if it is of the form
-    \(
-    \tm{\res{x_1}{x'_1}{\dots\res{x_n}{x'_n}{(\child\;M_1\parallel\dots\parallel\child\;M_m\parallel\main\;V)}}}
-    \)
+
+    $\tm{\res{x_1}{x'_1}{\dots\res{x_n}{x'_n}{(\child\;M_1\parallel\dots\parallel\child\;M_m\parallel\main\;V)}}}$
+
     where each $\tm{M_i}$ is ready to act on $\tm{x_i}$.
   \end{definition}
-  \begin{restatabletheorem}{thmpgvclosedprogressconfs}[Progress, $\cred$]%
+
+  \begin{lemma}
+    \label{lem:pgv-ready-priority}
+    If $\tseq[\cs{p}]{\ty{\Gamma}}{L}{T}$ is ready to act on $\tmty{x}{S}\in\ty{\Gamma}$, then the priority bound $\cs{p}$ is some priority $\cs{o}$, \ie not $\cs{\pbot}$ or $\cs{\ptop}$.
+  \end{lemma}
+  \begin{proof}
+    Let $\tm{L}=\tm{\plug{E}{M}}$.
+By induction on the structure of $\tm{E}$.
+$\tm{M}$ has priority $\pr({\ty{S}})$, and each constructor of the evaluation context $\tm{E}$ passes on the \emph{maximum} of the priorities of its premises.
+No rule introduces the priority bound $\cs{\ptop}$ on the sequent.
+  \end{proof}
+
+  \begin{theorem}[Progress, $\cred$]%
     \label{thm:pgv-closed-progress-confs}
     If $\cseq[\main]{\emptyenv}{\conf{C}}$ and $\tm{\conf{C}}$ is in canonical form, then either $\tm{\conf{C}}\cred\tm{\conf{D}}$ for some $\tm{\conf{D}}$; or $\tm{\conf{C}\equiv\conf{D}}$ for some $\tm{\conf{D}}$ in normal form.
-  \end{restatabletheorem}
-  \begin{proof}[Sketch]
-    Our proof follows that of Kobayashi~\cite[theorem 2]{kobayashi06}.
+  \end{theorem}
+  % \begin{proof}[Sketch]
+  %   Our proof follows that of Kobayashi~\cite[theorem 2]{kobayashi06}.
 We apply \cref{lem:pgv-open-progress-terms} to each thread.
@@
 We move $\tm{L}$ into the $i^{\text{th}}$ position and repeat until we either find a reduction or reach normal form.
-  \end{proof}
+  % \end{proof}
+  \input{prf/thm-pgv-closed-progress-confs}
+  \input{fig/pgv-typing-sugar}
+
   \begin{corollary}%
@@
   \end{corollary}
-\end{compacttheorems}
 It follows immediately from \cref{thm:pgv-closed-progress-confs} and \cref{cor:pgv-closed-progress} that a term which does not return an endpoint will complete all its communication actions, thus satisfying deadlock freedom.
-\end{document}
-
-%%% Local Variables:
-%%% TeX-master: "priorities"
-%%% End:

### 2021-12-03 lmcs/introduction.tex
Update introduction.tex

--- 2021-04-26 introduction.tex
@@
-\documentclass[main.tex]{subfiles}
-
-\begin{document}
 \section{Introduction}
-Session types~\cite{honda93,takeuchihonda94,hondavasconcelos98} are types for protocols.
-Regular types ensure functions are used according to their specification.
+Session types~\cite{honda93,takeuchihonda94,hondavasconcelos98} are a type formalism used to specify and verify correct ordering of operations, namely protocols.
+Regular types ensure, for example that functions are used according to their specification.
 Session types ensure \emph{communication channels} are used according to their protocols.
 Session types have been studied in many settings.
-For instance, in the $\pi$-calculus~\cite{honda93,takeuchihonda94,hondavasconcelos98}, a foundational calculus for communication and concurrency, and in concurrent $\lambda$-calculi~\cite{gayvasconcelos10}, including the focus of our paper: Good Variation~\cite[GV]{wadler15,lindleymorris15}.
+Most notably, they have been defined for the $\pi$-calculus~\cite{honda93,takeuchihonda94,hondavasconcelos98}, a foundational calculus for communication and concurrency, and the concurrent $\lambda$-calculi~\cite{gayvasconcelos10}, including the main focus of our paper: Good Variation~\cite[GV]{wadler15,lindleymorris15}.
 
@@
 The benefit of working specifically with GV, as opposed to other concurrent $\lambda$-calculi, is its relation to CP~\cite{wadler14}, and its formal properties, including deadlock freedom.
+
 We thus pose our research question for GV:
+
 \begin{quotation}
@@
 \begin{enumerate}[labelindent=0pt,labelwidth=1.333\parindent,align=left,labelsep*=0pt,leftmargin=!]
-\item[(\secref{sec:pgv})] \textbf{Priority GV}.
-We present Priority GV (\secref{sec:pgv}, PGV), a session-typed functional language with priorities, and prove subject reduction (\cref{thm:pgv-subject-reduction-confs}) and progress (\cref{thm:pgv-closed-progress-confs}).
+\item[(\cref{sec:pgv})] \textbf{Priority GV}.
+We present Priority GV (\cref{sec:pgv}, PGV), a session-typed functional language with priorities, and prove \emph{subject reduction} (\cref{thm:pgv-subject-reduction-confs}) and \emph{progress} (\cref{thm:pgv-closed-progress-confs}).
 We addresses several problems in the original GV language, most notably:
-  \begin{enumerate*}
+  \begin{enumerate}
   \item PGV does not require the pseudo-type $S^\sharp$; and
   \item its structural congruence is type preserving.
-  \end{enumerate*}
+  \end{enumerate}
   PGV answers our research question positively as it allows cyclic-structured binary session-typed processes that are deadlock free.
-\item[(\secref{sec:pcp})] \textbf{Translation from PCP to PGV}.
-  We present a \emph{sound and complete encoding} of Priority CP~\cite{dardhagay18extended} in PGV (\secref{sec:pcp}).
+\item[(\cref{sec:pcp})] \textbf{Translation from PCP to PGV}.
+  We present a \emph{sound and complete encoding} of PCP~\cite{dardhagay18extended} in PGV (\cref{sec:pcp}).
 We prove the encoding preserves typing (\cref{thm:pcp-to-pgv-confs-preservation}) and satisfies operational correspondence (\cref{thm:pcp-to-pgv-operational-correspondence-soundness,thm:pcp-to-pgv-operational-correspondence-completeness}).
@@
 \end{enumerate}
-\end{document}

### 2021-12-03 lmcs/appendix.tex
Update appendix.tex

--- 2021-04-14 appendix.tex
@@
 % !TeX root = priorities.tex
-\section{Priority GV}
-\begingroup
-\usingnamespace{pgv}
-\restatelemma{lempgvvaluedone}
-\input{prf/lem-pgv-value-done}
+% \section{Priority GV}
+% \begingroup
+% \usingnamespace{pgv}
+% \restatelemma{lempgvvaluedone}
+% \input{prf/lem-pgv-value-done}
 
-\restatelemma{lempgvsubstitution}
-\input{prf/lem-pgv-substitution}
+% \restatelemma{lempgvsubstitution}
+% \input{prf/lem-pgv-substitution}
 
-\restatelemma{lempgvsubjectreductionterms}
-\input{prf/lem-pgv-subject-reduction-terms}
+% \restatelemma{lempgvsubjectreductionterms}
+% \input{prf/lem-pgv-subject-reduction-terms}
 
-\restatelemma{lempgvsubjectcongruence}
-\input{prf/lem-pgv-subject-congruence}
+% \restatelemma{lempgvsubjectcongruence}
+% \input{prf/lem-pgv-subject-congruence}
 
-\restatetheorem{thmpgvsubjectreductionconfs}
-\input{prf/thm-pgv-subject-reduction-confs}
+% \restatetheorem{thmpgvsubjectreductionconfs}
+% \input{prf/thm-pgv-subject-reduction-confs}
 
-\begin{restatablelemma}{lempgvreadypriority}
-  \label{lem:pgv-ready-priority}
-  If $\tseq[\cs{p}]{\ty{\Gamma}}{L}{T}$ is ready to act on $\tmty{x}{S}\in\ty{\Gamma}$, then the priority bound $\cs{p}$ is some priority $\cs{o}$, \ie not $\cs{\pbot}$ or $\cs{\ptop}$.
-\end{restatablelemma}
-\begin{proof}
-  Let $\tm{L}=\tm{\plug{E}{M}}$.
+% \begin{restatablelemma}{lempgvreadypriority}
+%   \label{lem:pgv-ready-priority}
+%   If $\tseq[\cs{p}]{\ty{\Gamma}}{L}{T}$ is ready to act on $\tmty{x}{S}\in\ty{\Gamma}$, then the priority bound $\cs{p}$ is some priority $\cs{o}$, \ie not $\cs{\pbot}$ or $\cs{\ptop}$.
+% \end{restatablelemma}
+% \begin{proof}
+%   Let $\tm{L}=\tm{\plug{E}{M}}$.
 By induction on the structure of $\tm{E}$.
@@
 No rule introduces the priority bound $\cs{\ptop}$ on the sequent.
-\end{proof}
+% \end{proof}
 
-\restatelemma{lempgvcanonicalforms}
-\begin{proof}
-  We move any $\nu$-binders to the top using \LabTirName{SC-ResExt}, discard any superfluous occurrences of $\tm{\child\;\unit}$ using \LabTirName{SC-ParNil}, and move the main thread to the rightmost position using \LabTirName{SC-ParComm} and \LabTirName{SC-ParAssoc}.
-\end{proof}
+% \restatelemma{lempgvcanonicalforms}
+% \begin{proof}
+%   We move any $\nu$-binders to the top using \LabTirName{SC-ResExt}, discard any superfluous occurrences of $\tm{\child\;\unit}$ using \LabTirName{SC-ParNil}, and move the main thread to the rightmost position using \LabTirName{SC-ParComm} and \LabTirName{SC-ParAssoc}.
+% \end{proof}
 
-\restatetheorem{thmpgvclosedprogressconfs}
-\input{prf/thm-pgv-closed-progress-confs}
+% \restatetheorem{thmpgvclosedprogressconfs}
+% \input{prf/thm-pgv-closed-progress-confs}
 
-\input{fig/pgv-typing-sugar}
-\endgroup
+% \input{fig/pgv-typing-sugar}
+% \endgroup
 

### 2021-12-06 lmcs/priority-cp-long.tex
pcp long

--- 2021-12-03 lmcs/appendix.tex
@@
 \section{Relation to Priority CP}
+\label{sec:pcp}
+
 \begingroup
@@
 
-\subsubsection*{Environments}
+\subsubsection*{Typing Environments}
 \label{sec:pcp-environments}
@@
 
-\subsubsection*{Duality}
+\subsubsection*{Type Duality}
 \label{sec:pcp-duality}
@@
 \]
+
 Process $\pcp{\tm{\link{x}{y}}}$ links endpoints $\pcp{\tm{x}}$ and $\pcp{\tm{y}}$ and forwards communication from one to the other.
@@
 
-\subsubsection*{Operational Semantics}
+\subsection{Operational Semantics}
 The operational semantics for PCP, given in \cref{fig:pcp-operational-semantics}, is defined as a reduction relation $\pcp{\red}$ on processes (bottom) and uses structural congruence (top).
@@
 
-\subsubsection*{Typing}
+\subsection{Typing Rules}
 \Cref{fig:pcp-typing} gives the typing rules for our version of PCP.
@@
 
-\begin{compacttheorems}
-    \begin{restatabledefinition}{defpcpactions}[Actions]
+\begin{definition}[Actions]
 A~process acts on an endpoint $\tm{x}$ if it is $\tm{\link{x}{y}}$, $\tm{\link{y}{x}}$, $\tm{\send{x}{y}{P}}$, $\tm{\recv{x}{y}{P}}$, $\tm{\close{x}{P}}$, $\tm{\wait{x}{P}}$, $\tm{\inl{x}{P}}$, $\tm{\inr{x}{P}}$, $\tm{\offer{x}{P}{Q}}$, or $\tm{\absurd{x}}$.
 A~process is an action if it acts on some endpoint $\tm{x}$.
-  \end{restatabledefinition}
-  \begin{restatabledefinition}{defpcpcanonicalforms}[Canonical Forms]
+\end{definition}
+\begin{definition}[Canonical Forms]
 \label{def:pcp-canonical-forms}
 A~process $\tm{P}$ is in canonical form if it is either $\tm{\halt}$ or of the form $\tm{\res{x_1}{x'_1}{\dots\res{x_n}{x'_n}{(P_1 \parallel\dots\parallel P_m)}}}$ where $m>0$ and each $\tm{P_j}$ is an action.
-  \end{restatabledefinition}
-  \begin{restatablelemma}{lempcpcanonicalforms}[Canonical Forms]
+\end{definition}
+\begin{lemma}[Canonical Forms]
 \label{lem:pcp-canonical-forms}
 If $\seq{P}{\ty{\Gamma}}$, there exists some $\tm{Q}$ such that $\tm{P}\equiv\tm{Q}$ and $\tm{Q}$ is in canonical form.
-  \end{restatablelemma}
+\end{lemma}
 \begin{proof}
-    If $\tm{P}=\tm{\halt}$, we are done.
-Otherwise, we move any $\nu$-binders to the top using \LabTirName{SC-ResExt}, and discard any superfluous occurrences of $\tm{\halt}$ using \LabTirName{SC-ParNil}.
+If $\tm{P}=\tm{\halt}$, we are done. Otherwise, we move any $\nu$-binders to the top using \LabTirName{SC-ResExt}, and discard any superfluous occurrences of $\tm{\halt}$ using \LabTirName{SC-ParNil}.
 \end{proof}
-  \restatetheorem{thmpcpclosedprogress}
+
+%\restatetheorem{thmpcpclosedprogress}
+\begin{restatabletheorem}{thmpcpclosedprogress}[Progress, $\pcp{\red}$]%
+  \label{thm:pcp-closed-progress}
+  \hfill\\%newline before theorem statement
+  If $\pcp{\seq{P}{\emptyenv}}$, then either $\pcp{\tm{P}=\tm{\halt}}$ or there exists a $\pcp{\tm{Q}}$ such that $\pcp{\tm{P}\red\tm{Q}}$.
+\end{restatabletheorem}
 \input{prf/thm-pcp-closed-progress}
-\end{compacttheorems}
+
 \endgroup
@@
 \begingroup
-\restatelemma{lempcptopgvtermspreservation}
+%\restatelemma{lempcptopgvtermspreservation}
+\begin{lemma}[Preservation, ${\tm{\cpgvM{\cdot}}}$]%
+  \label{lem:pcp-to-pgv-terms-preservation}
+  If $\pcp{\seq{P}{\ty{\Gamma}}}$, then $\pgv{\tseq[\cs{p}]{\ty{\cpgvT{\Gamma}}}{\cpgvM{P}}{\tyunit}}$.
+\end{lemma}
 \input{prf/lem-pcp-to-pgv-terms-preservation}
 
-\restatetheorem{thmpcptopgvconfspreservation}
+%\restatetheorem{thmpcptopgvconfspreservation}
+\begin{theorem}[Preservation, ${\tm{\cpgvC{\cdot}}}$]%
+  \label{thm:pcp-to-pgv-confs-preservation}
+  If $\pcp{\seq{P}{\ty{\Gamma}}}$, then $\pgv{\cseq[\child]{\ty{\cpgvT{\Gamma}}}{\cpgvC{P}}}$.
+\end{theorem}
 \input{prf/thm-pcp-to-pgv-confs-preservation}
 
-\restatetheorem{thmpcptopgvoperationalcorrespondencesoundness}
+%\restatetheorem{thmpcptopgvoperationalcorrespondencesoundness}
+\begin{theorem}%
+  [Operational Correspondence, Soundness, ${\tm{\cpgvC{\cdot}}}$]%
+  \label{thm:pcp-to-pgv-operational-correspondence-soundness}
+  \hfill\\%newline before theorem statement
+  If $\pcp{\seq{P}{\ty{\Gamma}}}$ and $\pgv{\tm{\cpgvC{P}}\cred\tm{\conf{C}}}$, there exists a $\tm{Q}$ such that $\pcp{\tm{P}\red^+\tm{Q}}$ and $\pgv{\tm{\conf{C}}\cred^\star\tm{\cpgvC{Q}}}$
+\end{theorem}
 \input{prf/thm-pcp-to-pgv-operational-correspondence-soundness}
 
-\restatelemma{lempcptopgvcpgvMtocpgvC}
+%\restatelemma{lempcptopgvcpgvMtocpgvC}
+\begin{lemma}%
+  \label{lem:pcp-to-pgv-cpgvM-to-cpgvC}
+  For any $\pcp{\tm{P}}$, either:
+  \begin{itemize}
+  \item $\pgv{\tm{\child\;\cpgvM{P}}=\tm{\cpgvC{P}}}$; or
+  \item $\pgv{\tm{\child\;\cpgvM{P}}\cred^+\tm{\cpgvC{P}}}$, and for any $\pgv{\tm{\conf{C}}}$, if $\pgv{\tm{\child\;\cpgvM{P}}\cred\tm{\conf{C}}}$, then $\pgv{\tm{\conf{C}}\cred^\star\tm{\cpgvC{P}}}$.
+  \end{itemize}
+\end{lemma}
 \input{prf/lem-pcp-to-pgv-cpgvM-to-cpgvC}
 
-\begin{restatablelemma}{lempcptopgvoperationalcorrespondenceequiv}%
+\begin{lemma}%
   \label{lem:pcp-to-pgv-confs-operational-correspondence-equiv}
@@
   then $\pgv{\tm{\cpgvC{P}}\equiv\tm{\cpgvC{Q}}}$.
-\end{restatablelemma}
+\end{lemma}
 \begin{proof}
@@
 
-\restatetheorem{thmpcptopgvoperationalcorrespondencecompleteness}
+%\restatetheorem{thmpcptopgvoperationalcorrespondencecompleteness}
+\begin{theorem}%
+  [Operational Correspondence, Completeness, ${\tm{\cpgvC{\cdot}}}$]%
+  \label{thm:pcp-to-pgv-operational-correspondence-completeness}
+  \hfill\\%newline before theorem statement
+  If $\pcp{\seq{P}{\ty{\Gamma}}}$ and $\pcp{\tm{P}\red\tm{Q}}$,
+  then $\pgv{\tm{\cpgvC{P}}\cred^+\tm{\cpgvC{Q}}}$.
+\end{theorem}
 \input{prf/thm-pcp-to-pgv-operational-correspondence-completeness}

### 2021-12-06 lmcs/introduction.tex
polishing

@@
 
-\begin{quotation}
+\vspace{1em}
   \textbf{RQ:}
-  Can we design a more expressive GV which guarantees deadlock freedom for cyclic-structured processes?
-\end{quotation}
+  \emph{Can we design a more expressive GV which guarantees deadlock freedom for cyclic-structured processes?}
+\vspace{1em}
 
@@
 This closes the circle of the connection between CP and GV~\cite{wadler14}, and their priority-based versions, PCP~\cite{dardhagay18extended} and PGV.
+
 We make the following main contributions:
-\begin{enumerate}[labelindent=0pt,labelwidth=1.333\parindent,align=left,labelsep*=0pt,leftmargin=!]
-\item[(\cref{sec:pgv})] \textbf{Priority GV}.
+\begin{enumerate}
+\item \textbf{Priority GV}.
 We present Priority GV (\cref{sec:pgv}, PGV), a session-typed functional language with priorities, and prove \emph{subject reduction} (\cref{thm:pgv-subject-reduction-confs}) and \emph{progress} (\cref{thm:pgv-closed-progress-confs}).
+
 We addresses several problems in the original GV language, most notably:
   \begin{enumerate}
-  \item PGV does not require the pseudo-type $S^\sharp$; and
-  \item its structural congruence is type preserving.
+  \item PGV does not require the pseudo-type $S^\sharp$;
+  \item Structural congruence is type preserving.
   \end{enumerate}
   PGV answers our research question positively as it allows cyclic-structured binary session-typed processes that are deadlock free.
-\item[(\cref{sec:pcp})] \textbf{Translation from PCP to PGV}.
+\item \textbf{Translation from PCP to PGV}.
   We present a \emph{sound and complete encoding} of PCP~\cite{dardhagay18extended} in PGV (\cref{sec:pcp}).
 We prove the encoding preserves typing (\cref{thm:pcp-to-pgv-confs-preservation}) and satisfies operational correspondence (\cref{thm:pcp-to-pgv-operational-correspondence-soundness,thm:pcp-to-pgv-operational-correspondence-completeness}).
+
   To obtain a tight correspondence, we update PCP, moving away from commuting conversions and reduction as cut elimination towards reduction based on structural congruence, as it is standard in process calculi.

### 2021-12-06 lmcs/priority-gv.tex
work on pgv

@@
 
+\subsection{Syntax of Types and Terms}
 \subsubsection*{Session types}
@@
 
-\subsubsection*{Types}
+\subsubsection*{Regular Types}
 Types ($\ty{T}$, $\ty{U}$) are defined by the following grammar:
@@
 
-\subsubsection*{Environments}
+\subsubsection*{Typing Environments}
 Typing environments $\ty{\Gamma}$, $\ty{\Delta}$ associate types to names.
@@
 
-\subsubsection*{Duality}
+\subsubsection*{Type Duality}
 Duality plays a crucial role in session types.
@@
 \]
+
 As the syntax for units suggests, these are the binary and nullary forms of the more common n-ary choice constructs $\ty{{\oplus}^{\cs{o}}\{l_i:S_i\}_{i\in{I}}}$ and $\ty{{\with}^{\cs{o}}\{l_i:S_i\}_{i\in{I}}}$, which one may obtain generalising the sum types to variant types.
@@
 
-\subsubsection*{Operational Semantics}
+\subsection{Operational Semantics}
+\subsubsection*{Configurations}
 Priority GV terms are evaluated as part of a configuration of processes. Configurations are defined by the following grammar:
@@
 \]
+
 Configurations ($\tm{\conf{C}}$, $\tm{\conf{D}}$, $\tm{\conf{E}}$) consist of threads $\tm{\phi\;M}$, parallel compositions $\tm{\ppar{\conf{C}}{\conf{D}}}$, and name restrictions $\tm{\res{x}{x'}{\conf{C}}}$.
@@
 \]
+
 The use of $\tm{\child}$ for child threads~\cite{lindleymorris15} overlaps with the use of the meta-variable $\cs{o}$ for priorities~\cite{dardhagay18extended}.
@@
 
+\subsubsection*{Values}
 Values ($\tm{V}$, $\tm{W}$), evaluation contexts ($\tm{E}$), thread evaluation contexts ($\tm{\conf{F}}$), and configuration contexts ($\tm{\conf{G}}$) are defined by the following grammars:
@@
 
+\subsubsection*{Reduction Relation}
 \input{fig/pgv-operational-semantics}
@@
 
-\subsubsection*{Typing}
-\Cref{fig:pgv-typing} gives the typing rules for PGV.
+\subsection{Typing Rules}
+\input{fig/pgv-typing}
+
+\subsubsection*{Terms Typing}
+%\Cref{fig:pgv-typing} gives the typing rules for PGV.
 Typing rules for terms are at the top of \cref{fig:pgv-typing}.
@@
 However, the rules can easily be extended to allow values with unrestricted usage~\cite{wadler14}.
-
-\input{fig/pgv-typing}
 
@@
 \begin{mathpar}
-  \small
   \inferrule*{
@@
 \end{mathpar}
+
 Closures suspend communication, so \LabTirName{T-Lam} stores the priority bounds of the function body on the function type, and \LabTirName{T-App} restores them.
@@
 
+\subsubsection*{Configurations Typing}
 Typing rules for configurations are at the bottom of \cref{fig:pgv-typing}.
@@
 While our configuration typing is based on the standard typing for GV, it differs on two counts:
-\begin{enumerate*}[label= (\roman*) ]
+\begin{enumerate}[label= (\roman*) ]
 \item \emph{we require that child threads return the unit value}, as opposed to a terminated channel; and
 \item \emph{we simplify typing for parallel composition}.
-\end{enumerate*}
+\end{enumerate}
 In order to guarantee deadlock freedom, in GV each parallel composition must split \emph{exactly one} channel of the channel pseudo-type $\ty{S^\sharp}$ into two endpoints of type $\ty{S}$ and $\ty{\co{S}}$.
@@
 
+\subsubsection*{Syntactic Sugar Typing}
+\input{fig/pgv-typing-sugar}
+The following typing rules given in \cref{{fig:pgv-typing-sugar},fig:pgv-typing-sugar-select,fig:pgv-typing-sugar-offer}, cover syntactic sugar typing for PGV.
+
 \section{Technical Developments}
-\subsection{Subject reduction}
+\subsection{Subject Reduction}
 Unlike with previous versions of GV, structural congruence, term reduction, and configuration reduction are all type preserving.
@@
 
-\subsection{Progress and deadlock freedom}
+\subsection{Progress and Deadlock Freedom}
 PGV satisfies progress, as PGV configurations either reduce or are in normal form.
@@
 This behaviour is not new to PGV.
+
 Let us consider an example, adapted from Lindley and Morris~\cite{lindleymorris15}, in which a term returns an endpoint linked to an echo server.
@@
 \]
+
 If we reduce the above program, we get $\tm{\res{x}{x'}{(\ppar{\child\;\echo_{x}}{\main\;x'})}}$.
@@
 Progress for the term language is standard for GV, and deviates from progress for linear $\lambda$-calculus only in that terms may reduce to values or \emph{ready terms}:
-\begin{compacttheorems}
 \begin{definition}[Actions]%
@@
 \end{definition}
-  \begin{restatablelemma}{lempgvopenprogressterms}[Progress, $\tred$]%
+\begin{lemma}[Progress, $\tred$]%
   \label{lem:pgv-open-progress-terms}
@@
   \end{enumerate*}
-  \end{restatablelemma}
-\end{compacttheorems}
+\end{lemma}
 
@@
     A~configuration $\tm{\conf{C}}$ is in canonical form if it is of the form 
-
     $\tm{\res{x_1}{x'_1}{\dots\res{x_n}{x'_n}{(\child\;M_1\parallel\dots\parallel\child\;M_m\parallel\main\;N)}}}$ 
-
     where no term $\tm{M_i}$ is a value.
@@
     A~configuration $\tm{\conf{C}}$ is in normal form if it is of the form 
-
     $\tm{\res{x_1}{x'_1}{\dots\res{x_n}{x'_n}{(\child\;M_1\parallel\dots\parallel\child\;M_m\parallel\main\;V)}}}$ 
-
     where each $\tm{M_i}$ is ready to act on $\tm{x_i}$.
@@
   \input{prf/thm-pgv-closed-progress-confs}
-  \input{fig/pgv-typing-sugar}
 

### 2021-12-07 lmcs/priority-gv.tex
top down polish

@@
 \]
+
 Session types $\ty{\tysend[\cs{o}]{T}{S}}$ and $\ty{\tyrecv[\cs{o}]{T}{S}}$ describe the endpoints of a channel over which we send or receive a value of type $\ty{T}$, and then proceed as $\ty{S}$.
@@
 
-\subsubsection*{Regular Types}
+\subsubsection*{Types}
 Types ($\ty{T}$, $\ty{U}$) are defined by the following grammar:
@@
 \]
+
 Types $\ty{\typrod{T}{U}}$, $\ty{\tyunit}$, $\ty{\tysum{T}{U}}$, and $\ty{\tyvoid}$ are the standard linear $\lambda$-calculus product type, unit type, sum type, and empty type.
@@
 Constants are best understood in conjunction with their typing and reduction rules in~\cref{fig:pgv-typing,fig:pgv-operational-semantics}.
+
 Briefly, $\tm{\link}$ links two endpoints together, forwarding messages from one to the other, $\tm{\new}$ creates a new channel and returns a pair of its endpoints, and $\tm{\spawn}$ spawns off its argument as a new thread.
+
 The $\tm{\send}$ and $\tm{\recv}$ functions send and receive values on a channel.
 However, since the typing rules for PGV ensure the linear usage of endpoints, they also return a new copy of the endpoint to continue the session.
+
 The $\tm{\close}$ and $\tm{\wait}$ functions close a channel.
+
 We use syntactic sugar to make terms more readable: we write $\tm{\letbind{x}{M}{N}}$ in place of $\tm{(\lambda x.N)\;M}$, $\tm{\lambda\unit.M}$ in place of $\tm{\lambda z.\andthen{z}{M}}$, and $\tm{\lambda\pair{x}{y}.M}$ in place of $\tm{\lambda z.\letpair{x}{y}{z}{M}}$.
@@
 While our configuration reduction is based on the standard evaluation for GV, the increased expressiveness of PGV allows us to simplify the relation on two counts.
-\begin{enumerate*}[label=(\alph*)]
+\begin{enumerate}[label=(\alph*)]
 \item
@@
 Consequently, we also simplify the evaluation of the $\tm{\link}$ construct.
+
 Intuitively, evaluating $\tm{\link}$ causes a substitution: if we have a channel bound as $\tm{\res{x}{y}{}}$, then $\tm{\link\;\pair{w}{x}}$ replaces all occurrences of $\tm{y}$ by $\tm{w}$.
@@
 In GV, this is not the case, and subject reduction relies on proving that if $\equiv\cred$ ends up in an ill-typed configuration, we can rewrite it to a well-typed configuration using $\equiv$.
-\end{enumerate*}
+\end{enumerate}
 
@@
 \end{enumerate}
+
 In order to guarantee deadlock freedom, in GV each parallel composition must split \emph{exactly one} channel of the channel pseudo-type $\ty{S^\sharp}$ into two endpoints of type $\ty{S}$ and $\ty{\co{S}}$.
@@
 \end{corollary}
-It follows immediately from \cref{thm:pgv-closed-progress-confs} and \cref{cor:pgv-closed-progress} that a term which does not return an endpoint will complete all its communication actions, thus satisfying deadlock freedom.
+
+An immediate consequence of \cref{thm:pgv-closed-progress-confs} and \cref{cor:pgv-closed-progress} is that \emph{a term which does not return an endpoint will complete all its communication actions, thus satisfying deadlock freedom.}

### 2022-11-07 lmcs/conclusion.tex
restructuring before reworking
polishing
Created TODO.md, started fixes

--- 2021-04-26 conclusion.tex
@@
-% !TeX root = priorities.tex
-\documentclass[main.tex]{subfiles}
-
-\begin{document}
-
 \section{Related Work and Discussion}
@@
 For the `standard' typed $\pi$-calculus, an important line of work starts from Kobayashi's approach to deadlock freedom~\cite{kobayashi98}, where priorities are values from an abstract poset.
-Kobayashi~\cite{kobayashi06} simplifies the abstract poset to pairs of naturals, called \emph{obligations} and\emph{capabilities}.
+Kobayashi~\cite{kobayashi06} simplifies the abstract poset to pairs of naturals, called \emph{obligations} and \emph{capabilities}.
 Padovani simplifies these further to a single natural, called a \emph{priority}~\cite{padovani14}, and adapts obligations/capabilities to session types~\cite{padovani13}.
@@
 
-\subsubsection*{Acknowledgements}
-The authors would like to thank Simon Fowler, April Gon\c{c}alves, and Philip Wadler for their comments on the manuscript.
-
-\end{document}
+% \subsubsection*{Acknowledgements}
+% The authors would like to thank Simon Fowler, April Gon\c{c}alves, and Philip Wadler for their comments on the manuscript.

### 2022-11-07 lmcs/introduction.tex
Created TODO.md, started fixes

@@
 In the $\pi$-calculus literature, there have been several attempts at developing Curry-Howard correspondences between session-typed $\pi$-calculus and linear logic~\cite{girard87}: Caires and Pfenning's $\pi$DILL~\cite{cairespfenning10} corresponds to dual intuitionistic linear logic~\cite{barber96}, and Wadler's Classical Processes~\cite[CP]{wadler14} corresponds to classical linear logic~\cite[CLL]{girard87}.
-Both calculi guarantee deadlock freedom, which they achieve by restricting structure of processes and shared channels to \emph{trees}, by combing name restriction and parallel composition into a single construct, corresponding to the logical cut.
+Both calculi guarantee deadlock freedom, which they achieve by restricting structure of processes and shared channels to \emph{trees}, by combining name restriction and parallel composition into a single construct, corresponding to the logical cut.
 This ensures that two processes can only communicate via exactly one series of channels, which rules out interleavings of sessions, and guarantees deadlock freedom.

### 2022-11-07 lmcs/priority-gv.tex
Created TODO.md, started fixes

@@
     \begin{array}{l}
-      \letpair{x}{x'}{\new}{}
+      \letpair{x}{x'}{\new\unit}{}
       \\
-      \letpair{y}{y'}{\new}{}
+      \letpair{y}{y'}{\new\unit}{}
       \\
-      \spawn\left(%
+      \spawn\left(\lambda\unit.%
       \begin{array}{l}
@@
   \sep      \tm{\child}
-  \hfill
+  \hfill\qquad\hfill
   \tm{\conf{C}}, \tm{\conf{D}}, \tm{\conf{E}}
@@
   \sep      \tm{\res{x}{x'}{\conf{C}}}
-  \hfill
 \]
@@
   \tm{\main}  + \tm{\child} = \tm{\main}
-  \hfill
+  \hfill\qquad\hfill
   \tm{\child} + \tm{\main}  = \tm{\main}
-  \hfill
+  \hfill\qquad\hfill
   \tm{\child} + \tm{\child} = \tm{\child}
-  \hfill
+  \hfill\qquad\hfill
   (\tm{\main}  + \tm{\main} \; \text{is undefined})
-  \hfill
+  \hfill\qquad\hfill
 \]
@@
 Structural congruence satisfies the following axioms: $\LabTirName{SC-LinkSwap}$ allows swapping channels in the link process.
-$\LabTirName{SC-ResLink}$ allows restriction to applied to link which is structurally equivalent to the terminated process, thus allowing elimination of unnecessary restrictions.
+$\LabTirName{SC-ResLink}$ allows restriction to be applied to link which is structurally equivalent to the terminated process, thus allowing elimination of unnecessary restrictions.
 $\LabTirName{SC-ResSwap}$ allows swapping channels and $\LabTirName{SC-ResComm}$ states that restriction is commutative.
@@
 Consequently, we can embed it directly into the reduction relation.
-In GV, this is not the case, and subject reduction relies on proving that if $\equiv\cred$ ends up in an ill-typed configuration, we can rewrite it to a well-typed configuration using $\equiv$.
+In GV, this is not the case, and subject reduction relies on proving that if $\equiv\cred$ ends up in an ill-typed configuration, we can rewrite it to a well typed configuration using $\equiv$.
 \end{enumerate}
@@
 Crucially, it composes the terms below in parallel.
-While each of these terms itself is well-typed, they impose opposite conditions on the priorities, so their composition is ill-typed.
+While each of these terms itself is well typed, they impose opposite conditions on the priorities, so their composition is ill-typed.
 (We omit the priorities on $\ty{\tyends}$ and $\ty{\tyendr}$.)
@@
 PGV satisfies progress, as PGV configurations either reduce or are in normal form.
-However, the normal forms may seem surprising at first, as evaluating a well-typed PGV term does not necessarily produce \emph{just} a value.
+However, the normal forms may seem surprising at first, as evaluating a well typed PGV term does not necessarily produce \emph{just} a value.
 If a term returns an endpoint, then its normal form contains a thread which is ready to communicate on the dual of that endpoint.
@@
 Furthermore, no term $\tm{M_i}$ can perform a communication action by itself, since those are excluded by the definition of actions.
-Finally, as a corollary, we get that well-typed terms which do not return endpoints return \emph{just} a value:
+Finally, as a corollary, we get that well typed terms which do not return endpoints return \emph{just} a value:
 

### 2022-11-07 lmcs/conclusion.tex
Fix typos & more uses of hfill

@@
 Carbone~\etal~\cite{carbonedardha14} use catalysers to show that progress is a compositional form of lock-freedom and can be lifted to session types via the encoding of session types to linear types~\cite{kobayashi07,dardhagiachino12,dardha14beat}.
-Vieira and Vasconcelos~\cite{vieiravasconcelos13} use single priorities and an abstract partial order to guarantee deadlock freedom in a binary session-typed $\pi$-calculus and building on conservation types.
+Vieira and Vasconcelos~\cite{vieiravasconcelos13} use single priorities and an abstract partial order to guarantee deadlock freedom in a binary session-typed $\pi$-calculus and building on conversation types.
 

### 2022-11-07 lmcs/introduction.tex
Fix typos & more uses of hfill

@@
 GV is a concurrent $\lambda$-calculus with \emph{binary} session types, where each channel is shared between exactly two processes.
-Binary session types guarantee two crucial properties \emph{communication safety}---\eg, if the protocol says to transmit an integer, you transmit an integer---and \emph{session fidelity}---\eg, if the protocol says send, you send.
+Binary session types guarantee two crucial properties: \emph{communication safety}---\eg, if the protocol says to transmit an integer, you transmit an integer---and \emph{session fidelity}---\eg, if the protocol says send, you send.
 A third crucial property is \emph{deadlock freedom}, which ensures that processes do not have cyclic dependencies---\eg, when two processes wait for each other to send a value.

### 2022-11-07 lmcs/introduction.tex
Fixed typo. Rewrote first line.

@@
 \section{Introduction}
-Session types~\cite{honda93,takeuchihonda94,hondavasconcelos98} are a type formalism used to specify and verify correct ordering of operations, namely protocols.
-Regular types ensure, for example that functions are used according to their specification.
-Session types ensure \emph{communication channels} are used according to their protocols.
+Session types~\cite{honda93,takeuchihonda94,hondavasconcelos98} are a type formalism that ensures \emph{communication channels} are used according to their protocols, much like, \eg, data types ensure that functions are used according to their signature.
 Session types have been studied in many settings.

### 2022-11-07 lmcs/priority-cp-long.tex
top down polish
Created TODO.md, started fixes
Fixed typo. Rewrote first line.

@@
 \]
+
 Each connective is annotated with a priority $\cs{o}\in\mathbb{N}$.
+
 Types $\pcp{\ty{\tytens[\cs{o}]{A}{B}}}$ and $\pcp{\ty{\typarr[\cs{o}]{A}{B}}}$ type the endpoints of a channel over which we send or receive a channel of type $\pcp{\ty{A}}$, and then proceed as type $\pcp{\ty{B}}$.
@@
 These two types act as units for $\pcp{\ty{\tytens[\cs{o}]{A}{B}}}$ and $\pcp{\ty{\typarr[\cs{o}]{A}{B}}}$, respectively.
+
 Types $\pcp{\ty{\typlus[\cs{o}]{A}{B}}}$ and $\pcp{\ty{\tywith[\cs{o}]{A}{B}}}$ type the endpoints of a channel over which we can receive or send a choice between two branches $\pcp{\ty{A}}$ or $\pcp{\ty{B}}$.
 We have opted for a simplified version of choice and followed the original Wadler's CP \cite{wadler14}, however types $\ty{\oplus}$ and $\ty{\with}$ can be trivially generalised to $\pcp{\ty{\oplus^{\cs{o}}\{l_i:A_i\}_{i\in I}}}$ and $\pcp{\ty{\with^{\cs{o}}\{l_i:A_i\}_{i\in I}}}$, respectively, as in the original PCP \cite{dardhagay18extended}.
+
 Types $\pcp{\ty{\tynil[\cs{o}]}}$ and $\pcp{\ty{\tytop[\cs{o}]}}$ type the endpoints of a channel over which we can send or receive a choice between \emph{no options}.
@@
 $\pcp{\tm{\res{x}{y}{P}}}$, $\pcp{\tm{(\ppar{P}{Q})}}$ and $\pcp{\tm{\halt}}$ denote respectively the restriction processes where channel endpoints $\pcp{\tm{x}}$ and $\pcp{\tm{y}}$ are bound together and with scope $\pcp{\tm{P}}$, the parallel composition of processes $\pcp{\tm{P}}$ and $\pcp{\tm{Q}}$ and the terminated process.
+
 Processes $\pcp{\tm{\send{x}{y}{P}}}$ and $\pcp{\tm{\recv{x}{y}{P}}}$ send or receive over channel $\pcp{\tm{x}}$ a value $\pcp{\tm{y}}$ and proceed as process $\pcp{\tm{P}}$.
 Processes $\pcp{\tm{\close{x}{P}}}$ and $\pcp{\tm{\wait{x}{P}}}$ send and receive an empty value---denoting the closure of channel $\pcp{\tm{x}}$, and continue as $\pcp{\tm{P}}$.
+
 Processes $\pcp{\tm{\inl{x}{P}}}$ and $\pcp{\tm{\inr{x}{P}}}$ make a left and right choice, respectively and proceed as process $\pcp{\tm{P}}$.
 Dually, $\pcp{\tm{\offer{x}{P}{Q}}}$ offers both left and right branches, with continuations $\pcp{\tm{P}}$ and $\pcp{\tm{Q}}$, and $\pcp{\tm{\absurd{x}}}$ is the empty offer.
+
 We write \emph{unbound} send as $\pcp{\tm{\usend{x}{y}{P}}}$, which is syntactic sugar for $\pcp{\tm{\send{x}{z}{(\ppar{\link{y}{z}}{P})}}}$.
@@
 \textsc{T-Link} states that the link process $\pcp{\tm{\link{x}{y}}}$ is well typed under channels $\pcp{\tm{x}}$ and $\pcp{\tm{y}}$ having dual types, respectively $\pcp{\ty{A}}$ and $\pcp{\ty{\co{A}}}$.
-\textsc{T-Res} sates that the restriction process $\pcp{\tm{\res{x}{y}{P}}}$ is well typed under typing context $\pcp{\ty{\Gamma}}$ if process $\pcp{\tm{P}}$ is well typed in $\pcp{\ty{\Gamma}}$ augmented with channel endpoints $\pcp{\tm{x}}$ and $\pcp{\tm{y}}$ having dual types, respectively $\pcp{\ty{A}}$ and $\pcp{\ty{\co{A}}}$.
+\textsc{T-Res} states that the restriction process $\pcp{\tm{\res{x}{y}{P}}}$ is well typed under typing context $\pcp{\ty{\Gamma}}$ if process $\pcp{\tm{P}}$ is well typed in $\pcp{\ty{\Gamma}}$ augmented with channel endpoints $\pcp{\tm{x}}$ and $\pcp{\tm{y}}$ having dual types, respectively $\pcp{\ty{A}}$ and $\pcp{\ty{\co{A}}}$.
 \textsc{T-Par} states that the parallel composition of processes $\pcp{\tm{P}}$ and $\pcp{\tm{Q}}$ is well typed under the disjoint union of their respective typing contexts.
 \textsc{T-Halt} states that the terminated process $\pcp{\tm{\halt}}$ is well typed in the empty context.
+
 \textsc{T-Send} and \textsc{T-Recv} state that the sending and receiving of a bound name $\pcp{\tm{y}}$ over a channel $\pcp{\tm{x}}$ is well typed under $\pcp{\ty{\Gamma}}$ and $\pcp{\tm{x}}$ of type $\pcp{\ty{\tytens[\cs{o}]{A}{B}}}$, respectively $\pcp{\typarr[\cs{o}]{A}{B}}$.
 Priority $\cs{o}$ is the smallest among all priorities of the types used by the output or input process, captured by the side condition $\pcp{\cs{o}<\minpr(\ty{\Gamma},\ty{A},\ty{B})}$.
+
 Rules \textsc{T-Close} and \textsc{T-Wait} type the closure of channel $\pcp{\tm{x}}$ and are in the same lines as the previous two rules, requiring that the priority of channel $\pcp{\tm{x}}$ is the smallest among all priorities in $\pcp{\ty{\Gamma}}$.
+
 \textsc{T-Select-Inl} and \textsc{T-Select-Inr} type respectively the left $\pcp{\tm{\inl{x}{P}}}$ and right $\pcp{\tm{\inr{x}{P}}}$ choice performed on channel $\pcp{\tm{x}}$.
@@
 
+\Cref{fig:pcp-typing-sugar} shows how syntactic sugar in PCP is well typed.
+\input{fig/pcp-typing-sugar}
+
 Finally, since our reduction relation is a strict subset of the reduction relation in the original~\cite{dardhagay18extended}, we defer to their proofs.
@@
 
-% Proofs that the syntactic sugar is well-typed.
-\input{fig/pcp-typing-sugar}
-
+\subsection{Technical Developments}
 \begin{definition}[Actions]

### 2022-11-07 lmcs/priority-gv.tex
Fix typos & more uses of hfill
Fixed typo. Rewrote first line.

@@
 The upper bound is written on the sequent, whereas the lower bound is approximated from the typing environment.
-Typing rules for sequential constructs enforce sequentially, \eg, the typing for $\tm{\andthen{M}{N}}$ has a side condition which requires that the upper bound of $\tm{M}$ is smaller than the lower bound of $\tm{N}$, \ie, $\tm{M}$ finishes before $\tm{N}$ starts.
+Typing rules for sequential constructs enforce sequentiality, \eg, the typing for $\tm{\andthen{M}{N}}$ has a side condition which requires that the upper bound of $\tm{M}$ is smaller than the lower bound of $\tm{N}$, \ie, $\tm{M}$ finishes before $\tm{N}$ starts.
 The typing rule for $\tm{\new}$ ensures that both endpoints of a channel share the same priorities.
@@
   \end{array}
-  \hfill
+  \hfill\qquad\hfill
   \begin{array}{lcl}
@@
   \end{array}
-  \hfill
 \]

### 2022-11-16 lmcs/priority-cp-long.tex
text for sec 4

@@
 \label{sec:pcp}
+
+Thus far we have presented Priority GV (PGV) together with the relevant technical results.
+We remind the reader that this line of work of adding priorities, started with Priority CP (PCP) \cite{dardhagay18} where priorities are integrated in Wadler's Classical Processes (CP), which is a $\pi$-calculus leveraging the correspondence of session types as linear logic propositions \cite{wadler12}.
+In his work, Wadler presents a connection (via encoding) of CP and GV.
+Following that work, we sat out to understand the connection between the priority versions of CP and GV, thus comparing PGV and PCP.
+Before presenting our formal results, we will revisit PCP in the following section.
 

### 2022-11-16 lmcs/introduction.tex
LCC update

@@
 There are many downsides to combining name restriction and parallel composition, such as lack of modularity, difficulty typing structural congruence and formulating label-transition semantics, which have led to various approaches to decoupling these constructs.
-Hypersequent CP~\cite{MP18,kokkemontesi19popl,kokkemontesi19tlla} and Linear Compositional Choreographies~\cite{CarboneMS18} decouple them, but maintain the correspondence to CLL and allow only tree-structured processes.
+Hypersequent CP~\cite{MP18,kokkemontesi19popl,kokkemontesi19tlla} and Linear Compositional Choreographies~\cite{CarboneMS18} decouple them, but maintain the correspondence to linear logic and allow only tree-structured processes.
 Priority CP~\cite[PCP]{dardhagay18extended} weakens the correspondence to CLL in exchange for a more expressive language which allows cyclic-structured processes.
@@
 \end{enumerate}
+
+This paper is an improved and extended version of a paper published at FORTE 2021 international conference \cite{kokkedardha21pgv}.
+We present detailed examples and complete proofs of our technical results.

### 2022-11-17 lmcs/conclusion.tex
clarifying some related work

@@
 
-While our work focuses on \emph{binary} session types, it is worth to discuss related work on Multiparty Session Types (MPST).
-The line of work on MPST starts with Honda~\etal~\cite{hondayoshida08}, which guarantees deadlock freedom \emph{within a single session}, but not for session interleaving.
+While our work focuses on \emph{binary} session types, it is worth to discuss related work on Multiparty Session Types (MPST), which describe communication among multiple agents in a distributed setting.
+The line of work on MPST starts with Honda~\etal~\cite{hondayoshida08}, which guarantees deadlock freedom \emph{within a single session}, but not for several session interleaving.
 Bettini~\etal~\cite{bettinicoppo08} follow a technique similar to Kobayashi's for MPST.
 The main difference with our work is that we associate priorities with communication actions, where Bettini~\etal~\cite{bettinicoppo08} associate them with channels.
-Carbone and Montesi~\cite{carbonemontesi13} combine MPST with choreographies and obtain a formalism that satisfies deadlock freedom.
-Deni\'{e}lou and Yoshida~\cite{DenielouY13} introduce \emph{multiparty compatibility} which generalises duality in binary session types.
-They synthesise safe and deadlock-free global types from local types leveraging LTSs and communicating automata.
+Carbone and Montesi~\cite{carbonemontesi13} combine MPST with choreographic programming and obtain a formalism that satisfies deadlock freedom.
+In the same vein as MPST, choreographic programming specifies communication among all participants in a distributed system.
+While MPST target mainly protocol descriptions, choreographies have mainly targetted implementations and programming languages as they are suitable for describing concrete system implementations.
+Deni\'{e}lou and Yoshida~\cite{DenielouY13} introduce \emph{multiparty compatibility}, which generalises the notion of duality in binary session types.
+They synthesise safe and deadlock-free \emph{global types}--specifying communication among \emph{all} involved participants, from \emph{local session types}--specifying communication from the viewpoint of \emph{one} participant.
+To do so, they leverage Labelled Transition Systems (LTSs) and communicating automata.
 Castellani~\etal~\cite{CastellaniDGH20} guarantee lock freedom, a stronger property than deadlock freedom, for MPST with \emph{internal delegation}, where participants in the same session are allowed to delegate tasks to each other, and internal delegation is captured by the global type.

### 2022-11-17 lmcs/priority-cp-long.tex
PCP and PLL

@@
 
+\subsection{PCP and PLL}
+In this subsection, we highlight the connection between PCP and linear logic.
+Dardha and Gay \cite{dardhagay18} present PCP--consequently also PCP given in this paper--in a way which can be viewed both as Classical Processes with restriction (T-Res) and parallel composition (T-Par) typing rules, and as a new version of linear logic, which they call Priority Linear Logic (PLL).
+PLL builds on CLL by replacing the cut rule with two logical rules: a mix and a cycle rule---here corresponding to T-Par and T-Res, respectively.
+Dardha and Gay \cite[\S 4]{dardhagay18} prove cycle-elimination, in the same lines as cut-elimination for CLL.
+As a corollary of cycle-elimination for PLL, we obtain deadlock freedom for PCP (Theorem 3 in \cite[\S 4]{dardhagay18}).
+In summary, PLL is an extension of CLL and the authors show the correspondence of PCP and PLL.
+Notice however, that PCP is not in correspondence with CLL itself, since processes in PCP are graphs, whether CLL induces trees.
+
 \subsection{Technical Developments}

### 2022-11-17 lmcs/conclusion.tex
another pass on related work

@@
 Padovani simplifies these further to a single natural, called a \emph{priority}~\cite{padovani14}, and adapts obligations/capabilities to session types~\cite{padovani13}.
-
-% We can obtain deadlock freedom for the $\pi$-calculus with session types via the encoding of session types using linear types~\cite{kobayashi07,dardhagiachino12,dardha14beat} and linear $\pi$-calculus with priorities.
+Later work by Kobayashi and co-authors \cite{GiachinoKL14,kobayashilaneve17} addresses deadlock detection for a value-passing CCS (predecessor of the $\pi$-calculus) where the number of nodes is arbitrary, namely modelling unbounded networks.
+The authors define an inference algorithm for their type system which guarantees deadlock freedom for these more complex kinds of communication networks.
 
@@
 Carbone and Debois~\cite{carbonedebois10} define progress for session typed $\pi$-calculus in terms of a \emph{catalyser} which provides the missing counterpart to a process.
-Carbone~\etal~\cite{carbonedardha14} use catalysers to show that progress is a compositional form of lock-freedom and can be lifted to session types via the encoding of session types to linear types~\cite{kobayashi07,dardhagiachino12,dardha14beat}.
+Carbone~\etal~\cite{carbonedardha14} use catalysers to show that progress is a compositional form of lock-freedom and can be lifted to session types via the encoding of session types to linear types~\cite{kobayashi07,dardhagiachino12,dardha14beat,dardhaetal17}.
+
 Vieira and Vasconcelos~\cite{vieiravasconcelos13} use single priorities and an abstract partial order to guarantee deadlock freedom in a binary session-typed $\pi$-calculus and building on conversation types.
+Coppo \etal \cite{coppoetal13,coppoetal16} present a deterministic, sound and complete, and compositional type inference algorithm for an \emph{interaction type system}, which is used to guarantee global progress for processes in a calculus based on asynchronous as well as dynamically interleaved and interfered multiparty sessions.
+The interaction type system allows to infer \emph{causalities}---much in the line of Kobayashi's priorities ---of channels, thus guaranteeing that session-typed processes do not get stuck at intermediate stages of their sessions.
 
@@
 Many years later, several correspondences between linear logic and the $\pi$-calculus with binary session types were proposed.
-Caires and Pfenning~\cite{cairespfenning10} propose a correspondence with dual intuitionistic linear logic, while Wadler~\cite{wadler14} proposes a correspondence with classical linear logic.
-Both guarantee deadlock freedom as a consequence of cut elimination.
-Dardha and Gay~\cite{dardhagay18extended} integrate Kobayashi and Padovani's work on priorities~\cite{kobayashi06,padovani14} with CP, loosening its ties to linear logic in exchange for expressivity.
-Dardha and P\'{e}rez~\cite{dardhaperez15extended} compare priorities \`a la Kobayashi with tree restrictions \`a la CP, and show that the latter is a subsystem of the former.
-Balzer~\etal~\cite{balzerpfenning17} introduce sharing at the cost of deadlock freedom, which they restore using an approach similar to priorities~\cite{balzertoninho19}.
+Caires and Pfenning~\cite{cairespfenning10} propose a correspondence with dual intuitionistic linear logic, while Wadler~\cite{wadler12} proposes a correspondence with classical linear logic.
+Both works guarantee deadlock freedom as a consequence of adopting cut elimination.
+Building on a previous work \cite{cairespfenning10}, Toninho \etal \cite{toninhoetal13} present a Curry-Howard correspondence between session types and linear logic for functional language via linear contextual monads, which are first-class values, thus giving rise to a higher-order session typed  language.
+In addition to the more standard results, the authors also prove a global progress theorem.
+Dardha and Gay~\cite{dardhagay18extended} define Priority CP by integrating Kobayashi and Padovani's work on priorities~\cite{kobayashi06,padovani14} with CP, loosening its ties to linear logic in exchange for expressivity.
+However, they show how PCP can be also \emph{viewed as} a an extension of linear logic, which they call Priority Linear Logic (PLL), and uses mix and cycle rules as opposed to the cut rule.
+Dardha and P\'{e}rez~\cite{dardhaperez15extended,dardhaperez15,DardhaP22} compare priorities \`a la Kobayashi with tree restrictions \`a la CP, and show that the latter is a subsystem of the former.
+In addition, they give a detailed account of comparing several type systems for deadlock freedom spanning across session types, linear logic, and linear types.
 Carbone~\etal~\cite{CarboneMSY15,carbonelindley16} give a logical view of MPST with a generalised duality.
@@
 Horne~\cite{Horne20} give a system for subtyping and multiparty compatibility where compatible processes are race free and deadlock free using a Curry-Howard correspondence, similar to the approach in~\cite{CiobanuH15}.
+Balzer~\etal~\cite{balzerpfenning17} introduce sharing at the cost of deadlock freedom, which they restore using \emph{worlds}, an approach similar to priorities~\cite{balzertoninho19}.
+Staying on sharing, Rocha and Caires \cite{rochacaires21} introduce an imperative feature, that of shared mutable states into a functional language based on Curry-Howard correspondence with linear logic.
+Their type system is thus able to capture programs which were not possible in previous works.
+The authors prove extensive technical results, including session fidelity, progress, confluence and
+normalisation.
+Lastly, Jacobs \etal \cite{JacobsBK22a} present a novel technique to guarantee deadlock freedom based on the notion of connectivity graph, which is an abstract representation of the topology of concurrent systems, and separation logic used to substructurally treat connectivity graph edges and labels.
 
@@
 
-% \subsubsection*{Acknowledgements}
-% The authors would like to thank Simon Fowler, April Gon\c{c}alves, and Philip Wadler for their comments on the manuscript.
+\subsubsection*{Acknowledgements}
+The authors would like to thank the anonymous reviewers for their detailed feedback, which helped produce a more complete and polished work.
+Also, the authors would like to thank Simon Fowler, April Gon\c{c}alves, and Philip Wadler for their comments on the manuscript.

### 2022-11-17 lmcs/priority-gv.tex
another pass on related work

@@
 \emph{Actions} are terms which perform communication actions and which synchronise between two threads.
-\emph{Ready terms} are terms which perform communication actions, either by themselves, \eg creating a new channel or thread, or with another thread, \eg sending or receiving.
-Progress for the term language is standard for GV, and deviates from progress for linear $\lambda$-calculus only in that terms may reduce to values or \emph{ready terms}:
 \begin{definition}[Actions]%
@@
 \end{definition}
+
+\emph{Ready terms} are terms which perform communication actions, either by themselves, \eg creating a new channel or thread, or with another thread, \eg sending or receiving.
+It is worth mentioning that the notion of readiness presented here is akin to \emph{live} processes  introduced by Caires and Pfenning \cite{cairespfenning10,DardhaP22}, and \emph{poised} processes introduced by Pfenning and Griffith \cite{PfenningG15} and later used by Balzer \etal \cite{balzerpfenning17,balzertoninho19}.
+Ready processes like live/poised processes denote processes that are ready to communicate on their providing channel.
 \begin{definition}[Ready Terms]%
@@
 \end{definition}
+
+\emph{Progress} for the term language is standard for GV, and deviates from progress for linear $\lambda$-calculus only in that terms may reduce to values or \emph{ready terms}.
+
 \begin{lemma}[Progress, $\tred$]%

### 2022-11-18 lmcs/conclusion.tex
more in related work
polishing related work

@@
-\section{Related Work and Discussion}
+\section{Related Work and Conclusion}
+\usingnamespace{pcp}
 
@@
 Deadlock freedom and progress are well studied properties in the $\pi$-calculus.
-For the `standard' typed $\pi$-calculus, an important line of work starts from Kobayashi's approach to deadlock freedom~\cite{kobayashi98}, where priorities are values from an abstract poset.
+For the ``standard'' typed $\pi$-calculus---types for channels used in input and output, an important line of work stems from Kobayashi's approach to deadlock freedom~\cite{kobayashi98}, where priorities are values from an abstract poset.
 Kobayashi~\cite{kobayashi06} simplifies the abstract poset to pairs of naturals, called \emph{obligations} and \emph{capabilities}.
 Padovani simplifies these further to a single natural, called a \emph{priority}~\cite{padovani14}, and adapts obligations/capabilities to session types~\cite{padovani13}.
-Later work by Kobayashi and co-authors \cite{GiachinoKL14,kobayashilaneve17} addresses deadlock detection for a value-passing CCS (predecessor of the $\pi$-calculus) where the number of nodes is arbitrary, namely modelling unbounded networks.
-The authors define an inference algorithm for their type system which guarantees deadlock freedom for these more complex kinds of communication networks.
+Later work by Kobayashi and co-authors \cite{GiachinoKL14,kobayashilaneve17} address deadlock detection for a value-passing CCS (a predecessor of the $\pi$-calculus) where the number of nodes in a network is arbitrary, namely modelling unbounded networks.
+The authors define a sound inference algorithm for their type system, which guarantees deadlock freedom for these more complex kinds of communication networks.
+This type system is more expressive than previous work by Kobayashi.
 
-For the session-typed $\pi$-calculus, Dezani~\etal~\cite{dezani-ciancaglinimostrous06} guarantee progress by allowing only one active session at a time.
-Dezani~\cite{dezani-ciancagliniliguoro09progress} introduces a partial order on channels, similar to Kobayashi~\cite{kobayashi98}.
-Carbone and Debois~\cite{carbonedebois10} define progress for session typed $\pi$-calculus in terms of a \emph{catalyser} which provides the missing counterpart to a process.
-Carbone~\etal~\cite{carbonedardha14} use catalysers to show that progress is a compositional form of lock-freedom and can be lifted to session types via the encoding of session types to linear types~\cite{kobayashi07,dardhagiachino12,dardha14beat,dardhaetal17}.
+For the session-typed $\pi$-calculus, an important line of work stems from Dezani and co-authors.
+In their work, Dezani~\etal~\cite{dezani-ciancaglinimostrous06} guarantee progress by allowing only one active session at a time.
+As sessions do not interleave, consequently they do not deadlock.
+Later on, Dezani \etal~\cite{dezani-ciancagliniliguoro09progress} introduce a partial order on channels, similar to Kobayashi~\cite{kobayashi98} and produce a type system for progress.
 
-Vieira and Vasconcelos~\cite{vieiravasconcelos13} use single priorities and an abstract partial order to guarantee deadlock freedom in a binary session-typed $\pi$-calculus and building on conversation types.
+Carbone and Debois~\cite{carbonedebois10} define progress for session-typed $\pi$-calculus in terms of a \emph{catalyser}, which provides the missing counterpart to a process.
+Intuitively,  either the process is deadlock-free in which case the catalyser is structurally congruent to the inaction process, or if the process is ``stuck'' for \eg, an input process $\tm{\recv{x}{y}{\halt}}$ (using CP syntax) then the catalyser simply provides the missing output required for communication.
+Carbone~\etal~\cite{carbonedardha14} use catalysers to show that progress is a compositional form of livelock freedom and can be lifted to session types via the encoding of session types to linear types~\cite{kobayashi07,dardhagiachino12,dardha14beat,dardhaetal17}.
+In this work, Carbone \etal \cite{carbonedardha14} sistematise and compare different notions of liveness properties: progress, deadlock freedom and livelock freedom.
+Their technique of using the encoding of session types combined with Kobayashi's type systems for deadlock and livelock freedom allows for a more flexible deadlock/livelock detection.
+It is worth noting that livelock freedom is a stronger property than deadlock freedom in the presence of recursion, as the former also discards useless divergent processes; for finite processes the two properties coincide.
+
+Vieira and Vasconcelos~\cite{vieiravasconcelos13} use single priorities and an abstract partial order to guarantee deadlock freedom in a binary session-typed $\pi$-calculus and building on conversation types, which is akin to session types.
+
+While our work focuses on \emph{binary} session types, it is worth discussing related work on Multiparty Session Types (MPST), which describe communication among multiple agents in a distributed setting.
+The line of work on MPST starts with Honda~\etal~\cite{hondayoshida08}, which guarantees deadlock freedom \emph{within a single session} by design, but the property does not hold for session interleaving.
+Bettini~\etal~\cite{bettinicoppo08} follow a technique similar to Kobayashi's for MPST.
+The main difference with our work is that we associate priorities with communication actions, where Bettini~\etal~\cite{bettinicoppo08} associate them with channels.
 Coppo \etal \cite{coppoetal13,coppoetal16} present a deterministic, sound and complete, and compositional type inference algorithm for an \emph{interaction type system}, which is used to guarantee global progress for processes in a calculus based on asynchronous as well as dynamically interleaved and interfered multiparty sessions.
 The interaction type system allows to infer \emph{causalities}---much in the line of Kobayashi's priorities ---of channels, thus guaranteeing that session-typed processes do not get stuck at intermediate stages of their sessions.
-
-While our work focuses on \emph{binary} session types, it is worth to discuss related work on Multiparty Session Types (MPST), which describe communication among multiple agents in a distributed setting.
-The line of work on MPST starts with Honda~\etal~\cite{hondayoshida08}, which guarantees deadlock freedom \emph{within a single session}, but not for several session interleaving.
-Bettini~\etal~\cite{bettinicoppo08} follow a technique similar to Kobayashi's for MPST.
-The main difference with our work is that we associate priorities with communication actions, where Bettini~\etal~\cite{bettinicoppo08} associate them with channels.
 Carbone and Montesi~\cite{carbonemontesi13} combine MPST with choreographic programming and obtain a formalism that satisfies deadlock freedom.
@@
 To do so, they leverage Labelled Transition Systems (LTSs) and communicating automata.
-Castellani~\etal~\cite{CastellaniDGH20} guarantee lock freedom, a stronger property than deadlock freedom, for MPST with \emph{internal delegation}, where participants in the same session are allowed to delegate tasks to each other, and internal delegation is captured by the global type.
+Castellani~\etal~\cite{CastellaniDGH20} guarantee livelock freedom, a stronger property than deadlock freedom, for MPST with \emph{internal delegation}, where participants in the same session are allowed to delegate tasks to each other, and internal delegation is captured by the global type.
 Scalas and Yoshida~\cite{scalasyoshida19} provide a revision of the foundations for MPST, and offer a less complicated and more general theory, by removing duality/consistency.
@@
 Toninho~\etal~\cite{toninhocaires12} present a translation of simply-typed $\lambda$-calculus into session-typed $\pi$-calculus, but their focus is not on deadlock freedom.
+Fowler \etal \cite{fowleretal21} present Hypersequent GV (HGV), which is a variation of GV that uses hyper-environments, much in the same line as Hypersequent CP, and enjoys deadlock freedom, confluence, and strong normalisation.
 
@@
 Both works guarantee deadlock freedom as a consequence of adopting cut elimination.
-Building on a previous work \cite{cairespfenning10}, Toninho \etal \cite{toninhoetal13} present a Curry-Howard correspondence between session types and linear logic for functional language via linear contextual monads, which are first-class values, thus giving rise to a higher-order session typed  language.
+Building on a previous work \cite{cairespfenning10}, Toninho \etal \cite{toninhoetal13} present a Curry-Howard correspondence between session types and linear logic for functional language via linear contextual monads, which are first-class values, thus giving rise to a higher-order session-typed  language.
 In addition to the more standard results, the authors also prove a global progress theorem.
@@
 
-\subsubsection*{Conclusion}
+\subsubsection*{Conclusion and Future Work}
 We answered our research question by presenting Priority GV, a~session-typed functional language which allows cyclic communication structures and uses priorities to ensure deadlock freedom.
@@
 
-\subsubsection*{Future work}
 Our formalism so far only captures the core of GV.

### 2022-12-05 lmcs/introduction.tex
*reflection -> operational correspondence

@@
 In the $\pi$-calculus literature, there have been several attempts at developing Curry-Howard correspondences between session-typed $\pi$-calculus and linear logic~\cite{girard87}: Caires and Pfenning's $\pi$DILL~\cite{cairespfenning10} corresponds to dual intuitionistic linear logic~\cite{barber96}, and Wadler's Classical Processes~\cite[CP]{wadler14} corresponds to classical linear logic~\cite[CLL]{girard87}.
-Both calculi guarantee deadlock freedom, which they achieve by restricting structure of processes and shared channels to \emph{trees}, by combining name restriction and parallel composition into a single construct, corresponding to the logical cut.
+Both calculi guarantee deadlock freedom, which they achieve by restricting the structure of processes and shared channels to \emph{trees}, by combining name restriction and parallel composition into a single construct, corresponding to the logical cut.
 This ensures that two processes can only communicate via exactly one series of channels, which rules out interleavings of sessions, and guarantees deadlock freedom.
@@
 
-CP and GV are related via a pair of translations which satisfy simulation~\cite{lindleymorris16}, and which can be tweaked to satisfy reflection.
+CP and GV are related via a pair of translations which satisfy simulation~\cite{lindleymorris16}, and which can be tweaked to satisfy operational correspondence.
 The two calculi share the same strong guarantees.

### 2022-12-05 lmcs/priority-gv.tex
Process comments by Reviewer 2

@@
 \end{enumerate}
-PGV is more expressive than GV: \cref{ex:cycle} is typeable and guaranteed deadlock-free in PGV, but is not typeable in GV~\cite{wadler14} and not guaranteed deadlock-free in GV's predecessor~\cite{gayvasconcelos10}.
+PGV is more expressive than GV: \cref{ex:cycle} is typeable and guaranteed to be deadlock-free in PGV, but is not typeable in GV~\cite{wadler14} and not guaranteed deadlock-free in GV's predecessor~\cite{gayvasconcelos10}.
 We believe PGV is a non-conservative \emph{extension} of GV, as CP can be embedded in a Kobayashi-style system~\cite{dardhaperez15extended}.
@@
     \begin{array}{l}
-      \letpair{x}{x'}{\new\unit}{}
+      \letpair{x}{x'}{\new\;\unit}{}
       \\
-      \letpair{y}{y'}{\new\unit}{}
+      \letpair{y}{y'}{\new\;\unit}{}
       \\
@@
 \subsubsection*{Typing Environments}
-Typing environments $\ty{\Gamma}$, $\ty{\Delta}$ associate types to names.
+Typing environments $\ty{\Gamma}$, $\ty{\Delta}$, $\ty{\Theta}$ associate types to names.
 Environments are linear, so two environments can only be combined as $\ty{\Gamma}, \ty{\Delta}$ if their names are distinct, \ie $\fv(\ty{\Gamma})\cap\fv(\ty{\Delta})=\varnothing$.
@@
 We extend the function $\pr(\cdot)$ to types and typing contexts by returning the smallest priority in the type or context, or $\cs{\top}$ if there is no priority.
-We use $\sqcap$ and $\sqcup$ to denote the minimum and maximum:
+We use $\sqcap$ and $\sqcup$ to denote the minimum and maximum, respectively:
 \[
@@
 
-The use of $\tm{\child}$ for child threads~\cite{lindleymorris15} overlaps with the use of the meta-variable $\cs{o}$ for priorities~\cite{dardhagay18extended}.
-Both are used to annotate sequents: flags appear on the sequent in configuration typing, and priorities in term typing.
-To distinguish the two symbols, they are typeset in a different font and a different colour.
+To distinguish between the use of $\tm{\child}$ to mark child threads~\cite{lindleymorris15} and the use of the meta-variable $\cs{o}$ for priorities~\cite{dardhagay18extended}, they are typeset in a different font and colour.
 
@@
 
-In PGV, we keep track of a lower and upper bound on the priorities of a term, \ie, while evaluating the term, when does it start communicating, and when does it finish.
+In PGV, we keep track of a lower and upper bound on the priorities of a term, \ie, while evaluating the term, when it starts communicating, and it finishes, respectively.
 The upper bound is written on the sequent, whereas the lower bound is approximated from the typing environment.
@@
       }{
-        \tmty{\send}{\tylolli[\ptop,\cs{o}]{\typrod{A}{\tysend[\cs{o}]{A}{S}}}{S}}
+        \tmty{\send}{\tylolli[\ptop,\cs{o}]{\typrod{T}{\tysend[\cs{o}]{T}{S}}}{S}}
       }
@@
         }{
-          \tseq[\pbot]{\tmty{x}{A}}{x}{A}
+          \tseq[\pbot]{\tmty{x}{T}}{x}{T}
         }
@@
           \tseq[\pbot]
-          {\tmty{x}{A},\tmty{y}{\tysend[\cs{o}]{A}{S}}}
-          {y}{\tysend[\cs{o}]{A}{S}}
+          {\tmty{x}{T},\tmty{y}{\tysend[\cs{o}]{T}{S}}}
+          {y}{\tysend[\cs{o}]{T}{S}}
         }
@@
         \tseq[\pbot]
-        {\tmty{x}{A},\tmty{y}{\tysend[\cs{o}]{A}{S}}}
-        {\pair{x}{y}}{\typrod{A}{\tysend[\cs{o}]{A}{S}}}
+        {\tmty{x}{T},\tmty{y}{\tysend[\cs{o}]{T}{S}}}
+        {\pair{x}{y}}{\typrod{T}{\tysend[\cs{o}]{T}{S}}}
       }
@@
       \tseq[\cs{o}]
-      {\tmty{x}{A},\tmty{y}{\tysend[\cs{o}]{A}{S}}}
+      {\tmty{x}{T},\tmty{y}{\tysend[\cs{o}]{T}{S}}}
       {\send\;\pair{x}{y}}{S}
@@
     \tseq[\pbot]
-    {\tmty{y}{\tysend[\cs{o}]{A}{S}}}
-    {\lambda{x}.\send\;\pair{x}{y}}{\tylolli[\cs{o},\cs{o}]{A}{S}}
+    {\tmty{y}{\tysend[\cs{o}]{T}{S}}}
+    {\lambda{x}.\send\;\pair{x}{y}}{\tylolli[\cs{o},\cs{o}]{T}{S}}
   }
@@
 
-Canonical forms deviate from those for GV, in that we opt to move all $\nu$-binders to the top.
+\emph{Canonical forms} deviate from those for GV, in that we opt to move all $\nu$-binders to the top.
 The standard GV canonical form, alternating $\nu$-binders and their corresponding parallel compositions, does not work for PGV, since multiple channels may be split across a single parallel composition.

### 2022-12-05 lmcs/priority-cp-long.tex
Process comments by Reviewer 2
Restore the translation from PCP to PGV

@@
   \quad
-  \minpr(\ty{\Gamma},\tmty{x}{A}) = \minpr(\ty{\Gamma})\sqcap\minpr(\ty{A})
+  \minpr(\ty{\Gamma},\tmty{x}{T}) = \minpr(\ty{\Gamma})\sqcap\minpr(\ty{T})
 \]
@@
 \input{prf/thm-pcp-closed-progress}
-
 \endgroup
@@
 \begingroup
+We illustrate the relation between PCP and PGV by defining a translation from PCP to PGV. The translation on types is defined as follows:
+\[
+  \begin{array}{lcl}
+    \ty{\cpgvT{\pcp{\tytens[\cs{o}]{A}{B}}}}
+    & = & \pgv{\ty{\tysend[\cs{o}]{\co{\cpgvT{A}}}{\cpgvT{B}}}}
+    \\
+    \ty{\cpgvT{\pcp{\typlus[\cs{o}]{A}{B}}}}
+    & = & \pgv{\ty{\tyselect[\cs{o}]{\cpgvT{A}}{\cpgvT{B}}}}
+  \end{array}
+  \hfill\quad\hfill%
+  \begin{array}{lcl}
+    \ty{\cpgvT{\pcp{\tyone[\cs{o}]}}}
+    & = & \ty{\pgv{\tyends[\cs{o}]}}
+    \\
+    \ty{\cpgvT{\pcp{\tynil[\cs{o}]}}}
+    & = & \pgv{\ty{\tyselectemp[\cs{o}]}}
+  \end{array}
+\]%
+\[%
+  \begin{array}{lcl}
+    \ty{\cpgvT{\pcp{\typarr[\cs{o}]{A}{B}}}}
+    & = & \pgv{\ty{\tyrecv[\cs{o}]{\cpgvT{A}}{\cpgvT{B}}}}
+    \\
+    \ty{\cpgvT{\pcp{\tywith[\cs{o}]{A}{B}}}}
+    & = & \pgv{\ty{\tyoffer[\cs{o}]{\cpgvT{A}}{\cpgvT{B}}}}
+  \end{array}
+  \hfill\quad\hfill%
+  \begin{array}{lcl}
+    \ty{\cpgvT{\pcp{\tybot[\cs{o}]}}}
+    & = & \ty{\pgv{\tyendr[\cs{o}]}}
+    \\
+    \ty{\cpgvT{\pcp{\tytop[\cs{o}]}}}
+    & = & \pgv{\ty{\tyofferemp[\cs{o}]}}
+  \end{array}
+\]
+
+There are two separate translations on processes. The main translation, $\tm{\cpgvM{\cdot}}$, translates processes to \emph{terms}:
+\begin{align*}
+  &\pcp{\tm{\cpgvM{\link{x}{y}}}}
+  &&= \pgv{\tm{\link\;{\pair{x}{y}}}} \\
+  &\pcp{\tm{\cpgvM{\res{x}{y}{P}}}}
+  &&= \pgv{\tm{\letpair{x}{y}{\new}{\cpgvM{P}}}} \\
+  &\pcp{\tm{\cpgvM{\ppar{P}{Q}}}}
+  &&= \pgv{\tm{\andthen{\spawn\;{(\lambda\unit.\cpgvM{P})}}{\cpgvM{Q}}}} \\
+  &\pcp{\tm{\cpgvM{\halt}}}
+  &&= \pgv{\tm{\unit}} \\
+  &\pcp{\tm{\cpgvM{\close{x}{P}}}}
+  &&= \pgv{\tm{\andthen{\close\;{x}}{\cpgvM{P}}}} \\
+  &\pcp{\tm{\cpgvM{\wait{x}{P}}}}
+  &&= \pgv{\tm{\andthen{\wait\;{x}}{\cpgvM{P}}}} \\
+  &\pcp{\tm{\cpgvM{\send{x}{y}{P}}}}
+  &&= \pgv{\tm{\letpair{y}{z}{\new}{\letbind{x}{\send\;{\pair{z}{x}}}{\cpgvM{P}}}}} \\
+  &\pcp{\tm{\cpgvM{\recv{x}{y}{P}}}}
+  &&= \pgv{\tm{\letpair{y}{x}{\recv\;{x}}{\cpgvM{P}}}} \\
+  &\pcp{\tm{\cpgvM{\inl{x}{P}}}}
+  &&= \pgv{\tm{\letbind{x}{\select{\labinl}\;{x}}{\cpgvM{P}}}} \\
+  &\pcp{\tm{\cpgvM{\inr{x}{P}}}}
+  &&= \pgv{\tm{\letbind{x}{\select{\labinr}\;{x}}{\cpgvM{P}}}} \\
+  &\pcp{\tm{\cpgvM{\offer{x}{P}{Q}}}}
+  &&= \pgv{\tm{\offer{x}{x}{\cpgvM{P}}{x}{\cpgvM{Q}}}} \\
+  &\pcp{\tm{\cpgvM{\absurd{x}}}}
+  &&= \pgv{\tm{\offeremp{x}}}
+\end{align*}
+
+Unfortunately, the operational correspondence along $\tm{\cpgvM{\cdot}}$ is unsound, as it translates $\nu$-binders and parallel compositions to $\pgv{\tm{\new}}$ and $\pgv{\tm{\spawn}}$, which can reduce to their equivalent configuration constructs using \LabTirName{E-New} and \LabTirName{E-Spawn}.
+The same goes for $\nu$-binders which are inserted when translating bound send to unbound send.
+For instance, the process $\pcp{\tm{\send{x}{y}{P}}}$ is blocked, but its translation uses $\pgv{\tm{\new}}$ and can reduce.
+To address this issue, we use a second translation, $\tm{\cpgvC{\cdot}}$, which is equivalent to translating with $\tm{\cpgvM{\cdot}}$ then reducing with \LabTirName{E-New} and \LabTirName{E-Spawn}:
+\begin{align*}
+  &\pcp{\tm{\cpgvC{\res{x}{y}{P}}}}
+  &&= \pgv{\tm{\res{x}{y}{\cpgvC{P}}}}
+  \\
+  &\pcp{\tm{\cpgvC{\ppar{P}{Q}}}}
+  &&= \pgv{\tm{\ppar{\cpgvC{P}}{\cpgvC{Q}}}}
+  \\
+  &\pcp{\tm{\cpgvC{\send{x}{y}{P}}}}
+  &&= \pgv{\tm{\res{y}{z}{(\child\;\letbind{x}{\send\;\pair{z}{x}}{\cpgvM{P}})}}}
+  \\
+  &\pcp{\tm{\cpgvC{\inl{x}{P}}}}
+  &&= \pgv{\tm{\res{y}{z}{(\child\;\letbind{x}{\andthen{\close\;(\send\;\pair{\inl{y}}{x})}{z}}{\cpgvM{P}})}}}
+  \\
+  &\pcp{\tm{\cpgvC{\inr{x}{P}}}}
+  &&= \pgv{\tm{\res{y}{z}{(\child\;\letbind{x}{\andthen{\close\;(\send\;\pair{\inr{y}}{x})}{z}}{\cpgvM{P}})}}}
+  \\
+  &\pcp{\tm{\cpgvC{P}}}
+  &&= \pgv{\tm{\child{\cpgvM{P}}}},\quad\text{if none of the above apply}
+\end{align*}
+Typing environments are translated pointwise, and sequents $\pcp{\seq{P}{\ty{\Gamma}}}$ are translated as $\pgv{\cseq[\child]{\ty{\cpgvT{\ty{\Gamma}}}}{\cpgvC{P}}}$, where $\tm{\pgv{\child}}$ indicates a child thread, since translated processes do not have a main thread.
+The translations $\tm{\cpgvM{\cdot}}$ and $\tm{\cpgvC{\cdot}}$ preserve typing, and the latter induces a sound and complete operational correspondence.
+
 %\restatelemma{lempcptopgvtermspreservation}

### 2022-12-05 lmcs/priority-gv.tex
Further changes

@@
 The type of \emph{pure functions} $\ty{\tylolli{T}{U}}$, \ie those which perform no communications, is syntactic sugar for $\ty{\tylolli[\cs{\ptop},\cs{\pbot}]{T}{U}}$.
+The lower bound for a pure function is $\cs{\ptop}$ as pure functions never start communicating.
+For similar reasons, the upper bound for a pure function is $\cs{\pbot}$.
 
@@
 Constants $\tm{K}$ are typed using type schemas, and embedded using \LabTirName{T-Const} (mid of \cref{fig:pgv-typing}).
-The typing rules treat \emph{all variables} as linear resources, even those of non-linear types such as $\ty{\tyunit}$.
-However, the rules can easily be extended to allow values with unrestricted usage~\cite{wadler14}.
+The typing rules treat \emph{all variables} as linear resources, even those of non-linear types such as $\ty{\tyunit}$, though they can easily be extended to allow values with unrestricted usage~\cite{wadler14}.
 
@@
 
-In PGV, we keep track of a lower and upper bound on the priorities of a term, \ie, while evaluating the term, when it starts communicating, and it finishes, respectively.
-The upper bound is written on the sequent, whereas the lower bound is approximated from the typing environment.
+In PGV, we keep track of a lower and upper bound on the priorities of a term, \ie, while evaluating the term, when it starts communicating, and when it finishes, respectively.
+The upper bound is written on the sequent and the lower bound is approximated from the typing environment, \eg, for $\tseq[\cs{p}]{\ty{\Gamma}}{M}{T}$ the upper bound is $\cs{p}$ and the lower bound is at least $\minpr(\ty{\Gamma})$.
+The latter is \emph{correct} because a term cannot communicate at a priority earlier than the earliest priority amongst the channels it has access to.
+It is an \emph{approximation} on function terms, as these can ``skip'' communication by returning the corresponding channel unused.
+However, linearity prevents such functions from being used in well-typed configurations: once the unused channel's priority has passed, it can no longer be used.
+
 Typing rules for sequential constructs enforce sequentiality, \eg, the typing for $\tm{\andthen{M}{N}}$ has a side condition which requires that the upper bound of $\tm{M}$ is smaller than the lower bound of $\tm{N}$, \ie, $\tm{M}$ finishes before $\tm{N}$ starts.
@@
 Crucially, it composes the terms below in parallel.
-While each of these terms itself is well typed, they impose opposite conditions on the priorities, so their composition is ill-typed.
+While each of these terms itself is well typed, they impose opposite conditions on the priorities, so connecting $\ty{x}$ to $\ty{x'}$ and $\ty{y}$ to $\ty{y'}$ using \LabTirName{T-Res} is ill-typed, as there is no assignment to $\cs{o}$ and $\cs{o'}$ that can satsify both $\cs{o}<\cs{o'}$ and $\cs{o'}<\cs{o}$.
 (We omit the priorities on $\ty{\tyends}$ and $\ty{\tyendr}$.)
@@
 
+For simplicity, we assume priority annotations are not inferred, but provided as an input to type checking.
+However, for any term, priorities can be inferred, \eg, by using the topological ordering of the directed graph where the vertices are the priority meta-variables and the edges are the inequality constraints between the priority meta-variables in the typing derivation.
+
 \subsubsection*{Configurations Typing}
@@
     \tm{\main}
-    & \tm{\letpair{x}{x'}{\new}{}}
+    & \tm{\letpair{x}{x'}{\new\;\unit}{}}
     \\

### 2022-12-07 lmcs/prf/thm-pcp-closed-progress.tex
Process comments by Reviewer 2
Fix highlighting in Thm 4.4

--- 2021-11-17 coordination21/prf/thm-pcp-closed-progress.tex
@@
 Among them, we pick the process with the smallest priority $\minpr{(\ty{\Gamma_i})}$ for all $i$.
-Let this process be let this be $\tm{P_i}$ and the priority of the top prefix be $\cs o$.
+Let this process be $\tm{P_i}$ and the priority of the top prefix be $\cs o$.
 $\tm{P_i}$ acts on some endpoint $\tmty{y}{A}\in\ty{\Gamma_i}$.
-We must have $\minpr{(\ty{\Gamma_i})}=\pr{(\ty{A})} = \cs o$, since the other actions in $\tm{P_i}$ are guarded by the action on $\tmty{y}{A}$, thus satisfying law (i) of priorities.
+We must have $\minpr{(\ty{\Gamma_i})}=\pr{(\ty{A})} = \cs{o}$, since the other actions in $\tm{P_i}$ are guarded by the action on $\tmty{y}{A}$, thus satisfying law (i) of priorities.
 
@@
 
-  Otherwise, $\tm{P_i}$ is an input/branching or output/selection action on endpoint $\tm y$ of type $\ty A$ with priority $\cs o$.
-Since process $\tm P$ is closed and consequently it respects law (ii) of priorities, there must be a co-action $y'$ of type $\ty{\co{A}}$  where $\tm{y}$ and $\tm{y'}$ are dual endpoints of the same channel (by application of rule \textsc{T-Res}).
-By duality, $\pr{(\ty{A})}=\pr{(\ty{\co{A}})}= \cs o$.
-In the following we show that: $y'$ is the subject of a top level action of a process $P_j$ with $i\neq j$.
+  Otherwise, $\tm{P_i}$ is an input/branching or output/selection action on endpoint $\tm{y}$ of type $\ty{A}$ with priority $\cs{o}$.
+Since process $\tm{P}$ is closed and consequently it respects law (ii) of priorities, there must be a co-action $\tm{y'}$ of type $\ty{\co{A}}$  where $\tm{y}$ and $\tm{y'}$ are dual endpoints of the same channel (by application of rule \textsc{T-Res}).
+By duality, $\pr{(\ty{A})}=\pr{(\ty{\co{A}})}= \cs{o}$.
+In the following we show that: $\tm{y'}$ is the subject of a top level action of a process $\tm{P_j}$ with $i\neq j$.
 This allows for the communication among $\tm{P_i}$ and $\tm{P_j}$ to happen immediately over channel endpoints $\tm{y}$ and $\tm{y'}$.
 
-  Suppose that $\tm{y'}$ is an action not in a different parallel process $P_j$ but rather of $P_i$ itself.
+  Suppose that $\tm{y'}$ is an action not in a different parallel process $\tm{P_j}$ but rather of $\tm{P_i}$ itself.
 That means that the action on $\tm{y'}$ must be prefixed by the action on $\tm{y}$, which is top level in $\tm{P_i}$.
-To respect law (i) of priorities we must have $\cs o < \cs o$, which is absurd.
+To respect law (i) of priorities we must have $\cs{o}<\cs{o}$, which is absurd.
 This means that $\tm{y'}$ is in another parallel process $\tm{P_j}$ for $i\neq j$.
@@
   Suppose that $\tm{y'}$ in $\tm{P_j}$ is not at top level.
-In order to respect law (i) of priorities, it means that $\tm{y'}$ is prefixed by actions that are smaller than its priority $\cs o$.
-This leads to a contradiction because stated that $\cs o$ is the smallest priority.
-Hence, $y'$ must be the subject of a top level action.
+In order to respect law (i) of priorities, it means that $\tm{y'}$ is prefixed by actions that are smaller than its priority $\cs{o}$.
+This leads to a contradiction because stated that $\cs{o}$ is the smallest priority.
+Hence, $\tm{y'}$ must be the subject of a top level action.
 

### 2022-12-07 lmcs/prf/thm-pcp-closed-progress.tex
polishing text

--- 2021-11-17 coordination21/prf/thm-pcp-closed-progress.tex
@@
 
-  Our proof follows the same reasoning by Kobayashi~\cite{kobayashi06} used in the proof of deadlock freedom for closed processes (Theorem 2).
-
   Consider processes $\tm {P_1 \parallel\dots\parallel P_m}$.
 Among them, we pick the process with the smallest priority $\minpr{(\ty{\Gamma_i})}$ for all $i$.
-Let this process be let this be $\tm{P_i}$ and the priority of the top prefix be $\cs o$.
+Let this process be $\tm{P_i}$ and the priority of the top prefix be $\cs o$.
 $\tm{P_i}$ acts on some endpoint $\tmty{y}{A}\in\ty{\Gamma_i}$.

### 2022-12-07 lmcs/priority-cp-long.tex
Further changes
polishing text

@@
 
-Finally, since our reduction relation is a strict subset of the reduction relation in the original~\cite{dardhagay18extended}, we defer to their proofs.
+Finally, since our reduction relation is a strict subset of the reduction relation in the original~\cite{dardhagay18extended}, we defer to their proof of subject reduction (Theorem 2 in~\cite{dardhagay18extended}).
 We prove progress for our version of PCP, see~\cref{prf:thm-pcp-closed-progress}.
@@
 \end{proof}
+
+The proof for progress (below) follows the same reasoning by Kobayashi~\cite{kobayashi06} used in the proof of deadlock freedom for closed processes (Theorem 2).
 
@@
   &\pcp{\tm{\cpgvM{\res{x}{y}{P}}}}
-  &&= \pgv{\tm{\letpair{x}{y}{\new}{\cpgvM{P}}}} \\
+  &&= \pgv{\tm{\letpair{x}{y}{\new\;\unit}{\cpgvM{P}}}} \\
   &\pcp{\tm{\cpgvM{\ppar{P}{Q}}}}
@@
   &\pcp{\tm{\cpgvM{\send{x}{y}{P}}}}
-  &&= \pgv{\tm{\letpair{y}{z}{\new}{\letbind{x}{\send\;{\pair{z}{x}}}{\cpgvM{P}}}}} \\
+  &&= \pgv{\tm{\letpair{y}{z}{\new\;\unit}{\letbind{x}{\send\;{\pair{z}{x}}}{\cpgvM{P}}}}} \\
   &\pcp{\tm{\cpgvM{\recv{x}{y}{P}}}}

### 2022-12-07 lmcs/prf/thm-pcp-closed-progress.tex

--- 2022-12-07 lmcs/prf/thm-pcp-closed-progress.tex
@@
 
-  Our proof follows the same reasoning by Kobayashi~\cite{kobayashi06} used in the proof of deadlock freedom for closed processes (Theorem 2).
-
   Consider processes $\tm {P_1 \parallel\dots\parallel P_m}$.
@@
 $\tm{P_i}$ acts on some endpoint $\tmty{y}{A}\in\ty{\Gamma_i}$.
-We must have $\minpr{(\ty{\Gamma_i})}=\pr{(\ty{A})} = \cs{o}$, since the other actions in $\tm{P_i}$ are guarded by the action on $\tmty{y}{A}$, thus satisfying law (i) of priorities.
+We must have $\minpr{(\ty{\Gamma_i})}=\pr{(\ty{A})} = \cs o$, since the other actions in $\tm{P_i}$ are guarded by the action on $\tmty{y}{A}$, thus satisfying law (i) of priorities.
 

### 2022-12-09 lmcs/introduction.tex
Process comments by Reviewer 2
Clarify "series of channels"

@@
 In the $\pi$-calculus literature, there have been several attempts at developing Curry-Howard correspondences between session-typed $\pi$-calculus and linear logic~\cite{girard87}: Caires and Pfenning's $\pi$DILL~\cite{cairespfenning10} corresponds to dual intuitionistic linear logic~\cite{barber96}, and Wadler's Classical Processes~\cite[CP]{wadler14} corresponds to classical linear logic~\cite[CLL]{girard87}.
-Both calculi guarantee deadlock freedom, which they achieve by restricting the structure of processes and shared channels to \emph{trees}, by combining name restriction and parallel composition into a single construct, corresponding to the logical cut.
-This ensures that two processes can only communicate via exactly one series of channels, which rules out interleavings of sessions, and guarantees deadlock freedom.
+Both calculi guarantee deadlock freedom, which they achieve by the combination of binary session types with a restriction on the way processes can be connected by channels: two processes can share at most one channel, and, more generally, information transmitted between any two processes must pass through one \emph{unique} series of channels and intermediate processes.
+This property follows from a syntactic restriction: the combination of name restriction and parallel composition into a single construct, corresponding to the logical cut.
+
 There are many downsides to combining name restriction and parallel composition, such as lack of modularity, difficulty typing structural congruence and formulating label-transition semantics, which have led to various approaches to decoupling these constructs.
@@
 Our aim is to develop a more expressive version of GV while maintaining deadlock freedom.
-While process calculi have their advantages, \eg, their succinctness, we chose to work with GV for several reasons.
+While process calculi have their advantages, \eg, their succinctness compared to concurrent $\lambda$-calculi, we chose to work with GV for several reasons.
 In general, concurrent $\lambda$-calculi support higher-order functions, and have a capability for abstraction not usually present in process calculi.

### 2022-12-09 lmcs/introduction.tex
Further elaboration

@@
 In the $\pi$-calculus literature, there have been several attempts at developing Curry-Howard correspondences between session-typed $\pi$-calculus and linear logic~\cite{girard87}: Caires and Pfenning's $\pi$DILL~\cite{cairespfenning10} corresponds to dual intuitionistic linear logic~\cite{barber96}, and Wadler's Classical Processes~\cite[CP]{wadler14} corresponds to classical linear logic~\cite[CLL]{girard87}.
-Both calculi guarantee deadlock freedom, which they achieve by the combination of binary session types with a restriction on the way processes can be connected by channels: two processes can share at most one channel, and, more generally, information transmitted between any two processes must pass through one \emph{unique} series of channels and intermediate processes.
-This property follows from a syntactic restriction: the combination of name restriction and parallel composition into a single construct, corresponding to the logical cut.
+Both calculi guarantee deadlock freedom, which they achieve by the combination of binary session types with a restriction on the way processes can be connected by channels: two processes can share at most one channel, and, more generally, information transmitted between any two processes must pass through one unique series of channels and intermediate processes.
+We refer to such processes as \emph{tree-structured}, because the \emph{communication graph}---where the vertices are ready processes and two vertices are connected by an edge if and only if the corresponding processes share a channel---is a tree.
+For $\pi$DILL, CP, and GV, tree-structure follows from a syntactic restriction: the combination of name restriction and parallel composition into a single syntactic construct, corresponding to the logical cut.
 
-There are many downsides to combining name restriction and parallel composition, such as lack of modularity, difficulty typing structural congruence and formulating label-transition semantics, which have led to various approaches to decoupling these constructs.
-Hypersequent CP~\cite{MP18,kokkemontesi19popl,kokkemontesi19tlla} and Linear Compositional Choreographies~\cite{CarboneMS18} decouple them, but maintain the correspondence to linear logic and allow only tree-structured processes.
+There are many downsides to combining name restriction and parallel composition, such as lack of modularity, difficulty typing structural congruence and formulating label-transition semantics.
+GV, specifically, struggles with a complicated metatheory due to the mismatch between its term language---where restriction and parallel composition are combined---and its configuration language---where they are not.
+There have been various approaches to decoupling restriction and parallel composition.
+Hypersequent CP~\cite{MP18,kokkemontesi19popl,kokkemontesi19tlla}, Hypersequent GV~\cite{fowleretal21}, and Linear Compositional Choreographies~\cite{CarboneMS18} decouple them, but maintain the correspondence to linear logic and allow only tree-structured processes.
 Priority CP~\cite[PCP]{dardhagay18extended} weakens the correspondence to CLL in exchange for a more expressive language which allows cyclic-structured processes.

### 2022-12-12 lmcs/conclusion.tex
removing some emphs

@@
 
-While our work focuses on \emph{binary} session types, it is worth discussing related work on Multiparty Session Types (MPST), which describe communication among multiple agents in a distributed setting.
-The line of work on MPST starts with Honda~\etal~\cite{hondayoshida08}, which guarantees deadlock freedom \emph{within a single session} by design, but the property does not hold for session interleaving.
+While our work focuses on {binary} session types, it is worth discussing related work on Multiparty Session Types (MPST), which describe communication among multiple agents in a distributed setting.
+The line of work on MPST starts with Honda~\etal~\cite{hondayoshida08}, which guarantees deadlock freedom {within a single session} by design, but the property does not hold for session interleaving.
 Bettini~\etal~\cite{bettinicoppo08} follow a technique similar to Kobayashi's for MPST.
@@
 Deni\'{e}lou and Yoshida~\cite{DenielouY13} introduce \emph{multiparty compatibility}, which generalises the notion of duality in binary session types.
-They synthesise safe and deadlock-free \emph{global types}--specifying communication among \emph{all} involved participants, from \emph{local session types}--specifying communication from the viewpoint of \emph{one} participant.
+They synthesise safe and deadlock-free \emph{global types}--specifying communication among {all} involved participants, from \emph{local session types}--specifying communication from the viewpoint of {one} participant.
 To do so, they leverage Labelled Transition Systems (LTSs) and communicating automata.
-Castellani~\etal~\cite{CastellaniDGH20} guarantee livelock freedom, a stronger property than deadlock freedom, for MPST with \emph{internal delegation}, where participants in the same session are allowed to delegate tasks to each other, and internal delegation is captured by the global type.
+Castellani~\etal~\cite{CastellaniDGH20} guarantee livelock freedom, a stronger property than deadlock freedom, for MPST with {internal delegation}, where participants in the same session are allowed to delegate tasks to each other, and internal delegation is captured by the global type.
 Scalas and Yoshida~\cite{scalasyoshida19} provide a revision of the foundations for MPST, and offer a less complicated and more general theory, by removing duality/consistency.
@@
 Dardha and Gay~\cite{dardhagay18extended} define Priority CP by integrating Kobayashi and Padovani's work on priorities~\cite{kobayashi06,padovani14} with CP, loosening its ties to linear logic in exchange for expressivity.
-However, they show how PCP can be also \emph{viewed as} a an extension of linear logic, which they call Priority Linear Logic (PLL), and uses mix and cycle rules as opposed to the cut rule.
+However, they show how PCP can be also {viewed as} a an extension of linear logic, which they call Priority Linear Logic (PLL), and uses mix and cycle rules as opposed to the cut rule.
 Dardha and P\'{e}rez~\cite{dardhaperez15extended,dardhaperez15,DardhaP22} compare priorities \`a la Kobayashi with tree restrictions \`a la CP, and show that the latter is a subsystem of the former.

### 2022-12-12 lmcs/introduction.tex
removing some emphs

@@
 A third crucial property is \emph{deadlock freedom}, which ensures that processes do not have cyclic dependencies---\eg, when two processes wait for each other to send a value.
-Binary session types \emph{alone} are insufficient to rule out deadlocks arising from interleaved sessions, but several additional techniques have been developed to guarantee deadlock freedom in session-typed $\pi$-calculus and concurrent $\lambda$-calculus.
+Binary session types {alone} are insufficient to rule out deadlocks arising from interleaved sessions, but several additional techniques have been developed to guarantee deadlock freedom in session-typed $\pi$-calculus and concurrent $\lambda$-calculus.
 

### 2022-12-12 lmcs/priority-cp-long.tex
removing some emphs

@@
 
-Types $\pcp{\ty{\tynil[\cs{o}]}}$ and $\pcp{\ty{\tytop[\cs{o}]}}$ type the endpoints of a channel over which we can send or receive a choice between \emph{no options}.
+Types $\pcp{\ty{\tynil[\cs{o}]}}$ and $\pcp{\ty{\tytop[\cs{o}]}}$ type the endpoints of a channel over which we can send or receive a choice between {no options}.
 These two types act as units for $\pcp{\ty{\typlus[\cs{o}]{A}{B}}}$ and $\pcp{\ty{\tywith[\cs{o}]{A}{B}}}$, respectively.
@@
 
-The main change we make to PCP is \emph{removing commuting conversions}.
-Commuting conversions are necessary if we want our reduction strategy to correspond \emph{exactly} to cut (or cycle in \cite{dardhagay18extended}) elimination.
+The main change we make to PCP is {removing commuting conversions}.
+Commuting conversions are necessary if we want our reduction strategy to correspond {exactly} to cut (or cycle in \cite{dardhagay18extended}) elimination.
 However, as Lindley and Morris~\cite{lindleymorris15} show, all communications that can be performed \emph{with} the use of commuting conversions, can also be performed \emph{without} them, but using structural congruence.

### 2022-12-12 lmcs/priority-gv.tex
polishing text
removing some emphs

@@
 We present Priority GV (PGV), a session-typed functional language based on GV~\cite{wadler15,lindleymorris15} which uses priorities \`{a} la Kobayashi and Padovani~\cite{kobayashi06,padovaninovara15} to enforce deadlock freedom.
-Priority GV offers a more fine-grained analysis of communication structures, and by separating channel creation form thread spawning it allows cyclic structures.
+Priority GV is more flexible than GV because it allows processes to share more than one communication channel.
 
@@
 Given a function with type $\ty{\tylolli[\cs{p},\cs{q}]{T}{U}}$, $\cs{p}$ is a \emph{lower bound} on the priorities of the endpoints captured by the body of the function, and $\cs{q}$ is an \emph{upper bound} on the priority of the communications that take place as a result of applying the function.
-The type of \emph{pure functions} $\ty{\tylolli{T}{U}}$, \ie those which perform no communications, is syntactic sugar for $\ty{\tylolli[\cs{\ptop},\cs{\pbot}]{T}{U}}$.
+The type of {pure functions} $\ty{\tylolli{T}{U}}$, \ie those which perform no communications, is syntactic sugar for $\ty{\tylolli[\cs{\ptop},\cs{\pbot}]{T}{U}}$.
 The lower bound for a pure function is $\cs{\ptop}$ as pure functions never start communicating.
@@
 Duality plays a crucial role in session types.
-The two endpoints of a channel are assigned dual types, ensuring that, for instance, whenever one program \emph{sends} a value on a channel, the program on the other end is waiting to \emph{receive}.
+The two endpoints of a channel are assigned dual types, ensuring that, for instance, whenever one program {sends} a value on a channel, the program on the other end is waiting to {receive}.
 Each session type $\ty{S}$ has a dual, written $\ty{\co{S}}$.
-Duality is an involutive function which \emph{preserves priorities}:
+Duality is an involutive function which {preserves priorities}:
 \[
@@
 
-Term reduction is the standard call-by-value, left-to-right evaluation for GV, and only deviates from reduction for the linear $\lambda$-calculus in that it reduces terms to values \emph{or} ready terms waiting to perform a communication action.
+Term reduction is the standard call-by-value, left-to-right evaluation for GV, and only deviates from reduction for the linear $\lambda$-calculus in that it reduces terms to values {or} ready terms waiting to perform a communication action.
 
@@
 Intuitively, evaluating $\tm{\link}$ causes a substitution: if we have a channel bound as $\tm{\res{x}{y}{}}$, then $\tm{\link\;\pair{w}{x}}$ replaces all occurrences of $\tm{y}$ by $\tm{w}$.
-However, in GV, $\tm{\link}$ is required to return a terminated channel, which means that the semantics for $\tm{link}$ must \emph{create} a fresh channel of type $\ty{\tyends}/\ty{\tyendr}$.
-The endpoint of type $\ty{\tyends}$ is returned by the $\tm{link}$ construct, and a $\tm{\wait}$ on the other endpoint guards the \emph{actual} substitution.
+However, in GV, $\tm{\link}$ is required to return a terminated channel, which means that the semantics for $\tm{link}$ must {create} a fresh channel of type $\ty{\tyends}/\ty{\tyendr}$.
+The endpoint of type $\ty{\tyends}$ is returned by the $\tm{link}$ construct, and a $\tm{\wait}$ on the other endpoint guards the {actual} substitution.
 In PGV, evaluating $\tm{\link}$ simply causes a substitution.
@@
 Constants $\tm{K}$ are typed using type schemas, and embedded using \LabTirName{T-Const} (mid of \cref{fig:pgv-typing}).
-The typing rules treat \emph{all variables} as linear resources, even those of non-linear types such as $\ty{\tyunit}$, though they can easily be extended to allow values with unrestricted usage~\cite{wadler14}.
+The typing rules treat {all variables} as linear resources, even those of non-linear types such as $\ty{\tyunit}$, though they can easily be extended to allow values with unrestricted usage~\cite{wadler14}.
 
 The only non-standard feature of the typing rules is the priority annotations.
-Priorities are based on \emph{obligations/capabilities} used by Kobayashi~\cite{kobayashi06}, and simplified to single priorities following Padovani~\cite{padovani14}.
+Priorities are based on {obligations/capabilities} used by Kobayashi~\cite{kobayashi06}, and simplified to single priorities following Padovani~\cite{padovani14}.
 The integration of priorities into GV is adapted from Padovani and Novara~\cite{padovaninovara15}.
@@
 The upper bound is written on the sequent and the lower bound is approximated from the typing environment, \eg, for $\tseq[\cs{p}]{\ty{\Gamma}}{M}{T}$ the upper bound is $\cs{p}$ and the lower bound is at least $\minpr(\ty{\Gamma})$.
-The latter is \emph{correct} because a term cannot communicate at a priority earlier than the earliest priority amongst the channels it has access to.
+The latter is {correct} because a term cannot communicate at a priority earlier than the earliest priority amongst the channels it has access to.
 It is an \emph{approximation} on function terms, as these can ``skip'' communication by returning the corresponding channel unused.
@@
 Typing rules for configurations are at the bottom of \cref{fig:pgv-typing}.
-Configurations are typed by a judgement $\cseq[\phi]{\ty{\Gamma}}{\conf{C}}$ stating that ``a configuration $\tm{\conf{C}}$ with flag $\phi$ is well typed under typing environment $\ty{\Gamma}$''.
+Configurations are typed by a judgement $\cseq[\phi]{\ty{\Gamma}}{\conf{C}}$ stating that ``a configuration $\tm{\conf{C}}$ with flag $\tm\phi$ is well typed under typing environment $\ty{\Gamma}$''.
 Configuration typing is based on the standard typing for GV.
@@
 Main threads are allowed to return a value, whereas child threads are required to return the unit value.
-Sequents are annotated with a flag $\phi$, which ensures that there is at most one main thread.
+Sequents are annotated with a flag $\tm\phi$, which ensures that there is at most one main thread.
 
@@
 PGV satisfies progress, as PGV configurations either reduce or are in normal form.
-However, the normal forms may seem surprising at first, as evaluating a well typed PGV term does not necessarily produce \emph{just} a value.
+However, the normal forms may seem surprising at first, as evaluating a well typed PGV term does not necessarily produce {just} a value.
 If a term returns an endpoint, then its normal form contains a thread which is ready to communicate on the dual of that endpoint.
@@
 Clearly, no more evaluation is possible, even though the configuration contains the thread $\tm{\child\;\echo_{x}}$, which is blocked on $\tm{x}$.
-In~\cref{cor:pgv-closed-progress} we will show that if a term does not return an endpoint, it must produce \emph{only} a value.
+In~\cref{cor:pgv-closed-progress} we will show that if a term does not return an endpoint, it must produce {only} a value.
 
@@
 \emph{Ready terms} are terms which perform communication actions, either by themselves, \eg creating a new channel or thread, or with another thread, \eg sending or receiving.
-It is worth mentioning that the notion of readiness presented here is akin to \emph{live} processes  introduced by Caires and Pfenning \cite{cairespfenning10,DardhaP22}, and \emph{poised} processes introduced by Pfenning and Griffith \cite{PfenningG15} and later used by Balzer \etal \cite{balzerpfenning17,balzertoninho19}.
+It is worth mentioning that the notion of readiness presented here is akin to {live} processes introduced by Caires and Pfenning \cite{cairespfenning10,DardhaP22}, and {poised} processes introduced by Pfenning and Griffith \cite{PfenningG15} and later used by Balzer \etal \cite{balzerpfenning17,balzertoninho19}.
 Ready processes like live/poised processes denote processes that are ready to communicate on their providing channel.
@@
 Furthermore, no term $\tm{M_i}$ can perform a communication action by itself, since those are excluded by the definition of actions.
-Finally, as a corollary, we get that well typed terms which do not return endpoints return \emph{just} a value:
+Finally, as a corollary, we get that well typed terms which do not return endpoints return {just} a value:
 

### 2022-12-12 lmcs/introduction.tex
update Done

@@
 
-In the $\pi$-calculus literature, there have been several attempts at developing Curry-Howard correspondences between session-typed $\pi$-calculus and linear logic~\cite{girard87}: Caires and Pfenning's $\pi$DILL~\cite{cairespfenning10} corresponds to dual intuitionistic linear logic~\cite{barber96}, and Wadler's Classical Processes~\cite[CP]{wadler14} corresponds to classical linear logic~\cite[CLL]{girard87}.
+In the $\pi$-calculus literature, there have been several developments of Curry-Howard correspondences between session-typed $\pi$-calculus and linear logic~\cite{girard87}: Caires and Pfenning's $\pi$DILL~\cite{cairespfenning10} corresponds to dual intuitionistic linear logic~\cite{barber96}, and Wadler's Classical Processes~\cite[CP]{wadler14} corresponds to classical linear logic~\cite[CLL]{girard87}.
 Both calculi guarantee deadlock freedom, which they achieve by the combination of binary session types with a restriction on the way processes can be connected by channels: two processes can share at most one channel, and, more generally, information transmitted between any two processes must pass through one unique series of channels and intermediate processes.
@@
 Within a concurrent $\lambda$-calculus, one can derive extensions of the communication capabilities of the language via well-understood extensions of the functional fragment, \eg, we can derive internal/external choice from sum types.
-Concurrent $\lambda$-calculi maintain a clear separation between the program which the user writes and the configurations which represent the state of the system as it evaluates the program.
+Concurrent $\lambda$-calculi, among other languages, maintain a clear separation between the program which the user writes and the configurations which represent the state of the system as it evaluates the program.
 However, our main motivation is that results obtained for $\lambda$-calculi transfer more easily to real-world functional programming languages.

### 2022-12-12 lmcs/conclusion.tex
Address ties to logic

@@
 In addition to the more standard results, the authors also prove a global progress theorem.
-Dardha and Gay~\cite{dardhagay18extended} define Priority CP by integrating Kobayashi and Padovani's work on priorities~\cite{kobayashi06,padovani14} with CP, loosening its ties to linear logic in exchange for expressivity.
+Dardha and Gay~\cite{dardhagay18extended} define Priority CP by integrating Kobayashi and Padovani's work on priorities~\cite{kobayashi06,padovani14} with CP, weakening its ties to linear logic in exchange for expressivity.
 However, they show how PCP can be also {viewed as} a an extension of linear logic, which they call Priority Linear Logic (PLL), and uses mix and cycle rules as opposed to the cut rule.

### 2022-12-12 lmcs/introduction.tex
Address ties to logic

@@
 There have been various approaches to decoupling restriction and parallel composition.
-Hypersequent CP~\cite{MP18,kokkemontesi19popl,kokkemontesi19tlla}, Hypersequent GV~\cite{fowleretal21}, and Linear Compositional Choreographies~\cite{CarboneMS18} decouple them, but maintain the correspondence to linear logic and allow only tree-structured processes.
-Priority CP~\cite[PCP]{dardhagay18extended} weakens the correspondence to CLL in exchange for a more expressive language which allows cyclic-structured processes.
+Hypersequent CP~\cite[HCP]{MP18,kokkemontesi19popl,kokkemontesi19tlla}, Hypersequent GV~\cite{fowleretal21}, and Linear Compositional Choreographies~\cite{CarboneMS18} decouple them, but maintain the tree-structure of processes and a correspondence to linear logic, \eg, while the typing rules for HCP are no longer exactly the proof rules for CLL, every typing derivations in HCP is isomorphic a proof in CLL, and vice versa.
+Priority CP~\cite[PCP]{dardhagay18extended} weakens the correspondence to CLL.
+PCP is a non-conservative extension of CLL: every proof in CLL can be translated to a typing derivation in PCP~\cite{DardhaP22}, but PCP can prove strictly more theorems, including (partial) bijections between several CLL connectives.
+In exchange, PCP has a much more expressive language which allows cyclic-structured processes.
 PCP decouples CP's cut rule into two separate constructs: one for parallel composition via a mix rule, and one for name restriction via a cycle rule.

### 2022-12-12 lmcs/priority-gv.tex
Added text to subject reduction.

@@
 Linearity is ensured by splitting environments on branching rules, requiring that the environment in the variable rule consists of just the variable, and the environment in the constant and unit rules are empty.
-Constants $\tm{K}$ are typed using type schemas, and embedded using \LabTirName{T-Const} (mid of \cref{fig:pgv-typing}).
+Constants $\tm{K}$ are typed using type schemas, which hold for any concrete assignment of types and priorities to their meta-variables.
+Instantiated type schemas are embedded into typing derivations using \LabTirName{T-Const} in~\cref{fig:pgv-typing}, \eg, the type schema for $\tm{\send}$ can be instantiated with $\cs{o}=\cs{2}$, $\ty{T}=\ty{\tyunit}$, and $\ty{S}=\ty{\tyvoid}$, and embedded using \LabTirName{T-Const} to give the following typing derivation:
+\[
+  \inferrule*{
+  }{\tmty{\send}{\tylolli[\cs{\ptop},\cs{2}]{\typrod{\tyunit}{\tysend[\cs{2}]{\tyunit}{\tyvoid}}}{\tyvoid}}}
+\]
+
 The typing rules treat {all variables} as linear resources, even those of non-linear types such as $\ty{\tyunit}$, though they can easily be extended to allow values with unrestricted usage~\cite{wadler14}.

### 2022-12-13 lmcs/priority-gv.tex
well-typed

@@
 Consequently, we can embed it directly into the reduction relation.
-In GV, this is not the case, and subject reduction relies on proving that if $\equiv\cred$ ends up in an ill-typed configuration, we can rewrite it to a well typed configuration using $\equiv$.
+In GV, this is not the case, and subject reduction relies on proving that if $\equiv\cred$ ends up in an ill-typed configuration, we can rewrite it to a well-typed configuration using $\equiv$.
 \end{enumerate}
@@
 It is an \emph{approximation} on function terms, as these can ``skip'' communication by returning the corresponding channel unused.
-However, linearity prevents such functions from being used in well-typed configurations: once the unused channel's priority has passed, it can no longer be used.
+However, linearity prevents such functions from being used in well typed configurations: once the unused channel's priority has passed, it can no longer be used.
 
@@
 PGV satisfies progress, as PGV configurations either reduce or are in normal form.
-However, the normal forms may seem surprising at first, as evaluating a well typed PGV term does not necessarily produce {just} a value.
+However, the normal forms may seem surprising at first, as evaluating a well-typed PGV term does not necessarily produce {just} a value.
 If a term returns an endpoint, then its normal form contains a thread which is ready to communicate on the dual of that endpoint.
@@
 Furthermore, no term $\tm{M_i}$ can perform a communication action by itself, since those are excluded by the definition of actions.
-Finally, as a corollary, we get that well typed terms which do not return endpoints return {just} a value:
+Finally, as a corollary, we get that well-typed terms which do not return endpoints return {just} a value:
 

### 2022-12-13 lmcs/conclusion.tex
Update conclusion.tex

@@
 In addition to the more standard results, the authors also prove a global progress theorem.
-Dardha and Gay~\cite{dardhagay18extended} define Priority CP by integrating Kobayashi and Padovani's work on priorities~\cite{kobayashi06,padovani14} with CP, weakening its ties to linear logic in exchange for expressivity.
+Dardha and Gay~\cite{dardhagay18extended} define Priority CP by integrating Kobayashi and Padovani's work on priorities~\cite{kobayashi06,padovani14} with CP, which as described in the introduction, weakens its ties to linear logic in exchange for expressivity.
 However, they show how PCP can be also {viewed as} a an extension of linear logic, which they call Priority Linear Logic (PLL), and uses mix and cycle rules as opposed to the cut rule.

### 2022-12-13 lmcs/REBUTTAL-LETTER.md
response letter

# Response to the reviewers

We thank the reviewers for their useful feedback and comments. We addressed their concerns in the paper and respond below detailing how we went about these changes.

# Review 1

## Connection to logic

- 2nd paragraph: What do you mean when you say that Hypersequent CP and Linear Compositional Choreographies "maintain" the correspondence to CLL?

  The authors: We elaborated about this in page 2, now 3rd paragraph.

- 2nd paragraph: You say that PCP weakens the correspondence to CLL. What do you mean, exactly? Understanding this is important to gauge the relevance of PGV as well.

  The authors: We elaborated about this in page 2, now 3rd paragraph.

- page 39, 2nd line: Again, what do you mean by "loosening its ties to linear logic"?

  The authors: We adjusted this line and connected it to the explanation provided in the intro.

## Comments on technical developments
- page 5: Please explain the syntax of terms carefully. For example, you have a term for sequential composition (M;N) in a functional language, but you] don't even mention it in the paragraph that describes terms. The term does not even appear in the semantics.

- page 12: Where is the case for `M;N`? Is it obtained by composing N of type unit with something like what is typed in T-LamUnit?

- page 12, case T-LamUnit: I could not find a typing rule for a term `let () = ...`, which would be necessary here.

- 3rd paragraph: What is "reflection"?

  The authors: We do not need a typing rule for whichever one we end up using, as it'll be a part of the core language. We added informal definition for simulation and reflection. Also, we removed "reflection" and substituted "operational correspondence".

  @elektra85: @wenkokke please check the above and adjust

- You say in the abstract that deadlock-freedom is guaranteed in CP and GV by combining parallel composition/thread spawning with channel creation. Later in the Introduction, you say that deadlock-freedom actually depends on restricting the structure of processes and shared channels to trees, which is achieved by the aforementioned combination. But then you mention works (Hypersequent CP and Linear Compositional Choreographies) that achieve the tree structure without that combination. So the really important thing seems to be the "tree structure" of processes. What do you mean by that exactly? You say that "this ensures that two processes can only communicate via exactly one series of channels". What's a "series of channels"? A mini example on what can and cannot be written in CP / Hypersequent CP / Linear Compositional Choreographies would probably help.

  The authors: We clarifed this in page 2.

- The Introduction should mention that this article is the extended version of a published conference paper.

  The authors: We have now added a paragraph at the end of the introduction.

- Please write some introductory text for Section 4. At this point I had forgotten most about what PCP is and why it matters for this paper.

  The authors: We have added an opening to sec 4, clarifying what PCP is and why it matters to us.

- page 10: in "a configuration C with flag \phi" you highlight C but not \phi (and likewise in the second-last line of the same paragraph). What's the rationale for colouring things?

  The authors: we have fixed this now.

- The document is full of emphasised words, a bit too much for my taste. I'd suggest to highlight only keywords that should be remembered for later, not words that the reader should really read (the reader should read everything, and if not that part should be erased)---an example is the emphasis on "alone" in line 4 of page 2. This might be a matter of taste so it's not a strong suggestion.

  The authors: we have fixed this now.

- page 4: You use a different font than usual for the empty set symbol, so maybe write explicitly in parenthesis that it is what you mean when you use the symbol for the first time?

  The authors: The rationale for colouring terms is to communicate their syntactic sort: red for terms, blue for types, teal for priorities & priority constraints. This may not have been consistently applied.

## Related Work

- 2nd paragraph: You say that Linear Compositional Choreographies maintain a correspondence to CLL, but they are based on Intuitionistic Linear Logic.

  The authors: We have now fixed this.

- page 27: Rule T-Halt recalls standard session types rather than linear logic. Could you elaborate on what the link between PCP and linear logic is? See also my previous comment on the introduction.

  The authors: We have now added section 4.4 on PCP and PLL to clarify this.

- 3rd paragraph of Section 6: Similarly, I don't think the average reader will know what a "global type" is. Might be worth saying something about it.

  The authors: We have now clarified this.

- 3rd paragraph of Section 6: You say that [CM13] combine MPST with choreographies to achieve deadlock-freedom. But global types in MPST _are_ choreographies. Do you mean choreographic programming ("combine MPST with choreographic programming")?

  The authors: Yes, thank you. We have fixed this now.

- By the way, just like global types, I don't think that the average reader will know what choreographic programming is.

  The authors: We have added a description of choreographies in this paragraph.

- 2nd paragraph: You say that Linear Compositional Choreographies maintain a correspondence to CLL, but they are based on Intuitionistic Linear Logic.

  The authors: We have fixed this.

- 1st paragraph of Section 6: As far as I know one of the most powerful of Kobayashi's systems is the one in

  > Naoki Kobayashi, Cosimo Laneve
  > Deadlock analysis of unbounded process networks. Inf. Comput. 252: 48-70 (2017)

  so it might be worth citing. This system seems much more powerful than the one presented in this paper (but has no link to linear logic, as far as I know, so it doesn't detract from the authors' contribution).

  The authors: we have now added this related work and commented about it, thanks.

- Your work and GV in general reminded me a lot of this paper:

  > Bernardo Toninho, Luís Caires, Frank Pfenning:
  > Higher-Order Processes, Functions, and Sessions: A Monadic Integration. ESOP 2013: 350-369

  where the session-typed pi-calculus (with linear logic) is integrated with functions. Probably also worth mentioning.

  The authors: we have now added this related work and commented about it, thanks.

## Minor comments
- page 26: you sometimes say "well typed" and other times "well-typed".

  The authors: we fixed this and use consistently well typed.

- page 8: "to applied" -> "to be applied"

- 1st paragraph of Section 6: "andcapabilities" -> "and capabilities"

- page 6: The syntax of configuration needs a line break or some spacing, I think.

- page 6: I couldn't really make sense of the equations for the markings. I proceeded by guessing what spacing should have been there.

- last line of page 1: I think you lack a colon at the end.

- page 26: T-Res sTates

- 1st line: "verify correct ordering of operations, namely protocols". Don't session types verify _implementations_ of protocols, instead of the protocols themselves?

- 2nd line: "Regular types". I don't think you mean regular types in the technical sense. Perhaps you mean data types?

  The authors: thank you for pointing out the typos and minor issues. We have now fixed them in the paper.

# Review 2

## Presentation and text

- References: It would be nice if references included DOI links.

  The authors: we will add the DOIs to the final version; we kept the bibliography succinct because of the page limit enforced by the journal. We will discuss this with the editors.

- Abstract: It seems rather unusual to include citations.

  The authors: it is not uncommon for references to be included in the abstract. We think it gives a more complete picture, while allowing the abstract to remain succinct. Without the references we would need to expand on several notions, which defeats the purpose of a succinct abstract.

- P2, ”there have been several attempts at developing Curry-Howard correspondences”: Attempts, implies that it failed.

  The authors: we have fixed this now.

- P2, ”Concurrent lambda-calculi maintain a clear separation between the program ... and the configurations”: True also for some other session-typed languages.

  The authors: we have fixed this now.

- P3, ”Priority GV offers a more fine-grained analysis of communication structures”: Clarify?

  The authors: we have now clarified this in the paper.

- P4: Why does top denote a lower bound and bottom an upper bound? Shouldn’t it be the other way around?

  The authors: we have now clarified this in the paper in the typing rule section, right after pure functional types.

- P5, ”top-most connective”: Left-most?

  The authors: we chose top-most because it refers to the tree of the type. Also,  we believe that top-most is unambiguous.

- P6, grammar, bottom of page: Why is K (defined earlier in the term syntax) a value? Shouldn’t K be mentioned in rhs of E? Also, I do not understand the need for the thread evaluation context, nor do I understand how it fits with the remaining notions, E and G. F doesn’t show up there. Could it be that F should appear in the rhs of G? Maybe instead of the hole?~~

  The authors: Functions are values, K is the sort for builtin functions, so K is a value. The thread evaluation context is separately useful. The contexts F and G have holes of different types. They are used in different places.

- SC-ResComm: Why is the side condition necessary? Doesn’t name restriction generate fresh names?

  The authors: Following the operational semantics, sure, but it's not encoded in the term syntax. Your suggestion amounts to moving the side-condition to well-formedness of the term syntax.
  @elektra85: rephrase the above to make it clear and say it with more kindness.

- P8, ”We no longer require that every child thread finishes by returning a terminated channel”: Doesn’t that have implications for linearity?

  The authors: It doesn't. In previous settings, each child thread returned its terminated channel to be closed by the "runtime system". In our setting, it is enforced by the type system that each child thread must close its terminated channel.

- P8, middle of page, in (b): not sure how to parse this arrow/implication?

  @wenkokke: Which? The composition of equality and reduction?
  @elektra85: just say something, even if trial (add it below)
  The authors:

- Fig2: rule T-Absurd would admit weakening, so is wrong.

  @elektra85: clarify this in the paper and here.

- T-Pair: why should p < minpr(∆)?

  The authors: we need this condition becuase of the operational semantics evaluating products. They're ordered. First field goes first.

- Lemma 3.2: Also, shouldn’t there be a relationship on p and q?

  The authors: There is no need for a relationship between p and q because by Lemma 3.1 q=\bottom for all cases.

- The proof of Lemma 3.3 is sloppy. It omits the typing of evaluation contexts, which is all glossed over in the last case, dismissed as immediate. Also, it is stated for open terms. Presumably, Gamma can only contain channel names at this stage, but that is not made precise, nor guaranteed.

  The authors: Lemma 3.3 is subject reduction for a simply-typed lambda calculus with constants, which is standard. We clarified this in the paper.

- Also, I am not familiar with type schemas, some explanation would be helpful.

  The authors: we have now fixed use of A,B,C in type schemas and added further explanation.

- Lemma 3.2 proof, it is not clear what is meant by right arrow with label V/x. What is the semantics of it? Note, I didn’t check the remaining cases. But in general I find the proof not detailed enough. Several steps are omitted.

  @elektra85: comment on this, add something in the paper.

- Fig.10: What is the semantics of this arrow/implication?

  The authors: It is the unfolding of the definition of substitution/translation. We have rewritten these to be single-lined arrows instead, which is more consistent with the remainder of the paper and the usual notation for commuting diagrams.

- P22, explanations below Lemma 3.7: This seems important, but it should be elaborated, I wasn’t able to understand the details. Also, I don’t remember that the term canonical forms has been defined. Is a term in canonical form if it is ready? (Reading on, I see that the term canonical form is now defined. So, the sequencing here is probably not ideal.)

  The authors: we added emphasis to make it apparent at the start of the sentence that the term is being defined, as we did for the other terms.

- P22: I am surprised to read that progress seems to allow ready terms. Shouldn’t a stronger progress statement be possible due to linearity?

  The authors: I believe you may be reading progress for the term language as progress for the configuration language. we have defined two different progress properties, for terms and for the whole language. In the paper, the results are labelled accordingly.

- P22, ”we opt to move all nu-binders to the top”: Where is that done?

  The authors: This is done in the definition of canonical forms that's referenced in the reviewer's remark, as well as in the associated canonical forms proof by "using SC-ResExt".

- P24, pr(): here pr is used, earlier minpr. Not sure whether that is intended, but it also caused mistakes in an earlier proof (see earlier remark). Actually, this definition uses again minpr.

  The authors: We changed the definitions of \pr and \minpr. We now overload \pr to work on session types, data types and contexts, and remove \minpr.

- P26, ”Finally, since our reduction relation is a strict subset of the reduction relation in the original [DG18b], we defer to their proofs.”: I am not convinced by this argument, i.e., why should the proof that was carried out for a bigger system hold for a subset?

  The authors: thanks for noticing this, we meant subject reduction (Thm 2) specifically and we clarified this in the paper.

- Lemma 4.5: this lemma seems to rely on a translation (notation (.)), which has not been introduced yet, as far as I can tell! See also my remark in the previous section.

  The authors: We have included our definition for the translations from PCP to PGV, which had been erroneously removed from the journal submission.

  @wenkokke: Ensure that we consistently use either `M; N` or `let () = M in N`. In my opinion, all uses of the latter should be replaced by the former, and `let () = M in N` should not occur in the paper.

  @elektra85: what is the above comment and where should it go?

- Proof of Lemma 3.4: it does not seem to go by induction of C = C’ because structural congruence is not an inductive definition. Instead, it seems typing derivations of the structurally equal configurations are provided and it is shown that the same type is derived.

  The authors: Derivations of structural congruence are an inductive definition, and we say that we proceed "by induction on the derivation of C = C'". The language is finite, so there are only inductive definitions.

- Lemma 3.7: here the remark is made that Gamma contains only session types. Reminding of my earlier remark regarding closed terms, I think this remark should also be made for subject reduction?

  The authors: "Only contains session types" means that the term is closed wrt variables, but may become blocked on a channel, i.e., be ready. So this is essentially progress for variable-closed terms. If there's variables, obviously progress becomes trickier, since we may also become blocked on a variable. Subject reduction holds for all terms.

- Figures 3-5: I have not checked them for mistakes. What is the significance of providing this syntactic sugar?

  The authors: We use syntactic sugar in the examples. In addition, giving syntactic sugar for select and offer is a witness to the expressivity of our language. Select/offer are normally primitives in session typed pi-calculus.

- P8, last sentence: At this point it’s not clear to me how the lower bound is approximated.

  The authors: we addressed this at the third paragraph of sec 2.3.

- P6: again, the definitions for the flags lack space to separate them. Also, this seems not a proper inductive definition, i.e., base cases are missing. Moreover, what seems required is a left to right reading here, i.e., the right to left reading doesn’t seem to make sense. I think the proper way to define this would be as an inductive definition over the configuration, in particular decomposing C||D. The introduction of ”+” seems unnecessary, moreover the semantics of it is not defined. And, after having read the paper, these definitions are later on never used.

  The authors: We added the required space. "+" is used in ...
  @elektra85: wen pls complete the above.

- Another question I was left wondering about is whether priorities are inferred by typing or whether the programmer has to provide them. The former seems to be the case, but this should be made explicit. Also, it would be helpful to provide a complete, unsuccessful typing derivation for Example 2.2. Two individual derivations, for each thread, are given on page 10, which show that the derivations impose contradicting orderings on the priorities. It would then also be helpful to show the rule that composes the two derivations and fails because of the contradiction. Which rule would that be?

  The authors: we added the following passage:

  > For simplicity, we assume priority annotations are not inferred, but provided as an input to type checking. However, for any term, priorities can be inferred, e.g., by using the topological ordering of the directed graph where the vertices are the priority meta-variables and the edges are the inequality constraints between the priority meta-variables in the typing derivation.

  @wenkokke: Add an explanation to the type of pure functions.
  @elektra85: wen address your own comment above.

- About top and bottom on lolli: aren’t they used counter standard usage? I.e., shouldn’t the greatest upper bound be top? After all the priorities on the sequent are joined.

  The authors: In the paper we have the following sentence.
  > The lower bound for a pure function is top as pure functions never start communicating. For similar reasons, the upper bound for a pure function is bottom.

- Fig2, runtime typing: Shouldn’t these rules incorporate the earlier defined + operator to ensure that there is only one main thread?

  The authors: They literally do: T-Par.
  @elektra85: adjust the above.

- P9: At first, I didn’t see how the typing rule for New ensures that both endpoint have same priority. Maybe remind the reader that S and S requires priorities to be the same.

  The authors: In the paper we have the following sentence.

  > _(immediately after the figure)_
  >
  > "The typing rule for new ensures that both endpoints of a channel share the same priorities."

## Minor Changes

We have fixed all the minor comments below.

- P2, ”communicate via exactly one series of channels”: What is meant by series of channels?

- @Reviewer2: Fix occurances of bare `new`.

- @Reviewer2: P28, ”Let this process be let this be Pi...”: grammar

- @Reviewer2: P36: It would be helpful to already alert the reader to the figure on the next page when discussing the setup here.

- @Reviewer2: P24, pr(): here pr is used, earlier minpr. Not sure whether that is intended, but it also caused mistakes in an earlier proof (see earlier remark). Actually, this definition uses again minpr.

- Lemma 3.2 proof, 2nd case: pr() should be minpr().

- Lemma 3.2: Metavariable Theta has not been introduced earlier. For typing environments, only Gamma and Delta have been introduced.

- P2, ”processes and shared channels to trees”: What is meant by shared channels? Please clarify.

  The authors: A channel is shared by two processes if each holds an endpoint.

- Proof of Lemma 3.1 misses cases T-Var and T-Const.

- P10: Example lambda abstraction, I don’t think that the metavariable A was ever introduced.

  The authors: thanks, that should be T.

- P8, ”In PGV, we keep track of a lower and upper bound on the priorities of a term, i.e., while evaluating the term, when does it start communicating, and when does it finish”: Nitpicking, but this is not really properly formulated. What does the ”it start” refer to? Use of resp. would be helpful.

  The authors: we fixed it.

- P8, ”allows restriction to applied”: to be applied

- P6, ”overlap between flags and priorities”: Could you clarify what the implications of this overlap are? Or does it just mean that the color indicates what ”circle” means?

  The authors: Yes; we condensed that paragraph.

- P2, ”can be tweaked to satisfy reflection”: What is meant by reflection here?

- P3, ”is typeable and guaranteed deadlock-free”: Guaranteed to be

- P5: ”to denote the minimum and maximum”: ..., respectively. Also, that listing seems to be missing max!

  The authors: we defined max here as we need it and it is used later.

- P5: The programs in Example 2.1 are not well-formed wrt abstract syntax of terms defined on page 5.For example, the term ”recv x” cannot be parsed.

  The authors: the reviewer has misunderstood this. Note for eg: "recv x" is "M N" with "M = K = recv" and "N = x"
  @elektra85: wen say somthing more above, as atm it doesn't make much sense.

- P6: I am not sure how to parse the grammar of configurations. The two productions seem not clearly separated. I believe it should be phi ::= closed circle — open circle and then separately, C, D, E ::== ...?

- P2, ”succinctness of process calculi”: What is meant?

  The authors: we mean compared to concurrent lambda calculi.

- P2, ”they achieve by restricting structure of”: the structure of

## Related work

The reviewer suggested adding an comparing to the following related work.
- M. Coppo, M. Dezani-Ciancaglini, L. Padovani, and N. Yoshida. Inference of global progress prop- erties for dynamically interleaved multiparty sessions. In COORDINATION, 2013. doi: 10.1007/ 978-3-642-38493-6\ 4.

- M. Coppo, M. Dezani-Ciancaglini, N. Yoshida, and L. Padovani. Global progress for dynamically interleaved multiparty sessions. MSCS, 26(2):238–302, 2016. doi: 10.1017/S0960129514000188. URL https://doi.org/10.1017/S0960129514000188.

- B. Toninho, L. Caires, and F. Pfenning. Higher-order processes, functions, and sessions: A monadic integration. In ESOP, volume 7792 of LNCS, pages 350–369, 2013. doi: 10.1007/978-3-642-37036-6\ 20. URL https://doi.org/10.1007/978-3-642-37036-6\_20.

  The authors: many thanks for the suggestion, we have now added all the above related work and compared to them.

# Review 3

- I disagree with the (informal) definition of session interleaving used in e.g. paragraph 2, page 2, of the introduction. The authors state that:

  > [...] This ensures that two processes can only communicate via exactly one series of channels, which rules out interleavings of sessions, and guarantees deadlock freedom
  > To my understanding, session interleaving is the ability of a process to communicate on more than one channel, which is clearly supported by linear logic-based systems in a deadlock-free way. This is in sharp contrast with the original session type works where the sheer fact that a process can use more than one channel entails the possibility of deadlock. What is ruled out is cyclic dependencies of blocking actions, not interleaving.

  The authors: "What is ruled out is cyclic dependencies of blocking actions, not interleaving." That's actually not true, or not at least overly rose-tinted. What linear logic based systems ensure is that information can only flow from one process to another across one unique path. Since that path consists of session-typed channels, and session types rule out cyclic dependencies on session-typed channels, deadlock freedom follows. Sure, that *does* rule out "cyclic dependencies of blocking actions", but it rules out much more than that. Priorities rule out *exactly* all cyclic dependencies of blocking actions, and nothing else.

- The authors should state clearly what the downsides of coupling channel creation with thread spawning in a setting such as GV are (beyond the acyclicity constraint).

  The authors: we have added a clarification that GV suffers from a complicated metatheory as a consequence.

## Minor comments

- page 2, paragraph 2: combing -> combining

- page 6: formating of \phi ::= ... is broken

- Section 6, paragraph 2: conservation types -> conversation

- page 10: l1, enforce sequentially

The authors: many thanks for this. We have addressed typos and inconsistencies.

## Related work

- The issue raised in the abstract about CP is coupling thread spawning and channel creation is true also about the earlier work of Caires and Pfenning, and is generally true of all works in this space based on the core proof theory of linear logic. I suggest reflecting this in the abstract rather than singling out CP in this sense.

  The authors: We have now added a paragraph about this in the abstract.

- The notion of Ready Term (Def. 3.6) seems related to the notion of poised process of [BP17,BTP19] and live process of [CP10]. A remark to this effect would be helpful.

  The authors: we have now commented about it and restructured this part of the paper.

- Since the orignal publication of this work, a few works on deadlock-freedom by typing that go beyond the languages that stem directly from the logical interpretation of session types have been published that should be included in the related work discussion. Please comment on the recent work by Rocha and Caires on shared state in a language inspired by differential linear logic:

  > Pedro Rocha, Luís Caires:
  > Propositions-as-types and shared state. Proc. ACM Program. Lang. 5(ICFP): 1-30 (2021)

  The authors: thanks, we have now added this reference and commented about the work.

### 2022-12-13 lmcs/REBUTTAL-LETTER.md
Fix comment

@@
 
-- P8, middle of page, in (b): not sure how to parse this arrow/implication?
-
-  @wenkokke: Which? The composition of equality and reduction?
-  @elektra85: just say something, even if trial (add it below)
-  The authors:
-
 - Fig2: rule T-Absurd would admit weakening, so is wrong.
@@
 
-- Fig.10: What is the semantics of this arrow/implication?
+- P8, middle of page, in (b): not sure how to parse this arrow/implication?
 
-  The authors: It is the unfolding of the definition of substitution/translation.
+  Fig.10: What is the semantics of this arrow/implication?
+
+  The authors: It is the unfolding of the definition of substitution and translation, respectively.
 We have rewritten these to be single-lined arrows instead, which is more consistent with the remainder of the paper and the usual notation for commuting diagrams.

### 2022-12-13 lmcs/REBUTTAL-LETTER.md
Fix comment
Multiple fixes

@@
 
-- page 12: Where is the case for `M;N`? Is it obtained by composing N of type unit with something like what is typed in T-LamUnit?
+- page 12: Where is the case for `M;N`?
+Is it obtained by composing N of type unit with something like what is typed in T-LamUnit?
+Also, for the case T-LamUnit: I could not find a typing rule for a term `let () = ...`, which would be necessary here.
 
-- page 12, case T-LamUnit: I could not find a typing rule for a term `let () = ...`, which would be necessary here.
+  The authors: We have merged `M;N` and `let () = ...` as they are synonymous. All uses of the latter have been replaced by the former, and `let () = M in N` should no longer occur in the paper.
 
@@
 
-  The authors: We do not need a typing rule for whichever one we end up using, as it'll be a part of the core language.
-We added informal definition for simulation and reflection.
-Also, we removed "reflection" and substituted "operational correspondence".
-
-  @elektra85: @wenkokke please check the above and adjust
+  The authors: We added informal definition for simulation and reflection. Also, we removed "reflection" and substituted "operational correspondence".
 
@@
 
-  @wenkokke: Ensure that we consistently use either `M; N` or `let () = M in N`.
-In my opinion, all uses of the latter should be replaced by the former, and `let () = M in N` should not occur in the paper.
-
-  @elektra85: what is the above comment and where should it go?
-
 - Proof of Lemma 3.4: it does not seem to go by induction of C = C’ because structural congruence is not an inductive definition.
@@
 
-  @wenkokke: Add an explanation to the type of pure functions.
-  @elektra85: wen address your own comment above.
-
 - About top and bottom on lolli: aren’t they used counter standard usage? I.e., shouldn’t the greatest upper bound be top? After all the priorities on the sequent are joined.
@@
 
-  The authors: They literally do: T-Par.
-  @elektra85: adjust the above.
+  The authors: The T-Par incorporates the + operator.
 
@@
 
+- We have fixed all occurances of bare `new`.
+
 - P2, ”communicate via exactly one series of channels”: What is meant by series of channels?
-
-- @Reviewer2: Fix occurances of bare `new`.
 

### 2022-12-13 lmcs/REBUTTAL-LETTER.md
minor
thm 4.4

@@
 
+## General comments
+
+- The reviewer comments that they cannot write down derivation for `• let (x, x′) = new in spawn(λ().echo x); x′`.
+
+  The authors: we apologize, there is typo in the term and it should be `• let (x, x′) = new () in spawn(λ().echo x); x′`. We have fixed this in the paper.
+
 - P2, ”there have been several attempts at developing Curry-Howard correspondences”: Attempts, implies that it failed.
@@
 Subject reduction holds for all terms.
+
+- Proof of Thm 4.4: this is not a rigorous argument.
+I would like to understand better the architecture of the proof.
+It seems to be going by cases, so I want to be convinced that they are exhaustive.
+Also, various inferences are not properly justified.
+
+  The authors: we have marked this proof as a sketch. We have also added the following comment:
+  >  This proof follows the exact same reasoning and proof sketch given by Kobayashi in \cite{kobayashi06} and later adopted by Dardha and Gay for PCP in their technical report \cite{dardhagay18extended}.
 

### 2022-12-13 lmcs/prf/thm-pcp-closed-progress.tex
thm 4.4

@@
-\begin{proof}
+\begin{proof}(Sketch)
   \label{prf:thm-pcp-closed-progress}
+
+  This proof follows the exact same reasoning and proof sketch given by Kobayashi in \cite{kobayashi06} and later adopted by Dardha and Gay for PCP in their technical report \cite{dardhagay18extended}.
+
   By~\cref{lem:pcp-canonical-forms}, we rewrite $\tm{P}$ to canonical form. If the resulting process is $\tm{\halt}$, we are done. Otherwise, it is of the form

### 2022-12-14 lmcs/REBUTTAL-LETTER.md
No more TODO's in REBUTTAL_LETTER

@@
 
-  The authors: Following the operational semantics, sure, but it's not encoded in the term syntax. Your suggestion amounts to moving the side-condition to well-formedness of the term syntax.
-  @elektra85: rephrase the above to make it clear and say it with more kindness.
+  The authors: Yes, operationally, the `new` function generates name restrictions with fresh names.
+However, this is not encoded as a property of the _syntax_ of terms.
+Therefore, to ensure that the equivalence works for any configuration, and not just configurations which are reachable by the operational semantics, it needs that side-condition.
 
@@
 
-  @elektra85: clarify this in the paper and here.
+  The authors: The rule T-Absurd is exfalso, i.e., from a proof of false, or a term of the empty type, anything follows, including weakening.
+
+  We believe that T-Absurd can only be used to derive weakening in the presence of a proof of false.
+If the reviewer believes that T-Absurd admits general weakening, we would welcome them to elaborate.
 
@@
 
-  The authors: we need this condition becuase of the operational semantics evaluating products. They're ordered. First field goes first.
+  The authors: We need this condition because of the operational semantics evaluating products, which evaluate the fields of a pair in order.
 
@@
 
+- P8, middle of page, in (b): not sure how to parse this arrow/implication?
+
+  The authors: This is the composition of the structural congruence and the reduction relation. We split them up and elaborated in the text:
+
+  > In GV, this is not the case, and subject reduction relies on proving that if the result of rewriting via ≡ followed by reducing via −→C is an ill-typed configuration, we can rewrite it to a well-typed configuration via ≡.
+
 - Lemma 3.2 proof, it is not clear what is meant by right arrow with label V/x.
@@
 Several steps are omitted.
-
-  @elektra85: comment on this, add something in the paper.
-
-- P8, middle of page, in (b): not sure how to parse this arrow/implication?
 
@@
 
-  The authors: We added the required space. "+" is used in ...
-  @elektra85: wen pls complete the above.
+  The authors: We added the required space. "+" is used in the T-Par rule.
 
@@
 
-- @Reviewer2: P28, ”Let this process be let this be Pi...”: grammar
+- P28, ”Let this process be let this be Pi...”: grammar
 
-- @Reviewer2: P36: It would be helpful to already alert the reader to the figure on the next page when discussing the setup here.
+- P36: It would be helpful to already alert the reader to the figure on the next page when discussing the setup here.
 
-- @Reviewer2: P24, pr(): here pr is used, earlier minpr.
-Not sure whether that is intended, but it also caused mistakes in an earlier proof (see earlier remark).
-Actually, this definition uses again minpr.
+- P24, pr(): here pr is used, earlier minpr. Not sure whether that is intended, but it also caused mistakes in an earlier proof (see earlier remark). Actually, this definition uses again minpr.
 
@@
 
-- P5: The programs in Example 2.1 are not well-formed wrt abstract syntax of terms defined on page 5.For example, the term ”recv x” cannot be parsed.
+- P5: The programs in Example 2.1 are not well-formed wrt abstract syntax of terms defined on page 5. For example, the term ”recv x” cannot be parsed.
 
-  The authors: the reviewer has misunderstood this. Note for eg: "recv x" is "M N" with "M = K = recv" and "N = x"
-  @elektra85: wen say somthing more above, as atm it doesn't make much sense.
+  The authors: The term `recv x` is well-formed. Specifically, it is an application (of the form `M N`) where:
+    * `M` is a constant (of the form `K`), specifically, it is `recv`; and
+    * `N` is a variable (of the form `x`).
 

### 2022-12-14 lmcs/priority-gv.tex
No more TODO's in REBUTTAL_LETTER

@@
 While our configuration reduction is based on the standard evaluation for GV, the increased expressiveness of PGV allows us to simplify the relation on two counts.
-\begin{enumerate}[label=(\alph*)]
+\begin{enumerate}[label=(\roman*)]
 \item
@@
 Consequently, we can embed it directly into the reduction relation.
-In GV, this is not the case, and subject reduction relies on proving that if $\equiv\cred$ ends up in an ill-typed configuration, we can rewrite it to a well-typed configuration using $\equiv$.
+In GV, this is not the case, and subject reduction relies on proving that if the result of rewriting via $\equiv$ followed by reducing via $\cred$ is an ill-typed configuration, we can rewrite it to a well-typed configuration via $\equiv$.
 \end{enumerate}
@@
   If $\tseq[\cs{p}]{\ty{\Gamma}}{M}{T}$ and $\ty{\Gamma}$ contains only session types, then:
-  \begin{enumerate*}[label= (\alph*) ]
+  \begin{enumerate*}[label= (\roman*) ]
   \item $\tm{M}$ is a value;

### 2022-12-15 lmcs/REBUTTAL-LETTER.md
Did as many of the TODOs as I could

@@
 # Review 1
+
+- @Reviewer1: Did you find it straightforward to apply the concepts from PCP to PGV? What was challenging?
+
+  The authors: The main difficulty was due to the fact that PGV has to track starting & ending priorities, rather than just [whichever PCP stores], but we were aided in this by the work by Padovani & Novara.
+We added the following passage:
+
+  > We cannot straightforwardly adapt the priority typing from PCP to PGV, as PGV adds higher-order functions.
+Instead, the priority typing for PGV follow the work by Padovani and Novara~\cite{padovaninovara15}.
+
+- @Reviewer1: page 4: In your presentation of types, I wondered if there are any well-formedness conditions on priority bounds.
+
+  The authors: There are no well-formedness conditions, but there are particular lemmas that hold for them, which could be _taken as_ well-formedness conditions.
+In particular I believe the theorem is: for a sequent with lower bound p and upper bound q, either p < q, or p is top and q is bottom.
+This is relaxed somewhat by some extensions, in which case I believe we can also have "p is top and q is a concrete priority", but I don't recall to what extend that applies here.
+I believe it may have been due to recursion, in which case it wouldn't.
+We added the following passage:
+
+  > We postulate that the only function types---and, consequently, sequents---that are inhabited in PGV are pure functions and functions $\ty{\tylolli[\cs{p},\cs{q}]{T}{U}}$ for which $\cs{p}<\cs{q}$.
+
+- @Reviewer1: page 6: The paragraph on values and evaluation contexts is going to be clear only to a selected few.
+Please write some text that explains how contexts work (holes, notation for replacing a hole with a term, etc.), that values are a specialisation of terms, etc.
+
+  The authors:  We added the following passage:
+
+  > Values are the subset of terms which cannot reduce further.
+  > Evaluation contexts are one-hole term contexts, \ie, terms with exactly one hole, written $\tm{\hole}$.
+We write $\tm{\plug{E}{M}}$ for the evaluation context $\tm{E}$ with its hole replaced by the term $\tm{M}$.
+Evaluation contexts are specifically those one-hole term contexts under which term reduction can take place.
+  > Thread contexts are a convenient way to lift the notion of evaluation contexts to threads.
+We write $\tm{\plug{\conf{F}}{M}}$ for the thread context $\tm{\conf{F}}$ with its hole replaced by the term $\tm{M}$.
+  > Configuration contexts are one-hole configuration contexts, \ie, configurations with exactly one hole, written $\tm{\hole}$.
+Specifically, configuration contexts are those one-hole term contexts under which configuration reduction can take place.
+The definition for $\tm{\conf{G}}$ only gives the case in which the hole is in the left-most parallel process, \ie, it only defines $\tm{\ppar{\conf{G}}{\conf{C}}}$ and not $\tm{\ppar{\conf{C}}{\conf{G}}}$.
+The latter is not needed, as $\tm{\parallel}$ is symmetric under structural congruence, though it would be harmless to add.
+We write $\tm{\plug{\conf{G}}{\conf{C}}}$ for the evaluation context $\tm{\conf{G}}$ with its hole replaced by the term $\tm{\conf{C}}$.
+
+- @Reviewer1: Another example is that you say that you "recover fork", but at this point it's not clear what fork was in GV.
+
+  The authors: At this point the reader has been told that "[GV] combines channel creation and thread spawning into a single operation, called ``fork'', which is related to the cut construct in CP." This effectively summarizes all one needs to know about fork in GV.
+The fact that it is recoverable is only interesting to the readers familiar with GV, or interested in the exact relation to GV, and we hope that any reader unfamiliar with GV can gleam fork's original semantics from our definition.
+
+- @Reviewer1: page 22: Lemma 3.7 is fundamental and seems to be the key difference wrt PCP and Kobayashi's work.
+Is that the case?
+Is the proof trivial?
+Could you at least provide a sketch, or say at least _why_ it is trivial?
+
+  The authors: It's not _trivial_ but it's _boring_, as it's precisely the proof you'd give for any lambda calculus.
+We should mention that, essentially, this is just progress for the simply-typed linear lambda calculus, with "M is ready" falling out as extra stuck cases.
+We added the following passage:
+
+  > [Progress for PGV] deviates from progress for linear $\lambda$-calculus only in that terms may reduce to values or \emph{ready terms}, where the definition of ready terms encompasses all terms whose reduction is struck on some constant $\tm{K}$.
 
@@
 ## Minor comments
+
+- page 23: In case (4), before introducing the subcases (a) and (b), it'd be kind to say that L' is either a ready term or a value.
+
 - page 26: you sometimes say "well typed" and other times "well-typed".

### 2022-12-15 lmcs/priority-gv.tex
Did as many of the TODOs as I could

@@
 
+We postulate that the only function types---and, consequently, sequents---that are inhabited in PGV are pure functions and functions $\ty{\tylolli[\cs{p},\cs{q}]{T}{U}}$ for which $\cs{p}<\cs{q}$.
+
 \subsubsection*{Typing Environments}
@@
 We use syntactic sugar to make terms more readable: we write $\tm{\letbind{x}{M}{N}}$ in place of $\tm{(\lambda x.N)\;M}$, $\tm{\lambda\unit.M}$ in place of $\tm{\lambda z.\andthen{z}{M}}$, and $\tm{\lambda\pair{x}{y}.M}$ in place of $\tm{\lambda z.\letpair{x}{y}{z}{M}}$.
-We recover $\tm{\fork}$ as $\tm{\lambda x.\letpair{y}{z}{\new\;\unit}{\andthen{\spawn\;{(\lambda\unit.x\;y)}}{z}}}$.
+We can recover GV's $\tm{\fork}$ as $\tm{\lambda x.\letpair{y}{z}{\new\;\unit}{\andthen{\spawn\;{(\lambda\unit.x\;y)}}{z}}}$.
 
@@
 \]
+Values are the subset of terms which cannot reduce further.
+Evaluation contexts are one-hole term contexts, \ie, terms with exactly one hole, written $\tm{\hole}$.
+We write $\tm{\plug{E}{M}}$ for the evaluation context $\tm{E}$ with its hole replaced by the term $\tm{M}$.
+Evaluation contexts are specifically those one-hole term contexts under which term reduction can take place.
+Thread contexts are a convenient way to lift the notion of evaluation contexts to threads.
+We write $\tm{\plug{\conf{F}}{M}}$ for the thread context $\tm{\conf{F}}$ with its hole replaced by the term $\tm{M}$.
+Configuration contexts are one-hole configuration contexts, \ie, configurations with exactly one hole, written $\tm{\hole}$.
+Specifically, configuration contexts are those one-hole term contexts under which configuration reduction can take place.
+The definition for $\tm{\conf{G}}$ only gives the case in which the hole is in the left-most parallel process, \ie, it only defines $\tm{\ppar{\conf{G}}{\conf{C}}}$ and not $\tm{\ppar{\conf{C}}{\conf{G}}}$.
+The latter is not needed, as $\tm{\parallel}$ is symmetric under structural congruence, though it would be harmless to add.
+We write $\tm{\plug{\conf{G}}{\conf{C}}}$ for the evaluation context $\tm{\conf{G}}$ with its hole replaced by the term $\tm{\conf{C}}$.
 
@@
 
-\emph{Progress} for the term language is standard for GV, and deviates from progress for linear $\lambda$-calculus only in that terms may reduce to values or \emph{ready terms}.
+\emph{Progress} for the term language is standard for GV, and deviates from progress for linear $\lambda$-calculus only in that terms may reduce to values or \emph{ready terms}, where the definition of ready terms encompasses all terms whose reduction is struck on some constant $\tm{K}$.
 

### 2022-12-16 lmcs/REBUTTAL-LETTER.md
add a reference

@@
 
-  The authors: thanks, we have now added this reference and commented about the work.
+  Please comment on the work by Qian, Kavvos and Birkedal using coexponentials:
+
+  > Zesen Qian, G. A. Kavvos, Lars Birkedal:
+  > Client-server sessions in linear logic. Proc. ACM Program. Lang. 5(ICFP): 1-31 (2021)
+
+  The authors: thanks, we have now added these references and commented about the work.

### 2022-12-16 lmcs/REBUTTAL-LETTER.md
changing a derivation

@@
 
-- @Reviewer1: Did you find it straightforward to apply the concepts from PCP to PGV? What was challenging?
+- Did you find it straightforward to apply the concepts from PCP to PGV? What was challenging?
 
@@
 
-- @Reviewer1: page 4: In your presentation of types, I wondered if there are any well-formedness conditions on priority bounds.
+- page 4: In your presentation of types, I wondered if there are any well-formedness conditions on priority bounds.
 
@@
 
-- @Reviewer1: page 6: The paragraph on values and evaluation contexts is going to be clear only to a selected few.
+- page 6: The paragraph on values and evaluation contexts is going to be clear only to a selected few.
 Please write some text that explains how contexts work (holes, notation for replacing a hole with a term, etc.), that values are a specialisation of terms, etc.
@@
 
-- @Reviewer1: Another example is that you say that you "recover fork", but at this point it's not clear what fork was in GV.
+- Another example is that you say that you "recover fork", but at this point it's not clear what fork was in GV.
 
@@
 
-- @Reviewer1: page 22: Lemma 3.7 is fundamental and seems to be the key difference wrt PCP and Kobayashi's work.
+- page 20, case E-Close: I couldn't make sense of the derivation, where two subderivations seem to be organised vertically?.. Is that the right reading? Could you reorganise it to make it obvious?
+
+  The authors: we have now fixed this derivation.
+
+- page 22: Lemma 3.7 is fundamental and seems to be the key difference wrt PCP and Kobayashi's work.
 Is that the case?

### 2022-12-16 lmcs/REBUTTAL-LETTER.md
rebuttal and todos

@@
 The fact that it is recoverable is only interesting to the readers familiar with GV, or interested in the exact relation to GV, and we hope that any reader unfamiliar with GV can gleam fork's original semantics from our definition.
+
+- page 12: Here and in all other cases in the remainder, you write "= with \Delta" on top which usually stands for a definition.
+Is that what you mean?
+Don't you mean a simple "=" instead?
+Or perhaps you mean that "= with \Delta" stands for "which is defined as"?
+A clarification would be appreciated.
+
+  The authors: we intended to use the "= with \Delta" because the syntactic sugar rules are indeed definitions in terms of the core typing rules; they are macros for what comes after the "= with \Delta".
 
@@
 ## General comments
+
+- Another technical comment I have is on the distinction between variables and channels (which the authors refer to as names).
+The latter are generated at runtime (via new), and are also gathered, in the environment.
+I believe that’s also the reason why the usual theorems (preservation and progress) are stated relative to a non-empty context.
+However, the formalization should guarantee that the respective terms are closed.
+It would also be helpful if different meta variables were used for variables and names.
+Currently, x, y, z are used for both.
+
+  The authors: we use 'names' to refer to channel endpoints, not channels.
+Preservation results (lemma 3.3 and thm 3.1) is standard and stated on open terms.
+Term progress (lemma 3.7) is stated for terms which may contain free names but not variables.
+We agree that separating variable and channel contexts would make the theorem statement clearer however, it would complicate the type system and its meta-theory.
+Configuration progress (thm 3.2) is stated for closed configurations, which is standard.
+Results are labelled accordingly.
+We clarified this after lemma 3.7.
 

### 2022-12-16 lmcs/REBUTTAL-LETTER.md
Changes

@@
 I believe it may have been due to recursion, in which case it wouldn't.
-We added the following passage:
-
-  > We postulate that the only function types---and, consequently, sequents---that are inhabited in PGV are pure functions and functions $\ty{\tylolli[\cs{p},\cs{q}]{T}{U}}$ for which $\cs{p}<\cs{q}$.
+We added a passage to the relevant section (now page 5).
 
@@
 
-  The authors:  We added the following passage:
-
-  > Values are the subset of terms which cannot reduce further.
-  > Evaluation contexts are one-hole term contexts, \ie, terms with exactly one hole, written $\tm{\hole}$.
-We write $\tm{\plug{E}{M}}$ for the evaluation context $\tm{E}$ with its hole replaced by the term $\tm{M}$.
-Evaluation contexts are specifically those one-hole term contexts under which term reduction can take place.
-  > Thread contexts are a convenient way to lift the notion of evaluation contexts to threads.
-We write $\tm{\plug{\conf{F}}{M}}$ for the thread context $\tm{\conf{F}}$ with its hole replaced by the term $\tm{M}$.
-  > Configuration contexts are one-hole configuration contexts, \ie, configurations with exactly one hole, written $\tm{\hole}$.
-Specifically, configuration contexts are those one-hole term contexts under which configuration reduction can take place.
-The definition for $\tm{\conf{G}}$ only gives the case in which the hole is in the left-most parallel process, \ie, it only defines $\tm{\ppar{\conf{G}}{\conf{C}}}$ and not $\tm{\ppar{\conf{C}}{\conf{G}}}$.
-The latter is not needed, as $\tm{\parallel}$ is symmetric under structural congruence, though it would be harmless to add.
-We write $\tm{\plug{\conf{G}}{\conf{C}}}$ for the evaluation context $\tm{\conf{G}}$ with its hole replaced by the term $\tm{\conf{C}}$.
+  The authors: We added a passage to the relevant section (now page 7).
 
@@
 
-  > [Progress for PGV] deviates from progress for linear $\lambda$-calculus only in that terms may reduce to values or \emph{ready terms}, where the definition of ready terms encompasses all terms whose reduction is struck on some constant $\tm{K}$.
+  > [Progress for PGV] deviates from progress for linear lambda calculus only in that terms may reduce to values or _ready terms_, where the definition of ready terms encompasses all terms whose reduction is struck on some constant K.
+
+- pages 27-36: These sections are too dry.
+There's no explanation of what works as usual, what is different, how the translation from PCP to PGV works (it is not even really _defined_ before the proofs!), and what the reader should focus on.
+It almost reads like a file intended for a theorem prover.
+I suggest that you rewrite these sections by explaining these aspects, presenting selected rules and the translation with examples, and then putting what "works as usual" into an appendix.
+
+  The authors: We have added a section at the beginning of section 4.6 explaining how the translation works in prose.
 
@@
 This may not have been consistently applied.
+
+- page 23: In case (4.b), if L' is a value like an abstraction or a pair, what do you mean by "the endpoint that it is blocked on"?
+
+  The authors: We amended the definition of _ready terms_ to include the phrase "is blocked on" as an alias for "is ready to act on".
 

### 2022-12-16 lmcs/priority-gv.tex
Changes

@@
   A~term $\tm{L}$ is \emph{ready} if it is of the form $\tm{\plug{E}{M}}$, where $\tm{M}$ is of the form $\tm{\new}$, $\tm{\spawn\;N}$, $\tm{\link\;\pair{x}{y}}$, or $\tm{M}$ acts on $\tm{x}$.
-In the latter case, we say that $\tm{L}$ is \emph{ready to act on} $\tm{x}$.
+In the latter case, we say that $\tm{L}$ is \emph{ready to act on} $\tm{x}$ or is \emph{blocked on}.
 \end{definition}
@@
 \end{lemma}
+
+With ``$\ty{\Gamma}$ contains only session types'' we mean that for every $\tmty{x}{T}\in\ty{\Gamma}$, $\ty{T}$ is a session type, \ie, is of the form $\ty{S}$.
 
@@
   \end{theorem}
-  % \begin{proof}[Sketch]
-  %   Our proof follows that of Kobayashi~\cite[theorem 2]{kobayashi06}.
-We apply \cref{lem:pgv-open-progress-terms} to each thread.
-Either we obtain a reduction, or each child thread is ready and the main thread ready or a value.
-We pick the ready term $\tm{L}$ with the smallest priority bound.
-If $\tm{L}$ contains new, spawn, or a link, we apply \LabTirName{E-New}, \LabTirName{E-Spawn}, or \LabTirName{E-Link}.
-Otherwise, $\tm{L}$ must be ready on some $\tm{x_i}$.
-Linearity guarantees there is some thread  $\tm{L'}$ which acts on $\tm{x'_i}$.
-If $\tm{L'}$ is ready, priority typing guarantees it is ready on $\tm{x'_i}$, and we apply \LabTirName{E-Send} or \LabTirName{E-Close}.
-If $\tm{L'}$ is not ready, it must be the main thread returning a value.
-We move $\tm{L}$ into the $i^{\text{th}}$ position and repeat until we either find a reduction or reach normal form.
-  % \end{proof}
   \input{prf/thm-pgv-closed-progress-confs}

### 2022-12-16 lmcs/REBUTTAL-LETTER.md
Final version

@@
 
+- While I generally understand the use of priorities to rule out circular waiting dependency, I am unclear about what the invariants are that the type system maintains in terms of priorities.
+
+  The authors: We have added a comment explaining the invariant that we believe the type system maintains in terms of priorities in section 2.1 (2nd paragraph of page 5).
+
 ## Minor Changes

### 2022-12-16 lmcs/priority-cp-long.tex
Changes
Final version

@@
 \begingroup
-We illustrate the relation between PCP and PGV by defining a translation from PCP to PGV. The translation on types is defined as follows:
+We illustrate the relation between PCP and PGV by defining a translation $\cpgvC{\cdot}$ from PCP processes to PGV configurations which satisfies operational correspondence.
+
+The translation $\tm{\cpgvC{\cdot}}$ translates PCP processes to PGV configurations by translating as much as possible of the $\pi$-calculus constructs from PCP to the identical configuration constructs in PGV, \ie, all top-level $\tm{\nu}$ and $\tm{\parallel}$ constructs.
+When it encounters the first action, it translates the remainder of the processs to a term using $\tm{\cpgvM{\cdot}}$.
+The translation $\tm{\cpgvM{\cdot}}$ translates PCP processes to PGV terms by mapping the $\pi$-calculus constructs from PCP (\eg, $\tm{\nu}$, $\tm{\parallel}$, $\pcp{\tm{\send{\cdot}{\cdot}{\cdot}}}$, $\pcp{\tm{\send{\cdot}{\cdot}{\cdot}}}$, $\pcp{\tm{\recv{\cdot}{\cdot}{\cdot}}}$, \etc) to the corresponding constants in PGV (\eg, $\pgv{\tm{\new}}$, $\pgv{\tm{\spawn}}$, $\pgv{\tm{\send}}$, $\pgv{\tm{\recv}}$, \etc).
+Translating a process with $\tm{\cpgvC{\cdot}}$ is the same as translating that process with $\tm{\cpgvM{\cdot}}$ followed by several steps of configuration reduction.
+The translation $\tm{\cpgvT{\cdot}}$ tranlates session types from PCP to session types in PGV matching the translation on processes.
+
+The translation on types is defined as follows:
 \[
@@
 
-There are two separate translations on processes. The main translation, $\tm{\cpgvM{\cdot}}$, translates processes to \emph{terms}:
+The translation $\tm{\cpgvM{\cdot}}$ translates processes to \emph{terms} and maps the $\pi$-calculus constructs from PCP to the corresponding constants in PGV:
 \begin{align*}
@@
 For instance, the process $\pcp{\tm{\send{x}{y}{P}}}$ is blocked, but its translation uses $\pgv{\tm{\new}}$ and can reduce.
-To address this issue, we use a second translation, $\tm{\cpgvC{\cdot}}$, which is equivalent to translating with $\tm{\cpgvM{\cdot}}$ then reducing with \LabTirName{E-New} and \LabTirName{E-Spawn}:
+To address this issue, we introduce a second translation, $\tm{\cpgvC{\cdot}}$, which is equivalent to translating with $\tm{\cpgvM{\cdot}}$ then reducing with \LabTirName{E-New} and \LabTirName{E-Spawn}:
 \begin{align*}

### 2022-12-16 lmcs/response.md
Final response

--- 2022-12-16 lmcs/REBUTTAL-LETTER.md
@@
 
-  > We cannot straightforwardly adapt the priority typing from PCP to PGV, as PGV adds higher-order functions.
-Instead, the priority typing for PGV follow the work by Padovani and Novara~\cite{padovaninovara15}.
+  > We cannot straightforwardly adapt the priority typing from PCP to PGV, as PGV adds higher-order functions. Instead, the priority typing for PGV follow the work by Padovani and Novara.
 
@@
 
-- page 12: Here and in all other cases in the remainder, you write "= with \Delta" on top which usually stands for a definition.
+- page 12: Here and in all other cases in the remainder, you write "= with $\Delta$" on top which usually stands for a definition.
 Is that what you mean?
 Don't you mean a simple "=" instead?
-Or perhaps you mean that "= with \Delta" stands for "which is defined as"?
+Or perhaps you mean that "= with $\Delta$" stands for "which is defined as"?
 A clarification would be appreciated.
 
-  The authors: we intended to use the "= with \Delta" because the syntactic sugar rules are indeed definitions in terms of the core typing rules; they are macros for what comes after the "= with \Delta".
+  The authors: we intended to use the "= with $\Delta$" because the syntactic sugar rules are indeed definitions in terms of the core typing rules; they are macros for what comes after the "= with $\Delta$".
 
@@
 
-- page 10: in "a configuration C with flag \phi" you highlight C but not \phi (and likewise in the second-last line of the same paragraph). What's the rationale for colouring things?
+- page 10: in "a configuration C with flag $\phi$" you highlight C but not $\phi$ (and likewise in the second-last line of the same paragraph). What's the rationale for colouring things?
 
@@
 
-- The reviewer comments that they cannot write down derivation for `• let (x, x′) = new in spawn(λ().echo x); x′`.
+- The reviewer comments that they cannot write down derivation for `• let (x, x') = new in spawn(lambda ().echo x); x'`.
 
-  The authors: we apologize, there is typo in the term and it should be `• let (x, x′) = new () in spawn(λ().echo x); x′`. We have fixed this in the paper.
+  The authors: we apologize, there is typo in the term and it should be `• let (x, x') = new () in spawn(lambda ().echo x); x'`. We have fixed this in the paper.
 
@@
 
-- T-Pair: why should p < minpr(∆)?
+- T-Pair: why should p < minpr($\Delta$)?
 
@@
 
-  The authors: There is no need for a relationship between p and q because by Lemma 3.1 q=\bottom for all cases.
+  The authors: There is no need for a relationship between p and q because by Lemma 3.1 $q = \bot$ for all cases.
 
@@
 
-  > In GV, this is not the case, and subject reduction relies on proving that if the result of rewriting via ≡ followed by reducing via −→C is an ill-typed configuration, we can rewrite it to a well-typed configuration via ≡.
+  > In GV, this is not the case, and subject reduction relies on proving that if the result of rewriting via $\equiv$ followed by reducing via $\longrightarrow_C$ is an ill-typed configuration, we can rewrite it to a well-typed configuration via $\equiv$.
 
@@
 
-  The authors: We changed the definitions of \pr and \minpr. We now overload \pr to work on session types, data types and contexts, and remove \minpr.
+  The authors: We changed the definitions of pr and minpr. We now overload pr to work on session types, data types and contexts, and remove minpr.
 
@@
   The authors: we have marked this proof as a sketch. We have also added the following comment:
-  >  This proof follows the exact same reasoning and proof sketch given by Kobayashi in \cite{kobayashi06} and later adopted by Dardha and Gay for PCP in their technical report \cite{dardhagay18extended}.
+  >  This proof follows the exact same reasoning and proof sketch given by Kobayashi [Kobayashi06] and later adopted by Dardha and Gay for PCP in their technical report [DardhaGay18].
 
@@
 The reviewer suggested adding an comparing to the following related work.
-- M.
-Coppo, M.
-Dezani-Ciancaglini, L.
-Padovani, and N.
-Yoshida.
-Inference of global progress prop- erties for dynamically interleaved multiparty sessions.
-In COORDINATION, 2013.
-doi: 10.1007/ 978-3-642-38493-6\ 4.
+- M. Coppo, M. Dezani-Ciancaglini, L. Padovani, and N. Yoshida. Inference of global progress prop- erties for dynamically interleaved multiparty sessions. In COORDINATION, 2013.
 
-- M.
-Coppo, M.
-Dezani-Ciancaglini, N.
-Yoshida, and L.
-Padovani.
-Global progress for dynamically interleaved multiparty sessions.
-MSCS, 26(2):238–302, 2016.
-doi: 10.1017/S0960129514000188.
-URL https://doi.org/10.1017/S0960129514000188.
+- M. Coppo, M. Dezani-Ciancaglini, N. Yoshida, and L. Padovani. Global progress for dynamically interleaved multiparty sessions. MSCS, 26(2):238–302, 2016.
 
-- B.
-Toninho, L.
-Caires, and F.
-Pfenning.
-Higher-order processes, functions, and sessions: A monadic integration.
-In ESOP, volume 7792 of LNCS, pages 350–369, 2013.
-doi: 10.1007/978-3-642-37036-6\ 20.
-URL https://doi.org/10.1007/978-3-642-37036-6\_20.
+- B. Toninho, L. Caires, and F. Pfenning. Higher-order processes, functions, and sessions: A monadic integration. In ESOP, volume 7792 of LNCS, pages 350–369, 2013.
 
@@
 
-- page 6: formating of \phi ::= ... is broken
+- page 6: formating of $\phi ::= ...$ is broken
 

### 2023-11-13 lmcs/introduction.tex
Did as many of the TODOs as I could
Update lmcs.cls

@@
 This closes the circle of the connection between CP and GV~\cite{wadler14}, and their priority-based versions, PCP~\cite{dardhagay18extended} and PGV.
+We cannot straightforwardly adapt the priority typing from PCP to PGV, as PGV adds higher-order functions.
+Instead, the priority typing for PGV follow the work by Padovani and Novara~\cite{padovaninovara15}.
 
@@
         We present a \emph{sound and complete encoding} of PCP~\cite{dardhagay18extended} in PGV (\cref{sec:pcp}).
-We prove the encoding preserves typing (\cref{thm:pcp-to-pgv-confs-preservation}) and satisfies operational correspondence (\cref{thm:pcp-to-pgv-operational-correspondence-soundness,thm:pcp-to-pgv-operational-correspondence-completeness}).
+We prove the encoding preserves typing (\cref{thm:pcp-to-pgv-confs-preservation}) and satisfies operational correspondence (\cref{thm:pcp-to-pgv-operational-correspondence-soundness} and \cref{thm:pcp-to-pgv-operational-correspondence-completeness}).
 

### 2023-11-13 lmcs/priority-cp-long.tex
Update lmcs.cls

@@
 
-% \begin{restatablelemma}{lempgvreadypriority}
+% \begin{lem}
 %   \label{lem:pgv-ready-priority}
 %   If $\tseq[\cs{p}]{\ty{\Gamma}}{L}{T}$ is ready to act on $\tmty{x}{S}\in\ty{\Gamma}$, then the priority bound $\cs{p}$ is some priority $\cs{o}$, \ie not $\cs{\pbot}$ or $\cs{\ptop}$.
-% \end{restatablelemma}
-% \begin{proof}
+% \end{lem}
+% \proof
 %   Let $\tm{L}=\tm{\plug{E}{M}}$.
@@
 No rule introduces the priority bound $\cs{\ptop}$ on the sequent.
-% \end{proof}
+% \qed
 
 % \restatelemma{lempgvcanonicalforms}
-% \begin{proof}
+% \proof
 %   We move any $\nu$-binders to the top using \LabTirName{SC-ResExt}, discard any superfluous occurrences of $\tm{\child\;\unit}$ using \LabTirName{SC-ParNil}, and move the main thread to the rightmost position using \LabTirName{SC-ParComm} and \LabTirName{SC-ParAssoc}.
-% \end{proof}
+% \qed
 
@@
 \subsection{Technical Developments}
-\begin{definition}[Actions]
+\begin{defi}
   A~process acts on an endpoint $\tm{x}$ if it is $\tm{\link{x}{y}}$, $\tm{\link{y}{x}}$, $\tm{\send{x}{y}{P}}$, $\tm{\recv{x}{y}{P}}$, $\tm{\close{x}{P}}$, $\tm{\wait{x}{P}}$, $\tm{\inl{x}{P}}$, $\tm{\inr{x}{P}}$, $\tm{\offer{x}{P}{Q}}$, or $\tm{\absurd{x}}$.
 A~process is an action if it acts on some endpoint $\tm{x}$.
-\end{definition}
-\begin{definition}[Canonical Forms]
-\label{def:pcp-canonical-forms}
+\end{defi}
+\begin{defi}\label{def:pcp-canonical-forms}
+  %[Canonical Forms]
   A~process $\tm{P}$ is in canonical form if it is either $\tm{\halt}$ or of the form $\tm{\res{x_1}{x'_1}{\dots\res{x_n}{x'_n}{(P_1 \parallel\dots\parallel P_m)}}}$ where $m>0$ and each $\tm{P_j}$ is an action.
-\end{definition}
-\begin{lemma}[Canonical Forms]
-\label{lem:pcp-canonical-forms}
+\end{defi}
+\begin{lem}\label{lem:pcp-canonical-forms}
+  %[Canonical Forms]
   If $\seq{P}{\ty{\Gamma}}$, there exists some $\tm{Q}$ such that $\tm{P}\equiv\tm{Q}$ and $\tm{Q}$ is in canonical form.
-\end{lemma}
-\begin{proof}
+\end{lem}
+\proof
 If $\tm{P}=\tm{\halt}$, we are done. Otherwise, we move any $\nu$-binders to the top using \LabTirName{SC-ResExt}, and discard any superfluous occurrences of $\tm{\halt}$ using \LabTirName{SC-ParNil}.
-\end{proof}
+\qed
 
@@
 
-%\restatetheorem{thmpcpclosedprogress}
-\begin{restatabletheorem}{thmpcpclosedprogress}[Progress, $\pcp{\red}$]%
-  \label{thm:pcp-closed-progress}
-  \hfill\\%newline before theorem statement
+\begin{thm}\label{thm:pcp-closed-progress}
+  %[Progress, $\pcp{\red}$]%
   If $\pcp{\seq{P}{\emptyenv}}$, then either $\pcp{\tm{P}=\tm{\halt}}$ or there exists a $\pcp{\tm{Q}}$ such that $\pcp{\tm{P}\red\tm{Q}}$.
-\end{restatabletheorem}
+\end{thm}
 \input{prf/thm-pcp-closed-progress}
@@
 \begin{align*}
-  &\pcp{\tm{\cpgvM{\link{x}{y}}}}
-  &&= \pgv{\tm{\link\;{\pair{x}{y}}}} \\
-  &\pcp{\tm{\cpgvM{\res{x}{y}{P}}}}
-  &&= \pgv{\tm{\letpair{x}{y}{\new\;\unit}{\cpgvM{P}}}} \\
-  &\pcp{\tm{\cpgvM{\ppar{P}{Q}}}}
-  &&= \pgv{\tm{\andthen{\spawn\;{(\lambda\unit.\cpgvM{P})}}{\cpgvM{Q}}}} \\
-  &\pcp{\tm{\cpgvM{\halt}}}
-  &&= \pgv{\tm{\unit}} \\
-  &\pcp{\tm{\cpgvM{\close{x}{P}}}}
-  &&= \pgv{\tm{\andthen{\close\;{x}}{\cpgvM{P}}}} \\
-  &\pcp{\tm{\cpgvM{\wait{x}{P}}}}
-  &&= \pgv{\tm{\andthen{\wait\;{x}}{\cpgvM{P}}}} \\
-  &\pcp{\tm{\cpgvM{\send{x}{y}{P}}}}
-  &&= \pgv{\tm{\letpair{y}{z}{\new\;\unit}{\letbind{x}{\send\;{\pair{z}{x}}}{\cpgvM{P}}}}} \\
-  &\pcp{\tm{\cpgvM{\recv{x}{y}{P}}}}
-  &&= \pgv{\tm{\letpair{y}{x}{\recv\;{x}}{\cpgvM{P}}}} \\
-  &\pcp{\tm{\cpgvM{\inl{x}{P}}}}
-  &&= \pgv{\tm{\letbind{x}{\select{\labinl}\;{x}}{\cpgvM{P}}}} \\
-  &\pcp{\tm{\cpgvM{\inr{x}{P}}}}
-  &&= \pgv{\tm{\letbind{x}{\select{\labinr}\;{x}}{\cpgvM{P}}}} \\
-  &\pcp{\tm{\cpgvM{\offer{x}{P}{Q}}}}
-  &&= \pgv{\tm{\offer{x}{x}{\cpgvM{P}}{x}{\cpgvM{Q}}}} \\
-  &\pcp{\tm{\cpgvM{\absurd{x}}}}
-  &&= \pgv{\tm{\offeremp{x}}}
+   & \pcp{\tm{\cpgvM{\link{x}{y}}}}
+   &                                    & = \pgv{\tm{\link\;{\pair{x}{y}}}}                                                      \\
+   & \pcp{\tm{\cpgvM{\res{x}{y}{P}}}}
+   &                                    & = \pgv{\tm{\letpair{x}{y}{\new\;\unit}{\cpgvM{P}}}}                                    \\
+   & \pcp{\tm{\cpgvM{\ppar{P}{Q}}}}
+   &                                    & = \pgv{\tm{\andthen{\spawn\;{(\lambda\unit.\cpgvM{P})}}{\cpgvM{Q}}}}                   \\
+   & \pcp{\tm{\cpgvM{\halt}}}
+   &                                    & = \pgv{\tm{\unit}}                                                                     \\
+   & \pcp{\tm{\cpgvM{\close{x}{P}}}}
+   &                                    & = \pgv{\tm{\andthen{\close\;{x}}{\cpgvM{P}}}}                                          \\
+   & \pcp{\tm{\cpgvM{\wait{x}{P}}}}
+   &                                    & = \pgv{\tm{\andthen{\wait\;{x}}{\cpgvM{P}}}}                                           \\
+   & \pcp{\tm{\cpgvM{\send{x}{y}{P}}}}
+   &                                    & = \pgv{\tm{\letpair{y}{z}{\new\;\unit}{\letbind{x}{\send\;{\pair{z}{x}}}{\cpgvM{P}}}}} \\
+   & \pcp{\tm{\cpgvM{\recv{x}{y}{P}}}}
+   &                                    & = \pgv{\tm{\letpair{y}{x}{\recv\;{x}}{\cpgvM{P}}}}                                     \\
+   & \pcp{\tm{\cpgvM{\inl{x}{P}}}}
+   &                                    & = \pgv{\tm{\letbind{x}{\select{\labinl}\;{x}}{\cpgvM{P}}}}                             \\
+   & \pcp{\tm{\cpgvM{\inr{x}{P}}}}
+   &                                    & = \pgv{\tm{\letbind{x}{\select{\labinr}\;{x}}{\cpgvM{P}}}}                             \\
+   & \pcp{\tm{\cpgvM{\offer{x}{P}{Q}}}}
+   &                                    & = \pgv{\tm{\offer{x}{x}{\cpgvM{P}}{x}{\cpgvM{Q}}}}                                     \\
+   & \pcp{\tm{\cpgvM{\absurd{x}}}}
+   &                                    & = \pgv{\tm{\offeremp{x}}}
 \end{align*}
@@
 \begin{align*}
-  &\pcp{\tm{\cpgvC{\res{x}{y}{P}}}}
-  &&= \pgv{\tm{\res{x}{y}{\cpgvC{P}}}}
+   & \pcp{\tm{\cpgvC{\res{x}{y}{P}}}}
+   &                                   & = \pgv{\tm{\res{x}{y}{\cpgvC{P}}}}
   \\
-  &\pcp{\tm{\cpgvC{\ppar{P}{Q}}}}
-  &&= \pgv{\tm{\ppar{\cpgvC{P}}{\cpgvC{Q}}}}
+   & \pcp{\tm{\cpgvC{\ppar{P}{Q}}}}
+   &                                   & = \pgv{\tm{\ppar{\cpgvC{P}}{\cpgvC{Q}}}}
   \\
-  &\pcp{\tm{\cpgvC{\send{x}{y}{P}}}}
-  &&= \pgv{\tm{\res{y}{z}{(\child\;\letbind{x}{\send\;\pair{z}{x}}{\cpgvM{P}})}}}
+   & \pcp{\tm{\cpgvC{\send{x}{y}{P}}}}
+   &                                   & = \pgv{\tm{\res{y}{z}{(\child\;\letbind{x}{\send\;\pair{z}{x}}{\cpgvM{P}})}}}
   \\
-  &\pcp{\tm{\cpgvC{\inl{x}{P}}}}
-  &&= \pgv{\tm{\res{y}{z}{(\child\;\letbind{x}{\andthen{\close\;(\send\;\pair{\inl{y}}{x})}{z}}{\cpgvM{P}})}}}
+   & \pcp{\tm{\cpgvC{\inl{x}{P}}}}
+   &                                   & = \pgv{\tm{\res{y}{z}{(\child\;\letbind{x}{\andthen{\close\;(\send\;\pair{\inl{y}}{x})}{z}}{\cpgvM{P}})}}}
   \\
-  &\pcp{\tm{\cpgvC{\inr{x}{P}}}}
-  &&= \pgv{\tm{\res{y}{z}{(\child\;\letbind{x}{\andthen{\close\;(\send\;\pair{\inr{y}}{x})}{z}}{\cpgvM{P}})}}}
+   & \pcp{\tm{\cpgvC{\inr{x}{P}}}}
+   &                                   & = \pgv{\tm{\res{y}{z}{(\child\;\letbind{x}{\andthen{\close\;(\send\;\pair{\inr{y}}{x})}{z}}{\cpgvM{P}})}}}
   \\
-  &\pcp{\tm{\cpgvC{P}}}
-  &&= \pgv{\tm{\child{\cpgvM{P}}}},\quad\text{if none of the above apply}
+   & \pcp{\tm{\cpgvC{P}}}
+   &                                   & = \pgv{\tm{\child{\cpgvM{P}}}},\quad\text{if none of the above apply}
 \end{align*}
@@
 
-%\restatelemma{lempcptopgvtermspreservation}
-\begin{lemma}[Preservation, ${\tm{\cpgvM{\cdot}}}$]%
-  \label{lem:pcp-to-pgv-terms-preservation}
+\begin{lem}\label{lem:pcp-to-pgv-terms-preservation}
+  % [Preservation, ${\tm{\cpgvM{\cdot}}}$]%
   If $\pcp{\seq{P}{\ty{\Gamma}}}$, then $\pgv{\tseq[\cs{p}]{\ty{\cpgvT{\Gamma}}}{\cpgvM{P}}{\tyunit}}$.
-\end{lemma}
+\end{lem}
 \input{prf/lem-pcp-to-pgv-terms-preservation}
 
-%\restatetheorem{thmpcptopgvconfspreservation}
-\begin{theorem}[Preservation, ${\tm{\cpgvC{\cdot}}}$]%
-  \label{thm:pcp-to-pgv-confs-preservation}
+\begin{thm}\label{thm:pcp-to-pgv-confs-preservation}
+  %[Preservation, ${\tm{\cpgvC{\cdot}}}$]%
   If $\pcp{\seq{P}{\ty{\Gamma}}}$, then $\pgv{\cseq[\child]{\ty{\cpgvT{\Gamma}}}{\cpgvC{P}}}$.
-\end{theorem}
+\end{thm}
 \input{prf/thm-pcp-to-pgv-confs-preservation}
 
-%\restatetheorem{thmpcptopgvoperationalcorrespondencesoundness}
-\begin{theorem}%
-  [Operational Correspondence, Soundness, ${\tm{\cpgvC{\cdot}}}$]%
-  \label{thm:pcp-to-pgv-operational-correspondence-soundness}
-  \hfill\\%newline before theorem statement
+\begin{thm}\label{thm:pcp-to-pgv-operational-correspondence-soundness}
+  %[Operational Correspondence, Soundness, ${\tm{\cpgvC{\cdot}}}$]%
   If $\pcp{\seq{P}{\ty{\Gamma}}}$ and $\pgv{\tm{\cpgvC{P}}\cred\tm{\conf{C}}}$, there exists a $\tm{Q}$ such that $\pcp{\tm{P}\red^+\tm{Q}}$ and $\pgv{\tm{\conf{C}}\cred^\star\tm{\cpgvC{Q}}}$
-\end{theorem}
+\end{thm}
 \input{prf/thm-pcp-to-pgv-operational-correspondence-soundness}
 
-%\restatelemma{lempcptopgvcpgvMtocpgvC}
-\begin{lemma}%
-  \label{lem:pcp-to-pgv-cpgvM-to-cpgvC}
+\begin{lem}\label{lem:pcp-to-pgv-cpgvM-to-cpgvC}
   For any $\pcp{\tm{P}}$, either:
@@
   \end{itemize}
-\end{lemma}
+\end{lem}
 \input{prf/lem-pcp-to-pgv-cpgvM-to-cpgvC}
 
-\begin{lemma}%
-  \label{lem:pcp-to-pgv-confs-operational-correspondence-equiv}
+\begin{lem}\label{lem:pcp-to-pgv-confs-operational-correspondence-equiv}
   If $\pcp{\seq{P}{\ty{\Gamma}}}$ and $\pcp{\tm{P}\equiv\tm{Q}}$,
   then $\pgv{\tm{\cpgvC{P}}\equiv\tm{\cpgvC{Q}}}$.
-\end{lemma}
-\begin{proof}
+\end{lem}
+\proof
 Every axiom of the structural congruence in PCP maps directly to the axiom of the same name in PGV.
-\end{proof}
+\qed
 
-%\restatetheorem{thmpcptopgvoperationalcorrespondencecompleteness}
-\begin{theorem}%
-  [Operational Correspondence, Completeness, ${\tm{\cpgvC{\cdot}}}$]%
-  \label{thm:pcp-to-pgv-operational-correspondence-completeness}
+\begin{thm}\label{thm:pcp-to-pgv-operational-correspondence-completeness}
+  %[Operational Correspondence, Completeness, ${\tm{\cpgvC{\cdot}}}$]%
   \hfill\\%newline before theorem statement
@@
   then $\pgv{\tm{\cpgvC{P}}\cred^+\tm{\cpgvC{Q}}}$.
-\end{theorem}
+\end{thm}
 \input{prf/thm-pcp-to-pgv-operational-correspondence-completeness}
@@
 \usingnamespace{pcp}
-\begin{example}[Milner's Cyclic Scheduler, PCP]
-  \label{ex:pcp-scheduler}
+\begin{exa}\label{ex:pcp-scheduler}
   A set of processes $\tm{\proc{i}}$, $1\leq{i}\leq{n}$, is scheduled to perform some tasks in cyclic order, starting with $\tm{\proc1}$, ending with $\tm{\proc{n}}$, and notifying $\tm{\proc1}$ when all processes have finished.
@@
   \]
-\end{example}
+\end{exa}
 \endgroup
@@
 \usingnamespace{pgv}
-\begin{example}[Milner's Cyclic Scheduler, PGV]
-  \label{ex:pgv-scheduler}
+\begin{exa}\label{ex:pgv-scheduler}
+  %[Milner's Cyclic Scheduler, PGV]
   The PGV scheduler has exactly the same behaviour as the PCP version in~\cref{ex:pcp-scheduler}.

### 2023-11-13 lmcs/priority-cp.tex
restructuring before reworking
Update lmcs.cls

--- 2021-04-14 priority-cp.tex
@@
-% !TeX root = priorities.tex
-\documentclass[main.tex]{subfiles}
-
-\begin{document}
-
 \section{Relation to Priority CP}\label{sec:pcp}
@@
 Typing environments, duality, and the priority function $\pr(\cdot)$ are defined as expected.
-% (see~\cref{sec:pcp-environments,sec:pcp-duality,sec:pcp-priorities}).
 \[
@@
 PCP decomposes cut into \LabTirName{T-Res} and \LabTirName{T-Par} rules---corresponding to cycle and mix rules, respectively---and guarantees deadlock freedom by using priority constraints, \eg, as in \LabTirName{T-Send}.
-% The full typing rules and operational semantics are available in \cref{fig:pcp-operational-semantics,fig:pcp-typing} in \cref{app:revisiting-PCP}:
 \begin{mathpar}
@@
 We prove closed progress for our updated PCP.
-% The proof is in \cref{prf:thm-pcp-closed-progress}.
-\begin{compacttheorems}
-  \begin{restatabletheorem}{thmpcpclosedprogress}[Progress, $\pcp{\red}$]%
-    \label{thm:pcp-closed-progress}
-    \hfill\\%newline before theorem statement
+\begin{thm}\label{thm:pcp-closed-progress}
+  %[Progress, $\pcp{\red}$]%
   If $\pcp{\seq{P}{\emptyenv}}$, then either $\pcp{\tm{P}=\tm{\halt}}$ or there exists a $\pcp{\tm{Q}}$ such that $\pcp{\tm{P}\red\tm{Q}}$.
-  \end{restatabletheorem}
-\end{compacttheorems}
+\end{thm}
 \endgroup
@@
 \begin{align*}
-  &\pcp{\tm{\cpgvM{\link{x}{y}}}}
-  &&= \pgv{\tm{\link\;{\pair{x}{y}}}} \\
-  &\pcp{\tm{\cpgvM{\res{x}{y}{P}}}}
-  &&= \pgv{\tm{\letpair{x}{y}{\new}{\cpgvM{P}}}} \\
-  &\pcp{\tm{\cpgvM{\ppar{P}{Q}}}}
-  &&= \pgv{\tm{\andthen{\spawn\;{(\lambda\unit.\cpgvM{P})}}{\cpgvM{Q}}}} \\
-  &\pcp{\tm{\cpgvM{\halt}}}
-  &&= \pgv{\tm{\unit}} \\
-  &\pcp{\tm{\cpgvM{\close{x}{P}}}}
-  &&= \pgv{\tm{\andthen{\close\;{x}}{\cpgvM{P}}}} \\
-  &\pcp{\tm{\cpgvM{\wait{x}{P}}}}
-  &&= \pgv{\tm{\andthen{\wait\;{x}}{\cpgvM{P}}}} \\
-  &\pcp{\tm{\cpgvM{\send{x}{y}{P}}}}
-  &&= \pgv{\tm{\letpair{y}{z}{\new}{\letbind{x}{\send\;{\pair{z}{x}}}{\cpgvM{P}}}}} \\
-  &\pcp{\tm{\cpgvM{\recv{x}{y}{P}}}}
-  &&= \pgv{\tm{\letpair{y}{x}{\recv\;{x}}{\cpgvM{P}}}} \\
-  &\pcp{\tm{\cpgvM{\inl{x}{P}}}}
-  &&= \pgv{\tm{\letbind{x}{\select{\labinl}\;{x}}{\cpgvM{P}}}} \\
-  &\pcp{\tm{\cpgvM{\inr{x}{P}}}}
-  &&= \pgv{\tm{\letbind{x}{\select{\labinr}\;{x}}{\cpgvM{P}}}} \\
-  &\pcp{\tm{\cpgvM{\offer{x}{P}{Q}}}}
-  &&= \pgv{\tm{\offer{x}{x}{\cpgvM{P}}{x}{\cpgvM{Q}}}} \\
-  &\pcp{\tm{\cpgvM{\absurd{x}}}}
-  &&= \pgv{\tm{\offeremp{x}}}
+   & \pcp{\tm{\cpgvM{\link{x}{y}}}}
+   &                                    & = \pgv{\tm{\link\;{\pair{x}{y}}}}                                               \\
+   & \pcp{\tm{\cpgvM{\res{x}{y}{P}}}}
+   &                                    & = \pgv{\tm{\letpair{x}{y}{\new}{\cpgvM{P}}}}                                    \\
+   & \pcp{\tm{\cpgvM{\ppar{P}{Q}}}}
+   &                                    & = \pgv{\tm{\andthen{\spawn\;{(\lambda\unit.\cpgvM{P})}}{\cpgvM{Q}}}}            \\
+   & \pcp{\tm{\cpgvM{\halt}}}
+   &                                    & = \pgv{\tm{\unit}}                                                              \\
+   & \pcp{\tm{\cpgvM{\close{x}{P}}}}
+   &                                    & = \pgv{\tm{\andthen{\close\;{x}}{\cpgvM{P}}}}                                   \\
+   & \pcp{\tm{\cpgvM{\wait{x}{P}}}}
+   &                                    & = \pgv{\tm{\andthen{\wait\;{x}}{\cpgvM{P}}}}                                    \\
+   & \pcp{\tm{\cpgvM{\send{x}{y}{P}}}}
+   &                                    & = \pgv{\tm{\letpair{y}{z}{\new}{\letbind{x}{\send\;{\pair{z}{x}}}{\cpgvM{P}}}}} \\
+   & \pcp{\tm{\cpgvM{\recv{x}{y}{P}}}}
+   &                                    & = \pgv{\tm{\letpair{y}{x}{\recv\;{x}}{\cpgvM{P}}}}                              \\
+   & \pcp{\tm{\cpgvM{\inl{x}{P}}}}
+   &                                    & = \pgv{\tm{\letbind{x}{\select{\labinl}\;{x}}{\cpgvM{P}}}}                      \\
+   & \pcp{\tm{\cpgvM{\inr{x}{P}}}}
+   &                                    & = \pgv{\tm{\letbind{x}{\select{\labinr}\;{x}}{\cpgvM{P}}}}                      \\
+   & \pcp{\tm{\cpgvM{\offer{x}{P}{Q}}}}
+   &                                    & = \pgv{\tm{\offer{x}{x}{\cpgvM{P}}{x}{\cpgvM{Q}}}}                              \\
+   & \pcp{\tm{\cpgvM{\absurd{x}}}}
+   &                                    & = \pgv{\tm{\offeremp{x}}}
 \end{align*}
@@
 The translations $\tm{\cpgvM{\cdot}}$ and $\tm{\cpgvC{\cdot}}$ preserve typing, and the latter induces a sound and complete operational correspondence.
-% The proofs are by structural induction and can be found in~\cref{prf:lem-pcp-to-pgv-terms-preservation,prf:lem-pcp-to-pgv-terms-preservation,prf:lem-pcp-to-pgv-cpgvM-to-cpgvC,prf:thm-pcp-to-pgv-operational-correspondence-soundness,prf:thm-pcp-to-pgv-operational-correspondence-completeness}.
-\begin{compacttheorems}
-  \begin{restatablelemma}{lempcptopgvtermspreservation}[Preservation, ${\tm{\cpgvM{\cdot}}}$]%
-    \label{lem:pcp-to-pgv-terms-preservation}
+\begin{lem}\label{lem:pcp-to-pgv-terms-preservation}
+  %[Preservation, ${\tm{\cpgvM{\cdot}}}$]%
   If $\pcp{\seq{P}{\ty{\Gamma}}}$, then $\pgv{\tseq[\cs{p}]{\ty{\cpgvT{\Gamma}}}{\cpgvM{P}}{\tyunit}}$.
-  \end{restatablelemma}
-  \begin{restatabletheorem}{thmpcptopgvconfspreservation}[Preservation, ${\tm{\cpgvC{\cdot}}}$]%
-    \label{thm:pcp-to-pgv-confs-preservation}
+\end{lem}
+\begin{thm}\label{thm:pcp-to-pgv-confs-preservation}
+  %[Preservation, ${\tm{\cpgvC{\cdot}}}$]%
   If $\pcp{\seq{P}{\ty{\Gamma}}}$, then $\pgv{\cseq[\child]{\ty{\cpgvT{\Gamma}}}{\cpgvC{P}}}$.
-  \end{restatabletheorem}
-  \begin{restatablelemma}{lempcptopgvcpgvMtocpgvC}%
-    \label{lem:pcp-to-pgv-cpgvM-to-cpgvC}
+\end{thm}
+\begin{lem}\label{lem:pcp-to-pgv-cpgvM-to-cpgvC}
   For any $\pcp{\tm{P}}$, either:
@@
   \end{itemize}
-  \end{restatablelemma}
-  \begin{restatabletheorem}{thmpcptopgvoperationalcorrespondencesoundness}%
-    [Operational Correspondence, Soundness, ${\tm{\cpgvC{\cdot}}}$]%
-    \label{thm:pcp-to-pgv-operational-correspondence-soundness}
-    \hfill\\%newline before theorem statement
+\end{lem}
+\begin{thm}\label{thm:pcp-to-pgv-operational-correspondence-soundness}
+  %[Operational Correspondence, Soundness, ${\tm{\cpgvC{\cdot}}}$]%
   If $\pcp{\seq{P}{\ty{\Gamma}}}$ and $\pgv{\tm{\cpgvC{P}}\cred\tm{\conf{C}}}$, there exists a $\tm{Q}$ such that $\pcp{\tm{P}\red^+\tm{Q}}$ and $\pgv{\tm{\conf{C}}\cred^\star\tm{\cpgvC{Q}}}$
-  \end{restatabletheorem}
-  \begin{restatabletheorem}{thmpcptopgvoperationalcorrespondencecompleteness}%
-    [Operational Correspondence, Completeness, ${\tm{\cpgvC{\cdot}}}$]%
-    \label{thm:pcp-to-pgv-operational-correspondence-completeness}
-    \hfill\\%newline before theorem statement
+\end{thm}
+\begin{thm}\label{thm:pcp-to-pgv-operational-correspondence-completeness}
+  %[Operational Correspondence, Completeness, ${\tm{\cpgvC{\cdot}}}$]%
   If $\pcp{\seq{P}{\ty{\Gamma}}}$ and $\pcp{\tm{P}\red\tm{Q}}$,
   then $\pgv{\tm{\cpgvC{P}}\cred^+\tm{\cpgvC{Q}}}$.
-  \end{restatabletheorem}
-\end{compacttheorems}
+\end{thm}
 \endgroup
-
-\end{document}
-
-%%% Local Variables:
-%%% TeX-master: "priorities"
-%%% End:

### 2023-11-13 lmcs/priority-gv.tex
Update lmcs.cls

@@
 
-We illustrate this with two programs in PGV, \cref{ex:cycle,ex:deadlock}.
+We illustrate this with two programs in PGV, \cref{ex:cycle} and \cref{ex:deadlock}.
 Each program contains two processes---the main process, and the child process created by $\tm{\spawn}$---which communicate using \emph{two} channels.
@@
   \begin{minipage}{0.5\linewidth}
-\begin{example}[Cyclic Structure]%
-  \label{ex:cycle}
+    \begin{exa}\label{ex:cycle}
+      % [Cyclic Structure]%
       \[\tm{%
@@
         }\]
-\end{example}
+    \end{exa}
   \end{minipage}%
   \begin{minipage}{0.5\linewidth}
-\begin{example}[Deadlock]%
-  \label{ex:deadlock}
+    \begin{exa}\label{ex:deadlock}
+      % [Deadlock]%
       \[\tm{%
@@
         }\]
-\end{example}
+    \end{exa}
   \end{minipage}
@@
 For this, we prove~\cref{lem:pgv-value-done}, which shows that values have finished all their communication, and that any priorities in the type of the value come from the typing environment.
-%The proofs are by structural induction and can be found in \cref{prf:lem-pgv-value-done,prf:lem-pgv-substitution,prf:lem-pgv-subject-reduction-terms,prf:lem-pgv-subject-congruence,prf:thm-pgv-subject-reduction-confs}.
-  \begin{lemma}%
-    \label{lem:pgv-value-done}
+\begin{lem}\label{lem:pgv-value-done}
   If $\tseq[\cs{p}]{\ty{\Gamma}}{V}{T}$, then $\cs{p}=\cs{\pbot}$, and $\minpr(\ty{\Gamma})=\minpr(\ty{T})$.
-  \end{lemma}
+\end{lem}
 \input{prf/lem-pgv-value-done}
 
-  \begin{lemma}[Substitution]%
-    \label{lem:pgv-substitution}
-    \hfill\\%newline before theorem statement
+\begin{lem}\label{lem:pgv-substitution}
+  %[Substitution]
   If $\tseq[\cs{p}]{\ty{\Gamma},\tmty{x}{U'}}{M}{T}$ and $\tseq[\cs{q}]{\ty{\Theta}}{V}{U'}$, then $\tseq[\cs{p}]{\ty{\Gamma},\ty{\Theta}}{\subst{M}{V}{x}}{T}$.
-  \end{lemma}
+\end{lem}
 \input{prf/lem-pgv-substitution}
 
-  \begin{lemma}[Subject Reduction, $\tred$]%
-    \label{lem:pgv-subject-reduction-terms}
-    \hfill\\%newline before theorem statement
+\begin{lem}\label{lem:pgv-subject-reduction-terms}
+  %[Subject Reduction, $\tred$]%
   If $\tseq[\cs{p}]{\ty{\Gamma}}{M}{T}$ and $\tm{M}\tred\tm{M'}$,
   then $\tseq[\cs{p}]{\ty{\Gamma}}{M'}{T}$.
-  \end{lemma}
+\end{lem}
 \input{prf/lem-pgv-subject-reduction-terms}
 
-  \begin{lemma}[Subject Congruence, $\equiv$]%
-    \label{lem:pgv-subject-congruence}
-    \hfill\\%newline before theorem statement
+\begin{lem}\label{lem:pgv-subject-congruence}
+  %[Subject Congruence, $\equiv$]%
   If $\cseq[\phi]{\ty{\Gamma}}{\conf{C}}$ and $\tm{\conf{C}}\equiv\tm{\conf{C'}}$,
   then $\cseq[\phi]{\ty{\Gamma}}{\conf{C'}}$.
-  \end{lemma}
+\end{lem}
 \input{prf/lem-pgv-subject-congruence}
 
-  \begin{theorem}[Subject Reduction, $\cred$]%
-    \label{thm:pgv-subject-reduction-confs}
-    \hfill\\%newline before theorem statement
+\begin{thm}\label{thm:pgv-subject-reduction-confs}
+  %[Subject Reduction, $\cred$]%
   If $\cseq[\phi]{\ty{\Gamma}}{\conf{C}}$ and $\tm{\conf{C}}\cred\tm{\conf{C'}}$,
   then $\cseq[\phi]{\ty{\Gamma}}{\conf{C'}}$.
-  \end{theorem}
+\end{thm}
 \input{prf/thm-pgv-subject-reduction-confs}
@@
 \emph{Actions} are terms which perform communication actions and which synchronise between two threads.
-\begin{definition}[Actions]%
-  \label{def:pgv-actions}
+\begin{defi}\label{def:pgv-actions}
+  %[Actions]%
   A~term \emph{acts on} an endpoint $\tm{x}$ if it is $\tm{\send\;\pair{V}{x}}$, $\tm{\recv\;{x}}$, $\tm{\close\;{x}}$, or $\tm{\wait\;{x}}$.
 A~term is an \emph{action} if it acts on some endpoint $\tm{x}$.
-\end{definition}
+\end{defi}
 
@@
 Ready processes like live/poised processes denote processes that are ready to communicate on their providing channel.
-\begin{definition}[Ready Terms]%
-  \label{def:pgv-ready-actions}
+\begin{defi}\label{def:pgv-ready-actions}
+  % [Ready Terms]%
   A~term $\tm{L}$ is \emph{ready} if it is of the form $\tm{\plug{E}{M}}$, where $\tm{M}$ is of the form $\tm{\new}$, $\tm{\spawn\;N}$, $\tm{\link\;\pair{x}{y}}$, or $\tm{M}$ acts on $\tm{x}$.
 In the latter case, we say that $\tm{L}$ is \emph{ready to act on} $\tm{x}$ or is \emph{blocked on}.
-\end{definition}
+\end{defi}
 
@@
 
-\begin{lemma}[Progress, $\tred$]%
-  \label{lem:pgv-open-progress-terms}
+\begin{lem}\label{lem:pgv-open-progress-terms}
+  %[Progress, $\tred$]%
   If $\tseq[\cs{p}]{\ty{\Gamma}}{M}{T}$ and $\ty{\Gamma}$ contains only session types, then:
@@
   \end{enumerate*}
-\end{lemma}
+\end{lem}
 
@@
 
-  \begin{definition}[Canonical Forms]%
-    \label{def:pgv-canonical-forms}
+\begin{defi}\label{def:pgv-canonical-forms}
+  %[Canonical Forms]%
   A~configuration $\tm{\conf{C}}$ is in canonical form if it is of the form
@@
   where no term $\tm{M_i}$ is a value.
-  \end{definition}
+\end{defi}
 
-  \begin{lemma}[Canonical Forms]%
-    \label{lem:pgv-canonical-forms}
+\begin{lem}\label{lem:pgv-canonical-forms}
+  % [Canonical Forms]%
   If $\cseq[\main]{\ty{\Gamma}}{\conf{C}}$, there exists some $\tm{\conf{D}}$ such that $\tm{\conf{C}}\equiv\tm{\conf{D}}$ and $\tm{\conf{D}}$ is in canonical form.
-  \end{lemma}
-  \begin{proof}
+\end{lem}
+\proof
 We move any $\nu$-binders to the top using \LabTirName{SC-ResExt}, discard any superfluous occurrences of $\tm{\child\;\unit}$ using \LabTirName{SC-ParNil}, and move the main thread to the rightmost position using \LabTirName{SC-ParComm} and \LabTirName{SC-ParAssoc}.
-  \end{proof}
+\qed
 
-  \begin{definition}[Normal Forms]%
+\begin{defi}
+  % [Normal Forms]%
   A~configuration $\tm{\conf{C}}$ is in normal form if it is of the form
@@
   where each $\tm{M_i}$ is ready to act on $\tm{x_i}$.
-  \end{definition}
+\end{defi}
 
-  \begin{lemma}
-    \label{lem:pgv-ready-priority}
+\begin{lem}\label{lem:pgv-ready-priority}
   If $\tseq[\cs{p}]{\ty{\Gamma}}{L}{T}$ is ready to act on $\tmty{x}{S}\in\ty{\Gamma}$, then the priority bound $\cs{p}$ is some priority $\cs{o}$, \ie not $\cs{\pbot}$ or $\cs{\ptop}$.
-  \end{lemma}
-  \begin{proof}
+\end{lem}
+\proof
 Let $\tm{L}=\tm{\plug{E}{M}}$.
@@
 No rule introduces the priority bound $\cs{\ptop}$ on the sequent.
-  \end{proof}
+\qed
 
-  \begin{theorem}[Progress, $\cred$]%
-    \label{thm:pgv-closed-progress-confs}
+\begin{thm}\label{thm:pgv-closed-progress-confs}
+  % [Progress, $\cred$]%
   If $\cseq[\main]{\emptyenv}{\conf{C}}$ and $\tm{\conf{C}}$ is in canonical form, then either $\tm{\conf{C}}\cred\tm{\conf{D}}$ for some $\tm{\conf{D}}$; or $\tm{\conf{C}\equiv\conf{D}}$ for some $\tm{\conf{D}}$ in normal form.
-  \end{theorem}
+\end{thm}
 \input{prf/thm-pgv-closed-progress-confs}
 
-\begin{corollary}%
-  \label{cor:pgv-closed-progress}
+\begin{cor}\label{cor:pgv-closed-progress}
   If $\cseq[\phi]{\emptyenv}{\conf{C}}$, $\tm{\conf{C}}\centernot\cred$, and $\tm{\conf{C}}$ contains no endpoints, then $\tm{\conf{C}}\equiv\tm{\phi\;V}$ for some value $\tm{V}$.
-\end{corollary}
+\end{cor}
 
